### 2022-04-25 manuscript/main.tex
add manuscript template

%%%%%%%%%%TO FORMAT A PREPRINT FOR A PCI%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper]{article}
%%%%%%%%%   CHOOSE A PCI   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\newcommand{\PCI}{Peer Community In Zoology}
%\newcommand{\PCI}{Peer Community In Ecology}
%\newcommand{\PCI}{Peer Community In Evolutionary Biology}
%\newcommand{\PCI}{Peer Community In Paleontology}
%\newcommand{\PCI}{Peer Community In Archaeology}
%\newcommand{\PCI}{Peer Community In Animal Science}
\newcommand{\PCI}{Peer Community In Genomics}
%\newcommand{\PCI}{Peer Community In Neuroscience}
%\newcommand{\PCI}{Peer Community In Mathematical and Computational Biology}
%\newcommand{\PCI}{Peer Community In Forest and Wood Sciences}
%\newcommand{\PCI}{Peer Community In Ecotoxicology and Environmental Chemistry}
%\newcommand{\PCI}{Peer Community In Infections}
%\newcommand{\PCI}{Peer Community In Registered Reports}
%\newcommand{\PCI}{Peer Community In Network Science}
%\newcommand{\PCI}{Peer Community In Microbiology}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{preambule.tex}

%%%%%%%%%   CHOOSE A BADGE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Comment lines 131 and/or 133 in preambule.tex if you don't use data and/or code in your preprint
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%   SET THE TITLE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\preprinttitle}{This is the title of the preprint}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%   SET THE version of the preprint. Replace X by a number %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\version}{X}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%  SET THE LIST OF AUTHORS WITH CORRESPONDING AFFILIATIONS  use \& before last author %%%%%%%%%%%%%%%%%%%%
\newcommand{\listauthors}{\raggedright 
FirstnameOne FamilynameOne\textsuperscript{1}, \space
FirstnameTwo FamilynameTwo\textsuperscript{1}, \&
FirstnameThree FamilynameThree\textsuperscript{2}
%\fourthauthor \textsuperscript{i}
%etc...
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE LIST OF AFFILIATIONS  %%%%%%%%%%%%%%%%%%%%
\newcommand{\listinstitutions}{
\textsuperscript{1} DeptOne, InstitutionOne -- TownOne, CountryOne 
\\
\textsuperscript{2} DeptTwo, InstitutionTwo -- TownTwo, CoutryTwo
%\\
%\textsuperscript{3} \thirdinstitution
%etc
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE DATE OF UPLOAD on the preprint server %%%%%%%%%%%%%%
\newcommand{\datepub}{ddth Month yyyy}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE RECOMMENDER(s) NAME(s) %%%%%%%%%%%%%%
\newcommand{\recommender}{FirstName FamilyName}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE DOI of the RECOMMENDATION %%%%%%%%%%%%%%
\newcommand{\DOIrecommendation}{xxx/xxx}
%%%%%%%%%%%%% for example 10.24072/pci.mcb.100003 %%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE REVIEWERS' NAMES IF KNOWN and/or X anonymous reviewers %%%%%%%%%%%%%%
\newcommand{\reviewers}{FirstName FamilyName and two anonymous reviewer}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%  SET THE 'CITE AS' OF YOUR PREPRINT. paste the line we sent you by Email (XXX the "cite as") in place of the xxx %%%%%%%%%%%%
\newcommand{\citeas}{xxx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE 'CORRESPONDENCE TO' %%%%%%%%%%%%%%%%%
\newcommand{\email}{mail.mail@mail.xx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE ABSTRACT %%%%%%%%%%%%%%%%%
\newcommand{\preprintabstract}{This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract. This is the abstract.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE KEYWORDS %%%%%%%%%%%%%%%%%
\newcommand{\preprintkeywords}{KeywordOne; KeywordTwo; KeywordThree; KeywordFour; KeywordFive}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\beginingpreprint

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Text of the preprint %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% use \emph{} for italics and \parencite{} to cite a reference and \ref{XX} to cite the \label{XX}
% use \\ and a blank line to mark the end of paragraph and starting of a new one

\section*{\centering Introduction}

This is an example of text. This realization is especially important because it can flip around our expectations about which species expand fast, and how to manage them. We tend to think of initial colonization and long-term abundance as two independent axes of variation among species or indeed as two ends of a spectrum, in the classic competition-colonization tradeoff \parencite{levins1971regional}. When both play into invasion speed, good dispersers might not outrun good competitors. This is useful knowledge, whether we want to contain an invasion or secure a reintroduction. \\

In their study "When higher carrying capacities lead to faster propagation", \textcite{levins1971regional} combine mathematical analysis, Individual-Based simulations and experiments to show that various mechanisms can cause pushed fronts, whose speed increases with the carrying capacity K of the species. Rather than focus on one particular angle, the authors endeavor to demonstrate that this qualitative effect appears again and again in a variety of settings. \\

It is perhaps surprising that this notable and general connection between K and invasion speed has managed to garner so little fame in ecology. A large fraction of the literature employs the venerable Fisher-KPP reaction-diffusion model, which combines local logistic growth with linear diffusion in space. This model has prompted both considerable mathematical developments \parencite{crooks2004spatial} and many applications to modelling real invasions \parencite{shigesada1997biological}. But it only allows pulled fronts, driven by the small populations at the edge of a species range, with a speed that depends only on their initial growth rate r.\\

\section*{\centering Material and methods}

This classic setup is, however, singular in many ways. \textcite{levins1971regional} use it as a null model, and introduce three mechanisms or factors that each ensure a 1 role of K in invasion speed, while giving less importance to the pioneers at the border. \\

Two factors, the Allee effect and demographic stochasticity, make small edge populations slower to grow or less likely to survive. These two factors are studied theoretically, and to make their claims stronger, the authors stack the deck against K. When generalizing equations or simulations beyond the null case, it is easy to obtain functional forms where the parameter K does not only play the role of equilibrium carrying capacity, but also affects dynamical properties such as the maximum or mean growth rate. In that case, it can trivially change the propagation speed, without it meaning anything about the role of established populations behind the front. \textcite{levins1971regional} avoid this pitfall by disentangling these effects, at the cost of slightly more peculiar expressions, and show that varying essentially nothing but the carrying capacity can still impact the speed of the invasion front. \\

\section*{\centering Results}
The third factor, density-dependent dispersal, makes small populations less prone to disperse. It is well established empirically (see Fig~\ref{Fig1}) and theoretically that various biological mechanisms, from collective organization to behavioral switches, can prompt organisms in denser populations to disperse more, e.g. in such a way as to escape competition \parencite{matthysen2005density}. The authors demonstrate how this effect induces a link between carrying capacity and invasion speed, both theoretically and in a dispersal experiment on the parasitoid wasp, \emph{Trichogramma chilonis}.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%  EXAMPLE OF A FIGURE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h]
  \captionsetup{justification=centering}
  \caption{A nice picture of bidule.}
  \centering
  \includegraphics[scale=0.5]{fig.png}
  \label{Fig1}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{\centering Discussion}
Overall, this study carries a simple and clear message, supported by valuable contributions from different angles. Although some sections are clearly written for the theoretical ecology crowd, this article has something for everyone, from the stray physicist to the open-minded manager. The collaboration between theoreticians and experimentalists, while not central, is worthy of note. Because the narrative of this study is the variety of mechanisms that can lead to the same qualitative effect, the inclusion of various approaches is not a gimmick, but helps drive home its main message. The work is fairly self-contained, although one could always wish for further developments, especially in the direction of more quantitative testing of these mechanisms. \\

In conclusion, \textcite{levins1971regional} effectively convey the widely relevant message that, for some species, invading is not just about the destination, it is about the many offspring one makes along the way. \\

\section*{\centering Acknowledgements}
This is your acknowledgments. Version \version~ of this article has been peer-reviewed and recommended by 
\emph{\PCI} (\DOIrecommendationlink)

\section*{\centering Conflict of interest disclosure}
The authors of this preprint declare that they have no financial conflict of interest with the content of this article. XXX and XXX are recommenders for PCI XXX

\section*{\centering Data, script and code availability}
Data are available online: link or DOI of the webpage hosting the data

\section*{\centering Supplementary information availability}
Script and codes are available online: link or DOI of the webpage hosting the script and codes

\titleformat*{\section}{\bfseries\Large\centering}

\printbibliography[notcategory=ignore]

\section*{Annexes or Supplementary Information}
This is your annex 1

This is your annex 2

\end{document}

### 2022-04-25 manuscript/preambule.tex

%%%%%%%%%%TO FORMAT A PREPRINT%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[top=7cm,bottom=2.5cm,headheight=120pt,headsep=15pt,left=6cm,right=1.5cm,marginparwidth=4.2cm,marginparsep=0.5cm]{geometry}

\usepackage{marginnote}
\usepackage{ifthen}
\reversemarginpar  % sets margin notes to the left
\usepackage{lipsum} % Required to insert dummy text
\usepackage{calc}
\usepackage{siunitx}
\usepackage{lineno}
\usepackage{titlesec}
\usepackage{indentfirst}
%\usepackage[none]{hyphenat} % use only if there is a problem
% Use Unicode characters
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Clean citations with biblatex
\usepackage[
backend=biber,
natbib=true,
sortcites=true,
defernumbers=true,
style=authoryear,
citestyle=authoryear-comp,
maxnames=99,
maxcitenames=2,
giveninits=true,
terseinits=true,
url=false
]{biblatex}
\DeclareNameAlias{default}{family-given}
\renewcommand*{\revsdnamepunct}{} % no comma between family and given names
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}} % remove 'In:' before journal name
\DeclareFieldFormat[article]{pages}{#1} % remove pp.
\AtEveryBibitem{\ifentrytype{article}{\clearfield{number}}{}} % don't print issue numbers
\DeclareFieldFormat[article, inbook, incollection, inproceedings, misc, thesis, unpublished]{title}{#1} % title without quotes
\usepackage{csquotes}
\RequirePackage[english]{babel} % must be called after biblatex
\addbibresource{refs.bib}
\DeclareBibliographyCategory{ignore}
\addtocategory{ignore}{recommendation} % adding recommendation to 'ignore' category so that it does not appear in the References
% Clickable references. Use \url{www.example.com} or \href{www.example.com}{description} to add a clicky url
\usepackage{nameref}
\usepackage[pdfborder={0 0 0}]{hyperref}  % sets link border to white
\urlstyle{same}

% Include figures
\usepackage{graphbox}  % loads graphicx ppackage with extended options for vertical alignment of figures
% Line numbers
%\usepackage[right]{lineno}
% Improve typesetting in LaTex
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }
% Text layout and font (Open Sans)
\setlength{\parindent}{0.4cm}
\linespread{1.2}
\RequirePackage[default,scale=0.90]{opensans}
% Defining document colors
\usepackage{xcolor}
\definecolor{darkgray}{HTML}{808080}
\definecolor{mediumgray}{HTML}{6D6E70}
\definecolor{ligthgray}{HTML}{d9d9d9}
\definecolor{pciblue}{HTML}{74adca}
\definecolor{opengreen}{HTML}{77933c}
% Use adjustwidth environment to exceed text width
\usepackage{changepage}
% Adjust caption style
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,singlelinecheck=off]{caption}

% Headers and footers

% Headers and footers
\usepackage{fancyhdr}  % custom headers/footers
\pagestyle{fancy}  % enables customization of headers/footers
\fancyhfoffset[L]{4.5cm}  % offsets header and footer to the left to include margin
\renewcommand{\headrulewidth}{\ifnum\thepage=1 0.5pt \else 0pt \fi} % header ruler only on first page
%\renewcommand{\footrulewidth}{0.5pt}
\renewcommand{\footrulewidth}{\ifnum\thepage=1 0.5pt \else 0pt \fi}
% full logo on first page, then no logo on subsequent pages 
\lhead{\ifnum\thepage=1 \includegraphics[width=13.5cm]{\logoname}\else \fi}  
\chead{}
\rhead{}

\lfoot{\ifnum\thepage=1 \scriptsize \textsc{\color{mediumgray}\PCI} \else  \fi}
\cfoot{}
\rfoot{}
% End Headers and footers

% DOI's
\newcommand{\DOIlink}{\href{https://doi.org/\DOI}{https://doi.org/\DOI}}
\newcommand{\DOIrecommendationlink}{\href{https://doi.org/\DOIrecommendation}{https://doi.org/\DOIrecommendation}}

%which logo to display
\ifthenelse{\equal{\PCI}{Peer Community In Zoology}}{\newcommand{\logoname}{logo_PDF_zool.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Ecology}}{\newcommand{\logoname}{logo_PDF_ecology.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Evolutionary Biology}}{\newcommand{\logoname}{logo_PDF_evolbiol.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Animal Science}}{\newcommand{\logoname}{logo_PDF_animsci.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Archaeology}}{\newcommand{\logoname}{logo_PDF_archaeo.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Forest and Wood Sciences}}{\newcommand{\logoname}{logo_PDF_fws.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Genomics}}{\newcommand{\logoname}{logo_PDF_genomics.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Mathematical and Computational Biology}}{\newcommand{\logoname}{logo_PDF_mcb.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Network Science}}{\newcommand{\logoname}{logo_PDF_networksci.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Paleontology}}{\newcommand{\logoname}{logo_PDF_paleo.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Neuroscience}}{\newcommand{\logoname}{logo_PDF_neuro.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Ecotoxicology and Environmental Chemistry}}{\newcommand{\logoname}{logo_PDF_ecotoxenvchem.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Infections}}{\newcommand{\logoname}{logo_PDF_infections.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Registered Reports}}{\newcommand{\logoname}{logo_PDF_rr.jpg}}{}

\newcommand{\beginingpreprint}{

\vspace*{0.5cm}
\begin{flushleft}
\baselineskip=30pt
% Margin information
\marginpar{
\large\textnormal{\color{pciblue}\\RESEARCH ARTICLE}\\
\vspace*{0.5pt}
\\
\includegraphics[align=c,width=0.5cm]{OAlogoblue.png} \space \large\textbf{\color{pciblue}Open Access}\\
\\
\includegraphics[align=c,width=0.5cm]{o_peerreview.png} \space \large\textbf{\color{pciblue}Open Peer-Review}\\
\\
\includegraphics[align=c,width=0.5cm]{o_data3.png} \space \large\textbf{\color{pciblue}Open Data}\\
\\
\includegraphics[align=c,width=0.5cm]{o_script3.png} \space \large\textbf{\color{pciblue}Open Code}\\
\\
\\
\\
\\
\raggedright
%\vspace*{3.25cm}
\scriptsize\textbf{Cite as:}\space
\citeas\\
\vspace*{0.5cm}
\textbf{Posted:} \datepub\\
\vspace*{0.5cm}
\textbf{Recommender:}\\
\recommender\\
\vspace*{0.5cm}
\textbf{Reviewers:}\\
\reviewers\\
\vspace*{0.5cm}
\textbf{Correspondence:}\\
\href{mailto:\email}{\email}\\

}
{\Huge
\fontseries{sb}\selectfont{\preprinttitle}}
\end{flushleft}
\vspace*{0.25cm}
\begin{flushleft}
%\marginpar{

%\marginnote{
%\raggedright
%\vspace*{3.25cm}
%\scriptsize\textbf{Cite as:}\space
%\fullcite{preprint}\\
%\vspace*{0.5cm}
%\textbf{Published:} \datepub\\
%\vspace*{0.5cm}
%\textbf{Recommender:}\\
%\recommender\\
%\vspace*{0.5cm}
%\textbf{Reviewers:}\\
%\reviewers\\
%\vspace*{0.5cm}
%\textbf{Correspondence:}\\
%\href{mailto:\email}{\email}\\
%\vspace*{0.5cm} 
%\vspace*{3cm}
%\vspace*{0.2cm}
%}[0.8cm]

\Large
\listauthors
\end{flushleft}
\bigskip
{\raggedright
\listinstitutions}
% Recommended preprint box
\begin{flushleft}
%\noindent
\fcolorbox{lightgray}{lightgray}{
\parbox{\textwidth - 2\fboxsep}{
\centering\large{\fontseries{sb}\selectfont{This article has been peer-reviewed and recommended by\\
\emph{\PCI} (\DOIrecommendationlink)}}\\
%\raggedright\large{\fontseries{sb}\selectfont{An article peer-reviewed and recommended by \emph{\PCI}, edited by \recommender \space on the basis of the reviews by \reviewers \space (DOI: \DOIrecommendationlink)}}\\
%\small \fullcite{preprint}
}}
\end{flushleft}
\vspace*{0.5cm}
\fcolorbox{pciblue}{pciblue}{
\parbox{\textwidth - 2\fboxsep}{
\vspace{0.25cm}
\textbf{\large{\textsc{Abstract}}}\\
\preprintabstract\\

\footnotesize{\textbf{\emph{Keywords: }}\preprintkeywords}
\vspace{0.25cm}}
}
\newpage
\newgeometry{margin=1in}
}

### 2022-04-25 manuscript/main.tex
add materials / methods

@@
 
-\input{preambule.tex}
+\input{preamble.tex}
 
@@
 %%%%%%%%%   SET THE TITLE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\preprinttitle}{This is the title of the preprint}
+\newcommand{\preprinttitle}{Gut microbes and their genes or something}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%   SET THE version of the preprint. Replace X by a number %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\version}{X}
+\newcommand{\version}{1}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand{\listauthors}{\raggedright 
-FirstnameOne FamilynameOne\textsuperscript{1}, \space
-FirstnameTwo FamilynameTwo\textsuperscript{1}, \&
-FirstnameThree FamilynameThree\textsuperscript{2}
+Kevin S. Bonham\textsuperscript{1}, \space
+Guilherme Fahur Bottino\textsuperscript{1}, 
+Firstname Lastname\textsuperscript{2}, \&
+Vanja Klepac-Ceraj\textsuperscript{1}
 %\fourthauthor \textsuperscript{i}
@@
 \newcommand{\listinstitutions}{
-\textsuperscript{1} DeptOne, InstitutionOne -- TownOne, CountryOne 
+\textsuperscript{1} Department of Biological Sciences, Wellesley College -- Wellesley, MA, USA
 \\
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE 'CORRESPONDENCE TO' %%%%%%%%%%%%%%%%%
-\newcommand{\email}{mail.mail@mail.xx}
+\newcommand{\email}{vklepacc@wellesley.edu}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE ABSTRACT %%%%%%%%%%%%%%%%%
-\newcommand{\preprintabstract}{This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.}
+\newcommand{\preprintabstract}{
+  Both the brain and microbiome of humans develop rapidly in the first years of life,
+  enabling extensive signaling between the gut and central nervous system (dubbed the “microbiome-gut-brain axis”).
+  Emerging evidence implicates gut microorganisms and microbiota composition in cognitive outcomes
+  and neurodevelopmental disorders (e.g., autism),
+  but the influence of gut microbial metabolism on typical neurodevelopment has not been explored in detail.
+  We investigated the relationship of the microbiome with the neuroanatomy and cognitive function
+  of \hl{XXX} healthy children,
+  demonstrating that differences in gut microbial taxa and gene functions
+  are associated with the size of brain regions and with overall cognitive function.
+  Many species, including \hl{XXX} and \hl{XXX}, were associated with higher cognitive function, 
+  while some species such as \hl{XXX} was more commonly found in children with low cognitive scores.
+  Microbial enzymes involved in the metabolism of neuroactive compounds such as glutamate and GABA,
+  were also associated with structure of the brain \hl{eg XXX and YYY} and with overall cognitive function.
+}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE KEYWORDS %%%%%%%%%%%%%%%%%
-\newcommand{\preprintkeywords}{KeywordOne; KeywordTwo; KeywordThree; KeywordFour; KeywordFive}
+\newcommand{\preprintkeywords}{Microbiome; Cognitition; Metagenomes}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section*{\centering Introduction}
+\section*{Introduction}
 
-This is an example of text.
-This realization is especially important because it can flip around our expectations about which species expand fast, and how to manage them.
-We tend to think of initial colonization and long-term abundance as two independent axes of variation among species or indeed as two ends of a spectrum, in the classic competition-colonization tradeoff \parencite{levins1971regional}.
-When both play into invasion speed, good dispersers might not outrun good competitors.
-This is useful knowledge, whether we want to contain an invasion or secure a reintroduction.
-\\
+This is an example of text. Use \textcite{levins1971regional} for inline or \parencite{crooks2004spatial} for normal. \\ % adds between-paragraph space
 
-In their study "When higher carrying capacities lead to faster propagation", \textcite{levins1971regional} combine mathematical analysis, Individual-Based simulations and experiments to show that various mechanisms can cause pushed fronts, whose speed increases with the carrying capacity K of the species.
-Rather than focus on one particular angle, the authors endeavor to demonstrate that this qualitative effect appears again and again in a variety of settings.
-\\
+Here's some other text.
 
-It is perhaps surprising that this notable and general connection between K and invasion speed has managed to garner so little fame in ecology.
-A large fraction of the literature employs the venerable Fisher-KPP reaction-diffusion model, which combines local logistic growth with linear diffusion in space.
-This model has prompted both considerable mathematical developments \parencite{crooks2004spatial} and many applications to modelling real invasions \parencite{shigesada1997biological}.
-But it only allows pulled fronts, driven by the small populations at the edge of a species range, with a speed that depends only on their initial growth rate r.\\
+\section*{Material and methods}
 
-\section*{\centering Material and methods}
+\subsection*{Human Subjects}
 
-This classic setup is, however, singular in many ways.
-\textcite{levins1971regional} use it as a null model, and introduce three mechanisms or factors that each ensure a 1 role of K in invasion speed, while giving less importance to the pioneers at the border.
-\\
+Data used in this study were drawn from the ongoing longitudinal RESONANCE study
+of healthy and neurotypical brain and cognitive development,
+based at Brown University in Providence, RI, USA.
+RESONANCE is part of the NIH initiative Environmental influences on Child Health Outcomes (ECHO) \cite{Forrest2018-ud,Gillman2018-om},
+a longitudinal observational study of healthy and neurotypical brain development
+that spans the fetal and infant to adolescent life stages,
+combining neuroimaging (magnetic resonance imaging, MRI), neurocognitive assessments, bio-specimen analyses, subject genetics,
+environmental exposures such as lead, and rich demographic, socioeconomic, family and medical history information.
+From the RESONANCE cohort, 344 typically-developing children
+between the ages of 28 days and 15 years old were selected for analysis in this study.
 
-Two factors, the Allee effect and demographic stochasticity, make small edge populations slower to grow or less likely to survive.
-These two factors are studied theoretically, and to make their claims stronger, the authors stack the deck against K.
-When generalizing equations or simulations beyond the null case, it is easy to obtain functional forms where the parameter K does not only play the role of equilibrium carrying capacity, but also affects dynamical properties such as the maximum or mean growth rate.
-In that case, it can trivially change the propagation speed, without it meaning anything about the role of established populations behind the front.
-\textcite{levins1971regional} avoid this pitfall by disentangling these effects, at the cost of slightly more peculiar expressions, and show that varying essentially nothing but the carrying capacity can still impact the speed of the invasion front.
-\\
+General participant demographics are provided in Tables \ref{tab:demographics} and \ref{tab:agestats} and Figure \ref{fig:data}.
+Complete metadata are available in Data Records (see below), with children being representative of the RI population.
+As a broad background, children in the RESONANCE cohort were born full-term (>37 weeks gestation)
+with height and weight normal for gestational age, and from uncomplicated singleton pregnancies.
+Children with known major risk factors for developmental abnormalities at enrollment were excluded.
+In addition to screening at the time of enrollment,
+on-going screening for worrisome behaviors using validated tools was performed
+to identify at-risk children and remove them from subsequent analysis.
 
-\section*{\centering Results}
+Exclusion criteria included: \emph{in utero} exposure to alcohol, cigarette or illicit substance exposure;
+preterm (<37 wks gestation) birth; small for gestational age or less than 1500 g; fetal ultrasound abnormalities;
+preeclampsia, high blood pressure, or gestational diabetes; 5 minute APGAR scores <8;
+NICU admission; neurological disorder (e.g., head injury resulting in loss of consciousness, epilepsy);
+and psychiatric or learning disorder (including maternal depression) in the infant, parents, or siblings requiring medication in the year prior to pregnancy.
+
+Demographic and other non-biospecimen data such as race and ethnicity, parental education and occupation,
+feeding behavior (breast- and formula-feeding), child weight and height,
+were collected through questionnaires or direct examination as appropriate.
+All data were collected at every assessment visit.
+All procedures for this study were approved by the local institutional review board at Rhode Island Hospital,
+and all experiments adhered to the regulation of the review board.
+Written informed consent was obtained from all parents or legal guardians of enrolled participants.
+
+\subsection*{Cognitive Assessments}
+
+Overall cognitive function was assessed using age-appropriate methods.
+For children from birth to 30 months, we used an Early Learning Composite
+as assessed via the Mullen Scales of Early Learning (MSEL) \cite{Mullen1995-ty},
+a standardized and population-normed tool for assessing fine and gross motor,
+expressive and receptive language, and visual reception functioning in children from birth through 68 months of age.
+
+The third edition of the Bayley Scales of Infant and Toddler Development \cite{Bayley2006-wm}
+is a standard series of measures used primarily to assess the development of infants and toddlers,
+ranging from 1 to 42 months of age.
+
+The Wechsler Intelligence Quotient for Children (WISC) \cite{Wechsler2012-mi}
+is an individually administered standard intelligence test for children aged 6 to 16 years.
+It derives a full scale intelligence quotient (IQ) score, which we used to assess overall cognitive functioning.
+The fourth edition of the Wechsler Preschool and Primary Scale of Intelligence (WPPSI-IV) \cite{Wechsler2012-mi}
+is an individually administered standard intelligence test for children aged 2 years 6 months to 7 years 7 months,
+trying to meet the increasing need for the assessment of preschoolers.
+Just as the WISC, it derives a full scale IQ score, which we used to assess overall cognitive functioning.
+
+\subsection*{Stool Sample Collection and Sequencing}
+
+Stool samples (n=493) were collected by parents in OMR-200 tubes (OMNIgene GUT, DNA Genotek, Ottawa, Ontario, Canada),
+immediately stored on ice, and brought within 24 hrs to the lab in RI where they were immediately frozen at -80 $^{\circ}$C.
+Stool samples were not collected if the subject had taken antibiotics within the last two weeks.
+DNA extraction was performed at Wellesley College (Wellesley, MA).
+Nucleic acids were extracted from stool samples using the RNeasy PowerMicrobiome kit
+automated on the QIAcube (Qiagen, Germantown, MD), excluding the DNA degradation steps.
+Extracted DNA was sequenced at the Integrated Microbiome Resource (IMR, Dalhousie University, NS, Canada)
+
+Shotgun metagenomic sequencing was performed on all samples.
+A pooled library (max 96 samples per run) was prepared using the Illumina Nextera Flex Kit for MiSeq and NextSeq from 1 ng of each sample.
+Samples were then pooled onto a plate and sequenced
+on the Illumina NextSeq 550 platform using 150+150 bp paired-end “high output” chemistry,
+generating ~400 million raw reads and ~120 Gb of sequence per plate.
+
+\hl{Are we still planning to include 16S data here?}
+
+For sequencing 16S rRNA gene amplicons,
+the V4-V5 region of the 16S ribosomal RNA gene was sequenced according to the protocol
+described by Comeau et al. \cite{Comeau2017-jg}.
+Briefly, the V4-V5 region was amplified once using the Phusion High-Fidelity DNA polymerase
+(ThermoFisher Scientific, Waltham, MA) and universal bacterial primers
+515F: 5’-GTGYCAGCMGCCGCGGTAA-3’ and 926R: 5’-CCGYCAATTYMTTTRAGTTT-3’ \cite{Parada2016-uz,Walters2016-fi}.
+These primers had appropriate Illumina adapters and error-correcting barcodes unique to each sample
+to allow up to 380 samples to be simultaneously run per single flow cell.
+After being pooled into a single library and quantified fluorometrically,
+samples were cleaned-up and normalized using the high-throughput Charm Biotech Just-a-Plate 96-well Normalization Kit (Charm Biotech, Cape Girardeau, MO).
+The normalized samples were sequenced on the Illumina MiSeq platform (Illumina, San Diego, CA)
+using 300+300 bp paired-end V3 chemistry, producing ~55,000 raw reads per sample.
+
+\subsection*{Computational Analysis}
+
+Shotgun metagenomic sequences were analyzed using the bioBakery suite of computational tools \cite{McIver2018-yc}.
+First, \verb|KneadData| (v0.7.7) was used to perform quality control of raw sequence reads,
+such as read trimming and removal of reads matching a human genome reference.
+Next, \verb|MetaPhlAn| (v3.0.7, using database \verb|mpa_v30_CHOCOPhlAn_201901|) was used to generate taxonomic profiles
+by aligning reads to a reference database of marker genes.
+Finally, \verb|HUMAnN| (v3.0.0a4) was used to functionally profile the metagenomes.
+
+Raw amplicon sequences were profiled using Quantitative Insights in Microbial Ecology 2 (QIIME2) v2021.2.0 \cite{Bolyen2019-qq}.
+Briefly, primers flanking V4-V5 were removed from fastq reads using the cutadapt (v3.2) QIIME2 plugin \cite{Martin2011-zv}.
+Raw sequence reads of samples from enrichment cultures were denoised, filtered and clustered into amplicon sequence variants (ASVs) using the Divisive Amplicon Denoising Algorithm (DADA2) plugin in QIIME 2 \cite{Callahan2016-ol}.
+After denoising and filtering, 16270 total sequences were recovered with a mean length of 373 bases (270-465, standard deviation 13.21).
+Taxonomy was assigned to each ASV using a Naïve-Bayes classifier compared against SILVA v.138 reference database \cite{Yilmaz2013-rj,Quast2013-hc}
+trained on the 515F-806R region of the 16S rRNA gene \cite{Bokulich2018-dv}.
+
+Additional data processing, generation of summary statistics, 
+and generation of plots was performed using the julia programming language \cite{Bezanson2017-ud}.
+See Code Availability section for additional details.
+
+\section*{Results}
 The third factor, density-dependent dispersal, makes small populations less prone to disperse.
@@
 
-\section*{\centering Discussion}
+\section*{Discussion}
 Overall, this study carries a simple and clear message, supported by valuable contributions from different angles.
@@
 
-\section*{\centering Acknowledgements}
+\section*{Acknowledgements}
 This is your acknowledgments. Version \version~ of this article has been peer-reviewed and recommended by 
@@
 
-\section*{\centering Conflict of interest disclosure}
+\section*{Conflict of interest disclosure}
 The authors of this preprint declare that they have no financial conflict of interest with the content of this article. XXX and XXX are recommenders for PCI XXX
 
-\section*{\centering Data, script and code availability}
+\section*{Data, script and code availability}
 Data are available online: link or DOI of the webpage hosting the data
 
-\section*{\centering Supplementary information availability}
+\section*{Supplementary information availability}
 Script and codes are available online: link or DOI of the webpage hosting the script and codes

### 2022-04-25 manuscript/main.tex
try compiling latex

@@
 \section*{Results}
-The third factor, density-dependent dispersal, makes small populations less prone to disperse.
-It is well established empirically (see Fig~\ref{Fig1}) and theoretically that various biological mechanisms, from collective organization to behavioral switches, can prompt organisms in denser populations to disperse more, e.g.
-in such a way as to escape competition \parencite{matthysen2005density}.
-The authors demonstrate how this effect induces a link between carrying capacity and invasion speed, both theoretically and in a dispersal experiment on the parasitoid wasp, \emph{Trichogramma chilonis}.\\
+Figure example (see Fig~\ref{Fig1}). \\
+
+\subsection*{Outline}
+
+\begin{itemize}
+  \item Intial description: Cohort size, demographics, data types
+  \item Basic analysis: Summary statistics, ordinations, permanovas, mantel tests
+  \item Exploration of taxonomic profiles + cogscores / brain, cross-sectional. Linear models. \hl{Is this necessary??}
+  \item Try HAllA on multivariate data?
+  \item Predictive model (RF) - taxonomic profiles + cogscores.
+  \item Taxonomic profiles + brain structure
+  \item Gene functions + metabolomes
+\end{itemize}
 
@@
 \section*{Conflict of interest disclosure}
-The authors of this preprint declare that they have no financial conflict of interest with the content of this article. XXX and XXX are recommenders for PCI XXX
+The authors of this preprint declare that they have no financial conflict of interest with the content of this article. \hl{XXX} and \hl{XXX} are recommenders for PCI \hl{XXX}
 

### 2022-05-11 slides/subsections/metabolites.tex
create beamer presentation with figures

% -----------------------------------------------------------------------------

\begin{frame}{GABA and Glutamate}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.5\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/gaba-glutamate.png}
        \end{figure}

        \column{.4\textwidth} % Right column and width

        \textbf{Details}
        \begin{enumerate}
            \item Comparing mother and child samples
            \item Mean-centered and compositional
            \item Plotting $log_2(relative\_abundance)$
        \end{enumerate}

    \end{columns}

\end{frame}

% -----------------------------------------------------------------------------

\begin{frame}{GABA and Glutamate, by age}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.5\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/gaba-glutamate_age.png}
        \end{figure}

        \column{.4\textwidth} % Right column and width

        \textbf{Details}
        \begin{enumerate}
            \item Comparing mother and child samples
            \item Age in months colored, moms are gray
            \item Plotting $log_2(relative\_abundance)$
        \end{enumerate}

    \end{columns}

\end{frame}

% -----------------------------------------------------------------------------

\begin{frame}{Metabolites PCoA}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.5\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/metabolites_pcoa.png}
        \end{figure}

        \column{.4\textwidth} % Right column and width

        \textbf{Details}
        \begin{enumerate}
            \item PCoA based on BC dissimilartiy
        \end{enumerate}

    \end{columns}

\end{frame}

% -----------------------------------------------------------------------------

### 2022-05-11 slides/subsections/samples.tex
create beamer presentation with figures

% -----------------------------------------------------------------------------

\begin{frame}{Mother-child dyads}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.6\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/dyads.png}
        \end{figure}

        \column{.4\textwidth} % Right column and width

        \textbf{Details}
        \begin{enumerate}
            \item \textbf{Left} - Child samples that have at least 1 associated maternal sample
            \item \textbf{Right} - Maternal samples that have at least 1 associated child sample
        \end{enumerate}

    \end{columns}

\end{frame}

% -----------------------------------------------------------------------------

\begin{frame}{Upset - Brain scans + stool}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.6\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/upset_scan_stool.pdf}
        \end{figure}

        \column{.4\textwidth} % Right column and width

        \textbf{Details}
        \begin{enumerate}
            \item Overlap of sample collection and brain scans
            \item "Prev stool" means that there's stool sample collected prior to timepoint
            \item Not 100\% sure of this code - needs review
        \end{enumerate}

    \end{columns}

\end{frame}

% -----------------------------------------------------------------------------

\begin{frame}{Upset - By subject}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.6\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/upset_scan_stool_bysybject.pdf}
        \end{figure}

        \column{.4\textwidth} % Right column and width

        \textbf{Details}
        \begin{enumerate}
            \item Overlap of sample collection and brain scans
            \item Numbers reduced to unique subjects
        \end{enumerate}

    \end{columns}

\end{frame}

% -----------------------------------------------------------------------------

\begin{frame}{Upset - Overlaps with cogscore}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.6\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/upset_score_stool.pdf}
        \end{figure}

        \column{.4\textwidth} % Right column and width

        \textbf{Details}
        \begin{enumerate}
            \item Overlap of sample collection and cognitive score
        \end{enumerate}

    \end{columns}

\end{frame}

% -----------------------------------------------------------------------------
\begin{frame}{Upset - Kids under 3}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.6\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/upset_score_stool_under3.pdf}
        \end{figure}

        \column{.4\textwidth} % Right column and width

        \textbf{Details}
        \begin{enumerate}
            \item Overlap of sample collection and cognitive score for kids under 3
            \item "Under 3" row is all timepoints, regardless of stool collection
        \end{enumerate}

    \end{columns}

\end{frame}

% -----------------------------------------------------------------------------
\begin{frame}{Upset - By subject}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.6\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/upset_scan_stool_bysybject.pdf}
        \end{figure}

        \column{.4\textwidth} % Right column and width

        \textbf{Details}
        \begin{enumerate}
            \item Numbers reduced to unique subjects
        \end{enumerate}

    \end{columns}

\end{frame}

% -----------------------------------------------------------------------------

### 2022-05-11 notes/07-kneaddata_counts.jl

@@ -9,7 +9,7 @@ else
     df = DataFrame()
 
     kneads = readdir("/grace/echo/analysis/biobakery3/links/kneaddata/", join=true)
-    for k in kneads[946:end]
+    for k in kneads
         open(k) do io
             stream = GzipDecompressorStream(io)
             c = count(line-> startswith(line, "@"), eachline(stream))

### 2022-05-11 notes/Y_data_requests.jl

@@ -91,3 +91,95 @@ count(groupby(df, :subject)) do grp
 end
 
 CSV.write("/home/kevin/Desktop/RESONANCE_bp.csv", sort(df, ["subject", "timepoint"]))
+
+## Kirsty + 3rd quarter
+
+using Resonance
+omni, etoh, tps, complete_brain, metabolites, species = startup()
+
+mo3 = @subset tps begin
+    .!ismissing.(:ageMonths)
+    2 .< :ageMonths .< 5
+end
+
+@rsubset! omni :sample in samplenames(species)
+unique!(omni, [:subject, :timepoint])
+mo3 = leftjoin(mo3, select(omni, [:subject, :timepoint, :sample]), on=[:subject, :timepoint])
+@rsubset! mo3 !ismissing(:sample)
+mo3spec = species[:, in.(samplenames(species), Ref(mo3.sample))]
+
+set!(mo3spec, mo3)
+
+#-
+
+ginisimpson!(mo3spec)
+prev = prevalence(mo3spec) |> vec
+count(prev .> 0.56)
+#- 
+
+using CairoMakie
+using AlgebraOfGraphics
+
+#-
+
+fig = Figure(; resolution = (1100, 850))
+ax1 = Axis(fig[1,1]; title="Species prevalence", xlabel="prevalence", ylabel="number of taxa")
+prevhist = hist!(ax1, prev)
+
+ax2 = Axis(fig[2,1]; title="Age and diversity", xlabel="Age (months)", ylabel="Gini-Simpson Diversity")
+scatter!(ax2, get(mo3spec, :ageMonths), get(mo3spec, :ginisimpson))
+
+topprev = featurenames(mo3spec)[prev .> 0.55]
+
+
+lcol = GridLayout()
+fig.layout[1:2, 2] = lcol
+
+df = vcat(collect(
+            DataFrame("abundance" => vec(abundances(mo3spec[Regex(topprev[i]), :])),
+                      "ageMonths" => get(mo3spec, :ageMonths),
+                      "species"   => fill(topprev[i], nsamples(mo3spec))
+                      ) for i in eachindex(topprev)
+            )...
+)
+
+
+pl = data(df) * mapping(:ageMonths=> "Age (months)", :abundance=> "Relative abundance"; layout=:species)
[…]

### 2022-05-11 src/airtable.jl

@@ -29,7 +29,7 @@ function airtable_metadata()
 
     df = DataFrame()
     
-    for samp in Term.track(samples; description="[blue bold]Building DataFrame[/blue bold]")
+    @track for samp in samples
         # skip anything not in the "ECHO" project
         "rec8zdkF1pchFNiJC" ∈ get(samp, :Project, []) || continue
 

### 2022-05-11 slides/subsections/brain_figures.tex

% -----------------------------------------------------------------------------

\begin{frame}{Brain PCoA}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.45\textwidth} % Left column and width
        \textbf{Details}
        \begin{enumerate}
            \item PCA, Euclidean distance of brain regions, normalized by total brain size
            \item All kids with processed brain scans
            \item Top: colored by age, first 2 axes
            \item Bottom: PCA1 x Age, colored by cogscore

        \end{enumerate}

        \column{.5\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/brain_pcoa.png}
        \end{figure}

    \end{columns}
\end{frame}

% -----------------------------------------------------------------------------

\begin{frame}{Cognitive score summaries}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.45\textwidth} % Left column and width

        \begin{figure}
            \includegraphics[width=1\linewidth]{../figures/cogScore_quartiles.png}
        \end{figure}

        \column{.5\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/cogscores_hist.pdf}
        \end{figure}

    \end{columns}

    \textbf{Details}
    \begin{enumerate}
        \item Quartiles calculated based on all measurements
    \end{enumerate}

\end{frame}

% -----------------------------------------------------------------------------

### 2022-05-23 notes/06-plot_intersections.jl
update slides /figures

@@ -135,7 +135,7 @@ hist!(scan_age_young, subset(tps, AsTable([:ageMonths, :has_segmentation]) => By
 Label(fig[1, 0], text = "All kids", tellheight=false)
 Label(fig[2, 1], text = "Young kids", tellheight=false)
 
-save("figures/age_hists.pdf")
+save("figures/age_hists.pdf", fig)
 fig
 
 ##
@@ -157,6 +157,7 @@ tightlimits!(ass_bydate, Left(), Right(), Bottom())
 
 linkyaxes!(stools_bydate, ass_bydate)
 
+save("figures/collection_over_time.pdf", fig)
 fig
 ##
 

### 2022-05-23 notes/07-kneaddata_counts.jl

@@ -1,46 +1,6 @@
 using Resonance
-using CodecZlib
-using DataFramesMeta
 
-
-if isfile("data/read_counts.csv")
-    df = CSV.read("data/read_counts.csv", DataFrame)
-else
-    df = DataFrame()
-
-    kneads = readdir("/grace/echo/analysis/biobakery3/links/kneaddata/", join=true)
-    for k in kneads
-        open(k) do io
-            stream = GzipDecompressorStream(io)
-            c = count(line-> startswith(line, "@"), eachline(stream))
-            push!(df, (file=k, count=c))
-        end
-    end
-
-    @transform!(df, @byrow :sample = first(split(basename(:file), "_")))
-    CSV.write("data/read_counts.csv", df)
-end
-
-##
-
-human = DataFrame() 
-
-for (root, dirs, files) in walkdir("/grace/echo/analysis/biobakery3/")
-    contains(root, "kneaddata") || continue
-    for f in files
-        (contains(f, "hg37") || contains(f, "sapiens")) || continue
-        s = first(split(f, "_"))
-        try
-            open(joinpath(root, f)) do io
-                stream = GzipDecompressorStream(io)
-                c = count(line-> startswith(line, "@"), eachline(stream))
-                push!(human, (sample=s, count=c))
-            end
-        catch e
-            @error e
-        end
-    end
-end
+kneadread = Resonance.load_knead()
 
 ##
 
@@ -52,6 +12,8 @@ human = @chain human begin
 end
 
 hist(human.count_sum .* 150)
+
+
 using Statistics
 
 median(human.count_sum .* 150)

### 2022-05-23 src/Resonance.jl

@@ -28,6 +28,8 @@ using Term.progress
 using Arrow
 using Tables
 using SparseArrays
+using CodecZlib
+using FilePaths
 
 @reexport using BiobakeryUtils
 @reexport using DataFrames
@@ -42,5 +44,6 @@ include("genefamilies.jl")
 include("plotting.jl")
 include("brain.jl")
 include("startup.jl")
+include("kneaddata.jl")
 
 end

### 2022-05-23 src/kneaddata.jl

@@ -1,0 +1,32 @@
+
+function _read_knead_read_counts(dir=ENV["ANALYSIS_FILES"])
+    fileinfo = NamedTuple[]
+
+    paired_paths = Iterators.filter(walkpath(PosixPath(dir))) do file 
+        "kneaddata" in file.segments || return false
+        contains(filename(file), r"read_counts") || return false
+        return true
+    end
+
+    df = DataFrame()
+
+    for path in paired_paths
+        @info path
+        batch = path.segments[findfirst(s-> contains(s, "batch"), path.segments)]
+        logs = CSV.read(path, DataFrame; stringtype=String, missingstring=["", "NA"])
+        rename!(col-> replace(col, "hg37dec_v0.1"=>"Homo_sapiens"), logs)
+        subset!(logs, "raw pair1"=>ByRow(!ismissing))
+        foreach(col-> logs[!, col] = coalesce.(logs[!, col], 0), names(logs)[2:end])
+        logs[!, :batch] .= batch
+        append!(df, logs)
+    end
+
+    CSV.write("data/read_counts.csv", df)
+end
+
+##
+
+function load_knead(dir=ENV["ANALYSIS_FILES"])
+    !isfile("data/read_counts.csv") && _read_knead_read_counts(dir)
+    CSV.read("data/read_counts.csv", DataFrame)
+end

### 2022-05-23 src/startup.jl

@@ -97,7 +97,7 @@ function _gentps()
     
     select!(tps, ["subject", "timepoint", mainmeta..., brainmeta...])
     rename!(tps, Dict(k=> replace(k, "-"=>"_") for k in brainmeta))
-    foreach(i-> (brainmeta[i] = replace(brainmeta[i], "-"=>"_")), eachindex(brainmeta))
+    brainmeta = map(i-> replace(brainmeta[i], "-"=>"_"), eachindex(brainmeta))
     
     complete_brain = completecases(tps[:, brainmeta])
     return tps, complete_brain

### 2022-06-01 manuscript/main.tex
simplify manuscript template

--- 2022-04-25 manuscript/main.tex
@@
 \documentclass[a4paper]{article}
-%%%%%%%%%   CHOOSE A PCI   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\newcommand{\PCI}{Peer Community In Zoology}
-%\newcommand{\PCI}{Peer Community In Ecology}
-%\newcommand{\PCI}{Peer Community In Evolutionary Biology}
-%\newcommand{\PCI}{Peer Community In Paleontology}
-%\newcommand{\PCI}{Peer Community In Archaeology}
-%\newcommand{\PCI}{Peer Community In Animal Science}
-\newcommand{\PCI}{Peer Community In Genomics}
-%\newcommand{\PCI}{Peer Community In Neuroscience}
-%\newcommand{\PCI}{Peer Community In Mathematical and Computational Biology}
-%\newcommand{\PCI}{Peer Community In Forest and Wood Sciences}
-%\newcommand{\PCI}{Peer Community In Ecotoxicology and Environmental Chemistry}
-%\newcommand{\PCI}{Peer Community In Infections}
-%\newcommand{\PCI}{Peer Community In Registered Reports}
-%\newcommand{\PCI}{Peer Community In Network Science}
-%\newcommand{\PCI}{Peer Community In Microbiology}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\input{preamble.tex}
 
-\input{preambule.tex}
-
-%%%%%%%%%   CHOOSE A BADGE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%Comment lines 131 and/or 133 in preambule.tex if you don't use data and/or code in your preprint
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%   SET THE TITLE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\preprinttitle}{This is the title of the preprint}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%   SET THE version of the preprint. Replace X by a number %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\version}{X}
+\title{Gut microbes and their genes or something else}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%  SET THE LIST OF AUTHORS WITH CORRESPONDING AFFILIATIONS  use \& before last author %%%%%%%%%%%%%%%%%%%%
-\newcommand{\listauthors}{\raggedright 
-FirstnameOne FamilynameOne\textsuperscript{1}, \space
-FirstnameTwo FamilynameTwo\textsuperscript{1}, \&
-FirstnameThree FamilynameThree\textsuperscript{2}
+\author{Kevin S. Bonham\textsuperscript{1}, \space
+Guilherme Fahur Bottino\textsuperscript{1}, 
+Firstname Lastname\textsuperscript{2}, \&
+Vanja Klepac-Ceraj\textsuperscript{1}
 %\fourthauthor \textsuperscript{i}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE LIST OF AFFILIATIONS  %%%%%%%%%%%%%%%%%%%%
-\newcommand{\listinstitutions}{
-\textsuperscript{1} DeptOne, InstitutionOne -- TownOne, CountryOne 
-\\
-\textsuperscript{2} DeptTwo, InstitutionTwo -- TownTwo, CoutryTwo
-%\\
-%\textsuperscript{3} \thirdinstitution
-%etc
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE DATE OF UPLOAD on the preprint server %%%%%%%%%%%%%%
-\newcommand{\datepub}{ddth Month yyyy}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE RECOMMENDER(s) NAME(s) %%%%%%%%%%%%%%
-\newcommand{\recommender}{FirstName FamilyName}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE DOI of the RECOMMENDATION %%%%%%%%%%%%%%
-\newcommand{\DOIrecommendation}{xxx/xxx}
-%%%%%%%%%%%%% for example 10.24072/pci.mcb.100003 %%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE REVIEWERS' NAMES IF KNOWN and/or X anonymous reviewers %%%%%%%%%%%%%%
-\newcommand{\reviewers}{FirstName FamilyName and two anonymous reviewer}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%  SET THE 'CITE AS' OF YOUR PREPRINT. paste the line we sent you by Email (XXX the "cite as") in place of the xxx %%%%%%%%%%%%
-\newcommand{\citeas}{xxx}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE 'CORRESPONDENCE TO' %%%%%%%%%%%%%%%%%
-\newcommand{\email}{mail.mail@mail.xx}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE ABSTRACT %%%%%%%%%%%%%%%%%
-\newcommand{\preprintabstract}{This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.
-This is the abstract.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE KEYWORDS %%%%%%%%%%%%%%%%%
-\newcommand{\preprintkeywords}{KeywordOne; KeywordTwo; KeywordThree; KeywordFour; KeywordFive}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{document}
-\beginingpreprint
+
+\begin{abstract}
+\input{sections/abstract.tex}
+\end{abstract}
 
@@
 
-\section*{\centering Introduction}
+\section*{Introduction}
 
-This is an example of text.
-This realization is especially important because it can flip around our expectations about which species expand fast, and how to manage them.
-We tend to think of initial colonization and long-term abundance as two independent axes of variation among species or indeed as two ends of a spectrum, in the classic competition-colonization tradeoff \parencite{levins1971regional}.
-When both play into invasion speed, good dispersers might not outrun good competitors.
-This is useful knowledge, whether we want to contain an invasion or secure a reintroduction.
-\\
+\input{sections/intro}
 
-In their study "When higher carrying capacities lead to faster propagation", \textcite{levins1971regional} combine mathematical analysis, Individual-Based simulations and experiments to show that various mechanisms can cause pushed fronts, whose speed increases with the carrying capacity K of the species.
-Rather than focus on one particular angle, the authors endeavor to demonstrate that this qualitative effect appears again and again in a variety of settings.
-\\
+\section*{Material and methods}
 
-It is perhaps surprising that this notable and general connection between K and invasion speed has managed to garner so little fame in ecology.
-A large fraction of the literature employs the venerable Fisher-KPP reaction-diffusion model, which combines local logistic growth with linear diffusion in space.
-This model has prompted both considerable mathematical developments \parencite{crooks2004spatial} and many applications to modelling real invasions \parencite{shigesada1997biological}.
-But it only allows pulled fronts, driven by the small populations at the edge of a species range, with a speed that depends only on their initial growth rate r.\\
+\input{sections/methods}
 
-\section*{\centering Material and methods}
+\section*{Results}
 
-This classic setup is, however, singular in many ways.
-\textcite{levins1971regional} use it as a null model, and introduce three mechanisms or factors that each ensure a 1 role of K in invasion speed, while giving less importance to the pioneers at the border.
-\\
-
-Two factors, the Allee effect and demographic stochasticity, make small edge populations slower to grow or less likely to survive.
-These two factors are studied theoretically, and to make their claims stronger, the authors stack the deck against K.
-When generalizing equations or simulations beyond the null case, it is easy to obtain functional forms where the parameter K does not only play the role of equilibrium carrying capacity, but also affects dynamical properties such as the maximum or mean growth rate.
-In that case, it can trivially change the propagation speed, without it meaning anything about the role of established populations behind the front.
-\textcite{levins1971regional} avoid this pitfall by disentangling these effects, at the cost of slightly more peculiar expressions, and show that varying essentially nothing but the carrying capacity can still impact the speed of the invasion front.
-\\
-
-\section*{\centering Results}
-The third factor, density-dependent dispersal, makes small populations less prone to disperse.
-It is well established empirically (see Fig~\ref{Fig1}) and theoretically that various biological mechanisms, from collective organization to behavioral switches, can prompt organisms in denser populations to disperse more, e.g.
-in such a way as to escape competition \parencite{matthysen2005density}.
-The authors demonstrate how this effect induces a link between carrying capacity and invasion speed, both theoretically and in a dispersal experiment on the parasitoid wasp, \emph{Trichogramma chilonis}.\\
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  EXAMPLE OF A FIGURE   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[h]
-  \captionsetup{justification=centering}
-  \caption{A nice picture of bidule.}
-  \centering
-  \includegraphics[scale=0.5]{fig.png}
-  \label{Fig1}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section*{\centering Discussion}
-Overall, this study carries a simple and clear message, supported by valuable contributions from different angles.
-Although some sections are clearly written for the theoretical ecology crowd, this article has something for everyone, from the stray physicist to the open-minded manager.
-The collaboration between theoreticians and experimentalists, while not central, is worthy of note.
-Because the narrative of this study is the variety of mechanisms that can lead to the same qualitative effect, the inclusion of various approaches is not a gimmick, but helps drive home its main message.
-The work is fairly self-contained, although one could always wish for further developments, especially in the direction of more quantitative testing of these mechanisms.
-\\
-
-In conclusion, \textcite{levins1971regional} effectively convey the widely relevant message that, for some species, invading is not just about the destination, it is about the many offspring one makes along the way.
-\\
-
-\section*{\centering Acknowledgements}
-This is your acknowledgments. Version \version~ of this article has been peer-reviewed and recommended by 
-\emph{\PCI} (\DOIrecommendationlink)
-
-\section*{\centering Conflict of interest disclosure}
-The authors of this preprint declare that they have no financial conflict of interest with the content of this article. XXX and XXX are recommenders for PCI XXX
-
-\section*{\centering Data, script and code availability}
-Data are available online: link or DOI of the webpage hosting the data
-
-\section*{\centering Supplementary information availability}
-Script and codes are available online: link or DOI of the webpage hosting the script and codes
-
-\titleformat*{\section}{\bfseries\Large\centering}
-
-\printbibliography[notcategory=ignore]
-
-\section*{Annexes or Supplementary Information}
-This is your annex 1
-
-This is your annex 2
+\input{sections/results}
 

### 2022-06-01 manuscript/sections/abstract.tex
simplify manuscript template

Both the brain and microbiome of humans develop rapidly in the first years of life,
enabling extensive signaling between the gut and central nervous system (dubbed the “microbiome-gut-brain axis”).
Emerging evidence implicates gut microorganisms and microbiota composition in cognitive outcomes
and neurodevelopmental disorders (e.g., autism),
but the influence of gut microbial metabolism on typical neurodevelopment has not been explored in detail.
We investigated the relationship of the microbiome with the neuroanatomy and cognitive function
of \hl{XXX} healthy children,
demonstrating that differences in gut microbial taxa and gene functions
are associated with the size of brain regions and with overall cognitive function.
Many species, including \hl{XXX} and \hl{XXX}, were associated with higher cognitive function, 
while some species such as \hl{XXX} was more commonly found in children with low cognitive scores.
Microbial enzymes involved in the metabolism of neuroactive compounds such as glutamate and GABA,
were also associated with structure of the brain \hl{eg XXX and YYY} and with overall cognitive function.

### 2022-06-01 notes/Z-Khula-pres.jl

@@ -1,0 +1,171 @@
+using Resonance
+omni, etoh, tps, complete_brain, metabolites, species = startup()
+genes = Resonance.read_gfs_arrow()
+
+specu1 = species[:, [!ismissing(a) && 2 < a < 10 for a in get(species, :ageMonths)]]
+shannon!(specu1)
+
+set!(genes, leftjoin(tps, select(omni, [:sample, :subject, :timepoint]), on = [:subject, :timepoint]))
+genesu1 = genes[:, [!ismissing(a) && 2 < a < 10 for a in get(genes, :ageMonths)]]
+shannon!(genesu1)
+
+set!(metabolites, leftjoin(tps, select(etoh, [:sample, :subject, :timepoint]), on = [:subject, :timepoint]))
+metabolitesu1 = metabolites[:, [!ismissing(a) && 2 < a < 10 for a in get(metabolites, :ageMonths)]]
+
+pcspec = pcoa(specu1)
+
+pcgenes = pcoa(genesu1)
+#-
+
+using CairoMakie
+using AlgebraOfGraphics
+
+#-
+
+fig = Figure(resolution=(1000, 1000))
+ax1 = Axis(fig[1,1]; ylabel=mdsaxis(pcspec, 2))
+ax2 = Axis(fig[2,1]; xlabel=mdsaxis(pcspec, 1), ylabel=mdsaxis(pcspec, 2))
+ax3 = Axis(fig[1,2]; ylabel=mdsaxis(pcgenes, 2))
+ax4 = Axis(fig[2,2]; xlabel=mdsaxis(pcgenes, 1), ylabel=mdsaxis(pcgenes, 2))
+
+sc1 = scatter!(ax1, loadings(pcspec, 1), loadings(pcspec,2);
+                    xlabel=mdsaxis(pcspec, 1),
+                    ylabel=mdsaxis(pcspec, 2), 
+                    color=get(specu1, :ageMonths)
+)
+sc2 = scatter!(ax2, loadings(pcspec, 1), loadings(pcspec,2);
+                    color=get(specu1, :shannon),
+                    colormap=:plasma
+)
+
+sc3 = scatter!(ax3, loadings(pcgenes, 1), loadings(pcgenes,2);
+                    xlabel=mdsaxis(pcgenes, 1),
+                    ylabel=mdsaxis(pcgenes, 2), 
+                    color=get(genesu1, :ageMonths)
+)
[…]

### 2022-06-01 src/Resonance.jl

@@ -9,6 +9,7 @@ export airtable_metadata,
        pull_row,
        loadings,
        varexplained,
+       mdsaxis,
        commonname,
        countmap,
        codebreastfeeding!,

### 2022-06-01 manuscript/preamble.tex
add materials / methods
Generate presentation for Khula that includes only young kids (#36)
* start khula presentation

* simplify manuscript template

--- 2022-04-25 manuscript/preambule.tex
@@
 
-\usepackage[top=7cm,bottom=2.5cm,headheight=120pt,headsep=15pt,left=6cm,right=1.5cm,marginparwidth=4.2cm,marginparsep=0.5cm]{geometry}
-
-\usepackage{marginnote}
-\usepackage{ifthen}
-\reversemarginpar  % sets margin notes to the left
-\usepackage{lipsum} % Required to insert dummy text
-\usepackage{calc}
-\usepackage{siunitx}
 \usepackage{lineno}
 \usepackage{titlesec}
-\usepackage{indentfirst}
 %\usepackage[none]{hyphenat} % use only if there is a problem
@@
 \usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
 % Clean citations with biblatex
@@
 ]{biblatex}
-\DeclareNameAlias{default}{family-given}
-\renewcommand*{\revsdnamepunct}{} % no comma between family and given names
-\renewbibmacro{in:}{%
-  \ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}} % remove 'In:' before journal name
-\DeclareFieldFormat[article]{pages}{#1} % remove pp.
-\AtEveryBibitem{\ifentrytype{article}{\clearfield{number}}{}} % don't print issue numbers
-\DeclareFieldFormat[article, inbook, incollection, inproceedings, misc, thesis, unpublished]{title}{#1} % title without quotes
-\usepackage{csquotes}
-\RequirePackage[english]{babel} % must be called after biblatex
-\addbibresource{refs.bib}
-\DeclareBibliographyCategory{ignore}
-\addtocategory{ignore}{recommendation} % adding recommendation to 'ignore' category so that it does not appear in the References
-% Clickable references. Use \url{www.example.com} or \href{www.example.com}{description} to add a clicky url
-\usepackage{nameref}
-\usepackage[pdfborder={0 0 0}]{hyperref}  % sets link border to white
-\urlstyle{same}
 
-% Include figures
-\usepackage{graphbox}  % loads graphicx ppackage with extended options for vertical alignment of figures
-% Line numbers
-%\usepackage[right]{lineno}
-% Improve typesetting in LaTex
-\usepackage{microtype}
-\DisableLigatures[f]{encoding = *, family = * }
-% Text layout and font (Open Sans)
-\setlength{\parindent}{0.4cm}
-\linespread{1.2}
-\RequirePackage[default,scale=0.90]{opensans}
-% Defining document colors
-\usepackage{xcolor}
-\definecolor{darkgray}{HTML}{808080}
-\definecolor{mediumgray}{HTML}{6D6E70}
-\definecolor{ligthgray}{HTML}{d9d9d9}
-\definecolor{pciblue}{HTML}{74adca}
-\definecolor{opengreen}{HTML}{77933c}
-% Use adjustwidth environment to exceed text width
-\usepackage{changepage}
-% Adjust caption style
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,singlelinecheck=off]{caption}
+% Highlighting
+\usepackage{color}
+\usepackage{soul}
 
 % Headers and footers
-
-% Headers and footers
-\usepackage{fancyhdr}  % custom headers/footers
-\pagestyle{fancy}  % enables customization of headers/footers
-\fancyhfoffset[L]{4.5cm}  % offsets header and footer to the left to include margin
-\renewcommand{\headrulewidth}{\ifnum\thepage=1 0.5pt \else 0pt \fi} % header ruler only on first page
-%\renewcommand{\footrulewidth}{0.5pt}
-\renewcommand{\footrulewidth}{\ifnum\thepage=1 0.5pt \else 0pt \fi}
-% full logo on first page, then no logo on subsequent pages 
-\lhead{\ifnum\thepage=1 \includegraphics[width=13.5cm]{\logoname}\else \fi}  
-\chead{}
-\rhead{}
-
-\lfoot{\ifnum\thepage=1 \scriptsize \textsc{\color{mediumgray}\PCI} \else  \fi}
-\cfoot{}
-\rfoot{}
-% End Headers and footers
-
-% DOI's
-\newcommand{\DOIlink}{\href{https://doi.org/\DOI}{https://doi.org/\DOI}}
-\newcommand{\DOIrecommendationlink}{\href{https://doi.org/\DOIrecommendation}{https://doi.org/\DOIrecommendation}}
-
-%which logo to display
-\ifthenelse{\equal{\PCI}{Peer Community In Zoology}}{\newcommand{\logoname}{logo_PDF_zool.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Ecology}}{\newcommand{\logoname}{logo_PDF_ecology.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Evolutionary Biology}}{\newcommand{\logoname}{logo_PDF_evolbiol.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Animal Science}}{\newcommand{\logoname}{logo_PDF_animsci.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Archaeology}}{\newcommand{\logoname}{logo_PDF_archaeo.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Forest and Wood Sciences}}{\newcommand{\logoname}{logo_PDF_fws.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Genomics}}{\newcommand{\logoname}{logo_PDF_genomics.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Mathematical and Computational Biology}}{\newcommand{\logoname}{logo_PDF_mcb.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Network Science}}{\newcommand{\logoname}{logo_PDF_networksci.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Paleontology}}{\newcommand{\logoname}{logo_PDF_paleo.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Neuroscience}}{\newcommand{\logoname}{logo_PDF_neuro.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Ecotoxicology and Environmental Chemistry}}{\newcommand{\logoname}{logo_PDF_ecotoxenvchem.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Infections}}{\newcommand{\logoname}{logo_PDF_infections.jpg}}{}
-\ifthenelse{\equal{\PCI}{Peer Community In Registered Reports}}{\newcommand{\logoname}{logo_PDF_rr.jpg}}{}
-
-\newcommand{\beginingpreprint}{
-
-\vspace*{0.5cm}
-\begin{flushleft}
-\baselineskip=30pt
-% Margin information
-\marginpar{
-\large\textnormal{\color{pciblue}\\RESEARCH ARTICLE}\\
-\vspace*{0.5pt}
-\\
-\includegraphics[align=c,width=0.5cm]{OAlogoblue.png} \space \large\textbf{\color{pciblue}Open Access}\\
-\\
-\includegraphics[align=c,width=0.5cm]{o_peerreview.png} \space \large\textbf{\color{pciblue}Open Peer-Review}\\
-\\
-\includegraphics[align=c,width=0.5cm]{o_data3.png} \space \large\textbf{\color{pciblue}Open Data}\\
-\\
-\includegraphics[align=c,width=0.5cm]{o_script3.png} \space \large\textbf{\color{pciblue}Open Code}\\
-\\
-\\
-\\
-\\
-\raggedright
-%\vspace*{3.25cm}
-\scriptsize\textbf{Cite as:}\space
-\citeas\\
-\vspace*{0.5cm}
-\textbf{Posted:} \datepub\\
-\vspace*{0.5cm}
-\textbf{Recommender:}\\
-\recommender\\
-\vspace*{0.5cm}
-\textbf{Reviewers:}\\
-\reviewers\\
-\vspace*{0.5cm}
-\textbf{Correspondence:}\\
-\href{mailto:\email}{\email}\\
-
-}
-{\Huge
-\fontseries{sb}\selectfont{\preprinttitle}}
-\end{flushleft}
-\vspace*{0.25cm}
-\begin{flushleft}
-%\marginpar{
-
-%\marginnote{
-%\raggedright
-%\vspace*{3.25cm}
-%\scriptsize\textbf{Cite as:}\space
-%\fullcite{preprint}\\
-%\vspace*{0.5cm}
-%\textbf{Published:} \datepub\\
-%\vspace*{0.5cm}
-%\textbf{Recommender:}\\
-%\recommender\\
-%\vspace*{0.5cm}
-%\textbf{Reviewers:}\\
-%\reviewers\\
-%\vspace*{0.5cm}
-%\textbf{Correspondence:}\\
-%\href{mailto:\email}{\email}\\
-%\vspace*{0.5cm} 
-%\vspace*{3cm}
-%\vspace*{0.2cm}
-%}[0.8cm]
-
-\Large
-\listauthors
-\end{flushleft}
-\bigskip
-{\raggedright
-\listinstitutions}
-% Recommended preprint box
-\begin{flushleft}
-%\noindent
-\fcolorbox{lightgray}{lightgray}{
-\parbox{\textwidth - 2\fboxsep}{
-\centering\large{\fontseries{sb}\selectfont{This article has been peer-reviewed and recommended by\\
-\emph{\PCI} (\DOIrecommendationlink)}}\\
-%\raggedright\large{\fontseries{sb}\selectfont{An article peer-reviewed and recommended by \emph{\PCI}, edited by \recommender \space on the basis of the reviews by \reviewers \space (DOI: \DOIrecommendationlink)}}\\
-%\small \fullcite{preprint}
-}}
-\end{flushleft}
-\vspace*{0.5cm}
-\fcolorbox{pciblue}{pciblue}{
-\parbox{\textwidth - 2\fboxsep}{
-\vspace{0.25cm}
-\textbf{\large{\textsc{Abstract}}}\\
-\preprintabstract\\
-
-\footnotesize{\textbf{\emph{Keywords: }}\preprintkeywords}
-\vspace{0.25cm}}
-}
-\newpage
-\newgeometry{margin=1in}
-}

### 2022-06-01 manuscript/sections/methods.tex
Generate presentation for Khula that includes only young kids (#36)
* start khula presentation

* simplify manuscript template

\subsection*{Human Subjects}

Data used in this study were drawn from the ongoing longitudinal RESONANCE study
of healthy and neurotypical brain and cognitive development,
based at Brown University in Providence, RI, USA.
RESONANCE is part of the NIH initiative Environmental influences on Child Health Outcomes (ECHO) \cite{Forrest2018-ud,Gillman2018-om},
a longitudinal observational study of healthy and neurotypical brain development
that spans the fetal and infant to adolescent life stages,
combining neuroimaging (magnetic resonance imaging, MRI), neurocognitive assessments, bio-specimen analyses, subject genetics,
environmental exposures such as lead, and rich demographic, socioeconomic, family and medical history information.
From the RESONANCE cohort, 344 typically-developing children
between the ages of 28 days and 15 years old were selected for analysis in this study.

General participant demographics are provided in Tables \ref{tab:demographics} and \ref{tab:agestats} and Figure \ref{fig:data}.
Complete metadata are available in Data Records (see below), with children being representative of the RI population.
As a broad background, children in the RESONANCE cohort were born full-term (>37 weeks gestation)
with height and weight normal for gestational age, and from uncomplicated singleton pregnancies.
Children with known major risk factors for developmental abnormalities at enrollment were excluded.
In addition to screening at the time of enrollment,
on-going screening for worrisome behaviors using validated tools was performed
to identify at-risk children and remove them from subsequent analysis.

Exclusion criteria included: \emph{in utero} exposure to alcohol, cigarette or illicit substance exposure;
preterm (<37 wks gestation) birth; small for gestational age or less than 1500 g; fetal ultrasound abnormalities;
preeclampsia, high blood pressure, or gestational diabetes; 5 minute APGAR scores <8;
NICU admission; neurological disorder (e.g., head injury resulting in loss of consciousness, epilepsy);
and psychiatric or learning disorder (including maternal depression) in the infant, parents, or siblings requiring medication in the year prior to pregnancy.

Demographic and other non-biospecimen data such as race and ethnicity, parental education and occupation,
feeding behavior (breast- and formula-feeding), child weight and height,
were collected through questionnaires or direct examination as appropriate.
All data were collected at every assessment visit.
All procedures for this study were approved by the local institutional review board at Rhode Island Hospital,
and all experiments adhered to the regulation of the review board.
Written informed consent was obtained from all parents or legal guardians of enrolled participants.

\subsection*{Cognitive Assessments}

Overall cognitive function was assessed using age-appropriate methods.
For children from birth to 30 months, we used an Early Learning Composite
as assessed via the Mullen Scales of Early Learning (MSEL) \cite{Mullen1995-ty},
a standardized and population-normed tool for assessing fine and gross motor,
expressive and receptive language, and visual reception functioning in children from birth through 68 months of age.

The third edition of the Bayley Scales of Infant and Toddler Development \cite{Bayley2006-wm}
is a standard series of measures used primarily to assess the development of infants and toddlers,
ranging from 1 to 42 months of age.

The Wechsler Intelligence Quotient for Children (WISC) \cite{Wechsler2012-mi}
is an individually administered standard intelligence test for children aged 6 to 16 years.
It derives a full scale intelligence quotient (IQ) score, which we used to assess overall cognitive functioning.
The fourth edition of the Wechsler Preschool and Primary Scale of Intelligence (WPPSI-IV) \cite{Wechsler2012-mi}
is an individually administered standard intelligence test for children aged 2 years 6 months to 7 years 7 months,
trying to meet the increasing need for the assessment of preschoolers.
Just as the WISC, it derives a full scale IQ score, which we used to assess overall cognitive functioning.

\subsection*{Stool Sample Collection and Sequencing}

Stool samples (n=493) were collected by parents in OMR-200 tubes (OMNIgene GUT, DNA Genotek, Ottawa, Ontario, Canada),
immediately stored on ice, and brought within 24 hrs to the lab in RI where they were immediately frozen at -80 $^{\circ}$C.
Stool samples were not collected if the subject had taken antibiotics within the last two weeks.
DNA extraction was performed at Wellesley College (Wellesley, MA).
Nucleic acids were extracted from stool samples using the RNeasy PowerMicrobiome kit
automated on the QIAcube (Qiagen, Germantown, MD), excluding the DNA degradation steps.
Extracted DNA was sequenced at the Integrated Microbiome Resource (IMR, Dalhousie University, NS, Canada)

Shotgun metagenomic sequencing was performed on all samples.
A pooled library (max 96 samples per run) was prepared using the Illumina Nextera Flex Kit for MiSeq and NextSeq from 1 ng of each sample.
Samples were then pooled onto a plate and sequenced
on the Illumina NextSeq 550 platform using 150+150 bp paired-end “high output” chemistry,
generating ~400 million raw reads and ~120 Gb of sequence per plate.

\hl{Are we still planning to include 16S data here?}

For sequencing 16S rRNA gene amplicons,
the V4-V5 region of the 16S ribosomal RNA gene was sequenced according to the protocol
described by Comeau et al. \cite{Comeau2017-jg}.
Briefly, the V4-V5 region was amplified once using the Phusion High-Fidelity DNA polymerase
(ThermoFisher Scientific, Waltham, MA) and universal bacterial primers
515F: 5’-GTGYCAGCMGCCGCGGTAA-3’ and 926R: 5’-CCGYCAATTYMTTTRAGTTT-3’ \cite{Parada2016-uz,Walters2016-fi}.
These primers had appropriate Illumina adapters and error-correcting barcodes unique to each sample
to allow up to 380 samples to be simultaneously run per single flow cell.
After being pooled into a single library and quantified fluorometrically,
samples were cleaned-up and normalized using the high-throughput Charm Biotech Just-a-Plate 96-well Normalization Kit (Charm Biotech, Cape Girardeau, MO).
The normalized samples were sequenced on the Illumina MiSeq platform (Illumina, San Diego, CA)
using 300+300 bp paired-end V3 chemistry, producing ~55,000 raw reads per sample.

\subsection*{Computational Analysis}

Shotgun metagenomic sequences were analyzed using the bioBakery suite of computational tools \cite{McIver2018-yc}.
First, \verb|KneadData| (v0.7.7) was used to perform quality control of raw sequence reads,
such as read trimming and removal of reads matching a human genome reference.
Next, \verb|MetaPhlAn| (v3.0.7, using database \verb|mpa_v30_CHOCOPhlAn_201901|) was used to generate taxonomic profiles
by aligning reads to a reference database of marker genes.
Finally, \verb|HUMAnN| (v3.0.0a4) was used to functionally profile the metagenomes.

Raw amplicon sequences were profiled using Quantitative Insights in Microbial Ecology 2 (QIIME2) v2021.2.0 \cite{Bolyen2019-qq}.
Briefly, primers flanking V4-V5 were removed from fastq reads using the cutadapt (v3.2) QIIME2 plugin \cite{Martin2011-zv}.
Raw sequence reads of samples from enrichment cultures were denoised, filtered and clustered into amplicon sequence variants (ASVs) using the Divisive Amplicon Denoising Algorithm (DADA2) plugin in QIIME 2 \cite{Callahan2016-ol}.
After denoising and filtering, 16270 total sequences were recovered with a mean length of 373 bases (270-465, standard deviation 13.21).
Taxonomy was assigned to each ASV using a Naïve-Bayes classifier compared against SILVA v.138 reference database \cite{Yilmaz2013-rj,Quast2013-hc}
trained on the 515F-806R region of the 16S rRNA gene \cite{Bokulich2018-dv}.

Additional data processing, generation of summary statistics, 
and generation of plots was performed using the julia programming language \cite{Bezanson2017-ud}.
See Code Availability section for additional details.

### 2022-06-01 manuscript/sections/results.tex
Generate presentation for Khula that includes only young kids (#36)
* start khula presentation

* simplify manuscript template

\subsection*{Outline}

\begin{itemize}
  \item Intial description: Cohort size, demographics, data types
  \item Basic analysis: Summary statistics, ordinations, permanovas, mantel tests
  \item Exploration of taxonomic profiles + cogscores / brain, cross-sectional. Linear models. \hl{Is this necessary??}
  \item Try HAllA on multivariate data?
  \item Predictive model (RF) - taxonomic profiles + cogscores.
  \item Taxonomic profiles + brain structure
  \item Gene functions + metabolomes
\end{itemize}

### 2022-06-01 src/plotting.jl

@@ -23,7 +23,15 @@ end
 # https://github.com/JuliaStats/MultivariateStats.jl/pull/162
 function loadings(M::MultivariateStats.MDS)
     ev = eigvals(M)
-    ev' .* projection(M)[:, 1:length(ev)]
+    return ev' .* projection(M)[:, 1:length(ev)]
+end
+
+# https://github.com/JuliaStats/MultivariateStats.jl/pull/162
+function loadings(M::MultivariateStats.MDS, dim)
+    l = loadings(M)
+    return l[:, dim]
 end
 
 varexplained(M::MultivariateStats.MDS) = eigvals(M) ./ sum(eigvals(M))
+
+mdsaxis(M::MultivariateStats.MDS, dim::Int) = "MDS$dim ($(round(varexplained(M)[dim] * 100, digits=2))%)"

### 2022-06-01 src/startup.jl

@@ -90,7 +90,7 @@ function _gentps()
         all(ismissing, (t,p)) && return missing
         return max(coalesce(t, 0), coalesce(p, 0))
     end
-    
+    brainmeta = Resonance.brainmeta
     for m in brainmeta
         tps[!, m] .= tps[!, m] ./ map(x-> ismissing(x) || x == 0 ? missing : x, tps."EstimatedTotalIntraCranialVol")
     end

### 2022-06-13 notes/08-permanovas.jl
run mantel test

@@ -1,6 +1,7 @@
 using Resonance
 using CairoMakie
 
+
 omni, etoh, tps, complete_brain, metabolites, species = startup()
 genes = Resonance.read_gfs_arrow()
 ecs = Resonance.read_gfs_arrow(kind="ecs_rename")
@@ -8,11 +9,12 @@ ecs = filter(!hastaxon, ecs)
 kos = Resonance.read_gfs_arrow(kind="kos_rename")
 kos = filter(!hastaxon, kos)
 
-
 unique!(tps, ["subject", "timepoint"])
 
 set!(species, leftjoin(select(omni, ["subject", "timepoint", "sample"]), tps, on=[:subject, :timepoint]))
 set!(genes, leftjoin(select(omni, ["subject", "timepoint", "sample"]), tps, on=[:subject, :timepoint]))
+set!(ecs, leftjoin(select(omni, ["subject", "timepoint", "sample"]), tps, on=[:subject, :timepoint]))
+set!(kos, leftjoin(select(omni, ["subject", "timepoint", "sample"]), tps, on=[:subject, :timepoint]))
 set!(metabolites, leftjoin(select(etoh, ["subject", "timepoint", "sample"]), tps, on=[:subject, :timepoint]))
 
 #-
@@ -23,14 +25,20 @@ specu6mo = specu6mo[vec(prevalence(specu6mo)) .> 0, :]
 genesu6mo = genes[:, map(a-> !ismissing(a) && a < 6, get(genes, :ageMonths))]
 genesu6mo = genesu6mo[vec(prevalence(genesu6mo)) .> 0, :]
 
+ecsu6mo = ecs[:, map(a-> !ismissing(a) && a < 6, get(ecs, :ageMonths))]
+ecsu6mo = ecsu6mo[vec(prevalence(ecsu6mo)) .> 0, :]
+
+kosu6mo = kos[:, map(a-> !ismissing(a) && a < 6, get(kos, :ageMonths))]
+kosu6mo = kosu6mo[vec(prevalence(kosu6mo)) .> 0, :]
+
 metabu6mo = metabolites[:, map(a-> !ismissing(a) && a < 6, get(metabolites, :ageMonths))]
 metabu6mo = metabu6mo[vec(prevalence(metabu6mo)) .> 0, :]
 
 
 #-
 
-fig, ax, hm = plot_permanovas([specu6mo, genesu6mo, metabu6mo], [:ageMonths, :cogScore, :rce, :ed];
-    commlabels = ["species", "genes", "metabolites"],
[…]

### 2022-06-13 src/Resonance.jl

@@ -12,6 +12,7 @@ export airtable_metadata,
        mdsaxis,
        permanovas,
        plot_permanovas,
+       mantel,
        commonname,
        countmap,
        codebreastfeeding!,
@@ -34,6 +35,8 @@ using FilePaths
 using ProgressLogging
 using CategoricalArrays
 using PERMANOVA
+using LinearAlgebra
+using Random
 
 @reexport using BiobakeryUtils
 @reexport using DataFrames
@@ -51,5 +54,6 @@ include("plotting.jl")
 include("brain.jl")
 include("startup.jl")
 include("kneaddata.jl")
+include("mantel.jl")
 
 end

### 2022-06-13 src/genefamilies.jl

@@ -67,7 +67,7 @@ function read_gfs_arrow(; kind="genefamilies", stratified=false)
     @info "building sparse mat"
     mat = sparse(tbl.fidx, tbl.sidx, tbl.value)
     @info "getting features"
-    fs = [GeneFunction(line) for line in eachline(joinpath(scratch, "$(kind)_features.txt"))]
+    fs = [genefunction(line) for line in eachline(joinpath(scratch, "$(kind)_features.txt"))]
     @info "getting samples"
     ss = [MicrobiomeSample(line) for line in eachline(joinpath(scratch, "$(kind)_samples.txt"))]
     return CommunityProfile(mat, fs, ss)

### 2022-06-13 src/mantel.jl

@@ -1,0 +1,19 @@
+function mantel(mat1, mat2; n = 1000)
+    size(mat1) == size(mat2) || throw(ArgumentError("Matrices must be the same size"))
+    all(issymmetric, (mat1, mat2)) || error("Matries are not symetric")
+    
+    ord = size(mat1, 1)
+    
+    v1 = view(mat1, [CartesianIndex(i,j) for i in 1:ord for j in 1:ord if i < j])
+    v2 = view(mat2, [CartesianIndex(i,j) for i in 1:ord for j in 1:ord if i < j])
+
+    c_real = cor(v1, v2)
+    cs = Vector{Float64}(undef, n)
+
+    @inbounds for i in range(1, n)
+        cs[i] = cor(view(v1, randperm(ord)), view(v2, randperm(ord)))
+    end
+
+    return (c_real, sum(>(c_real), cs) / n)
+
+end

### 2022-06-13 src/plotting.jl

@@ -71,10 +71,11 @@ function plot_permanovas!(ax, comms, metadatums; commlabels=[], mdlabels=[], col
     hm = heatmap!(ax, vmat'; colormap, colorrange)
 
     for ci in CartesianIndices(pmat)
-        text!(string(round(vmat[ci], digits=2), "%"); position=(ci[2],ci[1]), align=(:center, :center))
+        c = vmat[ci] < (colorrange[2] - colorrange[1]) / 2 ? :black : :lightgray
+        text!(string(round(vmat[ci], digits=2), "%"); position=(ci[2],ci[1]), align=(:center, :center), color=c)
         p = pmat[ci]
         stars = p < 0.001 ? "***" : p < 0.01 ? "**" : p < 0.05 ? "*" : ""
-        text!(stars; position=(ci[2],ci[1]), align=(:center, :bottom))
+        text!(stars; position=(ci[2],ci[1]), align=(:center, :bottom), color=c)
     end
 
     ax.xticks = (1:length(comms), isempty(commlabels) ? ["comm$i" for i in 1:length(comms)] : commlabels)

### 2022-06-13 notes/05-demographics.jl
Permanovas (#37)
* update packages

* fix genefamilies function

* update kneaddata read counts

* more demo fixes

* oops, add permanovas

* write new plotting permanova function

* run mantel test

@@ -4,19 +4,7 @@ omni, tps = startup(; dfs = [:omni, :tps])
 using CairoMakie
 using AlgebraOfGraphics
 
-rce = (@chain tps begin
-    groupby(:subject)
-    combine(:simple_race => (r-> coalesce(r...)) => :race)
-end).race
-
-
 hist(collect(skipmissing(tps.ageMonths)))
-
-hhs = (@chain tps begin
-    groupby(:subject)
-    combine(:mother_HHS_Education => (r-> coalesce(r...)) => :hhs)
-end).hhs
-
 
 ##
 
@@ -30,12 +18,12 @@ labels = [
 ]
 
 counts = [
-    1142,
-    290,
-    1,
-    343,
-    30,
-    594
+    "White"   =>   1142,
+    "Black"   =>   290,
+    "Asian"   =>   1,
+    "Mixed"   =>   343,
+    "Other"   =>   30,
+    "Unknown" =>   594
 ]
 
 colors = [
@@ -57,33 +45,23 @@ Legend(fig[2,1], [MarkerElement(color=c, marker=:circle) for c in colors],
         tellheight=true, tellwidth=false, nbanks=3
 )
 
-save("figures/demo_race.png", fig)
+# save("figures/demo_race.png", fig)
 fig
 ##
 
-using CategoricalArrays
 
-ed = skipmissing(unique(tps, [:subject, :mother_HHS_Education]).mother_HHS_Education) |> collect
-filter!(!=(-8), ed)
-ed = categorical(sort(ed); ordered=true)
-ed = recode(ed, 
-     2 => "Junior high school",
-     3 => "Some high school",
-     4 => "High school grad",
-     5 => "Some college",
-     6 => "College grad",
-     7 => "Grad/professional school")
+ed = skipmissing(unique(tps, [:subject]).ed) |> collect
 
 fig, ax, p = pie(map(levels(ed)) do lev
         count(==(lev), ed)
     end;
-    color = Makie.to_colormap(:viridis, 6), axis=(;aspect=1)
+    color = Makie.resample_cmap(:viridis, 6), axis=(;aspect=1)
 )
 hidedecorations!(ax)
 hidespines!(ax)
 
 Legend(fig[2,1],
-    [MarkerElement(color=c, marker=:circle) for c in Makie.to_colormap(:viridis, 6)],
[…]

### 2022-06-13 notes/07-kneaddata_counts.jl

@@ -2,21 +2,44 @@ using Resonance
 
 kneadread = Resonance.load_knead()
 
+DataFrames.transform!(kneadread, 
+    AsTable(Cols(r"raw")) => ByRow(sum) => :raw,
+    AsTable(Cols(r"trimmed")) => ByRow(sum) => :trimmed,
+    AsTable(Cols(r"decontaminated")) => ByRow(sum) => :decontaminated,
+    AsTable(Cols(r"final")) => ByRow(sum) => :final
+)
+
+DataFrames.transform!(kneadread, 
+    AsTable([:raw, :trimmed]) => ByRow(x-> ismissing(x[2]) ? 0 : (x[1] - x[2]) / x[1] * 100) => :perc_trimmed,
+    AsTable([:raw, :decontaminated]) => ByRow(x-> ismissing(x[2]) ? 0 : (x[1] - x[2]) / x[1] * 100) => :perc_decontaminated
+)
+
+kneadlong = stack(select(kneadread, [:Sample, :raw, :trimmed, :final]), [:raw, :trimmed, :final])
+
 ##
 
 using CairoMakie
+using AlgebraOfGraphics
 
-human = @chain human begin
-    groupby(:sample)
-    combine(:count=> sum)    
-end
+##
 
-hist(human.count_sum .* 150)
+fig = Figure(; resolution=(1100, 600))
+lims = extrema(skipmissing(kneadread.raw))
 
+ax1 = Axis(fig[1,1]; xlabel="Raw reads", ylabel="Final reads", yscale=log10, xscale=log10)
+ax2 = Axis(fig[1,3]; xlabel="Raw reads", ylabel="Final reads", yscale=log10, xscale=log10)
 
-using Statistics
+sc1 = scatter!(ax1, kneadread.raw, kneadread.final; color = kneadread.perc_trimmed)
+lines!(ax1, Point2f.(zip(lims, lims)); color = :gray)
 
-median(human.count_sum .* 150)
-mean(human.count_sum .* 150)
-minimum(human.count_sum .* 150)
-quantile(human.count_sum .* 150, [0.10, 0.9])
+sc2 = scatter!(ax2, kneadread.raw, kneadread.final; color = kneadread.perc_decontaminated, colormap=:plasma)
+lines!(ax2, Point2f.(zip(lims, lims)); color = :gray)
+
+Colorbar(fig[1,2], sc1; label = "Percent trimmed")
[…]

### 2022-06-13 notes/08-permanovas.jl

@@ -1,0 +1,175 @@
+using Resonance
+using CairoMakie
+
+
+omni, etoh, tps, complete_brain, metabolites, species = startup()
+genes = Resonance.read_gfs_arrow()
+ecs = Resonance.read_gfs_arrow(kind="ecs_rename")
+ecs = filter(!hastaxon, ecs)
+kos = Resonance.read_gfs_arrow(kind="kos_rename")
+kos = filter(!hastaxon, kos)
+
+unique!(tps, ["subject", "timepoint"])
+
+set!(species, leftjoin(select(omni, ["subject", "timepoint", "sample"]), tps, on=[:subject, :timepoint]))
+set!(genes, leftjoin(select(omni, ["subject", "timepoint", "sample"]), tps, on=[:subject, :timepoint]))
+set!(ecs, leftjoin(select(omni, ["subject", "timepoint", "sample"]), tps, on=[:subject, :timepoint]))
+set!(kos, leftjoin(select(omni, ["subject", "timepoint", "sample"]), tps, on=[:subject, :timepoint]))
+set!(metabolites, leftjoin(select(etoh, ["subject", "timepoint", "sample"]), tps, on=[:subject, :timepoint]))
+
+#-
+
+specu6mo = species[:, map(a-> !ismissing(a) && a < 6, get(species, :ageMonths))]
+specu6mo = specu6mo[vec(prevalence(specu6mo)) .> 0, :]
+
+genesu6mo = genes[:, map(a-> !ismissing(a) && a < 6, get(genes, :ageMonths))]
+genesu6mo = genesu6mo[vec(prevalence(genesu6mo)) .> 0, :]
+
+ecsu6mo = ecs[:, map(a-> !ismissing(a) && a < 6, get(ecs, :ageMonths))]
+ecsu6mo = ecsu6mo[vec(prevalence(ecsu6mo)) .> 0, :]
+
+kosu6mo = kos[:, map(a-> !ismissing(a) && a < 6, get(kos, :ageMonths))]
+kosu6mo = kosu6mo[vec(prevalence(kosu6mo)) .> 0, :]
+
+metabu6mo = metabolites[:, map(a-> !ismissing(a) && a < 6, get(metabolites, :ageMonths))]
+metabu6mo = metabu6mo[vec(prevalence(metabu6mo)) .> 0, :]
+
+
+#-
+
[…]

### 2022-06-13 src/Resonance.jl

@@ -10,6 +10,9 @@ export airtable_metadata,
        loadings,
        varexplained,
        mdsaxis,
+       permanovas,
+       plot_permanovas,
+       mantel,
        commonname,
        countmap,
        codebreastfeeding!,
@@ -24,19 +27,24 @@ using MultivariateStats
 using Dictionaries
 using CodecZlib
 using Statistics
-using Term
-using Term.progress
 using Arrow
 using Tables
 using SparseArrays
 using CodecZlib
 using FilePaths
+using ProgressLogging
+using CategoricalArrays
+using PERMANOVA
+using LinearAlgebra
+using Random
 
 @reexport using BiobakeryUtils
 @reexport using DataFrames
 @reexport using CSV
 @reexport using XLSX
 @reexport using DataFramesMeta
+
+const transform = DataFrames.transform
 
 include("airtable.jl")
 include("wrangle.jl")
@@ -46,5 +54,6 @@ include("plotting.jl")
 include("brain.jl")
 include("startup.jl")
 include("kneaddata.jl")
+include("mantel.jl")
 
 end

### 2022-06-13 src/genefamilies.jl

@@ -1,10 +1,9 @@
-function write_gfs_arrow()
+function write_gfs_arrow(dir=ENV["ANALYSIS_FILES"]; kind="genefamilies", stratified=false)
     allfiles = String[]
 
     @info "getting files"
-    for (root, dirs, files) in walkdir("/grace/echo/analysis/biobakery3/")
-        contains(root, "batch") || continue
-        filter!(f-> contains(f, "genefamilies") && !contains(f, "relab") && !contains(f, "zymo") && !contains(f, r"^FE\d+"), files)
+    for (root, dirs, files) in walkdir(dir)
+        filter!(f-> contains(splitext(f)[1], Regex(string(kind, '$'))) && !contains(f, r"^FE\d+"), files)
         append!(allfiles, joinpath.(Ref(root), files))
     end
 
@@ -24,10 +23,10 @@ function write_gfs_arrow()
     @info "writing temporary arrow file"
     open(tmp, "w") do io
         tbls = Tables.partitioner(allfiles) do f
-            samplename = replace(splitext(basename(f))[1], r"_S\d+_genefamilies" => "")
+            samplename = replace(splitext(basename(f))[1], Regex(string(raw"_S\d+_", kind)) => "")
 
             sdf = CSV.read(f, DataFrame; header=["feature", "value"], skipto=2)
-            subset!(sdf, "feature"=> ByRow(f-> !contains(f, '|'))) # skip stratified features
+            stratified || subset!(sdf, "feature"=> ByRow(f-> !contains(f, '|'))) # skip stratified features
             sdf.sample .= samplename
             sdf.sidx .= sdic[samplename]
 
@@ -47,13 +46,13 @@ function write_gfs_arrow()
     df.fidx = [fdic[f] for f in df.feature]
 
     @info "Writing table"
-    Arrow.write(joinpath(scratch, "genefamilies.arrow"), df)
-    open(joinpath(scratch, "features.txt"), "w") do io
+    Arrow.write(joinpath(scratch, "$kind.arrow"), df)
+    open(joinpath(scratch, "$(kind)_features.txt"), "w") do io
         for f in keys(fdic)
             println(io, f)
         end
     end
-    open(joinpath(scratch, "samples.txt"), "w") do io
+    open(joinpath(scratch, "$(kind)_samples.txt"), "w") do io
         for s in keys(sdic)
             println(io, s)
         end
[…]

### 2022-06-13 src/kneaddata.jl

@@ -1,32 +1,19 @@
 
-function _read_knead_read_counts(dir=ENV["ANALYSIS_FILES"])
-    fileinfo = NamedTuple[]
-
-    paired_paths = Iterators.filter(walkpath(PosixPath(dir))) do file 
-        "kneaddata" in file.segments || return false
-        contains(filename(file), r"read_counts") || return false
-        return true
-    end
-
-    df = DataFrame()
-
-    for path in paired_paths
-        @info path
-        batch = path.segments[findfirst(s-> contains(s, "batch"), path.segments)]
-        logs = CSV.read(path, DataFrame; stringtype=String, missingstring=["", "NA"])
-        rename!(col-> replace(col, "hg37dec_v0.1"=>"Homo_sapiens"), logs)
-        subset!(logs, "raw pair1"=>ByRow(!ismissing))
-        foreach(col-> logs[!, col] = coalesce.(logs[!, col], 0), names(logs)[2:end])
-        logs[!, :batch] .= batch
-        append!(df, logs)
-    end
-
-    CSV.write("data/read_counts.csv", df)
-end
-
-##
-
 function load_knead(dir=ENV["ANALYSIS_FILES"])
-    !isfile("data/read_counts.csv") && _read_knead_read_counts(dir)
-    CSV.read("data/read_counts.csv", DataFrame)
+    run(Cmd([
+        "kneaddata_read_count_table",
+        "--input", joinpath(dir, "kneaddata/"),
+        "--output", joinpath(dir, "kneaddata", "kneaddata_read_counts.tsv")
+    ]))
+
+    df = CSV.read("/grace/sequencing/processed/mgx/kneaddata/kneaddata_read_counts.tsv", DataFrame; missingstring="NA")
+    for col in names(df, r"Homo_sapiens")
+        df[!, col] = map(p-> coalesce(p...), zip(df[!, col], df[!, replace(col, "Homo_sapiens"=>"hg37dec_v0.1")]))
+    end
+    df."decontaminated Homo_sapiens pair1" = map(x-> ismissing(x) ? missing :  x isa Real ? x : parse(Float64, x), df."decontaminated Homo_sapiens pair1")
+    df."decontaminated Homo_sapiens orphan2" = map(x-> ismissing(x) ? missing :  x isa Real ? x : parse(Float64, x), df."decontaminated Homo_sapiens orphan2")
+
+    df = select(df, Not(r"hg37"))
+    CSV.write("data/read_counts.csv", df)
+    return df
 end

### 2022-06-13 src/plotting.jl

@@ -35,3 +35,51 @@ end
 varexplained(M::MultivariateStats.MDS) = eigvals(M) ./ sum(eigvals(M))
 
 mdsaxis(M::MultivariateStats.MDS, dim::Int) = "MDS$dim ($(round(varexplained(M)[dim] * 100, digits=2))%)"
+
+varexpl(p::PERMANOVA.PSummary) = p.results[1, 3] * 100
+pvalue(p::PERMANOVA.PSummary) = p.results[1, 5]
+
+function permanovas(comm, metadatums)
+    permdf = DataFrame()
+    for md in metadatums
+        com_md = get(comm, md)
+        hasmd = findall(!ismissing, com_md)
+        df = DataFrame(test = com_md[hasmd])
+        disallowmissing!(df)
+        size(df, 1) < 20 && @warn "Very small number of not missing $md"
+
+        p = permanova(df, abundances(comm)[:, hasmd]', BrayCurtis, @formula(1 ~ test))
+        push!(permdf, (; metadatum = md, varexpl=varexpl(p), pvalue=pvalue(p)))
+    end
+    return permdf
+
+end
+
+function plot_permanovas(comms, metadatums; commlabels=[], mdlabels=[], colormap=:blues, colorrange=(0,10))
+    fig = Figure()
+    ax = Axis(fig[1,1])
+    hm = plot_permanovas!(ax, comms, metadatums; commlabels, mdlabels, colormap, colorrange)
+
+    return fig, ax, hm
+end
+
+function plot_permanovas!(ax, comms, metadatums; commlabels=[], mdlabels=[], colormap=:blues, colorrange=(0,10))
+    ps = [permanovas(comm, metadatums) for comm in comms]
+    vmat = mapreduce(df-> df.varexpl, hcat, ps)
+    pmat = mapreduce(df-> df.pvalue, hcat, ps)
+    
+    hm = heatmap!(ax, vmat'; colormap, colorrange)
+
+    for ci in CartesianIndices(pmat)
+        c = vmat[ci] < (colorrange[2] - colorrange[1]) / 2 ? :black : :lightgray
+        text!(string(round(vmat[ci], digits=2), "%"); position=(ci[2],ci[1]), align=(:center, :center), color=c)
+        p = pmat[ci]
+        stars = p < 0.001 ? "***" : p < 0.01 ? "**" : p < 0.05 ? "*" : ""
[…]

### 2022-06-13 src/startup.jl

@@ -5,12 +5,15 @@ const mainmeta = [
         "age6to12mo",
         "age12moplus",
         "mother_HHS_Education",
-        "simple_race",
+        "race",
         "cogScore",
         "has_segmentation",
         "ECHOTPCoded",
-        "assessmentDate"
+        "assessmentDate",
+        "ed",
+        "rce"
     ]
+
     
 const brainmeta = [
             #  "CortexVol",
@@ -79,6 +82,29 @@ end
 
 function _gentps()
     tps  = CSV.read(joinpath(@__DIR__, "..", "data/wrangled/timepoints.csv"), DataFrame)
+
+    DataFrames.transform!(groupby(tps, :subject), :mother_HHS_Education => (r->coalesce(r...)) => :hhs)
+    tps.ed = categorical(tps.hhs; levels=[-8 , 2:7...], ordered=true)
+    tps.ed = recode(tps.ed,
+        -8 => missing,
+        2 => "Junior high school",
+        3 => "Some high school",
+        4 => "High school grad",
+        5 => "Some college",
+        6 => "College grad",
+        7 => "Grad/professional school")
+
+    DataFrames.transform!(groupby(tps, :subject), :race => (r->coalesce(r...)) => :race)
+    tps.rce = categorical(tps.race; ordered=true)
+    tps.rce = recode(tps.rce, 
+        "American Indian or Alaska Native"=> "Other",
+        "Some other race"                 => "Other",
+        "Asian Indian"                    => "Asian",
+        "Black or African American"       => "Black",
+        missing                           => "Unknown"
+    )
+    levels!(tps.rce, ["White","Black","Asian","Mixed","Other","Unknown"])
+
     tps."Left-Thalamus" = map(eachrow(tps)) do row
         (t, p) = (row."Left-Thalamus", row."Left-Thalamus-Proper")
         all(ismissing, (t,p)) && return missing
@@ -94,6 +120,8 @@ function _gentps()
     for m in brainmeta
         tps[!, m] .= tps[!, m] ./ map(x-> ismissing(x) || x == 0 ? missing : x, tps."EstimatedTotalIntraCranialVol")
     end
+
+
     
     select!(tps, ["subject", "timepoint", mainmeta..., brainmeta...])
     rename!(tps, Dict(k=> replace(k, "-"=>"_") for k in brainmeta))

### 2022-06-13 src/wrangle.jl

@@ -32,7 +32,7 @@ function findprevstool(timepoints, has_stool; rev=false)
     return (timepoints .> fs) .& has_stool
 end
 
-countmap(v) = Dict(k=> count(x -> x === k, v) for k in unique(v))
+countmap(v) = Dict(k=> count(x -> !ismissing(x) && x == k, v) for k in unique(v))
 
 ## Breastfeeding
 

### 2022-06-13 slides/subsections/models.tex
create beamer presentation with figures
update slides /figures
include png and svgs in version control

--- 2022-05-11 slides/subsections/metabolites.tex
@@
-% -----------------------------------------------------------------------------
-
-\begin{frame}{GABA and Glutamate}
+\begin{frame}{PERMANOVAs - all kids}
     \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment
 
-        \column{.5\textwidth} % Right column and width
+        \column{.6\textwidth} % Right column and width
 
         \begin{figure}
-        \includegraphics[width=1\linewidth]{../figures/gaba-glutamate.png}
+        \includegraphics[width=1\linewidth]{../figures/kids_all_permanovas.png}
         \end{figure}
@@
         \textbf{Details}
-        \begin{enumerate}
-            \item Comparing mother and child samples
-            \item Mean-centered and compositional
-            \item Plotting $log_2(relative\_abundance)$
-        \end{enumerate}
+        \begin{itemize}
+            \item Single sample (first) per kid
+            \item Univariate permanova
+            \item For KOs / ECs, UNGROUPED features were excluded, UNMAPPED were included
+        \end{itemize}
 
@@
 
-% -----------------------------------------------------------------------------
-
-\begin{frame}{GABA and Glutamate, by age}
+\begin{frame}{PERMANOVAs - kids under 6 months old}
     \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment
 
-        \column{.5\textwidth} % Right column and width
+        \column{.6\textwidth} % Right column and width
 
         \begin{figure}
-        \includegraphics[width=1\linewidth]{../figures/gaba-glutamate_age.png}
+        \includegraphics[width=1\linewidth]{../figures/kids_all_permanovas.png}
         \end{figure}
@@
         \textbf{Details}
-        \begin{enumerate}
-            \item Comparing mother and child samples
-            \item Age in months colored, moms are gray
-            \item Plotting $log_2(relative\_abundance)$
-        \end{enumerate}
+        \begin{itemize}
+            \item Single sample per kid
+            \item Univariate permanova
+            \item For KOs / ECs, UNGROUPED features were excluded, UNMAPPED were included
+        \end{itemize}
 
@@
 
-% -----------------------------------------------------------------------------
-
-\begin{frame}{Metabolites PCoA}
+\begin{frame}{PERMANOVAs - kids over 1 year old}
     \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment
 
-        \column{.5\textwidth} % Right column and width
+        \column{.6\textwidth} % Right column and width
 
         \begin{figure}
-        \includegraphics[width=1\linewidth]{../figures/metabolites_pcoa.png}
+        \includegraphics[width=1\linewidth]{../figures/kids_all_permanovas.png}
         \end{figure}
@@
         \textbf{Details}
-        \begin{enumerate}
-            \item PCoA based on BC dissimilartiy
-        \end{enumerate}
+        \begin{itemize}
+            \item Single sample per kid
+            \item Univariate permanova
+            \item For KOs / ECs, UNGROUPED features were excluded, UNMAPPED were included
+            \item Metabolites not included because there are only a few samples over 1 year
+        \end{itemize}
 
@@
 
-% -----------------------------------------------------------------------------
+\begin{frame}{Linear models - taxa}
+    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment
+
+        \column{.6\textwidth} % Right column and width
+
+        \begin{figure}
+        \includegraphics[width=1\linewidth]{../figures/lms-split-2.png}
+        \end{figure}
+
+        \column{.4\textwidth} % Right column and width
+
+        \textbf{Details}
+        \begin{itemize}
+            \item Kids $> 18$ months old, species $> 10\%$ prevalence
+            \item GLM: $bug \sim cogScore + age$
+            \item There were ~15 bugs significant after FDR correction
+        \end{itemize}
+
+    \end{columns}
+
+\end{frame}
+
+\begin{frame}{Linear models - taxa}
+    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment
+
+        \column{.6\textwidth} % Right column and width
+
+        \begin{figure}
+        \includegraphics[width=1\linewidth]{../figures/lms-split-2.png}
+        \end{figure}
+
+        \column{.4\textwidth} % Right column and width
+
+        \textbf{Details}
+        \begin{itemize}
+            \item Kids $> 18$ months old, species $> 10\%$ prevalence
+            \item GLM: $bug \sim cogScore + age$
+            \item There were ~15 bugs significant after FDR correction
+        \end{itemize}
+
+    \end{columns}
+
+\end{frame}

### 2022-06-13 notes/06-plot_intersections.jl

@@ -81,7 +81,7 @@ linkxaxes!(dot_ax, intersection_ax)
 hidespines!(intersection_ax, :t, :r, :b)
 hidespines!(set_ax, :t, :r, :l)
 
-save("figures/upset_scan_stool.pdf", fig)
+save("figures/upset_scan_stool.png", fig)
 fig
 
 ##
@@ -135,7 +135,7 @@ hist!(scan_age_young, subset(tps, AsTable([:ageMonths, :has_segmentation]) => By
 Label(fig[1, 0], text = "All kids", tellheight=false)
 Label(fig[2, 1], text = "Young kids", tellheight=false)
 
-save("figures/age_hists.pdf", fig)
+save("figures/age_hists.png", fig)
 fig
 
 ##
@@ -157,7 +157,7 @@ tightlimits!(ass_bydate, Left(), Right(), Bottom())
 
 linkyaxes!(stools_bydate, ass_bydate)
 
-save("figures/collection_over_time.pdf", fig)
+save("figures/collection_over_time.png", fig)
 fig
 ##
 
@@ -221,7 +221,7 @@ linkxaxes!(dot_ax, intersection_ax)
 hidespines!(intersection_ax, :t, :r, :b)
 hidespines!(set_ax, :t, :r, :l)
 
-save("figures/upset_scan_stool_bysybject.pdf", fig)
+save("figures/upset_scan_stool_bysybject.png", fig)
 fig
 
 ##
@@ -274,7 +274,7 @@ linkxaxes!(dot_ax, intersection_ax)
 hidespines!(intersection_ax, :t, :r, :b)
 hidespines!(set_ax, :t, :r, :l)
 
-save("figures/upset_score_stool.pdf", fig)
+save("figures/upset_score_stool.png", fig)
 fig
 
 ##
@@ -283,7 +283,7 @@ fig
 fig, ax, plt = hist(collect(skipmissing(tps.cogScore)))
 plt2 = hist!(ax, collect(skipmissing(tps.cogScore[tps.has_stool .|| tps.has_prevstool])))
 Legend(fig[1,2], [plt, plt2], ["All scores", "Scores with stool"])
-save("figures/cogscores_hist.pdf", fig)
+save("figures/cogscores_hist.png", fig)
 fig
 
 ##
@@ -335,7 +335,7 @@ linkxaxes!(dot_ax, intersection_ax)
 hidespines!(intersection_ax, :t, :r, :b)
[…]

### 2022-06-13 notes/08-permanovas.jl
change svg to png

@@ -113,6 +113,7 @@ save("figures/kids_all_permanovas.png", fig)
 fig
 
 #-
+
 ## Mantel tests
 
 using Combinatorics
@@ -171,5 +172,5 @@ end
 ax.xticks = (1:n-1, isempty(labels) ? ["comm$i" for i in 1:n-1] : labels[2:end])
 ax.yticks = (1:n-1, isempty(labels) ? ["comm$i" for i in 1:n] : labels[1:end-1])
 
-save("figures/mantel.svg", fig)
+save("figures/mantel.png", fig)
 fig

### 2022-06-13 notes/08-permanovas.jl
some mantel tweaks

@@ -163,10 +163,11 @@ ax = Axis(fig[1,1], title="Mantel tests")
 hm = heatmap!(ax, vmat[1:n-1, 2:n]'; colormap=:viridis, colorrange = (0.01, 1), lowclip=:lightgray)
 
 for ci in manteldf.idx
-    text!(string(round(vmat[ci], digits=4)); position=(ci[2]-1,ci[1]), align=(:center, :center))
+    c = vmat[ci] < 0.5 ? :lightgray : :black
+    text!(string(round(vmat[ci], digits=4)); position=(ci[2]-1,ci[1]), align=(:center, :center), color=c)
     p = pmat[ci]
     stars = p < 0.001 ? "***" : p < 0.01 ? "**" : p < 0.05 ? "*" : ""
-    text!(stars; position=(ci[2]-1,ci[1]), align=(:center, :bottom))
+    text!(stars; position=(ci[2]-1,ci[1]), align=(:center, :bottom), color=c)
 end
 
 ax.xticks = (1:n-1, isempty(labels) ? ["comm$i" for i in 1:n-1] : labels[2:end])

### 2022-06-13 slides/subsections/samples.tex
update slides /figures
Permanovas (#37)
* update packages

* fix genefamilies function

* update kneaddata read counts

* more demo fixes

* oops, add permanovas

* write new plotting permanova function

* run mantel test
Presentation updates (#39)
* include png and svgs in version control

* change svg to png

* some mantel tweaks

* remove test workflow for now

@@
         \begin{figure}
-        \includegraphics[width=1\linewidth]{../figures/dyads.png}
+        \includegraphics[width=1\linewidth]{../figures/dyads}
         \end{figure}
@@
         \textbf{Details}
-        \begin{enumerate}
+        \begin{itemize}
             \item \textbf{Left} - Child samples that have at least 1 associated maternal sample
             \item \textbf{Right} - Maternal samples that have at least 1 associated child sample
-        \end{enumerate}
+        \end{itemize}
+
+    \end{columns}
+
+\end{frame}
+
+% -----------------------------------------------------------------------------
+
+\begin{frame}{Sample collection by age}
+    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment
+
+        \column{.6\textwidth} % Right column and width
+
+        \begin{figure}
+        \includegraphics[width=1\linewidth]{../figures/age_hists}
+        \end{figure}
+
+        \column{.4\textwidth} % Right column and width
+
+        \textbf{Details}
+        \begin{itemize}
+            \item Bottom row is same as top, but only from timepoints $<$ 24 months
+        \end{itemize}
+
+    \end{columns}
+
+\end{frame}
+
+% -----------------------------------------------------------------------------
+
+\begin{frame}{Sample collection over time}
+    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment
+
+        \column{.6\textwidth} % Right column and width
+
+        \begin{figure}
+        \includegraphics[width=1\linewidth]{../figures/collection_over_time}
+        \end{figure}
+
+        \column{.4\textwidth} % Right column and width
+
+        \textbf{Details}
+        \begin{itemize}
+            \item \textbf{Top} - Stool samples (just omni) collected by date
+            \item \textbf{Bottom} - Assessments (clinical visits) by date
+        \end{itemize}
+
+    \end{columns}
+
+\end{frame}
+
+% -----------------------------------------------------------------------------
+
+\begin{frame}{Sample QC}
+    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment
+
+        \column{.6\textwidth} % Right column and width
+
+        \begin{figure}
+        \includegraphics[width=1\linewidth]{../figures/kneaddata_counts}
+        \end{figure}
+
+        \column{.4\textwidth} % Right column and width
+
+        \textbf{Details}
+        \begin{itemize}
+            \item Raw reads (x-axis) vs final QC'd reads (y-axis) after $kneaddata$ QC
+            \item \textbf{Left} - Shows percent trimmed
+            \item \textbf{Right} - Shows percent human reads
+        \end{itemize}
 
@@
         \begin{figure}
-        \includegraphics[width=1\linewidth]{../figures/upset_scan_stool.pdf}
+        \includegraphics[width=1\linewidth]{../figures/upset_scan_stool.png}
         \end{figure}
@@
         \textbf{Details}
-        \begin{enumerate}
+        \begin{itemize}
             \item Overlap of sample collection and brain scans
@@
             \item Not 100\% sure of this code - needs review
-        \end{enumerate}
+        \end{itemize}
 
@@
         \begin{figure}
-        \includegraphics[width=1\linewidth]{../figures/upset_scan_stool_bysybject.pdf}
+        \includegraphics[width=1\linewidth]{../figures/upset_scan_stool_bysybject.png}
         \end{figure}
@@
         \textbf{Details}
-        \begin{enumerate}
+        \begin{itemize}
             \item Overlap of sample collection and brain scans
             \item Numbers reduced to unique subjects
-        \end{enumerate}
+        \end{itemize}
 
@@
         \begin{figure}
-        \includegraphics[width=1\linewidth]{../figures/upset_score_stool.pdf}
+        \includegraphics[width=1\linewidth]{../figures/upset_score_stool.png}
         \end{figure}
@@
         \textbf{Details}
-        \begin{enumerate}
+        \begin{itemize}
             \item Overlap of sample collection and cognitive score
-        \end{enumerate}
+        \end{itemize}
 
@@
         \begin{figure}
-        \includegraphics[width=1\linewidth]{../figures/upset_score_stool_under3.pdf}
+        \includegraphics[width=1\linewidth]{../figures/upset_score_stool_under3.png}
         \end{figure}
@@
         \textbf{Details}
-        \begin{enumerate}
+        \begin{itemize}
             \item Overlap of sample collection and cognitive score for kids under 3
             \item "Under 3" row is all timepoints, regardless of stool collection
-        \end{enumerate}
+        \end{itemize}
 
@@
         \textbf{Details}
-        \begin{enumerate}
+        \begin{itemize}
             \item Numbers reduced to unique subjects
-        \end{enumerate}
+        \end{itemize}
 

### 2022-06-13 notes/08-permanovas.jl

@@ -113,6 +113,7 @@ save("figures/kids_all_permanovas.png", fig)
 fig
 
 #-
+
 ## Mantel tests
 
 using Combinatorics
@@ -162,14 +163,15 @@ ax = Axis(fig[1,1], title="Mantel tests")
 hm = heatmap!(ax, vmat[1:n-1, 2:n]'; colormap=:viridis, colorrange = (0.01, 1), lowclip=:lightgray)
 
 for ci in manteldf.idx
-    text!(string(round(vmat[ci], digits=4)); position=(ci[2]-1,ci[1]), align=(:center, :center))
+    c = vmat[ci] < 0.5 ? :lightgray : :black
+    text!(string(round(vmat[ci], digits=4)); position=(ci[2]-1,ci[1]), align=(:center, :center), color=c)
     p = pmat[ci]
     stars = p < 0.001 ? "***" : p < 0.01 ? "**" : p < 0.05 ? "*" : ""
-    text!(stars; position=(ci[2]-1,ci[1]), align=(:center, :bottom))
+    text!(stars; position=(ci[2]-1,ci[1]), align=(:center, :bottom), color=c)
 end
 
 ax.xticks = (1:n-1, isempty(labels) ? ["comm$i" for i in 1:n-1] : labels[2:end])
 ax.yticks = (1:n-1, isempty(labels) ? ["comm$i" for i in 1:n] : labels[1:end-1])
 
-save("figures/mantel.svg", fig)
+save("figures/mantel.png", fig)
 fig

### 2022-06-29 slides/subsections/fsea.tex
update figures with FSEA

\begin{frame}{FSEA results}

    \textbf{Details}
    \begin{itemize}
        \item Feature set enrichment analysis on Neuroactive gene sets
        \item Using all kids
        \item Using Mann-Whitney U test
        \item Adjusted q via Benjamini-Hochberg FDR correction
    \end{itemize}
\end{frame}

\begin{frame}{FSEA results}

    \begin{table}
        \tiny
        \begin{tabular}{rrrrrrr}
            \hline\hline
            \textbf{geneset} & \textbf{U} & \textbf{median} & \textbf{mu} & \textbf{sigma} & \textbf{pvalue} & \textbf{qvalue} \\
            \texttt{String} & \texttt{Float64} & \texttt{Float64} & \texttt{Float64} & \texttt{Float64} & \texttt{Float64} & \texttt{Float64} \\\hline
            Menaquinone synthesis & 1.80133e8 & -0.0227514 & -5.3586e7 & 9.71341e6 & 3.45407e-8 & 1.03622e-6 \\
            Propionate degradation & 7.92215e6 & -0.0512878 & -1.02439e7 & 2.70804e6 & 0.0001551 & 0.0023265 \\
            Glutamate degradation & 1.90583e7 & -0.0372657 & -1.24293e7 & 3.56529e6 & 0.000489918 & 0.00489918 \\
            GABA synthesis & 2.3322e7 & -0.029258 & -1.30098e7 & 3.82973e6 & 0.000681137 & 0.00510853 \\
            Tryptophan synthesis & 2.56466e8 & -0.00748907 & -2.68982e7 & 1.06954e7 & 0.0119057 & 0.0605421 \\
            SAM synthesis & 6.56769e7 & -0.0162254 & -1.42519e7 & 5.68037e6 & 0.0121084 & 0.0605421 \\
            p-Cresol synthesis & 3.26506e7 & -0.0350193 & -8.52534e6 & 4.07706e6 & 0.0365233 & 0.148674 \\
            Acetate synthesis & 2.16053e8 & -0.0120984 & -2.00878e7 & 9.7636e6 & 0.0396464 & 0.148674 \\
            Quinolinic acid degradation & 1.48097e8 & -0.0162254 & -1.53888e7 & 8.12391e6 & 0.058191 & 0.193307 \\
            Butyrate synthesis & 4.47099e7 & -0.0306039 & -8.57642e6 & 4.63802e6 & 0.0644355 & 0.193307 \\\hline\hline
        \end{tabular}
    \end{table}

\end{frame}

\begin{frame}{Menaquinone synthesis}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.7\textwidth} % Right column and width
        \begin{figure}
            \includegraphics[width=1\linewidth]{../figures/fsea_Menaquinone-synthesis.png}
        \end{figure}

    \end{columns}

\end{frame}

\begin{frame}{Propionate degradation}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.7\textwidth} % Right column and width
        \begin{figure}
            \includegraphics[width=1\linewidth]{../figures/fsea_Propionate-degradation.png}
        \end{figure}

    \end{columns}

\end{frame}

\begin{frame}{Glutamate degradation}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.7\textwidth} % Right column and width
        \begin{figure}
            \includegraphics[width=1\linewidth]{../figures/fsea_Glutamate-degradation.png}
        \end{figure}

    \end{columns}

\end{frame}

\begin{frame}{GABA synthesis}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.7\textwidth} % Right column and width
        \begin{figure}
            \includegraphics[width=1\linewidth]{../figures/fsea_GABA-synthesis.png}
        \end{figure}

    \end{columns}

\end{frame}

\begin{frame}{Tryptophan synthesis}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.7\textwidth} % Right column and width
        \begin{figure}
            \includegraphics[width=1\linewidth]{../figures/fsea_Tryptophan-synthesis.png}
        \end{figure}

    \end{columns}

\end{frame}

\begin{frame}{SAM synthesis}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.7\textwidth} % Right column and width
        \begin{figure}
            \includegraphics[width=1\linewidth]{../figures/fsea_SAM-synthesis.png}
        \end{figure}

    \end{columns}

\end{frame}

\begin{frame}{p-Cresol synthesis}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.7\textwidth} % Right column and width
        \begin{figure}
            \includegraphics[width=1\linewidth]{../figures/fsea_p-Cresol-synthesis.png}
        \end{figure}

    \end{columns}

\end{frame}

\begin{frame}{Acetate synthesis}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.7\textwidth} % Right column and width
        \begin{figure}
            \includegraphics[width=1\linewidth]{../figures/fsea_Acetate-synthesis.png}
        \end{figure}

    \end{columns}

\end{frame}

\begin{frame}{Quinolinic acid degradation}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.7\textwidth} % Right column and width
        \begin{figure}
            \includegraphics[width=1\linewidth]{../figures/fsea_Quinolinic-acid-degradation.png}
        \end{figure}

    \end{columns}

\end{frame}

\begin{frame}{Butyrate synthesis}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.7\textwidth} % Right column and width
        \begin{figure}
            \includegraphics[width=1\linewidth]{../figures/fsea_Butyrate-synthesis.png}
        \end{figure}

    \end{columns}

\end{frame}

### 2022-06-29 notes/10-fsea.jl

@@ -10,6 +10,7 @@ using MultipleTesting
 using AlgebraOfGraphics
 using Statistics
 using DataFrames.InvertedIndices
+using DataFrames.PrettyTables
 using ThreadsX
 
 omni, etoh, tps, complete_brain, metabolites, species = startup()
@@ -29,7 +30,7 @@ gsgenes = genes[:, .!ismissing.(get(genes, :cogScore))]
 cscor = cor(abundances(gsgenes), get(gsgenes, :cogScore), dims=2)
 
 fsdf = DataFrame(
-    ThreadsX.map(collect(keys(neuroactive))) do gs
+    map(collect(keys(neuroactive))) do gs
         ixs = neuroactive[gs]
         isempty(ixs) && return (; geneset = gs, U = NaN, median = NaN, mu = NaN, sigma = NaN, pvalue = NaN)
 
@@ -39,6 +40,9 @@ fsdf = DataFrame(
         acs = filter(!isnan, cscor[Not(ixs)])
         mwu = MannWhitneyUTest(cs, acs)
 
+        fig = Resonance.plot_fsea(acs, cs; label = gs)
+        save("figures/fsea_$(replace(gs, ' '=>"-")).png", fig)
+
         return (; geneset = gs, U = mwu.U, median = mwu.median, mu = mwu.mu, sigma = mwu.sigma, pvalue=pvalue(mwu))
     end
 )
@@ -46,18 +50,10 @@ fsdf = DataFrame(
 subset!(fsdf, :pvalue=> ByRow(!isnan))
 fsdf.qvalue = adjust(fsdf.pvalue, BenjaminiHochberg())
 sort!(fsdf, :qvalue)
-CSV.write("data/fsea_all.csv")
+CSV.write("data/fsea_all.csv", fsdf)
 
+pretty_table(first(fsdf, 10); backend = Val(:latex))
 #- 
-
-gs = "Menaquinone synthesis"
-ixs = neuroactive[gs]
-cs = filter(!isnan, cscor[ixs])
-acs = filter(!isnan, cscor[Not(ixs)])
-
-Resonance.plot_fsea(acs, cs)
-
-#-
 
 nodupes_samples = unique(DataFrame(metadata(gsgenes)), :subject).sample
 

### 2022-06-29 src/plotting.jl

@@ -84,8 +84,10 @@ function plot_permanovas!(ax, comms, metadatums; commlabels=[], mdlabels=[], col
     return hm
 end
 
-function plot_fsea(setcors, notcors; label="geneset")
-    srt = sortperm([setcors; notcors])
+function plot_fsea(setcors, notcors; label="")
+    fullcors = [setcors; notcors]
+    
+    srt = sortperm(fullcors; rev=true)
     ranks = invperm(srt)
     setranks = Set(ranks[1:length(setcors)])
     
@@ -95,16 +97,35 @@ function plot_fsea(setcors, notcors; label="geneset")
     xs = 1:length(srt)
     ys = cumsum(i ∈ setranks ? setscore : notscore for i in eachindex(ranks))
     
+    t = "Enrichment score - $(round(maximum(ys), digits=3))"
+    !isempty(label) && (t = string(label, ": ", t))
+
     fig = Figure()
-    ax1 = Axis(fig[1,1]; title="Enrichment", ylabel="enrichment score")
+    ax1 = Axis(fig[1,1]; title=t, ylabel="enrichment score")
     hidexdecorations!(ax1)
-    ax2 = Axis(fig[2,1]; ylabel="correlation", xlabel="rank")
+
+    ax2 = Axis(fig[2,1])
+    hidedecorations!(ax2)
+
+    ax3 = Axis(fig[3,1]; ylabel="correlation", xlabel="rank")
     
     lines!(ax1, xs, ys)
-    stairs!(ax2, xs, [setcors; notcors][srt])
+
+    vlines!(ax2, ranks[1:length(setcors)]; color=:black)
+
+    blow, bup = extrema(fullcors)
+    band!(ax3, xs, fill(blow, length(srt)), fill(bup, length(srt)); color=fullcors[srt])
     
-    rowsize!(fig.layout, 2, Relative(1/4))
+    lower = [x < 0 ? x : 0.0 for x in fullcors[srt]]
+    upper = [x > 0 ? x : 0.0 for x in fullcors[srt]]
+    band!(ax3, xs, lower, upper; color=:lightgray)
+
     
+    rowsize!(fig.layout, 2, Relative(1/8))
+    rowsize!(fig.layout, 3, Relative(1/4))
+    
+    linkxaxes!(ax1, ax2, ax3)
+    tightlimits!.((ax1, ax2, ax3))
     fig
 end
 

### 2022-07-18 notes/10-fsea.jl
update brain region fsea

@@ -40,8 +40,8 @@ fsdf = DataFrame(
         acs = filter(!isnan, cscor[Not(ixs)])
         mwu = MannWhitneyUTest(cs, acs)
 
-        fig = Resonance.plot_fsea(acs, cs; label = gs)
-        save("figures/fsea_$(replace(gs, ' '=>"-")).png", fig)
+        # fig = Resonance.plot_fsea(acs, cs; label = gs)
+        # save("figures/fsea_$(replace(gs, ' '=>"-")).png", fig)
 
         return (; geneset = gs, U = mwu.U, median = mwu.median, mu = mwu.mu, sigma = mwu.sigma, pvalue=pvalue(mwu))
     end
@@ -137,3 +137,62 @@ subset!(fsdf, :pvalue=> ByRow(!isnan))
 fsdf.qvalue = adjust(fsdf.pvalue, BenjaminiHochberg())
 sort!(fsdf, :qvalue)
 CSV.write(datafiles("fsea_o12.csv"))
+
+
+#-
+
+brain_features = Symbol.(replace.(Resonance.brainmeta, "-"=>"_"))
+brgenes = genes[:, .!ismissing.(get(genes, :Left_Thalamus))]
+
+brfsea = DataFrame()
+
+for region in brain_features
+    @warn region
+    cscor = cor(abundances(brgenes), get(brgenes, region), dims=2)
+
+    fsdf = DataFrame(
+        map(collect(keys(neuroactive))) do gs
+            @info gs
+            ixs = neuroactive[gs]
+            isempty(ixs) && return (; region, geneset = gs, U = NaN, median = NaN, mu = NaN, sigma = NaN, pvalue = NaN, cs = Float64[], acs = Float64[])
+
+            cs = filter(!isnan, cscor[ixs])
+            isempty(cs) && return (; region, geneset = gs, U = NaN, median = NaN, mu = NaN, sigma = NaN, pvalue = NaN, cs = Float64[], acs = Float64[])
+
+            acs = filter(!isnan, cscor[Not(ixs)])
+            mwu = MannWhitneyUTest(cs, acs)
+
+            return (; region, geneset = gs, U = mwu.U, median = mwu.median, mu = mwu.mu, sigma = mwu.sigma, pvalue=pvalue(mwu), cs, acs)
+        end
+    )
+    subset!(fsdf, :pvalue=> ByRow(!isnan))
[…]

### 2022-07-18 src/plotting.jl

@@ -84,9 +84,10 @@ function plot_permanovas!(ax, comms, metadatums; commlabels=[], mdlabels=[], col
     return hm
 end
 
-function plot_fsea(setcors, notcors; label="")
+function plot_fsea(setcors, notcors; label="", bandres=5000)
     fullcors = [setcors; notcors]
-    
+    ncors = length(fullcors)
+
     srt = sortperm(fullcors; rev=true)
     ranks = invperm(srt)
     setranks = Set(ranks[1:length(setcors)])
@@ -94,10 +95,10 @@ function plot_fsea(setcors, notcors; label="")
     setscore =  1 / length(setcors)
     notscore = -1 / length(notcors)
     
-    xs = 1:length(srt)
+    xs = 1:ncors
     ys = cumsum(i ∈ setranks ? setscore : notscore for i in eachindex(ranks))
     
-    t = "Enrichment score - $(round(maximum(ys), digits=3))"
+    t = "Enrichment score - $(round(max(abs.(extrema(ys))...), digits=3))"
     !isempty(label) && (t = string(label, ": ", t))
 
     fig = Figure()
@@ -110,11 +111,12 @@ function plot_fsea(setcors, notcors; label="")
     ax3 = Axis(fig[3,1]; ylabel="correlation", xlabel="rank")
     
     lines!(ax1, xs, ys)
-
     vlines!(ax2, ranks[1:length(setcors)]; color=:black)
 
+
+    rn = ncors > bandres ? round.(Int, range(1, ncors; length=bandres)) : range(1, ncors)
     blow, bup = extrema(fullcors)
-    band!(ax3, xs, fill(blow, length(srt)), fill(bup, length(srt)); color=fullcors[srt])
+    band!(ax3, xs, fill(blow, rn), fill(bup, rn); color=fullcors[srt[rn]])
     
     lower = [x < 0 ? x : 0.0 for x in fullcors[srt]]
     upper = [x > 0 ? x : 0.0 for x in fullcors[srt]]

### 2022-07-21 slides/subsections/fsea.tex
Brain fsea (#43)
* update brain region fsea

* Keep working on plots

* complete brain fsea... for now

@@
               \texttt{String} & \texttt{Float64} & \texttt{Float64} & \texttt{Float64} & \texttt{Float64} & \texttt{Float64} & \texttt{Float64} \\\hline
-            Menaquinone synthesis & 1.80133e8 & -0.0227514 & -5.3586e7 & 9.71341e6 & 3.45407e-8 & 1.03622e-6 \\
-            Propionate degradation & 7.92215e6 & -0.0512878 & -1.02439e7 & 2.70804e6 & 0.0001551 & 0.0023265 \\
-            Glutamate degradation & 1.90583e7 & -0.0372657 & -1.24293e7 & 3.56529e6 & 0.000489918 & 0.00489918 \\
-            GABA synthesis & 2.3322e7 & -0.029258 & -1.30098e7 & 3.82973e6 & 0.000681137 & 0.00510853 \\
-            Tryptophan synthesis & 2.56466e8 & -0.00748907 & -2.68982e7 & 1.06954e7 & 0.0119057 & 0.0605421 \\
-            SAM synthesis & 6.56769e7 & -0.0162254 & -1.42519e7 & 5.68037e6 & 0.0121084 & 0.0605421 \\
-            p-Cresol synthesis & 3.26506e7 & -0.0350193 & -8.52534e6 & 4.07706e6 & 0.0365233 & 0.148674 \\
-            Acetate synthesis & 2.16053e8 & -0.0120984 & -2.00878e7 & 9.7636e6 & 0.0396464 & 0.148674 \\
-            Quinolinic acid degradation & 1.48097e8 & -0.0162254 & -1.53888e7 & 8.12391e6 & 0.058191 & 0.193307 \\
-            Butyrate synthesis & 4.47099e7 & -0.0306039 & -8.57642e6 & 4.63802e6 & 0.0644355 & 0.193307 \\\hline\hline
+              Menaquinone synthesis & 1.80593e8 & -0.0227227 & -5.30723e7 & 9.7112e6 & 4.62758e-8 & 1.38827e-6 \\
+              Propionate degradation & 7.8577e6 & -0.0520019 & -1.03042e7 & 2.70742e6 & 0.000141289 & 0.00211934 \\
+              GABA synthesis & 2.3124e7 & -0.0302445 & -1.31996e7 & 3.82886e6 & 0.000566035 & 0.00424526 \\
+              Glutamate degradation & 1.89577e7 & -0.0376948 & -1.25228e7 & 3.56448e6 & 0.000442716 & 0.00424526 \\
+              Tryptophan synthesis & 2.56154e8 & -0.00604063 & -2.71454e7 & 1.0693e7 & 0.0111292 & 0.0667749 \\
+              S-Adenosylmethionine synthesis & 6.5944e7 & -0.0124222 & -1.39665e7 & 5.67907e6 & 0.013921 & 0.069605 \\
+              Acetate synthesis & 2.1512e8 & -0.0140563 & -2.09669e7 & 9.76138e6 & 0.0317181 & 0.135935 \\
+              p-Cresol synthesis & 3.26509e7 & -0.0356679 & -8.51573e6 & 4.07613e6 & 0.0366924 & 0.137596 \\
+              Quinolinic acid degradation & 1.47932e8 & -0.0163907 & -1.55163e7 & 8.12206e6 & 0.0560839 & 0.168252 \\
+              Butyrate synthesis & 4.44021e7 & -0.0320785 & -8.87212e6 & 4.63697e6 & 0.0557039 & 0.168252 \\\hline\hline
+            \end{tabular}
+          \end{table}
+
+\end{frame}
+
+\begin{frame}{FSEA results}
+
+    \textbf{Details}
+    \begin{itemize}
+        \item Feature set enrichment analysis on Neuroactive gene sets
+        \item Using all kids with brain scans
+        \item Using Mann-Whitney U test
+        \item Adjusted q via Benjamini-Hochberg FDR correction
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{FSEA results}
+
+    \begin{table}
+        \tiny
+        \begin{tabular}{rrrrr}
+            \hline\hline
+            \textbf{region} & \textbf{geneset} & \textbf{U} & \textbf{pvalue} & \textbf{qvalue} \\
+            \texttt{Symbol} & \texttt{String} & \texttt{Float64} & \texttt{Float64} & \texttt{Float64} \\\hline
+            Right Cerebellum White Matter & Menaquinone synthesis & 7.74978e7 & 1.85854e-6 & 0.00135302 \\
+            Left Cerebellum White Matter & Menaquinone synthesis & 8.21281e7 & 8.03345e-5 & 0.0194945 \\
+            Left Caudate & GABA synthesis & 7.43746e6 & 6.34299e-5 & 0.0194945 \\
+            Right Cerebellum White Matter & GABA synthesis & 7.97035e6 & 0.000170094 & 0.0247657 \\
+            Right Cerebellum Cortex & Menaquinone synthesis & 8.30886e7 & 0.000161753 & 0.0247657 \\
+            Right Cerebellum Cortex & GABA synthesis & 8.1275e6 & 0.000225155 & 0.0273189 \\
+            Left Cerebellum Cortex & GABA synthesis & 8.63868e6 & 0.000542555 & 0.0529034 \\
+            Left Hippocampus & Menaquinone synthesis & 8.4953e7 & 0.000581357 & 0.0529034 \\
+            Brain Stem & Menaquinone synthesis & 8.55193e7 & 0.000839954 & 0.0679429 \\
+            Left Hippocampus & GABA synthesis & 9.08537e6 & 0.00112323 & 0.0817708 \\
+            Left Cerebellum White Matter & GABA synthesis & 9.24701e6 & 0.00144799 & 0.0921011 \\
+            Left Amygdala & GABA synthesis & 9.31564e6 & 0.00161044 & 0.0921011 \\
+            Right Accumbens area & Histamine synthesis & 8.48778e5 & 0.00164466 & 0.0921011 \\
+            Right Cerebellum White Matter & Glutamate degradation & 1.01619e7 & 0.00195892 & 0.101864 \\
+            Left Amygdala & Glutamate synthesis & 4.84973e7 & 0.00230519 & 0.102444 \\
+            Right Cerebellum White Matter & Inositol degradation & 3.41852e6 & 0.00253296 & 0.102444 \\
+            Brain Stem & Glutamate synthesis & 7.48954e7 & 0.00215846 & 0.102444 \\
+            CSF & Menaquinone synthesis & 1.21243e8 & 0.00249003 & 0.102444 \\
+            Left Cerebellum White Matter & Inositol degradation & 3.71948e6 & 0.0046878 & 0.126397 \\
+            Left Putamen & GABA synthesis & 1.00288e7 & 0.00461295 & 0.126397 \\
+            Left Pallidum & GABA synthesis & 9.86761e6 & 0.00366688 & 0.126397 \\
+            Left Amygdala & Tryptophan synthesis & 1.22323e8 & 0.00346543 & 0.126397 \\
+            Right Lateral Ventricle & p-Cresol synthesis & 2.70358e7 & 0.00388996 & 0.126397 \\
+            Right Caudate & GABA synthesis & 1.0038e7 & 0.00467257 & 0.126397 \\
+            Right Putamen & Glutamate synthesis & 7.39107e7 & 0.00451969 & 0.126397 \\
+            Right Accumbens area & Glutamate synthesis & 7.40534e7 & 0.00407287 & 0.126397 \\
+            CSF & Butyrate synthesis & 3.86017e7 & 0.00389266 & 0.126397 \\
+            Left Accumbens area & ClpB & 2.99465e7 & 0.00541846 & 0.14088 \\
+            Right Accumbens area & Butyrate synthesis & 3.80803e7 & 0.00666094 & 0.167213 \\
+            Right Cerebellum White Matter & Propionate degradation & 2.00749e6 & 0.00742702 & 0.180229 \\\hline\hline
         \end{tabular}

### 2022-07-21 notes/10-fsea.jl

@@ -137,3 +137,84 @@ subset!(fsdf, :pvalue=> ByRow(!isnan))
 fsdf.qvalue = adjust(fsdf.pvalue, BenjaminiHochberg())
 sort!(fsdf, :qvalue)
 CSV.write(datafiles("fsea_o12.csv"))
+
+
+#-
+
+brain_features = Symbol.(replace.(Resonance.brainmeta, "-"=>"_"))
+brgenes = genes[:, .!ismissing.(get(genes, :Left_Thalamus))]
+
+brfsea = DataFrame()
+
+for region in brain_features
+    @warn region
+    cscor = cor(abundances(brgenes), get(brgenes, region), dims=2)
+
+    fsdf = DataFrame(
+        map(collect(keys(neuroactive))) do gs
+            @info gs
+            ixs = neuroactive[gs]
+            isempty(ixs) && return (; region, geneset = gs, U = NaN, median = NaN, mu = NaN, sigma = NaN, pvalue = NaN)
+
+            cs = filter(!isnan, cscor[ixs])
+            isempty(cs) && return (; region, geneset = gs, U = NaN, median = NaN, mu = NaN, sigma = NaN, pvalue = NaN)
+
+            acs = filter(!isnan, cscor[Not(ixs)])
+            mwu = MannWhitneyUTest(cs, acs)
+
+            return (; region, geneset = gs, U = mwu.U, median = mwu.median, mu = mwu.mu, sigma = mwu.sigma, pvalue=pvalue(mwu))
+        end
+    )
+    subset!(fsdf, :pvalue=> ByRow(!isnan))
+    append!(brfsea, fsdf)
+end
+
+
+brfsea.qvalue = adjust(brfsea.pvalue, BenjaminiHochberg())
+sort!(brfsea, :qvalue)
+CSV.write(datafiles("fsea_brain_all.csv"), brfsea)
+
+pretty_table(first(brfsea, 10); backend = Val(:latex))
+
+#-
+
+using Connectomes
+using WGLMakie
+using ColorSchemes
+WGLMakie.activate!()
+
+c = Connectomes.connectomepath() |> Connectome
+brfsea = CSV.read(datafiles("fsea_brain_all.csv"), DataFrame)
+brgroup = groupby(brfsea, :geneset)
+
+#-
+
+gr = first(brgroup)
+gr = subset(gr, :pvalue => ByRow(<(0.05)))
[…]

### 2022-07-21 src/genefamilies.jl

@@ -81,9 +81,10 @@ function get_neuroactive_kos(neuroactivepath=datafiles("gbm.txt"))
        line = split(line, r"[\t,]")
        if startswith(line[1], "MGB")
            (mgb, desc) = line
-           desc = rstrip(replace(desc, r"\bI+\b.*$"=>""))
-           desc = replace(desc, r" \([\w\s]+\)$"=>"")
+           desc = rstrip(replace(desc, r"\b[IV]+\b.*$"=>""))
+           desc = replace(desc, r" \([\w\s\-]+\)"=>"")
            desc = replace(desc, r"^.+ \(([\w\-]+)\) (.+)$"=>s"\1 \2")
+           desc = replace(desc, " (AA"=>"")
            @info "getting unirefs for $desc"
            !in(desc, keys(neuroactive)) && insert!(neuroactive, desc, String[])
        else

### 2022-07-21 src/plotting.jl

@@ -84,9 +84,10 @@ function plot_permanovas!(ax, comms, metadatums; commlabels=[], mdlabels=[], col
     return hm
 end
 
-function plot_fsea(setcors, notcors; label="")
+function plot_fsea(setcors, notcors; label="", bandres=5000)
     fullcors = [setcors; notcors]
-    
+    ncors = length(fullcors)
+
     srt = sortperm(fullcors; rev=true)
     ranks = invperm(srt)
     setranks = Set(ranks[1:length(setcors)])
@@ -94,10 +95,10 @@ function plot_fsea(setcors, notcors; label="")
     setscore =  1 / length(setcors)
     notscore = -1 / length(notcors)
     
-    xs = 1:length(srt)
+    xs = 1:ncors
     ys = cumsum(i ∈ setranks ? setscore : notscore for i in eachindex(ranks))
     
-    t = "Enrichment score - $(round(maximum(ys), digits=3))"
+    t = "Enrichment score - $(round(max(abs.(extrema(ys))...), digits=3))"
     !isempty(label) && (t = string(label, ": ", t))
 
     fig = Figure()
@@ -110,15 +111,16 @@ function plot_fsea(setcors, notcors; label="")
     ax3 = Axis(fig[3,1]; ylabel="correlation", xlabel="rank")
     
     lines!(ax1, xs, ys)
-
     vlines!(ax2, ranks[1:length(setcors)]; color=:black)
 
+
+    rn = ncors > bandres ? round.(Int, range(1, ncors; length=bandres)) : range(1, ncors)
     blow, bup = extrema(fullcors)
-    band!(ax3, xs, fill(blow, length(srt)), fill(bup, length(srt)); color=fullcors[srt])
+    band!(ax3, rn, fill(blow, length(rn)), fill(bup, length(rn)); color=fullcors[srt[rn]])
     
     lower = [x < 0 ? x : 0.0 for x in fullcors[srt]]
     upper = [x > 0 ? x : 0.0 for x in fullcors[srt]]
-    band!(ax3, xs, lower, upper; color=:lightgray)
+    band!(ax3, xs[rn], lower[rn], upper[rn]; color=:lightgray)
 
     
     rowsize!(fig.layout, 2, Relative(1/8))

### 2022-07-30 docs/make.jl
add workflows and docs

@@ -1,0 +1,31 @@
+using Resonance
+using Documenter
+
+DocMeta.setdocmeta!(Resonance, :DocTestSetup, :(using Resonance); recursive=true)
+
+makedocs(;
+    modules=[Resonance],
+    authors="Kevin Bonham, PhD <kbonham@wellesley.edu> and contributors",
+    repo="https://github.com/Klepac-Ceraj-lab/Resonance/blob/{commit}{path}#{line}",
+    sitename="Resonance",
+    format=Documenter.HTML(;
+        prettyurls=get(ENV, "CI", "false") == "true",
+        canonical="https://Klepac-Ceraj-lab.github.io/Resonance",
+        edit_link="main",
+        assets=String[],
+    ),
+    pages=[
+        "Home" => "index.md",
+        "Setup" => "setup.md",
+        "Figures" => [
+            "Figure 1" => "Figures/figure1.md",
+            "Figure 2" => "Figures/figure1.md",
+            "Figure 3" => "Figures/figure1.md",
+        ]
+    ],
+)
+
+deploydocs(;
+    repo="github.com/Klepac-Ceraj-lab/Resonance",
+    devbranch="main",
+)

### 2022-07-30 notes/W-longform.jl

@@ -1,41 +1,0 @@
-using Resonance, CategoricalArrays
-
-#include(joinpath(pathof(Resonance), "..", "..", "notes", "01-metadata_wrangle.jl"))
-#include(joinpath(pathof(Resonance), "..", "..", "notes", "02-species.jl"))
-#include(joinpath(pathof(Resonance), "..", "..", "notes", "03-metabolites.jl"))
-#include(joinpath(pathof(Resonance), "..", "..", "notes", "04-functions.jl"))
-
-omni, etoh, tps, complete_brain, metabolites, species = startup()
-
-for col in names(tps)
-    tps[!, col] = collect(tps[!, col])
-end
-
-tps_specific = stack(tps, ["ageMonths", "cogScore", "ECHOTPCoded", Resonance.brainmeta_underscore...], [:subject, :timepoint])
-subj_specific = stack(tps, [:mother_HHS_Education, :rce], [:subject]) # changed simple_race to rce
-grp = groupby(subj_specific, [:subject, :variable])
-subj_specific = DataFrames.combine(grp, :value=> (v-> coalesce(v...))=> :value)
-
-##
-
-tps_specific = vcat(tps_specific, DataFrame(subject   = get(species, :subject), 
-                                timepoint = get(species, :timepoint),
-                                variable  = fill("sample", nsamples(species)),
-                                value     = samplenames(species))
-)
-
-for s in samples(species)
-    subject   = s.subject
-    timepoint = s.timepoint
-    for f in features(species)
-        variable = name(f)
-        value = species[f, s]
-        push!(tps_specific, (; subject, timepoint, variable, value))
-    end
-end
-
-CSV.write(datafiles("wrangled", "tidy_timepoints_with_brain.csv"), tps_specific)
-CSV.write(datafiles("wrangled", "tidy_subjects.csv"), subj_specific)
-
-
-##

### 2022-07-30 notes/Y-data_requests.jl

@@ -1,185 +1,0 @@
-# Vanja + Aditi CBCL
-
-using Resonance
-omni, etoh, tps, metabolites, species = startup()
-genes = Resonance.read_gfs_arrow()
-
-aditi_sids = readlines(datafiles("aditi_samples.txt"))
-
-CSV.write(joinpath(ENV["SCRATCH_SPACE"], "aditi_taxa.csv"), species[:, aditi_sids])
-CSV.write(joinpath(ENV["SCRATCH_SPACE"], "aditi_genefamilies.csv"), genes[:, aditi_sids])
-
- 
-
-## Korean group, blood pressure
-
-using Resonance
-
-fecalsamples = CSV.read(datafiles("wrangled", "samples.csv"), DataFrame)
-@rsubset! fecalsamples :Fecal_EtOH == "F"
-
-timepoints = CSV.read(datafiles("wrangled", "timepoints.csv"), DataFrame)
-
-function mapbp(label)
-    bp = Dict(
-        "E2a" => "Systolic",
-        "E2b" => "Diastolic",
-        "E2c" => "Heart rate"
-    )
-    for pair in pairs(bp)
-        bp = replace(label, pair)
-    end
-
-    # bp = replace(bp)
-end
-
-
-@assert size(subset(timepoints, "ChildAnthropometryOverTwo::E2a1"=> ByRow(!ismissing), "ChildAnthropometryUnderTwo::E2a1" => ByRow(!ismissing)), 1) == 0
-
-timepoints.Systolic_1 = map(eachrow(timepoints)) do row
-    coalesce(row."ChildAnthropometryOverTwo::E2a1", row."ChildAnthropometryUnderTwo::E2a1")
-end
-timepoints.Systolic_2 = map(eachrow(timepoints)) do row
-    coalesce(row."ChildAnthropometryOverTwo::E2a2", row."ChildAnthropometryUnderTwo::E2a2")
-end
-timepoints.Systolic_3 = map(eachrow(timepoints)) do row
-    coalesce(row."ChildAnthropometryOverTwo::E2a3", row."ChildAnthropometryUnderTwo::E2a3")
-end
-timepoints.Diastolic_1 = map(eachrow(timepoints)) do row
-    coalesce(row."ChildAnthropometryOverTwo::E2b1", row."ChildAnthropometryUnderTwo::E2b1")
-end
-timepoints.Diastolic_2 = map(eachrow(timepoints)) do row
[…]

### 2022-07-30 notes/Z-Khula-pres.jl

@@ -1,171 +1,0 @@
-using Resonance
-omni, etoh, tps, metabolites, species = startup()
-genes = Resonance.read_gfs_arrow()
-
-specu1 = species[:, [!ismissing(a) && 2 < a < 10 for a in get(species, :ageMonths)]]
-shannon!(specu1)
-
-set!(genes, leftjoin(tps, select(omni, [:sample, :subject, :timepoint]), on = [:subject, :timepoint]))
-genesu1 = genes[:, [!ismissing(a) && 2 < a < 10 for a in get(genes, :ageMonths)]]
-shannon!(genesu1)
-
-set!(metabolites, leftjoin(tps, select(etoh, [:sample, :subject, :timepoint]), on = [:subject, :timepoint]))
-metabolitesu1 = metabolites[:, [!ismissing(a) && 2 < a < 10 for a in get(metabolites, :ageMonths)]]
-
-pcspec = pcoa(specu1)
-
-pcgenes = pcoa(genesu1)
-#-
-
-using CairoMakie
-using AlgebraOfGraphics
-
-#-
-
-fig = Figure(resolution=(1000, 1000))
-ax1 = Axis(fig[1,1]; ylabel=mdsaxis(pcspec, 2))
-ax2 = Axis(fig[2,1]; xlabel=mdsaxis(pcspec, 1), ylabel=mdsaxis(pcspec, 2))
-ax3 = Axis(fig[1,2]; ylabel=mdsaxis(pcgenes, 2))
-ax4 = Axis(fig[2,2]; xlabel=mdsaxis(pcgenes, 1), ylabel=mdsaxis(pcgenes, 2))
-
-sc1 = scatter!(ax1, loadings(pcspec, 1), loadings(pcspec,2);
-                    xlabel=mdsaxis(pcspec, 1),
-                    ylabel=mdsaxis(pcspec, 2), 
-                    color=get(specu1, :ageMonths)
-)
-sc2 = scatter!(ax2, loadings(pcspec, 1), loadings(pcspec,2);
-                    color=get(specu1, :shannon),
-                    colormap=:plasma
-)
-
-sc3 = scatter!(ax3, loadings(pcgenes, 1), loadings(pcgenes,2);
-                    xlabel=mdsaxis(pcgenes, 1),
-                    ylabel=mdsaxis(pcgenes, 2), 
-                    color=get(genesu1, :ageMonths)
-)
-sc4 = scatter!(ax4, loadings(pcgenes, 1), loadings(pcgenes,2);
[…]

### 2022-07-30 notes/covid.jl

@@ -1,52 +1,0 @@
-using Resonance
-using AlgebraOfGraphics
-
-timepoints = CSV.read(datafiles("wrangled", "timepoints.csv"), DataFrame)
-
-omnisamples = CSV.read(datafiles("wrangled", "omnisamples.csv"), DataFrame)
-rename!(omnisamples, "DOC"=>"date")
-
-covid = CSV.read(datafiles("wrangled", "covid.csv"), DataFrame)
-rename!(covid, ["SampleID"=>"sample", "CollectionDate"=>"date"])
-
-spec = metaphlan_profiles(datafiles("wrangled", "species.csv"))
-spec = spec[:, intersect(samplenames(spec), omnisamples.sample)]
-
-insert!(spec, select(omnisamples, ["sample", "date"]))
-
-insert!(spec, leftjoin(
-                    select(omnisamples, ["sample", "subject", "timepoint"]),
-                    select(timepoints, ["subject", "timepoint", "ageMonths"]);
-                    on = ["subject", "timepoint"])
-)
-
-metadf = metadata(spec) |> DataFrame
-spec = spec[:, .!ismissing.(metadf.ageMonths)]
-
-p = pcoa(spec)
-
-using Dates
-using CategoricalArrays
-
-covdates = [d < Date(2020, 03, 01) ? "pre" : "post" for d in Date.(get(spec, :date))]
-
-df = DataFrame(loadings(p)[:, 1:6], [:pco1, :pco2, :pco3, :pco4, :pco5, :pco6])
-df.sample = samplenames(spec)
-df.date = Date.(get(spec, :date))
-df.covdates = categorical([d < Date(2020, 03, 01) ? "pre" : "post" for d in df.date])
-df.ageMonths = get(spec, :ageMonths)
-
-plt = data(df) * mapping(:pco1=> "PCo1 ($(round(varexplained(p)[1] * 100; digits=2))) %", 
-                   :pco2=> "PCo2 ($(round(varexplained(p)[2] * 100; digits=2))) %")
-
-draw(plt * mapping(color=:covdates))
-
-plt = data(df) * mapping(:pco3=> "PCo3 ($(round(varexplained(p)[3] * 100; digits=2))) %", 
-                         :pco4=> "PCo4 ($(round(varexplained(p)[4] * 100; digits=2))) %")
-
-draw(plt * mapping(color=:covdates))
-
[…]

### 2022-07-30 notes/genemetab.jl

@@ -1,8 +1,0 @@
-using GLMakie
-using CSV
-using DataFrames
-
-genemetab = CSV.read(datafiles("genemetab.csv"), DataFrame)
-
-scatter(log.(1 .+ genemetab.glutdegr), log.(1 .+ genemetab.glutsynth), genemetab.glutgut,
-    axis=(; xlabel="degradataion", ylabel="synthesis", zlabel="concentration"))

### 2022-07-30 notes/methylation.jl

@@ -1,17 +1,0 @@
-using Resonance
-
-methylation = CSV.read(datafiles("targets_passed.csv"), DataFrame)
-wrangled = CSV.read(datafiles("wrangled.csv"), DataFrame)
-
-jnd = leftjoin(methylation, wrangled, on=["Sample_ID"=>"subject", "Timepoint"=>"timepoint"], matchmissing=:notequal)
-
-jnd.Sample_Name = jnd.sample
-
-files = filter(f-> any(s-> !ismissing(s) && contains(f, s), jnd.sample), readdir("/augusta/echo/analysis/biobakery3/links/metaphlan/"))
-open(datafiles("methylation_files.txt"), "w") do io
-    println.(Ref(io), files)
-end
-
-met = metaphlan_profiles(joinpath.(Ref("/augusta/echo/analysis/biobakery3/links/metaphlan/"), files))
-CSV.write("/home/kevin/Desktop/methylation_taxa.csv", met)
-CSV.write("/home/kevin/Desktop/methylation_taxa_link.csv", select(jnd, ["Sample_ID", "Timepoint", "Sample_Name"]))

### 2022-07-30 notes/10-fsea.jl
get started on dataexport

@@ -175,45 +175,5 @@ CSV.write(datafiles("fsea_brain_all.csv"), brfsea)
 
 pretty_table(first(brfsea, 10); backend = Val(:latex))
 
-#-
-
-using Connectomes
-using WGLMakie
-using ColorSchemes
-WGLMakie.activate!()
-
-c = Connectomes.connectomepath() |> Connectome
-brfsea = CSV.read(datafiles("fsea_brain_all.csv"), DataFrame)
-brgroup = groupby(brfsea, :geneset)
-
-#-
-
-gr = first(brgroup)
-gr = subset(gr, :pvalue => ByRow(<(0.05)))
-
-
-
-#-
-fig = Figure(resolution=(1200,600))
-ax = Axis3(fig[1,1], aspect = :data, azimuth = 0.0pi, elevation=0.0pi)
-hidedecorations!(ax)
-hidespines!(ax)
-
-plotted = Int[]
-
-for (r, m) in zip(gr.region, gr.mu)
-    i = findfirst(l-> contains(l, replace(r, "_"=>"-")), c.parc.Label)
-    i = isnothing(i) && r == "Brain_Stem" ? 83 : i
-    isnothing(i) && continue
-    push!(plotted, i)
-    plot_roi!(i, get(ColorSchemes.viridis, m / 3e7))
-end
-
-for i in setdiff(c.parc.ID, plotted)
-    plot_roi!(i, (:gray, 0.2))
-end
-
-fig
-
 #- 
 

### 2022-07-30 notes/99-exports.jl

@@ -1,0 +1,56 @@
+using Resonance
+using Chain
+using Arrow
+
+#-
+
+metadata = let df = CSV.read("data/timepoints_final.csv", DataFrame)
+    df.has_segmentation = map(s-> s == "true", df.has_segmentation)
+    df[df.has_segmentation .| .!ismissing.(df.cogScore) .| .!ismissing.(df.omni)  .| .!ismissing.(df.etoh), :]
+end
+
+CSV.write("data/timepoint_metadata.csv", select(metadata,
+    "subject", "timepoint", # Identifiers
+    "ageMonths", "race", "ed" => "maternalEd", # Demographics
+    "assessmentDate", "scanDate", # Dates 
+    "cogScore", "omni", "etoh", "has_segmentation" # Measurements
+))
+
+CSV.write("data/brain_measures.csv", select(metadata, 
+    "subject", "timepoint",
+    Resonance.brainmeta...
+))
+
+keepomni = Set(skipmissing(metadata.omni))
+
+#-
+
+species = CSV.read(datafiles("species.csv"), DataFrame)
+select!(species, "features", names(species)[map(n-> n in keepomni, names(species))]...) # Missing 11 samples?!?!
+CSV.write("data/species.csv", species)
+
+
+#-
+
+allfiles = String[]
+
+@info "getting files"
+for (root, dirs, files) in walkdir(ENV["ANALYSIS_FILES"])
+    filter!(f-> contains(splitext(f)[1], Regex(string("genefamilies", '$'))) && !contains(f, r"^FE\d+"), files)
+    append!(allfiles, joinpath.(Ref(root), files))
+end
+
+unique!(allfiles) do f
+    first(split(basename(f), '_')) |> String
+end
+filter(f-> String(first(split(basename(f), '_'))) in keepomni, allfiles)
+
+if !isempty(keepsamples)
+    filter!(f-> String(first(split(basename(f), '_'))) in keepsamples, allfiles)
+    length(allfiles) > 0 || throw(ErrorException("No sample intersection found"))
+end
+
+#-
+
+Resonance.write_gfs_arrow("data"; keepsamples = keepomni)
+Resonance.write_gfs_arrow("data"; kind=keepsamples = keepomni)

### 2022-07-30 src/Resonance.jl

@@ -44,7 +44,6 @@ using Tables
 using SparseArrays
 using CodecZlib
 using FilePaths
-using ProgressLogging
 using CategoricalArrays
 using PERMANOVA
 using LinearAlgebra
@@ -55,7 +54,6 @@ using ThreadsX
 @reexport using DataFrames
 @reexport using CSV
 @reexport using XLSX
-@reexport using DataFramesMeta
 
 const transform = DataFrames.transform
 

### 2023-02-16 manuscript/tex/converted.tex
add initial transfer from gdocs

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{multirow}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\usepackage[normalem]{ulem}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\hypertarget{title}{%
\subsection{Title}\label{title}}

\begin{itemize}
\item
  \begin{quote}
  Gut microbes and their genes are associated with cognition and
  neuroanatomy in children
  \end{quote}
\item
  \begin{quote}
  Gut microbes and their genes are associated with cognition performance
  and brain structure in children
  \end{quote}
\item
  \begin{quote}
  Extensive relationships between the gut microbiota and normal
  neurocognitive development
  \end{quote}
\item
  \begin{quote}
  Extensive relationships between the gut microbiota and typical
  neurocognitive development
  \end{quote}
\item
  \begin{quote}
  The gut microbial metagenome is connected to neurocognition in typical
  child development
  \end{quote}
\item
  \begin{quote}
  The microbial potential in regulating early childhood neurocognitive
  development
  \end{quote}
\end{itemize}

\hypertarget{authors}{%
\subsubsection{Authors}\label{authors}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \begin{quote}
  Bonham 0000-0003-3200-7533
  \end{quote}
\item
  \begin{quote}
  Bottino 0000-0003-1953-1576
  \end{quote}
\item
  \begin{quote}
  McCann 0000-0002-9753-7968
  \end{quote}
\item
  \begin{quote}
  Beauchemin \{ORCID??\}
  \end{quote}
\item
  \begin{quote}
  Elizabeth Weisse \textless elizabeth.weisse2@gmail.com\textgreater{}
  \end{quote}
\item
  \begin{quote}
  \ldots{}
  \end{quote}
\item
  \begin{quote}
  Jonathan O\textquotesingle Muircheartaigh
  \textless{}\href{mailto:jonathanom@kcl.ac.uk}{\uline{jonathanom@kcl.ac.uk}}\textgreater{}
  \end{quote}
\item
  \begin{quote}
  Huttenhower 0000-0002-1110-0096
  \end{quote}
\item
  \begin{quote}
  Muriel Bruchhage
  \textless{}\href{mailto:mmk.bruchhage@gmail.com}{\uline{mmk.bruchhage@gmail.com}}\textgreater{}
  0000-0001-9637-0951
  \end{quote}
\item
  \begin{quote}
  D'Sa 0000-0002-3476-8419
  \end{quote}
\item
  \begin{quote}
  Deoni 0000-0002-0633-6328
  \end{quote}
\item
  \begin{quote}
  Klepac-Ceraj 0000-0001-5387-5706
  \end{quote}
\end{enumerate}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

The gastrointestinal tract, its resident microorganisms and the central
nervous system are connected by biochemical signaling, also known as
"microbiome-gut-brain-axis." Both the human brain and the gut microbiome
have critical developmental windows in the first three years of life,
raising the possibility that their development is co-occurring and
likely co-dependent. Emerging evidence implicates gut microorganisms and
microbiota composition in cognitive outcomes and neurodevelopmental
disorders (e.g., autism and anxiety), but the influence of gut microbial
metabolism on typical neurodevelopment has not been explored in detail.
We investigated the relationship of the microbiome with the neuroanatomy
and cognitive function of 361 healthy children, demonstrating that
differences in gut microbial taxa and gene functions are associated with
overall cognitive function as well as with differences in the size of
multiple brain regions. Using a combination of multivariate linear
models and machine learning (ML) models, we showed that many species,
including \emph{Gordonibacter pamelae} and \emph{Blautia wexlerae}, were
significantly associated with higher cognitive function, while some
species such as \emph{Ruminococcus gnavus} were more commonly found in
children with low cognitive scores regardless of age or maternal
education. Microbial genes for enzymes involved in the metabolism of
neuroactive compounds, particularly short-chain fatty acids such as
glutamate and propionate, were also associated with cognitive function.
In addition, ML models were able to use microbial taxa to predict the
volume of brain regions, with particular taxa often dominating the
feature importance metric for individual brain regions, such as \emph{B.
wexlerae} being the most dominant for the parahippocampal region or
GABA-producing \emph{Bacteroidetes ovatus} for left accumbens. These
findings provide potential biomarkers of neurocognition and may lead to
the future development of targets for early detection and early
intervention.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

The gut and the brain are intimately linked. Signals from the brain
reach the gut through the autonomic nervous system and the endocrine
system, and the gut can communicate with the brain through the vagus
nerve and through endocrine and immune (cytokine) signaling molecules
\href{https://www.zotero.org/google-docs/?i3SRPn}{(Cerdó et al., 2019;
Pronovost and Hsiao, 2019; Sharon et al., 2016; Tognini, 2017)}. In
addition, the products of microbial metabolism generated in the gut can
influence the brain, both indirectly by stimulating the enteric nervous
and immune systems, as well as directly through molecules that enter
circulation and cross the blood-brain barrier. Causal links between the
gut microbiome and neural development, particularly atypical
development, are increasingly being identified
\href{https://www.zotero.org/google-docs/?UsrFKB}{(Spichak et al.,
2021)}. Both human epidemiology and animal models point to the effects
of gut microbes on the development of autism spectrum disorder
\href{https://www.zotero.org/google-docs/?Y8l5A0}{(Laue et al., 2020;
Wan et al., 2021)} and particular microbial taxa have been associated
with depression
\href{https://www.zotero.org/google-docs/?mbaFE3}{(Mayneris-Perxachs et
al., 2022; Valles-Colomer et al., 2019)} and Alzheimer's disease
\href{https://www.zotero.org/google-docs/?0BEGUn}{(Fung et al., 2017;
Kim et al., 2021)}. But information about this ``microbiome-gut-brain
axis'' in normal neurocognitive development remains lacking,
particularly early in life.

The first years of life are critical developmental windows for both the
microbiome and the brain
\href{https://www.zotero.org/google-docs/?XTqhlm}{(Laue et al., 2022)}.
Fetal development \emph{in utero} is believed to be mostly sterile, but
is rapidly seeded at birth through contact with the birth canal (if
birthed vaginally), caregivers, food sources (breastmilk or formula),
and other environmental sources such as antibiotics
\href{https://www.zotero.org/google-docs/?X6N0f4}{(Bäckhed et al., 2015;
Bokulich et al., 2016; Dominguez-Bello et al., 2010; Louwies et al.,
2020)}. The early microbiome is characterized by low microbial
diversity, rapid succession and evolution, and domination by
Actinobacteria, particularly the genus \emph{Bifidobacterium},
Bacteroidetes, especially \emph{Bacteroides}, and Proteobacteria
\href{https://www.zotero.org/google-docs/?VG43uh}{(Koenig et al.,
2011)}. Many of these microbes have specialized metabolic capabilities
for digesting human breast milk, such as \emph{Bifidobacterium infantis}
and \emph{Bacteroides fragilis}
\href{https://www.zotero.org/google-docs/?1dKzKj}{(Sela et al., 2008;
Tso et al., 2021)}. Upon the introduction of solid foods, the gut
microbiome undergoes another categorical transformation, with most taxa
of the infant microbiome being replaced by taxa more reminiscent of
adult microbiomes
\href{https://www.zotero.org/google-docs/?V3lY3p}{(Bäckhed et al.,
2015)}. Many prior studies have focused on either infant microbiomes or
adult microbiomes, since performing statistical analyses across this
transition poses particular challenges. Nevertheless, since this
transition coincides with critical neural developmental windows and
neural synaptogenesis, investigation across this solid-food boundary is
incredibly important
\href{https://www.zotero.org/google-docs/?d29bv8}{(Tau and Peterson,
2010)}.

A child's brain undergoes remarkable anatomical, microstructural,
organizational, and functional changes in the first years of life. By
age 5, a child's brain has reached \textgreater85\% of its adult size,
has achieved near-adult levels of myelination, and the pattern of axonal
connections has been established
\href{https://www.zotero.org/google-docs/?1jUZOO}{(Silbereis et al.,
2016)}. Much of this development occurs in discrete windows or
``sensitive-'' or ``critical periods'' (CPs)
\href{https://www.zotero.org/google-docs/?w4Pcze}{(Knudsen, 2004)} when
neural plasticity is particularly high, and particular modes of learning
and skill development are preferred. The timing of these sensitive
periods is driven in part by genetics, but can also be affected by the
environment, including the gut microbiome
\href{https://www.zotero.org/google-docs/?KTm9lQ}{(Cowan et al., 2020)}.
In fact, emerging evidence suggests that the timing and duration of CPs
may be driven in part by cues from the developing gut microbiome
\href{https://www.zotero.org/google-docs/?mAD1ne}{(Callaghan, 2020)}. As
such, understanding the normal spectrum of healthy microbiome
development and how it relates to normal neurocognitive development may
provide opportunities for identifying atypical development earlier and
offer opportunities for intervention.

To begin to address this need, we investigated the gut microbiome and
neurocognitive development of children from infancy through 10 years of
age in an accelerated longitudinal study design. Gut microbial
communities were assessed using shotgun metagenomic sequencing, enabling
profiling at both the taxonomic and gene-functional level, and
neurocognitive development was measured using expert-assessment of
cognitive function and neuroimaging using magnetic resonance imaging
(MRI). Using a combination of classical statistical analysis and machine
learning, we showed that the development of the gut microbiome, the
brain, and children's cognitive abilities are intimately linked, with
both microbial taxa and gene functions able to predict cognitive
performance and brain structure.

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{cohort-overview-and-summary-data}{%
\subsubsection{Cohort overview and summary
data}\label{cohort-overview-and-summary-data}}

We investigated the co-development of the brain and the microbiome in
early childhood in a cohort of typically-developing children in their
first years of life using a variety of orthogonal microbial and
neurocognitive assessments, including shotgun metagenomic sequencing,
cognitive and behavioral assessments, and neuroimaging. In all, 361
children from RESONANCE, an accelerated longitudinal cohort of early
development \href{https://paperpile.com/c/dPbU4e/klxU}{(Forrest,
Blackwell, and Camargo 2018)} between the ages of 82 days and 10 years
old (median age 2.21 years, Supplementary Figure 1) were included in
this study (Table 1, Figure 1A). To measure cognitive function, we used
age-appropriate assessments that can be normalized to a common, IQ-like
scale (Figure 1B), including the Mullen scale of early learning (MSEL)
for children from birth to 3 years of age
\href{https://www.zotero.org/google-docs/?90hfYr}{(Mullen and others,
1995)}, Wechsler preschool and primary scale of intelligence (WPPSI) for
4--5-year-olds
\href{https://www.zotero.org/google-docs/?wCfx5h}{(Wechsler, 2012)}, and
the Wechsler intelligence scale for children (WISC) for children 6 years
and up \href{https://www.zotero.org/google-docs/?AcNN3S}{(Wechsler,
1949)}.

As expected, the greatest differences in microbial taxa were driven by
age (Figure 1B-C, PCoA axis 1), while older children were primarily
stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
high-\emph{Prevotella copri} (Supplementary Figure 2). Overall variation
in gut microbial genes and also in brain volume profiles was similarly
driven largely by subject age (Figure 1C-F).

\includegraphics[width=6.45833in,height=3.87221in]{media/image3.png}

\textbf{Figure 1: Cohort overview and summary data.}

(A) Study design and overview. Stool samples, cognitive assessments and
neuroimaging were collected from participants in an accelerated
longitudinal design. (B) Cognitive function scores are assessed by
different instruments, but may be normalized to a common IQ-like scale.
(C-D) Principal coordinates analysis using Bray-Curtis dissimilarity on
taxonomic profiles demonstrates high beta diversity, with much of the
first axis of variation explained by increasing age and alpha diversity.
Differences in gene function profiles (E) and neuroimaging (PCA based on
the Euclidean distance of brain region volumes) (F) are likewise
dominated by changes as children age. (G) Permutation analysis of
variance (PERMANOVA) of taxonomic profiles, functional profiles
(annotated as UniRef90s) and neuroimaging, against the metadata of
interest. Variations in taxonomic and functional profiles explain a
modest but significant percent of the variation in cognitive development
in children over 18 months of age. (H) Mantel testing of different
microbial feature matrices, shows overlapping but distinct patterns of
variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
used as cut-offs in some following models.

\textbf{Table1: Subject demographics}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2620}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Group}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Subgroup}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{All}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Under 6 mo}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Over 18 mo}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
N subjects
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
361
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
74
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
257
\end{minipage} \\
\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
Samples
\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
262 (72.6\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
74 (100.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
203 (79.0\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
72 (19.9\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 (0.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
46 (17.9\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
\textgreater2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
27 (7.5\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 (0.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
8 (3.1\%)
\end{minipage} \\
\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
Age (months)
\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
Min
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
2.8
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
2.8
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
18.0
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Max
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
119.3
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
6.0
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
119.3
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Median
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
25.8
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
3.7
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
45.9
\end{minipage} \\
\multirow{2}{*}{\begin{minipage}[b]{\linewidth}\raggedright
Sex
\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
F
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
163 (45.2\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
37 (50.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
114 (44.4\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
M
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
198 (54.8\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
37 (50.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
143 (55.6\%)
\end{minipage} \\
\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
Race
\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
White
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
252 (69.8\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
42 (56.8\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
191 (74.3\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Black
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
40 (11.1\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
16 (21.6\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
21 (8.2\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Asian
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (0.3\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 (0.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (0.4\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Mixed
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
7 (1.9\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (1.4\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
6 (2.3\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
48 (13.3\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
12 (16.2\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
29 (11.3\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Unknown/Declined
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
2 (0.6\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (1.4\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
2 (0.8\%)
\end{minipage} \\
\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
Maternal Ed.
\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
Junior high school
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (0.3\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 (0.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (0.4\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Some high school
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
6 (1.7\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
3 (4.1\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
4 (1.6\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
High school grad
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
37 (10.2\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
13 (17.6\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
16 (6.2\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Some college
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
95 (26.3\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
25 (33.8\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
57 (22.2\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
College grad
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
90 (24.9\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
17 (23.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
69 (26.8\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Grad/professional school
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
123 (34.1\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
15 (20.3\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
104 (40.5\%)
\end{minipage} \\
\midrule()
\endhead
\bottomrule()
\end{longtable}

Several studies have demonstrated links between specific taxa and
measures of anxiety and depression
\href{https://www.zotero.org/google-docs/?X5rfKj}{(Mayneris-Perxachs et
al., 2022; Needham et al., 2022)}, cognitive flexibility in adults
\href{https://www.zotero.org/google-docs/?RGIf8m}{(Magnusson et al.,
2015)}, and atypical neural development
\href{https://www.zotero.org/google-docs/?hwPZKh}{(Liu et al., 2019; Wan
et al., 2021)}. We, therefore, set out to identify whether specific taxa
or gene functions were linked with normal cognitive development in
children. In order to assess whether variations in gut microbial taxa,
their genes, or their metabolism are linked with neurocognitive
development, we tested whether the beta diversity of microbial taxa and
gene functions, as well as variation in brain development as assessed by
neuroimaging, were associated with these contemporaneous measures of
cognitive function using permutation analysis of variance (PERMANOVA).
Due to the large ecological shift in the microbiome that occurs upon the
introduction of solid food, and a relatively wide range of ages when
infants are transitioned to solid food, we considered the pre-transition
(less than 6 months old) and post-transition (over 18 months old)
microbiomes separately. Even still, age was a major driver of variation
in gut microbiomes in children over 18 months old (Figure 1G). We also
found that overall variation in microbial species in children over 18
months old was associated with a small but significant variation in
cognitive function score (Figure 1G, R\textsuperscript{2} = 0.0124, q
\textless{} 0.001), as was variation in microbial gene functions
(R\textsuperscript{2} = 0.0119, q \textless{} 0.001). Variation in
microbial taxa and genes was not significantly associated with cognitive
function in children under 6 months, though this may be due to the low
taxonomic diversity, and broad lack of overlap between taxa in infants.
As expected, age was significantly associated with microbial beta
diversity (taxa R\textsuperscript{2} = 0.0215, and UniRef90s
R\textsuperscript{2} = 0.0245, q \textless{} 0.001), and very strongly
associated with neuroimaging profiles (R\textsuperscript{2} = 0.256, q
\textless{} 0.001).

Consistent with prior studies, different microbial measurement types
captured overlapping variation, with species profiles and gene function
profiles (annotated with clusters of 90\% similarity
\href{https://www.zotero.org/google-docs/?b9rEJS}{(Suzek et al.,
2007)}), both generated from DNA sequencing data, being tightly coupled
(Figure 1H, p \textless{} 0.001). Interestingly, other functional
groupings (Enzyme commission level4 - ECs
\href{https://www.zotero.org/google-docs/?fs8LO7}{(Bairoch, 2000)}, and
KEGG orthologues - KOs
\href{https://www.zotero.org/google-docs/?Xx5duE}{(Kanehisa et al.,
2004)}) overlapped only slightly with taxonomic profiles in both age
cohorts, despite being derived from UniRef90 labels. In children over 18
months, some variation (15.9\%, p \textless{} 0.01) in neuroimaging
overlapped with microbial measures, though this may be due to the
residual variation due to age in both measures.

\hypertarget{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}{%
\subsubsection{Microbial species and neuroactive genes are associated
with cognitive
performance}\label{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}}

To assess whether individual microbial species were associated with
cognitive function, we fit multivariable linear regression
\href{https://www.zotero.org/google-docs/?4qS6UF}{(Mallick et al.,
2021)} to the relative abundance of each species that had at least 15\%
prevalence in a given age group (Figure 2A, N = 92 for 0--120 months, N
= 46 for 0--6 months, N = 97 for 18--120 months). Only \emph{Blautia
wexlerae} was significantly associated (q value = 0.14, β = 0.0015) with
cognitive function in children under 6 months old after adjusting for
age and maternal education (Figure 2B). \emph{B. wexlerae} was
previously shown to be depleted in children with diabetes
\href{https://www.zotero.org/google-docs/?6mU2nI}{(Benítez-Páez et al.,
2020)}, and that oral administration of \emph{B. wexlerae} partially
ameliorated weight gain and inflammation from a high-fat diet in a mouse
model of T2D \href{https://www.zotero.org/google-docs/?y83hpS}{(Hosomi
et al., 2022; Liu et al., 2021)}. In children over 18 months of age,
several microbial species were significantly enriched (q-value
\textless{} 0.20) in children with higher cognitive function scores,
including \emph{Gordonibacter pamelaeae}, which produces the
neuroprotective metabolite urolithin
\href{https://www.zotero.org/google-docs/?HGtTEY}{(Gong et al., 2022;
Selma et al., 2014)}, \emph{Asaccharobacter celatus} and
\emph{Adelcreutzia equolifaciens}, which produce phytoestrogen-derived
equol \href{https://www.zotero.org/google-docs/?HCKZqm}{(Maruo et al.,
2008; Thawornkuno et al., 2009)}, and the SCFA-producing probiotic
species such as \emph{Eubacterium eligens} and \emph{Faecalibacterium
prausnitzii} \href{https://www.zotero.org/google-docs/?3eyheK}{(Ghosh et
al., 2020; Lopez-Siles et al., 2017; Mukherjee et al., 2020)} (Figure
2B).

\includegraphics[width=6.5in,height=4.83333in]{media/image4.png}

\textbf{Figure 2 - Taxa and gene functional groups are associated with
cognitive function.}

(A) Volcano plot of multivariable linear model results showing the
relationship between individual taxa and cognitive function score in
children over 18 months of age, controlling for age, maternal education,
and sequencing depth. Taxa that were significant after FDR correction (q
\textless{} 0.2) are colored red. (B) For taxa that were significantly
associated with cognitive function, heatmaps of prevalence, mean
relative abundance, and correlation with cognitive function in different
age groups. (C-F) Enrichment plots for selected neuroactive gene sets
used in feature set enrichment analysis (FSEA). (G) summary of FSEA
results across all samples (left) as well as in the under 6 months
(middle) and over 18 months (right) subsets, colored based on the
significance of the association. Markers indicate the individual
correlation of genes within a gene set, and vertical bars represent the
median correlation of that gene set.

Given that different microbial species might occupy the same metabolic
niche in different individuals, we hypothesized that microbial genes
grouped by functional activity would be associated with cognition. To
test this, we performed feature set enrichment analysis (FSEA) on groups
of genes with neuroactive potential
\href{https://www.zotero.org/google-docs/?xQURsj}{(Valles-Colomer et
al., 2019)} and concurrent cognitive function score (Table 2,
\protect\hyperlink{_i7fd3z19jdvg}{\uline{Figure 2}}C-G) and found that
several metabolic pathways were significantly enriched or depleted in
children with higher cognitive function scores. This was true both when
considering all age groups together, though the enrichment of most
pathways was more pronounced in children under 6 months or over 18
months. For example, genes for degrading the 3-carbon SCFA propionate
were significantly depleted in children with higher cognitive function
scores across all age groups tested (Table 2; propionate degradation I,
under 6 months, enrichment score (E.S.) = -0.542, corrected p-value (q)
= 0.020; over 18 months, E.S. = -0.674, q = 0.041). Interestingly, genes
for propionate synthesis were also significantly depleted in higher
scoring children over 18 months (E.S. -0.676, q = 0.023), as were genes
for synthesizing the 2 carbon SCFA acetate in children under 6 months
old (acetate synthesis I, E.S. = -0.194, q = 0.153; acetate synthesis
II, E.S. = -0.342, q = 0.020; acetate synthesis III, E.S. = -0.31, q =
0.052). SCFAs are produced by anaerobic fermentation of dietary fiber
and have been linked with immune system regulation as well as directly
with brain function
\href{https://www.zotero.org/google-docs/?2Z0sKZ}{(Dalile et al.,
2019)}.

Synthesis of menaquinone (vitamin K) was also negatively associated with
cognitive function score in older children (menaquinone synthesis I,
E.S. = -0.170, q = 0.0183). Menaquinone has several isoforms, one of
which, MK-4, has been found to be decreased in children diagnosed with
ASD \href{https://www.zotero.org/google-docs/?dTA3NO}{(Dong et al.,
2021)} and is potentially neuroprotective in both rodents and humans
\href{https://paperpile.com/c/dPbU4e/cTO9}{(Elkattawy et al., 2022)}. In
children over 18 months old, genes for the synthesis of the amino acids
glutamate and tryptophan were significantly enriched in children with
higher cognitive function scores (Glutamate synthesis I, E.S. = 0.242, q
= 0.047; Tryptophan synthesis, E.S. = 0.119, q = 0.041). Glutamate is a
critical neurotransmitter controlling neuronal excitatory/inhibitory
signaling along with gamma-aminobutyric acid (GABA), and their balance
in the brain controls neural plasticity and learning, particularly in
the developing brain
\href{https://www.zotero.org/google-docs/?QubMhx}{(Cohen Kadosh et al.,
2015; Palomo-Buitrago et al., 2019)}. Tryptophan metabolism, including
microbial metabolism of tryptophan, has previously been linked with
autism in children
\href{https://www.zotero.org/google-docs/?GPaUZE}{(Hoshino et al., 1984;
Xiao et al., 2021)}. Taken together, these results suggest that
microbial metabolic activity, particularly the metabolism (synthesis and
degradation) of neuroactive compounds may have effects on cognitive
development.

\textbf{Table 2 - Feature set enrichment analysis on neuroactive
microbial gene sets}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2300}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1259}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1525}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} &
\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2785} + 2\tabcolsep}}{%
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{All}
\end{minipage}} &
\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2591} + 2\tabcolsep}}{%
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{under 6 months}
\end{minipage}} &
\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2324} + 2\tabcolsep}@{}}{%
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{over 18 months}
\end{minipage}} \\
\begin{minipage}[b]{\linewidth}\raggedright
geneset
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{E.S.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{q value}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{E.S.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{q value}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{E.S.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{q value}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Acetate syn. I
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1223
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.5549
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1944
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.1533}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.1269
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.6023
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Acetate syn. II
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2201
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0660}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.3424
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0198}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0946
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.9199
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Acetate syn. III
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2104
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0859}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.3098
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0518}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0857
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.9199
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Glutamate deg. I
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.4037
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.1802}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2840
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.8159
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2489
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.6023
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Glutamate syn. II
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0953
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.6022
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1342
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.9802
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.2421
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0466}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Menaquinone syn. I
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0853
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.5635
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0833
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.8159
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1699
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.1825}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Propionate deg. I
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.4612
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.1302}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5421
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0198}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6738
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0414}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Propionate syn. I
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.3344
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.5549
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.3201
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.5713
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6760
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0227}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Tryptophan syn.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0573
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.9742
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0725
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.8159
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.1186
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0414}
\end{minipage} \\
\midrule()
\endhead
\bottomrule()
\end{longtable}

\hypertarget{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}{%
\subsubsection{Gut microbial taxonomic and functional profiles predict
cognitive
function}\label{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}}

FSEA relies on understanding functional relationships between individual
genes. However, because the relationships between individual taxa are
still largely unknown, we turned to Random Forest (RF) models, an
unsupervised non-parametric machine learning (ML) method that enables
the identification of underlying patterns in large numbers of individual
features (here, microbial species). Previous studies have reported
successful use of RFs for processing highly-dimensional and sparse data
from the domain of genomics
\href{https://www.zotero.org/google-docs/?C3CSPB}{(Amaratunga et al.,
2008; Brieuc et al., 2018; Chen and Ishwaran, 2012; Franzosa et al.,
2019; Stephan et al., 2015)}, along with other works where it was used
to predict cognitive conditions related to Alzheimer's disease in
different scenarios
\href{https://www.zotero.org/google-docs/?GiadGy}{(Ardekani et al.,
2017; Velazquez et al., 2021)}. Additionally, given the sequential
nature of variable consideration in each tree, RFs are naturally able to
work out complex input feature interactions, such as those present in a
microbiome-wide study, without the necessity to explicitly compute
interaction terms.

\includegraphics[width=6.5in,height=5.19444in]{media/image2.png}

\textbf{Figure 3 - Random forest models predict concurrently measured
cognitive function.}

Comparison of RF predictor importance versus linear models for children
between birth and 6 months old (A), and for those older than 18 months
(B). Colors represent whether the species belong to the group of
top-important features that account for 80\% of the cumulative
importance on the RF model, if that species was significant (q
\textless{} 0.2) in linear models, both, or neither. Ranked feature
importance for taxa in RF models for children between birth and 6 months
old (C), and for those older than 18 months (D). Taxa that are important
for RF models both for children under 6 months and for those over 18
months (E), only for children under 6 months (F), and only for children
over 18 months (G)

Given that gut microbial profiles, as well as neurocognitive
development, may partially reflect socioeconomic and demographic
factors, we assessed the performance of RF regressors where maternal
education (a proxy of socioeconomic status (SES)), sex, and age were
included as possible predictors, either alone or in combination with
microbial taxonomic profiles (Table 3).

\textbf{Table 3. Benchmark metrics for the cognitive assessment score
prediction models. Confidence intervals are calculated from the
distribution of metrics from repeated CV at a confidence level of 95\%}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1456}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0960}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3584}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Subject Ages (months)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Microbial feature
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Demo.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Test set correlation (± C.I.)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Test set Root-mean-square error (± C.I.)
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
0 to 6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.14 ± 0.01
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
13.01 ± 0.05
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
0 to 6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
taxa
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.10 ± 0.01
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
12.66 ± 0.05
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
0 to 6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
taxa
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.13 ± 0.01
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
12.70 ± 0.05
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
0 to 6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
genes
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.01 ± 0.01
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
12.56 ± 0.05
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
0 to 6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
genes
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.01 ± 0.01
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
12.56 ± 0.05
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
18 to 120
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.506 ± 0.003
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
16.27 ± 0.04
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
18 to 120
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
taxa
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.363 ± 0.003
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
17.66 ± 0.04
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
18 to 120
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
taxa
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.429 ± 0.003
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
17.29 ± 0.04
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
18 to 120
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
genes
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.303 ± 0.003
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
18.01 ± 0.04
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
18 to 120
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
genes
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.333 ± 0.004
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
17.86 ± 0.04
\end{minipage} \\
\midrule()
\endhead
\bottomrule()
\end{longtable}

As with linear models, RF models for children under 6 months old were
less generalizable (mean test-set correlation -0.13, mean RMSE 12.70),
but RFs were consistently able to learn the relationship between taxa
and cognitive function scores in children over 18 months of age (mean
test-set correlation 0.429, mean RMSE 17.29). For both age groups,
species that were important in RF overlapped with those that were
significant when testing the relationship with linear models (Figure
3A-B, Supplementary Table XXA-B), though there were substantial
differences. For children under 6 months, only \emph{B. wexlerae} was
significantly associated with cognitive function in linear models, but
was ranked 20th in importance for RF models. The importance metric
employed (MDI) is a measure of how well the variable differentiates
(splits) sets of samples by creating subgroups that reduce the
intra-group deviations, while also accounting for the amount of samples
affected by a split. All taxa significantly associated with cognitive
function score in children over 18 months using LMs belong to the
top-ranking group responsible for 80\% of the total relative importance,
except for \emph{Eubacterium ramulus} (Table 3, Figure 3A).

Interestingly, several taxa highly ranked in importance in both age
groups, including several that were significant in LMs for children over
18 months old, including \emph{R. gnavus} (0--6 months, rank = 13;
18--120 months, rank = 7) and \emph{G. pamelaeae} (0--6 months, rank =
16; 18--120 months, rank = 20), while others such as \emph{Allistipes
finegoldii} were age-group specific. Several taxa important in RF models
were not statistically significant when using linear models after
multiple hypothesis correction. However, these taxa had small nominal
p-values. For example, \emph{Erysipelatoclostridium ramosum} was the
most important feature in RF models for children under 6 months old and
had an LM p-value of 0.04. Subject age was consistently ranked highly in
feature importance, which could indicate that decision branches based on
microbial taxa have increased purity when considering the subject's age
or that age itself is a useful predictor.

\hypertarget{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}{%
\subsubsection{Gut microbial taxonomic profiles predict brain structure
differences}\label{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}}

If there are causal effects of microbial metabolism on cognitive
function, they might be reflected in changes in neuroanatomy. We again
employed a Random Forest modeling approach to associate gut taxonomic
profiles with individual brain regions identified in MRI scans,
normalized to total brain volume. Some brain regions were more readily
predicted by RF models trained on microbial taxa (Table 4, Supplementary
Table XX), in particular those that were highly correlated with age.
These included the L/R lingual gyrus (mean RF correlations, Left =
0.421, Right = 0.434; relative age importances, Left = 7.5\%, Right =
8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
0.200, Right = 0.273; relative age importances, Left = 3.7\%, Right =
6.3\%). In many cases, however, age was not an important variable in
high-performing models, such as that for the left accumbens area (mean
RF correlations = 0.288; relative age importances = 1.2\%).

\textbf{Table 4. Summary statistics for the neuroanatomy prediction
benchmarks}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Statistic
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Mean absolute proportional error (MAPE)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Correlation coefficient (R)
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
mean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.072
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.137
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Standard deviation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.028
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.136
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
maximum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.206
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.605
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
75th percentile
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.082
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.195
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
50th percentile
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.067
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.134
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
25th percentile
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.056
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.062
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
minimum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.035
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.147
\end{minipage} \\
\midrule()
\endhead
\bottomrule()
\end{longtable}

We also observed that many brain regions had high accordance across the
left and right hemispheres in terms of both model performance and
microbial feature importance. In contrast, other regions had substantial
differences between the hemispheres. For example, the left accumbens
area, which plays an important role in reward circuits
\href{https://www.zotero.org/google-docs/?zVicU7}{(Ernst et al., 2005;
Yau et al., 2012)}, has one of the highest test-set correlations of our
brain region models (R = 0.288), as compared to the right accumbens
models which could not adequately generalize, and had a negative mean
test-set correlation (R = -0.041). Most healthy individuals have a
rightward asymmetry in the nucleus accumbens, and reduced asymmetry has
been linked to substance use disorder in young adults
\href{https://www.zotero.org/google-docs/?V45KyN}{(Cao et al., 2021)}.
Feature importance for models of the left accumbens area were dominated
by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.4\%
relative importance), \emph{B. ovatus} (3.7\% relative importance), and
\emph{B. uniformis} (3.0\% relative importance).. The accumbens area is
associated with reward control, and in individuals diagnosed with ADHD,
it has been shown to have a divergent neuromorphology
\href{https://www.zotero.org/google-docs/?kLfEQ5}{(Hoogman et al.,
2017)}. Independently, \emph{B. ovatus}, \emph{B. uniformis} and
\emph{B. vulgatus} have been linked to ADHD
\href{https://www.zotero.org/google-docs/?luO0zh}{(Wang et al., 2020)}.
In fact, there are studies showing that alterations on the striatal
dopamine transporters can cause effects resembling hyperactivity and
attention deficit
\href{https://www.zotero.org/google-docs/?UxS47e}{(Yael et al., 2019)},
and that \emph{B. uniformis} is gut-microbial modulator of the brain
dopamine transporter
\href{https://www.zotero.org/google-docs/?L1cWJS}{(Hartstra et al.,
2020)}.

Many of the taxa identified included taxa also identified in LMs and RF
models of cognitive function. Interesting to note, while RF models for
multiple brain regions had many important microbial taxa, others were
dominated by a small number of taxa. In general, for all segments, a
consistent number of species between 14 and 22 (median = 20) was
responsible for one third of the fitness-weighted cumulative importance
(see Methods), regardless of the model benchmarks. We proceeded to
select a subset of the most important taxa (averaged over all segments)
and the segments whose importance was more heavily loaded to further
analyze the multiple relationships unveiled by the RF models
(\textbf{Figure 4 A and B})

\includegraphics[width=6.5in,height=5.19444in]{media/image1.png}

\textbf{Figure 4: Microbial feature importance in predicting brain
volumes in children over 18 months of age.}

(A) Average test-set correlations for prediction of MRI segmentation
data from microbiome and demographics on select segments after repeated
cross validation. (B) Heatmap of average individual relative taxa
importances on each brain segment. Importances are reported as
proportions relative to the sum of importances for each model - since
every model is trained with 132 features, and even distribution of
importance would be 0.75\% for each feature.

. Segments and taxa ordered by HCA on a select list of species with high
ranks on average importances, and their respective highest-load
segments. (C)

Our analysis revealed two major patterns of importance distribution from
taxa over the brain segments; some species portrayed high contributions
to multiple different segments, while others contributed modestly to
just one or two brain segments. Notable cases of the first pattern
included seven species - \emph{Anaerostipes hadrus}, \emph{Bacteroides
vulgatus}, \emph{Fusicatenibacter saccharivorans}, \emph{Ruminococcus
torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
\emph{Blautia wexlerae} - that combined, account for approximately one
third of the cumulative relative importances, computed after subsetting
on the taxa of interest (Figure 4C).

Among these, the most important variable is \emph{Anaerostipes hadrus},
which is a butyrate-producing anaerobe that has been positively
associated with cognitive function
\href{https://www.zotero.org/google-docs/?AVJiwu}{(Kant et al., 2015; Li
et al., 2022)}. Its importance is, however, heavily loaded on the volume
of entire major lobes (frontal, parietal), which cannot be readily
correlated to specific cognitive functions. After accounting for the
major lobes, it is found to be highly important on the prediction of the
cerebellar vermal lobules VIII-X, pars opercularis, cuneus and
precuneus, and anterior-cingulate (Figure 4B).

To better understand this relationship, we performed hierarchical
cluster analysis, which revealed a well defined cluster of species with
high importance loadings on close segments of the lower temporal and
close occipital lobe, to which \emph{A. hadrus} belongs. The other
species on this cluster are \emph{B. wexlerae}, \emph{R. torques},
\emph{R. intestinalis, R. bicirculans} and \emph{F. saccahrivorans}, who
contribute heavily to the entorhinal, fusiform, lingual and
parahippocampal segments.

This group of taxa and their related segments drew our attention because
they contained \emph{B. wexlerae}, a highlight from the cognitive
assessment results. Previous works exploring Parkinson Disease patients
found out that issues in the cognitive task of confrontation naming were
positively correlated with thinning in the fusiform gyrus and
parahippocampal gyrus
\href{https://www.zotero.org/google-docs/?rlO6eq}{(Pagonabarraga et al.,
2013)}. Additionally, the left and right parahippocampal, where \emph{B.
wexlerae} had the highest importance in both models, are important in
visual/spatial processing and memory
\href{https://www.zotero.org/google-docs/?wkkWZT}{(Aminoff et al.,
2013)}. \emph{B. wexlerae} can also produce acetylcholine in the gut
\href{https://www.zotero.org/google-docs/?jBoxSm}{(Hosomi et al.,
2022)}, and this molecule plays an important role in modulating memory
function \href{https://www.zotero.org/google-docs/?BQ0j2c}{(Haam and
Yakel, 2017)}.

Another notable importance cluster contains taxa associated with the
basal forebrain, especially the cingulate and the accumbens. This
cluster contains the previously-discussed \emph{B. ovatus} and \emph{B.
uniformis,} but also \emph{Alistipes finegoldii and Streptococcus
salivaris.} Reduction in the nucleus accumbens has been associated with
depression symptoms
\href{https://www.zotero.org/google-docs/?2aSZIr}{(Wacker et al.,
2009)}, and increased levels of the \emph{Alistipes} genus have been
observed in patients with depressive disorder
\href{https://www.zotero.org/google-docs/?bhhSZB}{(Jiang et al., 2015)}.
While these reports are important, simultaneously probing of the gut
metagenome and brain structure relationship is novel.

Both \emph{A. equolofaciens} and \emph{A. celatus,} two closely-related
equol-producing species, are examples of the second contribution
pattern, and had high importance in predicting the relative volume of
the right anterior cingulate (respectively, 3.0\% and 2.6\% relative
importances), which has been linked to social cognition and reward-based
decision making \href{https://www.zotero.org/google-docs/?Wi8Bpq}{(Apps
et al., 2016; Boes et al., 2008; Bush et al., 2002)}. Equol has a strong
estrogenic effect
\href{https://www.zotero.org/google-docs/?PqgFU7}{(Setchell et al.,
2005)}, and in anterior cingulate, estrogen, has been shown to regulate
pain-related aversion
\href{https://www.zotero.org/google-docs/?5huRIR}{(Xiao et al., 2013)}.
Another example of this pattern is \emph{R. gnavus,} which was the only
species with significant negative association with cognitive function.
It is heavily associated with the left pars opercularis (2.7\% relative
importance), a relationship that may be explained by the emerging
understanding that this species is increased in individuals with insulin
resistance and obesity
\href{https://www.zotero.org/google-docs/?lwzWUv}{(Ley et al., 2006)},
conditions that are known to produce structural abnormalities in the
brain \href{https://www.zotero.org/google-docs/?NlkFr6}{(Opel et al.,
2021)}. Finally, \emph{Coprococcus comes} displays an importance
distribution that splits almost evenly among the two previously reported
clusters. Its loading on the prediction of the left posterior cingulate
is the highest for a microbe in RF models (4.0\% relative importance)
(only age had higher relative importance in any model), while also being
one of the most important predictors for neighboring areas of the
posterior cingulate such as the pars opercularis (relative importances,
Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
precentral (2.3\% relative importance) and paracentral lobes (relative
importances, Left = 2.6\%, Right = 1.8\%).

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

The relationship between the gut microbiome and brain function via the
gut-microbiome-brain axis has gained increasing acceptance largely as a
result of human epidemiological studies investigating atypical
neurocognition (eg anxiety and depression, neurodegeneration, attention
deficit / hyperactivity disorder, and autism) and mechanistic studies in
animal models. The results from these studies point to the possibility
that gut microbes and their metabolism may be causally implicated in
cognitive development, but this study is the first to our knowledge that
directly investigates microbial species and their genes in relation to
typical development in young children. Understanding the gut-brain-axis
in early life is particularly important, since differences or
interventions in early life can have outsized and longer-term
consequences than those at later ages. Further, even in the absence of
causal impacts of microbial metabolism, identifying risk factors that
could point to other early interventions would also have value.

The use of shotgun metagenomic sequencing enabled us to get
species-level resolution of microbial taxa. A previous study of
cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing,
and showed that genera from the Lachnospiraceae family as well as
unclassified Clostridiales (now Eubacteriales) were associated with
higher scores on the Ages and Stages Questionnaire
\href{https://www.zotero.org/google-docs/?SB5Lwa}{(Sordillo et al.,
2019)}. However, each of these clades encompass dozens of genera with
diverse functions, each of which may have different effects. Indeed,
several of the taxa that were positively associated with cognitive
function in this study, including \emph{B. wexlerae}, the only species
identified by linear models in children under 6 months, \emph{D.
longicatena}, \emph{R. faecis}, and \emph{A. finegoldii} are
Clostridiales, as is \emph{R. gnavus}, which we found was negatively
associated with cognitive function (Figure 2A-B). This kind of
species-level resolution is typically not possible with amplicon
sequencing.

We identified several species in the family Eggerthelaceae that were
associated with cognitive function, including \emph{Gordonibacter
pamelaeae, Aldercreutzia equolofaciens, Asaccharobacter celatus}
(formerly regarded a subspecies of \emph{A. equolofaciens}
\href{https://www.zotero.org/google-docs/?OEhguI}{(Takahashi et al.,
2021)}), and \emph{Eggerthella lenta}. Many members of this family are
known in part due to unique metabolic activities. For example, \emph{A.
equolofaciens} produces the nonsteroidal estrogen equol from isoflaven
found in soybeans
\href{https://www.zotero.org/google-docs/?I2qNyB}{(Wang et al., 2005)},
and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
(found in pomegranates and some berries) into urolithin, which has been
shown in some studies to have a neuroprotective effect
\href{https://www.zotero.org/google-docs/?oTN8Lw}{(Gong et al., 2022;
Selma et al., 2014)}. \emph{E. lenta} has been extensively studied for
its ability to metabolize drug compounds such as the plant-derived heart
medication digoxin
\href{https://www.zotero.org/google-docs/?tH2Qeq}{(Haiser et al.,
2013)}. The metabolic versatility of this clade, and the large number of
species that are associated with cognition make these microbes prime
targets for further mechanistic studies.

In addition to improved species-level resolution, shotgun metagenomic
sequencing also enables gene-functional insight. We showed here that
genes for the metabolism of SCFAs, both their degradation and synthesis,
are associated with cognitive function scores. However, while the
differential abundance of genes for the metabolism of neuroactive
compounds like these is suggestive, it is difficult to reason about the
relationship between levels of these genes and the gut concentrations of
the molecules their product enzymes act on. For example, while it might
be intuitive to reason that increased levels of menaquinone synthesis
genes is indicative of increased menaquinone, it could be the case that
menaquinone deficiency selects for microbes that can synthesize it. For
the same reason, increased propionate degradation genes may
counterintuitively be indicative of high levels of propionate in the gut
lumen, since high propionate would select for microbes that can
metabolize it. For this reason, future studies coupling shotgun
metagenomics with stool metabolomics could improve our understanding of
the relationship between microbial metabolism and cognitive development.
Further, strain-level analysis linking specific gene content in species
of interest could further refine targeted efforts at identifying
specific metabolic signatures of microbe-brain interactions.

\{\{Guilherme add a paragraph about the usefulness of ML\}\}

The use of multiple age-appropriate cognitive assessments that could be
normalized to a common scale enabled us to analyze microbial
associations across multiple developmental periods, but carries several
drawbacks. In particular, the test-retest reliability, as well as
systematic differences between test administrators may introduce
substantial noise into these observations, particularly in the youngest
children. In addition, our study period overlapped with the beginning of
the COVID-19 pandemic, and we and others have observed some reduction in
measured scores for children that were assessed after the implementation
of lockdowns. In our subject set for this study, these effects are more
pronounced in some age groups due to our sampling schedule
\href{https://www.zotero.org/google-docs/?L8GnRj}{(Blackwell et al.,
2022; Deoni et al., 2021)} (Supplementary Figure 3).

This analysis allowed us to establish links between microbial taxa and
their functional potential with cognition and brain structure. Although
we cannot test causality or the chemistry behind the interactions
between gut microbial taxa, gut, and brain, this study provides clear
and statistically significant associations between the infant and early
child gut microbiota and neurocognition. Future studies should focus on
characterizing the early-life microbiome and neurocognitive development
across different geographic regions and lifestyles such as covering
traditionally understudied low-resource urban, peri-urban and rural
communities to obtain the more comprehensive understanding of the
variability within the different gut microbiomes reflects on
neurocognition. These studies would also provide us with the wealth of
data on different strains from the same species to better understand the
effect of genes and their products. Furthermore, culturing and microbial
community enrichment studies combined with genetic manipulation and
genomic approaches to understand microbial metabolism at the molecular
level is the key, as the metabolic functions shape and influence the
human host and its health. The discovery of the neuroactive metabolites
could provide us with biomarkers for early detection or necessary
medicinally useful molecules that can be applied in intervention.

\hypertarget{material-and-methods}{%
\subsection{Material and Methods}\label{material-and-methods}}

\hypertarget{study-ethics}{%
\subsubsection{Study Ethics}\label{study-ethics}}

All procedures for this study were approved by the local institutional
review board at Rhode Island Hospital, and all experiments adhered to
the regulation of the review board. Written informed consent was
obtained from all parents or legal guardians of enrolled participants.

\hypertarget{participants}{%
\subsubsection{Participants}\label{participants}}

Data used in this study were drawn from the ongoing longitudinal
RESONANCE study of healthy and neurotypical brain and cognitive
development, based at Brown University in Providence, RI, USA. The
RESONANCE study is part of the NIH initiative Environmental influences
on Child Health Outcomes (ECHO)
\href{https://www.zotero.org/google-docs/?0tZanX}{(Forrest et al., 2018;
Gillman and Blaisdell, 2018)}, a longitudinal observational study of
healthy and neurotypical brain development that spans the fetal and
infant to adolescent life stages, combining neuroimaging (magnetic
resonance imaging, MRI), neurocognitive assessments, bio-specimen
analyses, subject genetics, environmental exposures such as lead, and
rich demographic, socioeconomic, family and medical history information.
From the RESONANCE cohort, 361 typically-developing children between the
ages of 2.8 months and 10 years old (median age 2 years, 2 months) were
selected for analysis in this study.

General participant demographics are provided in Table 1. Children are
representative of the RI population. Children in the RESONANCE cohort
were born full-term (\textless{} 37 weeks gestation) with height and
weight normal for gestational age, and from uncomplicated singleton
pregnancies. Children with known major risk factors for developmental
abnormalities at enrollment were excluded. In addition to screening at
the time of enrollment, on-going screening for worrisome behaviors using
validated tools was performed to identify at-risk children and remove
them from subsequent analysis.

Exclusion criteria included: \emph{in utero} exposure to alcohol,
cigarette or illicit substance exposure; preterm (\textless{} 37 weeks
gestation) birth; small for gestational age or less than 1500 g; fetal
ultrasound abnormalities; preeclampsia, high blood pressure, or
gestational diabetes; 5 minute APGAR scores \textless{} 8; NICU
admission; neurological disorder (e.g., head injury resulting in loss of
consciousness, epilepsy); and psychiatric or learning disorder
(including maternal depression) in the infant, parents, or siblings
requiring medication in the year prior to pregnancy.

Demographic and other non-biospecimen data such as race and ethnicity,
parental education and occupation, feeding behavior (breast- and
formula-feeding), child weight and height, were collected through
questionnaires or direct examination as appropriate. All data were
collected at every assessment visit, if possible.

\hypertarget{cognitive-assessments}{%
\subsubsection{Cognitive Assessments}\label{cognitive-assessments}}

Overall cognitive function was assessed using age-appropriate methods.
For children from birth to 30 months, we used an Early Learning
Composite as assessed via the Mullen Scales of Early Learning (MSEL)
\href{https://www.zotero.org/google-docs/?NSyykd}{(Mullen and others,
1995)}, a standardized and population-normed tool for assessing fine and
gross motor, expressive and receptive language, and visual reception
functioning in children from birth through 68 months of age. The
Wechsler Intelligence Quotient for Children (WISC)
\href{https://www.zotero.org/google-docs/?OopdA1}{(Wechsler, 2012)} is
an individually administered standard intelligence test for children
aged 6 to 16 years. It derives a full scale intelligence quotient (IQ)
score, which we used to assess overall cognitive functioning. The fourth
edition of the Wechsler Preschool and Primary Scale of Intelligence
(WPPSI-IV) is an individually administered standard intelligence test
for children aged 2 years 6 months to 7 years 7 months, trying to meet
the increasing need for the assessment of preschoolers. Just as the
WISC, it derives a full scale IQ score, which we used to assess overall
cognitive functioning.

\hypertarget{stool-sample-collection-and-sequencing}{%
\subsubsection{Stool Sample Collection and
Sequencing}\label{stool-sample-collection-and-sequencing}}

Stool samples (n=493) were collected by parents in OMR-200 tubes
(OMNIgene GUT, DNA Genotek, Ottawa, Ontario, Canada), stored on ice, and
brought within 24 hrs to the lab in RI where they were immediately
frozen at -80 ˚C. Stool samples were not collected if the subject had
taken antibiotics within the last two weeks. DNA extraction was
performed at Wellesley College (Wellesley, MA). Nucleic acids were
extracted from stool samples using the RNeasy PowerMicrobiome kit,
excluding the DNA degradation steps. Briefly, the samples were lysed by
bead beating using the Powerlyzer 24 Homogenizer (Qiagen, Germantown,
MD) at 2500 rpm for 45 s and then transferred to the QIAcube (Qiagen,
Germantown, MD) to complete the extraction protocol. Extracted DNA was
sequenced at the Integrated Microbiome Resource (IMR, Dalhousie
University, NS, Canada).

Shotgun metagenomic sequencing was performed on all samples. A pooled
library (max 96 samples per run) was prepared using the Illumina Nextera
Flex Kit for MiSeq and NextSeq from 1 ng of each sample. Samples were
then pooled onto a plate and sequenced on the Illumina NextSeq 550
platform using 150+150 bp paired-end ``high output'' chemistry,
generating 400 million raw reads and 120 Gb of sequence per plate.

\hypertarget{machine-learning-for-cognitive-development}{%
\subsubsection{Machine learning for cognitive
development}\label{machine-learning-for-cognitive-development}}

Prediction of cognitive scores was carried out as a set of regression
experiments targeting real-valued continuous assessment scores.
Different experiment sets were designed to probe how different
representations of the gut microbiome (taxonomic profiles, Functional
profiles encoded as ECs) would behave, with and without the addition of
demographics (sex and maternal education as a proxy of socioeconomic
status) on participants from different age groups. Age (in months) was
provided as a covariate for all models (Table 3).

\textbf{Table 5. Experimental design and input composition for Random
Forest experiments}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1797}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2915}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Input set
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Age bracket
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Microbiome encoding type
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Demographics

Provided?

(sex, education)
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 to 6 months
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Not provided
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
yes
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 to 6 months
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Taxonomic profile
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
no
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
3
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 to 6 months
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Taxonomic profile
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
yes
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
4
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 to 6 months
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functional Profile (ECs)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
no
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
5
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 to 6 months
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functional Profile (ECs)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
yes
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
18 to 120 months
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Not provided
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
yes
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
7
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
18 to 120 months
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Taxonomic profile
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
no
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
8
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
18 to 120 months
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Taxonomic profile
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
yes
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
9
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
18 to 120 months
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functional Profile (ECs)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
no
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
10
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
18 to 120 months
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functional Profile (ECs)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
yes
\end{minipage} \\
\midrule()
\endhead
\bottomrule()
\end{longtable}

Random Forests (RFs) \href{https://paperpile.com/c/dPbU4e/VDqU}{(Breiman
1996)} were selected as the prediction engine and processed using the
DecisionTree.jl
\href{https://www.zotero.org/google-docs/?dQLyDs}{(Sadeghi et al.,
2022)} implementation, inside the MLJ.jl
\href{https://www.zotero.org/google-docs/?y8ywGI}{(Blaom et al., 2020)}
Machine Learning framework. Independent RFs were trained for each
experiment, using a set of default regression hyperparameters from
Breiman and Cutler
\href{https://www.zotero.org/google-docs/?d3gqFh}{(Breiman, 2001)}, on a
repeated cross-validation approach with different RNG seeds. One hundred
repetitions of 3-fold CV with 10 different intra-fold RNG states each
were employed, for a total of 3000 experiments per input set.

After the training procedures, the root-mean-square error (RMSE) for
cognitive assessment scores and mean absolute proportional error (MAPE)
for the brain segmentation data, along with Pearson\textquotesingle s
correlation coefficient (R) were benchmarked on the validation and train
sets. MAPE was chosen as the metric for brain segments due to magnitude
differences between median volumes of each segment, which would hinder
interpretation of raw error values without additional reference.

To derive biological insight from the models, the covariate variable
importances for all the input features, measured by mean decrease in
impurity (MDI, or GINI importance), was also analyzed. Leveraging the
distribution of results from the extensive repeated cross validation
experiments, rather than electing a representative model or picking the
highest validation-set correlation, a measure of model fitness
(\textbf{Equation 1}) was designed to weight the importances from each
trained forest. The objective was to give more weight to those with
higher benchmarks on the validation sets (or higher generalizability),
while penalizing information from highly overfit models, drawing
inspiration from the approach used on another work employing repeated CV
on Random Forests with high-dimensional, low sample size microbiome
datasets \{***\}. The resulting fitness-weighted importances were used
to generate the values in Figure 3.

\(fitness =\)

\textbf{Equation 1.} Mathematical expression of the fitness measure used
to weight feature importances based on model benchmarks

\hypertarget{mri-segmentation}{%
\subsubsection{MRI / segmentation}\label{mri-segmentation}}

MRI data was acquired at a 3T Siemens Trio scanner with the following
parameters: TE=5.6msec, TR=1400msec, FA=15 degrees, 1.1x1.1x1.1mm
resolution, 160x160 matrix, with an average of 112 slices. FOV was
adjusted to infant size. Using a combination of linear and nonlinear
image registration, we created representative age-specific templates
using the ANTs package
\href{https://www.zotero.org/google-docs/?ZlWDfL}{(Avants et al.,
2014)}. After age specific templates were created, a single flow from
each age to the 12-month template was estimated and a final warp from
the 12-month template to standard adult MNI space was performed. For
each individual infant/ child brain image, we then calculated the warp
from their native T1w image space to their nearest-in-age template.
Using the resulting warps (native → nearest age template → 12-month
template → MNI template), we could move the standard adult brain atlas
to the space of an individual infant in a single step. In this case, we
used the Harvard-Oxford brain atlas to provide a coarse-grained
parcellation of individual brains into subcortical regions (e.g.,
thalamus, putamen) and total grey and white matter volumes (included as
part of the FSL package
\href{https://www.zotero.org/google-docs/?AapPN7}{(Jenkinson et al.,
2012)}. Total tissue and brainstem volume as well as left and right
hemisphere volumes were derived for total white and cortical gray
matter, lateral ventricle, thalamus, caudate, putamen, pallidum,
hippocampus, amygdala, and accumbens as well as total brainstem volume
\href{https://www.zotero.org/google-docs/?MCDDB9}{(Bruchhage et al.,
n.d.)}.

\hypertarget{computational-analysis-of-metagenomes}{%
\subsubsection{Computational analysis of
metagenomes}\label{computational-analysis-of-metagenomes}}

Shotgun metagenomic sequences were analyzed using the bioBakery suite of
computational tools \href{https://paperpile.com/c/dPbU4e/vDau}{(Beghini
et al. 2021)}. First, KneadData (v0.7.7) was used to perform quality
control of raw sequence reads, such as read trimming and removal of
reads matching a human genome reference. Next, MetaPhlAn (v3.0.7, using
database mpa\_v30\_CHOCOPhlAn\_201901) was used to generate taxonomic
profiles by aligning reads to a reference database of marker genes.
Finally, HUMAnN (v3.0.0a4) was used to functionally profile the
metagenomes.

\hypertarget{data-and-code-availability}{%
\subsubsection{Data and code
availability}\label{data-and-code-availability}}

Taxonomic and functional microbial profiles, as well as subject
demographics necessary for statistical analyses and machine learning are
available on the Open Science Framework
\href{https://www.zotero.org/google-docs/?61HLlM}{(Bonham et al.,
2022)}. Data processing, generation of summary statistics, and
generation of plots was performed using the julia programming language
\href{https://www.zotero.org/google-docs/?zj3HnY}{(Bezanson et al.,
2017; Bonham et al., 2021; Danisch and Krumbiegel, 2021)}. All code for
data analysis and figure generation, as well as scripts for automated
download of input files are available on github \{\{CITE zenodo\}\}.

\hypertarget{research-standards}{%
\subsection{Research Standards}\label{research-standards}}

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

Research funding: NIH UG3 OD023313 and Wellcome LEAP 1kD.

\hypertarget{references-and-citations}{%
\subsection{References and Citations}\label{references-and-citations}}

\url{https://doi.org/10.1093/cercor/bhs201}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Amaratunga, D.,
Cabrera, J., Lee, Y.-S., 2008. Enriched random forests. Bioinformatics
24, 2010--2014. https://doi.org/10.1093/bioinformatics/btn356}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Aminoff, E.M.,
Kveraga, K., Bar, M., 2013. The role of the parahippocampal cortex in
cognition. Trends Cogn. Sci. 17, 379--390.
https://doi.org/10.1016/j.tics.2013.06.009}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Apps, M.A.J.,
Rushworth, M.F.S., Chang, S.W.C., 2016. The Anterior Cingulate Gyrus and
Social Cognition: Tracking the Motivation of Others. Neuron 90,
692--707. https://doi.org/10.1016/j.neuron.2016.04.018}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Ardekani, B.A.,
Bermudez, E., Mubeen, A.M., Bachman, A.H., Initiative, for the A.D.N.,
2017. Prediction of Incipient Alzheimer's Disease Dementia in Patients
with Mild Cognitive Impairment. J. Alzheimers Dis. 55, 269--281.
https://doi.org/10.3233/JAD-160594}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Avants, B.B.,
Tustison, N.J., Stauffer, M., Song, G., Wu, B., Gee, J.C., 2014. The
insight ToolKit image registration framework. Front. Neuroinform. 8, 44.
https://doi.org/10.3389/fninf.2014.00044}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Bäckhed, F., Roswall,
J., Peng, Y., Feng, Q., Jia, H., Kovatcheva-Datchary, P., Li, Y., Xia,
Y., Xie, H., Zhong, H., Khan, M.T., Zhang, J., Li, J., Xiao, L.,
Al-Aama, J., Zhang, D., Lee, Y.S., Kotowska, D., Colding, C., Tremaroli,
V., Yin, Y., Bergman, S., Xu, X., Madsen, L., Kristiansen, K., Dahlgren,
J., Wang, J., 2015. Dynamics and stabilization of the human gut
microbiome during the first year of life. Cell Host Microbe 17,
690--703. https://doi.org/10.1016/j.chom.2015.04.004}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Bairoch, A., 2000. The
ENZYME database in 2000. Nucleic Acids Res. 28, 304--305.
https://doi.org/10.1093/nar/28.1.304}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Benítez-Páez, A.,
Gómez del Pugar, E.M., López-Almela, I., Moya-Pérez, Á., Codoñer-Franch,
P., Sanz, Y., 2020. Depletion of \emph{Blautia} Species in the
Microbiota of Obese Children Relates to Intestinal Inflammation and
Metabolic Phenotype Worsening. mSystems 5, e00857-19.
https://doi.org/10.1128/mSystems.00857-19}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Bezanson, J., Edelman,
A., Karpinski, S., Shah, V.B., 2017. Julia: A fresh approach to
numerical computing. Siam Rev. 59, 65--98.
https://doi.org/10.1137/141000671}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Blackwell, C.K.,
Mansolf, M., Sherlock, P., Ganiban, J., Hofheimer, J.A., Barone, C.J.,
Bekelman, T.A., Blair, C., Cella, D., Collazo, S., Croen, L.A., Deoni,
S., Elliott, A.J., Ferrara, A., Fry, R.C., Gershon, R., Herbstman, J.B.,
Karagas, M.R., LeWinn, K.Z., Margolis, A., Miller, R.L., O'Shea, T.M.,
Porucznik, C.A., Wright, R.J., 2022. Youth well-being during the
COVID-19 pandemic. Pediatrics 149.
https://doi.org/10.1542/peds.2021-054754}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Blaom, A.D., Kiraly,
F., Lienart, T., Simillides, Y., Arenas, D., Vollmer, S.J., 2020. MLJ: A
Julia package for composable machine learning. J. Open Source Softw. 5,
2704. https://doi.org/10.21105/joss.02704}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Boes, A.D., Tranel,
D., Anderson, S.W., Nopoulos, P., 2008. Right anterior cingulate cortex
volume is a neuroanatomical correlate of aggression and defiance in
boys. Behav. Neurosci. 122, 677--684.
https://doi.org/10.1037/0735-7044.122.3.677}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Bokulich, N.A., Chung,
J., Battaglia, T., Henderson, N., Jay, M., Li, H., D Lieber, A., Wu, F.,
Perez-Perez, G.I., Chen, Y., Schweizer, W., Zheng, X., Contreras, M.,
Dominguez-Bello, M.G., Blaser, M.J., 2016. Antibiotics, birth mode, and
diet shape microbiome maturation during early life. Sci. Transl. Med. 8,
343ra82. https://doi.org/10.1126/scitranslmed.aad7121}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Bottino,
G.Z.M.F., Klepac-Ceraj, V., 2022. ECHO RESONANCE Microbiome.
https://doi.org/10.17605/OSF.IO/YBS32}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Kayisire,
A., Luo, A., Klepac-Ceraj, V., 2021. Microbiome.jl and BiobakeryUtils.jl
- Julia packages for working with microbial community data. J Open
Source Softw 6, 3876. https://doi.org/10.21105/joss.03876}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Breiman, L., 2001.
Random Forests. Mach. Learn. 45, 5--32.
https://doi.org/10.1023/A:1010933404324}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Brieuc, M.S.O.,
Waters, C.D., Drinan, D.P., Naish, K.A., 2018. A practical introduction
to Random Forest for genetic association studies in ecology and
evolution. Mol. Ecol. Resour. 18, 755--766.
https://doi.org/10.1111/1755-0998.12773}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Bruchhage, Muriel,
Chen, Y., Cataldo, A.G., Müller, H.-G., Weisee, E., Wilson, S., Pietsch,
M., D'Sa, V., Marquand, A., Madhow, S., Bouchard, K., Cole, J.H.,
Biondo, F., Elison, J., O'Muirchheartaigh, J., Deoni, C, S., n.d.
Longitudinal brain and cognitive development of the first 1000 Days:A
large Multi-Cohort Multi-Scanner study. Presented at the International
Society of Magnetic Resonance Medicine.}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Bush, G., Vogt, B.A.,
Holmes, J., Dale, A.M., Greve, D., Jenike, M.A., Rosen, B.R., 2002.
Dorsal anterior cingulate cortex: A role in reward-based decision
making. Proc. Natl. Acad. Sci. 99, 523--528.
https://doi.org/10.1073/pnas.012470999}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Callaghan, B., 2020.
Nested sensitive periods: how plasticity across the microbiota-gut-brain
axis interacts to affect the development of learning and memory. Curr.
Opin. Behav. Sci. 36, 55--62.
https://doi.org/10.1016/j.cobeha.2020.07.011}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Cao, Z.,
Ottino-Gonzalez, J., Cupertino, R.B., Schwab, N., Hoke, C., Catherine,
O., Cousijn, J., Dagher, A., Foxe, J.J., Goudriaan, A.E., Hester, R.,
Hutchison, K., Li, C.-S.R., London, E.D., Lorenzetti, V., Luijten, M.,
Martin-Santos, R., Momenan, R., Paulus, M.P., Schmaal, L., Sinha, R.,
Sjoerds, Z., Solowij, N., Stein, D.J., Stein, E.A., Uhlmann, A., van
Holst, R.J., Veltman, D.J., Wiers, R.W., Yücel, M., Zhang, S.,
Jahanshad, N., Thompson, P.M., Conrod, P., Mackey, S., Garavan, H.,
2021. Mapping cortical and subcortical asymmetries in substance
dependence: Findings from the ENIGMA Addiction Working Group. Addict.
Biol. 26, e13010. https://doi.org/10.1111/adb.13010}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Cerdó, T., Diéguez,
E., Campoy, C., 2019. Early nutrition and gut microbiome:
interrelationship between bacterial metabolism, immune system, brain
structure, and neurodevelopment. Am. J. Physiol. Endocrinol. Metab. 317,
E617--E630. https://doi.org/10.1152/ajpendo.00188.2019}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Chen, X., Ishwaran,
H., 2012. Random forests for genomic data analysis. Genomics 99,
323--329. https://doi.org/10.1016/j.ygeno.2012.04.003}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Cohen Kadosh, K.,
Krause, B., King, A.J., Near, J., Cohen Kadosh, R., 2015. Linking GABA
and glutamate levels to cognitive skill acquisition during development.
Hum. Brain Mapp. 36, 4334--4345. https://doi.org/10.1002/hbm.22921}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Cowan, C.S.M., Dinan,
T.G., Cryan, J.F., 2020. Annual Research Review: Critical windows - the
microbiota-gut-brain axis in neurocognitive development. J. Child
Psychol. Psychiatry 61, 353--371. https://doi.org/10.1111/jcpp.13156}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Dalile, B., Van
Oudenhove, L., Vervliet, B., Verbeke, K., 2019. The role of short-chain
fatty acids in microbiota-gut-brain communication. Nat. Rev.
Gastroenterol. Hepatol. 16, 461--478.
https://doi.org/10.1038/s41575-019-0157-3}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Danisch, S.,
Krumbiegel, J., 2021. Makie.jl: Flexible high-performance data
visualization for Julia. J. Open Source Softw. 6, 3349.
https://doi.org/10.21105/joss.03349}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Deoni, S.C.,
Beauchemin, J., Volpe, A., Dâ Sa, V., RESONANCE Consortium, 2021. Impact
of the COVID-19 pandemic on early child cognitive development: Initial
findings in a longitudinal observational study of child health. MedRxiv
Prepr. Serv. Health Sci. https://doi.org/10.1101/2021.08.10.21261846}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Dominguez-Bello, M.G.,
Costello, E.K., Contreras, M., Magris, M., Hidalgo, G., Fierer, N.,
Knight, R., 2010. Delivery mode shapes the acquisition and structure of
the initial microbiota across multiple body habitats in newborns. Proc.
Natl. Acad. Sci. U. S. A. 107, 11971--11975.
https://doi.org/10.1073/pnas.1002601107}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Dong, H., Wang, B.,
Feng, J., Yue, X., Jia, F., 2021. Correlation between serum
concentrations of menaquinone-4 and developmental quotients in children
with autism spectrum disorder. Front. Nutr. 8, 748513.
https://doi.org/10.3389/fnut.2021.748513}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Ernst, M., Nelson,
E.E., Jazbec, S., McClure, E.B., Monk, C.S., Leibenluft, E., Blair, J.,
Pine, D.S., 2005. Amygdala and nucleus accumbens in responses to receipt
and omission of gains in adults and adolescents. NeuroImage 25,
1279--1291. https://doi.org/10.1016/j.neuroimage.2004.12.038}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Forrest, C.B.,
Blackwell, C.K., Camargo, C.A., 2018. Advancing the science of
children's positive health in the national institutes of health
environmental influences on child health outcomes (ECHO) research
program. J. Pediatr. 196, 298--300.
https://doi.org/10.1016/j.jpeds.2018.02.004}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Franzosa, E.A.,
Sirota-Madi, A., Avila-Pacheco, J., Fornelos, N., Haiser, H.J., Reinker,
S., Vatanen, T., Hall, A.B., Mallick, H., McIver, L.J., Sauk, J.S.,
Wilson, R.G., Stevens, B.W., Scott, J.M., Pierce, K., Deik, A.A.,
Bullock, K., Imhann, F., Porter, J.A., Zhernakova, A., Fu, J., Weersma,
R.K., Wijmenga, C., Clish, C.B., Vlamakis, H., Huttenhower, C., Xavier,
R.J., 2019. Gut microbiome structure and metabolic activity in
inflammatory bowel disease. Nat. Microbiol. 4, 293--305.
https://doi.org/10.1038/s41564-018-0306-4}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Fung, T.C., Olson,
C.A., Hsiao, E.Y., 2017. Interactions between the microbiota, immune and
nervous systems in health and disease. Nat. Neurosci. 20, 145--155.
https://doi.org/10.1038/nn.4476}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Ghosh, T.S., Rampelli,
S., Jeffery, I.B., Santoro, A., Neto, M., Capri, M., Giampieri, E.,
Jennings, A., Candela, M., Turroni, S., Zoetendal, E.G., Hermes, G.D.A.,
Elodie, C., Meunier, N., Brugere, C.M., Pujos-Guillot, E., Berendsen,
A.M., Groot, L.C.P.G.M.D., Feskins, E.J.M., Kaluza, J., Pietruszka, B.,
Bielak, M.J., Comte, B., Maijo-Ferre, M., Nicoletti, C., Vos, W.M.D.,
Fairweather-Tait, S., Cassidy, A., Brigidi, P., Franceschi, C., O'Toole,
P.W., 2020. Mediterranean diet intervention alters the gut microbiome in
older people reducing frailty and improving health status: the NU-AGE
1-year dietary intervention across five European countries. Gut 69,
1218--1228. https://doi.org/10.1136/gutjnl-2019-319654}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Gillman, M.W.,
Blaisdell, C.J., 2018. Environmental influences on child health
outcomes, a research program of the NIH. Curr. Opin. Pediatr. 30, 260.}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Gong, Q.-Y., Cai, L.,
Jing, Y., Wang, W., Yang, D.-X., Chen, S.-W., Tian, H.-L., 2022.
Urolithin A alleviates blood-brain barrier disruption and attenuates
neuronal apoptosis following traumatic brain injury in mice. Neural
Regen. Res. 17, 2007. https://doi.org/10.4103/1673-5374.335163}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Haam, J., Yakel, J.L.,
2017. Cholinergic modulation of the hippocampal region and memory
function. J. Neurochem. 142, 111--121.
https://doi.org/10.1111/jnc.14052}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Haiser, H.J.,
Gootenberg, D.B., Chatman, K., Sirasani, G., Balskus, E.P., Turnbaugh,
P.J., 2013. Predicting and Manipulating Cardiac Drug Inactivation by the
Human Gut Bacterium Eggerthella lenta. Science 341, 295--298.
https://doi.org/10.1126/science.1235872}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Hartstra, A.V.,
Schüppel, V., Imangaliyev, S., Schrantee, A., Prodan, A., Collard, D.,
Levin, E., Dallinga-Thie, G., Ackermans, M.T., Winkelmeijer, M., Havik,
S.R., Metwaly, A., Lagkouvardos, I., Nier, A., Bergheim, I.,
Heikenwalder, M., Dunkel, A., Nederveen, A.J., Liebisch, G., Mancano,
G., Claus, S.P., Benítez-Páez, A., la Fleur, S.E., Bergman, J.J.,
Gerdes, V., Sanz, Y., Booij, J., Kemper, E., Groen, A.K., Serlie, M.J.,
Haller, D., Nieuwdorp, M., 2020. Infusion of donor feces affects the
gut--brain axis in humans with metabolic syndrome. Mol. Metab. 42,
101076. https://doi.org/10.1016/j.molmet.2020.101076}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoogman, M., Bralten,
J., Hibar, D.P., Mennes, M., Zwiers, M.P., Schweren, L., van Hulzen,
K.J.E., Medland, S.E., Shumskaya, E., Jahanshad, N., de Zeeuw, P.,
Szekely, E., Sudre, G., Wolfers, T., Onnink, A.M.H., Dammers, J.T.,
Mostert, J.C., Vives-Gilabert, Y., Kohls, G., Oberwelland, E., Seitz,
J., Schulte-Rüther, M., di Bruttopilo, S.A., Doyle, A.E., Høvik, M.F.,
Dramsdahl, M., Tamm, L., van Erp, T.G.M., Dale, A., Schork, A.,
Conzelmann, A., Zierhut, K., Baur, R., McCarthy, H., Yoncheva, Y.N.,
Cubillo, A., Chantiluke, K., Mehta, M.A., Paloyelis, Y., Hohmann, S.,
Baumeister, S., Bramati, I., Mattos, P., Tovar-Moll, F., Douglas, P.,
Banaschewski, T., Brandeis, D., Kuntsi, J., Asherson, P., Rubia, K.,
Kelly, C., Di Martino, A., Milham, M.P., Castellanos, F.X., Frodl, T.,
Zentis, M., Lesch, K.-P., Reif, A., Pauli, P., Jernigan, T., Haavik, J.,
Plessen, K.J., Lundervold, A.J., Hugdahl, K., Seidman, L.J., Biederman,
J., Rommelse, N., Heslenfeld, D.J., Hartman, C., Hoekstra, P.J.,
Oosterlaan, J., von Polier, G., Konrad, K., Vilarroya, O.,
Ramos-Quiroga, J.-A., Soliva, J.C., Durston, S., Buitelaar, J.K.,
Faraone, S.V., Shaw, P., Thompson, P., Franke, B., 2017. Subcortical
brain volume differences of participants with ADHD across the lifespan:
an ENIGMA collaboration. Lancet Psychiatry 4, 310--319.
https://doi.org/10.1016/S2215-0366(17)30049-4}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoshino, Y., Yamamoto,
T., Kaneko, M., Tachibana, R., Watanabe, M., Ono, Y., Kumashiro, H.,
1984. Blood Serotonin and Free Tryptophan Concentration in Autistic
Children. Neuropsychobiology 11, 22--27.
https://doi.org/10.1159/000118045}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Hosomi, K., Saito, M.,
Park, J., Murakami, H., Shibata, N., Ando, M., Nagatake, T., Konishi,
K., Ohno, H., Tanisawa, K., Mohsen, A., Chen, Y.-A., Kawashima, H.,
Natsume-Kitatani, Y., Oka, Y., Shimizu, H., Furuta, M., Tojima, Y.,
Sawane, K., Saika, A., Kondo, S., Yonejima, Y., Takeyama, H., Matsutani,
A., Mizuguchi, K., Miyachi, M., Kunisawa, J., 2022. Oral administration
of Blautia wexlerae ameliorates obesity and type 2 diabetes via
metabolic remodeling of the gut microbiota. Nat. Commun. 13, 4477.
https://doi.org/10.1038/s41467-022-32015-7}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Jenkinson, M.,
Beckmann, C.F., Behrens, T.E.J., Woolrich, M.W., Smith, S.M., 2012. FSL.
Neuroimage 62, 782--790.
https://doi.org/10.1016/j.neuroimage.2011.09.015}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Jiang, H., Ling, Z.,
Zhang, Y., Mao, H., Ma, Z., Yin, Y., Wang, W., Tang, W., Tan, Z., Shi,
J., Li, L., Ruan, B., 2015. Altered fecal microbiota composition in
patients with major depressive disorder. Brain. Behav. Immun. 48,
186--194. https://doi.org/10.1016/j.bbi.2015.03.016}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Kanehisa, M., Goto,
S., Kawashima, S., Okuno, Y., Hattori, M., 2004. The KEGG resource for
deciphering the genome. Nucleic Acids Res. 32, D277--D280.
https://doi.org/10.1093/nar/gkh063}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Kant, R., Rasinkangas,
P., Satokari, R., Pietilä, T.E., Palva, A., 2015. Genome Sequence of the
Butyrate-Producing Anaerobic Bacterium Anaerostipes hadrus PEL 85.
Genome Announc. 3, e00224-15. https://doi.org/10.1128/genomeA.00224-15}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Kim, C.-S., Cha, L.,
Sim, M., Jung, S., Chun, W.Y., Baik, H.W., Shin, D.-M., 2021. Probiotic
supplementation improves cognitive function and mood with changes in gut
microbiota in Community-Dwelling older adults: A randomized,
Double-Blind, Placebo-Controlled, multicenter trial. J Gerontol Biol Sci
Med Sci 76, 32--40. https://doi.org/10.1093/gerona/glaa090}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Knudsen, E.I., 2004.
Sensitive periods in the development of the brain and behavior. J. Cogn.
Neurosci. 16, 1412--1425. https://doi.org/10.1162/0898929042304796}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Koenig, J.E., Spor,
A., Scalfone, N., Fricker, A.D., Stombaugh, J., Knight, R., Angenent,
L.T., Ley, R.E., 2011. Succession of microbial consortia in the
developing infant gut microbiome. Proc. Natl. Acad. Sci. U. S. A. 108
Suppl 1, 4578--4585. https://doi.org/10.1073/pnas.1000081107}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Coker,
M.O., Madan, J.C., 2022. The developing microbiome from birth to 3
years: The Gut-Brain axis and neurodevelopmental outcomes. Front.
Pediatr. 10, 815885. https://doi.org/10.3389/fped.2022.815885}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Korrick,
S.A., Baker, E.R., Karagas, M.R., Madan, J.C., 2020. Prospective
associations of the infant gut microbiome and microbial function with
social behaviors related to autism at age 3 years. Sci. Rep. 10, 15515.
https://doi.org/10.1038/s41598-020-72386-9}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Ley, R.E., Turnbaugh,
P.J., Klein, S., Gordon, J.I., 2006. Human gut microbes associated with
obesity. Nature 444, 1022--1023. https://doi.org/10.1038/4441022a}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Li, Y.-M., Yu, Q.,
Huang, L., Fu, J., Feng, R.-J., 2022. {[}Correlations Between Gut
Microbiota Changes and Cognitive Function in Patients with Post-Stroke
Cognitive Impairment in the Early Stage{]}. Sichuan Da Xue Xue Bao Yi
Xue Ban 53, 857--865. https://doi.org/10.12182/20220960105}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, S., Li, E., Sun,
Z., Fu, D., Duan, G., Jiang, M., Yu, Y., Mei, L., Yang, P., Tang, Y.,
Zheng, P., 2019. Altered gut microbiota and short chain fatty acids in
Chinese children with autism spectrum disorder. Sci. Rep. 9, 287.
https://doi.org/10.1038/s41598-018-36430-z}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, X., Mao, B., Gu,
J., Wu, J., Cui, S., Wang, G., Zhao, J., Zhang, H., Chen, W., 2021.
\emph{Blautia} ---a new functional genus with potential probiotic
properties? Gut Microbes 13, 1875796.
https://doi.org/10.1080/19490976.2021.1875796}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Lopez-Siles, M.,
Duncan, S.H., Garcia-Gil, L.J., Martinez-Medina, M., 2017.
Faecalibacterium prausnitzii: from microbiology to diagnostics and
prognostics. ISME J. 11, 841--852.
https://doi.org/10.1038/ismej.2016.176}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Louwies, T., Johnson,
A.C., Orock, A., Yuan, T., Greenwood-Van Meerveld, B., 2020. The
microbiota-gut-brain axis: An emerging role for the epigenome. Exp.
Biol. Med. 245, 138--145. https://doi.org/10.1177/1535370219891690}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Magnusson, K.R.,
Hauck, L., Jeffrey, B.M., Elias, V., Humphrey, A., Nath, R., Perrone,
A., Bermudez, L.E., 2015. Relationships between diet-related changes in
the gut microbiome and cognitive flexibility. Neuroscience 300,
128--140. https://doi.org/10.1016/j.neuroscience.2015.05.016}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Mallick, H.,
Rahnavard, A., McIver, L.J., Ma, S., Zhang, Y., Nguyen, L.H., Tickle,
T.L., Weingart, G., Ren, B., Schwager, E.H., Chatterjee, S., Thompson,
K.N., Wilkinson, J.E., Subramanian, A., Lu, Y., Waldron, L., Paulson,
J.N., Franzosa, E.A., Bravo, H.C., Huttenhower, C., 2021. Multivariable
association discovery in population-scale meta-omics studies.
https://doi.org/10.1101/2021.01.20.427420}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Maruo, T., Sakamoto,
M., Ito, C., Toda, T., Benno, Y., 2008. Adlercreutzia equolifaciens gen.
nov., sp. nov., an equol-producing bacterium isolated from human faeces,
and emended description of the genus Eggerthella. Int. J. Syst. Evol.
Microbiol. 58, 1221--1227. https://doi.org/10.1099/ijs.0.65404-0}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Mayneris-Perxachs, J.,
Castells-Nobau, A., Arnoriaga-Rodríguez, M., Martin, M., de la
Vega-Correa, L., Zapata, C., Burokas, A., Blasco, G., Coll, C.,
Escrichs, A., Biarnés, C., Moreno-Navarrete, J.M., Puig, J., Garre-Olmo,
J., Ramos, R., Pedraza, S., Brugada, R., Vilanova, J.C., Serena, J.,
Gich, J., Ramió-Torrentà, L., Pérez-Brocal, V., Moya, A., Pamplona, R.,
Sol, J., Jové, M., Ricart, W., Portero-Otin, M., Deco, G., Maldonado,
R., Fernández-Real, J.M., 2022. Microbiota alterations in proline
metabolism impact depression. Cell Metab. 34, 681-701.e10.
https://doi.org/10.1016/j.cmet.2022.04.001}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Mukherjee, A., Lordan,
C., Ross, R.P., Cotter, P.D., 2020. Gut microbes from the
phylogenetically diverse genus Eubacterium and their various
contributions to gut health. Gut Microbes 12, 1802866.
https://doi.org/10.1080/19490976.2020.1802866}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Mullen, E.M., others,
1995. Mullen scales of early learning. AGS Circle Pines, MN.}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Needham, B.D.,
Funabashi, M., Adame, M.D., Wang, Z., Boktor, J.C., Haney, J., Wu,
W.-L., Rabut, C., Ladinsky, M.S., Hwang, S.-J., Guo, Y., Zhu, Q.,
Griffiths, J.A., Knight, R., Bjorkman, P.J., Shapiro, M.G., Geschwind,
D.H., Holschneider, D.P., Fischbach, M.A., Mazmanian, S.K., 2022. A
gut-derived metabolite alters brain activity and anxiety behaviour in
mice. Nature 602, 647--653. https://doi.org/10.1038/s41586-022-04396-8}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Opel, N., Thalamuthu,
A., Milaneschi, Y., Grotegerd, D., Flint, C., Leenings, R., Goltermann,
J., Richter, M., Hahn, T., Woditsch, G., Berger, K., Hermesdorf, M.,
McIntosh, A., Whalley, H.C., Harris, M.A., MacMaster, F.P., Walter, H.,
Veer, I.M., Frodl, T., Carballedo, A., Krug, A., Nenadic, I., Kircher,
T., Aleman, A., Groenewold, N.A., Stein, D.J., Soares, J.C.,
Zunta-Soares, G.B., Mwangi, B., Wu, M.-J., Walter, M., Li, M., Harrison,
B.J., Davey, C.G., Cullen, K.R., Klimes-Dougan, B., Mueller, B.A.,
Sämann, P.G., Penninx, B., Nawijn, L., Veltman, D.J., Aftanas, L., Brak,
I.V., Filimonova, E.A., Osipov, E.A., Reneman, L., Schrantee, A., Grabe,
H.J., Van der Auwera, S., Wittfeld, K., Hosten, N., Völzke, H., Sim, K.,
Gotlib, I.H., Sacchet, M.D., Lagopoulos, J., Hatton, S.N., Hickie, I.,
Pozzi, E., Thompson, P.M., Jahanshad, N., Schmaal, L., Baune, B.T.,
Dannlowski, U., 2021. Brain structural abnormalities in obesity:
relation to age, genetic risk, and common psychiatric disorders. Mol.
Psychiatry 26, 4839--4852. https://doi.org/10.1038/s41380-020-0774-9}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Pagonabarraga, J.,
Corcuera-Solano, I., Vives-Gilabert, Y., Llebaria, G., García-Sánchez,
C., Pascual-Sedano, B., Delfino, M., Kulisevsky, J., Gómez-Ansón, B.,
2013. Pattern of Regional Cortical Thinning Associated with Cognitive
Deterioration in Parkinson's Disease. PLoS ONE 8, e54980.
https://doi.org/10.1371/journal.pone.0054980}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Palomo-Buitrago, M.E.,
Sabater-Masdeu, M., Moreno-Navarrete, J.M., Caballano-Infantes, E.,
Arnoriaga-Rodríguez, M., Coll, C., Ramió, L., Palomino-Schätzlein, M.,
Gutiérrez-Carcedo, P., Pérez-Brocal, V., Simó, R., Moya, A., Ricart, W.,
Herance, J.R., Fernández-Real, J.M., 2019. Glutamate interactions with
obesity, insulin resistance, cognition and gut microbiota composition.
Acta Diabetol. 56, 569--579. https://doi.org/10.1007/s00592-019-01313-w}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Pronovost, G.N.,
Hsiao, E.Y., 2019. Perinatal interactions between the microbiome,
immunity, and neurodevelopment. Immunity 50, 18--36.
https://doi.org/10.1016/j.immuni.2018.11.016}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Sadeghi, B.,
Chiarawongse, P., Squire, K., Jones, D.C., Noack, A., St-Jean, C.,
Huijzer, R., Schätzle, R., Butterworth, I., Peng, Y.-F., Blaom, A.,
2022. DecisionTree.jl - A julia implementation of the CART decision tree
and random forest algorithms. https://doi.org/10.5281/zenodo.7359268}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Sela, D.A., Chapman,
J., Adeuya, A., Kim, J.H., Chen, F., Whitehead, T.R., Lapidus, A.,
Rokhsar, D.S., Lebrilla, C.B., German, J.B., Price, N.P., Richardson,
P.M., Mills, D.A., 2008. The genome sequence of \emph{Bifidobacterium
longum} subsp. \emph{infantis} reveals adaptations for milk utilization
within the infant microbiome. Proc. Natl. Acad. Sci. 105, 18964--18969.
https://doi.org/10.1073/pnas.0809584105}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Selma, M.V., Beltrán,
D., García-Villalba, R., Espín, J.C., Tomás-Barberán, F.A., 2014.
Description of urolithin production capacity from ellagic acid of two
human intestinal Gordonibacter species. Food Funct. 5, 1779--1784.
https://doi.org/10.1039/c4fo00092g}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Setchell, K.D.,
Clerici, C., Lephart, E.D., Cole, S.J., Heenan, C., Castellani, D.,
Wolfe, B.E., Nechemias-Zimmer, L., Brown, N.M., Lund, T.D., Handa, R.J.,
Heubi, J.E., 2005. S-Equol, a potent ligand for estrogen receptor β, is
the exclusive enantiomeric form of the soy isoflavone metabolite
produced by human intestinal bacterial flora. Am. J. Clin. Nutr. 81,
1072--1079. https://doi.org/10.1093/ajcn/81.5.1072}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Sharon, G., Sampson,
T.R., Geschwind, D.H., Mazmanian, S.K., 2016. The central nervous system
and the gut microbiome. Cell 167, 915--932.
https://doi.org/10.1016/j.cell.2016.10.027}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Silbereis, J.C.,
Pochareddy, S., Zhu, Y., Li, M., Sestan, N., 2016. The cellular and
molecular landscapes of the developing human central nervous system.
Neuron 89, 248--268.}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Sordillo, J.E.,
Korrick, S., Laranjo, N., Carey, V., Weinstock, G.M., Gold, D.R.,
O'Connor, G., Sandel, M., Bacharier, L.B., Beigelman, A., Zeiger, R.,
Litonjua, A.A., Weiss, S.T., 2019. Association of the infant gut
microbiome with early childhood neurodevelopmental outcomes: An
ancillary study to the VDAART randomized clinical trial. JAMA Netw. Open
2, e190905. https://doi.org/10.1001/jamanetworkopen.2019.0905}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Spichak, S.,
Bastiaanssen, T.F.S., Berding, K., Vlckova, K., Clarke, G., Dinan, T.G.,
Cryan, J.F., 2021. Mining microbes for mental health: Determining the
role of microbial metabolic pathways in human brain health and disease.
Neurosci. Biobehav. Rev. 125, 698--761.
https://doi.org/10.1016/j.neubiorev.2021.02.044}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Stephan, J., Stegle,
O., Beyer, A., 2015. A random forest approach to capture genetic effects
in the presence of population structure. Nat. Commun. 6, 7432.
https://doi.org/10.1038/ncomms8432}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Suzek, B.E., Huang,
H., McGarvey, P., Mazumder, R., Wu, C.H., 2007. UniRef: comprehensive
and non-redundant UniProt reference clusters. Bioinformatics 23,
1282--1288. https://doi.org/10.1093/bioinformatics/btm098}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Takahashi, H., Yang,
J., Yamamoto, H., Fukuda, S., Arakawa, K., 2021. Complete Genome
Sequence of Adlercreutzia equolifaciens subsp. celatus DSM 18785.
Microbiol. Resour. Announc. 10, e00354-21.
https://doi.org/10.1128/MRA.00354-21}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Tau, G.Z., Peterson,
B.S., 2010. Normal development of brain circuits. Neuropsychopharmacol.
Off. Publ. Am. Coll. Neuropsychopharmacol. 35, 147--168.
https://doi.org/10.1038/npp.2009.115}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Thawornkuno, C.,
Tanaka, M., Sone, T., Asano, K., 2009. Biotransformation of Daidzein to
Equol by Crude Enzyme from Asaccharobacter celatus AHU1763 Required an
Anaerobic Environment. Biosci. Biotechnol. Biochem. 73, 1435--1438.
https://doi.org/10.1271/bbb.80908}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Tognini, P., 2017. Gut
microbiota: a potential regulator of neurodevelopment. Front. Cell.
Neurosci. 11, 25. https://doi.org/10.3389/fncel.2017.00025}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Tso, L., Bonham, K.S.,
Fishbein, A., Rowland, S., Klepac-Ceraj, V., 2021. Targeted
High-Resolution taxonomic identification of bifidobacterium longum
subsp. infantis using human milk oligosaccharide metabolizing genes.
Nutrients 13, 2833. https://doi.org/10.3390/nu13082833}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Valles-Colomer, M.,
Falony, G., Darzi, Y., Tigchelaar, E.F., Wang, J., Tito, R.Y., Schiweck,
C., Kurilshikov, A., Joossens, M., Wijmenga, C., Claes, S., Van
Oudenhove, L., Zhernakova, A., Vieira-Silva, S., Raes, J., 2019. The
neuroactive potential of the human gut microbiota in quality of life and
depression. Nat. Microbiol. 4, 623--632.
https://doi.org/10.1038/s41564-018-0337-x}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Velazquez, M., Lee,
Y., Initiative, for the A.D.N., 2021. Random forest model for
feature-based Alzheimer's disease conversion prediction from early mild
cognitive impairment subjects. PLOS ONE 16, e0244773.
https://doi.org/10.1371/journal.pone.0244773}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Wacker, J., Dillon,
D.G., Pizzagalli, D.A., 2009. The role of the nucleus accumbens and
rostral anterior cingulate cortex in anhedonia: Integration of resting
EEG, fMRI, and volumetric techniques. NeuroImage 46, 327--337.
https://doi.org/10.1016/j.neuroimage.2009.01.058}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Wan, Y., Zuo, T., Xu,
Z., Zhang, F., Zhan, H., Chan, D., Leung, T.-F., Yeoh, Y.K., Chan,
F.K.L., Chan, R., Ng, S.C., 2021. Underdevelopment of the gut microbiota
and bacteria species as non-invasive markers of prediction in children
with autism spectrum disorder. Gut gutjnl-2020-324015.
https://doi.org/10.1136/gutjnl-2020-324015}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, L.-J., Yang,
C.-Y., Chou, W.-J., Lee, M.-J., Chou, M.-C., Kuo, H.-C., Yeh, Y.-M.,
Lee, S.-Y., Huang, L.-H., Li, S.-C., 2020. Gut microbiota and dietary
patterns in children with attention-deficit/hyperactivity disorder. Eur.
Child Adolesc. Psychiatry 29, 287--297.
https://doi.org/10.1007/s00787-019-01352-2}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, X.-L., Hur,
H.-G., Lee, J.H., Kim, K.T., Kim, S.-I., 2005. Enantioselective
Synthesis of S-Equol from Dihydrodaidzein by a Newly Isolated Anaerobic
Human Intestinal Bacterium. Appl. Environ. Microbiol. 71, 214--219.
https://doi.org/10.1128/AEM.71.1.214-219.2005}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 2012.
Wechsler preschool and primary scale of intelligence---fourth edition.
Psychol. Corp. San Antonio TX.}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 1949.
Wechsler intelligence scale for children.}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, L., Yan, J.,
Yang, T., Zhu, J., Li, T., Wei, H., Chen, J., 2021. Fecal Microbiome
Transplantation from Children with Autism Spectrum Disorder Modulates
Tryptophan and Serotonergic Synapse Metabolism and Induces Altered
Behaviors in Germ-Free Mice. mSystems 6, e01343-20.
https://doi.org/10.1128/mSystems.01343-20}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, X., Yang, Y.,
Zhang, Y., Zhang, X.-M., Zhao, Z.-Q., Zhang, Y.-Q., 2013. Estrogen in
the Anterior Cingulate Cortex Contributes to Pain-Related Aversion.
Cereb. Cortex 23, 2190--2203. https://doi.org/10.1093/cercor/bhs201}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Yael, D., Tahary, O.,
Gurovich, B., Belelovsky, K., Bar-Gad, I., 2019. Disinhibition of the
Nucleus Accumbens Leads to Macro-Scale Hyperactivity Consisting of
Micro-Scale Behavioral Segments Encoded by Striatal Activity. J.
Neurosci. 39, 5897--5909.
https://doi.org/10.1523/JNEUROSCI.3120-18.2019}

\href{https://www.zotero.org/google-docs/?uydJ8e}{Yau, W.-Y.W., Zubieta,
J.-K., Weiland, B.J., Samudra, P.G., Zucker, R.A., Heitzeg, M.M., 2012.
Nucleus Accumbens Response to Incentive Stimuli Anticipation in Children
of Alcoholics: Relationships with Precursive Behavioral Risk and
Lifetime Alcohol Use. J. Neurosci. 32, 2544--2551.
https://doi.org/10.1523/JNEUROSCI.1390-11.2012}

Supplementary Tables

\textbf{Supplementary Table XXA. List of most important taxa for Random
Forest models on the 0 to 6 months age bracket, microbiome +
demographics, ordered by relative weighted importance rank}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Taxon
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Rank
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Relative weighted importance
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Cumulative

relative weighted importance
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\midrule()
\endhead
\bottomrule()
\end{longtable}

\textbf{Supplementary Table XXA. List of most important taxa for Random
Forest models on the 18 to 120 months age bracket, microbiome +
demographics, ordered by relative weighted importance rank}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Taxon
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Rank
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Relative weighted importance
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Cumulative

relative weighted importance
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} \\
\midrule()
\endhead
\bottomrule()
\end{longtable}

\textbf{Supplementary Table XX. Benchmark metrics for each brain segment
prediction experiment. Confidence intervals are calculated from the
distribution of metrics from repeated CV at a confidence level of 95\%}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3933}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2440}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1846}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1782}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Brain segment
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Root-mean-square Error

(RMSE)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Mean Absolute Proportional Error (MAPE)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Correlation coefficient (R)
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Brain-stem
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0876
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.077
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.605
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
cerebellar-vermal-lobules-I-V
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0321
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.085
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.156
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
cerebellar-vermal-lobules-VI-VII
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0132
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.081
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.006
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
cerebellar-vermal-lobules-VIII-X
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0217
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.083
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.078
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
CSF
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0193
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.162
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.071
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-accumbens-area
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0020
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.094
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.288
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-accumbens-area
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0020
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.102
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.041
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-amygdala
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0037
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.063
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.288
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-amygdala
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0035
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.070
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.179
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-anterior-cingulate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0375
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.069
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.092
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-anterior-cingulate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0264
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.074
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.245
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-basal-forebrain
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0008
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.206
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.154
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-basal-forebrain
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0010
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.117
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.123
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-caudate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0257
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.099
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.029
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-caudate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0221
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.083
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.010
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-cerebellum-white-matter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0689
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.067
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.413
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-cerebellum-white-matter
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0694
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.070
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.434
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-cuneus
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0274
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.082
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.234
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-cuneus
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0264
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.075
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.065
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-entorhinal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0107
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.080
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.190
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-entorhinal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0105
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.082
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.099
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-fusiform
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0510
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.057
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.333
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-fusiform
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0334
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.053
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.248
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-hippocampus
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0128
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.054
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.075
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-hippocampus
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0150
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.059
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.094
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-insula
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0294
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.057
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.119
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-insula
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0303
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.056
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.077
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-isthmus-cingulate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0161
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.077
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.041
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-isthmus-cingulate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0166
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.085
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.074
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-lateral-occipital
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0661
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.062
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.224
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-lateral-occipital
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0716
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.054
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.221
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-lingual
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0410
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.069
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.421
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-lingual
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0440
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.072
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.434
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-middle-frontal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0875
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.052
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.000
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-middle-frontal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0636
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.039
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.083
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-orbitofrontal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0715
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.055
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.186
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-orbitofrontal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0814
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.061
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.133
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-pallidum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0075
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.064
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.297
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-pallidum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0081
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.061
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.264
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-paracentral
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0335
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.078
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.131
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-paracentral
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0364
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.084
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.023
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-parahippocampal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0098
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.058
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.154
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-parahippocampal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0092
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.060
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.134
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-parietal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0755
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.035
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.177
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-parietal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.1099
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.045
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.147
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-pars-opercularis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0197
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.057
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.146
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-pars-opercularis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0261
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.060
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.030
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-pars-orbitalis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0157
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.082
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.003
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-pars-orbitalis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0233
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.112
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.024
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-pars-triangularis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0226
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.063
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.174
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-pars-triangularis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0322
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.076
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.043
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-pericalcarine
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0178
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.111
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.200
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-pericalcarine
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0185
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.089
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.273
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-postcentral
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0495
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.052
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.168
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-postcentral
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0552
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.059
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.135
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-posterior-cingulate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0460
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.134
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.137
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-posterior-cingulate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0480
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.139
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.063
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-precentral
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0830
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.065
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.248
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-precentral
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0594
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.046
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.066
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-precuneus
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0489
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.056
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.144
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-precuneus
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0622
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.068
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.150
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-putamen
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0243
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.059
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.154
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-putamen
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0242
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.066
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.011
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-superior-frontal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0867
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.040
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.101
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-superior-frontal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.1206
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.044
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.096
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-supramarginal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0767
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.067
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.061
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-supramarginal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0734
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.068
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.001
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-temporal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.1530
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.040
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.128
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-temporal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.1503
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.039
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.068
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
left-thalamus-proper
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0347
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.053
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.170
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
right-thalamus-proper
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0304
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.050
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.162
\end{minipage} \\
\midrule()
\endhead
\bottomrule()
\end{longtable}

\end{document}

### 2023-02-17 manuscript/tex/main.tex
clean up

--- 2023-02-16 manuscript/tex/converted.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-]{article}
-\usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\usepackage{longtable,booktabs,array}
-\usepackage{multirow}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\usepackage[normalem]{ulem}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
+%%%%%%%%%%TO FORMAT A PREPRINT FOR A PCI%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\author{}
-\date{}
+\documentclass[a4paper]{article}
+
+\input{preamble.tex}
+
+\title{Gut microbes and their genes or something else}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%  SET THE LIST OF AUTHORS WITH CORRESPONDING AFFILIATIONS  use \& before last author %%%%%%%%%%%%%%%%%%%%
+\author{Kevin S. Bonham\textsuperscript{1}, \space
+Guilherme Fahur Bottino\textsuperscript{1}, 
+Firstname Lastname\textsuperscript{2}, \&
+Vanja Klepac-Ceraj\textsuperscript{1}
+%\fourthauthor \textsuperscript{i}
+%etc...
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE ABSTRACT %%%%%%%%%%%%%%%%%
 
@@
 
-\hypertarget{title}{%
-\subsection{Title}\label{title}}
-
-\begin{itemize}
-\item
-  \begin{quote}
-  Gut microbes and their genes are associated with cognition and
-  neuroanatomy in children
-  \end{quote}
-\item
-  \begin{quote}
-  Gut microbes and their genes are associated with cognition performance
-  and brain structure in children
-  \end{quote}
-\item
-  \begin{quote}
-  Extensive relationships between the gut microbiota and normal
-  neurocognitive development
-  \end{quote}
-\item
-  \begin{quote}
-  Extensive relationships between the gut microbiota and typical
-  neurocognitive development
-  \end{quote}
-\item
-  \begin{quote}
-  The gut microbial metagenome is connected to neurocognition in typical
-  child development
-  \end{quote}
-\item
-  \begin{quote}
-  The microbial potential in regulating early childhood neurocognitive
-  development
-  \end{quote}
-\end{itemize}
-
-\hypertarget{authors}{%
-\subsubsection{Authors}\label{authors}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \begin{quote}
-  Bonham 0000-0003-3200-7533
-  \end{quote}
-\item
-  \begin{quote}
-  Bottino 0000-0003-1953-1576
-  \end{quote}
-\item
-  \begin{quote}
-  McCann 0000-0002-9753-7968
-  \end{quote}
-\item
-  \begin{quote}
-  Beauchemin \{ORCID??\}
-  \end{quote}
-\item
-  \begin{quote}
-  Elizabeth Weisse \textless elizabeth.weisse2@gmail.com\textgreater{}
-  \end{quote}
-\item
-  \begin{quote}
-  \ldots{}
-  \end{quote}
-\item
-  \begin{quote}
-  Jonathan O\textquotesingle Muircheartaigh
-  \textless{}\href{mailto:jonathanom@kcl.ac.uk}{\uline{jonathanom@kcl.ac.uk}}\textgreater{}
-  \end{quote}
-\item
-  \begin{quote}
-  Huttenhower 0000-0002-1110-0096
-  \end{quote}
-\item
-  \begin{quote}
-  Muriel Bruchhage
-  \textless{}\href{mailto:mmk.bruchhage@gmail.com}{\uline{mmk.bruchhage@gmail.com}}\textgreater{}
-  0000-0001-9637-0951
-  \end{quote}
-\item
-  \begin{quote}
-  D'Sa 0000-0002-3476-8419
-  \end{quote}
-\item
-  \begin{quote}
-  Deoni 0000-0002-0633-6328
-  \end{quote}
-\item
-  \begin{quote}
-  Klepac-Ceraj 0000-0001-5387-5706
-  \end{quote}
-\end{enumerate}
-
-\hypertarget{abstract}{%
-\subsection{Abstract}\label{abstract}}
-
+\begin{abstract}
     The gastrointestinal tract, its resident microorganisms and the central
@@
     intervention.
+\end{abstract}
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%% Text of the preprint %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% use \emph{} for italics and \parencite{} to cite a reference and \ref{XX} to cite the \label{XX}
+% use \\ and a blank line to mark the end of paragraph and starting of a new one
 
-The gut and the brain are intimately linked. Signals from the brain
-reach the gut through the autonomic nervous system and the endocrine
-system, and the gut can communicate with the brain through the vagus
-nerve and through endocrine and immune (cytokine) signaling molecules
-\href{https://www.zotero.org/google-docs/?i3SRPn}{(Cerdó et al., 2019;
-Pronovost and Hsiao, 2019; Sharon et al., 2016; Tognini, 2017)}. In
-addition, the products of microbial metabolism generated in the gut can
-influence the brain, both indirectly by stimulating the enteric nervous
-and immune systems, as well as directly through molecules that enter
-circulation and cross the blood-brain barrier. Causal links between the
-gut microbiome and neural development, particularly atypical
-development, are increasingly being identified
-\href{https://www.zotero.org/google-docs/?UsrFKB}{(Spichak et al.,
-2021)}. Both human epidemiology and animal models point to the effects
-of gut microbes on the development of autism spectrum disorder
-\href{https://www.zotero.org/google-docs/?Y8l5A0}{(Laue et al., 2020;
-Wan et al., 2021)} and particular microbial taxa have been associated
-with depression
-\href{https://www.zotero.org/google-docs/?mbaFE3}{(Mayneris-Perxachs et
-al., 2022; Valles-Colomer et al., 2019)} and Alzheimer's disease
-\href{https://www.zotero.org/google-docs/?0BEGUn}{(Fung et al., 2017;
-Kim et al., 2021)}. But information about this ``microbiome-gut-brain
-axis'' in normal neurocognitive development remains lacking,
-particularly early in life.
+\section*{Introduction}
 
-The first years of life are critical developmental windows for both the
-microbiome and the brain
-\href{https://www.zotero.org/google-docs/?XTqhlm}{(Laue et al., 2022)}.
-Fetal development \emph{in utero} is believed to be mostly sterile, but
-is rapidly seeded at birth through contact with the birth canal (if
-birthed vaginally), caregivers, food sources (breastmilk or formula),
-and other environmental sources such as antibiotics
-\href{https://www.zotero.org/google-docs/?X6N0f4}{(Bäckhed et al., 2015;
-Bokulich et al., 2016; Dominguez-Bello et al., 2010; Louwies et al.,
-2020)}. The early microbiome is characterized by low microbial
-diversity, rapid succession and evolution, and domination by
-Actinobacteria, particularly the genus \emph{Bifidobacterium},
-Bacteroidetes, especially \emph{Bacteroides}, and Proteobacteria
-\href{https://www.zotero.org/google-docs/?VG43uh}{(Koenig et al.,
-2011)}. Many of these microbes have specialized metabolic capabilities
-for digesting human breast milk, such as \emph{Bifidobacterium infantis}
-and \emph{Bacteroides fragilis}
-\href{https://www.zotero.org/google-docs/?1dKzKj}{(Sela et al., 2008;
-Tso et al., 2021)}. Upon the introduction of solid foods, the gut
-microbiome undergoes another categorical transformation, with most taxa
-of the infant microbiome being replaced by taxa more reminiscent of
-adult microbiomes
-\href{https://www.zotero.org/google-docs/?V3lY3p}{(Bäckhed et al.,
-2015)}. Many prior studies have focused on either infant microbiomes or
-adult microbiomes, since performing statistical analyses across this
-transition poses particular challenges. Nevertheless, since this
-transition coincides with critical neural developmental windows and
-neural synaptogenesis, investigation across this solid-food boundary is
-incredibly important
-\href{https://www.zotero.org/google-docs/?d29bv8}{(Tau and Peterson,
-2010)}.
+\input{sections/intro}
 
-A child's brain undergoes remarkable anatomical, microstructural,
-organizational, and functional changes in the first years of life. By
-age 5, a child's brain has reached \textgreater85\% of its adult size,
-has achieved near-adult levels of myelination, and the pattern of axonal
-connections has been established
-\href{https://www.zotero.org/google-docs/?1jUZOO}{(Silbereis et al.,
-2016)}. Much of this development occurs in discrete windows or
-``sensitive-'' or ``critical periods'' (CPs)
-\href{https://www.zotero.org/google-docs/?w4Pcze}{(Knudsen, 2004)} when
-neural plasticity is particularly high, and particular modes of learning
-and skill development are preferred. The timing of these sensitive
-periods is driven in part by genetics, but can also be affected by the
-environment, including the gut microbiome
-\href{https://www.zotero.org/google-docs/?KTm9lQ}{(Cowan et al., 2020)}.
-In fact, emerging evidence suggests that the timing and duration of CPs
-may be driven in part by cues from the developing gut microbiome
-\href{https://www.zotero.org/google-docs/?mAD1ne}{(Callaghan, 2020)}. As
-such, understanding the normal spectrum of healthy microbiome
-development and how it relates to normal neurocognitive development may
-provide opportunities for identifying atypical development earlier and
-offer opportunities for intervention.
+\section*{Results}
 
-To begin to address this need, we investigated the gut microbiome and
-neurocognitive development of children from infancy through 10 years of
-age in an accelerated longitudinal study design. Gut microbial
-communities were assessed using shotgun metagenomic sequencing, enabling
-profiling at both the taxonomic and gene-functional level, and
-neurocognitive development was measured using expert-assessment of
-cognitive function and neuroimaging using magnetic resonance imaging
-(MRI). Using a combination of classical statistical analysis and machine
-learning, we showed that the development of the gut microbiome, the
-brain, and children's cognitive abilities are intimately linked, with
-both microbial taxa and gene functions able to predict cognitive
-performance and brain structure.
+\input{sections/results}
 
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
+\section*{Discussion}
 
-\hypertarget{cohort-overview-and-summary-data}{%
-\subsubsection{Cohort overview and summary
-data}\label{cohort-overview-and-summary-data}}
+\input{sections/discussion}
 
-We investigated the co-development of the brain and the microbiome in
-early childhood in a cohort of typically-developing children in their
-first years of life using a variety of orthogonal microbial and
-neurocognitive assessments, including shotgun metagenomic sequencing,
-cognitive and behavioral assessments, and neuroimaging. In all, 361
-children from RESONANCE, an accelerated longitudinal cohort of early
-development \href{https://paperpile.com/c/dPbU4e/klxU}{(Forrest,
-Blackwell, and Camargo 2018)} between the ages of 82 days and 10 years
-old (median age 2.21 years, Supplementary Figure 1) were included in
-this study (Table 1, Figure 1A). To measure cognitive function, we used
-age-appropriate assessments that can be normalized to a common, IQ-like
-scale (Figure 1B), including the Mullen scale of early learning (MSEL)
-for children from birth to 3 years of age
-\href{https://www.zotero.org/google-docs/?90hfYr}{(Mullen and others,
-1995)}, Wechsler preschool and primary scale of intelligence (WPPSI) for
-4--5-year-olds
-\href{https://www.zotero.org/google-docs/?wCfx5h}{(Wechsler, 2012)}, and
-the Wechsler intelligence scale for children (WISC) for children 6 years
-and up \href{https://www.zotero.org/google-docs/?AcNN3S}{(Wechsler,
-1949)}.
+\section*{Material and methods}
 
-As expected, the greatest differences in microbial taxa were driven by
-age (Figure 1B-C, PCoA axis 1), while older children were primarily
-stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
-high-\emph{Prevotella copri} (Supplementary Figure 2). Overall variation
-in gut microbial genes and also in brain volume profiles was similarly
-driven largely by subject age (Figure 1C-F).
-
-\includegraphics[width=6.45833in,height=3.87221in]{media/image3.png}
-
-\textbf{Figure 1: Cohort overview and summary data.}
-
-(A) Study design and overview. Stool samples, cognitive assessments and
-neuroimaging were collected from participants in an accelerated
-longitudinal design. (B) Cognitive function scores are assessed by
-different instruments, but may be normalized to a common IQ-like scale.
-(C-D) Principal coordinates analysis using Bray-Curtis dissimilarity on
-taxonomic profiles demonstrates high beta diversity, with much of the
-first axis of variation explained by increasing age and alpha diversity.
-Differences in gene function profiles (E) and neuroimaging (PCA based on
-the Euclidean distance of brain region volumes) (F) are likewise
-dominated by changes as children age. (G) Permutation analysis of
-variance (PERMANOVA) of taxonomic profiles, functional profiles
-(annotated as UniRef90s) and neuroimaging, against the metadata of
-interest. Variations in taxonomic and functional profiles explain a
-modest but significant percent of the variation in cognitive development
-in children over 18 months of age. (H) Mantel testing of different
-microbial feature matrices, shows overlapping but distinct patterns of
-variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
-used as cut-offs in some following models.
-
-\textbf{Table1: Subject demographics}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2620}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Group}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Subgroup}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{All}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Under 6 mo}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Over 18 mo}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-N subjects
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-361
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-74
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-257
-\end{minipage} \\
-\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Samples
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-1
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-262 (72.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-74 (100.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-203 (79.0\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-72 (19.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-46 (17.9\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-\textgreater2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-27 (7.5\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-8 (3.1\%)
-\end{minipage} \\
-\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Age (months)
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-Min
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18.0
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Max
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-119.3
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6.0
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-119.3
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Median
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-25.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-3.7
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-45.9
-\end{minipage} \\
-\multirow{2}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Sex
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-F
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-163 (45.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (50.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-114 (44.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-M
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-198 (54.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (50.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-143 (55.6\%)
-\end{minipage} \\
-\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Race
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-White
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-252 (69.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-42 (56.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-191 (74.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Black
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-40 (11.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16 (21.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-21 (8.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Asian
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Mixed
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-7 (1.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (1.4\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6 (2.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Other
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-48 (13.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12 (16.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-29 (11.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Unknown/Declined
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2 (0.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (1.4\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2 (0.8\%)
-\end{minipage} \\
-\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Maternal Ed.
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-Junior high school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Some high school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6 (1.7\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-3 (4.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-4 (1.6\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-High school grad
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (10.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-13 (17.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16 (6.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Some college
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-95 (26.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-25 (33.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-57 (22.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-College grad
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-90 (24.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17 (23.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-69 (26.8\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Grad/professional school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-123 (34.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-15 (20.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-104 (40.5\%)
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-Several studies have demonstrated links between specific taxa and
-measures of anxiety and depression
-\href{https://www.zotero.org/google-docs/?X5rfKj}{(Mayneris-Perxachs et
-al., 2022; Needham et al., 2022)}, cognitive flexibility in adults
-\href{https://www.zotero.org/google-docs/?RGIf8m}{(Magnusson et al.,
-2015)}, and atypical neural development
-\href{https://www.zotero.org/google-docs/?hwPZKh}{(Liu et al., 2019; Wan
-et al., 2021)}. We, therefore, set out to identify whether specific taxa
-or gene functions were linked with normal cognitive development in
-children. In order to assess whether variations in gut microbial taxa,
-their genes, or their metabolism are linked with neurocognitive
-development, we tested whether the beta diversity of microbial taxa and
-gene functions, as well as variation in brain development as assessed by
-neuroimaging, were associated with these contemporaneous measures of
-cognitive function using permutation analysis of variance (PERMANOVA).
-Due to the large ecological shift in the microbiome that occurs upon the
-introduction of solid food, and a relatively wide range of ages when
-infants are transitioned to solid food, we considered the pre-transition
-(less than 6 months old) and post-transition (over 18 months old)
-microbiomes separately. Even still, age was a major driver of variation
-in gut microbiomes in children over 18 months old (Figure 1G). We also
-found that overall variation in microbial species in children over 18
-months old was associated with a small but significant variation in
-cognitive function score (Figure 1G, R\textsuperscript{2} = 0.0124, q
-\textless{} 0.001), as was variation in microbial gene functions
-(R\textsuperscript{2} = 0.0119, q \textless{} 0.001). Variation in
-microbial taxa and genes was not significantly associated with cognitive
-function in children under 6 months, though this may be due to the low
-taxonomic diversity, and broad lack of overlap between taxa in infants.
-As expected, age was significantly associated with microbial beta
-diversity (taxa R\textsuperscript{2} = 0.0215, and UniRef90s
-R\textsuperscript{2} = 0.0245, q \textless{} 0.001), and very strongly
-associated with neuroimaging profiles (R\textsuperscript{2} = 0.256, q
-\textless{} 0.001).
-
-Consistent with prior studies, different microbial measurement types
-captured overlapping variation, with species profiles and gene function
-profiles (annotated with clusters of 90\% similarity
-\href{https://www.zotero.org/google-docs/?b9rEJS}{(Suzek et al.,
-2007)}), both generated from DNA sequencing data, being tightly coupled
-(Figure 1H, p \textless{} 0.001). Interestingly, other functional
-groupings (Enzyme commission level4 - ECs
-\href{https://www.zotero.org/google-docs/?fs8LO7}{(Bairoch, 2000)}, and
-KEGG orthologues - KOs
-\href{https://www.zotero.org/google-docs/?Xx5duE}{(Kanehisa et al.,
-2004)}) overlapped only slightly with taxonomic profiles in both age
-cohorts, despite being derived from UniRef90 labels. In children over 18
-months, some variation (15.9\%, p \textless{} 0.01) in neuroimaging
-overlapped with microbial measures, though this may be due to the
-residual variation due to age in both measures.
-
-\hypertarget{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}{%
-\subsubsection{Microbial species and neuroactive genes are associated
-with cognitive
-performance}\label{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}}
-
-To assess whether individual microbial species were associated with
-cognitive function, we fit multivariable linear regression
-\href{https://www.zotero.org/google-docs/?4qS6UF}{(Mallick et al.,
-2021)} to the relative abundance of each species that had at least 15\%
-prevalence in a given age group (Figure 2A, N = 92 for 0--120 months, N
-= 46 for 0--6 months, N = 97 for 18--120 months). Only \emph{Blautia
-wexlerae} was significantly associated (q value = 0.14, β = 0.0015) with
-cognitive function in children under 6 months old after adjusting for
-age and maternal education (Figure 2B). \emph{B. wexlerae} was
-previously shown to be depleted in children with diabetes
-\href{https://www.zotero.org/google-docs/?6mU2nI}{(Benítez-Páez et al.,
-2020)}, and that oral administration of \emph{B. wexlerae} partially
-ameliorated weight gain and inflammation from a high-fat diet in a mouse
-model of T2D \href{https://www.zotero.org/google-docs/?y83hpS}{(Hosomi
-et al., 2022; Liu et al., 2021)}. In children over 18 months of age,
-several microbial species were significantly enriched (q-value
-\textless{} 0.20) in children with higher cognitive function scores,
-including \emph{Gordonibacter pamelaeae}, which produces the
-neuroprotective metabolite urolithin
-\href{https://www.zotero.org/google-docs/?HGtTEY}{(Gong et al., 2022;
-Selma et al., 2014)}, \emph{Asaccharobacter celatus} and
-\emph{Adelcreutzia equolifaciens}, which produce phytoestrogen-derived
-equol \href{https://www.zotero.org/google-docs/?HCKZqm}{(Maruo et al.,
-2008; Thawornkuno et al., 2009)}, and the SCFA-producing probiotic
-species such as \emph{Eubacterium eligens} and \emph{Faecalibacterium
-prausnitzii} \href{https://www.zotero.org/google-docs/?3eyheK}{(Ghosh et
-al., 2020; Lopez-Siles et al., 2017; Mukherjee et al., 2020)} (Figure
-2B).
-
-\includegraphics[width=6.5in,height=4.83333in]{media/image4.png}
-
-\textbf{Figure 2 - Taxa and gene functional groups are associated with
-cognitive function.}
-
-(A) Volcano plot of multivariable linear model results showing the
-relationship between individual taxa and cognitive function score in
-children over 18 months of age, controlling for age, maternal education,
-and sequencing depth. Taxa that were significant after FDR correction (q
-\textless{} 0.2) are colored red. (B) For taxa that were significantly
-associated with cognitive function, heatmaps of prevalence, mean
-relative abundance, and correlation with cognitive function in different
-age groups. (C-F) Enrichment plots for selected neuroactive gene sets
-used in feature set enrichment analysis (FSEA). (G) summary of FSEA
-results across all samples (left) as well as in the under 6 months
-(middle) and over 18 months (right) subsets, colored based on the
-significance of the association. Markers indicate the individual
-correlation of genes within a gene set, and vertical bars represent the
-median correlation of that gene set.
-
-Given that different microbial species might occupy the same metabolic
-niche in different individuals, we hypothesized that microbial genes
-grouped by functional activity would be associated with cognition. To
-test this, we performed feature set enrichment analysis (FSEA) on groups
-of genes with neuroactive potential
-\href{https://www.zotero.org/google-docs/?xQURsj}{(Valles-Colomer et
-al., 2019)} and concurrent cognitive function score (Table 2,
-\protect\hyperlink{_i7fd3z19jdvg}{\uline{Figure 2}}C-G) and found that
-several metabolic pathways were significantly enriched or depleted in
-children with higher cognitive function scores. This was true both when
-considering all age groups together, though the enrichment of most
-pathways was more pronounced in children under 6 months or over 18
-months. For example, genes for degrading the 3-carbon SCFA propionate
-were significantly depleted in children with higher cognitive function
-scores across all age groups tested (Table 2; propionate degradation I,
-under 6 months, enrichment score (E.S.) = -0.542, corrected p-value (q)
-= 0.020; over 18 months, E.S. = -0.674, q = 0.041). Interestingly, genes
-for propionate synthesis were also significantly depleted in higher
-scoring children over 18 months (E.S. -0.676, q = 0.023), as were genes
-for synthesizing the 2 carbon SCFA acetate in children under 6 months
-old (acetate synthesis I, E.S. = -0.194, q = 0.153; acetate synthesis
-II, E.S. = -0.342, q = 0.020; acetate synthesis III, E.S. = -0.31, q =
-0.052). SCFAs are produced by anaerobic fermentation of dietary fiber
-and have been linked with immune system regulation as well as directly
-with brain function
-\href{https://www.zotero.org/google-docs/?2Z0sKZ}{(Dalile et al.,
-2019)}.
-
-Synthesis of menaquinone (vitamin K) was also negatively associated with
-cognitive function score in older children (menaquinone synthesis I,
-E.S. = -0.170, q = 0.0183). Menaquinone has several isoforms, one of
-which, MK-4, has been found to be decreased in children diagnosed with
-ASD \href{https://www.zotero.org/google-docs/?dTA3NO}{(Dong et al.,
-2021)} and is potentially neuroprotective in both rodents and humans
-\href{https://paperpile.com/c/dPbU4e/cTO9}{(Elkattawy et al., 2022)}. In
-children over 18 months old, genes for the synthesis of the amino acids
-glutamate and tryptophan were significantly enriched in children with
-higher cognitive function scores (Glutamate synthesis I, E.S. = 0.242, q
-= 0.047; Tryptophan synthesis, E.S. = 0.119, q = 0.041). Glutamate is a
-critical neurotransmitter controlling neuronal excitatory/inhibitory
-signaling along with gamma-aminobutyric acid (GABA), and their balance
-in the brain controls neural plasticity and learning, particularly in
-the developing brain
-\href{https://www.zotero.org/google-docs/?QubMhx}{(Cohen Kadosh et al.,
-2015; Palomo-Buitrago et al., 2019)}. Tryptophan metabolism, including
-microbial metabolism of tryptophan, has previously been linked with
-autism in children
-\href{https://www.zotero.org/google-docs/?GPaUZE}{(Hoshino et al., 1984;
-Xiao et al., 2021)}. Taken together, these results suggest that
-microbial metabolic activity, particularly the metabolism (synthesis and
-degradation) of neuroactive compounds may have effects on cognitive
-development.
-
-\textbf{Table 2 - Feature set enrichment analysis on neuroactive
-microbial gene sets}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2300}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1259}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1525}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1429}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2785} + 2\tabcolsep}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{All}
-\end{minipage}} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2591} + 2\tabcolsep}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{under 6 months}
-\end{minipage}} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2324} + 2\tabcolsep}@{}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{over 18 months}
-\end{minipage}} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-geneset
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1223
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5549
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1944
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1533}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1269
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. II
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2201
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0660}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3424
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0198}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0946
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9199
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. III
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2104
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0859}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3098
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0518}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0857
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9199
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Glutamate deg. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.4037
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1802}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2840
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2489
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Glutamate syn. II
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0953
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6022
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1342
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9802
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2421
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0466}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Menaquinone syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0853
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5635
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0833
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1699
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1825}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Propionate deg. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.4612
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1302}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.5421
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0198}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.6738
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0414}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Propionate syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3344
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5549
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3201
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5713
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.6760
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0227}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Tryptophan syn.
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0573
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9742
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0725
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1186
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0414}
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\hypertarget{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}{%
-\subsubsection{Gut microbial taxonomic and functional profiles predict
-cognitive
-function}\label{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}}
-
-FSEA relies on understanding functional relationships between individual
-genes. However, because the relationships between individual taxa are
-still largely unknown, we turned to Random Forest (RF) models, an
-unsupervised non-parametric machine learning (ML) method that enables
-the identification of underlying patterns in large numbers of individual
-features (here, microbial species). Previous studies have reported
-successful use of RFs for processing highly-dimensional and sparse data
-from the domain of genomics
-\href{https://www.zotero.org/google-docs/?C3CSPB}{(Amaratunga et al.,
-2008; Brieuc et al., 2018; Chen and Ishwaran, 2012; Franzosa et al.,
-2019; Stephan et al., 2015)}, along with other works where it was used
-to predict cognitive conditions related to Alzheimer's disease in
-different scenarios
-\href{https://www.zotero.org/google-docs/?GiadGy}{(Ardekani et al.,
-2017; Velazquez et al., 2021)}. Additionally, given the sequential
-nature of variable consideration in each tree, RFs are naturally able to
-work out complex input feature interactions, such as those present in a
-microbiome-wide study, without the necessity to explicitly compute
-interaction terms.
-
-\includegraphics[width=6.5in,height=5.19444in]{media/image2.png}
-
-\textbf{Figure 3 - Random forest models predict concurrently measured
-cognitive function.}
-
-Comparison of RF predictor importance versus linear models for children
-between birth and 6 months old (A), and for those older than 18 months
-(B). Colors represent whether the species belong to the group of
-top-important features that account for 80\% of the cumulative
-importance on the RF model, if that species was significant (q
-\textless{} 0.2) in linear models, both, or neither. Ranked feature
-importance for taxa in RF models for children between birth and 6 months
-old (C), and for those older than 18 months (D). Taxa that are important
-for RF models both for children under 6 months and for those over 18
-months (E), only for children under 6 months (F), and only for children
-over 18 months (G)
-
-Given that gut microbial profiles, as well as neurocognitive
-development, may partially reflect socioeconomic and demographic
-factors, we assessed the performance of RF regressors where maternal
-education (a proxy of socioeconomic status (SES)), sex, and age were
-included as possible predictors, either alone or in combination with
-microbial taxonomic profiles (Table 3).
-
-\textbf{Table 3. Benchmark metrics for the cognitive assessment score
-prediction models. Confidence intervals are calculated from the
-distribution of metrics from repeated CV at a confidence level of 95\%}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1456}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0960}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3584}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Subject Ages (months)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Microbial feature
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Demo.
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Test set correlation (± C.I.)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Test set Root-mean-square error (± C.I.)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.14 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-13.01 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.10 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.66 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.13 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.70 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.01 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.56 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.01 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.56 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.506 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16.27 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.363 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.66 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.429 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.29 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.303 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18.01 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.333 ± 0.004
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.86 ± 0.04
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-As with linear models, RF models for children under 6 months old were
-less generalizable (mean test-set correlation -0.13, mean RMSE 12.70),
-but RFs were consistently able to learn the relationship between taxa
-and cognitive function scores in children over 18 months of age (mean
-test-set correlation 0.429, mean RMSE 17.29). For both age groups,
-species that were important in RF overlapped with those that were
-significant when testing the relationship with linear models (Figure
-3A-B, Supplementary Table XXA-B), though there were substantial
-differences. For children under 6 months, only \emph{B. wexlerae} was
-significantly associated with cognitive function in linear models, but
-was ranked 20th in importance for RF models. The importance metric
-employed (MDI) is a measure of how well the variable differentiates
-(splits) sets of samples by creating subgroups that reduce the
-intra-group deviations, while also accounting for the amount of samples
-affected by a split. All taxa significantly associated with cognitive
-function score in children over 18 months using LMs belong to the
-top-ranking group responsible for 80\% of the total relative importance,
-except for \emph{Eubacterium ramulus} (Table 3, Figure 3A).
-
-Interestingly, several taxa highly ranked in importance in both age
-groups, including several that were significant in LMs for children over
-18 months old, including \emph{R. gnavus} (0--6 months, rank = 13;
-18--120 months, rank = 7) and \emph{G. pamelaeae} (0--6 months, rank =
-16; 18--120 months, rank = 20), while others such as \emph{Allistipes
-finegoldii} were age-group specific. Several taxa important in RF models
-were not statistically significant when using linear models after
-multiple hypothesis correction. However, these taxa had small nominal
-p-values. For example, \emph{Erysipelatoclostridium ramosum} was the
-most important feature in RF models for children under 6 months old and
-had an LM p-value of 0.04. Subject age was consistently ranked highly in
-feature importance, which could indicate that decision branches based on
-microbial taxa have increased purity when considering the subject's age
-or that age itself is a useful predictor.
-
-\hypertarget{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}{%
-\subsubsection{Gut microbial taxonomic profiles predict brain structure
-differences}\label{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}}
-
-If there are causal effects of microbial metabolism on cognitive
-function, they might be reflected in changes in neuroanatomy. We again
-employed a Random Forest modeling approach to associate gut taxonomic
-profiles with individual brain regions identified in MRI scans,
-normalized to total brain volume. Some brain regions were more readily
-predicted by RF models trained on microbial taxa (Table 4, Supplementary
-Table XX), in particular those that were highly correlated with age.
-These included the L/R lingual gyrus (mean RF correlations, Left =
-0.421, Right = 0.434; relative age importances, Left = 7.5\%, Right =
-8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
-0.200, Right = 0.273; relative age importances, Left = 3.7\%, Right =
-6.3\%). In many cases, however, age was not an important variable in
-high-performing models, such as that for the left accumbens area (mean
-RF correlations = 0.288; relative age importances = 1.2\%).
-
-\textbf{Table 4. Summary statistics for the neuroanatomy prediction
-benchmarks}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Statistic
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Mean absolute proportional error (MAPE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Correlation coefficient (R)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-mean
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.072
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.137
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Standard deviation
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.028
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.136
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-maximum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.605
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-75th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.195
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-50th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-25th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.062
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-minimum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.147
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-We also observed that many brain regions had high accordance across the
-left and right hemispheres in terms of both model performance and
-microbial feature importance. In contrast, other regions had substantial
-differences between the hemispheres. For example, the left accumbens
-area, which plays an important role in reward circuits
-\href{https://www.zotero.org/google-docs/?zVicU7}{(Ernst et al., 2005;
-Yau et al., 2012)}, has one of the highest test-set correlations of our
-brain region models (R = 0.288), as compared to the right accumbens
-models which could not adequately generalize, and had a negative mean
-test-set correlation (R = -0.041). Most healthy individuals have a
-rightward asymmetry in the nucleus accumbens, and reduced asymmetry has
-been linked to substance use disorder in young adults
-\href{https://www.zotero.org/google-docs/?V45KyN}{(Cao et al., 2021)}.
-Feature importance for models of the left accumbens area were dominated
-by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.4\%
-relative importance), \emph{B. ovatus} (3.7\% relative importance), and
-\emph{B. uniformis} (3.0\% relative importance).. The accumbens area is
-associated with reward control, and in individuals diagnosed with ADHD,
-it has been shown to have a divergent neuromorphology
-\href{https://www.zotero.org/google-docs/?kLfEQ5}{(Hoogman et al.,
-2017)}. Independently, \emph{B. ovatus}, \emph{B. uniformis} and
-\emph{B. vulgatus} have been linked to ADHD
-\href{https://www.zotero.org/google-docs/?luO0zh}{(Wang et al., 2020)}.
-In fact, there are studies showing that alterations on the striatal
-dopamine transporters can cause effects resembling hyperactivity and
-attention deficit
-\href{https://www.zotero.org/google-docs/?UxS47e}{(Yael et al., 2019)},
-and that \emph{B. uniformis} is gut-microbial modulator of the brain
-dopamine transporter
-\href{https://www.zotero.org/google-docs/?L1cWJS}{(Hartstra et al.,
-2020)}.
-
-Many of the taxa identified included taxa also identified in LMs and RF
-models of cognitive function. Interesting to note, while RF models for
-multiple brain regions had many important microbial taxa, others were
-dominated by a small number of taxa. In general, for all segments, a
-consistent number of species between 14 and 22 (median = 20) was
-responsible for one third of the fitness-weighted cumulative importance
-(see Methods), regardless of the model benchmarks. We proceeded to
-select a subset of the most important taxa (averaged over all segments)
-and the segments whose importance was more heavily loaded to further
-analyze the multiple relationships unveiled by the RF models
-(\textbf{Figure 4 A and B})
-
-\includegraphics[width=6.5in,height=5.19444in]{media/image1.png}
-
-\textbf{Figure 4: Microbial feature importance in predicting brain
-volumes in children over 18 months of age.}
-
-(A) Average test-set correlations for prediction of MRI segmentation
-data from microbiome and demographics on select segments after repeated
-cross validation. (B) Heatmap of average individual relative taxa
-importances on each brain segment. Importances are reported as
-proportions relative to the sum of importances for each model - since
-every model is trained with 132 features, and even distribution of
-importance would be 0.75\% for each feature.
-
-. Segments and taxa ordered by HCA on a select list of species with high
-ranks on average importances, and their respective highest-load
-segments. (C)
-
-Our analysis revealed two major patterns of importance distribution from
-taxa over the brain segments; some species portrayed high contributions
-to multiple different segments, while others contributed modestly to
-just one or two brain segments. Notable cases of the first pattern
-included seven species - \emph{Anaerostipes hadrus}, \emph{Bacteroides
-vulgatus}, \emph{Fusicatenibacter saccharivorans}, \emph{Ruminococcus
-torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
-\emph{Blautia wexlerae} - that combined, account for approximately one
-third of the cumulative relative importances, computed after subsetting
-on the taxa of interest (Figure 4C).
-
-Among these, the most important variable is \emph{Anaerostipes hadrus},
-which is a butyrate-producing anaerobe that has been positively
-associated with cognitive function
-\href{https://www.zotero.org/google-docs/?AVJiwu}{(Kant et al., 2015; Li
-et al., 2022)}. Its importance is, however, heavily loaded on the volume
-of entire major lobes (frontal, parietal), which cannot be readily
-correlated to specific cognitive functions. After accounting for the
-major lobes, it is found to be highly important on the prediction of the
-cerebellar vermal lobules VIII-X, pars opercularis, cuneus and
-precuneus, and anterior-cingulate (Figure 4B).
-
-To better understand this relationship, we performed hierarchical
-cluster analysis, which revealed a well defined cluster of species with
-high importance loadings on close segments of the lower temporal and
-close occipital lobe, to which \emph{A. hadrus} belongs. The other
-species on this cluster are \emph{B. wexlerae}, \emph{R. torques},
-\emph{R. intestinalis, R. bicirculans} and \emph{F. saccahrivorans}, who
-contribute heavily to the entorhinal, fusiform, lingual and
-parahippocampal segments.
-
-This group of taxa and their related segments drew our attention because
-they contained \emph{B. wexlerae}, a highlight from the cognitive
-assessment results. Previous works exploring Parkinson Disease patients
-found out that issues in the cognitive task of confrontation naming were
-positively correlated with thinning in the fusiform gyrus and
-parahippocampal gyrus
-\href{https://www.zotero.org/google-docs/?rlO6eq}{(Pagonabarraga et al.,
-2013)}. Additionally, the left and right parahippocampal, where \emph{B.
-wexlerae} had the highest importance in both models, are important in
-visual/spatial processing and memory
-\href{https://www.zotero.org/google-docs/?wkkWZT}{(Aminoff et al.,
-2013)}. \emph{B. wexlerae} can also produce acetylcholine in the gut
-\href{https://www.zotero.org/google-docs/?jBoxSm}{(Hosomi et al.,
-2022)}, and this molecule plays an important role in modulating memory
-function \href{https://www.zotero.org/google-docs/?BQ0j2c}{(Haam and
-Yakel, 2017)}.
-
-Another notable importance cluster contains taxa associated with the
-basal forebrain, especially the cingulate and the accumbens. This
-cluster contains the previously-discussed \emph{B. ovatus} and \emph{B.
-uniformis,} but also \emph{Alistipes finegoldii and Streptococcus
-salivaris.} Reduction in the nucleus accumbens has been associated with
-depression symptoms
-\href{https://www.zotero.org/google-docs/?2aSZIr}{(Wacker et al.,
-2009)}, and increased levels of the \emph{Alistipes} genus have been
-observed in patients with depressive disorder
-\href{https://www.zotero.org/google-docs/?bhhSZB}{(Jiang et al., 2015)}.
-While these reports are important, simultaneously probing of the gut
-metagenome and brain structure relationship is novel.
-
-Both \emph{A. equolofaciens} and \emph{A. celatus,} two closely-related
-equol-producing species, are examples of the second contribution
-pattern, and had high importance in predicting the relative volume of
-the right anterior cingulate (respectively, 3.0\% and 2.6\% relative
-importances), which has been linked to social cognition and reward-based
-decision making \href{https://www.zotero.org/google-docs/?Wi8Bpq}{(Apps
-et al., 2016; Boes et al., 2008; Bush et al., 2002)}. Equol has a strong
-estrogenic effect
-\href{https://www.zotero.org/google-docs/?PqgFU7}{(Setchell et al.,
-2005)}, and in anterior cingulate, estrogen, has been shown to regulate
-pain-related aversion
-\href{https://www.zotero.org/google-docs/?5huRIR}{(Xiao et al., 2013)}.
-Another example of this pattern is \emph{R. gnavus,} which was the only
-species with significant negative association with cognitive function.
-It is heavily associated with the left pars opercularis (2.7\% relative
-importance), a relationship that may be explained by the emerging
-understanding that this species is increased in individuals with insulin
-resistance and obesity
-\href{https://www.zotero.org/google-docs/?lwzWUv}{(Ley et al., 2006)},
-conditions that are known to produce structural abnormalities in the
-brain \href{https://www.zotero.org/google-docs/?NlkFr6}{(Opel et al.,
-2021)}. Finally, \emph{Coprococcus comes} displays an importance
-distribution that splits almost evenly among the two previously reported
-clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (4.0\% relative importance)
-(only age had higher relative importance in any model), while also being
-one of the most important predictors for neighboring areas of the
-posterior cingulate such as the pars opercularis (relative importances,
-Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
-precentral (2.3\% relative importance) and paracentral lobes (relative
-importances, Left = 2.6\%, Right = 1.8\%).
-
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
-
-The relationship between the gut microbiome and brain function via the
-gut-microbiome-brain axis has gained increasing acceptance largely as a
-result of human epidemiological studies investigating atypical
-neurocognition (eg anxiety and depression, neurodegeneration, attention
-deficit / hyperactivity disorder, and autism) and mechanistic studies in
-animal models. The results from these studies point to the possibility
-that gut microbes and their metabolism may be causally implicated in
-cognitive development, but this study is the first to our knowledge that
-directly investigates microbial species and their genes in relation to
-typical development in young children. Understanding the gut-brain-axis
-in early life is particularly important, since differences or
-interventions in early life can have outsized and longer-term
-consequences than those at later ages. Further, even in the absence of
-causal impacts of microbial metabolism, identifying risk factors that
-could point to other early interventions would also have value.
-
-The use of shotgun metagenomic sequencing enabled us to get
-species-level resolution of microbial taxa. A previous study of
-cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing,
-and showed that genera from the Lachnospiraceae family as well as
-unclassified Clostridiales (now Eubacteriales) were associated with
-higher scores on the Ages and Stages Questionnaire
-\href{https://www.zotero.org/google-docs/?SB5Lwa}{(Sordillo et al.,
-2019)}. However, each of these clades encompass dozens of genera with
-diverse functions, each of which may have different effects. Indeed,
-several of the taxa that were positively associated with cognitive
-function in this study, including \emph{B. wexlerae}, the only species
-identified by linear models in children under 6 months, \emph{D.
-longicatena}, \emph{R. faecis}, and \emph{A. finegoldii} are
-Clostridiales, as is \emph{R. gnavus}, which we found was negatively
-associated with cognitive function (Figure 2A-B). This kind of
-species-level resolution is typically not possible with amplicon
-sequencing.
-
-We identified several species in the family Eggerthelaceae that were
-associated with cognitive function, including \emph{Gordonibacter
-pamelaeae, Aldercreutzia equolofaciens, Asaccharobacter celatus}
-(formerly regarded a subspecies of \emph{A. equolofaciens}
-\href{https://www.zotero.org/google-docs/?OEhguI}{(Takahashi et al.,
-2021)}), and \emph{Eggerthella lenta}. Many members of this family are
-known in part due to unique metabolic activities. For example, \emph{A.
-equolofaciens} produces the nonsteroidal estrogen equol from isoflaven
-found in soybeans
-\href{https://www.zotero.org/google-docs/?I2qNyB}{(Wang et al., 2005)},
-and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
-(found in pomegranates and some berries) into urolithin, which has been
-shown in some studies to have a neuroprotective effect
-\href{https://www.zotero.org/google-docs/?oTN8Lw}{(Gong et al., 2022;
-Selma et al., 2014)}. \emph{E. lenta} has been extensively studied for
-its ability to metabolize drug compounds such as the plant-derived heart
-medication digoxin
-\href{https://www.zotero.org/google-docs/?tH2Qeq}{(Haiser et al.,
-2013)}. The metabolic versatility of this clade, and the large number of
-species that are associated with cognition make these microbes prime
-targets for further mechanistic studies.
-
-In addition to improved species-level resolution, shotgun metagenomic
-sequencing also enables gene-functional insight. We showed here that
-genes for the metabolism of SCFAs, both their degradation and synthesis,
-are associated with cognitive function scores. However, while the
-differential abundance of genes for the metabolism of neuroactive
-compounds like these is suggestive, it is difficult to reason about the
-relationship between levels of these genes and the gut concentrations of
-the molecules their product enzymes act on. For example, while it might
-be intuitive to reason that increased levels of menaquinone synthesis
-genes is indicative of increased menaquinone, it could be the case that
-menaquinone deficiency selects for microbes that can synthesize it. For
-the same reason, increased propionate degradation genes may
-counterintuitively be indicative of high levels of propionate in the gut
-lumen, since high propionate would select for microbes that can
-metabolize it. For this reason, future studies coupling shotgun
-metagenomics with stool metabolomics could improve our understanding of
-the relationship between microbial metabolism and cognitive development.
-Further, strain-level analysis linking specific gene content in species
-of interest could further refine targeted efforts at identifying
-specific metabolic signatures of microbe-brain interactions.
-
-\{\{Guilherme add a paragraph about the usefulness of ML\}\}
-
-The use of multiple age-appropriate cognitive assessments that could be
-normalized to a common scale enabled us to analyze microbial
-associations across multiple developmental periods, but carries several
-drawbacks. In particular, the test-retest reliability, as well as
-systematic differences between test administrators may introduce
-substantial noise into these observations, particularly in the youngest
-children. In addition, our study period overlapped with the beginning of
-the COVID-19 pandemic, and we and others have observed some reduction in
-measured scores for children that were assessed after the implementation
-of lockdowns. In our subject set for this study, these effects are more
-pronounced in some age groups due to our sampling schedule
-\href{https://www.zotero.org/google-docs/?L8GnRj}{(Blackwell et al.,
-2022; Deoni et al., 2021)} (Supplementary Figure 3).
-
-This analysis allowed us to establish links between microbial taxa and
-their functional potential with cognition and brain structure. Although
-we cannot test causality or the chemistry behind the interactions
-between gut microbial taxa, gut, and brain, this study provides clear
-and statistically significant associations between the infant and early
-child gut microbiota and neurocognition. Future studies should focus on
-characterizing the early-life microbiome and neurocognitive development
-across different geographic regions and lifestyles such as covering
-traditionally understudied low-resource urban, peri-urban and rural
-communities to obtain the more comprehensive understanding of the
-variability within the different gut microbiomes reflects on
-neurocognition. These studies would also provide us with the wealth of
-data on different strains from the same species to better understand the
-effect of genes and their products. Furthermore, culturing and microbial
-community enrichment studies combined with genetic manipulation and
-genomic approaches to understand microbial metabolism at the molecular
-level is the key, as the metabolic functions shape and influence the
-human host and its health. The discovery of the neuroactive metabolites
-could provide us with biomarkers for early detection or necessary
-medicinally useful molecules that can be applied in intervention.
-
-\hypertarget{material-and-methods}{%
-\subsection{Material and Methods}\label{material-and-methods}}
-
-\hypertarget{study-ethics}{%
-\subsubsection{Study Ethics}\label{study-ethics}}
-
-All procedures for this study were approved by the local institutional
-review board at Rhode Island Hospital, and all experiments adhered to
-the regulation of the review board. Written informed consent was
-obtained from all parents or legal guardians of enrolled participants.
-
-\hypertarget{participants}{%
-\subsubsection{Participants}\label{participants}}
-
-Data used in this study were drawn from the ongoing longitudinal
-RESONANCE study of healthy and neurotypical brain and cognitive
-development, based at Brown University in Providence, RI, USA. The
-RESONANCE study is part of the NIH initiative Environmental influences
-on Child Health Outcomes (ECHO)
-\href{https://www.zotero.org/google-docs/?0tZanX}{(Forrest et al., 2018;
-Gillman and Blaisdell, 2018)}, a longitudinal observational study of
-healthy and neurotypical brain development that spans the fetal and
-infant to adolescent life stages, combining neuroimaging (magnetic
-resonance imaging, MRI), neurocognitive assessments, bio-specimen
-analyses, subject genetics, environmental exposures such as lead, and
-rich demographic, socioeconomic, family and medical history information.
-From the RESONANCE cohort, 361 typically-developing children between the
-ages of 2.8 months and 10 years old (median age 2 years, 2 months) were
-selected for analysis in this study.
-
-General participant demographics are provided in Table 1. Children are
-representative of the RI population. Children in the RESONANCE cohort
-were born full-term (\textless{} 37 weeks gestation) with height and
-weight normal for gestational age, and from uncomplicated singleton
-pregnancies. Children with known major risk factors for developmental
-abnormalities at enrollment were excluded. In addition to screening at
-the time of enrollment, on-going screening for worrisome behaviors using
-validated tools was performed to identify at-risk children and remove
-them from subsequent analysis.
-
-Exclusion criteria included: \emph{in utero} exposure to alcohol,
-cigarette or illicit substance exposure; preterm (\textless{} 37 weeks
-gestation) birth; small for gestational age or less than 1500 g; fetal
-ultrasound abnormalities; preeclampsia, high blood pressure, or
-gestational diabetes; 5 minute APGAR scores \textless{} 8; NICU
-admission; neurological disorder (e.g., head injury resulting in loss of
-consciousness, epilepsy); and psychiatric or learning disorder
-(including maternal depression) in the infant, parents, or siblings
-requiring medication in the year prior to pregnancy.
-
-Demographic and other non-biospecimen data such as race and ethnicity,
-parental education and occupation, feeding behavior (breast- and
-formula-feeding), child weight and height, were collected through
-questionnaires or direct examination as appropriate. All data were
-collected at every assessment visit, if possible.
-
-\hypertarget{cognitive-assessments}{%
-\subsubsection{Cognitive Assessments}\label{cognitive-assessments}}
-
-Overall cognitive function was assessed using age-appropriate methods.
-For children from birth to 30 months, we used an Early Learning
-Composite as assessed via the Mullen Scales of Early Learning (MSEL)
-\href{https://www.zotero.org/google-docs/?NSyykd}{(Mullen and others,
-1995)}, a standardized and population-normed tool for assessing fine and
-gross motor, expressive and receptive language, and visual reception
-functioning in children from birth through 68 months of age. The
-Wechsler Intelligence Quotient for Children (WISC)
-\href{https://www.zotero.org/google-docs/?OopdA1}{(Wechsler, 2012)} is
-an individually administered standard intelligence test for children
-aged 6 to 16 years. It derives a full scale intelligence quotient (IQ)
-score, which we used to assess overall cognitive functioning. The fourth
-edition of the Wechsler Preschool and Primary Scale of Intelligence
-(WPPSI-IV) is an individually administered standard intelligence test
-for children aged 2 years 6 months to 7 years 7 months, trying to meet
-the increasing need for the assessment of preschoolers. Just as the
-WISC, it derives a full scale IQ score, which we used to assess overall
-cognitive functioning.
-
-\hypertarget{stool-sample-collection-and-sequencing}{%
-\subsubsection{Stool Sample Collection and
-Sequencing}\label{stool-sample-collection-and-sequencing}}
-
-Stool samples (n=493) were collected by parents in OMR-200 tubes
-(OMNIgene GUT, DNA Genotek, Ottawa, Ontario, Canada), stored on ice, and
-brought within 24 hrs to the lab in RI where they were immediately
-frozen at -80 ˚C. Stool samples were not collected if the subject had
-taken antibiotics within the last two weeks. DNA extraction was
-performed at Wellesley College (Wellesley, MA). Nucleic acids were
-extracted from stool samples using the RNeasy PowerMicrobiome kit,
-excluding the DNA degradation steps. Briefly, the samples were lysed by
-bead beating using the Powerlyzer 24 Homogenizer (Qiagen, Germantown,
-MD) at 2500 rpm for 45 s and then transferred to the QIAcube (Qiagen,
-Germantown, MD) to complete the extraction protocol. Extracted DNA was
-sequenced at the Integrated Microbiome Resource (IMR, Dalhousie
-University, NS, Canada).
-
-Shotgun metagenomic sequencing was performed on all samples. A pooled
-library (max 96 samples per run) was prepared using the Illumina Nextera
-Flex Kit for MiSeq and NextSeq from 1 ng of each sample. Samples were
-then pooled onto a plate and sequenced on the Illumina NextSeq 550
-platform using 150+150 bp paired-end ``high output'' chemistry,
-generating 400 million raw reads and 120 Gb of sequence per plate.
-
-\hypertarget{machine-learning-for-cognitive-development}{%
-\subsubsection{Machine learning for cognitive
-development}\label{machine-learning-for-cognitive-development}}
-
-Prediction of cognitive scores was carried out as a set of regression
-experiments targeting real-valued continuous assessment scores.
-Different experiment sets were designed to probe how different
-representations of the gut microbiome (taxonomic profiles, Functional
-profiles encoded as ECs) would behave, with and without the addition of
-demographics (sex and maternal education as a proxy of socioeconomic
-status) on participants from different age groups. Age (in months) was
-provided as a covariate for all models (Table 3).
-
-\textbf{Table 5. Experimental design and input composition for Random
-Forest experiments}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1797}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2915}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Input set
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Age bracket
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Microbiome encoding type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Demographics
-
-Provided?
-
-(sex, education)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-1
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Not provided
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-3
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-4
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-5
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Not provided
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-7
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-9
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-10
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-Random Forests (RFs) \href{https://paperpile.com/c/dPbU4e/VDqU}{(Breiman
-1996)} were selected as the prediction engine and processed using the
-DecisionTree.jl
-\href{https://www.zotero.org/google-docs/?dQLyDs}{(Sadeghi et al.,
-2022)} implementation, inside the MLJ.jl
-\href{https://www.zotero.org/google-docs/?y8ywGI}{(Blaom et al., 2020)}
-Machine Learning framework. Independent RFs were trained for each
-experiment, using a set of default regression hyperparameters from
-Breiman and Cutler
-\href{https://www.zotero.org/google-docs/?d3gqFh}{(Breiman, 2001)}, on a
-repeated cross-validation approach with different RNG seeds. One hundred
-repetitions of 3-fold CV with 10 different intra-fold RNG states each
-were employed, for a total of 3000 experiments per input set.
-
-After the training procedures, the root-mean-square error (RMSE) for
-cognitive assessment scores and mean absolute proportional error (MAPE)
-for the brain segmentation data, along with Pearson\textquotesingle s
-correlation coefficient (R) were benchmarked on the validation and train
-sets. MAPE was chosen as the metric for brain segments due to magnitude
-differences between median volumes of each segment, which would hinder
-interpretation of raw error values without additional reference.
-
-To derive biological insight from the models, the covariate variable
-importances for all the input features, measured by mean decrease in
-impurity (MDI, or GINI importance), was also analyzed. Leveraging the
-distribution of results from the extensive repeated cross validation
-experiments, rather than electing a representative model or picking the
-highest validation-set correlation, a measure of model fitness
-(\textbf{Equation 1}) was designed to weight the importances from each
-trained forest. The objective was to give more weight to those with
-higher benchmarks on the validation sets (or higher generalizability),
-while penalizing information from highly overfit models, drawing
-inspiration from the approach used on another work employing repeated CV
-on Random Forests with high-dimensional, low sample size microbiome
-datasets \{***\}. The resulting fitness-weighted importances were used
-to generate the values in Figure 3.
-
-\(fitness =\)
-
-\textbf{Equation 1.} Mathematical expression of the fitness measure used
-to weight feature importances based on model benchmarks
-
-\hypertarget{mri-segmentation}{%
-\subsubsection{MRI / segmentation}\label{mri-segmentation}}
-
-MRI data was acquired at a 3T Siemens Trio scanner with the following
-parameters: TE=5.6msec, TR=1400msec, FA=15 degrees, 1.1x1.1x1.1mm
-resolution, 160x160 matrix, with an average of 112 slices. FOV was
-adjusted to infant size. Using a combination of linear and nonlinear
-image registration, we created representative age-specific templates
-using the ANTs package
-\href{https://www.zotero.org/google-docs/?ZlWDfL}{(Avants et al.,
-2014)}. After age specific templates were created, a single flow from
-each age to the 12-month template was estimated and a final warp from
-the 12-month template to standard adult MNI space was performed. For
-each individual infant/ child brain image, we then calculated the warp
-from their native T1w image space to their nearest-in-age template.
-Using the resulting warps (native → nearest age template → 12-month
-template → MNI template), we could move the standard adult brain atlas
-to the space of an individual infant in a single step. In this case, we
-used the Harvard-Oxford brain atlas to provide a coarse-grained
-parcellation of individual brains into subcortical regions (e.g.,
-thalamus, putamen) and total grey and white matter volumes (included as
-part of the FSL package
-\href{https://www.zotero.org/google-docs/?AapPN7}{(Jenkinson et al.,
-2012)}. Total tissue and brainstem volume as well as left and right
-hemisphere volumes were derived for total white and cortical gray
-matter, lateral ventricle, thalamus, caudate, putamen, pallidum,
-hippocampus, amygdala, and accumbens as well as total brainstem volume
-\href{https://www.zotero.org/google-docs/?MCDDB9}{(Bruchhage et al.,
-n.d.)}.
-
-\hypertarget{computational-analysis-of-metagenomes}{%
-\subsubsection{Computational analysis of
-metagenomes}\label{computational-analysis-of-metagenomes}}
-
-Shotgun metagenomic sequences were analyzed using the bioBakery suite of
-computational tools \href{https://paperpile.com/c/dPbU4e/vDau}{(Beghini
-et al. 2021)}. First, KneadData (v0.7.7) was used to perform quality
-control of raw sequence reads, such as read trimming and removal of
-reads matching a human genome reference. Next, MetaPhlAn (v3.0.7, using
-database mpa\_v30\_CHOCOPhlAn\_201901) was used to generate taxonomic
-profiles by aligning reads to a reference database of marker genes.
-Finally, HUMAnN (v3.0.0a4) was used to functionally profile the
-metagenomes.
-
-\hypertarget{data-and-code-availability}{%
-\subsubsection{Data and code
-availability}\label{data-and-code-availability}}
-
-Taxonomic and functional microbial profiles, as well as subject
-demographics necessary for statistical analyses and machine learning are
-available on the Open Science Framework
-\href{https://www.zotero.org/google-docs/?61HLlM}{(Bonham et al.,
-2022)}. Data processing, generation of summary statistics, and
-generation of plots was performed using the julia programming language
-\href{https://www.zotero.org/google-docs/?zj3HnY}{(Bezanson et al.,
-2017; Bonham et al., 2021; Danisch and Krumbiegel, 2021)}. All code for
-data analysis and figure generation, as well as scripts for automated
-download of input files are available on github \{\{CITE zenodo\}\}.
-
-\hypertarget{research-standards}{%
-\subsection{Research Standards}\label{research-standards}}
-
-\hypertarget{acknowledgements}{%
-\subsection{Acknowledgements}\label{acknowledgements}}
-
-Research funding: NIH UG3 OD023313 and Wellcome LEAP 1kD.
-
-\hypertarget{references-and-citations}{%
-\subsection{References and Citations}\label{references-and-citations}}
-
-\url{https://doi.org/10.1093/cercor/bhs201}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Amaratunga, D.,
-Cabrera, J., Lee, Y.-S., 2008. Enriched random forests. Bioinformatics
-24, 2010--2014. https://doi.org/10.1093/bioinformatics/btn356}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Aminoff, E.M.,
-Kveraga, K., Bar, M., 2013. The role of the parahippocampal cortex in
-cognition. Trends Cogn. Sci. 17, 379--390.
-https://doi.org/10.1016/j.tics.2013.06.009}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Apps, M.A.J.,
-Rushworth, M.F.S., Chang, S.W.C., 2016. The Anterior Cingulate Gyrus and
-Social Cognition: Tracking the Motivation of Others. Neuron 90,
-692--707. https://doi.org/10.1016/j.neuron.2016.04.018}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ardekani, B.A.,
-Bermudez, E., Mubeen, A.M., Bachman, A.H., Initiative, for the A.D.N.,
-2017. Prediction of Incipient Alzheimer's Disease Dementia in Patients
-with Mild Cognitive Impairment. J. Alzheimers Dis. 55, 269--281.
-https://doi.org/10.3233/JAD-160594}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Avants, B.B.,
-Tustison, N.J., Stauffer, M., Song, G., Wu, B., Gee, J.C., 2014. The
-insight ToolKit image registration framework. Front. Neuroinform. 8, 44.
-https://doi.org/10.3389/fninf.2014.00044}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bäckhed, F., Roswall,
-J., Peng, Y., Feng, Q., Jia, H., Kovatcheva-Datchary, P., Li, Y., Xia,
-Y., Xie, H., Zhong, H., Khan, M.T., Zhang, J., Li, J., Xiao, L.,
-Al-Aama, J., Zhang, D., Lee, Y.S., Kotowska, D., Colding, C., Tremaroli,
-V., Yin, Y., Bergman, S., Xu, X., Madsen, L., Kristiansen, K., Dahlgren,
-J., Wang, J., 2015. Dynamics and stabilization of the human gut
-microbiome during the first year of life. Cell Host Microbe 17,
-690--703. https://doi.org/10.1016/j.chom.2015.04.004}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bairoch, A., 2000. The
-ENZYME database in 2000. Nucleic Acids Res. 28, 304--305.
-https://doi.org/10.1093/nar/28.1.304}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Benítez-Páez, A.,
-Gómez del Pugar, E.M., López-Almela, I., Moya-Pérez, Á., Codoñer-Franch,
-P., Sanz, Y., 2020. Depletion of \emph{Blautia} Species in the
-Microbiota of Obese Children Relates to Intestinal Inflammation and
-Metabolic Phenotype Worsening. mSystems 5, e00857-19.
-https://doi.org/10.1128/mSystems.00857-19}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bezanson, J., Edelman,
-A., Karpinski, S., Shah, V.B., 2017. Julia: A fresh approach to
-numerical computing. Siam Rev. 59, 65--98.
-https://doi.org/10.1137/141000671}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Blackwell, C.K.,
-Mansolf, M., Sherlock, P., Ganiban, J., Hofheimer, J.A., Barone, C.J.,
-Bekelman, T.A., Blair, C., Cella, D., Collazo, S., Croen, L.A., Deoni,
-S., Elliott, A.J., Ferrara, A., Fry, R.C., Gershon, R., Herbstman, J.B.,
-Karagas, M.R., LeWinn, K.Z., Margolis, A., Miller, R.L., O'Shea, T.M.,
-Porucznik, C.A., Wright, R.J., 2022. Youth well-being during the
-COVID-19 pandemic. Pediatrics 149.
-https://doi.org/10.1542/peds.2021-054754}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Blaom, A.D., Kiraly,
-F., Lienart, T., Simillides, Y., Arenas, D., Vollmer, S.J., 2020. MLJ: A
-Julia package for composable machine learning. J. Open Source Softw. 5,
-2704. https://doi.org/10.21105/joss.02704}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Boes, A.D., Tranel,
-D., Anderson, S.W., Nopoulos, P., 2008. Right anterior cingulate cortex
-volume is a neuroanatomical correlate of aggression and defiance in
-boys. Behav. Neurosci. 122, 677--684.
-https://doi.org/10.1037/0735-7044.122.3.677}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bokulich, N.A., Chung,
-J., Battaglia, T., Henderson, N., Jay, M., Li, H., D Lieber, A., Wu, F.,
-Perez-Perez, G.I., Chen, Y., Schweizer, W., Zheng, X., Contreras, M.,
-Dominguez-Bello, M.G., Blaser, M.J., 2016. Antibiotics, birth mode, and
-diet shape microbiome maturation during early life. Sci. Transl. Med. 8,
-343ra82. https://doi.org/10.1126/scitranslmed.aad7121}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Bottino,
-G.Z.M.F., Klepac-Ceraj, V., 2022. ECHO RESONANCE Microbiome.
-https://doi.org/10.17605/OSF.IO/YBS32}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Kayisire,
-A., Luo, A., Klepac-Ceraj, V., 2021. Microbiome.jl and BiobakeryUtils.jl
-- Julia packages for working with microbial community data. J Open
-Source Softw 6, 3876. https://doi.org/10.21105/joss.03876}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Breiman, L., 2001.
-Random Forests. Mach. Learn. 45, 5--32.
-https://doi.org/10.1023/A:1010933404324}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Brieuc, M.S.O.,
-Waters, C.D., Drinan, D.P., Naish, K.A., 2018. A practical introduction
-to Random Forest for genetic association studies in ecology and
-evolution. Mol. Ecol. Resour. 18, 755--766.
-https://doi.org/10.1111/1755-0998.12773}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bruchhage, Muriel,
-Chen, Y., Cataldo, A.G., Müller, H.-G., Weisee, E., Wilson, S., Pietsch,
-M., D'Sa, V., Marquand, A., Madhow, S., Bouchard, K., Cole, J.H.,
-Biondo, F., Elison, J., O'Muirchheartaigh, J., Deoni, C, S., n.d.
-Longitudinal brain and cognitive development of the first 1000 Days:A
-large Multi-Cohort Multi-Scanner study. Presented at the International
-Society of Magnetic Resonance Medicine.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bush, G., Vogt, B.A.,
-Holmes, J., Dale, A.M., Greve, D., Jenike, M.A., Rosen, B.R., 2002.
-Dorsal anterior cingulate cortex: A role in reward-based decision
-making. Proc. Natl. Acad. Sci. 99, 523--528.
-https://doi.org/10.1073/pnas.012470999}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Callaghan, B., 2020.
-Nested sensitive periods: how plasticity across the microbiota-gut-brain
-axis interacts to affect the development of learning and memory. Curr.
-Opin. Behav. Sci. 36, 55--62.
-https://doi.org/10.1016/j.cobeha.2020.07.011}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cao, Z.,
-Ottino-Gonzalez, J., Cupertino, R.B., Schwab, N., Hoke, C., Catherine,
-O., Cousijn, J., Dagher, A., Foxe, J.J., Goudriaan, A.E., Hester, R.,
-Hutchison, K., Li, C.-S.R., London, E.D., Lorenzetti, V., Luijten, M.,
-Martin-Santos, R., Momenan, R., Paulus, M.P., Schmaal, L., Sinha, R.,
-Sjoerds, Z., Solowij, N., Stein, D.J., Stein, E.A., Uhlmann, A., van
-Holst, R.J., Veltman, D.J., Wiers, R.W., Yücel, M., Zhang, S.,
-Jahanshad, N., Thompson, P.M., Conrod, P., Mackey, S., Garavan, H.,
-2021. Mapping cortical and subcortical asymmetries in substance
-dependence: Findings from the ENIGMA Addiction Working Group. Addict.
-Biol. 26, e13010. https://doi.org/10.1111/adb.13010}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cerdó, T., Diéguez,
-E., Campoy, C., 2019. Early nutrition and gut microbiome:
-interrelationship between bacterial metabolism, immune system, brain
-structure, and neurodevelopment. Am. J. Physiol. Endocrinol. Metab. 317,
-E617--E630. https://doi.org/10.1152/ajpendo.00188.2019}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Chen, X., Ishwaran,
-H., 2012. Random forests for genomic data analysis. Genomics 99,
-323--329. https://doi.org/10.1016/j.ygeno.2012.04.003}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cohen Kadosh, K.,
-Krause, B., King, A.J., Near, J., Cohen Kadosh, R., 2015. Linking GABA
-and glutamate levels to cognitive skill acquisition during development.
-Hum. Brain Mapp. 36, 4334--4345. https://doi.org/10.1002/hbm.22921}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cowan, C.S.M., Dinan,
-T.G., Cryan, J.F., 2020. Annual Research Review: Critical windows - the
-microbiota-gut-brain axis in neurocognitive development. J. Child
-Psychol. Psychiatry 61, 353--371. https://doi.org/10.1111/jcpp.13156}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dalile, B., Van
-Oudenhove, L., Vervliet, B., Verbeke, K., 2019. The role of short-chain
-fatty acids in microbiota-gut-brain communication. Nat. Rev.
-Gastroenterol. Hepatol. 16, 461--478.
-https://doi.org/10.1038/s41575-019-0157-3}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Danisch, S.,
-Krumbiegel, J., 2021. Makie.jl: Flexible high-performance data
-visualization for Julia. J. Open Source Softw. 6, 3349.
-https://doi.org/10.21105/joss.03349}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Deoni, S.C.,
-Beauchemin, J., Volpe, A., Dâ Sa, V., RESONANCE Consortium, 2021. Impact
-of the COVID-19 pandemic on early child cognitive development: Initial
-findings in a longitudinal observational study of child health. MedRxiv
-Prepr. Serv. Health Sci. https://doi.org/10.1101/2021.08.10.21261846}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dominguez-Bello, M.G.,
-Costello, E.K., Contreras, M., Magris, M., Hidalgo, G., Fierer, N.,
-Knight, R., 2010. Delivery mode shapes the acquisition and structure of
-the initial microbiota across multiple body habitats in newborns. Proc.
-Natl. Acad. Sci. U. S. A. 107, 11971--11975.
-https://doi.org/10.1073/pnas.1002601107}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dong, H., Wang, B.,
-Feng, J., Yue, X., Jia, F., 2021. Correlation between serum
-concentrations of menaquinone-4 and developmental quotients in children
-with autism spectrum disorder. Front. Nutr. 8, 748513.
-https://doi.org/10.3389/fnut.2021.748513}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ernst, M., Nelson,
-E.E., Jazbec, S., McClure, E.B., Monk, C.S., Leibenluft, E., Blair, J.,
-Pine, D.S., 2005. Amygdala and nucleus accumbens in responses to receipt
-and omission of gains in adults and adolescents. NeuroImage 25,
-1279--1291. https://doi.org/10.1016/j.neuroimage.2004.12.038}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Forrest, C.B.,
-Blackwell, C.K., Camargo, C.A., 2018. Advancing the science of
-children's positive health in the national institutes of health
-environmental influences on child health outcomes (ECHO) research
-program. J. Pediatr. 196, 298--300.
-https://doi.org/10.1016/j.jpeds.2018.02.004}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Franzosa, E.A.,
-Sirota-Madi, A., Avila-Pacheco, J., Fornelos, N., Haiser, H.J., Reinker,
-S., Vatanen, T., Hall, A.B., Mallick, H., McIver, L.J., Sauk, J.S.,
-Wilson, R.G., Stevens, B.W., Scott, J.M., Pierce, K., Deik, A.A.,
-Bullock, K., Imhann, F., Porter, J.A., Zhernakova, A., Fu, J., Weersma,
-R.K., Wijmenga, C., Clish, C.B., Vlamakis, H., Huttenhower, C., Xavier,
-R.J., 2019. Gut microbiome structure and metabolic activity in
-inflammatory bowel disease. Nat. Microbiol. 4, 293--305.
-https://doi.org/10.1038/s41564-018-0306-4}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Fung, T.C., Olson,
-C.A., Hsiao, E.Y., 2017. Interactions between the microbiota, immune and
-nervous systems in health and disease. Nat. Neurosci. 20, 145--155.
-https://doi.org/10.1038/nn.4476}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ghosh, T.S., Rampelli,
-S., Jeffery, I.B., Santoro, A., Neto, M., Capri, M., Giampieri, E.,
-Jennings, A., Candela, M., Turroni, S., Zoetendal, E.G., Hermes, G.D.A.,
-Elodie, C., Meunier, N., Brugere, C.M., Pujos-Guillot, E., Berendsen,
-A.M., Groot, L.C.P.G.M.D., Feskins, E.J.M., Kaluza, J., Pietruszka, B.,
-Bielak, M.J., Comte, B., Maijo-Ferre, M., Nicoletti, C., Vos, W.M.D.,
-Fairweather-Tait, S., Cassidy, A., Brigidi, P., Franceschi, C., O'Toole,
-P.W., 2020. Mediterranean diet intervention alters the gut microbiome in
-older people reducing frailty and improving health status: the NU-AGE
-1-year dietary intervention across five European countries. Gut 69,
-1218--1228. https://doi.org/10.1136/gutjnl-2019-319654}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Gillman, M.W.,
-Blaisdell, C.J., 2018. Environmental influences on child health
-outcomes, a research program of the NIH. Curr. Opin. Pediatr. 30, 260.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Gong, Q.-Y., Cai, L.,
-Jing, Y., Wang, W., Yang, D.-X., Chen, S.-W., Tian, H.-L., 2022.
-Urolithin A alleviates blood-brain barrier disruption and attenuates
-neuronal apoptosis following traumatic brain injury in mice. Neural
-Regen. Res. 17, 2007. https://doi.org/10.4103/1673-5374.335163}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Haam, J., Yakel, J.L.,
-2017. Cholinergic modulation of the hippocampal region and memory
-function. J. Neurochem. 142, 111--121.
-https://doi.org/10.1111/jnc.14052}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Haiser, H.J.,
-Gootenberg, D.B., Chatman, K., Sirasani, G., Balskus, E.P., Turnbaugh,
-P.J., 2013. Predicting and Manipulating Cardiac Drug Inactivation by the
-Human Gut Bacterium Eggerthella lenta. Science 341, 295--298.
-https://doi.org/10.1126/science.1235872}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hartstra, A.V.,
-Schüppel, V., Imangaliyev, S., Schrantee, A., Prodan, A., Collard, D.,
-Levin, E., Dallinga-Thie, G., Ackermans, M.T., Winkelmeijer, M., Havik,
-S.R., Metwaly, A., Lagkouvardos, I., Nier, A., Bergheim, I.,
-Heikenwalder, M., Dunkel, A., Nederveen, A.J., Liebisch, G., Mancano,
-G., Claus, S.P., Benítez-Páez, A., la Fleur, S.E., Bergman, J.J.,
-Gerdes, V., Sanz, Y., Booij, J., Kemper, E., Groen, A.K., Serlie, M.J.,
-Haller, D., Nieuwdorp, M., 2020. Infusion of donor feces affects the
-gut--brain axis in humans with metabolic syndrome. Mol. Metab. 42,
-101076. https://doi.org/10.1016/j.molmet.2020.101076}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoogman, M., Bralten,
-J., Hibar, D.P., Mennes, M., Zwiers, M.P., Schweren, L., van Hulzen,
-K.J.E., Medland, S.E., Shumskaya, E., Jahanshad, N., de Zeeuw, P.,
-Szekely, E., Sudre, G., Wolfers, T., Onnink, A.M.H., Dammers, J.T.,
-Mostert, J.C., Vives-Gilabert, Y., Kohls, G., Oberwelland, E., Seitz,
-J., Schulte-Rüther, M., di Bruttopilo, S.A., Doyle, A.E., Høvik, M.F.,
-Dramsdahl, M., Tamm, L., van Erp, T.G.M., Dale, A., Schork, A.,
-Conzelmann, A., Zierhut, K., Baur, R., McCarthy, H., Yoncheva, Y.N.,
-Cubillo, A., Chantiluke, K., Mehta, M.A., Paloyelis, Y., Hohmann, S.,
-Baumeister, S., Bramati, I., Mattos, P., Tovar-Moll, F., Douglas, P.,
-Banaschewski, T., Brandeis, D., Kuntsi, J., Asherson, P., Rubia, K.,
-Kelly, C., Di Martino, A., Milham, M.P., Castellanos, F.X., Frodl, T.,
-Zentis, M., Lesch, K.-P., Reif, A., Pauli, P., Jernigan, T., Haavik, J.,
-Plessen, K.J., Lundervold, A.J., Hugdahl, K., Seidman, L.J., Biederman,
-J., Rommelse, N., Heslenfeld, D.J., Hartman, C., Hoekstra, P.J.,
-Oosterlaan, J., von Polier, G., Konrad, K., Vilarroya, O.,
-Ramos-Quiroga, J.-A., Soliva, J.C., Durston, S., Buitelaar, J.K.,
-Faraone, S.V., Shaw, P., Thompson, P., Franke, B., 2017. Subcortical
-brain volume differences of participants with ADHD across the lifespan:
-an ENIGMA collaboration. Lancet Psychiatry 4, 310--319.
-https://doi.org/10.1016/S2215-0366(17)30049-4}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoshino, Y., Yamamoto,
-T., Kaneko, M., Tachibana, R., Watanabe, M., Ono, Y., Kumashiro, H.,
-1984. Blood Serotonin and Free Tryptophan Concentration in Autistic
-Children. Neuropsychobiology 11, 22--27.
-https://doi.org/10.1159/000118045}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hosomi, K., Saito, M.,
-Park, J., Murakami, H., Shibata, N., Ando, M., Nagatake, T., Konishi,
-K., Ohno, H., Tanisawa, K., Mohsen, A., Chen, Y.-A., Kawashima, H.,
-Natsume-Kitatani, Y., Oka, Y., Shimizu, H., Furuta, M., Tojima, Y.,
-Sawane, K., Saika, A., Kondo, S., Yonejima, Y., Takeyama, H., Matsutani,
-A., Mizuguchi, K., Miyachi, M., Kunisawa, J., 2022. Oral administration
-of Blautia wexlerae ameliorates obesity and type 2 diabetes via
-metabolic remodeling of the gut microbiota. Nat. Commun. 13, 4477.
-https://doi.org/10.1038/s41467-022-32015-7}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Jenkinson, M.,
-Beckmann, C.F., Behrens, T.E.J., Woolrich, M.W., Smith, S.M., 2012. FSL.
-Neuroimage 62, 782--790.
-https://doi.org/10.1016/j.neuroimage.2011.09.015}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Jiang, H., Ling, Z.,
-Zhang, Y., Mao, H., Ma, Z., Yin, Y., Wang, W., Tang, W., Tan, Z., Shi,
-J., Li, L., Ruan, B., 2015. Altered fecal microbiota composition in
-patients with major depressive disorder. Brain. Behav. Immun. 48,
-186--194. https://doi.org/10.1016/j.bbi.2015.03.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kanehisa, M., Goto,
-S., Kawashima, S., Okuno, Y., Hattori, M., 2004. The KEGG resource for
-deciphering the genome. Nucleic Acids Res. 32, D277--D280.
-https://doi.org/10.1093/nar/gkh063}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kant, R., Rasinkangas,
-P., Satokari, R., Pietilä, T.E., Palva, A., 2015. Genome Sequence of the
-Butyrate-Producing Anaerobic Bacterium Anaerostipes hadrus PEL 85.
-Genome Announc. 3, e00224-15. https://doi.org/10.1128/genomeA.00224-15}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kim, C.-S., Cha, L.,
-Sim, M., Jung, S., Chun, W.Y., Baik, H.W., Shin, D.-M., 2021. Probiotic
-supplementation improves cognitive function and mood with changes in gut
-microbiota in Community-Dwelling older adults: A randomized,
-Double-Blind, Placebo-Controlled, multicenter trial. J Gerontol Biol Sci
-Med Sci 76, 32--40. https://doi.org/10.1093/gerona/glaa090}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Knudsen, E.I., 2004.
-Sensitive periods in the development of the brain and behavior. J. Cogn.
-Neurosci. 16, 1412--1425. https://doi.org/10.1162/0898929042304796}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Koenig, J.E., Spor,
-A., Scalfone, N., Fricker, A.D., Stombaugh, J., Knight, R., Angenent,
-L.T., Ley, R.E., 2011. Succession of microbial consortia in the
-developing infant gut microbiome. Proc. Natl. Acad. Sci. U. S. A. 108
-Suppl 1, 4578--4585. https://doi.org/10.1073/pnas.1000081107}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Coker,
-M.O., Madan, J.C., 2022. The developing microbiome from birth to 3
-years: The Gut-Brain axis and neurodevelopmental outcomes. Front.
-Pediatr. 10, 815885. https://doi.org/10.3389/fped.2022.815885}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Korrick,
-S.A., Baker, E.R., Karagas, M.R., Madan, J.C., 2020. Prospective
-associations of the infant gut microbiome and microbial function with
-social behaviors related to autism at age 3 years. Sci. Rep. 10, 15515.
-https://doi.org/10.1038/s41598-020-72386-9}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ley, R.E., Turnbaugh,
-P.J., Klein, S., Gordon, J.I., 2006. Human gut microbes associated with
-obesity. Nature 444, 1022--1023. https://doi.org/10.1038/4441022a}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Li, Y.-M., Yu, Q.,
-Huang, L., Fu, J., Feng, R.-J., 2022. {[}Correlations Between Gut
-Microbiota Changes and Cognitive Function in Patients with Post-Stroke
-Cognitive Impairment in the Early Stage{]}. Sichuan Da Xue Xue Bao Yi
-Xue Ban 53, 857--865. https://doi.org/10.12182/20220960105}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, S., Li, E., Sun,
-Z., Fu, D., Duan, G., Jiang, M., Yu, Y., Mei, L., Yang, P., Tang, Y.,
-Zheng, P., 2019. Altered gut microbiota and short chain fatty acids in
-Chinese children with autism spectrum disorder. Sci. Rep. 9, 287.
-https://doi.org/10.1038/s41598-018-36430-z}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, X., Mao, B., Gu,
-J., Wu, J., Cui, S., Wang, G., Zhao, J., Zhang, H., Chen, W., 2021.
-\emph{Blautia} ---a new functional genus with potential probiotic
-properties? Gut Microbes 13, 1875796.
-https://doi.org/10.1080/19490976.2021.1875796}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Lopez-Siles, M.,
-Duncan, S.H., Garcia-Gil, L.J., Martinez-Medina, M., 2017.
-Faecalibacterium prausnitzii: from microbiology to diagnostics and
-prognostics. ISME J. 11, 841--852.
-https://doi.org/10.1038/ismej.2016.176}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Louwies, T., Johnson,
-A.C., Orock, A., Yuan, T., Greenwood-Van Meerveld, B., 2020. The
-microbiota-gut-brain axis: An emerging role for the epigenome. Exp.
-Biol. Med. 245, 138--145. https://doi.org/10.1177/1535370219891690}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Magnusson, K.R.,
-Hauck, L., Jeffrey, B.M., Elias, V., Humphrey, A., Nath, R., Perrone,
-A., Bermudez, L.E., 2015. Relationships between diet-related changes in
-the gut microbiome and cognitive flexibility. Neuroscience 300,
-128--140. https://doi.org/10.1016/j.neuroscience.2015.05.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mallick, H.,
-Rahnavard, A., McIver, L.J., Ma, S., Zhang, Y., Nguyen, L.H., Tickle,
-T.L., Weingart, G., Ren, B., Schwager, E.H., Chatterjee, S., Thompson,
-K.N., Wilkinson, J.E., Subramanian, A., Lu, Y., Waldron, L., Paulson,
-J.N., Franzosa, E.A., Bravo, H.C., Huttenhower, C., 2021. Multivariable
-association discovery in population-scale meta-omics studies.
-https://doi.org/10.1101/2021.01.20.427420}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Maruo, T., Sakamoto,
-M., Ito, C., Toda, T., Benno, Y., 2008. Adlercreutzia equolifaciens gen.
-nov., sp. nov., an equol-producing bacterium isolated from human faeces,
-and emended description of the genus Eggerthella. Int. J. Syst. Evol.
-Microbiol. 58, 1221--1227. https://doi.org/10.1099/ijs.0.65404-0}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mayneris-Perxachs, J.,
-Castells-Nobau, A., Arnoriaga-Rodríguez, M., Martin, M., de la
-Vega-Correa, L., Zapata, C., Burokas, A., Blasco, G., Coll, C.,
-Escrichs, A., Biarnés, C., Moreno-Navarrete, J.M., Puig, J., Garre-Olmo,
-J., Ramos, R., Pedraza, S., Brugada, R., Vilanova, J.C., Serena, J.,
-Gich, J., Ramió-Torrentà, L., Pérez-Brocal, V., Moya, A., Pamplona, R.,
-Sol, J., Jové, M., Ricart, W., Portero-Otin, M., Deco, G., Maldonado,
-R., Fernández-Real, J.M., 2022. Microbiota alterations in proline
-metabolism impact depression. Cell Metab. 34, 681-701.e10.
-https://doi.org/10.1016/j.cmet.2022.04.001}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mukherjee, A., Lordan,
-C., Ross, R.P., Cotter, P.D., 2020. Gut microbes from the
-phylogenetically diverse genus Eubacterium and their various
-contributions to gut health. Gut Microbes 12, 1802866.
-https://doi.org/10.1080/19490976.2020.1802866}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mullen, E.M., others,
-1995. Mullen scales of early learning. AGS Circle Pines, MN.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Needham, B.D.,
-Funabashi, M., Adame, M.D., Wang, Z., Boktor, J.C., Haney, J., Wu,
-W.-L., Rabut, C., Ladinsky, M.S., Hwang, S.-J., Guo, Y., Zhu, Q.,
-Griffiths, J.A., Knight, R., Bjorkman, P.J., Shapiro, M.G., Geschwind,
-D.H., Holschneider, D.P., Fischbach, M.A., Mazmanian, S.K., 2022. A
-gut-derived metabolite alters brain activity and anxiety behaviour in
-mice. Nature 602, 647--653. https://doi.org/10.1038/s41586-022-04396-8}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Opel, N., Thalamuthu,
-A., Milaneschi, Y., Grotegerd, D., Flint, C., Leenings, R., Goltermann,
-J., Richter, M., Hahn, T., Woditsch, G., Berger, K., Hermesdorf, M.,
-McIntosh, A., Whalley, H.C., Harris, M.A., MacMaster, F.P., Walter, H.,
-Veer, I.M., Frodl, T., Carballedo, A., Krug, A., Nenadic, I., Kircher,
-T., Aleman, A., Groenewold, N.A., Stein, D.J., Soares, J.C.,
-Zunta-Soares, G.B., Mwangi, B., Wu, M.-J., Walter, M., Li, M., Harrison,
-B.J., Davey, C.G., Cullen, K.R., Klimes-Dougan, B., Mueller, B.A.,
-Sämann, P.G., Penninx, B., Nawijn, L., Veltman, D.J., Aftanas, L., Brak,
-I.V., Filimonova, E.A., Osipov, E.A., Reneman, L., Schrantee, A., Grabe,
-H.J., Van der Auwera, S., Wittfeld, K., Hosten, N., Völzke, H., Sim, K.,
-Gotlib, I.H., Sacchet, M.D., Lagopoulos, J., Hatton, S.N., Hickie, I.,
-Pozzi, E., Thompson, P.M., Jahanshad, N., Schmaal, L., Baune, B.T.,
-Dannlowski, U., 2021. Brain structural abnormalities in obesity:
-relation to age, genetic risk, and common psychiatric disorders. Mol.
-Psychiatry 26, 4839--4852. https://doi.org/10.1038/s41380-020-0774-9}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Pagonabarraga, J.,
-Corcuera-Solano, I., Vives-Gilabert, Y., Llebaria, G., García-Sánchez,
-C., Pascual-Sedano, B., Delfino, M., Kulisevsky, J., Gómez-Ansón, B.,
-2013. Pattern of Regional Cortical Thinning Associated with Cognitive
-Deterioration in Parkinson's Disease. PLoS ONE 8, e54980.
-https://doi.org/10.1371/journal.pone.0054980}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Palomo-Buitrago, M.E.,
-Sabater-Masdeu, M., Moreno-Navarrete, J.M., Caballano-Infantes, E.,
-Arnoriaga-Rodríguez, M., Coll, C., Ramió, L., Palomino-Schätzlein, M.,
-Gutiérrez-Carcedo, P., Pérez-Brocal, V., Simó, R., Moya, A., Ricart, W.,
-Herance, J.R., Fernández-Real, J.M., 2019. Glutamate interactions with
-obesity, insulin resistance, cognition and gut microbiota composition.
-Acta Diabetol. 56, 569--579. https://doi.org/10.1007/s00592-019-01313-w}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Pronovost, G.N.,
-Hsiao, E.Y., 2019. Perinatal interactions between the microbiome,
-immunity, and neurodevelopment. Immunity 50, 18--36.
-https://doi.org/10.1016/j.immuni.2018.11.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sadeghi, B.,
-Chiarawongse, P., Squire, K., Jones, D.C., Noack, A., St-Jean, C.,
-Huijzer, R., Schätzle, R., Butterworth, I., Peng, Y.-F., Blaom, A.,
-2022. DecisionTree.jl - A julia implementation of the CART decision tree
-and random forest algorithms. https://doi.org/10.5281/zenodo.7359268}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sela, D.A., Chapman,
-J., Adeuya, A., Kim, J.H., Chen, F., Whitehead, T.R., Lapidus, A.,
-Rokhsar, D.S., Lebrilla, C.B., German, J.B., Price, N.P., Richardson,
-P.M., Mills, D.A., 2008. The genome sequence of \emph{Bifidobacterium
-longum} subsp. \emph{infantis} reveals adaptations for milk utilization
-within the infant microbiome. Proc. Natl. Acad. Sci. 105, 18964--18969.
-https://doi.org/10.1073/pnas.0809584105}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Selma, M.V., Beltrán,
-D., García-Villalba, R., Espín, J.C., Tomás-Barberán, F.A., 2014.
-Description of urolithin production capacity from ellagic acid of two
-human intestinal Gordonibacter species. Food Funct. 5, 1779--1784.
-https://doi.org/10.1039/c4fo00092g}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Setchell, K.D.,
-Clerici, C., Lephart, E.D., Cole, S.J., Heenan, C., Castellani, D.,
-Wolfe, B.E., Nechemias-Zimmer, L., Brown, N.M., Lund, T.D., Handa, R.J.,
-Heubi, J.E., 2005. S-Equol, a potent ligand for estrogen receptor β, is
-the exclusive enantiomeric form of the soy isoflavone metabolite
-produced by human intestinal bacterial flora. Am. J. Clin. Nutr. 81,
-1072--1079. https://doi.org/10.1093/ajcn/81.5.1072}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sharon, G., Sampson,
-T.R., Geschwind, D.H., Mazmanian, S.K., 2016. The central nervous system
-and the gut microbiome. Cell 167, 915--932.
-https://doi.org/10.1016/j.cell.2016.10.027}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Silbereis, J.C.,
-Pochareddy, S., Zhu, Y., Li, M., Sestan, N., 2016. The cellular and
-molecular landscapes of the developing human central nervous system.
-Neuron 89, 248--268.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sordillo, J.E.,
-Korrick, S., Laranjo, N., Carey, V., Weinstock, G.M., Gold, D.R.,
-O'Connor, G., Sandel, M., Bacharier, L.B., Beigelman, A., Zeiger, R.,
-Litonjua, A.A., Weiss, S.T., 2019. Association of the infant gut
-microbiome with early childhood neurodevelopmental outcomes: An
-ancillary study to the VDAART randomized clinical trial. JAMA Netw. Open
-2, e190905. https://doi.org/10.1001/jamanetworkopen.2019.0905}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Spichak, S.,
-Bastiaanssen, T.F.S., Berding, K., Vlckova, K., Clarke, G., Dinan, T.G.,
-Cryan, J.F., 2021. Mining microbes for mental health: Determining the
-role of microbial metabolic pathways in human brain health and disease.
-Neurosci. Biobehav. Rev. 125, 698--761.
-https://doi.org/10.1016/j.neubiorev.2021.02.044}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Stephan, J., Stegle,
-O., Beyer, A., 2015. A random forest approach to capture genetic effects
-in the presence of population structure. Nat. Commun. 6, 7432.
-https://doi.org/10.1038/ncomms8432}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Suzek, B.E., Huang,
-H., McGarvey, P., Mazumder, R., Wu, C.H., 2007. UniRef: comprehensive
-and non-redundant UniProt reference clusters. Bioinformatics 23,
-1282--1288. https://doi.org/10.1093/bioinformatics/btm098}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Takahashi, H., Yang,
-J., Yamamoto, H., Fukuda, S., Arakawa, K., 2021. Complete Genome
-Sequence of Adlercreutzia equolifaciens subsp. celatus DSM 18785.
-Microbiol. Resour. Announc. 10, e00354-21.
-https://doi.org/10.1128/MRA.00354-21}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tau, G.Z., Peterson,
-B.S., 2010. Normal development of brain circuits. Neuropsychopharmacol.
-Off. Publ. Am. Coll. Neuropsychopharmacol. 35, 147--168.
-https://doi.org/10.1038/npp.2009.115}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Thawornkuno, C.,
-Tanaka, M., Sone, T., Asano, K., 2009. Biotransformation of Daidzein to
-Equol by Crude Enzyme from Asaccharobacter celatus AHU1763 Required an
-Anaerobic Environment. Biosci. Biotechnol. Biochem. 73, 1435--1438.
-https://doi.org/10.1271/bbb.80908}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tognini, P., 2017. Gut
-microbiota: a potential regulator of neurodevelopment. Front. Cell.
-Neurosci. 11, 25. https://doi.org/10.3389/fncel.2017.00025}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tso, L., Bonham, K.S.,
-Fishbein, A., Rowland, S., Klepac-Ceraj, V., 2021. Targeted
-High-Resolution taxonomic identification of bifidobacterium longum
-subsp. infantis using human milk oligosaccharide metabolizing genes.
-Nutrients 13, 2833. https://doi.org/10.3390/nu13082833}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Valles-Colomer, M.,
-Falony, G., Darzi, Y., Tigchelaar, E.F., Wang, J., Tito, R.Y., Schiweck,
-C., Kurilshikov, A., Joossens, M., Wijmenga, C., Claes, S., Van
-Oudenhove, L., Zhernakova, A., Vieira-Silva, S., Raes, J., 2019. The
-neuroactive potential of the human gut microbiota in quality of life and
-depression. Nat. Microbiol. 4, 623--632.
-https://doi.org/10.1038/s41564-018-0337-x}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Velazquez, M., Lee,
-Y., Initiative, for the A.D.N., 2021. Random forest model for
-feature-based Alzheimer's disease conversion prediction from early mild
-cognitive impairment subjects. PLOS ONE 16, e0244773.
-https://doi.org/10.1371/journal.pone.0244773}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wacker, J., Dillon,
-D.G., Pizzagalli, D.A., 2009. The role of the nucleus accumbens and
-rostral anterior cingulate cortex in anhedonia: Integration of resting
-EEG, fMRI, and volumetric techniques. NeuroImage 46, 327--337.
-https://doi.org/10.1016/j.neuroimage.2009.01.058}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wan, Y., Zuo, T., Xu,
-Z., Zhang, F., Zhan, H., Chan, D., Leung, T.-F., Yeoh, Y.K., Chan,
-F.K.L., Chan, R., Ng, S.C., 2021. Underdevelopment of the gut microbiota
-and bacteria species as non-invasive markers of prediction in children
-with autism spectrum disorder. Gut gutjnl-2020-324015.
-https://doi.org/10.1136/gutjnl-2020-324015}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, L.-J., Yang,
-C.-Y., Chou, W.-J., Lee, M.-J., Chou, M.-C., Kuo, H.-C., Yeh, Y.-M.,
-Lee, S.-Y., Huang, L.-H., Li, S.-C., 2020. Gut microbiota and dietary
-patterns in children with attention-deficit/hyperactivity disorder. Eur.
-Child Adolesc. Psychiatry 29, 287--297.
-https://doi.org/10.1007/s00787-019-01352-2}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, X.-L., Hur,
-H.-G., Lee, J.H., Kim, K.T., Kim, S.-I., 2005. Enantioselective
-Synthesis of S-Equol from Dihydrodaidzein by a Newly Isolated Anaerobic
-Human Intestinal Bacterium. Appl. Environ. Microbiol. 71, 214--219.
-https://doi.org/10.1128/AEM.71.1.214-219.2005}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 2012.
-Wechsler preschool and primary scale of intelligence---fourth edition.
-Psychol. Corp. San Antonio TX.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 1949.
-Wechsler intelligence scale for children.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, L., Yan, J.,
-Yang, T., Zhu, J., Li, T., Wei, H., Chen, J., 2021. Fecal Microbiome
-Transplantation from Children with Autism Spectrum Disorder Modulates
-Tryptophan and Serotonergic Synapse Metabolism and Induces Altered
-Behaviors in Germ-Free Mice. mSystems 6, e01343-20.
-https://doi.org/10.1128/mSystems.01343-20}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, X., Yang, Y.,
-Zhang, Y., Zhang, X.-M., Zhao, Z.-Q., Zhang, Y.-Q., 2013. Estrogen in
-the Anterior Cingulate Cortex Contributes to Pain-Related Aversion.
-Cereb. Cortex 23, 2190--2203. https://doi.org/10.1093/cercor/bhs201}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Yael, D., Tahary, O.,
-Gurovich, B., Belelovsky, K., Bar-Gad, I., 2019. Disinhibition of the
-Nucleus Accumbens Leads to Macro-Scale Hyperactivity Consisting of
-Micro-Scale Behavioral Segments Encoded by Striatal Activity. J.
-Neurosci. 39, 5897--5909.
-https://doi.org/10.1523/JNEUROSCI.3120-18.2019}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Yau, W.-Y.W., Zubieta,
-J.-K., Weiland, B.J., Samudra, P.G., Zucker, R.A., Heitzeg, M.M., 2012.
-Nucleus Accumbens Response to Incentive Stimuli Anticipation in Children
-of Alcoholics: Relationships with Precursive Behavioral Risk and
-Lifetime Alcohol Use. J. Neurosci. 32, 2544--2551.
-https://doi.org/10.1523/JNEUROSCI.1390-11.2012}
-
-Supplementary Tables
-
-\textbf{Supplementary Table XXA. List of most important taxa for Random
-Forest models on the 0 to 6 months age bracket, microbiome +
-demographics, ordered by relative weighted importance rank}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Taxon
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Rank
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Relative weighted importance
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Cumulative
-
-relative weighted importance
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\textbf{Supplementary Table XXA. List of most important taxa for Random
-Forest models on the 18 to 120 months age bracket, microbiome +
-demographics, ordered by relative weighted importance rank}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Taxon
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Rank
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Relative weighted importance
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Cumulative
-
-relative weighted importance
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\textbf{Supplementary Table XX. Benchmark metrics for each brain segment
-prediction experiment. Confidence intervals are calculated from the
-distribution of metrics from repeated CV at a confidence level of 95\%}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3933}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2440}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1846}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1782}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Brain segment
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Root-mean-square Error
-
-(RMSE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Mean Absolute Proportional Error (MAPE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Correlation coefficient (R)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Brain-stem
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0876
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.605
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-I-V
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0321
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.085
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.156
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-VI-VII
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0132
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.081
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.006
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-VIII-X
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0217
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.078
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-CSF
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0193
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.162
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.071
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-accumbens-area
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0020
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.094
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.288
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-accumbens-area
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0020
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.102
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.041
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-amygdala
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0037
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.288
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-amygdala
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.070
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.179
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-anterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0375
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.069
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.092
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-anterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0264
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.074
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.245
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-basal-forebrain
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0008
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-basal-forebrain
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0010
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.117
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.123
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-caudate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0257
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.099
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.029
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-caudate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0221
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.010
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-cerebellum-white-matter
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0689
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.413
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-cerebellum-white-matter
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0694
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.070
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.434
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-cuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0274
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.234
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-cuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0264
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.075
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.065
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-entorhinal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0107
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.080
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.190
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-entorhinal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0105
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.099
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-fusiform
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0510
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.333
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-fusiform
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0334
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.053
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.248
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-hippocampus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0128
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.054
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.075
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-hippocampus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0150
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.094
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-insula
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0294
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.119
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-insula
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0303
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-isthmus-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0161
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.041
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-isthmus-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0166
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.085
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.074
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-lateral-occipital
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0661
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.062
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.224
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-lateral-occipital
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0716
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.054
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.221
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-lingual
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0410
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.069
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.421
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-lingual
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0440
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.072
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.434
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-middle-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0875
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.052
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.000
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-middle-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0636
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.039
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-orbitofrontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0715
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.055
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.186
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-orbitofrontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0814
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.133
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pallidum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0075
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.064
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.297
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pallidum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0081
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.264
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-paracentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0335
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.078
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.131
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-paracentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0364
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.084
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-parahippocampal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0098
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.058
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-parahippocampal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0092
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.060
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-parietal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0755
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.177
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-parietal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1099
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.045
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.147
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-opercularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0197
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.146
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-opercularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0261
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.060
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.030
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-orbitalis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0157
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.003
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-orbitalis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0233
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.112
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.024
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-triangularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0226
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.174
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-triangularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0322
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.076
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.043
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pericalcarine
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0178
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.111
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.200
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pericalcarine
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0185
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.089
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.273
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-postcentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0495
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.052
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.168
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-postcentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0552
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.135
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-posterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0460
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.137
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-posterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0480
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.139
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-precentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0830
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.065
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.248
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-precentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0594
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.046
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.066
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-precuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0489
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.144
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-precuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0622
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.150
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-putamen
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0243
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-putamen
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0242
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.066
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.011
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-superior-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0867
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.040
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.101
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-superior-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.044
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.096
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-supramarginal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0767
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-supramarginal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0734
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.001
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-temporal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1530
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.040
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.128
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-temporal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1503
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.039
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-thalamus-proper
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0347
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.053
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.170
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-thalamus-proper
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0304
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.050
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.162
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
+\input{sections/methods}
 

### 2023-02-17 manuscript/tex/sections/discussion.tex
clean up

--- 2023-02-16 manuscript/tex/converted.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-]{article}
-\usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\usepackage{longtable,booktabs,array}
-\usepackage{multirow}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\usepackage[normalem]{ulem}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-
-\author{}
-\date{}
-
-\begin{document}
-
-\hypertarget{title}{%
-\subsection{Title}\label{title}}
-
-\begin{itemize}
-\item
-  \begin{quote}
-  Gut microbes and their genes are associated with cognition and
-  neuroanatomy in children
-  \end{quote}
-\item
-  \begin{quote}
-  Gut microbes and their genes are associated with cognition performance
-  and brain structure in children
-  \end{quote}
-\item
-  \begin{quote}
-  Extensive relationships between the gut microbiota and normal
-  neurocognitive development
-  \end{quote}
-\item
-  \begin{quote}
-  Extensive relationships between the gut microbiota and typical
-  neurocognitive development
-  \end{quote}
-\item
-  \begin{quote}
-  The gut microbial metagenome is connected to neurocognition in typical
-  child development
-  \end{quote}
-\item
-  \begin{quote}
-  The microbial potential in regulating early childhood neurocognitive
-  development
-  \end{quote}
-\end{itemize}
-
-\hypertarget{authors}{%
-\subsubsection{Authors}\label{authors}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \begin{quote}
-  Bonham 0000-0003-3200-7533
-  \end{quote}
-\item
-  \begin{quote}
-  Bottino 0000-0003-1953-1576
-  \end{quote}
-\item
-  \begin{quote}
-  McCann 0000-0002-9753-7968
-  \end{quote}
-\item
-  \begin{quote}
-  Beauchemin \{ORCID??\}
-  \end{quote}
-\item
-  \begin{quote}
-  Elizabeth Weisse \textless elizabeth.weisse2@gmail.com\textgreater{}
-  \end{quote}
-\item
-  \begin{quote}
-  \ldots{}
-  \end{quote}
-\item
-  \begin{quote}
-  Jonathan O\textquotesingle Muircheartaigh
-  \textless{}\href{mailto:jonathanom@kcl.ac.uk}{\uline{jonathanom@kcl.ac.uk}}\textgreater{}
-  \end{quote}
-\item
-  \begin{quote}
-  Huttenhower 0000-0002-1110-0096
-  \end{quote}
-\item
-  \begin{quote}
-  Muriel Bruchhage
-  \textless{}\href{mailto:mmk.bruchhage@gmail.com}{\uline{mmk.bruchhage@gmail.com}}\textgreater{}
-  0000-0001-9637-0951
-  \end{quote}
-\item
-  \begin{quote}
-  D'Sa 0000-0002-3476-8419
-  \end{quote}
-\item
-  \begin{quote}
-  Deoni 0000-0002-0633-6328
-  \end{quote}
-\item
-  \begin{quote}
-  Klepac-Ceraj 0000-0001-5387-5706
-  \end{quote}
-\end{enumerate}
-
-\hypertarget{abstract}{%
-\subsection{Abstract}\label{abstract}}
-
-The gastrointestinal tract, its resident microorganisms and the central
-nervous system are connected by biochemical signaling, also known as
-"microbiome-gut-brain-axis." Both the human brain and the gut microbiome
-have critical developmental windows in the first three years of life,
-raising the possibility that their development is co-occurring and
-likely co-dependent. Emerging evidence implicates gut microorganisms and
-microbiota composition in cognitive outcomes and neurodevelopmental
-disorders (e.g., autism and anxiety), but the influence of gut microbial
-metabolism on typical neurodevelopment has not been explored in detail.
-We investigated the relationship of the microbiome with the neuroanatomy
-and cognitive function of 361 healthy children, demonstrating that
-differences in gut microbial taxa and gene functions are associated with
-overall cognitive function as well as with differences in the size of
-multiple brain regions. Using a combination of multivariate linear
-models and machine learning (ML) models, we showed that many species,
-including \emph{Gordonibacter pamelae} and \emph{Blautia wexlerae}, were
-significantly associated with higher cognitive function, while some
-species such as \emph{Ruminococcus gnavus} were more commonly found in
-children with low cognitive scores regardless of age or maternal
-education. Microbial genes for enzymes involved in the metabolism of
-neuroactive compounds, particularly short-chain fatty acids such as
-glutamate and propionate, were also associated with cognitive function.
-In addition, ML models were able to use microbial taxa to predict the
-volume of brain regions, with particular taxa often dominating the
-feature importance metric for individual brain regions, such as \emph{B.
-wexlerae} being the most dominant for the parahippocampal region or
-GABA-producing \emph{Bacteroidetes ovatus} for left accumbens. These
-findings provide potential biomarkers of neurocognition and may lead to
-the future development of targets for early detection and early
-intervention.
-
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
-
-The gut and the brain are intimately linked. Signals from the brain
-reach the gut through the autonomic nervous system and the endocrine
-system, and the gut can communicate with the brain through the vagus
-nerve and through endocrine and immune (cytokine) signaling molecules
-\href{https://www.zotero.org/google-docs/?i3SRPn}{(Cerdó et al., 2019;
-Pronovost and Hsiao, 2019; Sharon et al., 2016; Tognini, 2017)}. In
-addition, the products of microbial metabolism generated in the gut can
-influence the brain, both indirectly by stimulating the enteric nervous
-and immune systems, as well as directly through molecules that enter
-circulation and cross the blood-brain barrier. Causal links between the
-gut microbiome and neural development, particularly atypical
-development, are increasingly being identified
-\href{https://www.zotero.org/google-docs/?UsrFKB}{(Spichak et al.,
-2021)}. Both human epidemiology and animal models point to the effects
-of gut microbes on the development of autism spectrum disorder
-\href{https://www.zotero.org/google-docs/?Y8l5A0}{(Laue et al., 2020;
-Wan et al., 2021)} and particular microbial taxa have been associated
-with depression
-\href{https://www.zotero.org/google-docs/?mbaFE3}{(Mayneris-Perxachs et
-al., 2022; Valles-Colomer et al., 2019)} and Alzheimer's disease
-\href{https://www.zotero.org/google-docs/?0BEGUn}{(Fung et al., 2017;
-Kim et al., 2021)}. But information about this ``microbiome-gut-brain
-axis'' in normal neurocognitive development remains lacking,
-particularly early in life.
-
-The first years of life are critical developmental windows for both the
-microbiome and the brain
-\href{https://www.zotero.org/google-docs/?XTqhlm}{(Laue et al., 2022)}.
-Fetal development \emph{in utero} is believed to be mostly sterile, but
-is rapidly seeded at birth through contact with the birth canal (if
-birthed vaginally), caregivers, food sources (breastmilk or formula),
-and other environmental sources such as antibiotics
-\href{https://www.zotero.org/google-docs/?X6N0f4}{(Bäckhed et al., 2015;
-Bokulich et al., 2016; Dominguez-Bello et al., 2010; Louwies et al.,
-2020)}. The early microbiome is characterized by low microbial
-diversity, rapid succession and evolution, and domination by
-Actinobacteria, particularly the genus \emph{Bifidobacterium},
-Bacteroidetes, especially \emph{Bacteroides}, and Proteobacteria
-\href{https://www.zotero.org/google-docs/?VG43uh}{(Koenig et al.,
-2011)}. Many of these microbes have specialized metabolic capabilities
-for digesting human breast milk, such as \emph{Bifidobacterium infantis}
-and \emph{Bacteroides fragilis}
-\href{https://www.zotero.org/google-docs/?1dKzKj}{(Sela et al., 2008;
-Tso et al., 2021)}. Upon the introduction of solid foods, the gut
-microbiome undergoes another categorical transformation, with most taxa
-of the infant microbiome being replaced by taxa more reminiscent of
-adult microbiomes
-\href{https://www.zotero.org/google-docs/?V3lY3p}{(Bäckhed et al.,
-2015)}. Many prior studies have focused on either infant microbiomes or
-adult microbiomes, since performing statistical analyses across this
-transition poses particular challenges. Nevertheless, since this
-transition coincides with critical neural developmental windows and
-neural synaptogenesis, investigation across this solid-food boundary is
-incredibly important
-\href{https://www.zotero.org/google-docs/?d29bv8}{(Tau and Peterson,
-2010)}.
-
-A child's brain undergoes remarkable anatomical, microstructural,
-organizational, and functional changes in the first years of life. By
-age 5, a child's brain has reached \textgreater85\% of its adult size,
-has achieved near-adult levels of myelination, and the pattern of axonal
-connections has been established
-\href{https://www.zotero.org/google-docs/?1jUZOO}{(Silbereis et al.,
-2016)}. Much of this development occurs in discrete windows or
-``sensitive-'' or ``critical periods'' (CPs)
-\href{https://www.zotero.org/google-docs/?w4Pcze}{(Knudsen, 2004)} when
-neural plasticity is particularly high, and particular modes of learning
-and skill development are preferred. The timing of these sensitive
-periods is driven in part by genetics, but can also be affected by the
-environment, including the gut microbiome
-\href{https://www.zotero.org/google-docs/?KTm9lQ}{(Cowan et al., 2020)}.
-In fact, emerging evidence suggests that the timing and duration of CPs
-may be driven in part by cues from the developing gut microbiome
-\href{https://www.zotero.org/google-docs/?mAD1ne}{(Callaghan, 2020)}. As
-such, understanding the normal spectrum of healthy microbiome
-development and how it relates to normal neurocognitive development may
-provide opportunities for identifying atypical development earlier and
-offer opportunities for intervention.
-
-To begin to address this need, we investigated the gut microbiome and
-neurocognitive development of children from infancy through 10 years of
-age in an accelerated longitudinal study design. Gut microbial
-communities were assessed using shotgun metagenomic sequencing, enabling
-profiling at both the taxonomic and gene-functional level, and
-neurocognitive development was measured using expert-assessment of
-cognitive function and neuroimaging using magnetic resonance imaging
-(MRI). Using a combination of classical statistical analysis and machine
-learning, we showed that the development of the gut microbiome, the
-brain, and children's cognitive abilities are intimately linked, with
-both microbial taxa and gene functions able to predict cognitive
-performance and brain structure.
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-\hypertarget{cohort-overview-and-summary-data}{%
-\subsubsection{Cohort overview and summary
-data}\label{cohort-overview-and-summary-data}}
-
-We investigated the co-development of the brain and the microbiome in
-early childhood in a cohort of typically-developing children in their
-first years of life using a variety of orthogonal microbial and
-neurocognitive assessments, including shotgun metagenomic sequencing,
-cognitive and behavioral assessments, and neuroimaging. In all, 361
-children from RESONANCE, an accelerated longitudinal cohort of early
-development \href{https://paperpile.com/c/dPbU4e/klxU}{(Forrest,
-Blackwell, and Camargo 2018)} between the ages of 82 days and 10 years
-old (median age 2.21 years, Supplementary Figure 1) were included in
-this study (Table 1, Figure 1A). To measure cognitive function, we used
-age-appropriate assessments that can be normalized to a common, IQ-like
-scale (Figure 1B), including the Mullen scale of early learning (MSEL)
-for children from birth to 3 years of age
-\href{https://www.zotero.org/google-docs/?90hfYr}{(Mullen and others,
-1995)}, Wechsler preschool and primary scale of intelligence (WPPSI) for
-4--5-year-olds
-\href{https://www.zotero.org/google-docs/?wCfx5h}{(Wechsler, 2012)}, and
-the Wechsler intelligence scale for children (WISC) for children 6 years
-and up \href{https://www.zotero.org/google-docs/?AcNN3S}{(Wechsler,
-1949)}.
-
-As expected, the greatest differences in microbial taxa were driven by
-age (Figure 1B-C, PCoA axis 1), while older children were primarily
-stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
-high-\emph{Prevotella copri} (Supplementary Figure 2). Overall variation
-in gut microbial genes and also in brain volume profiles was similarly
-driven largely by subject age (Figure 1C-F).
-
-\includegraphics[width=6.45833in,height=3.87221in]{media/image3.png}
-
-\textbf{Figure 1: Cohort overview and summary data.}
-
-(A) Study design and overview. Stool samples, cognitive assessments and
-neuroimaging were collected from participants in an accelerated
-longitudinal design. (B) Cognitive function scores are assessed by
-different instruments, but may be normalized to a common IQ-like scale.
-(C-D) Principal coordinates analysis using Bray-Curtis dissimilarity on
-taxonomic profiles demonstrates high beta diversity, with much of the
-first axis of variation explained by increasing age and alpha diversity.
-Differences in gene function profiles (E) and neuroimaging (PCA based on
-the Euclidean distance of brain region volumes) (F) are likewise
-dominated by changes as children age. (G) Permutation analysis of
-variance (PERMANOVA) of taxonomic profiles, functional profiles
-(annotated as UniRef90s) and neuroimaging, against the metadata of
-interest. Variations in taxonomic and functional profiles explain a
-modest but significant percent of the variation in cognitive development
-in children over 18 months of age. (H) Mantel testing of different
-microbial feature matrices, shows overlapping but distinct patterns of
-variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
-used as cut-offs in some following models.
-
-\textbf{Table1: Subject demographics}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2620}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Group}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Subgroup}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{All}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Under 6 mo}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Over 18 mo}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-N subjects
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-361
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-74
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-257
-\end{minipage} \\
-\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Samples
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-1
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-262 (72.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-74 (100.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-203 (79.0\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-72 (19.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-46 (17.9\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-\textgreater2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-27 (7.5\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-8 (3.1\%)
-\end{minipage} \\
-\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Age (months)
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-Min
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18.0
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Max
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-119.3
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6.0
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-119.3
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Median
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-25.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-3.7
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-45.9
-\end{minipage} \\
-\multirow{2}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Sex
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-F
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-163 (45.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (50.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-114 (44.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-M
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-198 (54.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (50.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-143 (55.6\%)
-\end{minipage} \\
-\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Race
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-White
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-252 (69.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-42 (56.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-191 (74.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Black
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-40 (11.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16 (21.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-21 (8.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Asian
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Mixed
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-7 (1.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (1.4\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6 (2.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Other
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-48 (13.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12 (16.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-29 (11.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Unknown/Declined
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2 (0.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (1.4\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2 (0.8\%)
-\end{minipage} \\
-\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Maternal Ed.
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-Junior high school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Some high school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6 (1.7\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-3 (4.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-4 (1.6\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-High school grad
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (10.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-13 (17.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16 (6.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Some college
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-95 (26.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-25 (33.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-57 (22.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-College grad
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-90 (24.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17 (23.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-69 (26.8\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Grad/professional school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-123 (34.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-15 (20.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-104 (40.5\%)
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-Several studies have demonstrated links between specific taxa and
-measures of anxiety and depression
-\href{https://www.zotero.org/google-docs/?X5rfKj}{(Mayneris-Perxachs et
-al., 2022; Needham et al., 2022)}, cognitive flexibility in adults
-\href{https://www.zotero.org/google-docs/?RGIf8m}{(Magnusson et al.,
-2015)}, and atypical neural development
-\href{https://www.zotero.org/google-docs/?hwPZKh}{(Liu et al., 2019; Wan
-et al., 2021)}. We, therefore, set out to identify whether specific taxa
-or gene functions were linked with normal cognitive development in
-children. In order to assess whether variations in gut microbial taxa,
-their genes, or their metabolism are linked with neurocognitive
-development, we tested whether the beta diversity of microbial taxa and
-gene functions, as well as variation in brain development as assessed by
-neuroimaging, were associated with these contemporaneous measures of
-cognitive function using permutation analysis of variance (PERMANOVA).
-Due to the large ecological shift in the microbiome that occurs upon the
-introduction of solid food, and a relatively wide range of ages when
-infants are transitioned to solid food, we considered the pre-transition
-(less than 6 months old) and post-transition (over 18 months old)
-microbiomes separately. Even still, age was a major driver of variation
-in gut microbiomes in children over 18 months old (Figure 1G). We also
-found that overall variation in microbial species in children over 18
-months old was associated with a small but significant variation in
-cognitive function score (Figure 1G, R\textsuperscript{2} = 0.0124, q
-\textless{} 0.001), as was variation in microbial gene functions
-(R\textsuperscript{2} = 0.0119, q \textless{} 0.001). Variation in
-microbial taxa and genes was not significantly associated with cognitive
-function in children under 6 months, though this may be due to the low
-taxonomic diversity, and broad lack of overlap between taxa in infants.
-As expected, age was significantly associated with microbial beta
-diversity (taxa R\textsuperscript{2} = 0.0215, and UniRef90s
-R\textsuperscript{2} = 0.0245, q \textless{} 0.001), and very strongly
-associated with neuroimaging profiles (R\textsuperscript{2} = 0.256, q
-\textless{} 0.001).
-
-Consistent with prior studies, different microbial measurement types
-captured overlapping variation, with species profiles and gene function
-profiles (annotated with clusters of 90\% similarity
-\href{https://www.zotero.org/google-docs/?b9rEJS}{(Suzek et al.,
-2007)}), both generated from DNA sequencing data, being tightly coupled
-(Figure 1H, p \textless{} 0.001). Interestingly, other functional
-groupings (Enzyme commission level4 - ECs
-\href{https://www.zotero.org/google-docs/?fs8LO7}{(Bairoch, 2000)}, and
-KEGG orthologues - KOs
-\href{https://www.zotero.org/google-docs/?Xx5duE}{(Kanehisa et al.,
-2004)}) overlapped only slightly with taxonomic profiles in both age
-cohorts, despite being derived from UniRef90 labels. In children over 18
-months, some variation (15.9\%, p \textless{} 0.01) in neuroimaging
-overlapped with microbial measures, though this may be due to the
-residual variation due to age in both measures.
-
-\hypertarget{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}{%
-\subsubsection{Microbial species and neuroactive genes are associated
-with cognitive
-performance}\label{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}}
-
-To assess whether individual microbial species were associated with
-cognitive function, we fit multivariable linear regression
-\href{https://www.zotero.org/google-docs/?4qS6UF}{(Mallick et al.,
-2021)} to the relative abundance of each species that had at least 15\%
-prevalence in a given age group (Figure 2A, N = 92 for 0--120 months, N
-= 46 for 0--6 months, N = 97 for 18--120 months). Only \emph{Blautia
-wexlerae} was significantly associated (q value = 0.14, β = 0.0015) with
-cognitive function in children under 6 months old after adjusting for
-age and maternal education (Figure 2B). \emph{B. wexlerae} was
-previously shown to be depleted in children with diabetes
-\href{https://www.zotero.org/google-docs/?6mU2nI}{(Benítez-Páez et al.,
-2020)}, and that oral administration of \emph{B. wexlerae} partially
-ameliorated weight gain and inflammation from a high-fat diet in a mouse
-model of T2D \href{https://www.zotero.org/google-docs/?y83hpS}{(Hosomi
-et al., 2022; Liu et al., 2021)}. In children over 18 months of age,
-several microbial species were significantly enriched (q-value
-\textless{} 0.20) in children with higher cognitive function scores,
-including \emph{Gordonibacter pamelaeae}, which produces the
-neuroprotective metabolite urolithin
-\href{https://www.zotero.org/google-docs/?HGtTEY}{(Gong et al., 2022;
-Selma et al., 2014)}, \emph{Asaccharobacter celatus} and
-\emph{Adelcreutzia equolifaciens}, which produce phytoestrogen-derived
-equol \href{https://www.zotero.org/google-docs/?HCKZqm}{(Maruo et al.,
-2008; Thawornkuno et al., 2009)}, and the SCFA-producing probiotic
-species such as \emph{Eubacterium eligens} and \emph{Faecalibacterium
-prausnitzii} \href{https://www.zotero.org/google-docs/?3eyheK}{(Ghosh et
-al., 2020; Lopez-Siles et al., 2017; Mukherjee et al., 2020)} (Figure
-2B).
-
-\includegraphics[width=6.5in,height=4.83333in]{media/image4.png}
-
-\textbf{Figure 2 - Taxa and gene functional groups are associated with
-cognitive function.}
-
-(A) Volcano plot of multivariable linear model results showing the
-relationship between individual taxa and cognitive function score in
-children over 18 months of age, controlling for age, maternal education,
-and sequencing depth. Taxa that were significant after FDR correction (q
-\textless{} 0.2) are colored red. (B) For taxa that were significantly
-associated with cognitive function, heatmaps of prevalence, mean
-relative abundance, and correlation with cognitive function in different
-age groups. (C-F) Enrichment plots for selected neuroactive gene sets
-used in feature set enrichment analysis (FSEA). (G) summary of FSEA
-results across all samples (left) as well as in the under 6 months
-(middle) and over 18 months (right) subsets, colored based on the
-significance of the association. Markers indicate the individual
-correlation of genes within a gene set, and vertical bars represent the
-median correlation of that gene set.
-
-Given that different microbial species might occupy the same metabolic
-niche in different individuals, we hypothesized that microbial genes
-grouped by functional activity would be associated with cognition. To
-test this, we performed feature set enrichment analysis (FSEA) on groups
-of genes with neuroactive potential
-\href{https://www.zotero.org/google-docs/?xQURsj}{(Valles-Colomer et
-al., 2019)} and concurrent cognitive function score (Table 2,
-\protect\hyperlink{_i7fd3z19jdvg}{\uline{Figure 2}}C-G) and found that
-several metabolic pathways were significantly enriched or depleted in
-children with higher cognitive function scores. This was true both when
-considering all age groups together, though the enrichment of most
-pathways was more pronounced in children under 6 months or over 18
-months. For example, genes for degrading the 3-carbon SCFA propionate
-were significantly depleted in children with higher cognitive function
-scores across all age groups tested (Table 2; propionate degradation I,
-under 6 months, enrichment score (E.S.) = -0.542, corrected p-value (q)
-= 0.020; over 18 months, E.S. = -0.674, q = 0.041). Interestingly, genes
-for propionate synthesis were also significantly depleted in higher
-scoring children over 18 months (E.S. -0.676, q = 0.023), as were genes
-for synthesizing the 2 carbon SCFA acetate in children under 6 months
-old (acetate synthesis I, E.S. = -0.194, q = 0.153; acetate synthesis
-II, E.S. = -0.342, q = 0.020; acetate synthesis III, E.S. = -0.31, q =
-0.052). SCFAs are produced by anaerobic fermentation of dietary fiber
-and have been linked with immune system regulation as well as directly
-with brain function
-\href{https://www.zotero.org/google-docs/?2Z0sKZ}{(Dalile et al.,
-2019)}.
-
-Synthesis of menaquinone (vitamin K) was also negatively associated with
-cognitive function score in older children (menaquinone synthesis I,
-E.S. = -0.170, q = 0.0183). Menaquinone has several isoforms, one of
-which, MK-4, has been found to be decreased in children diagnosed with
-ASD \href{https://www.zotero.org/google-docs/?dTA3NO}{(Dong et al.,
-2021)} and is potentially neuroprotective in both rodents and humans
-\href{https://paperpile.com/c/dPbU4e/cTO9}{(Elkattawy et al., 2022)}. In
-children over 18 months old, genes for the synthesis of the amino acids
-glutamate and tryptophan were significantly enriched in children with
-higher cognitive function scores (Glutamate synthesis I, E.S. = 0.242, q
-= 0.047; Tryptophan synthesis, E.S. = 0.119, q = 0.041). Glutamate is a
-critical neurotransmitter controlling neuronal excitatory/inhibitory
-signaling along with gamma-aminobutyric acid (GABA), and their balance
-in the brain controls neural plasticity and learning, particularly in
-the developing brain
-\href{https://www.zotero.org/google-docs/?QubMhx}{(Cohen Kadosh et al.,
-2015; Palomo-Buitrago et al., 2019)}. Tryptophan metabolism, including
-microbial metabolism of tryptophan, has previously been linked with
-autism in children
-\href{https://www.zotero.org/google-docs/?GPaUZE}{(Hoshino et al., 1984;
-Xiao et al., 2021)}. Taken together, these results suggest that
-microbial metabolic activity, particularly the metabolism (synthesis and
-degradation) of neuroactive compounds may have effects on cognitive
-development.
-
-\textbf{Table 2 - Feature set enrichment analysis on neuroactive
-microbial gene sets}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2300}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1259}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1525}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1429}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2785} + 2\tabcolsep}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{All}
-\end{minipage}} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2591} + 2\tabcolsep}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{under 6 months}
-\end{minipage}} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2324} + 2\tabcolsep}@{}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{over 18 months}
-\end{minipage}} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-geneset
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1223
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5549
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1944
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1533}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1269
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. II
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2201
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0660}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3424
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0198}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0946
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9199
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. III
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2104
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0859}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3098
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0518}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0857
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9199
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Glutamate deg. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.4037
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1802}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2840
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2489
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Glutamate syn. II
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0953
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6022
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1342
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9802
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2421
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0466}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Menaquinone syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0853
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5635
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0833
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1699
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1825}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Propionate deg. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.4612
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1302}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.5421
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0198}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.6738
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0414}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Propionate syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3344
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5549
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3201
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5713
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.6760
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0227}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Tryptophan syn.
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0573
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9742
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0725
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1186
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0414}
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\hypertarget{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}{%
-\subsubsection{Gut microbial taxonomic and functional profiles predict
-cognitive
-function}\label{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}}
-
-FSEA relies on understanding functional relationships between individual
-genes. However, because the relationships between individual taxa are
-still largely unknown, we turned to Random Forest (RF) models, an
-unsupervised non-parametric machine learning (ML) method that enables
-the identification of underlying patterns in large numbers of individual
-features (here, microbial species). Previous studies have reported
-successful use of RFs for processing highly-dimensional and sparse data
-from the domain of genomics
-\href{https://www.zotero.org/google-docs/?C3CSPB}{(Amaratunga et al.,
-2008; Brieuc et al., 2018; Chen and Ishwaran, 2012; Franzosa et al.,
-2019; Stephan et al., 2015)}, along with other works where it was used
-to predict cognitive conditions related to Alzheimer's disease in
-different scenarios
-\href{https://www.zotero.org/google-docs/?GiadGy}{(Ardekani et al.,
-2017; Velazquez et al., 2021)}. Additionally, given the sequential
-nature of variable consideration in each tree, RFs are naturally able to
-work out complex input feature interactions, such as those present in a
-microbiome-wide study, without the necessity to explicitly compute
-interaction terms.
-
-\includegraphics[width=6.5in,height=5.19444in]{media/image2.png}
-
-\textbf{Figure 3 - Random forest models predict concurrently measured
-cognitive function.}
-
-Comparison of RF predictor importance versus linear models for children
-between birth and 6 months old (A), and for those older than 18 months
-(B). Colors represent whether the species belong to the group of
-top-important features that account for 80\% of the cumulative
-importance on the RF model, if that species was significant (q
-\textless{} 0.2) in linear models, both, or neither. Ranked feature
-importance for taxa in RF models for children between birth and 6 months
-old (C), and for those older than 18 months (D). Taxa that are important
-for RF models both for children under 6 months and for those over 18
-months (E), only for children under 6 months (F), and only for children
-over 18 months (G)
-
-Given that gut microbial profiles, as well as neurocognitive
-development, may partially reflect socioeconomic and demographic
-factors, we assessed the performance of RF regressors where maternal
-education (a proxy of socioeconomic status (SES)), sex, and age were
-included as possible predictors, either alone or in combination with
-microbial taxonomic profiles (Table 3).
-
-\textbf{Table 3. Benchmark metrics for the cognitive assessment score
-prediction models. Confidence intervals are calculated from the
-distribution of metrics from repeated CV at a confidence level of 95\%}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1456}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0960}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3584}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Subject Ages (months)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Microbial feature
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Demo.
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Test set correlation (± C.I.)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Test set Root-mean-square error (± C.I.)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.14 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-13.01 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.10 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.66 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.13 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.70 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.01 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.56 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.01 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.56 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.506 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16.27 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.363 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.66 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.429 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.29 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.303 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18.01 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.333 ± 0.004
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.86 ± 0.04
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-As with linear models, RF models for children under 6 months old were
-less generalizable (mean test-set correlation -0.13, mean RMSE 12.70),
-but RFs were consistently able to learn the relationship between taxa
-and cognitive function scores in children over 18 months of age (mean
-test-set correlation 0.429, mean RMSE 17.29). For both age groups,
-species that were important in RF overlapped with those that were
-significant when testing the relationship with linear models (Figure
-3A-B, Supplementary Table XXA-B), though there were substantial
-differences. For children under 6 months, only \emph{B. wexlerae} was
-significantly associated with cognitive function in linear models, but
-was ranked 20th in importance for RF models. The importance metric
-employed (MDI) is a measure of how well the variable differentiates
-(splits) sets of samples by creating subgroups that reduce the
-intra-group deviations, while also accounting for the amount of samples
-affected by a split. All taxa significantly associated with cognitive
-function score in children over 18 months using LMs belong to the
-top-ranking group responsible for 80\% of the total relative importance,
-except for \emph{Eubacterium ramulus} (Table 3, Figure 3A).
-
-Interestingly, several taxa highly ranked in importance in both age
-groups, including several that were significant in LMs for children over
-18 months old, including \emph{R. gnavus} (0--6 months, rank = 13;
-18--120 months, rank = 7) and \emph{G. pamelaeae} (0--6 months, rank =
-16; 18--120 months, rank = 20), while others such as \emph{Allistipes
-finegoldii} were age-group specific. Several taxa important in RF models
-were not statistically significant when using linear models after
-multiple hypothesis correction. However, these taxa had small nominal
-p-values. For example, \emph{Erysipelatoclostridium ramosum} was the
-most important feature in RF models for children under 6 months old and
-had an LM p-value of 0.04. Subject age was consistently ranked highly in
-feature importance, which could indicate that decision branches based on
-microbial taxa have increased purity when considering the subject's age
-or that age itself is a useful predictor.
-
-\hypertarget{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}{%
-\subsubsection{Gut microbial taxonomic profiles predict brain structure
-differences}\label{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}}
-
-If there are causal effects of microbial metabolism on cognitive
-function, they might be reflected in changes in neuroanatomy. We again
-employed a Random Forest modeling approach to associate gut taxonomic
-profiles with individual brain regions identified in MRI scans,
-normalized to total brain volume. Some brain regions were more readily
-predicted by RF models trained on microbial taxa (Table 4, Supplementary
-Table XX), in particular those that were highly correlated with age.
-These included the L/R lingual gyrus (mean RF correlations, Left =
-0.421, Right = 0.434; relative age importances, Left = 7.5\%, Right =
-8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
-0.200, Right = 0.273; relative age importances, Left = 3.7\%, Right =
-6.3\%). In many cases, however, age was not an important variable in
-high-performing models, such as that for the left accumbens area (mean
-RF correlations = 0.288; relative age importances = 1.2\%).
-
-\textbf{Table 4. Summary statistics for the neuroanatomy prediction
-benchmarks}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Statistic
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Mean absolute proportional error (MAPE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Correlation coefficient (R)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-mean
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.072
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.137
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Standard deviation
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.028
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.136
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-maximum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.605
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-75th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.195
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-50th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-25th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.062
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-minimum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.147
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-We also observed that many brain regions had high accordance across the
-left and right hemispheres in terms of both model performance and
-microbial feature importance. In contrast, other regions had substantial
-differences between the hemispheres. For example, the left accumbens
-area, which plays an important role in reward circuits
-\href{https://www.zotero.org/google-docs/?zVicU7}{(Ernst et al., 2005;
-Yau et al., 2012)}, has one of the highest test-set correlations of our
-brain region models (R = 0.288), as compared to the right accumbens
-models which could not adequately generalize, and had a negative mean
-test-set correlation (R = -0.041). Most healthy individuals have a
-rightward asymmetry in the nucleus accumbens, and reduced asymmetry has
-been linked to substance use disorder in young adults
-\href{https://www.zotero.org/google-docs/?V45KyN}{(Cao et al., 2021)}.
-Feature importance for models of the left accumbens area were dominated
-by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.4\%
-relative importance), \emph{B. ovatus} (3.7\% relative importance), and
-\emph{B. uniformis} (3.0\% relative importance).. The accumbens area is
-associated with reward control, and in individuals diagnosed with ADHD,
-it has been shown to have a divergent neuromorphology
-\href{https://www.zotero.org/google-docs/?kLfEQ5}{(Hoogman et al.,
-2017)}. Independently, \emph{B. ovatus}, \emph{B. uniformis} and
-\emph{B. vulgatus} have been linked to ADHD
-\href{https://www.zotero.org/google-docs/?luO0zh}{(Wang et al., 2020)}.
-In fact, there are studies showing that alterations on the striatal
-dopamine transporters can cause effects resembling hyperactivity and
-attention deficit
-\href{https://www.zotero.org/google-docs/?UxS47e}{(Yael et al., 2019)},
-and that \emph{B. uniformis} is gut-microbial modulator of the brain
-dopamine transporter
-\href{https://www.zotero.org/google-docs/?L1cWJS}{(Hartstra et al.,
-2020)}.
-
-Many of the taxa identified included taxa also identified in LMs and RF
-models of cognitive function. Interesting to note, while RF models for
-multiple brain regions had many important microbial taxa, others were
-dominated by a small number of taxa. In general, for all segments, a
-consistent number of species between 14 and 22 (median = 20) was
-responsible for one third of the fitness-weighted cumulative importance
-(see Methods), regardless of the model benchmarks. We proceeded to
-select a subset of the most important taxa (averaged over all segments)
-and the segments whose importance was more heavily loaded to further
-analyze the multiple relationships unveiled by the RF models
-(\textbf{Figure 4 A and B})
-
-\includegraphics[width=6.5in,height=5.19444in]{media/image1.png}
-
-\textbf{Figure 4: Microbial feature importance in predicting brain
-volumes in children over 18 months of age.}
-
-(A) Average test-set correlations for prediction of MRI segmentation
-data from microbiome and demographics on select segments after repeated
-cross validation. (B) Heatmap of average individual relative taxa
-importances on each brain segment. Importances are reported as
-proportions relative to the sum of importances for each model - since
-every model is trained with 132 features, and even distribution of
-importance would be 0.75\% for each feature.
-
-. Segments and taxa ordered by HCA on a select list of species with high
-ranks on average importances, and their respective highest-load
-segments. (C)
-
-Our analysis revealed two major patterns of importance distribution from
-taxa over the brain segments; some species portrayed high contributions
-to multiple different segments, while others contributed modestly to
-just one or two brain segments. Notable cases of the first pattern
-included seven species - \emph{Anaerostipes hadrus}, \emph{Bacteroides
-vulgatus}, \emph{Fusicatenibacter saccharivorans}, \emph{Ruminococcus
-torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
-\emph{Blautia wexlerae} - that combined, account for approximately one
-third of the cumulative relative importances, computed after subsetting
-on the taxa of interest (Figure 4C).
-
-Among these, the most important variable is \emph{Anaerostipes hadrus},
-which is a butyrate-producing anaerobe that has been positively
-associated with cognitive function
-\href{https://www.zotero.org/google-docs/?AVJiwu}{(Kant et al., 2015; Li
-et al., 2022)}. Its importance is, however, heavily loaded on the volume
-of entire major lobes (frontal, parietal), which cannot be readily
-correlated to specific cognitive functions. After accounting for the
-major lobes, it is found to be highly important on the prediction of the
-cerebellar vermal lobules VIII-X, pars opercularis, cuneus and
-precuneus, and anterior-cingulate (Figure 4B).
-
-To better understand this relationship, we performed hierarchical
-cluster analysis, which revealed a well defined cluster of species with
-high importance loadings on close segments of the lower temporal and
-close occipital lobe, to which \emph{A. hadrus} belongs. The other
-species on this cluster are \emph{B. wexlerae}, \emph{R. torques},
-\emph{R. intestinalis, R. bicirculans} and \emph{F. saccahrivorans}, who
-contribute heavily to the entorhinal, fusiform, lingual and
-parahippocampal segments.
-
-This group of taxa and their related segments drew our attention because
-they contained \emph{B. wexlerae}, a highlight from the cognitive
-assessment results. Previous works exploring Parkinson Disease patients
-found out that issues in the cognitive task of confrontation naming were
-positively correlated with thinning in the fusiform gyrus and
-parahippocampal gyrus
-\href{https://www.zotero.org/google-docs/?rlO6eq}{(Pagonabarraga et al.,
-2013)}. Additionally, the left and right parahippocampal, where \emph{B.
-wexlerae} had the highest importance in both models, are important in
-visual/spatial processing and memory
-\href{https://www.zotero.org/google-docs/?wkkWZT}{(Aminoff et al.,
-2013)}. \emph{B. wexlerae} can also produce acetylcholine in the gut
-\href{https://www.zotero.org/google-docs/?jBoxSm}{(Hosomi et al.,
-2022)}, and this molecule plays an important role in modulating memory
-function \href{https://www.zotero.org/google-docs/?BQ0j2c}{(Haam and
-Yakel, 2017)}.
-
-Another notable importance cluster contains taxa associated with the
-basal forebrain, especially the cingulate and the accumbens. This
-cluster contains the previously-discussed \emph{B. ovatus} and \emph{B.
-uniformis,} but also \emph{Alistipes finegoldii and Streptococcus
-salivaris.} Reduction in the nucleus accumbens has been associated with
-depression symptoms
-\href{https://www.zotero.org/google-docs/?2aSZIr}{(Wacker et al.,
-2009)}, and increased levels of the \emph{Alistipes} genus have been
-observed in patients with depressive disorder
-\href{https://www.zotero.org/google-docs/?bhhSZB}{(Jiang et al., 2015)}.
-While these reports are important, simultaneously probing of the gut
-metagenome and brain structure relationship is novel.
-
-Both \emph{A. equolofaciens} and \emph{A. celatus,} two closely-related
-equol-producing species, are examples of the second contribution
-pattern, and had high importance in predicting the relative volume of
-the right anterior cingulate (respectively, 3.0\% and 2.6\% relative
-importances), which has been linked to social cognition and reward-based
-decision making \href{https://www.zotero.org/google-docs/?Wi8Bpq}{(Apps
-et al., 2016; Boes et al., 2008; Bush et al., 2002)}. Equol has a strong
-estrogenic effect
-\href{https://www.zotero.org/google-docs/?PqgFU7}{(Setchell et al.,
-2005)}, and in anterior cingulate, estrogen, has been shown to regulate
-pain-related aversion
-\href{https://www.zotero.org/google-docs/?5huRIR}{(Xiao et al., 2013)}.
-Another example of this pattern is \emph{R. gnavus,} which was the only
-species with significant negative association with cognitive function.
-It is heavily associated with the left pars opercularis (2.7\% relative
-importance), a relationship that may be explained by the emerging
-understanding that this species is increased in individuals with insulin
-resistance and obesity
-\href{https://www.zotero.org/google-docs/?lwzWUv}{(Ley et al., 2006)},
-conditions that are known to produce structural abnormalities in the
-brain \href{https://www.zotero.org/google-docs/?NlkFr6}{(Opel et al.,
-2021)}. Finally, \emph{Coprococcus comes} displays an importance
-distribution that splits almost evenly among the two previously reported
-clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (4.0\% relative importance)
-(only age had higher relative importance in any model), while also being
-one of the most important predictors for neighboring areas of the
-posterior cingulate such as the pars opercularis (relative importances,
-Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
-precentral (2.3\% relative importance) and paracentral lobes (relative
-importances, Left = 2.6\%, Right = 1.8\%).
-
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
-
 The relationship between the gut microbiome and brain function via the
@@
 
-\{\{Guilherme add a paragraph about the usefulness of ML\}\}
-
 The use of multiple age-appropriate cognitive assessments that could be
@@
 medicinally useful molecules that can be applied in intervention.
-
-\hypertarget{material-and-methods}{%
-\subsection{Material and Methods}\label{material-and-methods}}
-
-\hypertarget{study-ethics}{%
-\subsubsection{Study Ethics}\label{study-ethics}}
-
-All procedures for this study were approved by the local institutional
-review board at Rhode Island Hospital, and all experiments adhered to
-the regulation of the review board. Written informed consent was
-obtained from all parents or legal guardians of enrolled participants.
-
-\hypertarget{participants}{%
-\subsubsection{Participants}\label{participants}}
-
-Data used in this study were drawn from the ongoing longitudinal
-RESONANCE study of healthy and neurotypical brain and cognitive
-development, based at Brown University in Providence, RI, USA. The
-RESONANCE study is part of the NIH initiative Environmental influences
-on Child Health Outcomes (ECHO)
-\href{https://www.zotero.org/google-docs/?0tZanX}{(Forrest et al., 2018;
-Gillman and Blaisdell, 2018)}, a longitudinal observational study of
-healthy and neurotypical brain development that spans the fetal and
-infant to adolescent life stages, combining neuroimaging (magnetic
-resonance imaging, MRI), neurocognitive assessments, bio-specimen
-analyses, subject genetics, environmental exposures such as lead, and
-rich demographic, socioeconomic, family and medical history information.
-From the RESONANCE cohort, 361 typically-developing children between the
-ages of 2.8 months and 10 years old (median age 2 years, 2 months) were
-selected for analysis in this study.
-
-General participant demographics are provided in Table 1. Children are
-representative of the RI population. Children in the RESONANCE cohort
-were born full-term (\textless{} 37 weeks gestation) with height and
-weight normal for gestational age, and from uncomplicated singleton
-pregnancies. Children with known major risk factors for developmental
-abnormalities at enrollment were excluded. In addition to screening at
-the time of enrollment, on-going screening for worrisome behaviors using
-validated tools was performed to identify at-risk children and remove
-them from subsequent analysis.
-
-Exclusion criteria included: \emph{in utero} exposure to alcohol,
-cigarette or illicit substance exposure; preterm (\textless{} 37 weeks
-gestation) birth; small for gestational age or less than 1500 g; fetal
-ultrasound abnormalities; preeclampsia, high blood pressure, or
-gestational diabetes; 5 minute APGAR scores \textless{} 8; NICU
-admission; neurological disorder (e.g., head injury resulting in loss of
-consciousness, epilepsy); and psychiatric or learning disorder
-(including maternal depression) in the infant, parents, or siblings
-requiring medication in the year prior to pregnancy.
-
-Demographic and other non-biospecimen data such as race and ethnicity,
-parental education and occupation, feeding behavior (breast- and
-formula-feeding), child weight and height, were collected through
-questionnaires or direct examination as appropriate. All data were
-collected at every assessment visit, if possible.
-
-\hypertarget{cognitive-assessments}{%
-\subsubsection{Cognitive Assessments}\label{cognitive-assessments}}
-
-Overall cognitive function was assessed using age-appropriate methods.
-For children from birth to 30 months, we used an Early Learning
-Composite as assessed via the Mullen Scales of Early Learning (MSEL)
-\href{https://www.zotero.org/google-docs/?NSyykd}{(Mullen and others,
-1995)}, a standardized and population-normed tool for assessing fine and
-gross motor, expressive and receptive language, and visual reception
-functioning in children from birth through 68 months of age. The
-Wechsler Intelligence Quotient for Children (WISC)
-\href{https://www.zotero.org/google-docs/?OopdA1}{(Wechsler, 2012)} is
-an individually administered standard intelligence test for children
-aged 6 to 16 years. It derives a full scale intelligence quotient (IQ)
-score, which we used to assess overall cognitive functioning. The fourth
-edition of the Wechsler Preschool and Primary Scale of Intelligence
-(WPPSI-IV) is an individually administered standard intelligence test
-for children aged 2 years 6 months to 7 years 7 months, trying to meet
-the increasing need for the assessment of preschoolers. Just as the
-WISC, it derives a full scale IQ score, which we used to assess overall
-cognitive functioning.
-
-\hypertarget{stool-sample-collection-and-sequencing}{%
-\subsubsection{Stool Sample Collection and
-Sequencing}\label{stool-sample-collection-and-sequencing}}
-
-Stool samples (n=493) were collected by parents in OMR-200 tubes
-(OMNIgene GUT, DNA Genotek, Ottawa, Ontario, Canada), stored on ice, and
-brought within 24 hrs to the lab in RI where they were immediately
-frozen at -80 ˚C. Stool samples were not collected if the subject had
-taken antibiotics within the last two weeks. DNA extraction was
-performed at Wellesley College (Wellesley, MA). Nucleic acids were
-extracted from stool samples using the RNeasy PowerMicrobiome kit,
-excluding the DNA degradation steps. Briefly, the samples were lysed by
-bead beating using the Powerlyzer 24 Homogenizer (Qiagen, Germantown,
-MD) at 2500 rpm for 45 s and then transferred to the QIAcube (Qiagen,
-Germantown, MD) to complete the extraction protocol. Extracted DNA was
-sequenced at the Integrated Microbiome Resource (IMR, Dalhousie
-University, NS, Canada).
-
-Shotgun metagenomic sequencing was performed on all samples. A pooled
-library (max 96 samples per run) was prepared using the Illumina Nextera
-Flex Kit for MiSeq and NextSeq from 1 ng of each sample. Samples were
-then pooled onto a plate and sequenced on the Illumina NextSeq 550
-platform using 150+150 bp paired-end ``high output'' chemistry,
-generating 400 million raw reads and 120 Gb of sequence per plate.
-
-\hypertarget{machine-learning-for-cognitive-development}{%
-\subsubsection{Machine learning for cognitive
-development}\label{machine-learning-for-cognitive-development}}
-
-Prediction of cognitive scores was carried out as a set of regression
-experiments targeting real-valued continuous assessment scores.
-Different experiment sets were designed to probe how different
-representations of the gut microbiome (taxonomic profiles, Functional
-profiles encoded as ECs) would behave, with and without the addition of
-demographics (sex and maternal education as a proxy of socioeconomic
-status) on participants from different age groups. Age (in months) was
-provided as a covariate for all models (Table 3).
-
-\textbf{Table 5. Experimental design and input composition for Random
-Forest experiments}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1797}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2915}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Input set
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Age bracket
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Microbiome encoding type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Demographics
-
-Provided?
-
-(sex, education)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-1
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Not provided
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-3
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-4
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-5
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Not provided
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-7
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-9
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-10
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-Random Forests (RFs) \href{https://paperpile.com/c/dPbU4e/VDqU}{(Breiman
-1996)} were selected as the prediction engine and processed using the
-DecisionTree.jl
-\href{https://www.zotero.org/google-docs/?dQLyDs}{(Sadeghi et al.,
-2022)} implementation, inside the MLJ.jl
-\href{https://www.zotero.org/google-docs/?y8ywGI}{(Blaom et al., 2020)}
-Machine Learning framework. Independent RFs were trained for each
-experiment, using a set of default regression hyperparameters from
-Breiman and Cutler
-\href{https://www.zotero.org/google-docs/?d3gqFh}{(Breiman, 2001)}, on a
-repeated cross-validation approach with different RNG seeds. One hundred
-repetitions of 3-fold CV with 10 different intra-fold RNG states each
-were employed, for a total of 3000 experiments per input set.
-
-After the training procedures, the root-mean-square error (RMSE) for
-cognitive assessment scores and mean absolute proportional error (MAPE)
-for the brain segmentation data, along with Pearson\textquotesingle s
-correlation coefficient (R) were benchmarked on the validation and train
-sets. MAPE was chosen as the metric for brain segments due to magnitude
-differences between median volumes of each segment, which would hinder
-interpretation of raw error values without additional reference.
-
-To derive biological insight from the models, the covariate variable
-importances for all the input features, measured by mean decrease in
-impurity (MDI, or GINI importance), was also analyzed. Leveraging the
-distribution of results from the extensive repeated cross validation
-experiments, rather than electing a representative model or picking the
-highest validation-set correlation, a measure of model fitness
-(\textbf{Equation 1}) was designed to weight the importances from each
-trained forest. The objective was to give more weight to those with
-higher benchmarks on the validation sets (or higher generalizability),
-while penalizing information from highly overfit models, drawing
-inspiration from the approach used on another work employing repeated CV
-on Random Forests with high-dimensional, low sample size microbiome
-datasets \{***\}. The resulting fitness-weighted importances were used
-to generate the values in Figure 3.
-
-\(fitness =\)
-
-\textbf{Equation 1.} Mathematical expression of the fitness measure used
-to weight feature importances based on model benchmarks
-
-\hypertarget{mri-segmentation}{%
-\subsubsection{MRI / segmentation}\label{mri-segmentation}}
-
-MRI data was acquired at a 3T Siemens Trio scanner with the following
-parameters: TE=5.6msec, TR=1400msec, FA=15 degrees, 1.1x1.1x1.1mm
-resolution, 160x160 matrix, with an average of 112 slices. FOV was
-adjusted to infant size. Using a combination of linear and nonlinear
-image registration, we created representative age-specific templates
-using the ANTs package
-\href{https://www.zotero.org/google-docs/?ZlWDfL}{(Avants et al.,
-2014)}. After age specific templates were created, a single flow from
-each age to the 12-month template was estimated and a final warp from
-the 12-month template to standard adult MNI space was performed. For
-each individual infant/ child brain image, we then calculated the warp
-from their native T1w image space to their nearest-in-age template.
-Using the resulting warps (native → nearest age template → 12-month
-template → MNI template), we could move the standard adult brain atlas
-to the space of an individual infant in a single step. In this case, we
-used the Harvard-Oxford brain atlas to provide a coarse-grained
-parcellation of individual brains into subcortical regions (e.g.,
-thalamus, putamen) and total grey and white matter volumes (included as
-part of the FSL package
-\href{https://www.zotero.org/google-docs/?AapPN7}{(Jenkinson et al.,
-2012)}. Total tissue and brainstem volume as well as left and right
-hemisphere volumes were derived for total white and cortical gray
-matter, lateral ventricle, thalamus, caudate, putamen, pallidum,
-hippocampus, amygdala, and accumbens as well as total brainstem volume
-\href{https://www.zotero.org/google-docs/?MCDDB9}{(Bruchhage et al.,
-n.d.)}.
-
-\hypertarget{computational-analysis-of-metagenomes}{%
-\subsubsection{Computational analysis of
-metagenomes}\label{computational-analysis-of-metagenomes}}
-
-Shotgun metagenomic sequences were analyzed using the bioBakery suite of
-computational tools \href{https://paperpile.com/c/dPbU4e/vDau}{(Beghini
-et al. 2021)}. First, KneadData (v0.7.7) was used to perform quality
-control of raw sequence reads, such as read trimming and removal of
-reads matching a human genome reference. Next, MetaPhlAn (v3.0.7, using
-database mpa\_v30\_CHOCOPhlAn\_201901) was used to generate taxonomic
-profiles by aligning reads to a reference database of marker genes.
-Finally, HUMAnN (v3.0.0a4) was used to functionally profile the
-metagenomes.
-
-\hypertarget{data-and-code-availability}{%
-\subsubsection{Data and code
-availability}\label{data-and-code-availability}}
-
-Taxonomic and functional microbial profiles, as well as subject
-demographics necessary for statistical analyses and machine learning are
-available on the Open Science Framework
-\href{https://www.zotero.org/google-docs/?61HLlM}{(Bonham et al.,
-2022)}. Data processing, generation of summary statistics, and
-generation of plots was performed using the julia programming language
-\href{https://www.zotero.org/google-docs/?zj3HnY}{(Bezanson et al.,
-2017; Bonham et al., 2021; Danisch and Krumbiegel, 2021)}. All code for
-data analysis and figure generation, as well as scripts for automated
-download of input files are available on github \{\{CITE zenodo\}\}.
-
-\hypertarget{research-standards}{%
-\subsection{Research Standards}\label{research-standards}}
-
-\hypertarget{acknowledgements}{%
-\subsection{Acknowledgements}\label{acknowledgements}}
-
-Research funding: NIH UG3 OD023313 and Wellcome LEAP 1kD.
-
-\hypertarget{references-and-citations}{%
-\subsection{References and Citations}\label{references-and-citations}}
-
-\url{https://doi.org/10.1093/cercor/bhs201}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Amaratunga, D.,
-Cabrera, J., Lee, Y.-S., 2008. Enriched random forests. Bioinformatics
-24, 2010--2014. https://doi.org/10.1093/bioinformatics/btn356}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Aminoff, E.M.,
-Kveraga, K., Bar, M., 2013. The role of the parahippocampal cortex in
-cognition. Trends Cogn. Sci. 17, 379--390.
-https://doi.org/10.1016/j.tics.2013.06.009}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Apps, M.A.J.,
-Rushworth, M.F.S., Chang, S.W.C., 2016. The Anterior Cingulate Gyrus and
-Social Cognition: Tracking the Motivation of Others. Neuron 90,
-692--707. https://doi.org/10.1016/j.neuron.2016.04.018}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ardekani, B.A.,
-Bermudez, E., Mubeen, A.M., Bachman, A.H., Initiative, for the A.D.N.,
-2017. Prediction of Incipient Alzheimer's Disease Dementia in Patients
-with Mild Cognitive Impairment. J. Alzheimers Dis. 55, 269--281.
-https://doi.org/10.3233/JAD-160594}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Avants, B.B.,
-Tustison, N.J., Stauffer, M., Song, G., Wu, B., Gee, J.C., 2014. The
-insight ToolKit image registration framework. Front. Neuroinform. 8, 44.
-https://doi.org/10.3389/fninf.2014.00044}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bäckhed, F., Roswall,
-J., Peng, Y., Feng, Q., Jia, H., Kovatcheva-Datchary, P., Li, Y., Xia,
-Y., Xie, H., Zhong, H., Khan, M.T., Zhang, J., Li, J., Xiao, L.,
-Al-Aama, J., Zhang, D., Lee, Y.S., Kotowska, D., Colding, C., Tremaroli,
-V., Yin, Y., Bergman, S., Xu, X., Madsen, L., Kristiansen, K., Dahlgren,
-J., Wang, J., 2015. Dynamics and stabilization of the human gut
-microbiome during the first year of life. Cell Host Microbe 17,
-690--703. https://doi.org/10.1016/j.chom.2015.04.004}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bairoch, A., 2000. The
-ENZYME database in 2000. Nucleic Acids Res. 28, 304--305.
-https://doi.org/10.1093/nar/28.1.304}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Benítez-Páez, A.,
-Gómez del Pugar, E.M., López-Almela, I., Moya-Pérez, Á., Codoñer-Franch,
-P., Sanz, Y., 2020. Depletion of \emph{Blautia} Species in the
-Microbiota of Obese Children Relates to Intestinal Inflammation and
-Metabolic Phenotype Worsening. mSystems 5, e00857-19.
-https://doi.org/10.1128/mSystems.00857-19}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bezanson, J., Edelman,
-A., Karpinski, S., Shah, V.B., 2017. Julia: A fresh approach to
-numerical computing. Siam Rev. 59, 65--98.
-https://doi.org/10.1137/141000671}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Blackwell, C.K.,
-Mansolf, M., Sherlock, P., Ganiban, J., Hofheimer, J.A., Barone, C.J.,
-Bekelman, T.A., Blair, C., Cella, D., Collazo, S., Croen, L.A., Deoni,
-S., Elliott, A.J., Ferrara, A., Fry, R.C., Gershon, R., Herbstman, J.B.,
-Karagas, M.R., LeWinn, K.Z., Margolis, A., Miller, R.L., O'Shea, T.M.,
-Porucznik, C.A., Wright, R.J., 2022. Youth well-being during the
-COVID-19 pandemic. Pediatrics 149.
-https://doi.org/10.1542/peds.2021-054754}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Blaom, A.D., Kiraly,
-F., Lienart, T., Simillides, Y., Arenas, D., Vollmer, S.J., 2020. MLJ: A
-Julia package for composable machine learning. J. Open Source Softw. 5,
-2704. https://doi.org/10.21105/joss.02704}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Boes, A.D., Tranel,
-D., Anderson, S.W., Nopoulos, P., 2008. Right anterior cingulate cortex
-volume is a neuroanatomical correlate of aggression and defiance in
-boys. Behav. Neurosci. 122, 677--684.
-https://doi.org/10.1037/0735-7044.122.3.677}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bokulich, N.A., Chung,
-J., Battaglia, T., Henderson, N., Jay, M., Li, H., D Lieber, A., Wu, F.,
-Perez-Perez, G.I., Chen, Y., Schweizer, W., Zheng, X., Contreras, M.,
-Dominguez-Bello, M.G., Blaser, M.J., 2016. Antibiotics, birth mode, and
-diet shape microbiome maturation during early life. Sci. Transl. Med. 8,
-343ra82. https://doi.org/10.1126/scitranslmed.aad7121}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Bottino,
-G.Z.M.F., Klepac-Ceraj, V., 2022. ECHO RESONANCE Microbiome.
-https://doi.org/10.17605/OSF.IO/YBS32}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Kayisire,
-A., Luo, A., Klepac-Ceraj, V., 2021. Microbiome.jl and BiobakeryUtils.jl
-- Julia packages for working with microbial community data. J Open
-Source Softw 6, 3876. https://doi.org/10.21105/joss.03876}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Breiman, L., 2001.
-Random Forests. Mach. Learn. 45, 5--32.
-https://doi.org/10.1023/A:1010933404324}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Brieuc, M.S.O.,
-Waters, C.D., Drinan, D.P., Naish, K.A., 2018. A practical introduction
-to Random Forest for genetic association studies in ecology and
-evolution. Mol. Ecol. Resour. 18, 755--766.
-https://doi.org/10.1111/1755-0998.12773}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bruchhage, Muriel,
-Chen, Y., Cataldo, A.G., Müller, H.-G., Weisee, E., Wilson, S., Pietsch,
-M., D'Sa, V., Marquand, A., Madhow, S., Bouchard, K., Cole, J.H.,
-Biondo, F., Elison, J., O'Muirchheartaigh, J., Deoni, C, S., n.d.
-Longitudinal brain and cognitive development of the first 1000 Days:A
-large Multi-Cohort Multi-Scanner study. Presented at the International
-Society of Magnetic Resonance Medicine.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bush, G., Vogt, B.A.,
-Holmes, J., Dale, A.M., Greve, D., Jenike, M.A., Rosen, B.R., 2002.
-Dorsal anterior cingulate cortex: A role in reward-based decision
-making. Proc. Natl. Acad. Sci. 99, 523--528.
-https://doi.org/10.1073/pnas.012470999}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Callaghan, B., 2020.
-Nested sensitive periods: how plasticity across the microbiota-gut-brain
-axis interacts to affect the development of learning and memory. Curr.
-Opin. Behav. Sci. 36, 55--62.
-https://doi.org/10.1016/j.cobeha.2020.07.011}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cao, Z.,
-Ottino-Gonzalez, J., Cupertino, R.B., Schwab, N., Hoke, C., Catherine,
-O., Cousijn, J., Dagher, A., Foxe, J.J., Goudriaan, A.E., Hester, R.,
-Hutchison, K., Li, C.-S.R., London, E.D., Lorenzetti, V., Luijten, M.,
-Martin-Santos, R., Momenan, R., Paulus, M.P., Schmaal, L., Sinha, R.,
-Sjoerds, Z., Solowij, N., Stein, D.J., Stein, E.A., Uhlmann, A., van
-Holst, R.J., Veltman, D.J., Wiers, R.W., Yücel, M., Zhang, S.,
-Jahanshad, N., Thompson, P.M., Conrod, P., Mackey, S., Garavan, H.,
-2021. Mapping cortical and subcortical asymmetries in substance
-dependence: Findings from the ENIGMA Addiction Working Group. Addict.
-Biol. 26, e13010. https://doi.org/10.1111/adb.13010}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cerdó, T., Diéguez,
-E., Campoy, C., 2019. Early nutrition and gut microbiome:
-interrelationship between bacterial metabolism, immune system, brain
-structure, and neurodevelopment. Am. J. Physiol. Endocrinol. Metab. 317,
-E617--E630. https://doi.org/10.1152/ajpendo.00188.2019}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Chen, X., Ishwaran,
-H., 2012. Random forests for genomic data analysis. Genomics 99,
-323--329. https://doi.org/10.1016/j.ygeno.2012.04.003}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cohen Kadosh, K.,
-Krause, B., King, A.J., Near, J., Cohen Kadosh, R., 2015. Linking GABA
-and glutamate levels to cognitive skill acquisition during development.
-Hum. Brain Mapp. 36, 4334--4345. https://doi.org/10.1002/hbm.22921}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cowan, C.S.M., Dinan,
-T.G., Cryan, J.F., 2020. Annual Research Review: Critical windows - the
-microbiota-gut-brain axis in neurocognitive development. J. Child
-Psychol. Psychiatry 61, 353--371. https://doi.org/10.1111/jcpp.13156}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dalile, B., Van
-Oudenhove, L., Vervliet, B., Verbeke, K., 2019. The role of short-chain
-fatty acids in microbiota-gut-brain communication. Nat. Rev.
-Gastroenterol. Hepatol. 16, 461--478.
-https://doi.org/10.1038/s41575-019-0157-3}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Danisch, S.,
-Krumbiegel, J., 2021. Makie.jl: Flexible high-performance data
-visualization for Julia. J. Open Source Softw. 6, 3349.
-https://doi.org/10.21105/joss.03349}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Deoni, S.C.,
-Beauchemin, J., Volpe, A., Dâ Sa, V., RESONANCE Consortium, 2021. Impact
-of the COVID-19 pandemic on early child cognitive development: Initial
-findings in a longitudinal observational study of child health. MedRxiv
-Prepr. Serv. Health Sci. https://doi.org/10.1101/2021.08.10.21261846}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dominguez-Bello, M.G.,
-Costello, E.K., Contreras, M., Magris, M., Hidalgo, G., Fierer, N.,
-Knight, R., 2010. Delivery mode shapes the acquisition and structure of
-the initial microbiota across multiple body habitats in newborns. Proc.
-Natl. Acad. Sci. U. S. A. 107, 11971--11975.
-https://doi.org/10.1073/pnas.1002601107}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dong, H., Wang, B.,
-Feng, J., Yue, X., Jia, F., 2021. Correlation between serum
-concentrations of menaquinone-4 and developmental quotients in children
-with autism spectrum disorder. Front. Nutr. 8, 748513.
-https://doi.org/10.3389/fnut.2021.748513}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ernst, M., Nelson,
-E.E., Jazbec, S., McClure, E.B., Monk, C.S., Leibenluft, E., Blair, J.,
-Pine, D.S., 2005. Amygdala and nucleus accumbens in responses to receipt
-and omission of gains in adults and adolescents. NeuroImage 25,
-1279--1291. https://doi.org/10.1016/j.neuroimage.2004.12.038}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Forrest, C.B.,
-Blackwell, C.K., Camargo, C.A., 2018. Advancing the science of
-children's positive health in the national institutes of health
-environmental influences on child health outcomes (ECHO) research
-program. J. Pediatr. 196, 298--300.
-https://doi.org/10.1016/j.jpeds.2018.02.004}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Franzosa, E.A.,
-Sirota-Madi, A., Avila-Pacheco, J., Fornelos, N., Haiser, H.J., Reinker,
-S., Vatanen, T., Hall, A.B., Mallick, H., McIver, L.J., Sauk, J.S.,
-Wilson, R.G., Stevens, B.W., Scott, J.M., Pierce, K., Deik, A.A.,
-Bullock, K., Imhann, F., Porter, J.A., Zhernakova, A., Fu, J., Weersma,
-R.K., Wijmenga, C., Clish, C.B., Vlamakis, H., Huttenhower, C., Xavier,
-R.J., 2019. Gut microbiome structure and metabolic activity in
-inflammatory bowel disease. Nat. Microbiol. 4, 293--305.
-https://doi.org/10.1038/s41564-018-0306-4}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Fung, T.C., Olson,
-C.A., Hsiao, E.Y., 2017. Interactions between the microbiota, immune and
-nervous systems in health and disease. Nat. Neurosci. 20, 145--155.
-https://doi.org/10.1038/nn.4476}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ghosh, T.S., Rampelli,
-S., Jeffery, I.B., Santoro, A., Neto, M., Capri, M., Giampieri, E.,
-Jennings, A., Candela, M., Turroni, S., Zoetendal, E.G., Hermes, G.D.A.,
-Elodie, C., Meunier, N., Brugere, C.M., Pujos-Guillot, E., Berendsen,
-A.M., Groot, L.C.P.G.M.D., Feskins, E.J.M., Kaluza, J., Pietruszka, B.,
-Bielak, M.J., Comte, B., Maijo-Ferre, M., Nicoletti, C., Vos, W.M.D.,
-Fairweather-Tait, S., Cassidy, A., Brigidi, P., Franceschi, C., O'Toole,
-P.W., 2020. Mediterranean diet intervention alters the gut microbiome in
-older people reducing frailty and improving health status: the NU-AGE
-1-year dietary intervention across five European countries. Gut 69,
-1218--1228. https://doi.org/10.1136/gutjnl-2019-319654}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Gillman, M.W.,
-Blaisdell, C.J., 2018. Environmental influences on child health
-outcomes, a research program of the NIH. Curr. Opin. Pediatr. 30, 260.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Gong, Q.-Y., Cai, L.,
-Jing, Y., Wang, W., Yang, D.-X., Chen, S.-W., Tian, H.-L., 2022.
-Urolithin A alleviates blood-brain barrier disruption and attenuates
-neuronal apoptosis following traumatic brain injury in mice. Neural
-Regen. Res. 17, 2007. https://doi.org/10.4103/1673-5374.335163}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Haam, J., Yakel, J.L.,
-2017. Cholinergic modulation of the hippocampal region and memory
-function. J. Neurochem. 142, 111--121.
-https://doi.org/10.1111/jnc.14052}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Haiser, H.J.,
-Gootenberg, D.B., Chatman, K., Sirasani, G., Balskus, E.P., Turnbaugh,
-P.J., 2013. Predicting and Manipulating Cardiac Drug Inactivation by the
-Human Gut Bacterium Eggerthella lenta. Science 341, 295--298.
-https://doi.org/10.1126/science.1235872}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hartstra, A.V.,
-Schüppel, V., Imangaliyev, S., Schrantee, A., Prodan, A., Collard, D.,
-Levin, E., Dallinga-Thie, G., Ackermans, M.T., Winkelmeijer, M., Havik,
-S.R., Metwaly, A., Lagkouvardos, I., Nier, A., Bergheim, I.,
-Heikenwalder, M., Dunkel, A., Nederveen, A.J., Liebisch, G., Mancano,
-G., Claus, S.P., Benítez-Páez, A., la Fleur, S.E., Bergman, J.J.,
-Gerdes, V., Sanz, Y., Booij, J., Kemper, E., Groen, A.K., Serlie, M.J.,
-Haller, D., Nieuwdorp, M., 2020. Infusion of donor feces affects the
-gut--brain axis in humans with metabolic syndrome. Mol. Metab. 42,
-101076. https://doi.org/10.1016/j.molmet.2020.101076}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoogman, M., Bralten,
-J., Hibar, D.P., Mennes, M., Zwiers, M.P., Schweren, L., van Hulzen,
-K.J.E., Medland, S.E., Shumskaya, E., Jahanshad, N., de Zeeuw, P.,
-Szekely, E., Sudre, G., Wolfers, T., Onnink, A.M.H., Dammers, J.T.,
-Mostert, J.C., Vives-Gilabert, Y., Kohls, G., Oberwelland, E., Seitz,
-J., Schulte-Rüther, M., di Bruttopilo, S.A., Doyle, A.E., Høvik, M.F.,
-Dramsdahl, M., Tamm, L., van Erp, T.G.M., Dale, A., Schork, A.,
-Conzelmann, A., Zierhut, K., Baur, R., McCarthy, H., Yoncheva, Y.N.,
-Cubillo, A., Chantiluke, K., Mehta, M.A., Paloyelis, Y., Hohmann, S.,
-Baumeister, S., Bramati, I., Mattos, P., Tovar-Moll, F., Douglas, P.,
-Banaschewski, T., Brandeis, D., Kuntsi, J., Asherson, P., Rubia, K.,
-Kelly, C., Di Martino, A., Milham, M.P., Castellanos, F.X., Frodl, T.,
-Zentis, M., Lesch, K.-P., Reif, A., Pauli, P., Jernigan, T., Haavik, J.,
-Plessen, K.J., Lundervold, A.J., Hugdahl, K., Seidman, L.J., Biederman,
-J., Rommelse, N., Heslenfeld, D.J., Hartman, C., Hoekstra, P.J.,
-Oosterlaan, J., von Polier, G., Konrad, K., Vilarroya, O.,
-Ramos-Quiroga, J.-A., Soliva, J.C., Durston, S., Buitelaar, J.K.,
-Faraone, S.V., Shaw, P., Thompson, P., Franke, B., 2017. Subcortical
-brain volume differences of participants with ADHD across the lifespan:
-an ENIGMA collaboration. Lancet Psychiatry 4, 310--319.
-https://doi.org/10.1016/S2215-0366(17)30049-4}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoshino, Y., Yamamoto,
-T., Kaneko, M., Tachibana, R., Watanabe, M., Ono, Y., Kumashiro, H.,
-1984. Blood Serotonin and Free Tryptophan Concentration in Autistic
-Children. Neuropsychobiology 11, 22--27.
-https://doi.org/10.1159/000118045}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hosomi, K., Saito, M.,
-Park, J., Murakami, H., Shibata, N., Ando, M., Nagatake, T., Konishi,
-K., Ohno, H., Tanisawa, K., Mohsen, A., Chen, Y.-A., Kawashima, H.,
-Natsume-Kitatani, Y., Oka, Y., Shimizu, H., Furuta, M., Tojima, Y.,
-Sawane, K., Saika, A., Kondo, S., Yonejima, Y., Takeyama, H., Matsutani,
-A., Mizuguchi, K., Miyachi, M., Kunisawa, J., 2022. Oral administration
-of Blautia wexlerae ameliorates obesity and type 2 diabetes via
-metabolic remodeling of the gut microbiota. Nat. Commun. 13, 4477.
-https://doi.org/10.1038/s41467-022-32015-7}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Jenkinson, M.,
-Beckmann, C.F., Behrens, T.E.J., Woolrich, M.W., Smith, S.M., 2012. FSL.
-Neuroimage 62, 782--790.
-https://doi.org/10.1016/j.neuroimage.2011.09.015}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Jiang, H., Ling, Z.,
-Zhang, Y., Mao, H., Ma, Z., Yin, Y., Wang, W., Tang, W., Tan, Z., Shi,
-J., Li, L., Ruan, B., 2015. Altered fecal microbiota composition in
-patients with major depressive disorder. Brain. Behav. Immun. 48,
-186--194. https://doi.org/10.1016/j.bbi.2015.03.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kanehisa, M., Goto,
-S., Kawashima, S., Okuno, Y., Hattori, M., 2004. The KEGG resource for
-deciphering the genome. Nucleic Acids Res. 32, D277--D280.
-https://doi.org/10.1093/nar/gkh063}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kant, R., Rasinkangas,
-P., Satokari, R., Pietilä, T.E., Palva, A., 2015. Genome Sequence of the
-Butyrate-Producing Anaerobic Bacterium Anaerostipes hadrus PEL 85.
-Genome Announc. 3, e00224-15. https://doi.org/10.1128/genomeA.00224-15}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kim, C.-S., Cha, L.,
-Sim, M., Jung, S., Chun, W.Y., Baik, H.W., Shin, D.-M., 2021. Probiotic
-supplementation improves cognitive function and mood with changes in gut
-microbiota in Community-Dwelling older adults: A randomized,
-Double-Blind, Placebo-Controlled, multicenter trial. J Gerontol Biol Sci
-Med Sci 76, 32--40. https://doi.org/10.1093/gerona/glaa090}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Knudsen, E.I., 2004.
-Sensitive periods in the development of the brain and behavior. J. Cogn.
-Neurosci. 16, 1412--1425. https://doi.org/10.1162/0898929042304796}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Koenig, J.E., Spor,
-A., Scalfone, N., Fricker, A.D., Stombaugh, J., Knight, R., Angenent,
-L.T., Ley, R.E., 2011. Succession of microbial consortia in the
-developing infant gut microbiome. Proc. Natl. Acad. Sci. U. S. A. 108
-Suppl 1, 4578--4585. https://doi.org/10.1073/pnas.1000081107}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Coker,
-M.O., Madan, J.C., 2022. The developing microbiome from birth to 3
-years: The Gut-Brain axis and neurodevelopmental outcomes. Front.
-Pediatr. 10, 815885. https://doi.org/10.3389/fped.2022.815885}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Korrick,
-S.A., Baker, E.R., Karagas, M.R., Madan, J.C., 2020. Prospective
-associations of the infant gut microbiome and microbial function with
-social behaviors related to autism at age 3 years. Sci. Rep. 10, 15515.
-https://doi.org/10.1038/s41598-020-72386-9}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ley, R.E., Turnbaugh,
-P.J., Klein, S., Gordon, J.I., 2006. Human gut microbes associated with
-obesity. Nature 444, 1022--1023. https://doi.org/10.1038/4441022a}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Li, Y.-M., Yu, Q.,
-Huang, L., Fu, J., Feng, R.-J., 2022. {[}Correlations Between Gut
-Microbiota Changes and Cognitive Function in Patients with Post-Stroke
-Cognitive Impairment in the Early Stage{]}. Sichuan Da Xue Xue Bao Yi
-Xue Ban 53, 857--865. https://doi.org/10.12182/20220960105}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, S., Li, E., Sun,
-Z., Fu, D., Duan, G., Jiang, M., Yu, Y., Mei, L., Yang, P., Tang, Y.,
-Zheng, P., 2019. Altered gut microbiota and short chain fatty acids in
-Chinese children with autism spectrum disorder. Sci. Rep. 9, 287.
-https://doi.org/10.1038/s41598-018-36430-z}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, X., Mao, B., Gu,
-J., Wu, J., Cui, S., Wang, G., Zhao, J., Zhang, H., Chen, W., 2021.
-\emph{Blautia} ---a new functional genus with potential probiotic
-properties? Gut Microbes 13, 1875796.
-https://doi.org/10.1080/19490976.2021.1875796}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Lopez-Siles, M.,
-Duncan, S.H., Garcia-Gil, L.J., Martinez-Medina, M., 2017.
-Faecalibacterium prausnitzii: from microbiology to diagnostics and
-prognostics. ISME J. 11, 841--852.
-https://doi.org/10.1038/ismej.2016.176}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Louwies, T., Johnson,
-A.C., Orock, A., Yuan, T., Greenwood-Van Meerveld, B., 2020. The
-microbiota-gut-brain axis: An emerging role for the epigenome. Exp.
-Biol. Med. 245, 138--145. https://doi.org/10.1177/1535370219891690}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Magnusson, K.R.,
-Hauck, L., Jeffrey, B.M., Elias, V., Humphrey, A., Nath, R., Perrone,
-A., Bermudez, L.E., 2015. Relationships between diet-related changes in
-the gut microbiome and cognitive flexibility. Neuroscience 300,
-128--140. https://doi.org/10.1016/j.neuroscience.2015.05.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mallick, H.,
-Rahnavard, A., McIver, L.J., Ma, S., Zhang, Y., Nguyen, L.H., Tickle,
-T.L., Weingart, G., Ren, B., Schwager, E.H., Chatterjee, S., Thompson,
-K.N., Wilkinson, J.E., Subramanian, A., Lu, Y., Waldron, L., Paulson,
-J.N., Franzosa, E.A., Bravo, H.C., Huttenhower, C., 2021. Multivariable
-association discovery in population-scale meta-omics studies.
-https://doi.org/10.1101/2021.01.20.427420}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Maruo, T., Sakamoto,
-M., Ito, C., Toda, T., Benno, Y., 2008. Adlercreutzia equolifaciens gen.
-nov., sp. nov., an equol-producing bacterium isolated from human faeces,
-and emended description of the genus Eggerthella. Int. J. Syst. Evol.
-Microbiol. 58, 1221--1227. https://doi.org/10.1099/ijs.0.65404-0}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mayneris-Perxachs, J.,
-Castells-Nobau, A., Arnoriaga-Rodríguez, M., Martin, M., de la
-Vega-Correa, L., Zapata, C., Burokas, A., Blasco, G., Coll, C.,
-Escrichs, A., Biarnés, C., Moreno-Navarrete, J.M., Puig, J., Garre-Olmo,
-J., Ramos, R., Pedraza, S., Brugada, R., Vilanova, J.C., Serena, J.,
-Gich, J., Ramió-Torrentà, L., Pérez-Brocal, V., Moya, A., Pamplona, R.,
-Sol, J., Jové, M., Ricart, W., Portero-Otin, M., Deco, G., Maldonado,
-R., Fernández-Real, J.M., 2022. Microbiota alterations in proline
-metabolism impact depression. Cell Metab. 34, 681-701.e10.
-https://doi.org/10.1016/j.cmet.2022.04.001}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mukherjee, A., Lordan,
-C., Ross, R.P., Cotter, P.D., 2020. Gut microbes from the
-phylogenetically diverse genus Eubacterium and their various
-contributions to gut health. Gut Microbes 12, 1802866.
-https://doi.org/10.1080/19490976.2020.1802866}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mullen, E.M., others,
-1995. Mullen scales of early learning. AGS Circle Pines, MN.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Needham, B.D.,
-Funabashi, M., Adame, M.D., Wang, Z., Boktor, J.C., Haney, J., Wu,
-W.-L., Rabut, C., Ladinsky, M.S., Hwang, S.-J., Guo, Y., Zhu, Q.,
-Griffiths, J.A., Knight, R., Bjorkman, P.J., Shapiro, M.G., Geschwind,
-D.H., Holschneider, D.P., Fischbach, M.A., Mazmanian, S.K., 2022. A
-gut-derived metabolite alters brain activity and anxiety behaviour in
-mice. Nature 602, 647--653. https://doi.org/10.1038/s41586-022-04396-8}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Opel, N., Thalamuthu,
-A., Milaneschi, Y., Grotegerd, D., Flint, C., Leenings, R., Goltermann,
-J., Richter, M., Hahn, T., Woditsch, G., Berger, K., Hermesdorf, M.,
-McIntosh, A., Whalley, H.C., Harris, M.A., MacMaster, F.P., Walter, H.,
-Veer, I.M., Frodl, T., Carballedo, A., Krug, A., Nenadic, I., Kircher,
-T., Aleman, A., Groenewold, N.A., Stein, D.J., Soares, J.C.,
-Zunta-Soares, G.B., Mwangi, B., Wu, M.-J., Walter, M., Li, M., Harrison,
-B.J., Davey, C.G., Cullen, K.R., Klimes-Dougan, B., Mueller, B.A.,
-Sämann, P.G., Penninx, B., Nawijn, L., Veltman, D.J., Aftanas, L., Brak,
-I.V., Filimonova, E.A., Osipov, E.A., Reneman, L., Schrantee, A., Grabe,
-H.J., Van der Auwera, S., Wittfeld, K., Hosten, N., Völzke, H., Sim, K.,
-Gotlib, I.H., Sacchet, M.D., Lagopoulos, J., Hatton, S.N., Hickie, I.,
-Pozzi, E., Thompson, P.M., Jahanshad, N., Schmaal, L., Baune, B.T.,
-Dannlowski, U., 2021. Brain structural abnormalities in obesity:
-relation to age, genetic risk, and common psychiatric disorders. Mol.
-Psychiatry 26, 4839--4852. https://doi.org/10.1038/s41380-020-0774-9}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Pagonabarraga, J.,
-Corcuera-Solano, I., Vives-Gilabert, Y., Llebaria, G., García-Sánchez,
-C., Pascual-Sedano, B., Delfino, M., Kulisevsky, J., Gómez-Ansón, B.,
-2013. Pattern of Regional Cortical Thinning Associated with Cognitive
-Deterioration in Parkinson's Disease. PLoS ONE 8, e54980.
-https://doi.org/10.1371/journal.pone.0054980}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Palomo-Buitrago, M.E.,
-Sabater-Masdeu, M., Moreno-Navarrete, J.M., Caballano-Infantes, E.,
-Arnoriaga-Rodríguez, M., Coll, C., Ramió, L., Palomino-Schätzlein, M.,
-Gutiérrez-Carcedo, P., Pérez-Brocal, V., Simó, R., Moya, A., Ricart, W.,
-Herance, J.R., Fernández-Real, J.M., 2019. Glutamate interactions with
-obesity, insulin resistance, cognition and gut microbiota composition.
-Acta Diabetol. 56, 569--579. https://doi.org/10.1007/s00592-019-01313-w}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Pronovost, G.N.,
-Hsiao, E.Y., 2019. Perinatal interactions between the microbiome,
-immunity, and neurodevelopment. Immunity 50, 18--36.
-https://doi.org/10.1016/j.immuni.2018.11.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sadeghi, B.,
-Chiarawongse, P., Squire, K., Jones, D.C., Noack, A., St-Jean, C.,
-Huijzer, R., Schätzle, R., Butterworth, I., Peng, Y.-F., Blaom, A.,
-2022. DecisionTree.jl - A julia implementation of the CART decision tree
-and random forest algorithms. https://doi.org/10.5281/zenodo.7359268}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sela, D.A., Chapman,
-J., Adeuya, A., Kim, J.H., Chen, F., Whitehead, T.R., Lapidus, A.,
-Rokhsar, D.S., Lebrilla, C.B., German, J.B., Price, N.P., Richardson,
-P.M., Mills, D.A., 2008. The genome sequence of \emph{Bifidobacterium
-longum} subsp. \emph{infantis} reveals adaptations for milk utilization
-within the infant microbiome. Proc. Natl. Acad. Sci. 105, 18964--18969.
-https://doi.org/10.1073/pnas.0809584105}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Selma, M.V., Beltrán,
-D., García-Villalba, R., Espín, J.C., Tomás-Barberán, F.A., 2014.
-Description of urolithin production capacity from ellagic acid of two
-human intestinal Gordonibacter species. Food Funct. 5, 1779--1784.
-https://doi.org/10.1039/c4fo00092g}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Setchell, K.D.,
-Clerici, C., Lephart, E.D., Cole, S.J., Heenan, C., Castellani, D.,
-Wolfe, B.E., Nechemias-Zimmer, L., Brown, N.M., Lund, T.D., Handa, R.J.,
-Heubi, J.E., 2005. S-Equol, a potent ligand for estrogen receptor β, is
-the exclusive enantiomeric form of the soy isoflavone metabolite
-produced by human intestinal bacterial flora. Am. J. Clin. Nutr. 81,
-1072--1079. https://doi.org/10.1093/ajcn/81.5.1072}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sharon, G., Sampson,
-T.R., Geschwind, D.H., Mazmanian, S.K., 2016. The central nervous system
-and the gut microbiome. Cell 167, 915--932.
-https://doi.org/10.1016/j.cell.2016.10.027}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Silbereis, J.C.,
-Pochareddy, S., Zhu, Y., Li, M., Sestan, N., 2016. The cellular and
-molecular landscapes of the developing human central nervous system.
-Neuron 89, 248--268.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sordillo, J.E.,
-Korrick, S., Laranjo, N., Carey, V., Weinstock, G.M., Gold, D.R.,
-O'Connor, G., Sandel, M., Bacharier, L.B., Beigelman, A., Zeiger, R.,
-Litonjua, A.A., Weiss, S.T., 2019. Association of the infant gut
-microbiome with early childhood neurodevelopmental outcomes: An
-ancillary study to the VDAART randomized clinical trial. JAMA Netw. Open
-2, e190905. https://doi.org/10.1001/jamanetworkopen.2019.0905}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Spichak, S.,
-Bastiaanssen, T.F.S., Berding, K., Vlckova, K., Clarke, G., Dinan, T.G.,
-Cryan, J.F., 2021. Mining microbes for mental health: Determining the
-role of microbial metabolic pathways in human brain health and disease.
-Neurosci. Biobehav. Rev. 125, 698--761.
-https://doi.org/10.1016/j.neubiorev.2021.02.044}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Stephan, J., Stegle,
-O., Beyer, A., 2015. A random forest approach to capture genetic effects
-in the presence of population structure. Nat. Commun. 6, 7432.
-https://doi.org/10.1038/ncomms8432}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Suzek, B.E., Huang,
-H., McGarvey, P., Mazumder, R., Wu, C.H., 2007. UniRef: comprehensive
-and non-redundant UniProt reference clusters. Bioinformatics 23,
-1282--1288. https://doi.org/10.1093/bioinformatics/btm098}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Takahashi, H., Yang,
-J., Yamamoto, H., Fukuda, S., Arakawa, K., 2021. Complete Genome
-Sequence of Adlercreutzia equolifaciens subsp. celatus DSM 18785.
-Microbiol. Resour. Announc. 10, e00354-21.
-https://doi.org/10.1128/MRA.00354-21}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tau, G.Z., Peterson,
-B.S., 2010. Normal development of brain circuits. Neuropsychopharmacol.
-Off. Publ. Am. Coll. Neuropsychopharmacol. 35, 147--168.
-https://doi.org/10.1038/npp.2009.115}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Thawornkuno, C.,
-Tanaka, M., Sone, T., Asano, K., 2009. Biotransformation of Daidzein to
-Equol by Crude Enzyme from Asaccharobacter celatus AHU1763 Required an
-Anaerobic Environment. Biosci. Biotechnol. Biochem. 73, 1435--1438.
-https://doi.org/10.1271/bbb.80908}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tognini, P., 2017. Gut
-microbiota: a potential regulator of neurodevelopment. Front. Cell.
-Neurosci. 11, 25. https://doi.org/10.3389/fncel.2017.00025}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tso, L., Bonham, K.S.,
-Fishbein, A., Rowland, S., Klepac-Ceraj, V., 2021. Targeted
-High-Resolution taxonomic identification of bifidobacterium longum
-subsp. infantis using human milk oligosaccharide metabolizing genes.
-Nutrients 13, 2833. https://doi.org/10.3390/nu13082833}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Valles-Colomer, M.,
-Falony, G., Darzi, Y., Tigchelaar, E.F., Wang, J., Tito, R.Y., Schiweck,
-C., Kurilshikov, A., Joossens, M., Wijmenga, C., Claes, S., Van
-Oudenhove, L., Zhernakova, A., Vieira-Silva, S., Raes, J., 2019. The
-neuroactive potential of the human gut microbiota in quality of life and
-depression. Nat. Microbiol. 4, 623--632.
-https://doi.org/10.1038/s41564-018-0337-x}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Velazquez, M., Lee,
-Y., Initiative, for the A.D.N., 2021. Random forest model for
-feature-based Alzheimer's disease conversion prediction from early mild
-cognitive impairment subjects. PLOS ONE 16, e0244773.
-https://doi.org/10.1371/journal.pone.0244773}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wacker, J., Dillon,
-D.G., Pizzagalli, D.A., 2009. The role of the nucleus accumbens and
-rostral anterior cingulate cortex in anhedonia: Integration of resting
-EEG, fMRI, and volumetric techniques. NeuroImage 46, 327--337.
-https://doi.org/10.1016/j.neuroimage.2009.01.058}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wan, Y., Zuo, T., Xu,
-Z., Zhang, F., Zhan, H., Chan, D., Leung, T.-F., Yeoh, Y.K., Chan,
-F.K.L., Chan, R., Ng, S.C., 2021. Underdevelopment of the gut microbiota
-and bacteria species as non-invasive markers of prediction in children
-with autism spectrum disorder. Gut gutjnl-2020-324015.
-https://doi.org/10.1136/gutjnl-2020-324015}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, L.-J., Yang,
-C.-Y., Chou, W.-J., Lee, M.-J., Chou, M.-C., Kuo, H.-C., Yeh, Y.-M.,
-Lee, S.-Y., Huang, L.-H., Li, S.-C., 2020. Gut microbiota and dietary
-patterns in children with attention-deficit/hyperactivity disorder. Eur.
-Child Adolesc. Psychiatry 29, 287--297.
-https://doi.org/10.1007/s00787-019-01352-2}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, X.-L., Hur,
-H.-G., Lee, J.H., Kim, K.T., Kim, S.-I., 2005. Enantioselective
-Synthesis of S-Equol from Dihydrodaidzein by a Newly Isolated Anaerobic
-Human Intestinal Bacterium. Appl. Environ. Microbiol. 71, 214--219.
-https://doi.org/10.1128/AEM.71.1.214-219.2005}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 2012.
-Wechsler preschool and primary scale of intelligence---fourth edition.
-Psychol. Corp. San Antonio TX.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 1949.
-Wechsler intelligence scale for children.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, L., Yan, J.,
-Yang, T., Zhu, J., Li, T., Wei, H., Chen, J., 2021. Fecal Microbiome
-Transplantation from Children with Autism Spectrum Disorder Modulates
-Tryptophan and Serotonergic Synapse Metabolism and Induces Altered
-Behaviors in Germ-Free Mice. mSystems 6, e01343-20.
-https://doi.org/10.1128/mSystems.01343-20}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, X., Yang, Y.,
-Zhang, Y., Zhang, X.-M., Zhao, Z.-Q., Zhang, Y.-Q., 2013. Estrogen in
-the Anterior Cingulate Cortex Contributes to Pain-Related Aversion.
-Cereb. Cortex 23, 2190--2203. https://doi.org/10.1093/cercor/bhs201}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Yael, D., Tahary, O.,
-Gurovich, B., Belelovsky, K., Bar-Gad, I., 2019. Disinhibition of the
-Nucleus Accumbens Leads to Macro-Scale Hyperactivity Consisting of
-Micro-Scale Behavioral Segments Encoded by Striatal Activity. J.
-Neurosci. 39, 5897--5909.
-https://doi.org/10.1523/JNEUROSCI.3120-18.2019}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Yau, W.-Y.W., Zubieta,
-J.-K., Weiland, B.J., Samudra, P.G., Zucker, R.A., Heitzeg, M.M., 2012.
-Nucleus Accumbens Response to Incentive Stimuli Anticipation in Children
-of Alcoholics: Relationships with Precursive Behavioral Risk and
-Lifetime Alcohol Use. J. Neurosci. 32, 2544--2551.
-https://doi.org/10.1523/JNEUROSCI.1390-11.2012}
-
-Supplementary Tables
-
-\textbf{Supplementary Table XXA. List of most important taxa for Random
-Forest models on the 0 to 6 months age bracket, microbiome +
-demographics, ordered by relative weighted importance rank}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Taxon
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Rank
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Relative weighted importance
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Cumulative
-
-relative weighted importance
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\textbf{Supplementary Table XXA. List of most important taxa for Random
-Forest models on the 18 to 120 months age bracket, microbiome +
-demographics, ordered by relative weighted importance rank}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Taxon
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Rank
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Relative weighted importance
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Cumulative
-
-relative weighted importance
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\textbf{Supplementary Table XX. Benchmark metrics for each brain segment
-prediction experiment. Confidence intervals are calculated from the
-distribution of metrics from repeated CV at a confidence level of 95\%}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3933}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2440}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1846}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1782}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Brain segment
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Root-mean-square Error
-
-(RMSE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Mean Absolute Proportional Error (MAPE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Correlation coefficient (R)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Brain-stem
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0876
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.605
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-I-V
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0321
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.085
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.156
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-VI-VII
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0132
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.081
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.006
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-VIII-X
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0217
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.078
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-CSF
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0193
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.162
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.071
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-accumbens-area
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0020
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.094
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.288
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-accumbens-area
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0020
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.102
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.041
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-amygdala
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0037
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.288
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-amygdala
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.070
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.179
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-anterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0375
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.069
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.092
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-anterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0264
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.074
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.245
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-basal-forebrain
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0008
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-basal-forebrain
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0010
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.117
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.123
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-caudate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0257
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.099
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.029
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-caudate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0221
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.010
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-cerebellum-white-matter
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0689
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.413
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-cerebellum-white-matter
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0694
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.070
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.434
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-cuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0274
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.234
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-cuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0264
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.075
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.065
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-entorhinal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0107
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.080
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.190
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-entorhinal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0105
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.099
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-fusiform
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0510
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.333
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-fusiform
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0334
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.053
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.248
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-hippocampus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0128
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.054
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.075
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-hippocampus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0150
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.094
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-insula
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0294
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.119
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-insula
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0303
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-isthmus-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0161
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.041
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-isthmus-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0166
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.085
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.074
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-lateral-occipital
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0661
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.062
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.224
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-lateral-occipital
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0716
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.054
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.221
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-lingual
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0410
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.069
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.421
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-lingual
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0440
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.072
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.434
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-middle-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0875
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.052
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.000
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-middle-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0636
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.039
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-orbitofrontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0715
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.055
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.186
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-orbitofrontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0814
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.133
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pallidum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0075
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.064
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.297
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pallidum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0081
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.264
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-paracentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0335
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.078
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.131
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-paracentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0364
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.084
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-parahippocampal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0098
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.058
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-parahippocampal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0092
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.060
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-parietal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0755
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.177
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-parietal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1099
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.045
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.147
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-opercularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0197
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.146
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-opercularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0261
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.060
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.030
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-orbitalis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0157
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.003
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-orbitalis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0233
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.112
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.024
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-triangularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0226
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.174
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-triangularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0322
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.076
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.043
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pericalcarine
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0178
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.111
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.200
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pericalcarine
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0185
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.089
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.273
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-postcentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0495
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.052
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.168
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-postcentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0552
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.135
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-posterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0460
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.137
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-posterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0480
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.139
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-precentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0830
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.065
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.248
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-precentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0594
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.046
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.066
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-precuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0489
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.144
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-precuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0622
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.150
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-putamen
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0243
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-putamen
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0242
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.066
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.011
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-superior-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0867
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.040
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.101
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-superior-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.044
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.096
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-supramarginal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0767
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-supramarginal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0734
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.001
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-temporal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1530
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.040
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.128
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-temporal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1503
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.039
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-thalamus-proper
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0347
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.053
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.170
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-thalamus-proper
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0304
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.050
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.162
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\end{document}

### 2023-02-17 manuscript/tex/sections/intro.tex
clean up

--- 2023-02-16 manuscript/tex/converted.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-]{article}
-\usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\usepackage{longtable,booktabs,array}
-\usepackage{multirow}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\usepackage[normalem]{ulem}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-
-\author{}
-\date{}
-
-\begin{document}
-
-\hypertarget{title}{%
-\subsection{Title}\label{title}}
-
-\begin{itemize}
-\item
-  \begin{quote}
-  Gut microbes and their genes are associated with cognition and
-  neuroanatomy in children
-  \end{quote}
-\item
-  \begin{quote}
-  Gut microbes and their genes are associated with cognition performance
-  and brain structure in children
-  \end{quote}
-\item
-  \begin{quote}
-  Extensive relationships between the gut microbiota and normal
-  neurocognitive development
-  \end{quote}
-\item
-  \begin{quote}
-  Extensive relationships between the gut microbiota and typical
-  neurocognitive development
-  \end{quote}
-\item
-  \begin{quote}
-  The gut microbial metagenome is connected to neurocognition in typical
-  child development
-  \end{quote}
-\item
-  \begin{quote}
-  The microbial potential in regulating early childhood neurocognitive
-  development
-  \end{quote}
-\end{itemize}
-
-\hypertarget{authors}{%
-\subsubsection{Authors}\label{authors}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \begin{quote}
-  Bonham 0000-0003-3200-7533
-  \end{quote}
-\item
-  \begin{quote}
-  Bottino 0000-0003-1953-1576
-  \end{quote}
-\item
-  \begin{quote}
-  McCann 0000-0002-9753-7968
-  \end{quote}
-\item
-  \begin{quote}
-  Beauchemin \{ORCID??\}
-  \end{quote}
-\item
-  \begin{quote}
-  Elizabeth Weisse \textless elizabeth.weisse2@gmail.com\textgreater{}
-  \end{quote}
-\item
-  \begin{quote}
-  \ldots{}
-  \end{quote}
-\item
-  \begin{quote}
-  Jonathan O\textquotesingle Muircheartaigh
-  \textless{}\href{mailto:jonathanom@kcl.ac.uk}{\uline{jonathanom@kcl.ac.uk}}\textgreater{}
-  \end{quote}
-\item
-  \begin{quote}
-  Huttenhower 0000-0002-1110-0096
-  \end{quote}
-\item
-  \begin{quote}
-  Muriel Bruchhage
-  \textless{}\href{mailto:mmk.bruchhage@gmail.com}{\uline{mmk.bruchhage@gmail.com}}\textgreater{}
-  0000-0001-9637-0951
-  \end{quote}
-\item
-  \begin{quote}
-  D'Sa 0000-0002-3476-8419
-  \end{quote}
-\item
-  \begin{quote}
-  Deoni 0000-0002-0633-6328
-  \end{quote}
-\item
-  \begin{quote}
-  Klepac-Ceraj 0000-0001-5387-5706
-  \end{quote}
-\end{enumerate}
-
-\hypertarget{abstract}{%
-\subsection{Abstract}\label{abstract}}
-
-The gastrointestinal tract, its resident microorganisms and the central
-nervous system are connected by biochemical signaling, also known as
-"microbiome-gut-brain-axis." Both the human brain and the gut microbiome
-have critical developmental windows in the first three years of life,
-raising the possibility that their development is co-occurring and
-likely co-dependent. Emerging evidence implicates gut microorganisms and
-microbiota composition in cognitive outcomes and neurodevelopmental
-disorders (e.g., autism and anxiety), but the influence of gut microbial
-metabolism on typical neurodevelopment has not been explored in detail.
-We investigated the relationship of the microbiome with the neuroanatomy
-and cognitive function of 361 healthy children, demonstrating that
-differences in gut microbial taxa and gene functions are associated with
-overall cognitive function as well as with differences in the size of
-multiple brain regions. Using a combination of multivariate linear
-models and machine learning (ML) models, we showed that many species,
-including \emph{Gordonibacter pamelae} and \emph{Blautia wexlerae}, were
-significantly associated with higher cognitive function, while some
-species such as \emph{Ruminococcus gnavus} were more commonly found in
-children with low cognitive scores regardless of age or maternal
-education. Microbial genes for enzymes involved in the metabolism of
-neuroactive compounds, particularly short-chain fatty acids such as
-glutamate and propionate, were also associated with cognitive function.
-In addition, ML models were able to use microbial taxa to predict the
-volume of brain regions, with particular taxa often dominating the
-feature importance metric for individual brain regions, such as \emph{B.
-wexlerae} being the most dominant for the parahippocampal region or
-GABA-producing \emph{Bacteroidetes ovatus} for left accumbens. These
-findings provide potential biomarkers of neurocognition and may lead to
-the future development of targets for early detection and early
-intervention.
-
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
-
 The gut and the brain are intimately linked. Signals from the brain
@@
 performance and brain structure.
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-\hypertarget{cohort-overview-and-summary-data}{%
-\subsubsection{Cohort overview and summary
-data}\label{cohort-overview-and-summary-data}}
-
-We investigated the co-development of the brain and the microbiome in
-early childhood in a cohort of typically-developing children in their
-first years of life using a variety of orthogonal microbial and
-neurocognitive assessments, including shotgun metagenomic sequencing,
-cognitive and behavioral assessments, and neuroimaging. In all, 361
-children from RESONANCE, an accelerated longitudinal cohort of early
-development \href{https://paperpile.com/c/dPbU4e/klxU}{(Forrest,
-Blackwell, and Camargo 2018)} between the ages of 82 days and 10 years
-old (median age 2.21 years, Supplementary Figure 1) were included in
-this study (Table 1, Figure 1A). To measure cognitive function, we used
-age-appropriate assessments that can be normalized to a common, IQ-like
-scale (Figure 1B), including the Mullen scale of early learning (MSEL)
-for children from birth to 3 years of age
-\href{https://www.zotero.org/google-docs/?90hfYr}{(Mullen and others,
-1995)}, Wechsler preschool and primary scale of intelligence (WPPSI) for
-4--5-year-olds
-\href{https://www.zotero.org/google-docs/?wCfx5h}{(Wechsler, 2012)}, and
-the Wechsler intelligence scale for children (WISC) for children 6 years
-and up \href{https://www.zotero.org/google-docs/?AcNN3S}{(Wechsler,
-1949)}.
-
-As expected, the greatest differences in microbial taxa were driven by
-age (Figure 1B-C, PCoA axis 1), while older children were primarily
-stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
-high-\emph{Prevotella copri} (Supplementary Figure 2). Overall variation
-in gut microbial genes and also in brain volume profiles was similarly
-driven largely by subject age (Figure 1C-F).
-
-\includegraphics[width=6.45833in,height=3.87221in]{media/image3.png}
-
-\textbf{Figure 1: Cohort overview and summary data.}
-
-(A) Study design and overview. Stool samples, cognitive assessments and
-neuroimaging were collected from participants in an accelerated
-longitudinal design. (B) Cognitive function scores are assessed by
-different instruments, but may be normalized to a common IQ-like scale.
-(C-D) Principal coordinates analysis using Bray-Curtis dissimilarity on
-taxonomic profiles demonstrates high beta diversity, with much of the
-first axis of variation explained by increasing age and alpha diversity.
-Differences in gene function profiles (E) and neuroimaging (PCA based on
-the Euclidean distance of brain region volumes) (F) are likewise
-dominated by changes as children age. (G) Permutation analysis of
-variance (PERMANOVA) of taxonomic profiles, functional profiles
-(annotated as UniRef90s) and neuroimaging, against the metadata of
-interest. Variations in taxonomic and functional profiles explain a
-modest but significant percent of the variation in cognitive development
-in children over 18 months of age. (H) Mantel testing of different
-microbial feature matrices, shows overlapping but distinct patterns of
-variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
-used as cut-offs in some following models.
-
-\textbf{Table1: Subject demographics}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2620}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Group}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Subgroup}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{All}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Under 6 mo}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Over 18 mo}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-N subjects
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-361
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-74
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-257
-\end{minipage} \\
-\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Samples
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-1
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-262 (72.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-74 (100.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-203 (79.0\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-72 (19.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-46 (17.9\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-\textgreater2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-27 (7.5\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-8 (3.1\%)
-\end{minipage} \\
-\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Age (months)
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-Min
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18.0
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Max
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-119.3
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6.0
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-119.3
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Median
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-25.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-3.7
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-45.9
-\end{minipage} \\
-\multirow{2}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Sex
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-F
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-163 (45.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (50.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-114 (44.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-M
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-198 (54.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (50.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-143 (55.6\%)
-\end{minipage} \\
-\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Race
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-White
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-252 (69.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-42 (56.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-191 (74.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Black
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-40 (11.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16 (21.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-21 (8.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Asian
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Mixed
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-7 (1.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (1.4\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6 (2.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Other
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-48 (13.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12 (16.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-29 (11.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Unknown/Declined
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2 (0.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (1.4\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2 (0.8\%)
-\end{minipage} \\
-\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Maternal Ed.
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-Junior high school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Some high school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6 (1.7\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-3 (4.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-4 (1.6\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-High school grad
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (10.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-13 (17.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16 (6.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Some college
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-95 (26.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-25 (33.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-57 (22.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-College grad
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-90 (24.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17 (23.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-69 (26.8\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Grad/professional school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-123 (34.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-15 (20.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-104 (40.5\%)
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-Several studies have demonstrated links between specific taxa and
-measures of anxiety and depression
-\href{https://www.zotero.org/google-docs/?X5rfKj}{(Mayneris-Perxachs et
-al., 2022; Needham et al., 2022)}, cognitive flexibility in adults
-\href{https://www.zotero.org/google-docs/?RGIf8m}{(Magnusson et al.,
-2015)}, and atypical neural development
-\href{https://www.zotero.org/google-docs/?hwPZKh}{(Liu et al., 2019; Wan
-et al., 2021)}. We, therefore, set out to identify whether specific taxa
-or gene functions were linked with normal cognitive development in
-children. In order to assess whether variations in gut microbial taxa,
-their genes, or their metabolism are linked with neurocognitive
-development, we tested whether the beta diversity of microbial taxa and
-gene functions, as well as variation in brain development as assessed by
-neuroimaging, were associated with these contemporaneous measures of
-cognitive function using permutation analysis of variance (PERMANOVA).
-Due to the large ecological shift in the microbiome that occurs upon the
-introduction of solid food, and a relatively wide range of ages when
-infants are transitioned to solid food, we considered the pre-transition
-(less than 6 months old) and post-transition (over 18 months old)
-microbiomes separately. Even still, age was a major driver of variation
-in gut microbiomes in children over 18 months old (Figure 1G). We also
-found that overall variation in microbial species in children over 18
-months old was associated with a small but significant variation in
-cognitive function score (Figure 1G, R\textsuperscript{2} = 0.0124, q
-\textless{} 0.001), as was variation in microbial gene functions
-(R\textsuperscript{2} = 0.0119, q \textless{} 0.001). Variation in
-microbial taxa and genes was not significantly associated with cognitive
-function in children under 6 months, though this may be due to the low
-taxonomic diversity, and broad lack of overlap between taxa in infants.
-As expected, age was significantly associated with microbial beta
-diversity (taxa R\textsuperscript{2} = 0.0215, and UniRef90s
-R\textsuperscript{2} = 0.0245, q \textless{} 0.001), and very strongly
-associated with neuroimaging profiles (R\textsuperscript{2} = 0.256, q
-\textless{} 0.001).
-
-Consistent with prior studies, different microbial measurement types
-captured overlapping variation, with species profiles and gene function
-profiles (annotated with clusters of 90\% similarity
-\href{https://www.zotero.org/google-docs/?b9rEJS}{(Suzek et al.,
-2007)}), both generated from DNA sequencing data, being tightly coupled
-(Figure 1H, p \textless{} 0.001). Interestingly, other functional
-groupings (Enzyme commission level4 - ECs
-\href{https://www.zotero.org/google-docs/?fs8LO7}{(Bairoch, 2000)}, and
-KEGG orthologues - KOs
-\href{https://www.zotero.org/google-docs/?Xx5duE}{(Kanehisa et al.,
-2004)}) overlapped only slightly with taxonomic profiles in both age
-cohorts, despite being derived from UniRef90 labels. In children over 18
-months, some variation (15.9\%, p \textless{} 0.01) in neuroimaging
-overlapped with microbial measures, though this may be due to the
-residual variation due to age in both measures.
-
-\hypertarget{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}{%
-\subsubsection{Microbial species and neuroactive genes are associated
-with cognitive
-performance}\label{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}}
-
-To assess whether individual microbial species were associated with
-cognitive function, we fit multivariable linear regression
-\href{https://www.zotero.org/google-docs/?4qS6UF}{(Mallick et al.,
-2021)} to the relative abundance of each species that had at least 15\%
-prevalence in a given age group (Figure 2A, N = 92 for 0--120 months, N
-= 46 for 0--6 months, N = 97 for 18--120 months). Only \emph{Blautia
-wexlerae} was significantly associated (q value = 0.14, β = 0.0015) with
-cognitive function in children under 6 months old after adjusting for
-age and maternal education (Figure 2B). \emph{B. wexlerae} was
-previously shown to be depleted in children with diabetes
-\href{https://www.zotero.org/google-docs/?6mU2nI}{(Benítez-Páez et al.,
-2020)}, and that oral administration of \emph{B. wexlerae} partially
-ameliorated weight gain and inflammation from a high-fat diet in a mouse
-model of T2D \href{https://www.zotero.org/google-docs/?y83hpS}{(Hosomi
-et al., 2022; Liu et al., 2021)}. In children over 18 months of age,
-several microbial species were significantly enriched (q-value
-\textless{} 0.20) in children with higher cognitive function scores,
-including \emph{Gordonibacter pamelaeae}, which produces the
-neuroprotective metabolite urolithin
-\href{https://www.zotero.org/google-docs/?HGtTEY}{(Gong et al., 2022;
-Selma et al., 2014)}, \emph{Asaccharobacter celatus} and
-\emph{Adelcreutzia equolifaciens}, which produce phytoestrogen-derived
-equol \href{https://www.zotero.org/google-docs/?HCKZqm}{(Maruo et al.,
-2008; Thawornkuno et al., 2009)}, and the SCFA-producing probiotic
-species such as \emph{Eubacterium eligens} and \emph{Faecalibacterium
-prausnitzii} \href{https://www.zotero.org/google-docs/?3eyheK}{(Ghosh et
-al., 2020; Lopez-Siles et al., 2017; Mukherjee et al., 2020)} (Figure
-2B).
-
-\includegraphics[width=6.5in,height=4.83333in]{media/image4.png}
-
-\textbf{Figure 2 - Taxa and gene functional groups are associated with
-cognitive function.}
-
-(A) Volcano plot of multivariable linear model results showing the
-relationship between individual taxa and cognitive function score in
-children over 18 months of age, controlling for age, maternal education,
-and sequencing depth. Taxa that were significant after FDR correction (q
-\textless{} 0.2) are colored red. (B) For taxa that were significantly
-associated with cognitive function, heatmaps of prevalence, mean
-relative abundance, and correlation with cognitive function in different
-age groups. (C-F) Enrichment plots for selected neuroactive gene sets
-used in feature set enrichment analysis (FSEA). (G) summary of FSEA
-results across all samples (left) as well as in the under 6 months
-(middle) and over 18 months (right) subsets, colored based on the
-significance of the association. Markers indicate the individual
-correlation of genes within a gene set, and vertical bars represent the
-median correlation of that gene set.
-
-Given that different microbial species might occupy the same metabolic
-niche in different individuals, we hypothesized that microbial genes
-grouped by functional activity would be associated with cognition. To
-test this, we performed feature set enrichment analysis (FSEA) on groups
-of genes with neuroactive potential
-\href{https://www.zotero.org/google-docs/?xQURsj}{(Valles-Colomer et
-al., 2019)} and concurrent cognitive function score (Table 2,
-\protect\hyperlink{_i7fd3z19jdvg}{\uline{Figure 2}}C-G) and found that
-several metabolic pathways were significantly enriched or depleted in
-children with higher cognitive function scores. This was true both when
-considering all age groups together, though the enrichment of most
-pathways was more pronounced in children under 6 months or over 18
-months. For example, genes for degrading the 3-carbon SCFA propionate
-were significantly depleted in children with higher cognitive function
-scores across all age groups tested (Table 2; propionate degradation I,
-under 6 months, enrichment score (E.S.) = -0.542, corrected p-value (q)
-= 0.020; over 18 months, E.S. = -0.674, q = 0.041). Interestingly, genes
-for propionate synthesis were also significantly depleted in higher
-scoring children over 18 months (E.S. -0.676, q = 0.023), as were genes
-for synthesizing the 2 carbon SCFA acetate in children under 6 months
-old (acetate synthesis I, E.S. = -0.194, q = 0.153; acetate synthesis
-II, E.S. = -0.342, q = 0.020; acetate synthesis III, E.S. = -0.31, q =
-0.052). SCFAs are produced by anaerobic fermentation of dietary fiber
-and have been linked with immune system regulation as well as directly
-with brain function
-\href{https://www.zotero.org/google-docs/?2Z0sKZ}{(Dalile et al.,
-2019)}.
-
-Synthesis of menaquinone (vitamin K) was also negatively associated with
-cognitive function score in older children (menaquinone synthesis I,
-E.S. = -0.170, q = 0.0183). Menaquinone has several isoforms, one of
-which, MK-4, has been found to be decreased in children diagnosed with
-ASD \href{https://www.zotero.org/google-docs/?dTA3NO}{(Dong et al.,
-2021)} and is potentially neuroprotective in both rodents and humans
-\href{https://paperpile.com/c/dPbU4e/cTO9}{(Elkattawy et al., 2022)}. In
-children over 18 months old, genes for the synthesis of the amino acids
-glutamate and tryptophan were significantly enriched in children with
-higher cognitive function scores (Glutamate synthesis I, E.S. = 0.242, q
-= 0.047; Tryptophan synthesis, E.S. = 0.119, q = 0.041). Glutamate is a
-critical neurotransmitter controlling neuronal excitatory/inhibitory
-signaling along with gamma-aminobutyric acid (GABA), and their balance
-in the brain controls neural plasticity and learning, particularly in
-the developing brain
-\href{https://www.zotero.org/google-docs/?QubMhx}{(Cohen Kadosh et al.,
-2015; Palomo-Buitrago et al., 2019)}. Tryptophan metabolism, including
-microbial metabolism of tryptophan, has previously been linked with
-autism in children
-\href{https://www.zotero.org/google-docs/?GPaUZE}{(Hoshino et al., 1984;
-Xiao et al., 2021)}. Taken together, these results suggest that
-microbial metabolic activity, particularly the metabolism (synthesis and
-degradation) of neuroactive compounds may have effects on cognitive
-development.
-
-\textbf{Table 2 - Feature set enrichment analysis on neuroactive
-microbial gene sets}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2300}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1259}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1525}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1429}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2785} + 2\tabcolsep}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{All}
-\end{minipage}} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2591} + 2\tabcolsep}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{under 6 months}
-\end{minipage}} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2324} + 2\tabcolsep}@{}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{over 18 months}
-\end{minipage}} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-geneset
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1223
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5549
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1944
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1533}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1269
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. II
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2201
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0660}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3424
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0198}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0946
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9199
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. III
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2104
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0859}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3098
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0518}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0857
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9199
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Glutamate deg. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.4037
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1802}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2840
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2489
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Glutamate syn. II
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0953
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6022
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1342
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9802
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2421
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0466}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Menaquinone syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0853
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5635
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0833
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1699
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1825}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Propionate deg. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.4612
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1302}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.5421
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0198}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.6738
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0414}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Propionate syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3344
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5549
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3201
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5713
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.6760
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0227}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Tryptophan syn.
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0573
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9742
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0725
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1186
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0414}
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\hypertarget{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}{%
-\subsubsection{Gut microbial taxonomic and functional profiles predict
-cognitive
-function}\label{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}}
-
-FSEA relies on understanding functional relationships between individual
-genes. However, because the relationships between individual taxa are
-still largely unknown, we turned to Random Forest (RF) models, an
-unsupervised non-parametric machine learning (ML) method that enables
-the identification of underlying patterns in large numbers of individual
-features (here, microbial species). Previous studies have reported
-successful use of RFs for processing highly-dimensional and sparse data
-from the domain of genomics
-\href{https://www.zotero.org/google-docs/?C3CSPB}{(Amaratunga et al.,
-2008; Brieuc et al., 2018; Chen and Ishwaran, 2012; Franzosa et al.,
-2019; Stephan et al., 2015)}, along with other works where it was used
-to predict cognitive conditions related to Alzheimer's disease in
-different scenarios
-\href{https://www.zotero.org/google-docs/?GiadGy}{(Ardekani et al.,
-2017; Velazquez et al., 2021)}. Additionally, given the sequential
-nature of variable consideration in each tree, RFs are naturally able to
-work out complex input feature interactions, such as those present in a
-microbiome-wide study, without the necessity to explicitly compute
-interaction terms.
-
-\includegraphics[width=6.5in,height=5.19444in]{media/image2.png}
-
-\textbf{Figure 3 - Random forest models predict concurrently measured
-cognitive function.}
-
-Comparison of RF predictor importance versus linear models for children
-between birth and 6 months old (A), and for those older than 18 months
-(B). Colors represent whether the species belong to the group of
-top-important features that account for 80\% of the cumulative
-importance on the RF model, if that species was significant (q
-\textless{} 0.2) in linear models, both, or neither. Ranked feature
-importance for taxa in RF models for children between birth and 6 months
-old (C), and for those older than 18 months (D). Taxa that are important
-for RF models both for children under 6 months and for those over 18
-months (E), only for children under 6 months (F), and only for children
-over 18 months (G)
-
-Given that gut microbial profiles, as well as neurocognitive
-development, may partially reflect socioeconomic and demographic
-factors, we assessed the performance of RF regressors where maternal
-education (a proxy of socioeconomic status (SES)), sex, and age were
-included as possible predictors, either alone or in combination with
-microbial taxonomic profiles (Table 3).
-
-\textbf{Table 3. Benchmark metrics for the cognitive assessment score
-prediction models. Confidence intervals are calculated from the
-distribution of metrics from repeated CV at a confidence level of 95\%}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1456}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0960}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3584}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Subject Ages (months)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Microbial feature
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Demo.
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Test set correlation (± C.I.)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Test set Root-mean-square error (± C.I.)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.14 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-13.01 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.10 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.66 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.13 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.70 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.01 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.56 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.01 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.56 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.506 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16.27 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.363 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.66 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.429 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.29 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.303 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18.01 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.333 ± 0.004
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.86 ± 0.04
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-As with linear models, RF models for children under 6 months old were
-less generalizable (mean test-set correlation -0.13, mean RMSE 12.70),
-but RFs were consistently able to learn the relationship between taxa
-and cognitive function scores in children over 18 months of age (mean
-test-set correlation 0.429, mean RMSE 17.29). For both age groups,
-species that were important in RF overlapped with those that were
-significant when testing the relationship with linear models (Figure
-3A-B, Supplementary Table XXA-B), though there were substantial
-differences. For children under 6 months, only \emph{B. wexlerae} was
-significantly associated with cognitive function in linear models, but
-was ranked 20th in importance for RF models. The importance metric
-employed (MDI) is a measure of how well the variable differentiates
-(splits) sets of samples by creating subgroups that reduce the
-intra-group deviations, while also accounting for the amount of samples
-affected by a split. All taxa significantly associated with cognitive
-function score in children over 18 months using LMs belong to the
-top-ranking group responsible for 80\% of the total relative importance,
-except for \emph{Eubacterium ramulus} (Table 3, Figure 3A).
-
-Interestingly, several taxa highly ranked in importance in both age
-groups, including several that were significant in LMs for children over
-18 months old, including \emph{R. gnavus} (0--6 months, rank = 13;
-18--120 months, rank = 7) and \emph{G. pamelaeae} (0--6 months, rank =
-16; 18--120 months, rank = 20), while others such as \emph{Allistipes
-finegoldii} were age-group specific. Several taxa important in RF models
-were not statistically significant when using linear models after
-multiple hypothesis correction. However, these taxa had small nominal
-p-values. For example, \emph{Erysipelatoclostridium ramosum} was the
-most important feature in RF models for children under 6 months old and
-had an LM p-value of 0.04. Subject age was consistently ranked highly in
-feature importance, which could indicate that decision branches based on
-microbial taxa have increased purity when considering the subject's age
-or that age itself is a useful predictor.
-
-\hypertarget{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}{%
-\subsubsection{Gut microbial taxonomic profiles predict brain structure
-differences}\label{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}}
-
-If there are causal effects of microbial metabolism on cognitive
-function, they might be reflected in changes in neuroanatomy. We again
-employed a Random Forest modeling approach to associate gut taxonomic
-profiles with individual brain regions identified in MRI scans,
-normalized to total brain volume. Some brain regions were more readily
-predicted by RF models trained on microbial taxa (Table 4, Supplementary
-Table XX), in particular those that were highly correlated with age.
-These included the L/R lingual gyrus (mean RF correlations, Left =
-0.421, Right = 0.434; relative age importances, Left = 7.5\%, Right =
-8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
-0.200, Right = 0.273; relative age importances, Left = 3.7\%, Right =
-6.3\%). In many cases, however, age was not an important variable in
-high-performing models, such as that for the left accumbens area (mean
-RF correlations = 0.288; relative age importances = 1.2\%).
-
-\textbf{Table 4. Summary statistics for the neuroanatomy prediction
-benchmarks}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Statistic
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Mean absolute proportional error (MAPE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Correlation coefficient (R)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-mean
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.072
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.137
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Standard deviation
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.028
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.136
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-maximum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.605
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-75th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.195
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-50th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-25th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.062
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-minimum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.147
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-We also observed that many brain regions had high accordance across the
-left and right hemispheres in terms of both model performance and
-microbial feature importance. In contrast, other regions had substantial
-differences between the hemispheres. For example, the left accumbens
-area, which plays an important role in reward circuits
-\href{https://www.zotero.org/google-docs/?zVicU7}{(Ernst et al., 2005;
-Yau et al., 2012)}, has one of the highest test-set correlations of our
-brain region models (R = 0.288), as compared to the right accumbens
-models which could not adequately generalize, and had a negative mean
-test-set correlation (R = -0.041). Most healthy individuals have a
-rightward asymmetry in the nucleus accumbens, and reduced asymmetry has
-been linked to substance use disorder in young adults
-\href{https://www.zotero.org/google-docs/?V45KyN}{(Cao et al., 2021)}.
-Feature importance for models of the left accumbens area were dominated
-by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.4\%
-relative importance), \emph{B. ovatus} (3.7\% relative importance), and
-\emph{B. uniformis} (3.0\% relative importance).. The accumbens area is
-associated with reward control, and in individuals diagnosed with ADHD,
-it has been shown to have a divergent neuromorphology
-\href{https://www.zotero.org/google-docs/?kLfEQ5}{(Hoogman et al.,
-2017)}. Independently, \emph{B. ovatus}, \emph{B. uniformis} and
-\emph{B. vulgatus} have been linked to ADHD
-\href{https://www.zotero.org/google-docs/?luO0zh}{(Wang et al., 2020)}.
-In fact, there are studies showing that alterations on the striatal
-dopamine transporters can cause effects resembling hyperactivity and
-attention deficit
-\href{https://www.zotero.org/google-docs/?UxS47e}{(Yael et al., 2019)},
-and that \emph{B. uniformis} is gut-microbial modulator of the brain
-dopamine transporter
-\href{https://www.zotero.org/google-docs/?L1cWJS}{(Hartstra et al.,
-2020)}.
-
-Many of the taxa identified included taxa also identified in LMs and RF
-models of cognitive function. Interesting to note, while RF models for
-multiple brain regions had many important microbial taxa, others were
-dominated by a small number of taxa. In general, for all segments, a
-consistent number of species between 14 and 22 (median = 20) was
-responsible for one third of the fitness-weighted cumulative importance
-(see Methods), regardless of the model benchmarks. We proceeded to
-select a subset of the most important taxa (averaged over all segments)
-and the segments whose importance was more heavily loaded to further
-analyze the multiple relationships unveiled by the RF models
-(\textbf{Figure 4 A and B})
-
-\includegraphics[width=6.5in,height=5.19444in]{media/image1.png}
-
-\textbf{Figure 4: Microbial feature importance in predicting brain
-volumes in children over 18 months of age.}
-
-(A) Average test-set correlations for prediction of MRI segmentation
-data from microbiome and demographics on select segments after repeated
-cross validation. (B) Heatmap of average individual relative taxa
-importances on each brain segment. Importances are reported as
-proportions relative to the sum of importances for each model - since
-every model is trained with 132 features, and even distribution of
-importance would be 0.75\% for each feature.
-
-. Segments and taxa ordered by HCA on a select list of species with high
-ranks on average importances, and their respective highest-load
-segments. (C)
-
-Our analysis revealed two major patterns of importance distribution from
-taxa over the brain segments; some species portrayed high contributions
-to multiple different segments, while others contributed modestly to
-just one or two brain segments. Notable cases of the first pattern
-included seven species - \emph{Anaerostipes hadrus}, \emph{Bacteroides
-vulgatus}, \emph{Fusicatenibacter saccharivorans}, \emph{Ruminococcus
-torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
-\emph{Blautia wexlerae} - that combined, account for approximately one
-third of the cumulative relative importances, computed after subsetting
-on the taxa of interest (Figure 4C).
-
-Among these, the most important variable is \emph{Anaerostipes hadrus},
-which is a butyrate-producing anaerobe that has been positively
-associated with cognitive function
-\href{https://www.zotero.org/google-docs/?AVJiwu}{(Kant et al., 2015; Li
-et al., 2022)}. Its importance is, however, heavily loaded on the volume
-of entire major lobes (frontal, parietal), which cannot be readily
-correlated to specific cognitive functions. After accounting for the
-major lobes, it is found to be highly important on the prediction of the
-cerebellar vermal lobules VIII-X, pars opercularis, cuneus and
-precuneus, and anterior-cingulate (Figure 4B).
-
-To better understand this relationship, we performed hierarchical
-cluster analysis, which revealed a well defined cluster of species with
-high importance loadings on close segments of the lower temporal and
-close occipital lobe, to which \emph{A. hadrus} belongs. The other
-species on this cluster are \emph{B. wexlerae}, \emph{R. torques},
-\emph{R. intestinalis, R. bicirculans} and \emph{F. saccahrivorans}, who
-contribute heavily to the entorhinal, fusiform, lingual and
-parahippocampal segments.
-
-This group of taxa and their related segments drew our attention because
-they contained \emph{B. wexlerae}, a highlight from the cognitive
-assessment results. Previous works exploring Parkinson Disease patients
-found out that issues in the cognitive task of confrontation naming were
-positively correlated with thinning in the fusiform gyrus and
-parahippocampal gyrus
-\href{https://www.zotero.org/google-docs/?rlO6eq}{(Pagonabarraga et al.,
-2013)}. Additionally, the left and right parahippocampal, where \emph{B.
-wexlerae} had the highest importance in both models, are important in
-visual/spatial processing and memory
-\href{https://www.zotero.org/google-docs/?wkkWZT}{(Aminoff et al.,
-2013)}. \emph{B. wexlerae} can also produce acetylcholine in the gut
-\href{https://www.zotero.org/google-docs/?jBoxSm}{(Hosomi et al.,
-2022)}, and this molecule plays an important role in modulating memory
-function \href{https://www.zotero.org/google-docs/?BQ0j2c}{(Haam and
-Yakel, 2017)}.
-
-Another notable importance cluster contains taxa associated with the
-basal forebrain, especially the cingulate and the accumbens. This
-cluster contains the previously-discussed \emph{B. ovatus} and \emph{B.
-uniformis,} but also \emph{Alistipes finegoldii and Streptococcus
-salivaris.} Reduction in the nucleus accumbens has been associated with
-depression symptoms
-\href{https://www.zotero.org/google-docs/?2aSZIr}{(Wacker et al.,
-2009)}, and increased levels of the \emph{Alistipes} genus have been
-observed in patients with depressive disorder
-\href{https://www.zotero.org/google-docs/?bhhSZB}{(Jiang et al., 2015)}.
-While these reports are important, simultaneously probing of the gut
-metagenome and brain structure relationship is novel.
-
-Both \emph{A. equolofaciens} and \emph{A. celatus,} two closely-related
-equol-producing species, are examples of the second contribution
-pattern, and had high importance in predicting the relative volume of
-the right anterior cingulate (respectively, 3.0\% and 2.6\% relative
-importances), which has been linked to social cognition and reward-based
-decision making \href{https://www.zotero.org/google-docs/?Wi8Bpq}{(Apps
-et al., 2016; Boes et al., 2008; Bush et al., 2002)}. Equol has a strong
-estrogenic effect
-\href{https://www.zotero.org/google-docs/?PqgFU7}{(Setchell et al.,
-2005)}, and in anterior cingulate, estrogen, has been shown to regulate
-pain-related aversion
-\href{https://www.zotero.org/google-docs/?5huRIR}{(Xiao et al., 2013)}.
-Another example of this pattern is \emph{R. gnavus,} which was the only
-species with significant negative association with cognitive function.
-It is heavily associated with the left pars opercularis (2.7\% relative
-importance), a relationship that may be explained by the emerging
-understanding that this species is increased in individuals with insulin
-resistance and obesity
-\href{https://www.zotero.org/google-docs/?lwzWUv}{(Ley et al., 2006)},
-conditions that are known to produce structural abnormalities in the
-brain \href{https://www.zotero.org/google-docs/?NlkFr6}{(Opel et al.,
-2021)}. Finally, \emph{Coprococcus comes} displays an importance
-distribution that splits almost evenly among the two previously reported
-clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (4.0\% relative importance)
-(only age had higher relative importance in any model), while also being
-one of the most important predictors for neighboring areas of the
-posterior cingulate such as the pars opercularis (relative importances,
-Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
-precentral (2.3\% relative importance) and paracentral lobes (relative
-importances, Left = 2.6\%, Right = 1.8\%).
-
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
-
-The relationship between the gut microbiome and brain function via the
-gut-microbiome-brain axis has gained increasing acceptance largely as a
-result of human epidemiological studies investigating atypical
-neurocognition (eg anxiety and depression, neurodegeneration, attention
-deficit / hyperactivity disorder, and autism) and mechanistic studies in
-animal models. The results from these studies point to the possibility
-that gut microbes and their metabolism may be causally implicated in
-cognitive development, but this study is the first to our knowledge that
-directly investigates microbial species and their genes in relation to
-typical development in young children. Understanding the gut-brain-axis
-in early life is particularly important, since differences or
-interventions in early life can have outsized and longer-term
-consequences than those at later ages. Further, even in the absence of
-causal impacts of microbial metabolism, identifying risk factors that
-could point to other early interventions would also have value.
-
-The use of shotgun metagenomic sequencing enabled us to get
-species-level resolution of microbial taxa. A previous study of
-cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing,
-and showed that genera from the Lachnospiraceae family as well as
-unclassified Clostridiales (now Eubacteriales) were associated with
-higher scores on the Ages and Stages Questionnaire
-\href{https://www.zotero.org/google-docs/?SB5Lwa}{(Sordillo et al.,
-2019)}. However, each of these clades encompass dozens of genera with
-diverse functions, each of which may have different effects. Indeed,
-several of the taxa that were positively associated with cognitive
-function in this study, including \emph{B. wexlerae}, the only species
-identified by linear models in children under 6 months, \emph{D.
-longicatena}, \emph{R. faecis}, and \emph{A. finegoldii} are
-Clostridiales, as is \emph{R. gnavus}, which we found was negatively
-associated with cognitive function (Figure 2A-B). This kind of
-species-level resolution is typically not possible with amplicon
-sequencing.
-
-We identified several species in the family Eggerthelaceae that were
-associated with cognitive function, including \emph{Gordonibacter
-pamelaeae, Aldercreutzia equolofaciens, Asaccharobacter celatus}
-(formerly regarded a subspecies of \emph{A. equolofaciens}
-\href{https://www.zotero.org/google-docs/?OEhguI}{(Takahashi et al.,
-2021)}), and \emph{Eggerthella lenta}. Many members of this family are
-known in part due to unique metabolic activities. For example, \emph{A.
-equolofaciens} produces the nonsteroidal estrogen equol from isoflaven
-found in soybeans
-\href{https://www.zotero.org/google-docs/?I2qNyB}{(Wang et al., 2005)},
-and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
-(found in pomegranates and some berries) into urolithin, which has been
-shown in some studies to have a neuroprotective effect
-\href{https://www.zotero.org/google-docs/?oTN8Lw}{(Gong et al., 2022;
-Selma et al., 2014)}. \emph{E. lenta} has been extensively studied for
-its ability to metabolize drug compounds such as the plant-derived heart
-medication digoxin
-\href{https://www.zotero.org/google-docs/?tH2Qeq}{(Haiser et al.,
-2013)}. The metabolic versatility of this clade, and the large number of
-species that are associated with cognition make these microbes prime
-targets for further mechanistic studies.
-
-In addition to improved species-level resolution, shotgun metagenomic
-sequencing also enables gene-functional insight. We showed here that
-genes for the metabolism of SCFAs, both their degradation and synthesis,
-are associated with cognitive function scores. However, while the
-differential abundance of genes for the metabolism of neuroactive
-compounds like these is suggestive, it is difficult to reason about the
-relationship between levels of these genes and the gut concentrations of
-the molecules their product enzymes act on. For example, while it might
-be intuitive to reason that increased levels of menaquinone synthesis
-genes is indicative of increased menaquinone, it could be the case that
-menaquinone deficiency selects for microbes that can synthesize it. For
-the same reason, increased propionate degradation genes may
-counterintuitively be indicative of high levels of propionate in the gut
-lumen, since high propionate would select for microbes that can
-metabolize it. For this reason, future studies coupling shotgun
-metagenomics with stool metabolomics could improve our understanding of
-the relationship between microbial metabolism and cognitive development.
-Further, strain-level analysis linking specific gene content in species
-of interest could further refine targeted efforts at identifying
-specific metabolic signatures of microbe-brain interactions.
-
-\{\{Guilherme add a paragraph about the usefulness of ML\}\}
-
-The use of multiple age-appropriate cognitive assessments that could be
-normalized to a common scale enabled us to analyze microbial
-associations across multiple developmental periods, but carries several
-drawbacks. In particular, the test-retest reliability, as well as
-systematic differences between test administrators may introduce
-substantial noise into these observations, particularly in the youngest
-children. In addition, our study period overlapped with the beginning of
-the COVID-19 pandemic, and we and others have observed some reduction in
-measured scores for children that were assessed after the implementation
-of lockdowns. In our subject set for this study, these effects are more
-pronounced in some age groups due to our sampling schedule
-\href{https://www.zotero.org/google-docs/?L8GnRj}{(Blackwell et al.,
-2022; Deoni et al., 2021)} (Supplementary Figure 3).
-
-This analysis allowed us to establish links between microbial taxa and
-their functional potential with cognition and brain structure. Although
-we cannot test causality or the chemistry behind the interactions
-between gut microbial taxa, gut, and brain, this study provides clear
-and statistically significant associations between the infant and early
-child gut microbiota and neurocognition. Future studies should focus on
-characterizing the early-life microbiome and neurocognitive development
-across different geographic regions and lifestyles such as covering
-traditionally understudied low-resource urban, peri-urban and rural
-communities to obtain the more comprehensive understanding of the
-variability within the different gut microbiomes reflects on
-neurocognition. These studies would also provide us with the wealth of
-data on different strains from the same species to better understand the
-effect of genes and their products. Furthermore, culturing and microbial
-community enrichment studies combined with genetic manipulation and
-genomic approaches to understand microbial metabolism at the molecular
-level is the key, as the metabolic functions shape and influence the
-human host and its health. The discovery of the neuroactive metabolites
-could provide us with biomarkers for early detection or necessary
-medicinally useful molecules that can be applied in intervention.
-
-\hypertarget{material-and-methods}{%
-\subsection{Material and Methods}\label{material-and-methods}}
-
-\hypertarget{study-ethics}{%
-\subsubsection{Study Ethics}\label{study-ethics}}
-
-All procedures for this study were approved by the local institutional
-review board at Rhode Island Hospital, and all experiments adhered to
-the regulation of the review board. Written informed consent was
-obtained from all parents or legal guardians of enrolled participants.
-
-\hypertarget{participants}{%
-\subsubsection{Participants}\label{participants}}
-
-Data used in this study were drawn from the ongoing longitudinal
-RESONANCE study of healthy and neurotypical brain and cognitive
-development, based at Brown University in Providence, RI, USA. The
-RESONANCE study is part of the NIH initiative Environmental influences
-on Child Health Outcomes (ECHO)
-\href{https://www.zotero.org/google-docs/?0tZanX}{(Forrest et al., 2018;
-Gillman and Blaisdell, 2018)}, a longitudinal observational study of
-healthy and neurotypical brain development that spans the fetal and
-infant to adolescent life stages, combining neuroimaging (magnetic
-resonance imaging, MRI), neurocognitive assessments, bio-specimen
-analyses, subject genetics, environmental exposures such as lead, and
-rich demographic, socioeconomic, family and medical history information.
-From the RESONANCE cohort, 361 typically-developing children between the
-ages of 2.8 months and 10 years old (median age 2 years, 2 months) were
-selected for analysis in this study.
-
-General participant demographics are provided in Table 1. Children are
-representative of the RI population. Children in the RESONANCE cohort
-were born full-term (\textless{} 37 weeks gestation) with height and
-weight normal for gestational age, and from uncomplicated singleton
-pregnancies. Children with known major risk factors for developmental
-abnormalities at enrollment were excluded. In addition to screening at
-the time of enrollment, on-going screening for worrisome behaviors using
-validated tools was performed to identify at-risk children and remove
-them from subsequent analysis.
-
-Exclusion criteria included: \emph{in utero} exposure to alcohol,
-cigarette or illicit substance exposure; preterm (\textless{} 37 weeks
-gestation) birth; small for gestational age or less than 1500 g; fetal
-ultrasound abnormalities; preeclampsia, high blood pressure, or
-gestational diabetes; 5 minute APGAR scores \textless{} 8; NICU
-admission; neurological disorder (e.g., head injury resulting in loss of
-consciousness, epilepsy); and psychiatric or learning disorder
-(including maternal depression) in the infant, parents, or siblings
-requiring medication in the year prior to pregnancy.
-
-Demographic and other non-biospecimen data such as race and ethnicity,
-parental education and occupation, feeding behavior (breast- and
-formula-feeding), child weight and height, were collected through
-questionnaires or direct examination as appropriate. All data were
-collected at every assessment visit, if possible.
-
-\hypertarget{cognitive-assessments}{%
-\subsubsection{Cognitive Assessments}\label{cognitive-assessments}}
-
-Overall cognitive function was assessed using age-appropriate methods.
-For children from birth to 30 months, we used an Early Learning
-Composite as assessed via the Mullen Scales of Early Learning (MSEL)
-\href{https://www.zotero.org/google-docs/?NSyykd}{(Mullen and others,
-1995)}, a standardized and population-normed tool for assessing fine and
-gross motor, expressive and receptive language, and visual reception
-functioning in children from birth through 68 months of age. The
-Wechsler Intelligence Quotient for Children (WISC)
-\href{https://www.zotero.org/google-docs/?OopdA1}{(Wechsler, 2012)} is
-an individually administered standard intelligence test for children
-aged 6 to 16 years. It derives a full scale intelligence quotient (IQ)
-score, which we used to assess overall cognitive functioning. The fourth
-edition of the Wechsler Preschool and Primary Scale of Intelligence
-(WPPSI-IV) is an individually administered standard intelligence test
-for children aged 2 years 6 months to 7 years 7 months, trying to meet
-the increasing need for the assessment of preschoolers. Just as the
-WISC, it derives a full scale IQ score, which we used to assess overall
-cognitive functioning.
-
-\hypertarget{stool-sample-collection-and-sequencing}{%
-\subsubsection{Stool Sample Collection and
-Sequencing}\label{stool-sample-collection-and-sequencing}}
-
-Stool samples (n=493) were collected by parents in OMR-200 tubes
-(OMNIgene GUT, DNA Genotek, Ottawa, Ontario, Canada), stored on ice, and
-brought within 24 hrs to the lab in RI where they were immediately
-frozen at -80 ˚C. Stool samples were not collected if the subject had
-taken antibiotics within the last two weeks. DNA extraction was
-performed at Wellesley College (Wellesley, MA). Nucleic acids were
-extracted from stool samples using the RNeasy PowerMicrobiome kit,
-excluding the DNA degradation steps. Briefly, the samples were lysed by
-bead beating using the Powerlyzer 24 Homogenizer (Qiagen, Germantown,
-MD) at 2500 rpm for 45 s and then transferred to the QIAcube (Qiagen,
-Germantown, MD) to complete the extraction protocol. Extracted DNA was
-sequenced at the Integrated Microbiome Resource (IMR, Dalhousie
-University, NS, Canada).
-
-Shotgun metagenomic sequencing was performed on all samples. A pooled
-library (max 96 samples per run) was prepared using the Illumina Nextera
-Flex Kit for MiSeq and NextSeq from 1 ng of each sample. Samples were
-then pooled onto a plate and sequenced on the Illumina NextSeq 550
-platform using 150+150 bp paired-end ``high output'' chemistry,
-generating 400 million raw reads and 120 Gb of sequence per plate.
-
-\hypertarget{machine-learning-for-cognitive-development}{%
-\subsubsection{Machine learning for cognitive
-development}\label{machine-learning-for-cognitive-development}}
-
-Prediction of cognitive scores was carried out as a set of regression
-experiments targeting real-valued continuous assessment scores.
-Different experiment sets were designed to probe how different
-representations of the gut microbiome (taxonomic profiles, Functional
-profiles encoded as ECs) would behave, with and without the addition of
-demographics (sex and maternal education as a proxy of socioeconomic
-status) on participants from different age groups. Age (in months) was
-provided as a covariate for all models (Table 3).
-
-\textbf{Table 5. Experimental design and input composition for Random
-Forest experiments}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1797}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2915}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Input set
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Age bracket
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Microbiome encoding type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Demographics
-
-Provided?
-
-(sex, education)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-1
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Not provided
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-3
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-4
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-5
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Not provided
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-7
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-9
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-10
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-Random Forests (RFs) \href{https://paperpile.com/c/dPbU4e/VDqU}{(Breiman
-1996)} were selected as the prediction engine and processed using the
-DecisionTree.jl
-\href{https://www.zotero.org/google-docs/?dQLyDs}{(Sadeghi et al.,
-2022)} implementation, inside the MLJ.jl
-\href{https://www.zotero.org/google-docs/?y8ywGI}{(Blaom et al., 2020)}
-Machine Learning framework. Independent RFs were trained for each
-experiment, using a set of default regression hyperparameters from
-Breiman and Cutler
-\href{https://www.zotero.org/google-docs/?d3gqFh}{(Breiman, 2001)}, on a
-repeated cross-validation approach with different RNG seeds. One hundred
-repetitions of 3-fold CV with 10 different intra-fold RNG states each
-were employed, for a total of 3000 experiments per input set.
-
-After the training procedures, the root-mean-square error (RMSE) for
-cognitive assessment scores and mean absolute proportional error (MAPE)
-for the brain segmentation data, along with Pearson\textquotesingle s
-correlation coefficient (R) were benchmarked on the validation and train
-sets. MAPE was chosen as the metric for brain segments due to magnitude
-differences between median volumes of each segment, which would hinder
-interpretation of raw error values without additional reference.
-
-To derive biological insight from the models, the covariate variable
-importances for all the input features, measured by mean decrease in
-impurity (MDI, or GINI importance), was also analyzed. Leveraging the
-distribution of results from the extensive repeated cross validation
-experiments, rather than electing a representative model or picking the
-highest validation-set correlation, a measure of model fitness
-(\textbf{Equation 1}) was designed to weight the importances from each
-trained forest. The objective was to give more weight to those with
-higher benchmarks on the validation sets (or higher generalizability),
-while penalizing information from highly overfit models, drawing
-inspiration from the approach used on another work employing repeated CV
-on Random Forests with high-dimensional, low sample size microbiome
-datasets \{***\}. The resulting fitness-weighted importances were used
-to generate the values in Figure 3.
-
-\(fitness =\)
-
-\textbf{Equation 1.} Mathematical expression of the fitness measure used
-to weight feature importances based on model benchmarks
-
-\hypertarget{mri-segmentation}{%
-\subsubsection{MRI / segmentation}\label{mri-segmentation}}
-
-MRI data was acquired at a 3T Siemens Trio scanner with the following
-parameters: TE=5.6msec, TR=1400msec, FA=15 degrees, 1.1x1.1x1.1mm
-resolution, 160x160 matrix, with an average of 112 slices. FOV was
-adjusted to infant size. Using a combination of linear and nonlinear
-image registration, we created representative age-specific templates
-using the ANTs package
-\href{https://www.zotero.org/google-docs/?ZlWDfL}{(Avants et al.,
-2014)}. After age specific templates were created, a single flow from
-each age to the 12-month template was estimated and a final warp from
-the 12-month template to standard adult MNI space was performed. For
-each individual infant/ child brain image, we then calculated the warp
-from their native T1w image space to their nearest-in-age template.
-Using the resulting warps (native → nearest age template → 12-month
-template → MNI template), we could move the standard adult brain atlas
-to the space of an individual infant in a single step. In this case, we
-used the Harvard-Oxford brain atlas to provide a coarse-grained
-parcellation of individual brains into subcortical regions (e.g.,
-thalamus, putamen) and total grey and white matter volumes (included as
-part of the FSL package
-\href{https://www.zotero.org/google-docs/?AapPN7}{(Jenkinson et al.,
-2012)}. Total tissue and brainstem volume as well as left and right
-hemisphere volumes were derived for total white and cortical gray
-matter, lateral ventricle, thalamus, caudate, putamen, pallidum,
-hippocampus, amygdala, and accumbens as well as total brainstem volume
-\href{https://www.zotero.org/google-docs/?MCDDB9}{(Bruchhage et al.,
-n.d.)}.
-
-\hypertarget{computational-analysis-of-metagenomes}{%
-\subsubsection{Computational analysis of
-metagenomes}\label{computational-analysis-of-metagenomes}}
-
-Shotgun metagenomic sequences were analyzed using the bioBakery suite of
-computational tools \href{https://paperpile.com/c/dPbU4e/vDau}{(Beghini
-et al. 2021)}. First, KneadData (v0.7.7) was used to perform quality
-control of raw sequence reads, such as read trimming and removal of
-reads matching a human genome reference. Next, MetaPhlAn (v3.0.7, using
-database mpa\_v30\_CHOCOPhlAn\_201901) was used to generate taxonomic
-profiles by aligning reads to a reference database of marker genes.
-Finally, HUMAnN (v3.0.0a4) was used to functionally profile the
-metagenomes.
-
-\hypertarget{data-and-code-availability}{%
-\subsubsection{Data and code
-availability}\label{data-and-code-availability}}
-
-Taxonomic and functional microbial profiles, as well as subject
-demographics necessary for statistical analyses and machine learning are
-available on the Open Science Framework
-\href{https://www.zotero.org/google-docs/?61HLlM}{(Bonham et al.,
-2022)}. Data processing, generation of summary statistics, and
-generation of plots was performed using the julia programming language
-\href{https://www.zotero.org/google-docs/?zj3HnY}{(Bezanson et al.,
-2017; Bonham et al., 2021; Danisch and Krumbiegel, 2021)}. All code for
-data analysis and figure generation, as well as scripts for automated
-download of input files are available on github \{\{CITE zenodo\}\}.
-
-\hypertarget{research-standards}{%
-\subsection{Research Standards}\label{research-standards}}
-
-\hypertarget{acknowledgements}{%
-\subsection{Acknowledgements}\label{acknowledgements}}
-
-Research funding: NIH UG3 OD023313 and Wellcome LEAP 1kD.
-
-\hypertarget{references-and-citations}{%
-\subsection{References and Citations}\label{references-and-citations}}
-
-\url{https://doi.org/10.1093/cercor/bhs201}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Amaratunga, D.,
-Cabrera, J., Lee, Y.-S., 2008. Enriched random forests. Bioinformatics
-24, 2010--2014. https://doi.org/10.1093/bioinformatics/btn356}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Aminoff, E.M.,
-Kveraga, K., Bar, M., 2013. The role of the parahippocampal cortex in
-cognition. Trends Cogn. Sci. 17, 379--390.
-https://doi.org/10.1016/j.tics.2013.06.009}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Apps, M.A.J.,
-Rushworth, M.F.S., Chang, S.W.C., 2016. The Anterior Cingulate Gyrus and
-Social Cognition: Tracking the Motivation of Others. Neuron 90,
-692--707. https://doi.org/10.1016/j.neuron.2016.04.018}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ardekani, B.A.,
-Bermudez, E., Mubeen, A.M., Bachman, A.H., Initiative, for the A.D.N.,
-2017. Prediction of Incipient Alzheimer's Disease Dementia in Patients
-with Mild Cognitive Impairment. J. Alzheimers Dis. 55, 269--281.
-https://doi.org/10.3233/JAD-160594}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Avants, B.B.,
-Tustison, N.J., Stauffer, M., Song, G., Wu, B., Gee, J.C., 2014. The
-insight ToolKit image registration framework. Front. Neuroinform. 8, 44.
-https://doi.org/10.3389/fninf.2014.00044}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bäckhed, F., Roswall,
-J., Peng, Y., Feng, Q., Jia, H., Kovatcheva-Datchary, P., Li, Y., Xia,
-Y., Xie, H., Zhong, H., Khan, M.T., Zhang, J., Li, J., Xiao, L.,
-Al-Aama, J., Zhang, D., Lee, Y.S., Kotowska, D., Colding, C., Tremaroli,
-V., Yin, Y., Bergman, S., Xu, X., Madsen, L., Kristiansen, K., Dahlgren,
-J., Wang, J., 2015. Dynamics and stabilization of the human gut
-microbiome during the first year of life. Cell Host Microbe 17,
-690--703. https://doi.org/10.1016/j.chom.2015.04.004}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bairoch, A., 2000. The
-ENZYME database in 2000. Nucleic Acids Res. 28, 304--305.
-https://doi.org/10.1093/nar/28.1.304}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Benítez-Páez, A.,
-Gómez del Pugar, E.M., López-Almela, I., Moya-Pérez, Á., Codoñer-Franch,
-P., Sanz, Y., 2020. Depletion of \emph{Blautia} Species in the
-Microbiota of Obese Children Relates to Intestinal Inflammation and
-Metabolic Phenotype Worsening. mSystems 5, e00857-19.
-https://doi.org/10.1128/mSystems.00857-19}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bezanson, J., Edelman,
-A., Karpinski, S., Shah, V.B., 2017. Julia: A fresh approach to
-numerical computing. Siam Rev. 59, 65--98.
-https://doi.org/10.1137/141000671}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Blackwell, C.K.,
-Mansolf, M., Sherlock, P., Ganiban, J., Hofheimer, J.A., Barone, C.J.,
-Bekelman, T.A., Blair, C., Cella, D., Collazo, S., Croen, L.A., Deoni,
-S., Elliott, A.J., Ferrara, A., Fry, R.C., Gershon, R., Herbstman, J.B.,
-Karagas, M.R., LeWinn, K.Z., Margolis, A., Miller, R.L., O'Shea, T.M.,
-Porucznik, C.A., Wright, R.J., 2022. Youth well-being during the
-COVID-19 pandemic. Pediatrics 149.
-https://doi.org/10.1542/peds.2021-054754}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Blaom, A.D., Kiraly,
-F., Lienart, T., Simillides, Y., Arenas, D., Vollmer, S.J., 2020. MLJ: A
-Julia package for composable machine learning. J. Open Source Softw. 5,
-2704. https://doi.org/10.21105/joss.02704}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Boes, A.D., Tranel,
-D., Anderson, S.W., Nopoulos, P., 2008. Right anterior cingulate cortex
-volume is a neuroanatomical correlate of aggression and defiance in
-boys. Behav. Neurosci. 122, 677--684.
-https://doi.org/10.1037/0735-7044.122.3.677}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bokulich, N.A., Chung,
-J., Battaglia, T., Henderson, N., Jay, M., Li, H., D Lieber, A., Wu, F.,
-Perez-Perez, G.I., Chen, Y., Schweizer, W., Zheng, X., Contreras, M.,
-Dominguez-Bello, M.G., Blaser, M.J., 2016. Antibiotics, birth mode, and
-diet shape microbiome maturation during early life. Sci. Transl. Med. 8,
-343ra82. https://doi.org/10.1126/scitranslmed.aad7121}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Bottino,
-G.Z.M.F., Klepac-Ceraj, V., 2022. ECHO RESONANCE Microbiome.
-https://doi.org/10.17605/OSF.IO/YBS32}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Kayisire,
-A., Luo, A., Klepac-Ceraj, V., 2021. Microbiome.jl and BiobakeryUtils.jl
-- Julia packages for working with microbial community data. J Open
-Source Softw 6, 3876. https://doi.org/10.21105/joss.03876}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Breiman, L., 2001.
-Random Forests. Mach. Learn. 45, 5--32.
-https://doi.org/10.1023/A:1010933404324}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Brieuc, M.S.O.,
-Waters, C.D., Drinan, D.P., Naish, K.A., 2018. A practical introduction
-to Random Forest for genetic association studies in ecology and
-evolution. Mol. Ecol. Resour. 18, 755--766.
-https://doi.org/10.1111/1755-0998.12773}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bruchhage, Muriel,
-Chen, Y., Cataldo, A.G., Müller, H.-G., Weisee, E., Wilson, S., Pietsch,
-M., D'Sa, V., Marquand, A., Madhow, S., Bouchard, K., Cole, J.H.,
-Biondo, F., Elison, J., O'Muirchheartaigh, J., Deoni, C, S., n.d.
-Longitudinal brain and cognitive development of the first 1000 Days:A
-large Multi-Cohort Multi-Scanner study. Presented at the International
-Society of Magnetic Resonance Medicine.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bush, G., Vogt, B.A.,
-Holmes, J., Dale, A.M., Greve, D., Jenike, M.A., Rosen, B.R., 2002.
-Dorsal anterior cingulate cortex: A role in reward-based decision
-making. Proc. Natl. Acad. Sci. 99, 523--528.
-https://doi.org/10.1073/pnas.012470999}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Callaghan, B., 2020.
-Nested sensitive periods: how plasticity across the microbiota-gut-brain
-axis interacts to affect the development of learning and memory. Curr.
-Opin. Behav. Sci. 36, 55--62.
-https://doi.org/10.1016/j.cobeha.2020.07.011}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cao, Z.,
-Ottino-Gonzalez, J., Cupertino, R.B., Schwab, N., Hoke, C., Catherine,
-O., Cousijn, J., Dagher, A., Foxe, J.J., Goudriaan, A.E., Hester, R.,
-Hutchison, K., Li, C.-S.R., London, E.D., Lorenzetti, V., Luijten, M.,
-Martin-Santos, R., Momenan, R., Paulus, M.P., Schmaal, L., Sinha, R.,
-Sjoerds, Z., Solowij, N., Stein, D.J., Stein, E.A., Uhlmann, A., van
-Holst, R.J., Veltman, D.J., Wiers, R.W., Yücel, M., Zhang, S.,
-Jahanshad, N., Thompson, P.M., Conrod, P., Mackey, S., Garavan, H.,
-2021. Mapping cortical and subcortical asymmetries in substance
-dependence: Findings from the ENIGMA Addiction Working Group. Addict.
-Biol. 26, e13010. https://doi.org/10.1111/adb.13010}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cerdó, T., Diéguez,
-E., Campoy, C., 2019. Early nutrition and gut microbiome:
-interrelationship between bacterial metabolism, immune system, brain
-structure, and neurodevelopment. Am. J. Physiol. Endocrinol. Metab. 317,
-E617--E630. https://doi.org/10.1152/ajpendo.00188.2019}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Chen, X., Ishwaran,
-H., 2012. Random forests for genomic data analysis. Genomics 99,
-323--329. https://doi.org/10.1016/j.ygeno.2012.04.003}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cohen Kadosh, K.,
-Krause, B., King, A.J., Near, J., Cohen Kadosh, R., 2015. Linking GABA
-and glutamate levels to cognitive skill acquisition during development.
-Hum. Brain Mapp. 36, 4334--4345. https://doi.org/10.1002/hbm.22921}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cowan, C.S.M., Dinan,
-T.G., Cryan, J.F., 2020. Annual Research Review: Critical windows - the
-microbiota-gut-brain axis in neurocognitive development. J. Child
-Psychol. Psychiatry 61, 353--371. https://doi.org/10.1111/jcpp.13156}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dalile, B., Van
-Oudenhove, L., Vervliet, B., Verbeke, K., 2019. The role of short-chain
-fatty acids in microbiota-gut-brain communication. Nat. Rev.
-Gastroenterol. Hepatol. 16, 461--478.
-https://doi.org/10.1038/s41575-019-0157-3}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Danisch, S.,
-Krumbiegel, J., 2021. Makie.jl: Flexible high-performance data
-visualization for Julia. J. Open Source Softw. 6, 3349.
-https://doi.org/10.21105/joss.03349}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Deoni, S.C.,
-Beauchemin, J., Volpe, A., Dâ Sa, V., RESONANCE Consortium, 2021. Impact
-of the COVID-19 pandemic on early child cognitive development: Initial
-findings in a longitudinal observational study of child health. MedRxiv
-Prepr. Serv. Health Sci. https://doi.org/10.1101/2021.08.10.21261846}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dominguez-Bello, M.G.,
-Costello, E.K., Contreras, M., Magris, M., Hidalgo, G., Fierer, N.,
-Knight, R., 2010. Delivery mode shapes the acquisition and structure of
-the initial microbiota across multiple body habitats in newborns. Proc.
-Natl. Acad. Sci. U. S. A. 107, 11971--11975.
-https://doi.org/10.1073/pnas.1002601107}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dong, H., Wang, B.,
-Feng, J., Yue, X., Jia, F., 2021. Correlation between serum
-concentrations of menaquinone-4 and developmental quotients in children
-with autism spectrum disorder. Front. Nutr. 8, 748513.
-https://doi.org/10.3389/fnut.2021.748513}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ernst, M., Nelson,
-E.E., Jazbec, S., McClure, E.B., Monk, C.S., Leibenluft, E., Blair, J.,
-Pine, D.S., 2005. Amygdala and nucleus accumbens in responses to receipt
-and omission of gains in adults and adolescents. NeuroImage 25,
-1279--1291. https://doi.org/10.1016/j.neuroimage.2004.12.038}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Forrest, C.B.,
-Blackwell, C.K., Camargo, C.A., 2018. Advancing the science of
-children's positive health in the national institutes of health
-environmental influences on child health outcomes (ECHO) research
-program. J. Pediatr. 196, 298--300.
-https://doi.org/10.1016/j.jpeds.2018.02.004}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Franzosa, E.A.,
-Sirota-Madi, A., Avila-Pacheco, J., Fornelos, N., Haiser, H.J., Reinker,
-S., Vatanen, T., Hall, A.B., Mallick, H., McIver, L.J., Sauk, J.S.,
-Wilson, R.G., Stevens, B.W., Scott, J.M., Pierce, K., Deik, A.A.,
-Bullock, K., Imhann, F., Porter, J.A., Zhernakova, A., Fu, J., Weersma,
-R.K., Wijmenga, C., Clish, C.B., Vlamakis, H., Huttenhower, C., Xavier,
-R.J., 2019. Gut microbiome structure and metabolic activity in
-inflammatory bowel disease. Nat. Microbiol. 4, 293--305.
-https://doi.org/10.1038/s41564-018-0306-4}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Fung, T.C., Olson,
-C.A., Hsiao, E.Y., 2017. Interactions between the microbiota, immune and
-nervous systems in health and disease. Nat. Neurosci. 20, 145--155.
-https://doi.org/10.1038/nn.4476}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ghosh, T.S., Rampelli,
-S., Jeffery, I.B., Santoro, A., Neto, M., Capri, M., Giampieri, E.,
-Jennings, A., Candela, M., Turroni, S., Zoetendal, E.G., Hermes, G.D.A.,
-Elodie, C., Meunier, N., Brugere, C.M., Pujos-Guillot, E., Berendsen,
-A.M., Groot, L.C.P.G.M.D., Feskins, E.J.M., Kaluza, J., Pietruszka, B.,
-Bielak, M.J., Comte, B., Maijo-Ferre, M., Nicoletti, C., Vos, W.M.D.,
-Fairweather-Tait, S., Cassidy, A., Brigidi, P., Franceschi, C., O'Toole,
-P.W., 2020. Mediterranean diet intervention alters the gut microbiome in
-older people reducing frailty and improving health status: the NU-AGE
-1-year dietary intervention across five European countries. Gut 69,
-1218--1228. https://doi.org/10.1136/gutjnl-2019-319654}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Gillman, M.W.,
-Blaisdell, C.J., 2018. Environmental influences on child health
-outcomes, a research program of the NIH. Curr. Opin. Pediatr. 30, 260.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Gong, Q.-Y., Cai, L.,
-Jing, Y., Wang, W., Yang, D.-X., Chen, S.-W., Tian, H.-L., 2022.
-Urolithin A alleviates blood-brain barrier disruption and attenuates
-neuronal apoptosis following traumatic brain injury in mice. Neural
-Regen. Res. 17, 2007. https://doi.org/10.4103/1673-5374.335163}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Haam, J., Yakel, J.L.,
-2017. Cholinergic modulation of the hippocampal region and memory
-function. J. Neurochem. 142, 111--121.
-https://doi.org/10.1111/jnc.14052}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Haiser, H.J.,
-Gootenberg, D.B., Chatman, K., Sirasani, G., Balskus, E.P., Turnbaugh,
-P.J., 2013. Predicting and Manipulating Cardiac Drug Inactivation by the
-Human Gut Bacterium Eggerthella lenta. Science 341, 295--298.
-https://doi.org/10.1126/science.1235872}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hartstra, A.V.,
-Schüppel, V., Imangaliyev, S., Schrantee, A., Prodan, A., Collard, D.,
-Levin, E., Dallinga-Thie, G., Ackermans, M.T., Winkelmeijer, M., Havik,
-S.R., Metwaly, A., Lagkouvardos, I., Nier, A., Bergheim, I.,
-Heikenwalder, M., Dunkel, A., Nederveen, A.J., Liebisch, G., Mancano,
-G., Claus, S.P., Benítez-Páez, A., la Fleur, S.E., Bergman, J.J.,
-Gerdes, V., Sanz, Y., Booij, J., Kemper, E., Groen, A.K., Serlie, M.J.,
-Haller, D., Nieuwdorp, M., 2020. Infusion of donor feces affects the
-gut--brain axis in humans with metabolic syndrome. Mol. Metab. 42,
-101076. https://doi.org/10.1016/j.molmet.2020.101076}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoogman, M., Bralten,
-J., Hibar, D.P., Mennes, M., Zwiers, M.P., Schweren, L., van Hulzen,
-K.J.E., Medland, S.E., Shumskaya, E., Jahanshad, N., de Zeeuw, P.,
-Szekely, E., Sudre, G., Wolfers, T., Onnink, A.M.H., Dammers, J.T.,
-Mostert, J.C., Vives-Gilabert, Y., Kohls, G., Oberwelland, E., Seitz,
-J., Schulte-Rüther, M., di Bruttopilo, S.A., Doyle, A.E., Høvik, M.F.,
-Dramsdahl, M., Tamm, L., van Erp, T.G.M., Dale, A., Schork, A.,
-Conzelmann, A., Zierhut, K., Baur, R., McCarthy, H., Yoncheva, Y.N.,
-Cubillo, A., Chantiluke, K., Mehta, M.A., Paloyelis, Y., Hohmann, S.,
-Baumeister, S., Bramati, I., Mattos, P., Tovar-Moll, F., Douglas, P.,
-Banaschewski, T., Brandeis, D., Kuntsi, J., Asherson, P., Rubia, K.,
-Kelly, C., Di Martino, A., Milham, M.P., Castellanos, F.X., Frodl, T.,
-Zentis, M., Lesch, K.-P., Reif, A., Pauli, P., Jernigan, T., Haavik, J.,
-Plessen, K.J., Lundervold, A.J., Hugdahl, K., Seidman, L.J., Biederman,
-J., Rommelse, N., Heslenfeld, D.J., Hartman, C., Hoekstra, P.J.,
-Oosterlaan, J., von Polier, G., Konrad, K., Vilarroya, O.,
-Ramos-Quiroga, J.-A., Soliva, J.C., Durston, S., Buitelaar, J.K.,
-Faraone, S.V., Shaw, P., Thompson, P., Franke, B., 2017. Subcortical
-brain volume differences of participants with ADHD across the lifespan:
-an ENIGMA collaboration. Lancet Psychiatry 4, 310--319.
-https://doi.org/10.1016/S2215-0366(17)30049-4}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoshino, Y., Yamamoto,
-T., Kaneko, M., Tachibana, R., Watanabe, M., Ono, Y., Kumashiro, H.,
-1984. Blood Serotonin and Free Tryptophan Concentration in Autistic
-Children. Neuropsychobiology 11, 22--27.
-https://doi.org/10.1159/000118045}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hosomi, K., Saito, M.,
-Park, J., Murakami, H., Shibata, N., Ando, M., Nagatake, T., Konishi,
-K., Ohno, H., Tanisawa, K., Mohsen, A., Chen, Y.-A., Kawashima, H.,
-Natsume-Kitatani, Y., Oka, Y., Shimizu, H., Furuta, M., Tojima, Y.,
-Sawane, K., Saika, A., Kondo, S., Yonejima, Y., Takeyama, H., Matsutani,
-A., Mizuguchi, K., Miyachi, M., Kunisawa, J., 2022. Oral administration
-of Blautia wexlerae ameliorates obesity and type 2 diabetes via
-metabolic remodeling of the gut microbiota. Nat. Commun. 13, 4477.
-https://doi.org/10.1038/s41467-022-32015-7}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Jenkinson, M.,
-Beckmann, C.F., Behrens, T.E.J., Woolrich, M.W., Smith, S.M., 2012. FSL.
-Neuroimage 62, 782--790.
-https://doi.org/10.1016/j.neuroimage.2011.09.015}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Jiang, H., Ling, Z.,
-Zhang, Y., Mao, H., Ma, Z., Yin, Y., Wang, W., Tang, W., Tan, Z., Shi,
-J., Li, L., Ruan, B., 2015. Altered fecal microbiota composition in
-patients with major depressive disorder. Brain. Behav. Immun. 48,
-186--194. https://doi.org/10.1016/j.bbi.2015.03.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kanehisa, M., Goto,
-S., Kawashima, S., Okuno, Y., Hattori, M., 2004. The KEGG resource for
-deciphering the genome. Nucleic Acids Res. 32, D277--D280.
-https://doi.org/10.1093/nar/gkh063}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kant, R., Rasinkangas,
-P., Satokari, R., Pietilä, T.E., Palva, A., 2015. Genome Sequence of the
-Butyrate-Producing Anaerobic Bacterium Anaerostipes hadrus PEL 85.
-Genome Announc. 3, e00224-15. https://doi.org/10.1128/genomeA.00224-15}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kim, C.-S., Cha, L.,
-Sim, M., Jung, S., Chun, W.Y., Baik, H.W., Shin, D.-M., 2021. Probiotic
-supplementation improves cognitive function and mood with changes in gut
-microbiota in Community-Dwelling older adults: A randomized,
-Double-Blind, Placebo-Controlled, multicenter trial. J Gerontol Biol Sci
-Med Sci 76, 32--40. https://doi.org/10.1093/gerona/glaa090}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Knudsen, E.I., 2004.
-Sensitive periods in the development of the brain and behavior. J. Cogn.
-Neurosci. 16, 1412--1425. https://doi.org/10.1162/0898929042304796}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Koenig, J.E., Spor,
-A., Scalfone, N., Fricker, A.D., Stombaugh, J., Knight, R., Angenent,
-L.T., Ley, R.E., 2011. Succession of microbial consortia in the
-developing infant gut microbiome. Proc. Natl. Acad. Sci. U. S. A. 108
-Suppl 1, 4578--4585. https://doi.org/10.1073/pnas.1000081107}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Coker,
-M.O., Madan, J.C., 2022. The developing microbiome from birth to 3
-years: The Gut-Brain axis and neurodevelopmental outcomes. Front.
-Pediatr. 10, 815885. https://doi.org/10.3389/fped.2022.815885}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Korrick,
-S.A., Baker, E.R., Karagas, M.R., Madan, J.C., 2020. Prospective
-associations of the infant gut microbiome and microbial function with
-social behaviors related to autism at age 3 years. Sci. Rep. 10, 15515.
-https://doi.org/10.1038/s41598-020-72386-9}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ley, R.E., Turnbaugh,
-P.J., Klein, S., Gordon, J.I., 2006. Human gut microbes associated with
-obesity. Nature 444, 1022--1023. https://doi.org/10.1038/4441022a}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Li, Y.-M., Yu, Q.,
-Huang, L., Fu, J., Feng, R.-J., 2022. {[}Correlations Between Gut
-Microbiota Changes and Cognitive Function in Patients with Post-Stroke
-Cognitive Impairment in the Early Stage{]}. Sichuan Da Xue Xue Bao Yi
-Xue Ban 53, 857--865. https://doi.org/10.12182/20220960105}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, S., Li, E., Sun,
-Z., Fu, D., Duan, G., Jiang, M., Yu, Y., Mei, L., Yang, P., Tang, Y.,
-Zheng, P., 2019. Altered gut microbiota and short chain fatty acids in
-Chinese children with autism spectrum disorder. Sci. Rep. 9, 287.
-https://doi.org/10.1038/s41598-018-36430-z}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, X., Mao, B., Gu,
-J., Wu, J., Cui, S., Wang, G., Zhao, J., Zhang, H., Chen, W., 2021.
-\emph{Blautia} ---a new functional genus with potential probiotic
-properties? Gut Microbes 13, 1875796.
-https://doi.org/10.1080/19490976.2021.1875796}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Lopez-Siles, M.,
-Duncan, S.H., Garcia-Gil, L.J., Martinez-Medina, M., 2017.
-Faecalibacterium prausnitzii: from microbiology to diagnostics and
-prognostics. ISME J. 11, 841--852.
-https://doi.org/10.1038/ismej.2016.176}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Louwies, T., Johnson,
-A.C., Orock, A., Yuan, T., Greenwood-Van Meerveld, B., 2020. The
-microbiota-gut-brain axis: An emerging role for the epigenome. Exp.
-Biol. Med. 245, 138--145. https://doi.org/10.1177/1535370219891690}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Magnusson, K.R.,
-Hauck, L., Jeffrey, B.M., Elias, V., Humphrey, A., Nath, R., Perrone,
-A., Bermudez, L.E., 2015. Relationships between diet-related changes in
-the gut microbiome and cognitive flexibility. Neuroscience 300,
-128--140. https://doi.org/10.1016/j.neuroscience.2015.05.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mallick, H.,
-Rahnavard, A., McIver, L.J., Ma, S., Zhang, Y., Nguyen, L.H., Tickle,
-T.L., Weingart, G., Ren, B., Schwager, E.H., Chatterjee, S., Thompson,
-K.N., Wilkinson, J.E., Subramanian, A., Lu, Y., Waldron, L., Paulson,
-J.N., Franzosa, E.A., Bravo, H.C., Huttenhower, C., 2021. Multivariable
-association discovery in population-scale meta-omics studies.
-https://doi.org/10.1101/2021.01.20.427420}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Maruo, T., Sakamoto,
-M., Ito, C., Toda, T., Benno, Y., 2008. Adlercreutzia equolifaciens gen.
-nov., sp. nov., an equol-producing bacterium isolated from human faeces,
-and emended description of the genus Eggerthella. Int. J. Syst. Evol.
-Microbiol. 58, 1221--1227. https://doi.org/10.1099/ijs.0.65404-0}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mayneris-Perxachs, J.,
-Castells-Nobau, A., Arnoriaga-Rodríguez, M., Martin, M., de la
-Vega-Correa, L., Zapata, C., Burokas, A., Blasco, G., Coll, C.,
-Escrichs, A., Biarnés, C., Moreno-Navarrete, J.M., Puig, J., Garre-Olmo,
-J., Ramos, R., Pedraza, S., Brugada, R., Vilanova, J.C., Serena, J.,
-Gich, J., Ramió-Torrentà, L., Pérez-Brocal, V., Moya, A., Pamplona, R.,
-Sol, J., Jové, M., Ricart, W., Portero-Otin, M., Deco, G., Maldonado,
-R., Fernández-Real, J.M., 2022. Microbiota alterations in proline
-metabolism impact depression. Cell Metab. 34, 681-701.e10.
-https://doi.org/10.1016/j.cmet.2022.04.001}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mukherjee, A., Lordan,
-C., Ross, R.P., Cotter, P.D., 2020. Gut microbes from the
-phylogenetically diverse genus Eubacterium and their various
-contributions to gut health. Gut Microbes 12, 1802866.
-https://doi.org/10.1080/19490976.2020.1802866}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mullen, E.M., others,
-1995. Mullen scales of early learning. AGS Circle Pines, MN.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Needham, B.D.,
-Funabashi, M., Adame, M.D., Wang, Z., Boktor, J.C., Haney, J., Wu,
-W.-L., Rabut, C., Ladinsky, M.S., Hwang, S.-J., Guo, Y., Zhu, Q.,
-Griffiths, J.A., Knight, R., Bjorkman, P.J., Shapiro, M.G., Geschwind,
-D.H., Holschneider, D.P., Fischbach, M.A., Mazmanian, S.K., 2022. A
-gut-derived metabolite alters brain activity and anxiety behaviour in
-mice. Nature 602, 647--653. https://doi.org/10.1038/s41586-022-04396-8}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Opel, N., Thalamuthu,
-A., Milaneschi, Y., Grotegerd, D., Flint, C., Leenings, R., Goltermann,
-J., Richter, M., Hahn, T., Woditsch, G., Berger, K., Hermesdorf, M.,
-McIntosh, A., Whalley, H.C., Harris, M.A., MacMaster, F.P., Walter, H.,
-Veer, I.M., Frodl, T., Carballedo, A., Krug, A., Nenadic, I., Kircher,
-T., Aleman, A., Groenewold, N.A., Stein, D.J., Soares, J.C.,
-Zunta-Soares, G.B., Mwangi, B., Wu, M.-J., Walter, M., Li, M., Harrison,
-B.J., Davey, C.G., Cullen, K.R., Klimes-Dougan, B., Mueller, B.A.,
-Sämann, P.G., Penninx, B., Nawijn, L., Veltman, D.J., Aftanas, L., Brak,
-I.V., Filimonova, E.A., Osipov, E.A., Reneman, L., Schrantee, A., Grabe,
-H.J., Van der Auwera, S., Wittfeld, K., Hosten, N., Völzke, H., Sim, K.,
-Gotlib, I.H., Sacchet, M.D., Lagopoulos, J., Hatton, S.N., Hickie, I.,
-Pozzi, E., Thompson, P.M., Jahanshad, N., Schmaal, L., Baune, B.T.,
-Dannlowski, U., 2021. Brain structural abnormalities in obesity:
-relation to age, genetic risk, and common psychiatric disorders. Mol.
-Psychiatry 26, 4839--4852. https://doi.org/10.1038/s41380-020-0774-9}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Pagonabarraga, J.,
-Corcuera-Solano, I., Vives-Gilabert, Y., Llebaria, G., García-Sánchez,
-C., Pascual-Sedano, B., Delfino, M., Kulisevsky, J., Gómez-Ansón, B.,
-2013. Pattern of Regional Cortical Thinning Associated with Cognitive
-Deterioration in Parkinson's Disease. PLoS ONE 8, e54980.
-https://doi.org/10.1371/journal.pone.0054980}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Palomo-Buitrago, M.E.,
-Sabater-Masdeu, M., Moreno-Navarrete, J.M., Caballano-Infantes, E.,
-Arnoriaga-Rodríguez, M., Coll, C., Ramió, L., Palomino-Schätzlein, M.,
-Gutiérrez-Carcedo, P., Pérez-Brocal, V., Simó, R., Moya, A., Ricart, W.,
-Herance, J.R., Fernández-Real, J.M., 2019. Glutamate interactions with
-obesity, insulin resistance, cognition and gut microbiota composition.
-Acta Diabetol. 56, 569--579. https://doi.org/10.1007/s00592-019-01313-w}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Pronovost, G.N.,
-Hsiao, E.Y., 2019. Perinatal interactions between the microbiome,
-immunity, and neurodevelopment. Immunity 50, 18--36.
-https://doi.org/10.1016/j.immuni.2018.11.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sadeghi, B.,
-Chiarawongse, P., Squire, K., Jones, D.C., Noack, A., St-Jean, C.,
-Huijzer, R., Schätzle, R., Butterworth, I., Peng, Y.-F., Blaom, A.,
-2022. DecisionTree.jl - A julia implementation of the CART decision tree
-and random forest algorithms. https://doi.org/10.5281/zenodo.7359268}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sela, D.A., Chapman,
-J., Adeuya, A., Kim, J.H., Chen, F., Whitehead, T.R., Lapidus, A.,
-Rokhsar, D.S., Lebrilla, C.B., German, J.B., Price, N.P., Richardson,
-P.M., Mills, D.A., 2008. The genome sequence of \emph{Bifidobacterium
-longum} subsp. \emph{infantis} reveals adaptations for milk utilization
-within the infant microbiome. Proc. Natl. Acad. Sci. 105, 18964--18969.
-https://doi.org/10.1073/pnas.0809584105}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Selma, M.V., Beltrán,
-D., García-Villalba, R., Espín, J.C., Tomás-Barberán, F.A., 2014.
-Description of urolithin production capacity from ellagic acid of two
-human intestinal Gordonibacter species. Food Funct. 5, 1779--1784.
-https://doi.org/10.1039/c4fo00092g}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Setchell, K.D.,
-Clerici, C., Lephart, E.D., Cole, S.J., Heenan, C., Castellani, D.,
-Wolfe, B.E., Nechemias-Zimmer, L., Brown, N.M., Lund, T.D., Handa, R.J.,
-Heubi, J.E., 2005. S-Equol, a potent ligand for estrogen receptor β, is
-the exclusive enantiomeric form of the soy isoflavone metabolite
-produced by human intestinal bacterial flora. Am. J. Clin. Nutr. 81,
-1072--1079. https://doi.org/10.1093/ajcn/81.5.1072}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sharon, G., Sampson,
-T.R., Geschwind, D.H., Mazmanian, S.K., 2016. The central nervous system
-and the gut microbiome. Cell 167, 915--932.
-https://doi.org/10.1016/j.cell.2016.10.027}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Silbereis, J.C.,
-Pochareddy, S., Zhu, Y., Li, M., Sestan, N., 2016. The cellular and
-molecular landscapes of the developing human central nervous system.
-Neuron 89, 248--268.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sordillo, J.E.,
-Korrick, S., Laranjo, N., Carey, V., Weinstock, G.M., Gold, D.R.,
-O'Connor, G., Sandel, M., Bacharier, L.B., Beigelman, A., Zeiger, R.,
-Litonjua, A.A., Weiss, S.T., 2019. Association of the infant gut
-microbiome with early childhood neurodevelopmental outcomes: An
-ancillary study to the VDAART randomized clinical trial. JAMA Netw. Open
-2, e190905. https://doi.org/10.1001/jamanetworkopen.2019.0905}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Spichak, S.,
-Bastiaanssen, T.F.S., Berding, K., Vlckova, K., Clarke, G., Dinan, T.G.,
-Cryan, J.F., 2021. Mining microbes for mental health: Determining the
-role of microbial metabolic pathways in human brain health and disease.
-Neurosci. Biobehav. Rev. 125, 698--761.
-https://doi.org/10.1016/j.neubiorev.2021.02.044}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Stephan, J., Stegle,
-O., Beyer, A., 2015. A random forest approach to capture genetic effects
-in the presence of population structure. Nat. Commun. 6, 7432.
-https://doi.org/10.1038/ncomms8432}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Suzek, B.E., Huang,
-H., McGarvey, P., Mazumder, R., Wu, C.H., 2007. UniRef: comprehensive
-and non-redundant UniProt reference clusters. Bioinformatics 23,
-1282--1288. https://doi.org/10.1093/bioinformatics/btm098}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Takahashi, H., Yang,
-J., Yamamoto, H., Fukuda, S., Arakawa, K., 2021. Complete Genome
-Sequence of Adlercreutzia equolifaciens subsp. celatus DSM 18785.
-Microbiol. Resour. Announc. 10, e00354-21.
-https://doi.org/10.1128/MRA.00354-21}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tau, G.Z., Peterson,
-B.S., 2010. Normal development of brain circuits. Neuropsychopharmacol.
-Off. Publ. Am. Coll. Neuropsychopharmacol. 35, 147--168.
-https://doi.org/10.1038/npp.2009.115}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Thawornkuno, C.,
-Tanaka, M., Sone, T., Asano, K., 2009. Biotransformation of Daidzein to
-Equol by Crude Enzyme from Asaccharobacter celatus AHU1763 Required an
-Anaerobic Environment. Biosci. Biotechnol. Biochem. 73, 1435--1438.
-https://doi.org/10.1271/bbb.80908}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tognini, P., 2017. Gut
-microbiota: a potential regulator of neurodevelopment. Front. Cell.
-Neurosci. 11, 25. https://doi.org/10.3389/fncel.2017.00025}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tso, L., Bonham, K.S.,
-Fishbein, A., Rowland, S., Klepac-Ceraj, V., 2021. Targeted
-High-Resolution taxonomic identification of bifidobacterium longum
-subsp. infantis using human milk oligosaccharide metabolizing genes.
-Nutrients 13, 2833. https://doi.org/10.3390/nu13082833}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Valles-Colomer, M.,
-Falony, G., Darzi, Y., Tigchelaar, E.F., Wang, J., Tito, R.Y., Schiweck,
-C., Kurilshikov, A., Joossens, M., Wijmenga, C., Claes, S., Van
-Oudenhove, L., Zhernakova, A., Vieira-Silva, S., Raes, J., 2019. The
-neuroactive potential of the human gut microbiota in quality of life and
-depression. Nat. Microbiol. 4, 623--632.
-https://doi.org/10.1038/s41564-018-0337-x}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Velazquez, M., Lee,
-Y., Initiative, for the A.D.N., 2021. Random forest model for
-feature-based Alzheimer's disease conversion prediction from early mild
-cognitive impairment subjects. PLOS ONE 16, e0244773.
-https://doi.org/10.1371/journal.pone.0244773}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wacker, J., Dillon,
-D.G., Pizzagalli, D.A., 2009. The role of the nucleus accumbens and
-rostral anterior cingulate cortex in anhedonia: Integration of resting
-EEG, fMRI, and volumetric techniques. NeuroImage 46, 327--337.
-https://doi.org/10.1016/j.neuroimage.2009.01.058}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wan, Y., Zuo, T., Xu,
-Z., Zhang, F., Zhan, H., Chan, D., Leung, T.-F., Yeoh, Y.K., Chan,
-F.K.L., Chan, R., Ng, S.C., 2021. Underdevelopment of the gut microbiota
-and bacteria species as non-invasive markers of prediction in children
-with autism spectrum disorder. Gut gutjnl-2020-324015.
-https://doi.org/10.1136/gutjnl-2020-324015}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, L.-J., Yang,
-C.-Y., Chou, W.-J., Lee, M.-J., Chou, M.-C., Kuo, H.-C., Yeh, Y.-M.,
-Lee, S.-Y., Huang, L.-H., Li, S.-C., 2020. Gut microbiota and dietary
-patterns in children with attention-deficit/hyperactivity disorder. Eur.
-Child Adolesc. Psychiatry 29, 287--297.
-https://doi.org/10.1007/s00787-019-01352-2}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, X.-L., Hur,
-H.-G., Lee, J.H., Kim, K.T., Kim, S.-I., 2005. Enantioselective
-Synthesis of S-Equol from Dihydrodaidzein by a Newly Isolated Anaerobic
-Human Intestinal Bacterium. Appl. Environ. Microbiol. 71, 214--219.
-https://doi.org/10.1128/AEM.71.1.214-219.2005}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 2012.
-Wechsler preschool and primary scale of intelligence---fourth edition.
-Psychol. Corp. San Antonio TX.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 1949.
-Wechsler intelligence scale for children.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, L., Yan, J.,
-Yang, T., Zhu, J., Li, T., Wei, H., Chen, J., 2021. Fecal Microbiome
-Transplantation from Children with Autism Spectrum Disorder Modulates
-Tryptophan and Serotonergic Synapse Metabolism and Induces Altered
-Behaviors in Germ-Free Mice. mSystems 6, e01343-20.
-https://doi.org/10.1128/mSystems.01343-20}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, X., Yang, Y.,
-Zhang, Y., Zhang, X.-M., Zhao, Z.-Q., Zhang, Y.-Q., 2013. Estrogen in
-the Anterior Cingulate Cortex Contributes to Pain-Related Aversion.
-Cereb. Cortex 23, 2190--2203. https://doi.org/10.1093/cercor/bhs201}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Yael, D., Tahary, O.,
-Gurovich, B., Belelovsky, K., Bar-Gad, I., 2019. Disinhibition of the
-Nucleus Accumbens Leads to Macro-Scale Hyperactivity Consisting of
-Micro-Scale Behavioral Segments Encoded by Striatal Activity. J.
-Neurosci. 39, 5897--5909.
-https://doi.org/10.1523/JNEUROSCI.3120-18.2019}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Yau, W.-Y.W., Zubieta,
-J.-K., Weiland, B.J., Samudra, P.G., Zucker, R.A., Heitzeg, M.M., 2012.
-Nucleus Accumbens Response to Incentive Stimuli Anticipation in Children
-of Alcoholics: Relationships with Precursive Behavioral Risk and
-Lifetime Alcohol Use. J. Neurosci. 32, 2544--2551.
-https://doi.org/10.1523/JNEUROSCI.1390-11.2012}
-
-Supplementary Tables
-
-\textbf{Supplementary Table XXA. List of most important taxa for Random
-Forest models on the 0 to 6 months age bracket, microbiome +
-demographics, ordered by relative weighted importance rank}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Taxon
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Rank
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Relative weighted importance
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Cumulative
-
-relative weighted importance
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\textbf{Supplementary Table XXA. List of most important taxa for Random
-Forest models on the 18 to 120 months age bracket, microbiome +
-demographics, ordered by relative weighted importance rank}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Taxon
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Rank
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Relative weighted importance
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Cumulative
-
-relative weighted importance
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\textbf{Supplementary Table XX. Benchmark metrics for each brain segment
-prediction experiment. Confidence intervals are calculated from the
-distribution of metrics from repeated CV at a confidence level of 95\%}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3933}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2440}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1846}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1782}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Brain segment
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Root-mean-square Error
-
-(RMSE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Mean Absolute Proportional Error (MAPE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Correlation coefficient (R)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Brain-stem
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0876
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.605
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-I-V
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0321
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.085
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.156
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-VI-VII
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0132
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.081
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.006
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-VIII-X
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0217
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.078
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-CSF
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0193
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.162
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.071
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-accumbens-area
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0020
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.094
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.288
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-accumbens-area
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0020
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.102
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.041
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-amygdala
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0037
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.288
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-amygdala
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.070
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.179
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-anterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0375
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.069
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.092
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-anterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0264
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.074
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.245
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-basal-forebrain
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0008
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-basal-forebrain
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0010
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.117
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.123
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-caudate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0257
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.099
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.029
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-caudate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0221
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.010
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-cerebellum-white-matter
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0689
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.413
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-cerebellum-white-matter
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0694
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.070
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.434
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-cuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0274
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.234
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-cuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0264
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.075
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.065
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-entorhinal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0107
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.080
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.190
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-entorhinal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0105
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.099
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-fusiform
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0510
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.333
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-fusiform
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0334
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.053
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.248
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-hippocampus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0128
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.054
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.075
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-hippocampus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0150
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.094
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-insula
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0294
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.119
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-insula
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0303
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-isthmus-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0161
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.041
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-isthmus-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0166
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.085
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.074
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-lateral-occipital
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0661
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.062
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.224
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-lateral-occipital
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0716
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.054
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.221
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-lingual
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0410
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.069
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.421
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-lingual
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0440
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.072
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.434
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-middle-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0875
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.052
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.000
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-middle-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0636
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.039
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-orbitofrontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0715
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.055
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.186
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-orbitofrontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0814
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.133
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pallidum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0075
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.064
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.297
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pallidum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0081
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.264
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-paracentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0335
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.078
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.131
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-paracentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0364
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.084
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-parahippocampal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0098
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.058
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-parahippocampal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0092
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.060
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-parietal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0755
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.177
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-parietal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1099
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.045
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.147
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-opercularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0197
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.146
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-opercularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0261
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.060
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.030
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-orbitalis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0157
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.003
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-orbitalis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0233
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.112
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.024
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-triangularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0226
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.174
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-triangularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0322
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.076
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.043
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pericalcarine
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0178
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.111
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.200
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pericalcarine
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0185
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.089
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.273
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-postcentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0495
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.052
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.168
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-postcentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0552
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.135
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-posterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0460
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.137
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-posterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0480
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.139
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-precentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0830
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.065
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.248
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-precentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0594
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.046
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.066
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-precuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0489
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.144
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-precuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0622
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.150
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-putamen
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0243
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-putamen
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0242
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.066
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.011
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-superior-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0867
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.040
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.101
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-superior-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.044
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.096
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-supramarginal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0767
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-supramarginal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0734
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.001
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-temporal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1530
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.040
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.128
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-temporal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1503
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.039
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-thalamus-proper
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0347
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.053
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.170
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-thalamus-proper
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0304
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.050
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.162
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\end{document}

### 2023-02-17 manuscript/tex/sections/methods.tex
clean up

--- 2023-02-16 manuscript/tex/converted.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-]{article}
-\usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\usepackage{longtable,booktabs,array}
-\usepackage{multirow}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\usepackage[normalem]{ulem}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-
-\author{}
-\date{}
-
-\begin{document}
-
-\hypertarget{title}{%
-\subsection{Title}\label{title}}
-
-\begin{itemize}
-\item
-  \begin{quote}
-  Gut microbes and their genes are associated with cognition and
-  neuroanatomy in children
-  \end{quote}
-\item
-  \begin{quote}
-  Gut microbes and their genes are associated with cognition performance
-  and brain structure in children
-  \end{quote}
-\item
-  \begin{quote}
-  Extensive relationships between the gut microbiota and normal
-  neurocognitive development
-  \end{quote}
-\item
-  \begin{quote}
-  Extensive relationships between the gut microbiota and typical
-  neurocognitive development
-  \end{quote}
-\item
-  \begin{quote}
-  The gut microbial metagenome is connected to neurocognition in typical
-  child development
-  \end{quote}
-\item
-  \begin{quote}
-  The microbial potential in regulating early childhood neurocognitive
-  development
-  \end{quote}
-\end{itemize}
-
-\hypertarget{authors}{%
-\subsubsection{Authors}\label{authors}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \begin{quote}
-  Bonham 0000-0003-3200-7533
-  \end{quote}
-\item
-  \begin{quote}
-  Bottino 0000-0003-1953-1576
-  \end{quote}
-\item
-  \begin{quote}
-  McCann 0000-0002-9753-7968
-  \end{quote}
-\item
-  \begin{quote}
-  Beauchemin \{ORCID??\}
-  \end{quote}
-\item
-  \begin{quote}
-  Elizabeth Weisse \textless elizabeth.weisse2@gmail.com\textgreater{}
-  \end{quote}
-\item
-  \begin{quote}
-  \ldots{}
-  \end{quote}
-\item
-  \begin{quote}
-  Jonathan O\textquotesingle Muircheartaigh
-  \textless{}\href{mailto:jonathanom@kcl.ac.uk}{\uline{jonathanom@kcl.ac.uk}}\textgreater{}
-  \end{quote}
-\item
-  \begin{quote}
-  Huttenhower 0000-0002-1110-0096
-  \end{quote}
-\item
-  \begin{quote}
-  Muriel Bruchhage
-  \textless{}\href{mailto:mmk.bruchhage@gmail.com}{\uline{mmk.bruchhage@gmail.com}}\textgreater{}
-  0000-0001-9637-0951
-  \end{quote}
-\item
-  \begin{quote}
-  D'Sa 0000-0002-3476-8419
-  \end{quote}
-\item
-  \begin{quote}
-  Deoni 0000-0002-0633-6328
-  \end{quote}
-\item
-  \begin{quote}
-  Klepac-Ceraj 0000-0001-5387-5706
-  \end{quote}
-\end{enumerate}
-
-\hypertarget{abstract}{%
-\subsection{Abstract}\label{abstract}}
-
-The gastrointestinal tract, its resident microorganisms and the central
-nervous system are connected by biochemical signaling, also known as
-"microbiome-gut-brain-axis." Both the human brain and the gut microbiome
-have critical developmental windows in the first three years of life,
-raising the possibility that their development is co-occurring and
-likely co-dependent. Emerging evidence implicates gut microorganisms and
-microbiota composition in cognitive outcomes and neurodevelopmental
-disorders (e.g., autism and anxiety), but the influence of gut microbial
-metabolism on typical neurodevelopment has not been explored in detail.
-We investigated the relationship of the microbiome with the neuroanatomy
-and cognitive function of 361 healthy children, demonstrating that
-differences in gut microbial taxa and gene functions are associated with
-overall cognitive function as well as with differences in the size of
-multiple brain regions. Using a combination of multivariate linear
-models and machine learning (ML) models, we showed that many species,
-including \emph{Gordonibacter pamelae} and \emph{Blautia wexlerae}, were
-significantly associated with higher cognitive function, while some
-species such as \emph{Ruminococcus gnavus} were more commonly found in
-children with low cognitive scores regardless of age or maternal
-education. Microbial genes for enzymes involved in the metabolism of
-neuroactive compounds, particularly short-chain fatty acids such as
-glutamate and propionate, were also associated with cognitive function.
-In addition, ML models were able to use microbial taxa to predict the
-volume of brain regions, with particular taxa often dominating the
-feature importance metric for individual brain regions, such as \emph{B.
-wexlerae} being the most dominant for the parahippocampal region or
-GABA-producing \emph{Bacteroidetes ovatus} for left accumbens. These
-findings provide potential biomarkers of neurocognition and may lead to
-the future development of targets for early detection and early
-intervention.
-
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
-
-The gut and the brain are intimately linked. Signals from the brain
-reach the gut through the autonomic nervous system and the endocrine
-system, and the gut can communicate with the brain through the vagus
-nerve and through endocrine and immune (cytokine) signaling molecules
-\href{https://www.zotero.org/google-docs/?i3SRPn}{(Cerdó et al., 2019;
-Pronovost and Hsiao, 2019; Sharon et al., 2016; Tognini, 2017)}. In
-addition, the products of microbial metabolism generated in the gut can
-influence the brain, both indirectly by stimulating the enteric nervous
-and immune systems, as well as directly through molecules that enter
-circulation and cross the blood-brain barrier. Causal links between the
-gut microbiome and neural development, particularly atypical
-development, are increasingly being identified
-\href{https://www.zotero.org/google-docs/?UsrFKB}{(Spichak et al.,
-2021)}. Both human epidemiology and animal models point to the effects
-of gut microbes on the development of autism spectrum disorder
-\href{https://www.zotero.org/google-docs/?Y8l5A0}{(Laue et al., 2020;
-Wan et al., 2021)} and particular microbial taxa have been associated
-with depression
-\href{https://www.zotero.org/google-docs/?mbaFE3}{(Mayneris-Perxachs et
-al., 2022; Valles-Colomer et al., 2019)} and Alzheimer's disease
-\href{https://www.zotero.org/google-docs/?0BEGUn}{(Fung et al., 2017;
-Kim et al., 2021)}. But information about this ``microbiome-gut-brain
-axis'' in normal neurocognitive development remains lacking,
-particularly early in life.
-
-The first years of life are critical developmental windows for both the
-microbiome and the brain
-\href{https://www.zotero.org/google-docs/?XTqhlm}{(Laue et al., 2022)}.
-Fetal development \emph{in utero} is believed to be mostly sterile, but
-is rapidly seeded at birth through contact with the birth canal (if
-birthed vaginally), caregivers, food sources (breastmilk or formula),
-and other environmental sources such as antibiotics
-\href{https://www.zotero.org/google-docs/?X6N0f4}{(Bäckhed et al., 2015;
-Bokulich et al., 2016; Dominguez-Bello et al., 2010; Louwies et al.,
-2020)}. The early microbiome is characterized by low microbial
-diversity, rapid succession and evolution, and domination by
-Actinobacteria, particularly the genus \emph{Bifidobacterium},
-Bacteroidetes, especially \emph{Bacteroides}, and Proteobacteria
-\href{https://www.zotero.org/google-docs/?VG43uh}{(Koenig et al.,
-2011)}. Many of these microbes have specialized metabolic capabilities
-for digesting human breast milk, such as \emph{Bifidobacterium infantis}
-and \emph{Bacteroides fragilis}
-\href{https://www.zotero.org/google-docs/?1dKzKj}{(Sela et al., 2008;
-Tso et al., 2021)}. Upon the introduction of solid foods, the gut
-microbiome undergoes another categorical transformation, with most taxa
-of the infant microbiome being replaced by taxa more reminiscent of
-adult microbiomes
-\href{https://www.zotero.org/google-docs/?V3lY3p}{(Bäckhed et al.,
-2015)}. Many prior studies have focused on either infant microbiomes or
-adult microbiomes, since performing statistical analyses across this
-transition poses particular challenges. Nevertheless, since this
-transition coincides with critical neural developmental windows and
-neural synaptogenesis, investigation across this solid-food boundary is
-incredibly important
-\href{https://www.zotero.org/google-docs/?d29bv8}{(Tau and Peterson,
-2010)}.
-
-A child's brain undergoes remarkable anatomical, microstructural,
-organizational, and functional changes in the first years of life. By
-age 5, a child's brain has reached \textgreater85\% of its adult size,
-has achieved near-adult levels of myelination, and the pattern of axonal
-connections has been established
-\href{https://www.zotero.org/google-docs/?1jUZOO}{(Silbereis et al.,
-2016)}. Much of this development occurs in discrete windows or
-``sensitive-'' or ``critical periods'' (CPs)
-\href{https://www.zotero.org/google-docs/?w4Pcze}{(Knudsen, 2004)} when
-neural plasticity is particularly high, and particular modes of learning
-and skill development are preferred. The timing of these sensitive
-periods is driven in part by genetics, but can also be affected by the
-environment, including the gut microbiome
-\href{https://www.zotero.org/google-docs/?KTm9lQ}{(Cowan et al., 2020)}.
-In fact, emerging evidence suggests that the timing and duration of CPs
-may be driven in part by cues from the developing gut microbiome
-\href{https://www.zotero.org/google-docs/?mAD1ne}{(Callaghan, 2020)}. As
-such, understanding the normal spectrum of healthy microbiome
-development and how it relates to normal neurocognitive development may
-provide opportunities for identifying atypical development earlier and
-offer opportunities for intervention.
-
-To begin to address this need, we investigated the gut microbiome and
-neurocognitive development of children from infancy through 10 years of
-age in an accelerated longitudinal study design. Gut microbial
-communities were assessed using shotgun metagenomic sequencing, enabling
-profiling at both the taxonomic and gene-functional level, and
-neurocognitive development was measured using expert-assessment of
-cognitive function and neuroimaging using magnetic resonance imaging
-(MRI). Using a combination of classical statistical analysis and machine
-learning, we showed that the development of the gut microbiome, the
-brain, and children's cognitive abilities are intimately linked, with
-both microbial taxa and gene functions able to predict cognitive
-performance and brain structure.
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-\hypertarget{cohort-overview-and-summary-data}{%
-\subsubsection{Cohort overview and summary
-data}\label{cohort-overview-and-summary-data}}
-
-We investigated the co-development of the brain and the microbiome in
-early childhood in a cohort of typically-developing children in their
-first years of life using a variety of orthogonal microbial and
-neurocognitive assessments, including shotgun metagenomic sequencing,
-cognitive and behavioral assessments, and neuroimaging. In all, 361
-children from RESONANCE, an accelerated longitudinal cohort of early
-development \href{https://paperpile.com/c/dPbU4e/klxU}{(Forrest,
-Blackwell, and Camargo 2018)} between the ages of 82 days and 10 years
-old (median age 2.21 years, Supplementary Figure 1) were included in
-this study (Table 1, Figure 1A). To measure cognitive function, we used
-age-appropriate assessments that can be normalized to a common, IQ-like
-scale (Figure 1B), including the Mullen scale of early learning (MSEL)
-for children from birth to 3 years of age
-\href{https://www.zotero.org/google-docs/?90hfYr}{(Mullen and others,
-1995)}, Wechsler preschool and primary scale of intelligence (WPPSI) for
-4--5-year-olds
-\href{https://www.zotero.org/google-docs/?wCfx5h}{(Wechsler, 2012)}, and
-the Wechsler intelligence scale for children (WISC) for children 6 years
-and up \href{https://www.zotero.org/google-docs/?AcNN3S}{(Wechsler,
-1949)}.
-
-As expected, the greatest differences in microbial taxa were driven by
-age (Figure 1B-C, PCoA axis 1), while older children were primarily
-stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
-high-\emph{Prevotella copri} (Supplementary Figure 2). Overall variation
-in gut microbial genes and also in brain volume profiles was similarly
-driven largely by subject age (Figure 1C-F).
-
-\includegraphics[width=6.45833in,height=3.87221in]{media/image3.png}
-
-\textbf{Figure 1: Cohort overview and summary data.}
-
-(A) Study design and overview. Stool samples, cognitive assessments and
-neuroimaging were collected from participants in an accelerated
-longitudinal design. (B) Cognitive function scores are assessed by
-different instruments, but may be normalized to a common IQ-like scale.
-(C-D) Principal coordinates analysis using Bray-Curtis dissimilarity on
-taxonomic profiles demonstrates high beta diversity, with much of the
-first axis of variation explained by increasing age and alpha diversity.
-Differences in gene function profiles (E) and neuroimaging (PCA based on
-the Euclidean distance of brain region volumes) (F) are likewise
-dominated by changes as children age. (G) Permutation analysis of
-variance (PERMANOVA) of taxonomic profiles, functional profiles
-(annotated as UniRef90s) and neuroimaging, against the metadata of
-interest. Variations in taxonomic and functional profiles explain a
-modest but significant percent of the variation in cognitive development
-in children over 18 months of age. (H) Mantel testing of different
-microbial feature matrices, shows overlapping but distinct patterns of
-variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
-used as cut-offs in some following models.
-
-\textbf{Table1: Subject demographics}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2620}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Group}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Subgroup}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{All}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Under 6 mo}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Over 18 mo}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-N subjects
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-361
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-74
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-257
-\end{minipage} \\
-\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Samples
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-1
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-262 (72.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-74 (100.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-203 (79.0\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-72 (19.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-46 (17.9\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-\textgreater2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-27 (7.5\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-8 (3.1\%)
-\end{minipage} \\
-\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Age (months)
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-Min
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18.0
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Max
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-119.3
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6.0
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-119.3
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Median
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-25.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-3.7
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-45.9
-\end{minipage} \\
-\multirow{2}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Sex
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-F
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-163 (45.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (50.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-114 (44.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-M
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-198 (54.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (50.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-143 (55.6\%)
-\end{minipage} \\
-\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Race
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-White
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-252 (69.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-42 (56.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-191 (74.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Black
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-40 (11.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16 (21.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-21 (8.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Asian
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Mixed
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-7 (1.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (1.4\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6 (2.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Other
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-48 (13.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12 (16.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-29 (11.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Unknown/Declined
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2 (0.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (1.4\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2 (0.8\%)
-\end{minipage} \\
-\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Maternal Ed.
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-Junior high school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Some high school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6 (1.7\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-3 (4.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-4 (1.6\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-High school grad
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (10.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-13 (17.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16 (6.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Some college
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-95 (26.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-25 (33.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-57 (22.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-College grad
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-90 (24.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17 (23.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-69 (26.8\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Grad/professional school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-123 (34.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-15 (20.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-104 (40.5\%)
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-Several studies have demonstrated links between specific taxa and
-measures of anxiety and depression
-\href{https://www.zotero.org/google-docs/?X5rfKj}{(Mayneris-Perxachs et
-al., 2022; Needham et al., 2022)}, cognitive flexibility in adults
-\href{https://www.zotero.org/google-docs/?RGIf8m}{(Magnusson et al.,
-2015)}, and atypical neural development
-\href{https://www.zotero.org/google-docs/?hwPZKh}{(Liu et al., 2019; Wan
-et al., 2021)}. We, therefore, set out to identify whether specific taxa
-or gene functions were linked with normal cognitive development in
-children. In order to assess whether variations in gut microbial taxa,
-their genes, or their metabolism are linked with neurocognitive
-development, we tested whether the beta diversity of microbial taxa and
-gene functions, as well as variation in brain development as assessed by
-neuroimaging, were associated with these contemporaneous measures of
-cognitive function using permutation analysis of variance (PERMANOVA).
-Due to the large ecological shift in the microbiome that occurs upon the
-introduction of solid food, and a relatively wide range of ages when
-infants are transitioned to solid food, we considered the pre-transition
-(less than 6 months old) and post-transition (over 18 months old)
-microbiomes separately. Even still, age was a major driver of variation
-in gut microbiomes in children over 18 months old (Figure 1G). We also
-found that overall variation in microbial species in children over 18
-months old was associated with a small but significant variation in
-cognitive function score (Figure 1G, R\textsuperscript{2} = 0.0124, q
-\textless{} 0.001), as was variation in microbial gene functions
-(R\textsuperscript{2} = 0.0119, q \textless{} 0.001). Variation in
-microbial taxa and genes was not significantly associated with cognitive
-function in children under 6 months, though this may be due to the low
-taxonomic diversity, and broad lack of overlap between taxa in infants.
-As expected, age was significantly associated with microbial beta
-diversity (taxa R\textsuperscript{2} = 0.0215, and UniRef90s
-R\textsuperscript{2} = 0.0245, q \textless{} 0.001), and very strongly
-associated with neuroimaging profiles (R\textsuperscript{2} = 0.256, q
-\textless{} 0.001).
-
-Consistent with prior studies, different microbial measurement types
-captured overlapping variation, with species profiles and gene function
-profiles (annotated with clusters of 90\% similarity
-\href{https://www.zotero.org/google-docs/?b9rEJS}{(Suzek et al.,
-2007)}), both generated from DNA sequencing data, being tightly coupled
-(Figure 1H, p \textless{} 0.001). Interestingly, other functional
-groupings (Enzyme commission level4 - ECs
-\href{https://www.zotero.org/google-docs/?fs8LO7}{(Bairoch, 2000)}, and
-KEGG orthologues - KOs
-\href{https://www.zotero.org/google-docs/?Xx5duE}{(Kanehisa et al.,
-2004)}) overlapped only slightly with taxonomic profiles in both age
-cohorts, despite being derived from UniRef90 labels. In children over 18
-months, some variation (15.9\%, p \textless{} 0.01) in neuroimaging
-overlapped with microbial measures, though this may be due to the
-residual variation due to age in both measures.
-
-\hypertarget{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}{%
-\subsubsection{Microbial species and neuroactive genes are associated
-with cognitive
-performance}\label{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}}
-
-To assess whether individual microbial species were associated with
-cognitive function, we fit multivariable linear regression
-\href{https://www.zotero.org/google-docs/?4qS6UF}{(Mallick et al.,
-2021)} to the relative abundance of each species that had at least 15\%
-prevalence in a given age group (Figure 2A, N = 92 for 0--120 months, N
-= 46 for 0--6 months, N = 97 for 18--120 months). Only \emph{Blautia
-wexlerae} was significantly associated (q value = 0.14, β = 0.0015) with
-cognitive function in children under 6 months old after adjusting for
-age and maternal education (Figure 2B). \emph{B. wexlerae} was
-previously shown to be depleted in children with diabetes
-\href{https://www.zotero.org/google-docs/?6mU2nI}{(Benítez-Páez et al.,
-2020)}, and that oral administration of \emph{B. wexlerae} partially
-ameliorated weight gain and inflammation from a high-fat diet in a mouse
-model of T2D \href{https://www.zotero.org/google-docs/?y83hpS}{(Hosomi
-et al., 2022; Liu et al., 2021)}. In children over 18 months of age,
-several microbial species were significantly enriched (q-value
-\textless{} 0.20) in children with higher cognitive function scores,
-including \emph{Gordonibacter pamelaeae}, which produces the
-neuroprotective metabolite urolithin
-\href{https://www.zotero.org/google-docs/?HGtTEY}{(Gong et al., 2022;
-Selma et al., 2014)}, \emph{Asaccharobacter celatus} and
-\emph{Adelcreutzia equolifaciens}, which produce phytoestrogen-derived
-equol \href{https://www.zotero.org/google-docs/?HCKZqm}{(Maruo et al.,
-2008; Thawornkuno et al., 2009)}, and the SCFA-producing probiotic
-species such as \emph{Eubacterium eligens} and \emph{Faecalibacterium
-prausnitzii} \href{https://www.zotero.org/google-docs/?3eyheK}{(Ghosh et
-al., 2020; Lopez-Siles et al., 2017; Mukherjee et al., 2020)} (Figure
-2B).
-
-\includegraphics[width=6.5in,height=4.83333in]{media/image4.png}
-
-\textbf{Figure 2 - Taxa and gene functional groups are associated with
-cognitive function.}
-
-(A) Volcano plot of multivariable linear model results showing the
-relationship between individual taxa and cognitive function score in
-children over 18 months of age, controlling for age, maternal education,
-and sequencing depth. Taxa that were significant after FDR correction (q
-\textless{} 0.2) are colored red. (B) For taxa that were significantly
-associated with cognitive function, heatmaps of prevalence, mean
-relative abundance, and correlation with cognitive function in different
-age groups. (C-F) Enrichment plots for selected neuroactive gene sets
-used in feature set enrichment analysis (FSEA). (G) summary of FSEA
-results across all samples (left) as well as in the under 6 months
-(middle) and over 18 months (right) subsets, colored based on the
-significance of the association. Markers indicate the individual
-correlation of genes within a gene set, and vertical bars represent the
-median correlation of that gene set.
-
-Given that different microbial species might occupy the same metabolic
-niche in different individuals, we hypothesized that microbial genes
-grouped by functional activity would be associated with cognition. To
-test this, we performed feature set enrichment analysis (FSEA) on groups
-of genes with neuroactive potential
-\href{https://www.zotero.org/google-docs/?xQURsj}{(Valles-Colomer et
-al., 2019)} and concurrent cognitive function score (Table 2,
-\protect\hyperlink{_i7fd3z19jdvg}{\uline{Figure 2}}C-G) and found that
-several metabolic pathways were significantly enriched or depleted in
-children with higher cognitive function scores. This was true both when
-considering all age groups together, though the enrichment of most
-pathways was more pronounced in children under 6 months or over 18
-months. For example, genes for degrading the 3-carbon SCFA propionate
-were significantly depleted in children with higher cognitive function
-scores across all age groups tested (Table 2; propionate degradation I,
-under 6 months, enrichment score (E.S.) = -0.542, corrected p-value (q)
-= 0.020; over 18 months, E.S. = -0.674, q = 0.041). Interestingly, genes
-for propionate synthesis were also significantly depleted in higher
-scoring children over 18 months (E.S. -0.676, q = 0.023), as were genes
-for synthesizing the 2 carbon SCFA acetate in children under 6 months
-old (acetate synthesis I, E.S. = -0.194, q = 0.153; acetate synthesis
-II, E.S. = -0.342, q = 0.020; acetate synthesis III, E.S. = -0.31, q =
-0.052). SCFAs are produced by anaerobic fermentation of dietary fiber
-and have been linked with immune system regulation as well as directly
-with brain function
-\href{https://www.zotero.org/google-docs/?2Z0sKZ}{(Dalile et al.,
-2019)}.
-
-Synthesis of menaquinone (vitamin K) was also negatively associated with
-cognitive function score in older children (menaquinone synthesis I,
-E.S. = -0.170, q = 0.0183). Menaquinone has several isoforms, one of
-which, MK-4, has been found to be decreased in children diagnosed with
-ASD \href{https://www.zotero.org/google-docs/?dTA3NO}{(Dong et al.,
-2021)} and is potentially neuroprotective in both rodents and humans
-\href{https://paperpile.com/c/dPbU4e/cTO9}{(Elkattawy et al., 2022)}. In
-children over 18 months old, genes for the synthesis of the amino acids
-glutamate and tryptophan were significantly enriched in children with
-higher cognitive function scores (Glutamate synthesis I, E.S. = 0.242, q
-= 0.047; Tryptophan synthesis, E.S. = 0.119, q = 0.041). Glutamate is a
-critical neurotransmitter controlling neuronal excitatory/inhibitory
-signaling along with gamma-aminobutyric acid (GABA), and their balance
-in the brain controls neural plasticity and learning, particularly in
-the developing brain
-\href{https://www.zotero.org/google-docs/?QubMhx}{(Cohen Kadosh et al.,
-2015; Palomo-Buitrago et al., 2019)}. Tryptophan metabolism, including
-microbial metabolism of tryptophan, has previously been linked with
-autism in children
-\href{https://www.zotero.org/google-docs/?GPaUZE}{(Hoshino et al., 1984;
-Xiao et al., 2021)}. Taken together, these results suggest that
-microbial metabolic activity, particularly the metabolism (synthesis and
-degradation) of neuroactive compounds may have effects on cognitive
-development.
-
-\textbf{Table 2 - Feature set enrichment analysis on neuroactive
-microbial gene sets}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2300}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1259}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1525}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1429}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2785} + 2\tabcolsep}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{All}
-\end{minipage}} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2591} + 2\tabcolsep}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{under 6 months}
-\end{minipage}} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2324} + 2\tabcolsep}@{}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{over 18 months}
-\end{minipage}} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-geneset
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1223
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5549
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1944
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1533}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1269
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. II
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2201
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0660}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3424
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0198}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0946
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9199
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. III
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2104
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0859}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3098
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0518}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0857
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9199
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Glutamate deg. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.4037
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1802}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2840
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2489
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Glutamate syn. II
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0953
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6022
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1342
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9802
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2421
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0466}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Menaquinone syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0853
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5635
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0833
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1699
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1825}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Propionate deg. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.4612
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1302}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.5421
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0198}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.6738
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0414}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Propionate syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3344
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5549
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3201
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5713
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.6760
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0227}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Tryptophan syn.
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0573
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9742
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0725
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1186
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0414}
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\hypertarget{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}{%
-\subsubsection{Gut microbial taxonomic and functional profiles predict
-cognitive
-function}\label{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}}
-
-FSEA relies on understanding functional relationships between individual
-genes. However, because the relationships between individual taxa are
-still largely unknown, we turned to Random Forest (RF) models, an
-unsupervised non-parametric machine learning (ML) method that enables
-the identification of underlying patterns in large numbers of individual
-features (here, microbial species). Previous studies have reported
-successful use of RFs for processing highly-dimensional and sparse data
-from the domain of genomics
-\href{https://www.zotero.org/google-docs/?C3CSPB}{(Amaratunga et al.,
-2008; Brieuc et al., 2018; Chen and Ishwaran, 2012; Franzosa et al.,
-2019; Stephan et al., 2015)}, along with other works where it was used
-to predict cognitive conditions related to Alzheimer's disease in
-different scenarios
-\href{https://www.zotero.org/google-docs/?GiadGy}{(Ardekani et al.,
-2017; Velazquez et al., 2021)}. Additionally, given the sequential
-nature of variable consideration in each tree, RFs are naturally able to
-work out complex input feature interactions, such as those present in a
-microbiome-wide study, without the necessity to explicitly compute
-interaction terms.
-
-\includegraphics[width=6.5in,height=5.19444in]{media/image2.png}
-
-\textbf{Figure 3 - Random forest models predict concurrently measured
-cognitive function.}
-
-Comparison of RF predictor importance versus linear models for children
-between birth and 6 months old (A), and for those older than 18 months
-(B). Colors represent whether the species belong to the group of
-top-important features that account for 80\% of the cumulative
-importance on the RF model, if that species was significant (q
-\textless{} 0.2) in linear models, both, or neither. Ranked feature
-importance for taxa in RF models for children between birth and 6 months
-old (C), and for those older than 18 months (D). Taxa that are important
-for RF models both for children under 6 months and for those over 18
-months (E), only for children under 6 months (F), and only for children
-over 18 months (G)
-
-Given that gut microbial profiles, as well as neurocognitive
-development, may partially reflect socioeconomic and demographic
-factors, we assessed the performance of RF regressors where maternal
-education (a proxy of socioeconomic status (SES)), sex, and age were
-included as possible predictors, either alone or in combination with
-microbial taxonomic profiles (Table 3).
-
-\textbf{Table 3. Benchmark metrics for the cognitive assessment score
-prediction models. Confidence intervals are calculated from the
-distribution of metrics from repeated CV at a confidence level of 95\%}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1456}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0960}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3584}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Subject Ages (months)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Microbial feature
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Demo.
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Test set correlation (± C.I.)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Test set Root-mean-square error (± C.I.)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.14 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-13.01 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.10 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.66 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.13 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.70 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.01 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.56 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.01 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.56 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.506 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16.27 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.363 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.66 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.429 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.29 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.303 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18.01 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.333 ± 0.004
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.86 ± 0.04
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-As with linear models, RF models for children under 6 months old were
-less generalizable (mean test-set correlation -0.13, mean RMSE 12.70),
-but RFs were consistently able to learn the relationship between taxa
-and cognitive function scores in children over 18 months of age (mean
-test-set correlation 0.429, mean RMSE 17.29). For both age groups,
-species that were important in RF overlapped with those that were
-significant when testing the relationship with linear models (Figure
-3A-B, Supplementary Table XXA-B), though there were substantial
-differences. For children under 6 months, only \emph{B. wexlerae} was
-significantly associated with cognitive function in linear models, but
-was ranked 20th in importance for RF models. The importance metric
-employed (MDI) is a measure of how well the variable differentiates
-(splits) sets of samples by creating subgroups that reduce the
-intra-group deviations, while also accounting for the amount of samples
-affected by a split. All taxa significantly associated with cognitive
-function score in children over 18 months using LMs belong to the
-top-ranking group responsible for 80\% of the total relative importance,
-except for \emph{Eubacterium ramulus} (Table 3, Figure 3A).
-
-Interestingly, several taxa highly ranked in importance in both age
-groups, including several that were significant in LMs for children over
-18 months old, including \emph{R. gnavus} (0--6 months, rank = 13;
-18--120 months, rank = 7) and \emph{G. pamelaeae} (0--6 months, rank =
-16; 18--120 months, rank = 20), while others such as \emph{Allistipes
-finegoldii} were age-group specific. Several taxa important in RF models
-were not statistically significant when using linear models after
-multiple hypothesis correction. However, these taxa had small nominal
-p-values. For example, \emph{Erysipelatoclostridium ramosum} was the
-most important feature in RF models for children under 6 months old and
-had an LM p-value of 0.04. Subject age was consistently ranked highly in
-feature importance, which could indicate that decision branches based on
-microbial taxa have increased purity when considering the subject's age
-or that age itself is a useful predictor.
-
-\hypertarget{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}{%
-\subsubsection{Gut microbial taxonomic profiles predict brain structure
-differences}\label{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}}
-
-If there are causal effects of microbial metabolism on cognitive
-function, they might be reflected in changes in neuroanatomy. We again
-employed a Random Forest modeling approach to associate gut taxonomic
-profiles with individual brain regions identified in MRI scans,
-normalized to total brain volume. Some brain regions were more readily
-predicted by RF models trained on microbial taxa (Table 4, Supplementary
-Table XX), in particular those that were highly correlated with age.
-These included the L/R lingual gyrus (mean RF correlations, Left =
-0.421, Right = 0.434; relative age importances, Left = 7.5\%, Right =
-8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
-0.200, Right = 0.273; relative age importances, Left = 3.7\%, Right =
-6.3\%). In many cases, however, age was not an important variable in
-high-performing models, such as that for the left accumbens area (mean
-RF correlations = 0.288; relative age importances = 1.2\%).
-
-\textbf{Table 4. Summary statistics for the neuroanatomy prediction
-benchmarks}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Statistic
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Mean absolute proportional error (MAPE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Correlation coefficient (R)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-mean
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.072
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.137
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Standard deviation
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.028
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.136
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-maximum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.605
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-75th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.195
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-50th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-25th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.062
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-minimum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.147
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-We also observed that many brain regions had high accordance across the
-left and right hemispheres in terms of both model performance and
-microbial feature importance. In contrast, other regions had substantial
-differences between the hemispheres. For example, the left accumbens
-area, which plays an important role in reward circuits
-\href{https://www.zotero.org/google-docs/?zVicU7}{(Ernst et al., 2005;
-Yau et al., 2012)}, has one of the highest test-set correlations of our
-brain region models (R = 0.288), as compared to the right accumbens
-models which could not adequately generalize, and had a negative mean
-test-set correlation (R = -0.041). Most healthy individuals have a
-rightward asymmetry in the nucleus accumbens, and reduced asymmetry has
-been linked to substance use disorder in young adults
-\href{https://www.zotero.org/google-docs/?V45KyN}{(Cao et al., 2021)}.
-Feature importance for models of the left accumbens area were dominated
-by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.4\%
-relative importance), \emph{B. ovatus} (3.7\% relative importance), and
-\emph{B. uniformis} (3.0\% relative importance).. The accumbens area is
-associated with reward control, and in individuals diagnosed with ADHD,
-it has been shown to have a divergent neuromorphology
-\href{https://www.zotero.org/google-docs/?kLfEQ5}{(Hoogman et al.,
-2017)}. Independently, \emph{B. ovatus}, \emph{B. uniformis} and
-\emph{B. vulgatus} have been linked to ADHD
-\href{https://www.zotero.org/google-docs/?luO0zh}{(Wang et al., 2020)}.
-In fact, there are studies showing that alterations on the striatal
-dopamine transporters can cause effects resembling hyperactivity and
-attention deficit
-\href{https://www.zotero.org/google-docs/?UxS47e}{(Yael et al., 2019)},
-and that \emph{B. uniformis} is gut-microbial modulator of the brain
-dopamine transporter
-\href{https://www.zotero.org/google-docs/?L1cWJS}{(Hartstra et al.,
-2020)}.
-
-Many of the taxa identified included taxa also identified in LMs and RF
-models of cognitive function. Interesting to note, while RF models for
-multiple brain regions had many important microbial taxa, others were
-dominated by a small number of taxa. In general, for all segments, a
-consistent number of species between 14 and 22 (median = 20) was
-responsible for one third of the fitness-weighted cumulative importance
-(see Methods), regardless of the model benchmarks. We proceeded to
-select a subset of the most important taxa (averaged over all segments)
-and the segments whose importance was more heavily loaded to further
-analyze the multiple relationships unveiled by the RF models
-(\textbf{Figure 4 A and B})
-
-\includegraphics[width=6.5in,height=5.19444in]{media/image1.png}
-
-\textbf{Figure 4: Microbial feature importance in predicting brain
-volumes in children over 18 months of age.}
-
-(A) Average test-set correlations for prediction of MRI segmentation
-data from microbiome and demographics on select segments after repeated
-cross validation. (B) Heatmap of average individual relative taxa
-importances on each brain segment. Importances are reported as
-proportions relative to the sum of importances for each model - since
-every model is trained with 132 features, and even distribution of
-importance would be 0.75\% for each feature.
-
-. Segments and taxa ordered by HCA on a select list of species with high
-ranks on average importances, and their respective highest-load
-segments. (C)
-
-Our analysis revealed two major patterns of importance distribution from
-taxa over the brain segments; some species portrayed high contributions
-to multiple different segments, while others contributed modestly to
-just one or two brain segments. Notable cases of the first pattern
-included seven species - \emph{Anaerostipes hadrus}, \emph{Bacteroides
-vulgatus}, \emph{Fusicatenibacter saccharivorans}, \emph{Ruminococcus
-torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
-\emph{Blautia wexlerae} - that combined, account for approximately one
-third of the cumulative relative importances, computed after subsetting
-on the taxa of interest (Figure 4C).
-
-Among these, the most important variable is \emph{Anaerostipes hadrus},
-which is a butyrate-producing anaerobe that has been positively
-associated with cognitive function
-\href{https://www.zotero.org/google-docs/?AVJiwu}{(Kant et al., 2015; Li
-et al., 2022)}. Its importance is, however, heavily loaded on the volume
-of entire major lobes (frontal, parietal), which cannot be readily
-correlated to specific cognitive functions. After accounting for the
-major lobes, it is found to be highly important on the prediction of the
-cerebellar vermal lobules VIII-X, pars opercularis, cuneus and
-precuneus, and anterior-cingulate (Figure 4B).
-
-To better understand this relationship, we performed hierarchical
-cluster analysis, which revealed a well defined cluster of species with
-high importance loadings on close segments of the lower temporal and
-close occipital lobe, to which \emph{A. hadrus} belongs. The other
-species on this cluster are \emph{B. wexlerae}, \emph{R. torques},
-\emph{R. intestinalis, R. bicirculans} and \emph{F. saccahrivorans}, who
-contribute heavily to the entorhinal, fusiform, lingual and
-parahippocampal segments.
-
-This group of taxa and their related segments drew our attention because
-they contained \emph{B. wexlerae}, a highlight from the cognitive
-assessment results. Previous works exploring Parkinson Disease patients
-found out that issues in the cognitive task of confrontation naming were
-positively correlated with thinning in the fusiform gyrus and
-parahippocampal gyrus
-\href{https://www.zotero.org/google-docs/?rlO6eq}{(Pagonabarraga et al.,
-2013)}. Additionally, the left and right parahippocampal, where \emph{B.
-wexlerae} had the highest importance in both models, are important in
-visual/spatial processing and memory
-\href{https://www.zotero.org/google-docs/?wkkWZT}{(Aminoff et al.,
-2013)}. \emph{B. wexlerae} can also produce acetylcholine in the gut
-\href{https://www.zotero.org/google-docs/?jBoxSm}{(Hosomi et al.,
-2022)}, and this molecule plays an important role in modulating memory
-function \href{https://www.zotero.org/google-docs/?BQ0j2c}{(Haam and
-Yakel, 2017)}.
-
-Another notable importance cluster contains taxa associated with the
-basal forebrain, especially the cingulate and the accumbens. This
-cluster contains the previously-discussed \emph{B. ovatus} and \emph{B.
-uniformis,} but also \emph{Alistipes finegoldii and Streptococcus
-salivaris.} Reduction in the nucleus accumbens has been associated with
-depression symptoms
-\href{https://www.zotero.org/google-docs/?2aSZIr}{(Wacker et al.,
-2009)}, and increased levels of the \emph{Alistipes} genus have been
-observed in patients with depressive disorder
-\href{https://www.zotero.org/google-docs/?bhhSZB}{(Jiang et al., 2015)}.
-While these reports are important, simultaneously probing of the gut
-metagenome and brain structure relationship is novel.
-
-Both \emph{A. equolofaciens} and \emph{A. celatus,} two closely-related
-equol-producing species, are examples of the second contribution
-pattern, and had high importance in predicting the relative volume of
-the right anterior cingulate (respectively, 3.0\% and 2.6\% relative
-importances), which has been linked to social cognition and reward-based
-decision making \href{https://www.zotero.org/google-docs/?Wi8Bpq}{(Apps
-et al., 2016; Boes et al., 2008; Bush et al., 2002)}. Equol has a strong
-estrogenic effect
-\href{https://www.zotero.org/google-docs/?PqgFU7}{(Setchell et al.,
-2005)}, and in anterior cingulate, estrogen, has been shown to regulate
-pain-related aversion
-\href{https://www.zotero.org/google-docs/?5huRIR}{(Xiao et al., 2013)}.
-Another example of this pattern is \emph{R. gnavus,} which was the only
-species with significant negative association with cognitive function.
-It is heavily associated with the left pars opercularis (2.7\% relative
-importance), a relationship that may be explained by the emerging
-understanding that this species is increased in individuals with insulin
-resistance and obesity
-\href{https://www.zotero.org/google-docs/?lwzWUv}{(Ley et al., 2006)},
-conditions that are known to produce structural abnormalities in the
-brain \href{https://www.zotero.org/google-docs/?NlkFr6}{(Opel et al.,
-2021)}. Finally, \emph{Coprococcus comes} displays an importance
-distribution that splits almost evenly among the two previously reported
-clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (4.0\% relative importance)
-(only age had higher relative importance in any model), while also being
-one of the most important predictors for neighboring areas of the
-posterior cingulate such as the pars opercularis (relative importances,
-Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
-precentral (2.3\% relative importance) and paracentral lobes (relative
-importances, Left = 2.6\%, Right = 1.8\%).
-
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
-
-The relationship between the gut microbiome and brain function via the
-gut-microbiome-brain axis has gained increasing acceptance largely as a
-result of human epidemiological studies investigating atypical
-neurocognition (eg anxiety and depression, neurodegeneration, attention
-deficit / hyperactivity disorder, and autism) and mechanistic studies in
-animal models. The results from these studies point to the possibility
-that gut microbes and their metabolism may be causally implicated in
-cognitive development, but this study is the first to our knowledge that
-directly investigates microbial species and their genes in relation to
-typical development in young children. Understanding the gut-brain-axis
-in early life is particularly important, since differences or
-interventions in early life can have outsized and longer-term
-consequences than those at later ages. Further, even in the absence of
-causal impacts of microbial metabolism, identifying risk factors that
-could point to other early interventions would also have value.
-
-The use of shotgun metagenomic sequencing enabled us to get
-species-level resolution of microbial taxa. A previous study of
-cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing,
-and showed that genera from the Lachnospiraceae family as well as
-unclassified Clostridiales (now Eubacteriales) were associated with
-higher scores on the Ages and Stages Questionnaire
-\href{https://www.zotero.org/google-docs/?SB5Lwa}{(Sordillo et al.,
-2019)}. However, each of these clades encompass dozens of genera with
-diverse functions, each of which may have different effects. Indeed,
-several of the taxa that were positively associated with cognitive
-function in this study, including \emph{B. wexlerae}, the only species
-identified by linear models in children under 6 months, \emph{D.
-longicatena}, \emph{R. faecis}, and \emph{A. finegoldii} are
-Clostridiales, as is \emph{R. gnavus}, which we found was negatively
-associated with cognitive function (Figure 2A-B). This kind of
-species-level resolution is typically not possible with amplicon
-sequencing.
-
-We identified several species in the family Eggerthelaceae that were
-associated with cognitive function, including \emph{Gordonibacter
-pamelaeae, Aldercreutzia equolofaciens, Asaccharobacter celatus}
-(formerly regarded a subspecies of \emph{A. equolofaciens}
-\href{https://www.zotero.org/google-docs/?OEhguI}{(Takahashi et al.,
-2021)}), and \emph{Eggerthella lenta}. Many members of this family are
-known in part due to unique metabolic activities. For example, \emph{A.
-equolofaciens} produces the nonsteroidal estrogen equol from isoflaven
-found in soybeans
-\href{https://www.zotero.org/google-docs/?I2qNyB}{(Wang et al., 2005)},
-and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
-(found in pomegranates and some berries) into urolithin, which has been
-shown in some studies to have a neuroprotective effect
-\href{https://www.zotero.org/google-docs/?oTN8Lw}{(Gong et al., 2022;
-Selma et al., 2014)}. \emph{E. lenta} has been extensively studied for
-its ability to metabolize drug compounds such as the plant-derived heart
-medication digoxin
-\href{https://www.zotero.org/google-docs/?tH2Qeq}{(Haiser et al.,
-2013)}. The metabolic versatility of this clade, and the large number of
-species that are associated with cognition make these microbes prime
-targets for further mechanistic studies.
-
-In addition to improved species-level resolution, shotgun metagenomic
-sequencing also enables gene-functional insight. We showed here that
-genes for the metabolism of SCFAs, both their degradation and synthesis,
-are associated with cognitive function scores. However, while the
-differential abundance of genes for the metabolism of neuroactive
-compounds like these is suggestive, it is difficult to reason about the
-relationship between levels of these genes and the gut concentrations of
-the molecules their product enzymes act on. For example, while it might
-be intuitive to reason that increased levels of menaquinone synthesis
-genes is indicative of increased menaquinone, it could be the case that
-menaquinone deficiency selects for microbes that can synthesize it. For
-the same reason, increased propionate degradation genes may
-counterintuitively be indicative of high levels of propionate in the gut
-lumen, since high propionate would select for microbes that can
-metabolize it. For this reason, future studies coupling shotgun
-metagenomics with stool metabolomics could improve our understanding of
-the relationship between microbial metabolism and cognitive development.
-Further, strain-level analysis linking specific gene content in species
-of interest could further refine targeted efforts at identifying
-specific metabolic signatures of microbe-brain interactions.
-
-\{\{Guilherme add a paragraph about the usefulness of ML\}\}
-
-The use of multiple age-appropriate cognitive assessments that could be
-normalized to a common scale enabled us to analyze microbial
-associations across multiple developmental periods, but carries several
-drawbacks. In particular, the test-retest reliability, as well as
-systematic differences between test administrators may introduce
-substantial noise into these observations, particularly in the youngest
-children. In addition, our study period overlapped with the beginning of
-the COVID-19 pandemic, and we and others have observed some reduction in
-measured scores for children that were assessed after the implementation
-of lockdowns. In our subject set for this study, these effects are more
-pronounced in some age groups due to our sampling schedule
-\href{https://www.zotero.org/google-docs/?L8GnRj}{(Blackwell et al.,
-2022; Deoni et al., 2021)} (Supplementary Figure 3).
-
-This analysis allowed us to establish links between microbial taxa and
-their functional potential with cognition and brain structure. Although
-we cannot test causality or the chemistry behind the interactions
-between gut microbial taxa, gut, and brain, this study provides clear
-and statistically significant associations between the infant and early
-child gut microbiota and neurocognition. Future studies should focus on
-characterizing the early-life microbiome and neurocognitive development
-across different geographic regions and lifestyles such as covering
-traditionally understudied low-resource urban, peri-urban and rural
-communities to obtain the more comprehensive understanding of the
-variability within the different gut microbiomes reflects on
-neurocognition. These studies would also provide us with the wealth of
-data on different strains from the same species to better understand the
-effect of genes and their products. Furthermore, culturing and microbial
-community enrichment studies combined with genetic manipulation and
-genomic approaches to understand microbial metabolism at the molecular
-level is the key, as the metabolic functions shape and influence the
-human host and its health. The discovery of the neuroactive metabolites
-could provide us with biomarkers for early detection or necessary
-medicinally useful molecules that can be applied in intervention.
-
-\hypertarget{material-and-methods}{%
-\subsection{Material and Methods}\label{material-and-methods}}
-
-\hypertarget{study-ethics}{%
-\subsubsection{Study Ethics}\label{study-ethics}}
+\subsubsection{Study Ethics}
 
@@
 
-\hypertarget{participants}{%
-\subsubsection{Participants}\label{participants}}
+\subsubsection{Participants}
 
@@
 
-\hypertarget{cognitive-assessments}{%
-\subsubsection{Cognitive Assessments}\label{cognitive-assessments}}
+\subsubsection{Cognitive Assessments}
 
@@
 
-\hypertarget{stool-sample-collection-and-sequencing}{%
-\subsubsection{Stool Sample Collection and
-Sequencing}\label{stool-sample-collection-and-sequencing}}
+\subsubsection{Stool Sample Collection an
+Sequencing}
 
@@
 
-\hypertarget{machine-learning-for-cognitive-development}{%
-\subsubsection{Machine learning for cognitive
-development}\label{machine-learning-for-cognitive-development}}
+\subsubsection{Machine learning for cognitiv
+development}
 
@@
 Forest experiments}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1797}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2915}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Input set
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Age bracket
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Microbiome encoding type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Demographics
-
-Provided?
-
-(sex, education)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-1
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Not provided
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-3
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-4
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-5
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Not provided
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-7
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-9
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-10
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
 
@@
 
-\hypertarget{mri-segmentation}{%
-\subsubsection{MRI / segmentation}\label{mri-segmentation}}
+\subsubsection{MRI / segmentation}
 
@@
 
-\hypertarget{computational-analysis-of-metagenomes}{%
-\subsubsection{Computational analysis of
-metagenomes}\label{computational-analysis-of-metagenomes}}
+\subsubsection{Computational analysis of metagenomes}
 
@@
 
-\hypertarget{data-and-code-availability}{%
-\subsubsection{Data and code
-availability}\label{data-and-code-availability}}
+\subsubsection{Data and code availability}
 
@@
 
-\hypertarget{research-standards}{%
-\subsection{Research Standards}\label{research-standards}}
+\subsection{Research Standards}
 
-\hypertarget{acknowledgements}{%
-\subsection{Acknowledgements}\label{acknowledgements}}
+\subsection{Acknowledgements}
 
 Research funding: NIH UG3 OD023313 and Wellcome LEAP 1kD.
-
-\hypertarget{references-and-citations}{%
-\subsection{References and Citations}\label{references-and-citations}}
-
-\url{https://doi.org/10.1093/cercor/bhs201}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Amaratunga, D.,
-Cabrera, J., Lee, Y.-S., 2008. Enriched random forests. Bioinformatics
-24, 2010--2014. https://doi.org/10.1093/bioinformatics/btn356}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Aminoff, E.M.,
-Kveraga, K., Bar, M., 2013. The role of the parahippocampal cortex in
-cognition. Trends Cogn. Sci. 17, 379--390.
-https://doi.org/10.1016/j.tics.2013.06.009}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Apps, M.A.J.,
-Rushworth, M.F.S., Chang, S.W.C., 2016. The Anterior Cingulate Gyrus and
-Social Cognition: Tracking the Motivation of Others. Neuron 90,
-692--707. https://doi.org/10.1016/j.neuron.2016.04.018}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ardekani, B.A.,
-Bermudez, E., Mubeen, A.M., Bachman, A.H., Initiative, for the A.D.N.,
-2017. Prediction of Incipient Alzheimer's Disease Dementia in Patients
-with Mild Cognitive Impairment. J. Alzheimers Dis. 55, 269--281.
-https://doi.org/10.3233/JAD-160594}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Avants, B.B.,
-Tustison, N.J., Stauffer, M., Song, G., Wu, B., Gee, J.C., 2014. The
-insight ToolKit image registration framework. Front. Neuroinform. 8, 44.
-https://doi.org/10.3389/fninf.2014.00044}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bäckhed, F., Roswall,
-J., Peng, Y., Feng, Q., Jia, H., Kovatcheva-Datchary, P., Li, Y., Xia,
-Y., Xie, H., Zhong, H., Khan, M.T., Zhang, J., Li, J., Xiao, L.,
-Al-Aama, J., Zhang, D., Lee, Y.S., Kotowska, D., Colding, C., Tremaroli,
-V., Yin, Y., Bergman, S., Xu, X., Madsen, L., Kristiansen, K., Dahlgren,
-J., Wang, J., 2015. Dynamics and stabilization of the human gut
-microbiome during the first year of life. Cell Host Microbe 17,
-690--703. https://doi.org/10.1016/j.chom.2015.04.004}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bairoch, A., 2000. The
-ENZYME database in 2000. Nucleic Acids Res. 28, 304--305.
-https://doi.org/10.1093/nar/28.1.304}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Benítez-Páez, A.,
-Gómez del Pugar, E.M., López-Almela, I., Moya-Pérez, Á., Codoñer-Franch,
-P., Sanz, Y., 2020. Depletion of \emph{Blautia} Species in the
-Microbiota of Obese Children Relates to Intestinal Inflammation and
-Metabolic Phenotype Worsening. mSystems 5, e00857-19.
-https://doi.org/10.1128/mSystems.00857-19}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bezanson, J., Edelman,
-A., Karpinski, S., Shah, V.B., 2017. Julia: A fresh approach to
-numerical computing. Siam Rev. 59, 65--98.
-https://doi.org/10.1137/141000671}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Blackwell, C.K.,
-Mansolf, M., Sherlock, P., Ganiban, J., Hofheimer, J.A., Barone, C.J.,
-Bekelman, T.A., Blair, C., Cella, D., Collazo, S., Croen, L.A., Deoni,
-S., Elliott, A.J., Ferrara, A., Fry, R.C., Gershon, R., Herbstman, J.B.,
-Karagas, M.R., LeWinn, K.Z., Margolis, A., Miller, R.L., O'Shea, T.M.,
-Porucznik, C.A., Wright, R.J., 2022. Youth well-being during the
-COVID-19 pandemic. Pediatrics 149.
-https://doi.org/10.1542/peds.2021-054754}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Blaom, A.D., Kiraly,
-F., Lienart, T., Simillides, Y., Arenas, D., Vollmer, S.J., 2020. MLJ: A
-Julia package for composable machine learning. J. Open Source Softw. 5,
-2704. https://doi.org/10.21105/joss.02704}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Boes, A.D., Tranel,
-D., Anderson, S.W., Nopoulos, P., 2008. Right anterior cingulate cortex
-volume is a neuroanatomical correlate of aggression and defiance in
-boys. Behav. Neurosci. 122, 677--684.
-https://doi.org/10.1037/0735-7044.122.3.677}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bokulich, N.A., Chung,
-J., Battaglia, T., Henderson, N., Jay, M., Li, H., D Lieber, A., Wu, F.,
-Perez-Perez, G.I., Chen, Y., Schweizer, W., Zheng, X., Contreras, M.,
-Dominguez-Bello, M.G., Blaser, M.J., 2016. Antibiotics, birth mode, and
-diet shape microbiome maturation during early life. Sci. Transl. Med. 8,
-343ra82. https://doi.org/10.1126/scitranslmed.aad7121}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Bottino,
-G.Z.M.F., Klepac-Ceraj, V., 2022. ECHO RESONANCE Microbiome.
-https://doi.org/10.17605/OSF.IO/YBS32}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Kayisire,
-A., Luo, A., Klepac-Ceraj, V., 2021. Microbiome.jl and BiobakeryUtils.jl
-- Julia packages for working with microbial community data. J Open
-Source Softw 6, 3876. https://doi.org/10.21105/joss.03876}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Breiman, L., 2001.
-Random Forests. Mach. Learn. 45, 5--32.
-https://doi.org/10.1023/A:1010933404324}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Brieuc, M.S.O.,
-Waters, C.D., Drinan, D.P., Naish, K.A., 2018. A practical introduction
-to Random Forest for genetic association studies in ecology and
-evolution. Mol. Ecol. Resour. 18, 755--766.
-https://doi.org/10.1111/1755-0998.12773}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bruchhage, Muriel,
-Chen, Y., Cataldo, A.G., Müller, H.-G., Weisee, E., Wilson, S., Pietsch,
-M., D'Sa, V., Marquand, A., Madhow, S., Bouchard, K., Cole, J.H.,
-Biondo, F., Elison, J., O'Muirchheartaigh, J., Deoni, C, S., n.d.
-Longitudinal brain and cognitive development of the first 1000 Days:A
-large Multi-Cohort Multi-Scanner study. Presented at the International
-Society of Magnetic Resonance Medicine.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bush, G., Vogt, B.A.,
-Holmes, J., Dale, A.M., Greve, D., Jenike, M.A., Rosen, B.R., 2002.
-Dorsal anterior cingulate cortex: A role in reward-based decision
-making. Proc. Natl. Acad. Sci. 99, 523--528.
-https://doi.org/10.1073/pnas.012470999}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Callaghan, B., 2020.
-Nested sensitive periods: how plasticity across the microbiota-gut-brain
-axis interacts to affect the development of learning and memory. Curr.
-Opin. Behav. Sci. 36, 55--62.
-https://doi.org/10.1016/j.cobeha.2020.07.011}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cao, Z.,
-Ottino-Gonzalez, J., Cupertino, R.B., Schwab, N., Hoke, C., Catherine,
-O., Cousijn, J., Dagher, A., Foxe, J.J., Goudriaan, A.E., Hester, R.,
-Hutchison, K., Li, C.-S.R., London, E.D., Lorenzetti, V., Luijten, M.,
-Martin-Santos, R., Momenan, R., Paulus, M.P., Schmaal, L., Sinha, R.,
-Sjoerds, Z., Solowij, N., Stein, D.J., Stein, E.A., Uhlmann, A., van
-Holst, R.J., Veltman, D.J., Wiers, R.W., Yücel, M., Zhang, S.,
-Jahanshad, N., Thompson, P.M., Conrod, P., Mackey, S., Garavan, H.,
-2021. Mapping cortical and subcortical asymmetries in substance
-dependence: Findings from the ENIGMA Addiction Working Group. Addict.
-Biol. 26, e13010. https://doi.org/10.1111/adb.13010}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cerdó, T., Diéguez,
-E., Campoy, C., 2019. Early nutrition and gut microbiome:
-interrelationship between bacterial metabolism, immune system, brain
-structure, and neurodevelopment. Am. J. Physiol. Endocrinol. Metab. 317,
-E617--E630. https://doi.org/10.1152/ajpendo.00188.2019}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Chen, X., Ishwaran,
-H., 2012. Random forests for genomic data analysis. Genomics 99,
-323--329. https://doi.org/10.1016/j.ygeno.2012.04.003}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cohen Kadosh, K.,
-Krause, B., King, A.J., Near, J., Cohen Kadosh, R., 2015. Linking GABA
-and glutamate levels to cognitive skill acquisition during development.
-Hum. Brain Mapp. 36, 4334--4345. https://doi.org/10.1002/hbm.22921}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cowan, C.S.M., Dinan,
-T.G., Cryan, J.F., 2020. Annual Research Review: Critical windows - the
-microbiota-gut-brain axis in neurocognitive development. J. Child
-Psychol. Psychiatry 61, 353--371. https://doi.org/10.1111/jcpp.13156}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dalile, B., Van
-Oudenhove, L., Vervliet, B., Verbeke, K., 2019. The role of short-chain
-fatty acids in microbiota-gut-brain communication. Nat. Rev.
-Gastroenterol. Hepatol. 16, 461--478.
-https://doi.org/10.1038/s41575-019-0157-3}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Danisch, S.,
-Krumbiegel, J., 2021. Makie.jl: Flexible high-performance data
-visualization for Julia. J. Open Source Softw. 6, 3349.
-https://doi.org/10.21105/joss.03349}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Deoni, S.C.,
-Beauchemin, J., Volpe, A., Dâ Sa, V., RESONANCE Consortium, 2021. Impact
-of the COVID-19 pandemic on early child cognitive development: Initial
-findings in a longitudinal observational study of child health. MedRxiv
-Prepr. Serv. Health Sci. https://doi.org/10.1101/2021.08.10.21261846}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dominguez-Bello, M.G.,
-Costello, E.K., Contreras, M., Magris, M., Hidalgo, G., Fierer, N.,
-Knight, R., 2010. Delivery mode shapes the acquisition and structure of
-the initial microbiota across multiple body habitats in newborns. Proc.
-Natl. Acad. Sci. U. S. A. 107, 11971--11975.
-https://doi.org/10.1073/pnas.1002601107}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dong, H., Wang, B.,
-Feng, J., Yue, X., Jia, F., 2021. Correlation between serum
-concentrations of menaquinone-4 and developmental quotients in children
-with autism spectrum disorder. Front. Nutr. 8, 748513.
-https://doi.org/10.3389/fnut.2021.748513}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ernst, M., Nelson,
-E.E., Jazbec, S., McClure, E.B., Monk, C.S., Leibenluft, E., Blair, J.,
-Pine, D.S., 2005. Amygdala and nucleus accumbens in responses to receipt
-and omission of gains in adults and adolescents. NeuroImage 25,
-1279--1291. https://doi.org/10.1016/j.neuroimage.2004.12.038}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Forrest, C.B.,
-Blackwell, C.K., Camargo, C.A., 2018. Advancing the science of
-children's positive health in the national institutes of health
-environmental influences on child health outcomes (ECHO) research
-program. J. Pediatr. 196, 298--300.
-https://doi.org/10.1016/j.jpeds.2018.02.004}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Franzosa, E.A.,
-Sirota-Madi, A., Avila-Pacheco, J., Fornelos, N., Haiser, H.J., Reinker,
-S., Vatanen, T., Hall, A.B., Mallick, H., McIver, L.J., Sauk, J.S.,
-Wilson, R.G., Stevens, B.W., Scott, J.M., Pierce, K., Deik, A.A.,
-Bullock, K., Imhann, F., Porter, J.A., Zhernakova, A., Fu, J., Weersma,
-R.K., Wijmenga, C., Clish, C.B., Vlamakis, H., Huttenhower, C., Xavier,
-R.J., 2019. Gut microbiome structure and metabolic activity in
-inflammatory bowel disease. Nat. Microbiol. 4, 293--305.
-https://doi.org/10.1038/s41564-018-0306-4}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Fung, T.C., Olson,
-C.A., Hsiao, E.Y., 2017. Interactions between the microbiota, immune and
-nervous systems in health and disease. Nat. Neurosci. 20, 145--155.
-https://doi.org/10.1038/nn.4476}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ghosh, T.S., Rampelli,
-S., Jeffery, I.B., Santoro, A., Neto, M., Capri, M., Giampieri, E.,
-Jennings, A., Candela, M., Turroni, S., Zoetendal, E.G., Hermes, G.D.A.,
-Elodie, C., Meunier, N., Brugere, C.M., Pujos-Guillot, E., Berendsen,
-A.M., Groot, L.C.P.G.M.D., Feskins, E.J.M., Kaluza, J., Pietruszka, B.,
-Bielak, M.J., Comte, B., Maijo-Ferre, M., Nicoletti, C., Vos, W.M.D.,
-Fairweather-Tait, S., Cassidy, A., Brigidi, P., Franceschi, C., O'Toole,
-P.W., 2020. Mediterranean diet intervention alters the gut microbiome in
-older people reducing frailty and improving health status: the NU-AGE
-1-year dietary intervention across five European countries. Gut 69,
-1218--1228. https://doi.org/10.1136/gutjnl-2019-319654}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Gillman, M.W.,
-Blaisdell, C.J., 2018. Environmental influences on child health
-outcomes, a research program of the NIH. Curr. Opin. Pediatr. 30, 260.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Gong, Q.-Y., Cai, L.,
-Jing, Y., Wang, W., Yang, D.-X., Chen, S.-W., Tian, H.-L., 2022.
-Urolithin A alleviates blood-brain barrier disruption and attenuates
-neuronal apoptosis following traumatic brain injury in mice. Neural
-Regen. Res. 17, 2007. https://doi.org/10.4103/1673-5374.335163}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Haam, J., Yakel, J.L.,
-2017. Cholinergic modulation of the hippocampal region and memory
-function. J. Neurochem. 142, 111--121.
-https://doi.org/10.1111/jnc.14052}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Haiser, H.J.,
-Gootenberg, D.B., Chatman, K., Sirasani, G., Balskus, E.P., Turnbaugh,
-P.J., 2013. Predicting and Manipulating Cardiac Drug Inactivation by the
-Human Gut Bacterium Eggerthella lenta. Science 341, 295--298.
-https://doi.org/10.1126/science.1235872}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hartstra, A.V.,
-Schüppel, V., Imangaliyev, S., Schrantee, A., Prodan, A., Collard, D.,
-Levin, E., Dallinga-Thie, G., Ackermans, M.T., Winkelmeijer, M., Havik,
-S.R., Metwaly, A., Lagkouvardos, I., Nier, A., Bergheim, I.,
-Heikenwalder, M., Dunkel, A., Nederveen, A.J., Liebisch, G., Mancano,
-G., Claus, S.P., Benítez-Páez, A., la Fleur, S.E., Bergman, J.J.,
-Gerdes, V., Sanz, Y., Booij, J., Kemper, E., Groen, A.K., Serlie, M.J.,
-Haller, D., Nieuwdorp, M., 2020. Infusion of donor feces affects the
-gut--brain axis in humans with metabolic syndrome. Mol. Metab. 42,
-101076. https://doi.org/10.1016/j.molmet.2020.101076}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoogman, M., Bralten,
-J., Hibar, D.P., Mennes, M., Zwiers, M.P., Schweren, L., van Hulzen,
-K.J.E., Medland, S.E., Shumskaya, E., Jahanshad, N., de Zeeuw, P.,
-Szekely, E., Sudre, G., Wolfers, T., Onnink, A.M.H., Dammers, J.T.,
-Mostert, J.C., Vives-Gilabert, Y., Kohls, G., Oberwelland, E., Seitz,
-J., Schulte-Rüther, M., di Bruttopilo, S.A., Doyle, A.E., Høvik, M.F.,
-Dramsdahl, M., Tamm, L., van Erp, T.G.M., Dale, A., Schork, A.,
-Conzelmann, A., Zierhut, K., Baur, R., McCarthy, H., Yoncheva, Y.N.,
-Cubillo, A., Chantiluke, K., Mehta, M.A., Paloyelis, Y., Hohmann, S.,
-Baumeister, S., Bramati, I., Mattos, P., Tovar-Moll, F., Douglas, P.,
-Banaschewski, T., Brandeis, D., Kuntsi, J., Asherson, P., Rubia, K.,
-Kelly, C., Di Martino, A., Milham, M.P., Castellanos, F.X., Frodl, T.,
-Zentis, M., Lesch, K.-P., Reif, A., Pauli, P., Jernigan, T., Haavik, J.,
-Plessen, K.J., Lundervold, A.J., Hugdahl, K., Seidman, L.J., Biederman,
-J., Rommelse, N., Heslenfeld, D.J., Hartman, C., Hoekstra, P.J.,
-Oosterlaan, J., von Polier, G., Konrad, K., Vilarroya, O.,
-Ramos-Quiroga, J.-A., Soliva, J.C., Durston, S., Buitelaar, J.K.,
-Faraone, S.V., Shaw, P., Thompson, P., Franke, B., 2017. Subcortical
-brain volume differences of participants with ADHD across the lifespan:
-an ENIGMA collaboration. Lancet Psychiatry 4, 310--319.
-https://doi.org/10.1016/S2215-0366(17)30049-4}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoshino, Y., Yamamoto,
-T., Kaneko, M., Tachibana, R., Watanabe, M., Ono, Y., Kumashiro, H.,
-1984. Blood Serotonin and Free Tryptophan Concentration in Autistic
-Children. Neuropsychobiology 11, 22--27.
-https://doi.org/10.1159/000118045}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hosomi, K., Saito, M.,
-Park, J., Murakami, H., Shibata, N., Ando, M., Nagatake, T., Konishi,
-K., Ohno, H., Tanisawa, K., Mohsen, A., Chen, Y.-A., Kawashima, H.,
-Natsume-Kitatani, Y., Oka, Y., Shimizu, H., Furuta, M., Tojima, Y.,
-Sawane, K., Saika, A., Kondo, S., Yonejima, Y., Takeyama, H., Matsutani,
-A., Mizuguchi, K., Miyachi, M., Kunisawa, J., 2022. Oral administration
-of Blautia wexlerae ameliorates obesity and type 2 diabetes via
-metabolic remodeling of the gut microbiota. Nat. Commun. 13, 4477.
-https://doi.org/10.1038/s41467-022-32015-7}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Jenkinson, M.,
-Beckmann, C.F., Behrens, T.E.J., Woolrich, M.W., Smith, S.M., 2012. FSL.
-Neuroimage 62, 782--790.
-https://doi.org/10.1016/j.neuroimage.2011.09.015}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Jiang, H., Ling, Z.,
-Zhang, Y., Mao, H., Ma, Z., Yin, Y., Wang, W., Tang, W., Tan, Z., Shi,
-J., Li, L., Ruan, B., 2015. Altered fecal microbiota composition in
-patients with major depressive disorder. Brain. Behav. Immun. 48,
-186--194. https://doi.org/10.1016/j.bbi.2015.03.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kanehisa, M., Goto,
-S., Kawashima, S., Okuno, Y., Hattori, M., 2004. The KEGG resource for
-deciphering the genome. Nucleic Acids Res. 32, D277--D280.
-https://doi.org/10.1093/nar/gkh063}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kant, R., Rasinkangas,
-P., Satokari, R., Pietilä, T.E., Palva, A., 2015. Genome Sequence of the
-Butyrate-Producing Anaerobic Bacterium Anaerostipes hadrus PEL 85.
-Genome Announc. 3, e00224-15. https://doi.org/10.1128/genomeA.00224-15}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kim, C.-S., Cha, L.,
-Sim, M., Jung, S., Chun, W.Y., Baik, H.W., Shin, D.-M., 2021. Probiotic
-supplementation improves cognitive function and mood with changes in gut
-microbiota in Community-Dwelling older adults: A randomized,
-Double-Blind, Placebo-Controlled, multicenter trial. J Gerontol Biol Sci
-Med Sci 76, 32--40. https://doi.org/10.1093/gerona/glaa090}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Knudsen, E.I., 2004.
-Sensitive periods in the development of the brain and behavior. J. Cogn.
-Neurosci. 16, 1412--1425. https://doi.org/10.1162/0898929042304796}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Koenig, J.E., Spor,
-A., Scalfone, N., Fricker, A.D., Stombaugh, J., Knight, R., Angenent,
-L.T., Ley, R.E., 2011. Succession of microbial consortia in the
-developing infant gut microbiome. Proc. Natl. Acad. Sci. U. S. A. 108
-Suppl 1, 4578--4585. https://doi.org/10.1073/pnas.1000081107}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Coker,
-M.O., Madan, J.C., 2022. The developing microbiome from birth to 3
-years: The Gut-Brain axis and neurodevelopmental outcomes. Front.
-Pediatr. 10, 815885. https://doi.org/10.3389/fped.2022.815885}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Korrick,
-S.A., Baker, E.R., Karagas, M.R., Madan, J.C., 2020. Prospective
-associations of the infant gut microbiome and microbial function with
-social behaviors related to autism at age 3 years. Sci. Rep. 10, 15515.
-https://doi.org/10.1038/s41598-020-72386-9}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ley, R.E., Turnbaugh,
-P.J., Klein, S., Gordon, J.I., 2006. Human gut microbes associated with
-obesity. Nature 444, 1022--1023. https://doi.org/10.1038/4441022a}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Li, Y.-M., Yu, Q.,
-Huang, L., Fu, J., Feng, R.-J., 2022. {[}Correlations Between Gut
-Microbiota Changes and Cognitive Function in Patients with Post-Stroke
-Cognitive Impairment in the Early Stage{]}. Sichuan Da Xue Xue Bao Yi
-Xue Ban 53, 857--865. https://doi.org/10.12182/20220960105}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, S., Li, E., Sun,
-Z., Fu, D., Duan, G., Jiang, M., Yu, Y., Mei, L., Yang, P., Tang, Y.,
-Zheng, P., 2019. Altered gut microbiota and short chain fatty acids in
-Chinese children with autism spectrum disorder. Sci. Rep. 9, 287.
-https://doi.org/10.1038/s41598-018-36430-z}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, X., Mao, B., Gu,
-J., Wu, J., Cui, S., Wang, G., Zhao, J., Zhang, H., Chen, W., 2021.
-\emph{Blautia} ---a new functional genus with potential probiotic
-properties? Gut Microbes 13, 1875796.
-https://doi.org/10.1080/19490976.2021.1875796}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Lopez-Siles, M.,
-Duncan, S.H., Garcia-Gil, L.J., Martinez-Medina, M., 2017.
-Faecalibacterium prausnitzii: from microbiology to diagnostics and
-prognostics. ISME J. 11, 841--852.
-https://doi.org/10.1038/ismej.2016.176}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Louwies, T., Johnson,
-A.C., Orock, A., Yuan, T., Greenwood-Van Meerveld, B., 2020. The
-microbiota-gut-brain axis: An emerging role for the epigenome. Exp.
-Biol. Med. 245, 138--145. https://doi.org/10.1177/1535370219891690}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Magnusson, K.R.,
-Hauck, L., Jeffrey, B.M., Elias, V., Humphrey, A., Nath, R., Perrone,
-A., Bermudez, L.E., 2015. Relationships between diet-related changes in
-the gut microbiome and cognitive flexibility. Neuroscience 300,
-128--140. https://doi.org/10.1016/j.neuroscience.2015.05.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mallick, H.,
-Rahnavard, A., McIver, L.J., Ma, S., Zhang, Y., Nguyen, L.H., Tickle,
-T.L., Weingart, G., Ren, B., Schwager, E.H., Chatterjee, S., Thompson,
-K.N., Wilkinson, J.E., Subramanian, A., Lu, Y., Waldron, L., Paulson,
-J.N., Franzosa, E.A., Bravo, H.C., Huttenhower, C., 2021. Multivariable
-association discovery in population-scale meta-omics studies.
-https://doi.org/10.1101/2021.01.20.427420}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Maruo, T., Sakamoto,
-M., Ito, C., Toda, T., Benno, Y., 2008. Adlercreutzia equolifaciens gen.
-nov., sp. nov., an equol-producing bacterium isolated from human faeces,
-and emended description of the genus Eggerthella. Int. J. Syst. Evol.
-Microbiol. 58, 1221--1227. https://doi.org/10.1099/ijs.0.65404-0}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mayneris-Perxachs, J.,
-Castells-Nobau, A., Arnoriaga-Rodríguez, M., Martin, M., de la
-Vega-Correa, L., Zapata, C., Burokas, A., Blasco, G., Coll, C.,
-Escrichs, A., Biarnés, C., Moreno-Navarrete, J.M., Puig, J., Garre-Olmo,
-J., Ramos, R., Pedraza, S., Brugada, R., Vilanova, J.C., Serena, J.,
-Gich, J., Ramió-Torrentà, L., Pérez-Brocal, V., Moya, A., Pamplona, R.,
-Sol, J., Jové, M., Ricart, W., Portero-Otin, M., Deco, G., Maldonado,
-R., Fernández-Real, J.M., 2022. Microbiota alterations in proline
-metabolism impact depression. Cell Metab. 34, 681-701.e10.
-https://doi.org/10.1016/j.cmet.2022.04.001}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mukherjee, A., Lordan,
-C., Ross, R.P., Cotter, P.D., 2020. Gut microbes from the
-phylogenetically diverse genus Eubacterium and their various
-contributions to gut health. Gut Microbes 12, 1802866.
-https://doi.org/10.1080/19490976.2020.1802866}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mullen, E.M., others,
-1995. Mullen scales of early learning. AGS Circle Pines, MN.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Needham, B.D.,
-Funabashi, M., Adame, M.D., Wang, Z., Boktor, J.C., Haney, J., Wu,
-W.-L., Rabut, C., Ladinsky, M.S., Hwang, S.-J., Guo, Y., Zhu, Q.,
-Griffiths, J.A., Knight, R., Bjorkman, P.J., Shapiro, M.G., Geschwind,
-D.H., Holschneider, D.P., Fischbach, M.A., Mazmanian, S.K., 2022. A
-gut-derived metabolite alters brain activity and anxiety behaviour in
-mice. Nature 602, 647--653. https://doi.org/10.1038/s41586-022-04396-8}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Opel, N., Thalamuthu,
-A., Milaneschi, Y., Grotegerd, D., Flint, C., Leenings, R., Goltermann,
-J., Richter, M., Hahn, T., Woditsch, G., Berger, K., Hermesdorf, M.,
-McIntosh, A., Whalley, H.C., Harris, M.A., MacMaster, F.P., Walter, H.,
-Veer, I.M., Frodl, T., Carballedo, A., Krug, A., Nenadic, I., Kircher,
-T., Aleman, A., Groenewold, N.A., Stein, D.J., Soares, J.C.,
-Zunta-Soares, G.B., Mwangi, B., Wu, M.-J., Walter, M., Li, M., Harrison,
-B.J., Davey, C.G., Cullen, K.R., Klimes-Dougan, B., Mueller, B.A.,
-Sämann, P.G., Penninx, B., Nawijn, L., Veltman, D.J., Aftanas, L., Brak,
-I.V., Filimonova, E.A., Osipov, E.A., Reneman, L., Schrantee, A., Grabe,
-H.J., Van der Auwera, S., Wittfeld, K., Hosten, N., Völzke, H., Sim, K.,
-Gotlib, I.H., Sacchet, M.D., Lagopoulos, J., Hatton, S.N., Hickie, I.,
-Pozzi, E., Thompson, P.M., Jahanshad, N., Schmaal, L., Baune, B.T.,
-Dannlowski, U., 2021. Brain structural abnormalities in obesity:
-relation to age, genetic risk, and common psychiatric disorders. Mol.
-Psychiatry 26, 4839--4852. https://doi.org/10.1038/s41380-020-0774-9}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Pagonabarraga, J.,
-Corcuera-Solano, I., Vives-Gilabert, Y., Llebaria, G., García-Sánchez,
-C., Pascual-Sedano, B., Delfino, M., Kulisevsky, J., Gómez-Ansón, B.,
-2013. Pattern of Regional Cortical Thinning Associated with Cognitive
-Deterioration in Parkinson's Disease. PLoS ONE 8, e54980.
-https://doi.org/10.1371/journal.pone.0054980}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Palomo-Buitrago, M.E.,
-Sabater-Masdeu, M., Moreno-Navarrete, J.M., Caballano-Infantes, E.,
-Arnoriaga-Rodríguez, M., Coll, C., Ramió, L., Palomino-Schätzlein, M.,
-Gutiérrez-Carcedo, P., Pérez-Brocal, V., Simó, R., Moya, A., Ricart, W.,
-Herance, J.R., Fernández-Real, J.M., 2019. Glutamate interactions with
-obesity, insulin resistance, cognition and gut microbiota composition.
-Acta Diabetol. 56, 569--579. https://doi.org/10.1007/s00592-019-01313-w}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Pronovost, G.N.,
-Hsiao, E.Y., 2019. Perinatal interactions between the microbiome,
-immunity, and neurodevelopment. Immunity 50, 18--36.
-https://doi.org/10.1016/j.immuni.2018.11.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sadeghi, B.,
-Chiarawongse, P., Squire, K., Jones, D.C., Noack, A., St-Jean, C.,
-Huijzer, R., Schätzle, R., Butterworth, I., Peng, Y.-F., Blaom, A.,
-2022. DecisionTree.jl - A julia implementation of the CART decision tree
-and random forest algorithms. https://doi.org/10.5281/zenodo.7359268}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sela, D.A., Chapman,
-J., Adeuya, A., Kim, J.H., Chen, F., Whitehead, T.R., Lapidus, A.,
-Rokhsar, D.S., Lebrilla, C.B., German, J.B., Price, N.P., Richardson,
-P.M., Mills, D.A., 2008. The genome sequence of \emph{Bifidobacterium
-longum} subsp. \emph{infantis} reveals adaptations for milk utilization
-within the infant microbiome. Proc. Natl. Acad. Sci. 105, 18964--18969.
-https://doi.org/10.1073/pnas.0809584105}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Selma, M.V., Beltrán,
-D., García-Villalba, R., Espín, J.C., Tomás-Barberán, F.A., 2014.
-Description of urolithin production capacity from ellagic acid of two
-human intestinal Gordonibacter species. Food Funct. 5, 1779--1784.
-https://doi.org/10.1039/c4fo00092g}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Setchell, K.D.,
-Clerici, C., Lephart, E.D., Cole, S.J., Heenan, C., Castellani, D.,
-Wolfe, B.E., Nechemias-Zimmer, L., Brown, N.M., Lund, T.D., Handa, R.J.,
-Heubi, J.E., 2005. S-Equol, a potent ligand for estrogen receptor β, is
-the exclusive enantiomeric form of the soy isoflavone metabolite
-produced by human intestinal bacterial flora. Am. J. Clin. Nutr. 81,
-1072--1079. https://doi.org/10.1093/ajcn/81.5.1072}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sharon, G., Sampson,
-T.R., Geschwind, D.H., Mazmanian, S.K., 2016. The central nervous system
-and the gut microbiome. Cell 167, 915--932.
-https://doi.org/10.1016/j.cell.2016.10.027}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Silbereis, J.C.,
-Pochareddy, S., Zhu, Y., Li, M., Sestan, N., 2016. The cellular and
-molecular landscapes of the developing human central nervous system.
-Neuron 89, 248--268.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sordillo, J.E.,
-Korrick, S., Laranjo, N., Carey, V., Weinstock, G.M., Gold, D.R.,
-O'Connor, G., Sandel, M., Bacharier, L.B., Beigelman, A., Zeiger, R.,
-Litonjua, A.A., Weiss, S.T., 2019. Association of the infant gut
-microbiome with early childhood neurodevelopmental outcomes: An
-ancillary study to the VDAART randomized clinical trial. JAMA Netw. Open
-2, e190905. https://doi.org/10.1001/jamanetworkopen.2019.0905}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Spichak, S.,
-Bastiaanssen, T.F.S., Berding, K., Vlckova, K., Clarke, G., Dinan, T.G.,
-Cryan, J.F., 2021. Mining microbes for mental health: Determining the
-role of microbial metabolic pathways in human brain health and disease.
-Neurosci. Biobehav. Rev. 125, 698--761.
-https://doi.org/10.1016/j.neubiorev.2021.02.044}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Stephan, J., Stegle,
-O., Beyer, A., 2015. A random forest approach to capture genetic effects
-in the presence of population structure. Nat. Commun. 6, 7432.
-https://doi.org/10.1038/ncomms8432}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Suzek, B.E., Huang,
-H., McGarvey, P., Mazumder, R., Wu, C.H., 2007. UniRef: comprehensive
-and non-redundant UniProt reference clusters. Bioinformatics 23,
-1282--1288. https://doi.org/10.1093/bioinformatics/btm098}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Takahashi, H., Yang,
-J., Yamamoto, H., Fukuda, S., Arakawa, K., 2021. Complete Genome
-Sequence of Adlercreutzia equolifaciens subsp. celatus DSM 18785.
-Microbiol. Resour. Announc. 10, e00354-21.
-https://doi.org/10.1128/MRA.00354-21}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tau, G.Z., Peterson,
-B.S., 2010. Normal development of brain circuits. Neuropsychopharmacol.
-Off. Publ. Am. Coll. Neuropsychopharmacol. 35, 147--168.
-https://doi.org/10.1038/npp.2009.115}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Thawornkuno, C.,
-Tanaka, M., Sone, T., Asano, K., 2009. Biotransformation of Daidzein to
-Equol by Crude Enzyme from Asaccharobacter celatus AHU1763 Required an
-Anaerobic Environment. Biosci. Biotechnol. Biochem. 73, 1435--1438.
-https://doi.org/10.1271/bbb.80908}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tognini, P., 2017. Gut
-microbiota: a potential regulator of neurodevelopment. Front. Cell.
-Neurosci. 11, 25. https://doi.org/10.3389/fncel.2017.00025}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tso, L., Bonham, K.S.,
-Fishbein, A., Rowland, S., Klepac-Ceraj, V., 2021. Targeted
-High-Resolution taxonomic identification of bifidobacterium longum
-subsp. infantis using human milk oligosaccharide metabolizing genes.
-Nutrients 13, 2833. https://doi.org/10.3390/nu13082833}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Valles-Colomer, M.,
-Falony, G., Darzi, Y., Tigchelaar, E.F., Wang, J., Tito, R.Y., Schiweck,
-C., Kurilshikov, A., Joossens, M., Wijmenga, C., Claes, S., Van
-Oudenhove, L., Zhernakova, A., Vieira-Silva, S., Raes, J., 2019. The
-neuroactive potential of the human gut microbiota in quality of life and
-depression. Nat. Microbiol. 4, 623--632.
-https://doi.org/10.1038/s41564-018-0337-x}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Velazquez, M., Lee,
-Y., Initiative, for the A.D.N., 2021. Random forest model for
-feature-based Alzheimer's disease conversion prediction from early mild
-cognitive impairment subjects. PLOS ONE 16, e0244773.
-https://doi.org/10.1371/journal.pone.0244773}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wacker, J., Dillon,
-D.G., Pizzagalli, D.A., 2009. The role of the nucleus accumbens and
-rostral anterior cingulate cortex in anhedonia: Integration of resting
-EEG, fMRI, and volumetric techniques. NeuroImage 46, 327--337.
-https://doi.org/10.1016/j.neuroimage.2009.01.058}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wan, Y., Zuo, T., Xu,
-Z., Zhang, F., Zhan, H., Chan, D., Leung, T.-F., Yeoh, Y.K., Chan,
-F.K.L., Chan, R., Ng, S.C., 2021. Underdevelopment of the gut microbiota
-and bacteria species as non-invasive markers of prediction in children
-with autism spectrum disorder. Gut gutjnl-2020-324015.
-https://doi.org/10.1136/gutjnl-2020-324015}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, L.-J., Yang,
-C.-Y., Chou, W.-J., Lee, M.-J., Chou, M.-C., Kuo, H.-C., Yeh, Y.-M.,
-Lee, S.-Y., Huang, L.-H., Li, S.-C., 2020. Gut microbiota and dietary
-patterns in children with attention-deficit/hyperactivity disorder. Eur.
-Child Adolesc. Psychiatry 29, 287--297.
-https://doi.org/10.1007/s00787-019-01352-2}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, X.-L., Hur,
-H.-G., Lee, J.H., Kim, K.T., Kim, S.-I., 2005. Enantioselective
-Synthesis of S-Equol from Dihydrodaidzein by a Newly Isolated Anaerobic
-Human Intestinal Bacterium. Appl. Environ. Microbiol. 71, 214--219.
-https://doi.org/10.1128/AEM.71.1.214-219.2005}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 2012.
-Wechsler preschool and primary scale of intelligence---fourth edition.
-Psychol. Corp. San Antonio TX.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 1949.
-Wechsler intelligence scale for children.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, L., Yan, J.,
-Yang, T., Zhu, J., Li, T., Wei, H., Chen, J., 2021. Fecal Microbiome
-Transplantation from Children with Autism Spectrum Disorder Modulates
-Tryptophan and Serotonergic Synapse Metabolism and Induces Altered
-Behaviors in Germ-Free Mice. mSystems 6, e01343-20.
-https://doi.org/10.1128/mSystems.01343-20}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, X., Yang, Y.,
-Zhang, Y., Zhang, X.-M., Zhao, Z.-Q., Zhang, Y.-Q., 2013. Estrogen in
-the Anterior Cingulate Cortex Contributes to Pain-Related Aversion.
-Cereb. Cortex 23, 2190--2203. https://doi.org/10.1093/cercor/bhs201}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Yael, D., Tahary, O.,
-Gurovich, B., Belelovsky, K., Bar-Gad, I., 2019. Disinhibition of the
-Nucleus Accumbens Leads to Macro-Scale Hyperactivity Consisting of
-Micro-Scale Behavioral Segments Encoded by Striatal Activity. J.
-Neurosci. 39, 5897--5909.
-https://doi.org/10.1523/JNEUROSCI.3120-18.2019}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Yau, W.-Y.W., Zubieta,
-J.-K., Weiland, B.J., Samudra, P.G., Zucker, R.A., Heitzeg, M.M., 2012.
-Nucleus Accumbens Response to Incentive Stimuli Anticipation in Children
-of Alcoholics: Relationships with Precursive Behavioral Risk and
-Lifetime Alcohol Use. J. Neurosci. 32, 2544--2551.
-https://doi.org/10.1523/JNEUROSCI.1390-11.2012}
-
-Supplementary Tables
-
-\textbf{Supplementary Table XXA. List of most important taxa for Random
-Forest models on the 0 to 6 months age bracket, microbiome +
-demographics, ordered by relative weighted importance rank}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Taxon
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Rank
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Relative weighted importance
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Cumulative
-
-relative weighted importance
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\textbf{Supplementary Table XXA. List of most important taxa for Random
-Forest models on the 18 to 120 months age bracket, microbiome +
-demographics, ordered by relative weighted importance rank}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Taxon
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Rank
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Relative weighted importance
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Cumulative
-
-relative weighted importance
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\textbf{Supplementary Table XX. Benchmark metrics for each brain segment
-prediction experiment. Confidence intervals are calculated from the
-distribution of metrics from repeated CV at a confidence level of 95\%}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3933}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2440}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1846}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1782}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Brain segment
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Root-mean-square Error
-
-(RMSE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Mean Absolute Proportional Error (MAPE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Correlation coefficient (R)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Brain-stem
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0876
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.605
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-I-V
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0321
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.085
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.156
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-VI-VII
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0132
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.081
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.006
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-VIII-X
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0217
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.078
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-CSF
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0193
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.162
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.071
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-accumbens-area
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0020
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.094
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.288
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-accumbens-area
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0020
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.102
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.041
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-amygdala
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0037
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.288
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-amygdala
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.070
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.179
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-anterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0375
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.069
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.092
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-anterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0264
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.074
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.245
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-basal-forebrain
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0008
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-basal-forebrain
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0010
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.117
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.123
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-caudate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0257
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.099
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.029
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-caudate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0221
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.010
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-cerebellum-white-matter
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0689
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.413
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-cerebellum-white-matter
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0694
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.070
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.434
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-cuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0274
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.234
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-cuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0264
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.075
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.065
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-entorhinal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0107
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.080
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.190
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-entorhinal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0105
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.099
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-fusiform
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0510
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.333
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-fusiform
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0334
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.053
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.248
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-hippocampus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0128
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.054
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.075
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-hippocampus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0150
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.094
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-insula
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0294
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.119
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-insula
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0303
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-isthmus-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0161
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.041
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-isthmus-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0166
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.085
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.074
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-lateral-occipital
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0661
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.062
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.224
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-lateral-occipital
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0716
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.054
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.221
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-lingual
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0410
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.069
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.421
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-lingual
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0440
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.072
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.434
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-middle-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0875
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.052
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.000
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-middle-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0636
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.039
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-orbitofrontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0715
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.055
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.186
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-orbitofrontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0814
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.133
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pallidum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0075
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.064
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.297
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pallidum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0081
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.264
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-paracentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0335
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.078
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.131
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-paracentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0364
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.084
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-parahippocampal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0098
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.058
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-parahippocampal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0092
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.060
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-parietal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0755
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.177
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-parietal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1099
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.045
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.147
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-opercularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0197
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.146
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-opercularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0261
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.060
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.030
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-orbitalis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0157
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.003
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-orbitalis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0233
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.112
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.024
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-triangularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0226
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.174
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-triangularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0322
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.076
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.043
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pericalcarine
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0178
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.111
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.200
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pericalcarine
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0185
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.089
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.273
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-postcentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0495
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.052
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.168
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-postcentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0552
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.135
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-posterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0460
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.137
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-posterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0480
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.139
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-precentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0830
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.065
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.248
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-precentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0594
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.046
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.066
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-precuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0489
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.144
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-precuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0622
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.150
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-putamen
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0243
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-putamen
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0242
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.066
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.011
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-superior-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0867
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.040
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.101
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-superior-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.044
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.096
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-supramarginal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0767
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-supramarginal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0734
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.001
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-temporal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1530
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.040
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.128
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-temporal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1503
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.039
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-thalamus-proper
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0347
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.053
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.170
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-thalamus-proper
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0304
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.050
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.162
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\end{document}

### 2023-02-17 manuscript/tex/sections/results.tex
clean up

--- 2023-02-16 manuscript/tex/converted.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-]{article}
-\usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\usepackage{longtable,booktabs,array}
-\usepackage{multirow}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\usepackage[normalem]{ulem}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-
-\author{}
-\date{}
-
-\begin{document}
-
-\hypertarget{title}{%
-\subsection{Title}\label{title}}
-
-\begin{itemize}
-\item
-  \begin{quote}
-  Gut microbes and their genes are associated with cognition and
-  neuroanatomy in children
-  \end{quote}
-\item
-  \begin{quote}
-  Gut microbes and their genes are associated with cognition performance
-  and brain structure in children
-  \end{quote}
-\item
-  \begin{quote}
-  Extensive relationships between the gut microbiota and normal
-  neurocognitive development
-  \end{quote}
-\item
-  \begin{quote}
-  Extensive relationships between the gut microbiota and typical
-  neurocognitive development
-  \end{quote}
-\item
-  \begin{quote}
-  The gut microbial metagenome is connected to neurocognition in typical
-  child development
-  \end{quote}
-\item
-  \begin{quote}
-  The microbial potential in regulating early childhood neurocognitive
-  development
-  \end{quote}
-\end{itemize}
-
-\hypertarget{authors}{%
-\subsubsection{Authors}\label{authors}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \begin{quote}
-  Bonham 0000-0003-3200-7533
-  \end{quote}
-\item
-  \begin{quote}
-  Bottino 0000-0003-1953-1576
-  \end{quote}
-\item
-  \begin{quote}
-  McCann 0000-0002-9753-7968
-  \end{quote}
-\item
-  \begin{quote}
-  Beauchemin \{ORCID??\}
-  \end{quote}
-\item
-  \begin{quote}
-  Elizabeth Weisse \textless elizabeth.weisse2@gmail.com\textgreater{}
-  \end{quote}
-\item
-  \begin{quote}
-  \ldots{}
-  \end{quote}
-\item
-  \begin{quote}
-  Jonathan O\textquotesingle Muircheartaigh
-  \textless{}\href{mailto:jonathanom@kcl.ac.uk}{\uline{jonathanom@kcl.ac.uk}}\textgreater{}
-  \end{quote}
-\item
-  \begin{quote}
-  Huttenhower 0000-0002-1110-0096
-  \end{quote}
-\item
-  \begin{quote}
-  Muriel Bruchhage
-  \textless{}\href{mailto:mmk.bruchhage@gmail.com}{\uline{mmk.bruchhage@gmail.com}}\textgreater{}
-  0000-0001-9637-0951
-  \end{quote}
-\item
-  \begin{quote}
-  D'Sa 0000-0002-3476-8419
-  \end{quote}
-\item
-  \begin{quote}
-  Deoni 0000-0002-0633-6328
-  \end{quote}
-\item
-  \begin{quote}
-  Klepac-Ceraj 0000-0001-5387-5706
-  \end{quote}
-\end{enumerate}
-
-\hypertarget{abstract}{%
-\subsection{Abstract}\label{abstract}}
-
-The gastrointestinal tract, its resident microorganisms and the central
-nervous system are connected by biochemical signaling, also known as
-"microbiome-gut-brain-axis." Both the human brain and the gut microbiome
-have critical developmental windows in the first three years of life,
-raising the possibility that their development is co-occurring and
-likely co-dependent. Emerging evidence implicates gut microorganisms and
-microbiota composition in cognitive outcomes and neurodevelopmental
-disorders (e.g., autism and anxiety), but the influence of gut microbial
-metabolism on typical neurodevelopment has not been explored in detail.
-We investigated the relationship of the microbiome with the neuroanatomy
-and cognitive function of 361 healthy children, demonstrating that
-differences in gut microbial taxa and gene functions are associated with
-overall cognitive function as well as with differences in the size of
-multiple brain regions. Using a combination of multivariate linear
-models and machine learning (ML) models, we showed that many species,
-including \emph{Gordonibacter pamelae} and \emph{Blautia wexlerae}, were
-significantly associated with higher cognitive function, while some
-species such as \emph{Ruminococcus gnavus} were more commonly found in
-children with low cognitive scores regardless of age or maternal
-education. Microbial genes for enzymes involved in the metabolism of
-neuroactive compounds, particularly short-chain fatty acids such as
-glutamate and propionate, were also associated with cognitive function.
-In addition, ML models were able to use microbial taxa to predict the
-volume of brain regions, with particular taxa often dominating the
-feature importance metric for individual brain regions, such as \emph{B.
-wexlerae} being the most dominant for the parahippocampal region or
-GABA-producing \emph{Bacteroidetes ovatus} for left accumbens. These
-findings provide potential biomarkers of neurocognition and may lead to
-the future development of targets for early detection and early
-intervention.
-
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
-
-The gut and the brain are intimately linked. Signals from the brain
-reach the gut through the autonomic nervous system and the endocrine
-system, and the gut can communicate with the brain through the vagus
-nerve and through endocrine and immune (cytokine) signaling molecules
-\href{https://www.zotero.org/google-docs/?i3SRPn}{(Cerdó et al., 2019;
-Pronovost and Hsiao, 2019; Sharon et al., 2016; Tognini, 2017)}. In
-addition, the products of microbial metabolism generated in the gut can
-influence the brain, both indirectly by stimulating the enteric nervous
-and immune systems, as well as directly through molecules that enter
-circulation and cross the blood-brain barrier. Causal links between the
-gut microbiome and neural development, particularly atypical
-development, are increasingly being identified
-\href{https://www.zotero.org/google-docs/?UsrFKB}{(Spichak et al.,
-2021)}. Both human epidemiology and animal models point to the effects
-of gut microbes on the development of autism spectrum disorder
-\href{https://www.zotero.org/google-docs/?Y8l5A0}{(Laue et al., 2020;
-Wan et al., 2021)} and particular microbial taxa have been associated
-with depression
-\href{https://www.zotero.org/google-docs/?mbaFE3}{(Mayneris-Perxachs et
-al., 2022; Valles-Colomer et al., 2019)} and Alzheimer's disease
-\href{https://www.zotero.org/google-docs/?0BEGUn}{(Fung et al., 2017;
-Kim et al., 2021)}. But information about this ``microbiome-gut-brain
-axis'' in normal neurocognitive development remains lacking,
-particularly early in life.
-
-The first years of life are critical developmental windows for both the
-microbiome and the brain
-\href{https://www.zotero.org/google-docs/?XTqhlm}{(Laue et al., 2022)}.
-Fetal development \emph{in utero} is believed to be mostly sterile, but
-is rapidly seeded at birth through contact with the birth canal (if
-birthed vaginally), caregivers, food sources (breastmilk or formula),
-and other environmental sources such as antibiotics
-\href{https://www.zotero.org/google-docs/?X6N0f4}{(Bäckhed et al., 2015;
-Bokulich et al., 2016; Dominguez-Bello et al., 2010; Louwies et al.,
-2020)}. The early microbiome is characterized by low microbial
-diversity, rapid succession and evolution, and domination by
-Actinobacteria, particularly the genus \emph{Bifidobacterium},
-Bacteroidetes, especially \emph{Bacteroides}, and Proteobacteria
-\href{https://www.zotero.org/google-docs/?VG43uh}{(Koenig et al.,
-2011)}. Many of these microbes have specialized metabolic capabilities
-for digesting human breast milk, such as \emph{Bifidobacterium infantis}
-and \emph{Bacteroides fragilis}
-\href{https://www.zotero.org/google-docs/?1dKzKj}{(Sela et al., 2008;
-Tso et al., 2021)}. Upon the introduction of solid foods, the gut
-microbiome undergoes another categorical transformation, with most taxa
-of the infant microbiome being replaced by taxa more reminiscent of
-adult microbiomes
-\href{https://www.zotero.org/google-docs/?V3lY3p}{(Bäckhed et al.,
-2015)}. Many prior studies have focused on either infant microbiomes or
-adult microbiomes, since performing statistical analyses across this
-transition poses particular challenges. Nevertheless, since this
-transition coincides with critical neural developmental windows and
-neural synaptogenesis, investigation across this solid-food boundary is
-incredibly important
-\href{https://www.zotero.org/google-docs/?d29bv8}{(Tau and Peterson,
-2010)}.
-
-A child's brain undergoes remarkable anatomical, microstructural,
-organizational, and functional changes in the first years of life. By
-age 5, a child's brain has reached \textgreater85\% of its adult size,
-has achieved near-adult levels of myelination, and the pattern of axonal
-connections has been established
-\href{https://www.zotero.org/google-docs/?1jUZOO}{(Silbereis et al.,
-2016)}. Much of this development occurs in discrete windows or
-``sensitive-'' or ``critical periods'' (CPs)
-\href{https://www.zotero.org/google-docs/?w4Pcze}{(Knudsen, 2004)} when
-neural plasticity is particularly high, and particular modes of learning
-and skill development are preferred. The timing of these sensitive
-periods is driven in part by genetics, but can also be affected by the
-environment, including the gut microbiome
-\href{https://www.zotero.org/google-docs/?KTm9lQ}{(Cowan et al., 2020)}.
-In fact, emerging evidence suggests that the timing and duration of CPs
-may be driven in part by cues from the developing gut microbiome
-\href{https://www.zotero.org/google-docs/?mAD1ne}{(Callaghan, 2020)}. As
-such, understanding the normal spectrum of healthy microbiome
-development and how it relates to normal neurocognitive development may
-provide opportunities for identifying atypical development earlier and
-offer opportunities for intervention.
-
-To begin to address this need, we investigated the gut microbiome and
-neurocognitive development of children from infancy through 10 years of
-age in an accelerated longitudinal study design. Gut microbial
-communities were assessed using shotgun metagenomic sequencing, enabling
-profiling at both the taxonomic and gene-functional level, and
-neurocognitive development was measured using expert-assessment of
-cognitive function and neuroimaging using magnetic resonance imaging
-(MRI). Using a combination of classical statistical analysis and machine
-learning, we showed that the development of the gut microbiome, the
-brain, and children's cognitive abilities are intimately linked, with
-both microbial taxa and gene functions able to predict cognitive
-performance and brain structure.
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-\hypertarget{cohort-overview-and-summary-data}{%
-\subsubsection{Cohort overview and summary
-data}\label{cohort-overview-and-summary-data}}
+\subsubsection{Cohort overview and summary data}
 
@@
 \textbf{Table1: Subject demographics}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2620}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Group}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Subgroup}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{All}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Under 6 mo}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Over 18 mo}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-N subjects
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-361
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-74
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-257
-\end{minipage} \\
-\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Samples
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-1
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-262 (72.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-74 (100.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-203 (79.0\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-72 (19.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-46 (17.9\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-\textgreater2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-27 (7.5\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-8 (3.1\%)
-\end{minipage} \\
-\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Age (months)
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-Min
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18.0
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Max
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-119.3
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6.0
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-119.3
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Median
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-25.8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-3.7
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-45.9
-\end{minipage} \\
-\multirow{2}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Sex
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-F
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-163 (45.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (50.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-114 (44.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-M
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-198 (54.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (50.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-143 (55.6\%)
-\end{minipage} \\
-\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Race
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-White
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-252 (69.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-42 (56.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-191 (74.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Black
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-40 (11.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16 (21.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-21 (8.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Asian
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Mixed
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-7 (1.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (1.4\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6 (2.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Other
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-48 (13.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12 (16.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-29 (11.3\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Unknown/Declined
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2 (0.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (1.4\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-2 (0.8\%)
-\end{minipage} \\
-\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
-Maternal Ed.
-\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
-Junior high school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 (0.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-1 (0.4\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Some high school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-6 (1.7\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-3 (4.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-4 (1.6\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-High school grad
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-37 (10.2\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-13 (17.6\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16 (6.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Some college
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-95 (26.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-25 (33.8\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-57 (22.2\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-College grad
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-90 (24.9\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17 (23.0\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-69 (26.8\%)
-\end{minipage} \\
-& \begin{minipage}[b]{\linewidth}\raggedright
-Grad/professional school
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-123 (34.1\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-15 (20.3\%)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-104 (40.5\%)
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
 
@@
 
-\hypertarget{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}{%
-\subsubsection{Microbial species and neuroactive genes are associated
-with cognitive
-performance}\label{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}}
+\subsubsection{Microbial species and neuroactive genes are associated with cognitive performance}
 
@@
 
-\textbf{Figure 2 - Taxa and gene functional groups are associated with
-cognitive function.}
+\textbf{Figure 2 - Taxa and gene functional groups are associated with cognitive function.}
 
@@
 
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2300}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1259}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1525}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1429}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2785} + 2\tabcolsep}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{All}
-\end{minipage}} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2591} + 2\tabcolsep}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{under 6 months}
-\end{minipage}} &
-\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2324} + 2\tabcolsep}@{}}{%
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{over 18 months}
-\end{minipage}} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-geneset
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{E.S.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{q value}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1223
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5549
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1944
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1533}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1269
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. II
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2201
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0660}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3424
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0198}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0946
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9199
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Acetate syn. III
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2104
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0859}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3098
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0518}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0857
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9199
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Glutamate deg. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.4037
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1802}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2840
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.2489
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Glutamate syn. II
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0953
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6022
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1342
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9802
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2421
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0466}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Menaquinone syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0853
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5635
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0833
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.1699
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1825}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Propionate deg. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.4612
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.1302}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.5421
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0198}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.6738
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0414}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Propionate syn. I
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3344
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5549
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.3201
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5713
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.6760
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0227}
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Tryptophan syn.
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0573
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.9742
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.0725
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.8159
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1186
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{0.0414}
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\hypertarget{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}{%
-\subsubsection{Gut microbial taxonomic and functional profiles predict
-cognitive
-function}\label{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}}
+\subsubsection{Gut microbial taxonomic and functional profiles predict cognitive function}
 
@@
 
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1456}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0960}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3584}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Subject Ages (months)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Microbial feature
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Demo.
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Test set correlation (± C.I.)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Test set Root-mean-square error (± C.I.)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.14 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-13.01 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.10 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.66 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.13 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.70 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.01 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.56 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-0 to 6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.01 ± 0.01
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-12.56 ± 0.05
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.506 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-16.27 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.363 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.66 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-taxa
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.429 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.29 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.303 ± 0.003
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18.01 ± 0.04
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-18 to 120
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-genes
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-+
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.333 ± 0.004
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-17.86 ± 0.04
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
 As with linear models, RF models for children under 6 months old were
@@
 
-\hypertarget{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}{%
-\subsubsection{Gut microbial taxonomic profiles predict brain structure
-differences}\label{gut-microbial-taxonomic-profiles-predict-brain-structure-differences}}
+\subsubsection{Gut microbial taxonomic profiles predict brain structure differences}
 
@@
 benchmarks}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Statistic
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Mean absolute proportional error (MAPE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Correlation coefficient (R)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-mean
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.072
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.137
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Standard deviation
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.028
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.136
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-maximum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.605
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-75th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.195
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-50th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-25th percentile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.062
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-minimum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.147
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
 
@@
 analyze the multiple relationships unveiled by the RF models
-(\textbf{Figure 4 A and B})
+(Figure 4 A and B)
 
@@
 every model is trained with 132 features, and even distribution of
-importance would be 0.75\% for each feature.
-
-. Segments and taxa ordered by HCA on a select list of species with high
-ranks on average importances, and their respective highest-load
-segments. (C)
+importance would be 0.75\% for each feature. Segments and taxa ordered
+by HCA on a select list of species with high ranks on average importances,
+and their respective highest-load segments (C).
 
@@
 importances, Left = 2.6\%, Right = 1.8\%).
-
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
-
-The relationship between the gut microbiome and brain function via the
-gut-microbiome-brain axis has gained increasing acceptance largely as a
-result of human epidemiological studies investigating atypical
-neurocognition (eg anxiety and depression, neurodegeneration, attention
-deficit / hyperactivity disorder, and autism) and mechanistic studies in
-animal models. The results from these studies point to the possibility
-that gut microbes and their metabolism may be causally implicated in
-cognitive development, but this study is the first to our knowledge that
-directly investigates microbial species and their genes in relation to
-typical development in young children. Understanding the gut-brain-axis
-in early life is particularly important, since differences or
-interventions in early life can have outsized and longer-term
-consequences than those at later ages. Further, even in the absence of
-causal impacts of microbial metabolism, identifying risk factors that
-could point to other early interventions would also have value.
-
-The use of shotgun metagenomic sequencing enabled us to get
-species-level resolution of microbial taxa. A previous study of
-cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing,
-and showed that genera from the Lachnospiraceae family as well as
-unclassified Clostridiales (now Eubacteriales) were associated with
-higher scores on the Ages and Stages Questionnaire
-\href{https://www.zotero.org/google-docs/?SB5Lwa}{(Sordillo et al.,
-2019)}. However, each of these clades encompass dozens of genera with
-diverse functions, each of which may have different effects. Indeed,
-several of the taxa that were positively associated with cognitive
-function in this study, including \emph{B. wexlerae}, the only species
-identified by linear models in children under 6 months, \emph{D.
-longicatena}, \emph{R. faecis}, and \emph{A. finegoldii} are
-Clostridiales, as is \emph{R. gnavus}, which we found was negatively
-associated with cognitive function (Figure 2A-B). This kind of
-species-level resolution is typically not possible with amplicon
-sequencing.
-
-We identified several species in the family Eggerthelaceae that were
-associated with cognitive function, including \emph{Gordonibacter
-pamelaeae, Aldercreutzia equolofaciens, Asaccharobacter celatus}
-(formerly regarded a subspecies of \emph{A. equolofaciens}
-\href{https://www.zotero.org/google-docs/?OEhguI}{(Takahashi et al.,
-2021)}), and \emph{Eggerthella lenta}. Many members of this family are
-known in part due to unique metabolic activities. For example, \emph{A.
-equolofaciens} produces the nonsteroidal estrogen equol from isoflaven
-found in soybeans
-\href{https://www.zotero.org/google-docs/?I2qNyB}{(Wang et al., 2005)},
-and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
-(found in pomegranates and some berries) into urolithin, which has been
-shown in some studies to have a neuroprotective effect
-\href{https://www.zotero.org/google-docs/?oTN8Lw}{(Gong et al., 2022;
-Selma et al., 2014)}. \emph{E. lenta} has been extensively studied for
-its ability to metabolize drug compounds such as the plant-derived heart
-medication digoxin
-\href{https://www.zotero.org/google-docs/?tH2Qeq}{(Haiser et al.,
-2013)}. The metabolic versatility of this clade, and the large number of
-species that are associated with cognition make these microbes prime
-targets for further mechanistic studies.
-
-In addition to improved species-level resolution, shotgun metagenomic
-sequencing also enables gene-functional insight. We showed here that
-genes for the metabolism of SCFAs, both their degradation and synthesis,
-are associated with cognitive function scores. However, while the
-differential abundance of genes for the metabolism of neuroactive
-compounds like these is suggestive, it is difficult to reason about the
-relationship between levels of these genes and the gut concentrations of
-the molecules their product enzymes act on. For example, while it might
-be intuitive to reason that increased levels of menaquinone synthesis
-genes is indicative of increased menaquinone, it could be the case that
-menaquinone deficiency selects for microbes that can synthesize it. For
-the same reason, increased propionate degradation genes may
-counterintuitively be indicative of high levels of propionate in the gut
-lumen, since high propionate would select for microbes that can
-metabolize it. For this reason, future studies coupling shotgun
-metagenomics with stool metabolomics could improve our understanding of
-the relationship between microbial metabolism and cognitive development.
-Further, strain-level analysis linking specific gene content in species
-of interest could further refine targeted efforts at identifying
-specific metabolic signatures of microbe-brain interactions.
-
-\{\{Guilherme add a paragraph about the usefulness of ML\}\}
-
-The use of multiple age-appropriate cognitive assessments that could be
-normalized to a common scale enabled us to analyze microbial
-associations across multiple developmental periods, but carries several
-drawbacks. In particular, the test-retest reliability, as well as
-systematic differences between test administrators may introduce
-substantial noise into these observations, particularly in the youngest
-children. In addition, our study period overlapped with the beginning of
-the COVID-19 pandemic, and we and others have observed some reduction in
-measured scores for children that were assessed after the implementation
-of lockdowns. In our subject set for this study, these effects are more
-pronounced in some age groups due to our sampling schedule
-\href{https://www.zotero.org/google-docs/?L8GnRj}{(Blackwell et al.,
-2022; Deoni et al., 2021)} (Supplementary Figure 3).
-
-This analysis allowed us to establish links between microbial taxa and
-their functional potential with cognition and brain structure. Although
-we cannot test causality or the chemistry behind the interactions
-between gut microbial taxa, gut, and brain, this study provides clear
-and statistically significant associations between the infant and early
-child gut microbiota and neurocognition. Future studies should focus on
-characterizing the early-life microbiome and neurocognitive development
-across different geographic regions and lifestyles such as covering
-traditionally understudied low-resource urban, peri-urban and rural
-communities to obtain the more comprehensive understanding of the
-variability within the different gut microbiomes reflects on
-neurocognition. These studies would also provide us with the wealth of
-data on different strains from the same species to better understand the
-effect of genes and their products. Furthermore, culturing and microbial
-community enrichment studies combined with genetic manipulation and
-genomic approaches to understand microbial metabolism at the molecular
-level is the key, as the metabolic functions shape and influence the
-human host and its health. The discovery of the neuroactive metabolites
-could provide us with biomarkers for early detection or necessary
-medicinally useful molecules that can be applied in intervention.
-
-\hypertarget{material-and-methods}{%
-\subsection{Material and Methods}\label{material-and-methods}}
-
-\hypertarget{study-ethics}{%
-\subsubsection{Study Ethics}\label{study-ethics}}
-
-All procedures for this study were approved by the local institutional
-review board at Rhode Island Hospital, and all experiments adhered to
-the regulation of the review board. Written informed consent was
-obtained from all parents or legal guardians of enrolled participants.
-
-\hypertarget{participants}{%
-\subsubsection{Participants}\label{participants}}
-
-Data used in this study were drawn from the ongoing longitudinal
-RESONANCE study of healthy and neurotypical brain and cognitive
-development, based at Brown University in Providence, RI, USA. The
-RESONANCE study is part of the NIH initiative Environmental influences
-on Child Health Outcomes (ECHO)
-\href{https://www.zotero.org/google-docs/?0tZanX}{(Forrest et al., 2018;
-Gillman and Blaisdell, 2018)}, a longitudinal observational study of
-healthy and neurotypical brain development that spans the fetal and
-infant to adolescent life stages, combining neuroimaging (magnetic
-resonance imaging, MRI), neurocognitive assessments, bio-specimen
-analyses, subject genetics, environmental exposures such as lead, and
-rich demographic, socioeconomic, family and medical history information.
-From the RESONANCE cohort, 361 typically-developing children between the
-ages of 2.8 months and 10 years old (median age 2 years, 2 months) were
-selected for analysis in this study.
-
-General participant demographics are provided in Table 1. Children are
-representative of the RI population. Children in the RESONANCE cohort
-were born full-term (\textless{} 37 weeks gestation) with height and
-weight normal for gestational age, and from uncomplicated singleton
-pregnancies. Children with known major risk factors for developmental
-abnormalities at enrollment were excluded. In addition to screening at
-the time of enrollment, on-going screening for worrisome behaviors using
-validated tools was performed to identify at-risk children and remove
-them from subsequent analysis.
-
-Exclusion criteria included: \emph{in utero} exposure to alcohol,
-cigarette or illicit substance exposure; preterm (\textless{} 37 weeks
-gestation) birth; small for gestational age or less than 1500 g; fetal
-ultrasound abnormalities; preeclampsia, high blood pressure, or
-gestational diabetes; 5 minute APGAR scores \textless{} 8; NICU
-admission; neurological disorder (e.g., head injury resulting in loss of
-consciousness, epilepsy); and psychiatric or learning disorder
-(including maternal depression) in the infant, parents, or siblings
-requiring medication in the year prior to pregnancy.
-
-Demographic and other non-biospecimen data such as race and ethnicity,
-parental education and occupation, feeding behavior (breast- and
-formula-feeding), child weight and height, were collected through
-questionnaires or direct examination as appropriate. All data were
-collected at every assessment visit, if possible.
-
-\hypertarget{cognitive-assessments}{%
-\subsubsection{Cognitive Assessments}\label{cognitive-assessments}}
-
-Overall cognitive function was assessed using age-appropriate methods.
-For children from birth to 30 months, we used an Early Learning
-Composite as assessed via the Mullen Scales of Early Learning (MSEL)
-\href{https://www.zotero.org/google-docs/?NSyykd}{(Mullen and others,
-1995)}, a standardized and population-normed tool for assessing fine and
-gross motor, expressive and receptive language, and visual reception
-functioning in children from birth through 68 months of age. The
-Wechsler Intelligence Quotient for Children (WISC)
-\href{https://www.zotero.org/google-docs/?OopdA1}{(Wechsler, 2012)} is
-an individually administered standard intelligence test for children
-aged 6 to 16 years. It derives a full scale intelligence quotient (IQ)
-score, which we used to assess overall cognitive functioning. The fourth
-edition of the Wechsler Preschool and Primary Scale of Intelligence
-(WPPSI-IV) is an individually administered standard intelligence test
-for children aged 2 years 6 months to 7 years 7 months, trying to meet
-the increasing need for the assessment of preschoolers. Just as the
-WISC, it derives a full scale IQ score, which we used to assess overall
-cognitive functioning.
-
-\hypertarget{stool-sample-collection-and-sequencing}{%
-\subsubsection{Stool Sample Collection and
-Sequencing}\label{stool-sample-collection-and-sequencing}}
-
-Stool samples (n=493) were collected by parents in OMR-200 tubes
-(OMNIgene GUT, DNA Genotek, Ottawa, Ontario, Canada), stored on ice, and
-brought within 24 hrs to the lab in RI where they were immediately
-frozen at -80 ˚C. Stool samples were not collected if the subject had
-taken antibiotics within the last two weeks. DNA extraction was
-performed at Wellesley College (Wellesley, MA). Nucleic acids were
-extracted from stool samples using the RNeasy PowerMicrobiome kit,
-excluding the DNA degradation steps. Briefly, the samples were lysed by
-bead beating using the Powerlyzer 24 Homogenizer (Qiagen, Germantown,
-MD) at 2500 rpm for 45 s and then transferred to the QIAcube (Qiagen,
-Germantown, MD) to complete the extraction protocol. Extracted DNA was
-sequenced at the Integrated Microbiome Resource (IMR, Dalhousie
-University, NS, Canada).
-
-Shotgun metagenomic sequencing was performed on all samples. A pooled
-library (max 96 samples per run) was prepared using the Illumina Nextera
-Flex Kit for MiSeq and NextSeq from 1 ng of each sample. Samples were
-then pooled onto a plate and sequenced on the Illumina NextSeq 550
-platform using 150+150 bp paired-end ``high output'' chemistry,
-generating 400 million raw reads and 120 Gb of sequence per plate.
-
-\hypertarget{machine-learning-for-cognitive-development}{%
-\subsubsection{Machine learning for cognitive
-development}\label{machine-learning-for-cognitive-development}}
-
-Prediction of cognitive scores was carried out as a set of regression
-experiments targeting real-valued continuous assessment scores.
-Different experiment sets were designed to probe how different
-representations of the gut microbiome (taxonomic profiles, Functional
-profiles encoded as ECs) would behave, with and without the addition of
-demographics (sex and maternal education as a proxy of socioeconomic
-status) on participants from different age groups. Age (in months) was
-provided as a covariate for all models (Table 3).
-
-\textbf{Table 5. Experimental design and input composition for Random
-Forest experiments}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1797}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2915}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Input set
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Age bracket
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Microbiome encoding type
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Demographics
-
-Provided?
-
-(sex, education)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-1
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Not provided
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-2
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-3
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-4
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-5
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0 to 6 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-6
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Not provided
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-7
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-8
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Taxonomic profile
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-9
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-no
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-10
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-18 to 120 months
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Functional Profile (ECs)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-yes
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-Random Forests (RFs) \href{https://paperpile.com/c/dPbU4e/VDqU}{(Breiman
-1996)} were selected as the prediction engine and processed using the
-DecisionTree.jl
-\href{https://www.zotero.org/google-docs/?dQLyDs}{(Sadeghi et al.,
-2022)} implementation, inside the MLJ.jl
-\href{https://www.zotero.org/google-docs/?y8ywGI}{(Blaom et al., 2020)}
-Machine Learning framework. Independent RFs were trained for each
-experiment, using a set of default regression hyperparameters from
-Breiman and Cutler
-\href{https://www.zotero.org/google-docs/?d3gqFh}{(Breiman, 2001)}, on a
-repeated cross-validation approach with different RNG seeds. One hundred
-repetitions of 3-fold CV with 10 different intra-fold RNG states each
-were employed, for a total of 3000 experiments per input set.
-
-After the training procedures, the root-mean-square error (RMSE) for
-cognitive assessment scores and mean absolute proportional error (MAPE)
-for the brain segmentation data, along with Pearson\textquotesingle s
-correlation coefficient (R) were benchmarked on the validation and train
-sets. MAPE was chosen as the metric for brain segments due to magnitude
-differences between median volumes of each segment, which would hinder
-interpretation of raw error values without additional reference.
-
-To derive biological insight from the models, the covariate variable
-importances for all the input features, measured by mean decrease in
-impurity (MDI, or GINI importance), was also analyzed. Leveraging the
-distribution of results from the extensive repeated cross validation
-experiments, rather than electing a representative model or picking the
-highest validation-set correlation, a measure of model fitness
-(\textbf{Equation 1}) was designed to weight the importances from each
-trained forest. The objective was to give more weight to those with
-higher benchmarks on the validation sets (or higher generalizability),
-while penalizing information from highly overfit models, drawing
-inspiration from the approach used on another work employing repeated CV
-on Random Forests with high-dimensional, low sample size microbiome
-datasets \{***\}. The resulting fitness-weighted importances were used
-to generate the values in Figure 3.
-
-\(fitness =\)
-
-\textbf{Equation 1.} Mathematical expression of the fitness measure used
-to weight feature importances based on model benchmarks
-
-\hypertarget{mri-segmentation}{%
-\subsubsection{MRI / segmentation}\label{mri-segmentation}}
-
-MRI data was acquired at a 3T Siemens Trio scanner with the following
-parameters: TE=5.6msec, TR=1400msec, FA=15 degrees, 1.1x1.1x1.1mm
-resolution, 160x160 matrix, with an average of 112 slices. FOV was
-adjusted to infant size. Using a combination of linear and nonlinear
-image registration, we created representative age-specific templates
-using the ANTs package
-\href{https://www.zotero.org/google-docs/?ZlWDfL}{(Avants et al.,
-2014)}. After age specific templates were created, a single flow from
-each age to the 12-month template was estimated and a final warp from
-the 12-month template to standard adult MNI space was performed. For
-each individual infant/ child brain image, we then calculated the warp
-from their native T1w image space to their nearest-in-age template.
-Using the resulting warps (native → nearest age template → 12-month
-template → MNI template), we could move the standard adult brain atlas
-to the space of an individual infant in a single step. In this case, we
-used the Harvard-Oxford brain atlas to provide a coarse-grained
-parcellation of individual brains into subcortical regions (e.g.,
-thalamus, putamen) and total grey and white matter volumes (included as
-part of the FSL package
-\href{https://www.zotero.org/google-docs/?AapPN7}{(Jenkinson et al.,
-2012)}. Total tissue and brainstem volume as well as left and right
-hemisphere volumes were derived for total white and cortical gray
-matter, lateral ventricle, thalamus, caudate, putamen, pallidum,
-hippocampus, amygdala, and accumbens as well as total brainstem volume
-\href{https://www.zotero.org/google-docs/?MCDDB9}{(Bruchhage et al.,
-n.d.)}.
-
-\hypertarget{computational-analysis-of-metagenomes}{%
-\subsubsection{Computational analysis of
-metagenomes}\label{computational-analysis-of-metagenomes}}
-
-Shotgun metagenomic sequences were analyzed using the bioBakery suite of
-computational tools \href{https://paperpile.com/c/dPbU4e/vDau}{(Beghini
-et al. 2021)}. First, KneadData (v0.7.7) was used to perform quality
-control of raw sequence reads, such as read trimming and removal of
-reads matching a human genome reference. Next, MetaPhlAn (v3.0.7, using
-database mpa\_v30\_CHOCOPhlAn\_201901) was used to generate taxonomic
-profiles by aligning reads to a reference database of marker genes.
-Finally, HUMAnN (v3.0.0a4) was used to functionally profile the
-metagenomes.
-
-\hypertarget{data-and-code-availability}{%
-\subsubsection{Data and code
-availability}\label{data-and-code-availability}}
-
-Taxonomic and functional microbial profiles, as well as subject
-demographics necessary for statistical analyses and machine learning are
-available on the Open Science Framework
-\href{https://www.zotero.org/google-docs/?61HLlM}{(Bonham et al.,
-2022)}. Data processing, generation of summary statistics, and
-generation of plots was performed using the julia programming language
-\href{https://www.zotero.org/google-docs/?zj3HnY}{(Bezanson et al.,
-2017; Bonham et al., 2021; Danisch and Krumbiegel, 2021)}. All code for
-data analysis and figure generation, as well as scripts for automated
-download of input files are available on github \{\{CITE zenodo\}\}.
-
-\hypertarget{research-standards}{%
-\subsection{Research Standards}\label{research-standards}}
-
-\hypertarget{acknowledgements}{%
-\subsection{Acknowledgements}\label{acknowledgements}}
-
-Research funding: NIH UG3 OD023313 and Wellcome LEAP 1kD.
-
-\hypertarget{references-and-citations}{%
-\subsection{References and Citations}\label{references-and-citations}}
-
-\url{https://doi.org/10.1093/cercor/bhs201}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Amaratunga, D.,
-Cabrera, J., Lee, Y.-S., 2008. Enriched random forests. Bioinformatics
-24, 2010--2014. https://doi.org/10.1093/bioinformatics/btn356}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Aminoff, E.M.,
-Kveraga, K., Bar, M., 2013. The role of the parahippocampal cortex in
-cognition. Trends Cogn. Sci. 17, 379--390.
-https://doi.org/10.1016/j.tics.2013.06.009}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Apps, M.A.J.,
-Rushworth, M.F.S., Chang, S.W.C., 2016. The Anterior Cingulate Gyrus and
-Social Cognition: Tracking the Motivation of Others. Neuron 90,
-692--707. https://doi.org/10.1016/j.neuron.2016.04.018}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ardekani, B.A.,
-Bermudez, E., Mubeen, A.M., Bachman, A.H., Initiative, for the A.D.N.,
-2017. Prediction of Incipient Alzheimer's Disease Dementia in Patients
-with Mild Cognitive Impairment. J. Alzheimers Dis. 55, 269--281.
-https://doi.org/10.3233/JAD-160594}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Avants, B.B.,
-Tustison, N.J., Stauffer, M., Song, G., Wu, B., Gee, J.C., 2014. The
-insight ToolKit image registration framework. Front. Neuroinform. 8, 44.
-https://doi.org/10.3389/fninf.2014.00044}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bäckhed, F., Roswall,
-J., Peng, Y., Feng, Q., Jia, H., Kovatcheva-Datchary, P., Li, Y., Xia,
-Y., Xie, H., Zhong, H., Khan, M.T., Zhang, J., Li, J., Xiao, L.,
-Al-Aama, J., Zhang, D., Lee, Y.S., Kotowska, D., Colding, C., Tremaroli,
-V., Yin, Y., Bergman, S., Xu, X., Madsen, L., Kristiansen, K., Dahlgren,
-J., Wang, J., 2015. Dynamics and stabilization of the human gut
-microbiome during the first year of life. Cell Host Microbe 17,
-690--703. https://doi.org/10.1016/j.chom.2015.04.004}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bairoch, A., 2000. The
-ENZYME database in 2000. Nucleic Acids Res. 28, 304--305.
-https://doi.org/10.1093/nar/28.1.304}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Benítez-Páez, A.,
-Gómez del Pugar, E.M., López-Almela, I., Moya-Pérez, Á., Codoñer-Franch,
-P., Sanz, Y., 2020. Depletion of \emph{Blautia} Species in the
-Microbiota of Obese Children Relates to Intestinal Inflammation and
-Metabolic Phenotype Worsening. mSystems 5, e00857-19.
-https://doi.org/10.1128/mSystems.00857-19}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bezanson, J., Edelman,
-A., Karpinski, S., Shah, V.B., 2017. Julia: A fresh approach to
-numerical computing. Siam Rev. 59, 65--98.
-https://doi.org/10.1137/141000671}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Blackwell, C.K.,
-Mansolf, M., Sherlock, P., Ganiban, J., Hofheimer, J.A., Barone, C.J.,
-Bekelman, T.A., Blair, C., Cella, D., Collazo, S., Croen, L.A., Deoni,
-S., Elliott, A.J., Ferrara, A., Fry, R.C., Gershon, R., Herbstman, J.B.,
-Karagas, M.R., LeWinn, K.Z., Margolis, A., Miller, R.L., O'Shea, T.M.,
-Porucznik, C.A., Wright, R.J., 2022. Youth well-being during the
-COVID-19 pandemic. Pediatrics 149.
-https://doi.org/10.1542/peds.2021-054754}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Blaom, A.D., Kiraly,
-F., Lienart, T., Simillides, Y., Arenas, D., Vollmer, S.J., 2020. MLJ: A
-Julia package for composable machine learning. J. Open Source Softw. 5,
-2704. https://doi.org/10.21105/joss.02704}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Boes, A.D., Tranel,
-D., Anderson, S.W., Nopoulos, P., 2008. Right anterior cingulate cortex
-volume is a neuroanatomical correlate of aggression and defiance in
-boys. Behav. Neurosci. 122, 677--684.
-https://doi.org/10.1037/0735-7044.122.3.677}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bokulich, N.A., Chung,
-J., Battaglia, T., Henderson, N., Jay, M., Li, H., D Lieber, A., Wu, F.,
-Perez-Perez, G.I., Chen, Y., Schweizer, W., Zheng, X., Contreras, M.,
-Dominguez-Bello, M.G., Blaser, M.J., 2016. Antibiotics, birth mode, and
-diet shape microbiome maturation during early life. Sci. Transl. Med. 8,
-343ra82. https://doi.org/10.1126/scitranslmed.aad7121}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Bottino,
-G.Z.M.F., Klepac-Ceraj, V., 2022. ECHO RESONANCE Microbiome.
-https://doi.org/10.17605/OSF.IO/YBS32}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bonham, K., Kayisire,
-A., Luo, A., Klepac-Ceraj, V., 2021. Microbiome.jl and BiobakeryUtils.jl
-- Julia packages for working with microbial community data. J Open
-Source Softw 6, 3876. https://doi.org/10.21105/joss.03876}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Breiman, L., 2001.
-Random Forests. Mach. Learn. 45, 5--32.
-https://doi.org/10.1023/A:1010933404324}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Brieuc, M.S.O.,
-Waters, C.D., Drinan, D.P., Naish, K.A., 2018. A practical introduction
-to Random Forest for genetic association studies in ecology and
-evolution. Mol. Ecol. Resour. 18, 755--766.
-https://doi.org/10.1111/1755-0998.12773}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bruchhage, Muriel,
-Chen, Y., Cataldo, A.G., Müller, H.-G., Weisee, E., Wilson, S., Pietsch,
-M., D'Sa, V., Marquand, A., Madhow, S., Bouchard, K., Cole, J.H.,
-Biondo, F., Elison, J., O'Muirchheartaigh, J., Deoni, C, S., n.d.
-Longitudinal brain and cognitive development of the first 1000 Days:A
-large Multi-Cohort Multi-Scanner study. Presented at the International
-Society of Magnetic Resonance Medicine.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Bush, G., Vogt, B.A.,
-Holmes, J., Dale, A.M., Greve, D., Jenike, M.A., Rosen, B.R., 2002.
-Dorsal anterior cingulate cortex: A role in reward-based decision
-making. Proc. Natl. Acad. Sci. 99, 523--528.
-https://doi.org/10.1073/pnas.012470999}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Callaghan, B., 2020.
-Nested sensitive periods: how plasticity across the microbiota-gut-brain
-axis interacts to affect the development of learning and memory. Curr.
-Opin. Behav. Sci. 36, 55--62.
-https://doi.org/10.1016/j.cobeha.2020.07.011}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cao, Z.,
-Ottino-Gonzalez, J., Cupertino, R.B., Schwab, N., Hoke, C., Catherine,
-O., Cousijn, J., Dagher, A., Foxe, J.J., Goudriaan, A.E., Hester, R.,
-Hutchison, K., Li, C.-S.R., London, E.D., Lorenzetti, V., Luijten, M.,
-Martin-Santos, R., Momenan, R., Paulus, M.P., Schmaal, L., Sinha, R.,
-Sjoerds, Z., Solowij, N., Stein, D.J., Stein, E.A., Uhlmann, A., van
-Holst, R.J., Veltman, D.J., Wiers, R.W., Yücel, M., Zhang, S.,
-Jahanshad, N., Thompson, P.M., Conrod, P., Mackey, S., Garavan, H.,
-2021. Mapping cortical and subcortical asymmetries in substance
-dependence: Findings from the ENIGMA Addiction Working Group. Addict.
-Biol. 26, e13010. https://doi.org/10.1111/adb.13010}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cerdó, T., Diéguez,
-E., Campoy, C., 2019. Early nutrition and gut microbiome:
-interrelationship between bacterial metabolism, immune system, brain
-structure, and neurodevelopment. Am. J. Physiol. Endocrinol. Metab. 317,
-E617--E630. https://doi.org/10.1152/ajpendo.00188.2019}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Chen, X., Ishwaran,
-H., 2012. Random forests for genomic data analysis. Genomics 99,
-323--329. https://doi.org/10.1016/j.ygeno.2012.04.003}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cohen Kadosh, K.,
-Krause, B., King, A.J., Near, J., Cohen Kadosh, R., 2015. Linking GABA
-and glutamate levels to cognitive skill acquisition during development.
-Hum. Brain Mapp. 36, 4334--4345. https://doi.org/10.1002/hbm.22921}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Cowan, C.S.M., Dinan,
-T.G., Cryan, J.F., 2020. Annual Research Review: Critical windows - the
-microbiota-gut-brain axis in neurocognitive development. J. Child
-Psychol. Psychiatry 61, 353--371. https://doi.org/10.1111/jcpp.13156}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dalile, B., Van
-Oudenhove, L., Vervliet, B., Verbeke, K., 2019. The role of short-chain
-fatty acids in microbiota-gut-brain communication. Nat. Rev.
-Gastroenterol. Hepatol. 16, 461--478.
-https://doi.org/10.1038/s41575-019-0157-3}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Danisch, S.,
-Krumbiegel, J., 2021. Makie.jl: Flexible high-performance data
-visualization for Julia. J. Open Source Softw. 6, 3349.
-https://doi.org/10.21105/joss.03349}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Deoni, S.C.,
-Beauchemin, J., Volpe, A., Dâ Sa, V., RESONANCE Consortium, 2021. Impact
-of the COVID-19 pandemic on early child cognitive development: Initial
-findings in a longitudinal observational study of child health. MedRxiv
-Prepr. Serv. Health Sci. https://doi.org/10.1101/2021.08.10.21261846}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dominguez-Bello, M.G.,
-Costello, E.K., Contreras, M., Magris, M., Hidalgo, G., Fierer, N.,
-Knight, R., 2010. Delivery mode shapes the acquisition and structure of
-the initial microbiota across multiple body habitats in newborns. Proc.
-Natl. Acad. Sci. U. S. A. 107, 11971--11975.
-https://doi.org/10.1073/pnas.1002601107}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Dong, H., Wang, B.,
-Feng, J., Yue, X., Jia, F., 2021. Correlation between serum
-concentrations of menaquinone-4 and developmental quotients in children
-with autism spectrum disorder. Front. Nutr. 8, 748513.
-https://doi.org/10.3389/fnut.2021.748513}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ernst, M., Nelson,
-E.E., Jazbec, S., McClure, E.B., Monk, C.S., Leibenluft, E., Blair, J.,
-Pine, D.S., 2005. Amygdala and nucleus accumbens in responses to receipt
-and omission of gains in adults and adolescents. NeuroImage 25,
-1279--1291. https://doi.org/10.1016/j.neuroimage.2004.12.038}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Forrest, C.B.,
-Blackwell, C.K., Camargo, C.A., 2018. Advancing the science of
-children's positive health in the national institutes of health
-environmental influences on child health outcomes (ECHO) research
-program. J. Pediatr. 196, 298--300.
-https://doi.org/10.1016/j.jpeds.2018.02.004}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Franzosa, E.A.,
-Sirota-Madi, A., Avila-Pacheco, J., Fornelos, N., Haiser, H.J., Reinker,
-S., Vatanen, T., Hall, A.B., Mallick, H., McIver, L.J., Sauk, J.S.,
-Wilson, R.G., Stevens, B.W., Scott, J.M., Pierce, K., Deik, A.A.,
-Bullock, K., Imhann, F., Porter, J.A., Zhernakova, A., Fu, J., Weersma,
-R.K., Wijmenga, C., Clish, C.B., Vlamakis, H., Huttenhower, C., Xavier,
-R.J., 2019. Gut microbiome structure and metabolic activity in
-inflammatory bowel disease. Nat. Microbiol. 4, 293--305.
-https://doi.org/10.1038/s41564-018-0306-4}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Fung, T.C., Olson,
-C.A., Hsiao, E.Y., 2017. Interactions between the microbiota, immune and
-nervous systems in health and disease. Nat. Neurosci. 20, 145--155.
-https://doi.org/10.1038/nn.4476}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ghosh, T.S., Rampelli,
-S., Jeffery, I.B., Santoro, A., Neto, M., Capri, M., Giampieri, E.,
-Jennings, A., Candela, M., Turroni, S., Zoetendal, E.G., Hermes, G.D.A.,
-Elodie, C., Meunier, N., Brugere, C.M., Pujos-Guillot, E., Berendsen,
-A.M., Groot, L.C.P.G.M.D., Feskins, E.J.M., Kaluza, J., Pietruszka, B.,
-Bielak, M.J., Comte, B., Maijo-Ferre, M., Nicoletti, C., Vos, W.M.D.,
-Fairweather-Tait, S., Cassidy, A., Brigidi, P., Franceschi, C., O'Toole,
-P.W., 2020. Mediterranean diet intervention alters the gut microbiome in
-older people reducing frailty and improving health status: the NU-AGE
-1-year dietary intervention across five European countries. Gut 69,
-1218--1228. https://doi.org/10.1136/gutjnl-2019-319654}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Gillman, M.W.,
-Blaisdell, C.J., 2018. Environmental influences on child health
-outcomes, a research program of the NIH. Curr. Opin. Pediatr. 30, 260.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Gong, Q.-Y., Cai, L.,
-Jing, Y., Wang, W., Yang, D.-X., Chen, S.-W., Tian, H.-L., 2022.
-Urolithin A alleviates blood-brain barrier disruption and attenuates
-neuronal apoptosis following traumatic brain injury in mice. Neural
-Regen. Res. 17, 2007. https://doi.org/10.4103/1673-5374.335163}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Haam, J., Yakel, J.L.,
-2017. Cholinergic modulation of the hippocampal region and memory
-function. J. Neurochem. 142, 111--121.
-https://doi.org/10.1111/jnc.14052}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Haiser, H.J.,
-Gootenberg, D.B., Chatman, K., Sirasani, G., Balskus, E.P., Turnbaugh,
-P.J., 2013. Predicting and Manipulating Cardiac Drug Inactivation by the
-Human Gut Bacterium Eggerthella lenta. Science 341, 295--298.
-https://doi.org/10.1126/science.1235872}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hartstra, A.V.,
-Schüppel, V., Imangaliyev, S., Schrantee, A., Prodan, A., Collard, D.,
-Levin, E., Dallinga-Thie, G., Ackermans, M.T., Winkelmeijer, M., Havik,
-S.R., Metwaly, A., Lagkouvardos, I., Nier, A., Bergheim, I.,
-Heikenwalder, M., Dunkel, A., Nederveen, A.J., Liebisch, G., Mancano,
-G., Claus, S.P., Benítez-Páez, A., la Fleur, S.E., Bergman, J.J.,
-Gerdes, V., Sanz, Y., Booij, J., Kemper, E., Groen, A.K., Serlie, M.J.,
-Haller, D., Nieuwdorp, M., 2020. Infusion of donor feces affects the
-gut--brain axis in humans with metabolic syndrome. Mol. Metab. 42,
-101076. https://doi.org/10.1016/j.molmet.2020.101076}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoogman, M., Bralten,
-J., Hibar, D.P., Mennes, M., Zwiers, M.P., Schweren, L., van Hulzen,
-K.J.E., Medland, S.E., Shumskaya, E., Jahanshad, N., de Zeeuw, P.,
-Szekely, E., Sudre, G., Wolfers, T., Onnink, A.M.H., Dammers, J.T.,
-Mostert, J.C., Vives-Gilabert, Y., Kohls, G., Oberwelland, E., Seitz,
-J., Schulte-Rüther, M., di Bruttopilo, S.A., Doyle, A.E., Høvik, M.F.,
-Dramsdahl, M., Tamm, L., van Erp, T.G.M., Dale, A., Schork, A.,
-Conzelmann, A., Zierhut, K., Baur, R., McCarthy, H., Yoncheva, Y.N.,
-Cubillo, A., Chantiluke, K., Mehta, M.A., Paloyelis, Y., Hohmann, S.,
-Baumeister, S., Bramati, I., Mattos, P., Tovar-Moll, F., Douglas, P.,
-Banaschewski, T., Brandeis, D., Kuntsi, J., Asherson, P., Rubia, K.,
-Kelly, C., Di Martino, A., Milham, M.P., Castellanos, F.X., Frodl, T.,
-Zentis, M., Lesch, K.-P., Reif, A., Pauli, P., Jernigan, T., Haavik, J.,
-Plessen, K.J., Lundervold, A.J., Hugdahl, K., Seidman, L.J., Biederman,
-J., Rommelse, N., Heslenfeld, D.J., Hartman, C., Hoekstra, P.J.,
-Oosterlaan, J., von Polier, G., Konrad, K., Vilarroya, O.,
-Ramos-Quiroga, J.-A., Soliva, J.C., Durston, S., Buitelaar, J.K.,
-Faraone, S.V., Shaw, P., Thompson, P., Franke, B., 2017. Subcortical
-brain volume differences of participants with ADHD across the lifespan:
-an ENIGMA collaboration. Lancet Psychiatry 4, 310--319.
-https://doi.org/10.1016/S2215-0366(17)30049-4}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hoshino, Y., Yamamoto,
-T., Kaneko, M., Tachibana, R., Watanabe, M., Ono, Y., Kumashiro, H.,
-1984. Blood Serotonin and Free Tryptophan Concentration in Autistic
-Children. Neuropsychobiology 11, 22--27.
-https://doi.org/10.1159/000118045}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Hosomi, K., Saito, M.,
-Park, J., Murakami, H., Shibata, N., Ando, M., Nagatake, T., Konishi,
-K., Ohno, H., Tanisawa, K., Mohsen, A., Chen, Y.-A., Kawashima, H.,
-Natsume-Kitatani, Y., Oka, Y., Shimizu, H., Furuta, M., Tojima, Y.,
-Sawane, K., Saika, A., Kondo, S., Yonejima, Y., Takeyama, H., Matsutani,
-A., Mizuguchi, K., Miyachi, M., Kunisawa, J., 2022. Oral administration
-of Blautia wexlerae ameliorates obesity and type 2 diabetes via
-metabolic remodeling of the gut microbiota. Nat. Commun. 13, 4477.
-https://doi.org/10.1038/s41467-022-32015-7}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Jenkinson, M.,
-Beckmann, C.F., Behrens, T.E.J., Woolrich, M.W., Smith, S.M., 2012. FSL.
-Neuroimage 62, 782--790.
-https://doi.org/10.1016/j.neuroimage.2011.09.015}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Jiang, H., Ling, Z.,
-Zhang, Y., Mao, H., Ma, Z., Yin, Y., Wang, W., Tang, W., Tan, Z., Shi,
-J., Li, L., Ruan, B., 2015. Altered fecal microbiota composition in
-patients with major depressive disorder. Brain. Behav. Immun. 48,
-186--194. https://doi.org/10.1016/j.bbi.2015.03.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kanehisa, M., Goto,
-S., Kawashima, S., Okuno, Y., Hattori, M., 2004. The KEGG resource for
-deciphering the genome. Nucleic Acids Res. 32, D277--D280.
-https://doi.org/10.1093/nar/gkh063}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kant, R., Rasinkangas,
-P., Satokari, R., Pietilä, T.E., Palva, A., 2015. Genome Sequence of the
-Butyrate-Producing Anaerobic Bacterium Anaerostipes hadrus PEL 85.
-Genome Announc. 3, e00224-15. https://doi.org/10.1128/genomeA.00224-15}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Kim, C.-S., Cha, L.,
-Sim, M., Jung, S., Chun, W.Y., Baik, H.W., Shin, D.-M., 2021. Probiotic
-supplementation improves cognitive function and mood with changes in gut
-microbiota in Community-Dwelling older adults: A randomized,
-Double-Blind, Placebo-Controlled, multicenter trial. J Gerontol Biol Sci
-Med Sci 76, 32--40. https://doi.org/10.1093/gerona/glaa090}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Knudsen, E.I., 2004.
-Sensitive periods in the development of the brain and behavior. J. Cogn.
-Neurosci. 16, 1412--1425. https://doi.org/10.1162/0898929042304796}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Koenig, J.E., Spor,
-A., Scalfone, N., Fricker, A.D., Stombaugh, J., Knight, R., Angenent,
-L.T., Ley, R.E., 2011. Succession of microbial consortia in the
-developing infant gut microbiome. Proc. Natl. Acad. Sci. U. S. A. 108
-Suppl 1, 4578--4585. https://doi.org/10.1073/pnas.1000081107}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Coker,
-M.O., Madan, J.C., 2022. The developing microbiome from birth to 3
-years: The Gut-Brain axis and neurodevelopmental outcomes. Front.
-Pediatr. 10, 815885. https://doi.org/10.3389/fped.2022.815885}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Laue, H.E., Korrick,
-S.A., Baker, E.R., Karagas, M.R., Madan, J.C., 2020. Prospective
-associations of the infant gut microbiome and microbial function with
-social behaviors related to autism at age 3 years. Sci. Rep. 10, 15515.
-https://doi.org/10.1038/s41598-020-72386-9}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Ley, R.E., Turnbaugh,
-P.J., Klein, S., Gordon, J.I., 2006. Human gut microbes associated with
-obesity. Nature 444, 1022--1023. https://doi.org/10.1038/4441022a}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Li, Y.-M., Yu, Q.,
-Huang, L., Fu, J., Feng, R.-J., 2022. {[}Correlations Between Gut
-Microbiota Changes and Cognitive Function in Patients with Post-Stroke
-Cognitive Impairment in the Early Stage{]}. Sichuan Da Xue Xue Bao Yi
-Xue Ban 53, 857--865. https://doi.org/10.12182/20220960105}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, S., Li, E., Sun,
-Z., Fu, D., Duan, G., Jiang, M., Yu, Y., Mei, L., Yang, P., Tang, Y.,
-Zheng, P., 2019. Altered gut microbiota and short chain fatty acids in
-Chinese children with autism spectrum disorder. Sci. Rep. 9, 287.
-https://doi.org/10.1038/s41598-018-36430-z}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Liu, X., Mao, B., Gu,
-J., Wu, J., Cui, S., Wang, G., Zhao, J., Zhang, H., Chen, W., 2021.
-\emph{Blautia} ---a new functional genus with potential probiotic
-properties? Gut Microbes 13, 1875796.
-https://doi.org/10.1080/19490976.2021.1875796}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Lopez-Siles, M.,
-Duncan, S.H., Garcia-Gil, L.J., Martinez-Medina, M., 2017.
-Faecalibacterium prausnitzii: from microbiology to diagnostics and
-prognostics. ISME J. 11, 841--852.
-https://doi.org/10.1038/ismej.2016.176}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Louwies, T., Johnson,
-A.C., Orock, A., Yuan, T., Greenwood-Van Meerveld, B., 2020. The
-microbiota-gut-brain axis: An emerging role for the epigenome. Exp.
-Biol. Med. 245, 138--145. https://doi.org/10.1177/1535370219891690}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Magnusson, K.R.,
-Hauck, L., Jeffrey, B.M., Elias, V., Humphrey, A., Nath, R., Perrone,
-A., Bermudez, L.E., 2015. Relationships between diet-related changes in
-the gut microbiome and cognitive flexibility. Neuroscience 300,
-128--140. https://doi.org/10.1016/j.neuroscience.2015.05.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mallick, H.,
-Rahnavard, A., McIver, L.J., Ma, S., Zhang, Y., Nguyen, L.H., Tickle,
-T.L., Weingart, G., Ren, B., Schwager, E.H., Chatterjee, S., Thompson,
-K.N., Wilkinson, J.E., Subramanian, A., Lu, Y., Waldron, L., Paulson,
-J.N., Franzosa, E.A., Bravo, H.C., Huttenhower, C., 2021. Multivariable
-association discovery in population-scale meta-omics studies.
-https://doi.org/10.1101/2021.01.20.427420}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Maruo, T., Sakamoto,
-M., Ito, C., Toda, T., Benno, Y., 2008. Adlercreutzia equolifaciens gen.
-nov., sp. nov., an equol-producing bacterium isolated from human faeces,
-and emended description of the genus Eggerthella. Int. J. Syst. Evol.
-Microbiol. 58, 1221--1227. https://doi.org/10.1099/ijs.0.65404-0}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mayneris-Perxachs, J.,
-Castells-Nobau, A., Arnoriaga-Rodríguez, M., Martin, M., de la
-Vega-Correa, L., Zapata, C., Burokas, A., Blasco, G., Coll, C.,
-Escrichs, A., Biarnés, C., Moreno-Navarrete, J.M., Puig, J., Garre-Olmo,
-J., Ramos, R., Pedraza, S., Brugada, R., Vilanova, J.C., Serena, J.,
-Gich, J., Ramió-Torrentà, L., Pérez-Brocal, V., Moya, A., Pamplona, R.,
-Sol, J., Jové, M., Ricart, W., Portero-Otin, M., Deco, G., Maldonado,
-R., Fernández-Real, J.M., 2022. Microbiota alterations in proline
-metabolism impact depression. Cell Metab. 34, 681-701.e10.
-https://doi.org/10.1016/j.cmet.2022.04.001}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mukherjee, A., Lordan,
-C., Ross, R.P., Cotter, P.D., 2020. Gut microbes from the
-phylogenetically diverse genus Eubacterium and their various
-contributions to gut health. Gut Microbes 12, 1802866.
-https://doi.org/10.1080/19490976.2020.1802866}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Mullen, E.M., others,
-1995. Mullen scales of early learning. AGS Circle Pines, MN.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Needham, B.D.,
-Funabashi, M., Adame, M.D., Wang, Z., Boktor, J.C., Haney, J., Wu,
-W.-L., Rabut, C., Ladinsky, M.S., Hwang, S.-J., Guo, Y., Zhu, Q.,
-Griffiths, J.A., Knight, R., Bjorkman, P.J., Shapiro, M.G., Geschwind,
-D.H., Holschneider, D.P., Fischbach, M.A., Mazmanian, S.K., 2022. A
-gut-derived metabolite alters brain activity and anxiety behaviour in
-mice. Nature 602, 647--653. https://doi.org/10.1038/s41586-022-04396-8}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Opel, N., Thalamuthu,
-A., Milaneschi, Y., Grotegerd, D., Flint, C., Leenings, R., Goltermann,
-J., Richter, M., Hahn, T., Woditsch, G., Berger, K., Hermesdorf, M.,
-McIntosh, A., Whalley, H.C., Harris, M.A., MacMaster, F.P., Walter, H.,
-Veer, I.M., Frodl, T., Carballedo, A., Krug, A., Nenadic, I., Kircher,
-T., Aleman, A., Groenewold, N.A., Stein, D.J., Soares, J.C.,
-Zunta-Soares, G.B., Mwangi, B., Wu, M.-J., Walter, M., Li, M., Harrison,
-B.J., Davey, C.G., Cullen, K.R., Klimes-Dougan, B., Mueller, B.A.,
-Sämann, P.G., Penninx, B., Nawijn, L., Veltman, D.J., Aftanas, L., Brak,
-I.V., Filimonova, E.A., Osipov, E.A., Reneman, L., Schrantee, A., Grabe,
-H.J., Van der Auwera, S., Wittfeld, K., Hosten, N., Völzke, H., Sim, K.,
-Gotlib, I.H., Sacchet, M.D., Lagopoulos, J., Hatton, S.N., Hickie, I.,
-Pozzi, E., Thompson, P.M., Jahanshad, N., Schmaal, L., Baune, B.T.,
-Dannlowski, U., 2021. Brain structural abnormalities in obesity:
-relation to age, genetic risk, and common psychiatric disorders. Mol.
-Psychiatry 26, 4839--4852. https://doi.org/10.1038/s41380-020-0774-9}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Pagonabarraga, J.,
-Corcuera-Solano, I., Vives-Gilabert, Y., Llebaria, G., García-Sánchez,
-C., Pascual-Sedano, B., Delfino, M., Kulisevsky, J., Gómez-Ansón, B.,
-2013. Pattern of Regional Cortical Thinning Associated with Cognitive
-Deterioration in Parkinson's Disease. PLoS ONE 8, e54980.
-https://doi.org/10.1371/journal.pone.0054980}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Palomo-Buitrago, M.E.,
-Sabater-Masdeu, M., Moreno-Navarrete, J.M., Caballano-Infantes, E.,
-Arnoriaga-Rodríguez, M., Coll, C., Ramió, L., Palomino-Schätzlein, M.,
-Gutiérrez-Carcedo, P., Pérez-Brocal, V., Simó, R., Moya, A., Ricart, W.,
-Herance, J.R., Fernández-Real, J.M., 2019. Glutamate interactions with
-obesity, insulin resistance, cognition and gut microbiota composition.
-Acta Diabetol. 56, 569--579. https://doi.org/10.1007/s00592-019-01313-w}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Pronovost, G.N.,
-Hsiao, E.Y., 2019. Perinatal interactions between the microbiome,
-immunity, and neurodevelopment. Immunity 50, 18--36.
-https://doi.org/10.1016/j.immuni.2018.11.016}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sadeghi, B.,
-Chiarawongse, P., Squire, K., Jones, D.C., Noack, A., St-Jean, C.,
-Huijzer, R., Schätzle, R., Butterworth, I., Peng, Y.-F., Blaom, A.,
-2022. DecisionTree.jl - A julia implementation of the CART decision tree
-and random forest algorithms. https://doi.org/10.5281/zenodo.7359268}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sela, D.A., Chapman,
-J., Adeuya, A., Kim, J.H., Chen, F., Whitehead, T.R., Lapidus, A.,
-Rokhsar, D.S., Lebrilla, C.B., German, J.B., Price, N.P., Richardson,
-P.M., Mills, D.A., 2008. The genome sequence of \emph{Bifidobacterium
-longum} subsp. \emph{infantis} reveals adaptations for milk utilization
-within the infant microbiome. Proc. Natl. Acad. Sci. 105, 18964--18969.
-https://doi.org/10.1073/pnas.0809584105}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Selma, M.V., Beltrán,
-D., García-Villalba, R., Espín, J.C., Tomás-Barberán, F.A., 2014.
-Description of urolithin production capacity from ellagic acid of two
-human intestinal Gordonibacter species. Food Funct. 5, 1779--1784.
-https://doi.org/10.1039/c4fo00092g}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Setchell, K.D.,
-Clerici, C., Lephart, E.D., Cole, S.J., Heenan, C., Castellani, D.,
-Wolfe, B.E., Nechemias-Zimmer, L., Brown, N.M., Lund, T.D., Handa, R.J.,
-Heubi, J.E., 2005. S-Equol, a potent ligand for estrogen receptor β, is
-the exclusive enantiomeric form of the soy isoflavone metabolite
-produced by human intestinal bacterial flora. Am. J. Clin. Nutr. 81,
-1072--1079. https://doi.org/10.1093/ajcn/81.5.1072}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sharon, G., Sampson,
-T.R., Geschwind, D.H., Mazmanian, S.K., 2016. The central nervous system
-and the gut microbiome. Cell 167, 915--932.
-https://doi.org/10.1016/j.cell.2016.10.027}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Silbereis, J.C.,
-Pochareddy, S., Zhu, Y., Li, M., Sestan, N., 2016. The cellular and
-molecular landscapes of the developing human central nervous system.
-Neuron 89, 248--268.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Sordillo, J.E.,
-Korrick, S., Laranjo, N., Carey, V., Weinstock, G.M., Gold, D.R.,
-O'Connor, G., Sandel, M., Bacharier, L.B., Beigelman, A., Zeiger, R.,
-Litonjua, A.A., Weiss, S.T., 2019. Association of the infant gut
-microbiome with early childhood neurodevelopmental outcomes: An
-ancillary study to the VDAART randomized clinical trial. JAMA Netw. Open
-2, e190905. https://doi.org/10.1001/jamanetworkopen.2019.0905}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Spichak, S.,
-Bastiaanssen, T.F.S., Berding, K., Vlckova, K., Clarke, G., Dinan, T.G.,
-Cryan, J.F., 2021. Mining microbes for mental health: Determining the
-role of microbial metabolic pathways in human brain health and disease.
-Neurosci. Biobehav. Rev. 125, 698--761.
-https://doi.org/10.1016/j.neubiorev.2021.02.044}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Stephan, J., Stegle,
-O., Beyer, A., 2015. A random forest approach to capture genetic effects
-in the presence of population structure. Nat. Commun. 6, 7432.
-https://doi.org/10.1038/ncomms8432}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Suzek, B.E., Huang,
-H., McGarvey, P., Mazumder, R., Wu, C.H., 2007. UniRef: comprehensive
-and non-redundant UniProt reference clusters. Bioinformatics 23,
-1282--1288. https://doi.org/10.1093/bioinformatics/btm098}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Takahashi, H., Yang,
-J., Yamamoto, H., Fukuda, S., Arakawa, K., 2021. Complete Genome
-Sequence of Adlercreutzia equolifaciens subsp. celatus DSM 18785.
-Microbiol. Resour. Announc. 10, e00354-21.
-https://doi.org/10.1128/MRA.00354-21}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tau, G.Z., Peterson,
-B.S., 2010. Normal development of brain circuits. Neuropsychopharmacol.
-Off. Publ. Am. Coll. Neuropsychopharmacol. 35, 147--168.
-https://doi.org/10.1038/npp.2009.115}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Thawornkuno, C.,
-Tanaka, M., Sone, T., Asano, K., 2009. Biotransformation of Daidzein to
-Equol by Crude Enzyme from Asaccharobacter celatus AHU1763 Required an
-Anaerobic Environment. Biosci. Biotechnol. Biochem. 73, 1435--1438.
-https://doi.org/10.1271/bbb.80908}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tognini, P., 2017. Gut
-microbiota: a potential regulator of neurodevelopment. Front. Cell.
-Neurosci. 11, 25. https://doi.org/10.3389/fncel.2017.00025}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Tso, L., Bonham, K.S.,
-Fishbein, A., Rowland, S., Klepac-Ceraj, V., 2021. Targeted
-High-Resolution taxonomic identification of bifidobacterium longum
-subsp. infantis using human milk oligosaccharide metabolizing genes.
-Nutrients 13, 2833. https://doi.org/10.3390/nu13082833}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Valles-Colomer, M.,
-Falony, G., Darzi, Y., Tigchelaar, E.F., Wang, J., Tito, R.Y., Schiweck,
-C., Kurilshikov, A., Joossens, M., Wijmenga, C., Claes, S., Van
-Oudenhove, L., Zhernakova, A., Vieira-Silva, S., Raes, J., 2019. The
-neuroactive potential of the human gut microbiota in quality of life and
-depression. Nat. Microbiol. 4, 623--632.
-https://doi.org/10.1038/s41564-018-0337-x}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Velazquez, M., Lee,
-Y., Initiative, for the A.D.N., 2021. Random forest model for
-feature-based Alzheimer's disease conversion prediction from early mild
-cognitive impairment subjects. PLOS ONE 16, e0244773.
-https://doi.org/10.1371/journal.pone.0244773}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wacker, J., Dillon,
-D.G., Pizzagalli, D.A., 2009. The role of the nucleus accumbens and
-rostral anterior cingulate cortex in anhedonia: Integration of resting
-EEG, fMRI, and volumetric techniques. NeuroImage 46, 327--337.
-https://doi.org/10.1016/j.neuroimage.2009.01.058}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wan, Y., Zuo, T., Xu,
-Z., Zhang, F., Zhan, H., Chan, D., Leung, T.-F., Yeoh, Y.K., Chan,
-F.K.L., Chan, R., Ng, S.C., 2021. Underdevelopment of the gut microbiota
-and bacteria species as non-invasive markers of prediction in children
-with autism spectrum disorder. Gut gutjnl-2020-324015.
-https://doi.org/10.1136/gutjnl-2020-324015}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, L.-J., Yang,
-C.-Y., Chou, W.-J., Lee, M.-J., Chou, M.-C., Kuo, H.-C., Yeh, Y.-M.,
-Lee, S.-Y., Huang, L.-H., Li, S.-C., 2020. Gut microbiota and dietary
-patterns in children with attention-deficit/hyperactivity disorder. Eur.
-Child Adolesc. Psychiatry 29, 287--297.
-https://doi.org/10.1007/s00787-019-01352-2}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wang, X.-L., Hur,
-H.-G., Lee, J.H., Kim, K.T., Kim, S.-I., 2005. Enantioselective
-Synthesis of S-Equol from Dihydrodaidzein by a Newly Isolated Anaerobic
-Human Intestinal Bacterium. Appl. Environ. Microbiol. 71, 214--219.
-https://doi.org/10.1128/AEM.71.1.214-219.2005}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 2012.
-Wechsler preschool and primary scale of intelligence---fourth edition.
-Psychol. Corp. San Antonio TX.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Wechsler, D., 1949.
-Wechsler intelligence scale for children.}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, L., Yan, J.,
-Yang, T., Zhu, J., Li, T., Wei, H., Chen, J., 2021. Fecal Microbiome
-Transplantation from Children with Autism Spectrum Disorder Modulates
-Tryptophan and Serotonergic Synapse Metabolism and Induces Altered
-Behaviors in Germ-Free Mice. mSystems 6, e01343-20.
-https://doi.org/10.1128/mSystems.01343-20}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Xiao, X., Yang, Y.,
-Zhang, Y., Zhang, X.-M., Zhao, Z.-Q., Zhang, Y.-Q., 2013. Estrogen in
-the Anterior Cingulate Cortex Contributes to Pain-Related Aversion.
-Cereb. Cortex 23, 2190--2203. https://doi.org/10.1093/cercor/bhs201}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Yael, D., Tahary, O.,
-Gurovich, B., Belelovsky, K., Bar-Gad, I., 2019. Disinhibition of the
-Nucleus Accumbens Leads to Macro-Scale Hyperactivity Consisting of
-Micro-Scale Behavioral Segments Encoded by Striatal Activity. J.
-Neurosci. 39, 5897--5909.
-https://doi.org/10.1523/JNEUROSCI.3120-18.2019}
-
-\href{https://www.zotero.org/google-docs/?uydJ8e}{Yau, W.-Y.W., Zubieta,
-J.-K., Weiland, B.J., Samudra, P.G., Zucker, R.A., Heitzeg, M.M., 2012.
-Nucleus Accumbens Response to Incentive Stimuli Anticipation in Children
-of Alcoholics: Relationships with Precursive Behavioral Risk and
-Lifetime Alcohol Use. J. Neurosci. 32, 2544--2551.
-https://doi.org/10.1523/JNEUROSCI.1390-11.2012}
-
-Supplementary Tables
-
-\textbf{Supplementary Table XXA. List of most important taxa for Random
-Forest models on the 0 to 6 months age bracket, microbiome +
-demographics, ordered by relative weighted importance rank}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Taxon
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Rank
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Relative weighted importance
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Cumulative
-
-relative weighted importance
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\textbf{Supplementary Table XXA. List of most important taxa for Random
-Forest models on the 18 to 120 months age bracket, microbiome +
-demographics, ordered by relative weighted importance rank}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.5334}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1599}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1533}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Taxon
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Rank
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Relative weighted importance
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Cumulative
-
-relative weighted importance
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\textbf{Supplementary Table XX. Benchmark metrics for each brain segment
-prediction experiment. Confidence intervals are calculated from the
-distribution of metrics from repeated CV at a confidence level of 95\%}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3933}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2440}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1846}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1782}}@{}}
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Brain segment
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Root-mean-square Error
-
-(RMSE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Mean Absolute Proportional Error (MAPE)
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Correlation coefficient (R)
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-Brain-stem
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0876
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.605
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-I-V
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0321
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.085
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.156
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-VI-VII
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0132
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.081
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.006
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-cerebellar-vermal-lobules-VIII-X
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0217
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.078
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-CSF
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0193
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.162
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.071
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-accumbens-area
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0020
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.094
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.288
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-accumbens-area
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0020
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.102
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.041
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-amygdala
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0037
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.288
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-amygdala
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.070
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.179
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-anterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0375
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.069
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.092
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-anterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0264
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.074
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.245
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-basal-forebrain
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0008
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-basal-forebrain
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0010
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.117
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.123
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-caudate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0257
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.099
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.029
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-caudate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0221
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.010
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-cerebellum-white-matter
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0689
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.413
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-cerebellum-white-matter
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0694
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.070
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.434
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-cuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0274
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.234
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-cuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0264
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.075
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.065
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-entorhinal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0107
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.080
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.190
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-entorhinal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0105
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.099
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-fusiform
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0510
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.333
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-fusiform
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0334
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.053
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.248
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-hippocampus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0128
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.054
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.075
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-hippocampus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0150
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.094
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-insula
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0294
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.119
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-insula
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0303
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-isthmus-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0161
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.077
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.041
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-isthmus-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0166
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.085
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.074
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-lateral-occipital
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0661
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.062
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.224
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-lateral-occipital
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0716
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.054
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.221
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-lingual
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0410
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.069
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.421
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-lingual
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0440
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.072
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.434
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-middle-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0875
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.052
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.000
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-middle-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0636
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.039
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.083
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-orbitofrontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0715
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.055
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.186
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-orbitofrontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0814
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.133
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pallidum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0075
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.064
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.297
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pallidum
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0081
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.264
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-paracentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0335
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.078
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.131
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-paracentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0364
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.084
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.023
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-parahippocampal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0098
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.058
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-parahippocampal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0092
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.060
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-parietal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0755
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.035
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.177
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-parietal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1099
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.045
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.147
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-opercularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0197
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.057
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.146
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-opercularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0261
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.060
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.030
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-orbitalis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0157
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.082
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.003
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-orbitalis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0233
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.112
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.024
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pars-triangularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0226
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.174
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pars-triangularis
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0322
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.076
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.043
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-pericalcarine
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0178
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.111
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.200
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-pericalcarine
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0185
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.089
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.273
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-postcentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0495
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.052
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.168
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-postcentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0552
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.135
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-posterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0460
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.134
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.137
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-posterior-cingulate
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0480
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.139
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.063
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-precentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0830
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.065
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.248
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-precentral
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0594
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.046
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.066
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-precuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0489
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.056
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.144
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-precuneus
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0622
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.150
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-putamen
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0243
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.059
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.154
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-putamen
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0242
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.066
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.011
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-superior-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0867
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.040
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.101
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-superior-frontal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1206
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.044
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.096
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-supramarginal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0767
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.067
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.061
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-supramarginal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0734
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
--0.001
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-temporal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1530
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.040
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.128
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-temporal
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1503
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.039
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.068
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-left-thalamus-proper
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0347
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.053
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.170
-\end{minipage} \\
-\begin{minipage}[b]{\linewidth}\raggedright
-right-thalamus-proper
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0304
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.050
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.162
-\end{minipage} \\
-\midrule()
-\endhead
-\bottomrule()
-\end{longtable}
-
-\end{document}

### 2023-02-17 manuscript/tex/sections/intro.tex
start working on refs

@@
 nerve and through endocrine and immune (cytokine) signaling molecules
-\href{https://www.zotero.org/google-docs/?i3SRPn}{(Cerdó et al., 2019;
-Pronovost and Hsiao, 2019; Sharon et al., 2016; Tognini, 2017)}. In
-addition, the products of microbial metabolism generated in the gut can
+\cite{cerdoEarlyNutritionGut2019,pronovostPerinatalInteractionsMicrobiome2019,sharonCentralNervousSystem2016,togniniGutMicrobiotaPotential2017}
+In addition, the products of microbial metabolism generated in the gut can
 influence the brain, both indirectly by stimulating the enteric nervous
@@
 gut microbiome and neural development, particularly atypical
-development, are increasingly being identified
-\href{https://www.zotero.org/google-docs/?UsrFKB}{(Spichak et al.,
-2021)}. Both human epidemiology and animal models point to the effects
+development, are increasingly being identified \cite{spichakMiningMicrobesMental2021}.
+Both human epidemiology and animal models point to the effects
 of gut microbes on the development of autism spectrum disorder
-\href{https://www.zotero.org/google-docs/?Y8l5A0}{(Laue et al., 2020;
-Wan et al., 2021)} and particular microbial taxa have been associated
-with depression
-\href{https://www.zotero.org/google-docs/?mbaFE3}{(Mayneris-Perxachs et
-al., 2022; Valles-Colomer et al., 2019)} and Alzheimer's disease
-\href{https://www.zotero.org/google-docs/?0BEGUn}{(Fung et al., 2017;
-Kim et al., 2021)}. But information about this ``microbiome-gut-brain
-axis'' in normal neurocognitive development remains lacking,
+\cite{laueProspectiveAssociationsInfant2020,wanUnderdevelopmentGutMicrobiota2021}
+and particular microbial taxa have been associated with depression
+\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,valles-colomerNeuroactivePotentialHuman2019}
+and Alzheimer's disease \cite{fungInteractionsMicrobiotaImmune2017,kimProbioticSupplementationImproves2021}.
+But information about this "microbiome-gut-brain
+axis" in normal neurocognitive development remains lacking,
 particularly early in life.

### 2023-02-17 manuscript/tex/sections/intro.tex
citations in intro

@@
 gut microbiome and neural development, particularly atypical
-development, are increasingly being identified \cite{spichakMiningMicrobesMental2021}.
+development, are increasingly being identified
+\cite{spichakMiningMicrobesMental2021}.
 Both human epidemiology and animal models point to the effects
@@
 \cite{mayneris-perxachsMicrobiotaAlterationsProline2022,valles-colomerNeuroactivePotentialHuman2019}
-and Alzheimer's disease \cite{fungInteractionsMicrobiotaImmune2017,kimProbioticSupplementationImproves2021}.
+and Alzheimer's disease
+\cite{fungInteractionsMicrobiotaImmune2017,kimProbioticSupplementationImproves2021}.
 But information about this "microbiome-gut-brain
@@
 microbiome and the brain
-\href{https://www.zotero.org/google-docs/?XTqhlm}{(Laue et al., 2022)}.
+\cite{laueDevelopingMicrobiomeBirth2022}.
 Fetal development \emph{in utero} is believed to be mostly sterile, but
@@
 and other environmental sources such as antibiotics
-\href{https://www.zotero.org/google-docs/?X6N0f4}{(Bäckhed et al., 2015;
-Bokulich et al., 2016; Dominguez-Bello et al., 2010; Louwies et al.,
-2020)}. The early microbiome is characterized by low microbial
+\cite{backhedDynamicsStabilizationHuman2015,bokulichAntibioticsBirthMode2016}.
+The early microbiome is characterized by low microbial
 diversity, rapid succession and evolution, and domination by
@@
 Bacteroidetes, especially \emph{Bacteroides}, and Proteobacteria
-\href{https://www.zotero.org/google-docs/?VG43uh}{(Koenig et al.,
-2011)}. Many of these microbes have specialized metabolic capabilities
+\cite{koenigSuccessionMicrobialConsortia2011}.
+Many of these microbes have specialized metabolic capabilities
 for digesting human breast milk, such as \emph{Bifidobacterium infantis}
 and \emph{Bacteroides fragilis}
-\href{https://www.zotero.org/google-docs/?1dKzKj}{(Sela et al., 2008;
-Tso et al., 2021)}. Upon the introduction of solid foods, the gut
+\cite{selaGenomeSequenceBifidobacterium2008}.
+Upon the introduction of solid foods, the gut
 microbiome undergoes another categorical transformation, with most taxa
@@
 adult microbiomes
-\href{https://www.zotero.org/google-docs/?V3lY3p}{(Bäckhed et al.,
-2015)}. Many prior studies have focused on either infant microbiomes or
+\cite{backhedDynamicsStabilizationHuman2015}.
+Many prior studies have focused on either infant microbiomes or
 adult microbiomes, since performing statistical analyses across this
@@
 incredibly important
-\href{https://www.zotero.org/google-docs/?d29bv8}{(Tau and Peterson,
-2010)}.
+\cite{tauNormalDevelopmentBrain2010}.
 
@@
 connections has been established
-\href{https://www.zotero.org/google-docs/?1jUZOO}{(Silbereis et al.,
-2016)}. Much of this development occurs in discrete windows or
+\cite{silbereisCellularMolecularLandscapes2016}.
+Much of this development occurs in discrete windows or
 ``sensitive-'' or ``critical periods'' (CPs)
-\href{https://www.zotero.org/google-docs/?w4Pcze}{(Knudsen, 2004)} when
-neural plasticity is particularly high, and particular modes of learning
+\cite{knudsenSensitivePeriodsDevelopment2004}
+when neural plasticity is particularly high,
+and particular modes of learning
 and skill development are preferred. The timing of these sensitive
@@
 environment, including the gut microbiome
-\href{https://www.zotero.org/google-docs/?KTm9lQ}{(Cowan et al., 2020)}.
+\cite{cowanAnnualResearchReview2020}.
 In fact, emerging evidence suggests that the timing and duration of CPs
 may be driven in part by cues from the developing gut microbiome
-\href{https://www.zotero.org/google-docs/?mAD1ne}{(Callaghan, 2020)}. As
-such, understanding the normal spectrum of healthy microbiome
+\cite{callaghanNestedSensitivePeriods2020}.
+As such,
+understanding the normal spectrum of healthy microbiome
 development and how it relates to normal neurocognitive development may

### 2023-02-17 manuscript/tex/sections/results.tex
citations in results

@@
 cognitive and behavioral assessments, and neuroimaging. In all, 361
-children from RESONANCE, an accelerated longitudinal cohort of early
-development \href{https://paperpile.com/c/dPbU4e/klxU}{(Forrest,
-Blackwell, and Camargo 2018)} between the ages of 82 days and 10 years
+children from RESONANCE,
+an accelerated longitudinal cohort of early development
+\cite{forrestAdvancingScienceChildren2018}
+between the ages of 82 days and 10 years
 old (median age 2.21 years, Supplementary Figure 1) were included in
@@
 for children from birth to 3 years of age
-\href{https://www.zotero.org/google-docs/?90hfYr}{(Mullen and others,
-1995)}, Wechsler preschool and primary scale of intelligence (WPPSI) for
-4--5-year-olds
-\href{https://www.zotero.org/google-docs/?wCfx5h}{(Wechsler, 2012)}, and
-the Wechsler intelligence scale for children (WISC) for children 6 years
-and up \href{https://www.zotero.org/google-docs/?AcNN3S}{(Wechsler,
-1949)}.
+\cite{mullenMullenScalesEarly1995},
+Wechsler preschool and primary scale of intelligence (WPPSI)
+for 4-5 year-olds
+\cite{wechslerWechslerPreschoolPrimary2012},
+and the Wechsler intelligence scale for children (WISC)
+for children 6 years and up
+\cite{wechslerWechslerIntelligenceScale1949}.
 
@@
 measures of anxiety and depression
-\href{https://www.zotero.org/google-docs/?X5rfKj}{(Mayneris-Perxachs et
-al., 2022; Needham et al., 2022)}, cognitive flexibility in adults
-\href{https://www.zotero.org/google-docs/?RGIf8m}{(Magnusson et al.,
-2015)}, and atypical neural development
-\href{https://www.zotero.org/google-docs/?hwPZKh}{(Liu et al., 2019; Wan
-et al., 2021)}. We, therefore, set out to identify whether specific taxa
+\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022},
+cognitive flexibility in adults
+\cite{magnussonRelationshipsDietrelatedChanges2015},
+and atypical neural development
+\cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021}.
+We, therefore, set out to identify whether specific taxa
 or gene functions were linked with normal cognitive development in
@@
 profiles (annotated with clusters of 90\% similarity
-\href{https://www.zotero.org/google-docs/?b9rEJS}{(Suzek et al.,
-2007)}), both generated from DNA sequencing data, being tightly coupled
+\cite{suzekUniRefComprehensiveNonredundant2007}
+), both generated from DNA sequencing data, being tightly coupled
 (Figure 1H, p \textless{} 0.001). Interestingly, other functional
 groupings (Enzyme commission level4 - ECs
-\href{https://www.zotero.org/google-docs/?fs8LO7}{(Bairoch, 2000)}, and
+\cite{bairochENZYMEDatabase20002000},
+and
 KEGG orthologues - KOs
-\href{https://www.zotero.org/google-docs/?Xx5duE}{(Kanehisa et al.,
-2004)}) overlapped only slightly with taxonomic profiles in both age
+\cite{kanehisaKEGGResourceDeciphering2004}
+) overlapped only slightly with taxonomic profiles in both age
 cohorts, despite being derived from UniRef90 labels. In children over 18
@@
 cognitive function, we fit multivariable linear regression
-\href{https://www.zotero.org/google-docs/?4qS6UF}{(Mallick et al.,
-2021)} to the relative abundance of each species that had at least 15\%
+\cite{mallickMultivariableAssociationDiscovery2021}
+to the relative abundance of each species that had at least 15\%
 prevalence in a given age group (Figure 2A, N = 92 for 0--120 months, N
@@
 previously shown to be depleted in children with diabetes
-\href{https://www.zotero.org/google-docs/?6mU2nI}{(Benítez-Páez et al.,
-2020)}, and that oral administration of \emph{B. wexlerae} partially
+\cite{benitez-paezDepletionBlautiaSpecies2020},
+and that oral administration of \emph{B. wexlerae} partially
 ameliorated weight gain and inflammation from a high-fat diet in a mouse
-model of T2D \href{https://www.zotero.org/google-docs/?y83hpS}{(Hosomi
-et al., 2022; Liu et al., 2021)}. In children over 18 months of age,
+model of T2D
+\cite{hosomiOralAdministrationBlautia2022,liuBlautiaNewFunctional2021}.
+In children over 18 months of age,
 several microbial species were significantly enriched (q-value
@@
 neuroprotective metabolite urolithin
-\href{https://www.zotero.org/google-docs/?HGtTEY}{(Gong et al., 2022;
-Selma et al., 2014)}, \emph{Asaccharobacter celatus} and
+\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014},
+\emph{Asaccharobacter celatus} and
 \emph{Adelcreutzia equolifaciens}, which produce phytoestrogen-derived
-equol \href{https://www.zotero.org/google-docs/?HCKZqm}{(Maruo et al.,
-2008; Thawornkuno et al., 2009)}, and the SCFA-producing probiotic
+equol
+\cite{maruoAdlercreutziaEquolifaciensGen2008,thawornkunoBiotransformationDaidzeinEquol2009},
+and the SCFA-producing probiotic
 species such as \emph{Eubacterium eligens} and \emph{Faecalibacterium
-prausnitzii} \href{https://www.zotero.org/google-docs/?3eyheK}{(Ghosh et
-al., 2020; Lopez-Siles et al., 2017; Mukherjee et al., 2020)} (Figure
+prausnitzii}
+\cite{ghoshMediterraneanDietIntervention2020}
+(Figure
 2B).
@@
 of genes with neuroactive potential
-\href{https://www.zotero.org/google-docs/?xQURsj}{(Valles-Colomer et
-al., 2019)} and concurrent cognitive function score (Table 2,
+\cite{valles-colomerNeuroactivePotentialHuman2019}
+and concurrent cognitive function score (Table 2,
 \protect\hyperlink{_i7fd3z19jdvg}{\uline{Figure 2}}C-G) and found that
@@
 with brain function
-\href{https://www.zotero.org/google-docs/?2Z0sKZ}{(Dalile et al.,
-2019)}.
+\cite{dalileRoleShortchainFatty2019}.
 
@@
 E.S. = -0.170, q = 0.0183). Menaquinone has several isoforms, one of
-which, MK-4, has been found to be decreased in children diagnosed with
-ASD \href{https://www.zotero.org/google-docs/?dTA3NO}{(Dong et al.,
-2021)} and is potentially neuroprotective in both rodents and humans
-\href{https://paperpile.com/c/dPbU4e/cTO9}{(Elkattawy et al., 2022)}. In
-children over 18 months old, genes for the synthesis of the amino acids
+which, MK-4, has been found to be decreased in children diagnosed with ASD
+\cite{dongCorrelationSerumConcentrations2021}
+and is potentially neuroprotective in both rodents and humans
+\cite{elkattawyVitaminK2Menaquinone72022}.
+In children over 18 months old,
+genes for the synthesis of the amino acids
 glutamate and tryptophan were significantly enriched in children with
@@
 the developing brain
-\href{https://www.zotero.org/google-docs/?QubMhx}{(Cohen Kadosh et al.,
-2015; Palomo-Buitrago et al., 2019)}. Tryptophan metabolism, including
+\cite{cohenkadoshLinkingGABAGlutamate2015,palomo-buitragoGlutamateInteractionsObesity2019}.
+Tryptophan metabolism, including
 microbial metabolism of tryptophan, has previously been linked with
 autism in children
-\href{https://www.zotero.org/google-docs/?GPaUZE}{(Hoshino et al., 1984;
-Xiao et al., 2021)}. Taken together, these results suggest that
+\cite{hoshinoBloodSerotoninFree1984,xiaoFecalMicrobiomeTransplantation2021}.
+Taken together, these results suggest that
 microbial metabolic activity, particularly the metabolism (synthesis and
@@
 from the domain of genomics
-\href{https://www.zotero.org/google-docs/?C3CSPB}{(Amaratunga et al.,
-2008; Brieuc et al., 2018; Chen and Ishwaran, 2012; Franzosa et al.,
-2019; Stephan et al., 2015)}, along with other works where it was used
+\cite{amaratungaEnrichedRandomForests2008,brieucPracticalIntroductionRandom2018,chenRandomForestsGenomic2012,franzosaGutMicrobiomeStructure2019,stephanRandomForestApproach2015},
+along with other works where it was used
 to predict cognitive conditions related to Alzheimer's disease in
 different scenarios
-\href{https://www.zotero.org/google-docs/?GiadGy}{(Ardekani et al.,
-2017; Velazquez et al., 2021)}. Additionally, given the sequential
+\cite{ardekaniPredictionIncipientAlzheimer2017,velazquezRandomForestModel2021}.
+Additionally, given the sequential
 nature of variable consideration in each tree, RFs are naturally able to
@@
 area, which plays an important role in reward circuits
-\href{https://www.zotero.org/google-docs/?zVicU7}{(Ernst et al., 2005;
-Yau et al., 2012)}, has one of the highest test-set correlations of our
+\cite{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012},
+has one of the highest test-set correlations of our
 brain region models (R = 0.288), as compared to the right accumbens
@@
 been linked to substance use disorder in young adults
-\href{https://www.zotero.org/google-docs/?V45KyN}{(Cao et al., 2021)}.
+\cite{caoMappingCorticalSubcortical2021}.
+
 Feature importance for models of the left accumbens area were dominated
@@
 it has been shown to have a divergent neuromorphology
-\href{https://www.zotero.org/google-docs/?kLfEQ5}{(Hoogman et al.,
-2017)}. Independently, \emph{B. ovatus}, \emph{B. uniformis} and
+\cite{hoogmanSubcorticalBrainVolume2017}.
+Independently, \emph{B. ovatus}, \emph{B. uniformis} and
 \emph{B. vulgatus} have been linked to ADHD
-\href{https://www.zotero.org/google-docs/?luO0zh}{(Wang et al., 2020)}.
+\cite{wangGutMicrobiotaDietary2020}.
+
 In fact, there are studies showing that alterations on the striatal
@@
 attention deficit
-\href{https://www.zotero.org/google-docs/?UxS47e}{(Yael et al., 2019)},
+\cite{yaelDisinhibitionNucleusAccumbens2019},
 and that \emph{B. uniformis} is gut-microbial modulator of the brain
 dopamine transporter
-\href{https://www.zotero.org/google-docs/?L1cWJS}{(Hartstra et al.,
-2020)}.
+\cite{hartstraInfusionDonorFeces2020}.
 
@@
 associated with cognitive function
-\href{https://www.zotero.org/google-docs/?AVJiwu}{(Kant et al., 2015; Li
-et al., 2022)}. Its importance is, however, heavily loaded on the volume
+\cite{kantGenomeSequenceButyrateProducing2015,liCorrelationsGutMicrobiota2022}.
+Its importance is, however, heavily loaded on the volume
 of entire major lobes (frontal, parietal), which cannot be readily
@@
 parahippocampal gyrus
-\href{https://www.zotero.org/google-docs/?rlO6eq}{(Pagonabarraga et al.,
-2013)}. Additionally, the left and right parahippocampal, where \emph{B.
+\cite{pagonabarragaPatternRegionalCortical2013}.
+Additionally, the left and right parahippocampal, where \emph{B.
 wexlerae} had the highest importance in both models, are important in
 visual/spatial processing and memory
-\href{https://www.zotero.org/google-docs/?wkkWZT}{(Aminoff et al.,
-2013)}. \emph{B. wexlerae} can also produce acetylcholine in the gut
-\href{https://www.zotero.org/google-docs/?jBoxSm}{(Hosomi et al.,
-2022)}, and this molecule plays an important role in modulating memory
-function \href{https://www.zotero.org/google-docs/?BQ0j2c}{(Haam and
-Yakel, 2017)}.
+\cite{aminoffRoleParahippocampalCortex2013}.
+\emph{B. wexlerae} can also produce acetylcholine in the gut
+\cite{hosomiOralAdministrationBlautia2022},
+and this molecule plays an important role in modulating memory
+function
+\cite{haamCholinergicModulationHippocampal2017}.
 
@@
 depression symptoms
-\href{https://www.zotero.org/google-docs/?2aSZIr}{(Wacker et al.,
-2009)}, and increased levels of the \emph{Alistipes} genus have been
+\cite{wackerRoleNucleusAccumbens2009},
+and increased levels of the \emph{Alistipes} genus have been
 observed in patients with depressive disorder
-\href{https://www.zotero.org/google-docs/?bhhSZB}{(Jiang et al., 2015)}.
+\cite{jiangAlteredFecalMicrobiota2015}.
+
 While these reports are important, simultaneously probing of the gut
@@
 importances), which has been linked to social cognition and reward-based
-decision making \href{https://www.zotero.org/google-docs/?Wi8Bpq}{(Apps
-et al., 2016; Boes et al., 2008; Bush et al., 2002)}. Equol has a strong
+decision making
+\cite{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002}.
+Equol has a strong
 estrogenic effect
-\href{https://www.zotero.org/google-docs/?PqgFU7}{(Setchell et al.,
-2005)}, and in anterior cingulate, estrogen, has been shown to regulate
+\cite{setchellSEquolPotentLigand2005},
+and in anterior cingulate, estrogen, has been shown to regulate
 pain-related aversion
-\href{https://www.zotero.org/google-docs/?5huRIR}{(Xiao et al., 2013)}.
+\cite{xiaoEstrogenAnteriorCingulate2013}.
+
 Another example of this pattern is \emph{R. gnavus,} which was the only
@@
 resistance and obesity
-\href{https://www.zotero.org/google-docs/?lwzWUv}{(Ley et al., 2006)},
+\cite{leyHumanGutMicrobes2006},
+
 conditions that are known to produce structural abnormalities in the
-brain \href{https://www.zotero.org/google-docs/?NlkFr6}{(Opel et al.,
-2021)}. Finally, \emph{Coprococcus comes} displays an importance
+brain
+\cite{opelBrainStructuralAbnormalities2021}.
+Finally, \emph{Coprococcus comes} displays an importance
 distribution that splits almost evenly among the two previously reported

### 2023-02-17 manuscript/tex/sections/discussion.tex
citations in discussion

@@
 higher scores on the Ages and Stages Questionnaire
-\href{https://www.zotero.org/google-docs/?SB5Lwa}{(Sordillo et al.,
-2019)}. However, each of these clades encompass dozens of genera with
+\cite{sordilloAssociationInfantGut2019}.
+However, each of these clades encompass dozens of genera with
 diverse functions, each of which may have different effects. Indeed,
@@
 (formerly regarded a subspecies of \emph{A. equolofaciens}
-\href{https://www.zotero.org/google-docs/?OEhguI}{(Takahashi et al.,
-2021)}), and \emph{Eggerthella lenta}. Many members of this family are
+\cite{takahashiCompleteGenomeSequence2021}),
+and \emph{Eggerthella lenta}. Many members of this family are
 known in part due to unique metabolic activities. For example, \emph{A.
@@
 found in soybeans
-\href{https://www.zotero.org/google-docs/?I2qNyB}{(Wang et al., 2005)},
+\cite{wangEnantioselectiveSynthesisSEquol2005},
 and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
@@
 shown in some studies to have a neuroprotective effect
-\href{https://www.zotero.org/google-docs/?oTN8Lw}{(Gong et al., 2022;
-Selma et al., 2014)}. \emph{E. lenta} has been extensively studied for
+\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
+\emph{E. lenta} has been extensively studied for
 its ability to metabolize drug compounds such as the plant-derived heart
 medication digoxin
-\href{https://www.zotero.org/google-docs/?tH2Qeq}{(Haiser et al.,
-2013)}. The metabolic versatility of this clade, and the large number of
+\cite{haiserPredictingManipulatingCardiac2013}.
+The metabolic versatility of this clade, and the large number of
 species that are associated with cognition make these microbes prime
@@
 pronounced in some age groups due to our sampling schedule
-\href{https://www.zotero.org/google-docs/?L8GnRj}{(Blackwell et al.,
-2022; Deoni et al., 2021)} (Supplementary Figure 3).
+\cite{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
+(Supplementary Figure 3).
 

### 2023-02-17 manuscript/main.tex
get started on dataexport
reorg

@@
 
-\input{preamble.tex}
+\usepackage{lineno}
+\usepackage{titlesec}
+%\usepackage[none]{hyphenat} % use only if there is a problem
+% Use Unicode characters
+\usepackage[utf8]{inputenc}
+% Clean citations with biblatex
+\usepackage[
+backend=biber,
+natbib=true,
+sortcites=true,
+defernumbers=true,
+style=authoryear,
+citestyle=authoryear-comp,
+maxnames=99,
+maxcitenames=2,
+giveninits=true,
+terseinits=true,
+url=false
+]{biblatex}
+
+\addbibresource{zotero_library}
+
+% Highlighting
+\usepackage{color}
+\usepackage{soul}
+\usepackage{xurl}
+\usepackage{longtable,booktabs,array}
+\usepackage{multirow}
+\usepackage{graphicx}
 
@@
 \begin{abstract}
-\input{sections/abstract.tex}
+The gastrointestinal tract, its resident microorganisms and the central
+nervous system are connected by biochemical signaling, also known as
+"microbiome-gut-brain-axis." Both the human brain and the gut microbiome
+have critical developmental windows in the first three years of life,
+raising the possibility that their development is co-occurring and
+likely co-dependent. Emerging evidence implicates gut microorganisms and
+microbiota composition in cognitive outcomes and neurodevelopmental
+disorders (e.g., autism and anxiety), but the influence of gut microbial
+metabolism on typical neurodevelopment has not been explored in detail.
+We investigated the relationship of the microbiome with the neuroanatomy
+and cognitive function of 361 healthy children, demonstrating that
+differences in gut microbial taxa and gene functions are associated with
+overall cognitive function as well as with differences in the size of
+multiple brain regions. Using a combination of multivariate linear
+models and machine learning (ML) models, we showed that many species,
+including \emph{Gordonibacter pamelae} and \emph{Blautia wexlerae}, were
+significantly associated with higher cognitive function, while some
+species such as \emph{Ruminococcus gnavus} were more commonly found in
+children with low cognitive scores regardless of age or maternal
+education. Microbial genes for enzymes involved in the metabolism of
+neuroactive compounds, particularly short-chain fatty acids such as
+glutamate and propionate, were also associated with cognitive function.
+In addition, ML models were able to use microbial taxa to predict the
+volume of brain regions, with particular taxa often dominating the
+feature importance metric for individual brain regions, such as \emph{B.
+wexlerae} being the most dominant for the parahippocampal region or
+GABA-producing \emph{Bacteroidetes ovatus} for left accumbens. These
+findings provide potential biomarkers of neurocognition and may lead to
+the future development of targets for early detection and early
+intervention.
 \end{abstract}
@@
 
-\input{sections/intro}
+The gut and the brain are intimately linked. Signals from the brain
+reach the gut through the autonomic nervous system and the endocrine
+system, and the gut can communicate with the brain through the vagus
+nerve and through endocrine and immune (cytokine) signaling molecules
+\cite{cerdoEarlyNutritionGut2019,pronovostPerinatalInteractionsMicrobiome2019,sharonCentralNervousSystem2016,togniniGutMicrobiotaPotential2017}
+In addition, the products of microbial metabolism generated in the gut can
+influence the brain, both indirectly by stimulating the enteric nervous
+and immune systems, as well as directly through molecules that enter
+circulation and cross the blood-brain barrier. Causal links between the
+gut microbiome and neural development, particularly atypical
+development, are increasingly being identified
+\cite{spichakMiningMicrobesMental2021}.
+Both human epidemiology and animal models point to the effects
+of gut microbes on the development of autism spectrum disorder
+\cite{laueProspectiveAssociationsInfant2020,wanUnderdevelopmentGutMicrobiota2021}
+and particular microbial taxa have been associated with depression
+\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,valles-colomerNeuroactivePotentialHuman2019}
+and Alzheimer's disease
+\cite{fungInteractionsMicrobiotaImmune2017,kimProbioticSupplementationImproves2021}.
+But information about this "microbiome-gut-brain
+axis" in normal neurocognitive development remains lacking,
+particularly early in life.
+
+The first years of life are critical developmental windows for both the
+microbiome and the brain
+\cite{laueDevelopingMicrobiomeBirth2022}.
+Fetal development \emph{in utero} is believed to be mostly sterile, but
+is rapidly seeded at birth through contact with the birth canal (if
+birthed vaginally), caregivers, food sources (breastmilk or formula),
+and other environmental sources such as antibiotics
+\cite{backhedDynamicsStabilizationHuman2015,bokulichAntibioticsBirthMode2016}.
+The early microbiome is characterized by low microbial
+diversity, rapid succession and evolution, and domination by
+Actinobacteria, particularly the genus \emph{Bifidobacterium},
+Bacteroidetes, especially \emph{Bacteroides}, and Proteobacteria
+\cite{koenigSuccessionMicrobialConsortia2011}.
+Many of these microbes have specialized metabolic capabilities
+for digesting human breast milk, such as \emph{Bifidobacterium infantis}
+and \emph{Bacteroides fragilis}
+\cite{selaGenomeSequenceBifidobacterium2008}.
+Upon the introduction of solid foods, the gut
+microbiome undergoes another categorical transformation, with most taxa
+of the infant microbiome being replaced by taxa more reminiscent of
+adult microbiomes
+\cite{backhedDynamicsStabilizationHuman2015}.
+Many prior studies have focused on either infant microbiomes or
+adult microbiomes, since performing statistical analyses across this
+transition poses particular challenges. Nevertheless, since this
+transition coincides with critical neural developmental windows and
+neural synaptogenesis, investigation across this solid-food boundary is
+incredibly important
+\cite{tauNormalDevelopmentBrain2010}.
+
+A child's brain undergoes remarkable anatomical, microstructural,
+organizational, and functional changes in the first years of life. By
+age 5, a child's brain has reached \textgreater85\% of its adult size,
+has achieved near-adult levels of myelination, and the pattern of axonal
+connections has been established
+\cite{silbereisCellularMolecularLandscapes2016}.
+Much of this development occurs in discrete windows or
+``sensitive-'' or ``critical periods'' (CPs)
+\cite{knudsenSensitivePeriodsDevelopment2004}
+when neural plasticity is particularly high,
+and particular modes of learning
+and skill development are preferred. The timing of these sensitive
+periods is driven in part by genetics, but can also be affected by the
+environment, including the gut microbiome
+\cite{cowanAnnualResearchReview2020}.
+In fact, emerging evidence suggests that the timing and duration of CPs
+may be driven in part by cues from the developing gut microbiome
+\cite{callaghanNestedSensitivePeriods2020}.
+As such,
+understanding the normal spectrum of healthy microbiome
+development and how it relates to normal neurocognitive development may
+provide opportunities for identifying atypical development earlier and
+offer opportunities for intervention.
+
+To begin to address this need, we investigated the gut microbiome and
+neurocognitive development of children from infancy through 10 years of
+age in an accelerated longitudinal study design. Gut microbial
+communities were assessed using shotgun metagenomic sequencing, enabling
+profiling at both the taxonomic and gene-functional level, and
+neurocognitive development was measured using expert-assessment of
+cognitive function and neuroimaging using magnetic resonance imaging
+(MRI). Using a combination of classical statistical analysis and machine
+learning, we showed that the development of the gut microbiome, the
+brain, and children's cognitive abilities are intimately linked, with
+both microbial taxa and gene functions able to predict cognitive
+performance and brain structure.
+
+\section*{Results}
+
+\subsubsection{Cohort overview and summary data}
+
+We investigated the co-development of the brain and the microbiome in
+early childhood in a cohort of typically-developing children in their
+first years of life using a variety of orthogonal microbial and
+neurocognitive assessments, including shotgun metagenomic sequencing,
+cognitive and behavioral assessments, and neuroimaging. In all, 361
+children from RESONANCE,
+an accelerated longitudinal cohort of early development
+\cite{forrestAdvancingScienceChildren2018}
+between the ages of 82 days and 10 years
+old (median age 2.21 years, Supplementary Figure 1) were included in
+this study (Table 1, Figure 1A). To measure cognitive function, we used
+age-appropriate assessments that can be normalized to a common, IQ-like
+scale (Figure 1B), including the Mullen scale of early learning (MSEL)
+for children from birth to 3 years of age
+\cite{mullenMullenScalesEarly1995},
+Wechsler preschool and primary scale of intelligence (WPPSI)
+for 4-5 year-olds
+\cite{wechslerWechslerPreschoolPrimary2012},
+and the Wechsler intelligence scale for children (WISC)
+for children 6 years and up
+\cite{wechslerWechslerIntelligenceScale1949}.
+
+As expected, the greatest differences in microbial taxa were driven by
+age (Figure 1B-C, PCoA axis 1), while older children were primarily
+stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
+high-\emph{Prevotella copri} (Supplementary Figure 2). Overall variation
+in gut microbial genes and also in brain volume profiles was similarly
+driven largely by subject age (Figure 1C-F).
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{graph_a}
+    \caption{An example graph}
+    \label{fig:x cubed graph}
+\end{figure}
+
+(A) Study design and overview. Stool samples, cognitive assessments and
+neuroimaging were collected from participants in an accelerated
+longitudinal design. (B) Cognitive function scores are assessed by
+different instruments, but may be normalized to a common IQ-like scale.
+(C-D) Principal coordinates analysis using Bray-Curtis dissimilarity on
+taxonomic profiles demonstrates high beta diversity, with much of the
+first axis of variation explained by increasing age and alpha diversity.
+Differences in gene function profiles (E) and neuroimaging (PCA based on
+the Euclidean distance of brain region volumes) (F) are likewise
+dominated by changes as children age. (G) Permutation analysis of
+variance (PERMANOVA) of taxonomic profiles, functional profiles
+(annotated as UniRef90s) and neuroimaging, against the metadata of
+interest. Variations in taxonomic and functional profiles explain a
+modest but significant percent of the variation in cognitive development
+in children over 18 months of age. (H) Mantel testing of different
+microbial feature matrices, shows overlapping but distinct patterns of
+variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
+used as cut-offs in some following models.
+
+\textbf{Table1: Subject demographics}
+
+Several studies have demonstrated links between specific taxa and
+measures of anxiety and depression
+\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022},
+cognitive flexibility in adults
+\cite{magnussonRelationshipsDietrelatedChanges2015},
+and atypical neural development
+\cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021}.
+We, therefore, set out to identify whether specific taxa
+or gene functions were linked with normal cognitive development in
+children. In order to assess whether variations in gut microbial taxa,
+their genes, or their metabolism are linked with neurocognitive
+development, we tested whether the beta diversity of microbial taxa and
+gene functions, as well as variation in brain development as assessed by
+neuroimaging, were associated with these contemporaneous measures of
+cognitive function using permutation analysis of variance (PERMANOVA).
+Due to the large ecological shift in the microbiome that occurs upon the
+introduction of solid food, and a relatively wide range of ages when
+infants are transitioned to solid food, we considered the pre-transition
+(less than 6 months old) and post-transition (over 18 months old)
+microbiomes separately. Even still, age was a major driver of variation
+in gut microbiomes in children over 18 months old (Figure 1G). We also
+found that overall variation in microbial species in children over 18
+months old was associated with a small but significant variation in
+cognitive function score (Figure 1G, R\textsuperscript{2} = 0.0124, q
+\textless{} 0.001), as was variation in microbial gene functions
+(R\textsuperscript{2} = 0.0119, q \textless{} 0.001). Variation in
+microbial taxa and genes was not significantly associated with cognitive
+function in children under 6 months, though this may be due to the low
+taxonomic diversity, and broad lack of overlap between taxa in infants.
+As expected, age was significantly associated with microbial beta
+diversity (taxa R\textsuperscript{2} = 0.0215, and UniRef90s
+R\textsuperscript{2} = 0.0245, q \textless{} 0.001), and very strongly
+associated with neuroimaging profiles (R\textsuperscript{2} = 0.256, q
+\textless{} 0.001).
+
+Consistent with prior studies, different microbial measurement types
+captured overlapping variation, with species profiles and gene function
+profiles (annotated with clusters of 90\% similarity
+\cite{suzekUniRefComprehensiveNonredundant2007}
+), both generated from DNA sequencing data, being tightly coupled
+(Figure 1H, p \textless{} 0.001). Interestingly, other functional
+groupings (Enzyme commission level4 - ECs
+\cite{bairochENZYMEDatabase20002000},
+and
+KEGG orthologues - KOs
+\cite{kanehisaKEGGResourceDeciphering2004}
+) overlapped only slightly with taxonomic profiles in both age
+cohorts, despite being derived from UniRef90 labels. In children over 18
+months, some variation (15.9\%, p \textless{} 0.01) in neuroimaging
+overlapped with microbial measures, though this may be due to the
+residual variation due to age in both measures.
+
+\subsubsection{Microbial species and neuroactive genes are associated with cognitive performance}
+
+To assess whether individual microbial species were associated with
+cognitive function, we fit multivariable linear regression
+\cite{mallickMultivariableAssociationDiscovery2021}
+to the relative abundance of each species that had at least 15\%
+prevalence in a given age group (Figure 2A, N = 92 for 0--120 months, N
+= 46 for 0--6 months, N = 97 for 18--120 months). Only \emph{Blautia
+wexlerae} was significantly associated (q value = 0.14, β = 0.0015) with
+cognitive function in children under 6 months old after adjusting for
+age and maternal education (Figure 2B). \emph{B. wexlerae} was
+previously shown to be depleted in children with diabetes
+\cite{benitez-paezDepletionBlautiaSpecies2020},
+and that oral administration of \emph{B. wexlerae} partially
+ameliorated weight gain and inflammation from a high-fat diet in a mouse
+model of T2D
+\cite{hosomiOralAdministrationBlautia2022,liuBlautiaNewFunctional2021}.
+In children over 18 months of age,
+several microbial species were significantly enriched (q-value
+\textless{} 0.20) in children with higher cognitive function scores,
+including \emph{Gordonibacter pamelaeae}, which produces the
+neuroprotective metabolite urolithin
+\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014},
+\emph{Asaccharobacter celatus} and
+\emph{Adelcreutzia equolifaciens}, which produce phytoestrogen-derived
+equol
+\cite{maruoAdlercreutziaEquolifaciensGen2008,thawornkunoBiotransformationDaidzeinEquol2009},
+and the SCFA-producing probiotic
+species such as \emph{Eubacterium eligens} and \emph{Faecalibacterium
+prausnitzii}
+\cite{ghoshMediterraneanDietIntervention2020}
+(Figure
+2B).
+
+\includegraphics[width=6.5in,height=4.83333in]{media/image4.png}
+
+\textbf{Figure 2 - Taxa and gene functional groups are associated with cognitive function.}
+
+(A) Volcano plot of multivariable linear model results showing the
+relationship between individual taxa and cognitive function score in
+children over 18 months of age, controlling for age, maternal education,
+and sequencing depth. Taxa that were significant after FDR correction (q
+\textless{} 0.2) are colored red. (B) For taxa that were significantly
+associated with cognitive function, heatmaps of prevalence, mean
+relative abundance, and correlation with cognitive function in different
+age groups. (C-F) Enrichment plots for selected neuroactive gene sets
+used in feature set enrichment analysis (FSEA). (G) summary of FSEA
+results across all samples (left) as well as in the under 6 months
+(middle) and over 18 months (right) subsets, colored based on the
+significance of the association. Markers indicate the individual
+correlation of genes within a gene set, and vertical bars represent the
+median correlation of that gene set.
+
+Given that different microbial species might occupy the same metabolic
+niche in different individuals, we hypothesized that microbial genes
+grouped by functional activity would be associated with cognition. To
+test this, we performed feature set enrichment analysis (FSEA) on groups
+of genes with neuroactive potential
+\cite{valles-colomerNeuroactivePotentialHuman2019}
+and concurrent cognitive function score (Table 2,
+\protect\hyperlink{_i7fd3z19jdvg}{\uline{Figure 2}}C-G) and found that
+several metabolic pathways were significantly enriched or depleted in
+children with higher cognitive function scores. This was true both when
+considering all age groups together, though the enrichment of most
+pathways was more pronounced in children under 6 months or over 18
+months. For example, genes for degrading the 3-carbon SCFA propionate
+were significantly depleted in children with higher cognitive function
+scores across all age groups tested (Table 2; propionate degradation I,
+under 6 months, enrichment score (E.S.) = -0.542, corrected p-value (q)
+= 0.020; over 18 months, E.S. = -0.674, q = 0.041). Interestingly, genes
+for propionate synthesis were also significantly depleted in higher
+scoring children over 18 months (E.S. -0.676, q = 0.023), as were genes
+for synthesizing the 2 carbon SCFA acetate in children under 6 months
+old (acetate synthesis I, E.S. = -0.194, q = 0.153; acetate synthesis
+II, E.S. = -0.342, q = 0.020; acetate synthesis III, E.S. = -0.31, q =
+0.052). SCFAs are produced by anaerobic fermentation of dietary fiber
+and have been linked with immune system regulation as well as directly
+with brain function
+\cite{dalileRoleShortchainFatty2019}.
+
+Synthesis of menaquinone (vitamin K) was also negatively associated with
+cognitive function score in older children (menaquinone synthesis I,
+E.S. = -0.170, q = 0.0183). Menaquinone has several isoforms, one of
+which, MK-4, has been found to be decreased in children diagnosed with ASD
+\cite{dongCorrelationSerumConcentrations2021}
+and is potentially neuroprotective in both rodents and humans
+\cite{elkattawyVitaminK2Menaquinone72022}.
+In children over 18 months old,
+genes for the synthesis of the amino acids
+glutamate and tryptophan were significantly enriched in children with
+higher cognitive function scores (Glutamate synthesis I, E.S. = 0.242, q
+= 0.047; Tryptophan synthesis, E.S. = 0.119, q = 0.041). Glutamate is a
+critical neurotransmitter controlling neuronal excitatory/inhibitory
+signaling along with gamma-aminobutyric acid (GABA), and their balance
+in the brain controls neural plasticity and learning, particularly in
+the developing brain
+\cite{cohenkadoshLinkingGABAGlutamate2015,palomo-buitragoGlutamateInteractionsObesity2019}.
+Tryptophan metabolism, including
+microbial metabolism of tryptophan, has previously been linked with
+autism in children
+\cite{hoshinoBloodSerotoninFree1984,xiaoFecalMicrobiomeTransplantation2021}.
+Taken together, these results suggest that
+microbial metabolic activity, particularly the metabolism (synthesis and
+degradation) of neuroactive compounds may have effects on cognitive
+development.
+
+\textbf{Table 2 - Feature set enrichment analysis on neuroactive
+microbial gene sets}
+
+\subsubsection{Gut microbial taxonomic and functional profiles predict cognitive function}
+
+FSEA relies on understanding functional relationships between individual
+genes. However, because the relationships between individual taxa are
+still largely unknown, we turned to Random Forest (RF) models, an
+unsupervised non-parametric machine learning (ML) method that enables
+the identification of underlying patterns in large numbers of individual
+features (here, microbial species). Previous studies have reported
+successful use of RFs for processing highly-dimensional and sparse data
+from the domain of genomics
+\cite{amaratungaEnrichedRandomForests2008,brieucPracticalIntroductionRandom2018,chenRandomForestsGenomic2012,franzosaGutMicrobiomeStructure2019,stephanRandomForestApproach2015},
+along with other works where it was used
+to predict cognitive conditions related to Alzheimer's disease in
+different scenarios
+\cite{ardekaniPredictionIncipientAlzheimer2017,velazquezRandomForestModel2021}.
+Additionally, given the sequential
+nature of variable consideration in each tree, RFs are naturally able to
+work out complex input feature interactions, such as those present in a
+microbiome-wide study, without the necessity to explicitly compute
+interaction terms.
+
+\includegraphics[width=6.5in,height=5.19444in]{media/image2.png}
+
+\textbf{Figure 3 - Random forest models predict concurrently measured
+cognitive function.}
+
+Comparison of RF predictor importance versus linear models for children
+between birth and 6 months old (A), and for those older than 18 months
+(B). Colors represent whether the species belong to the group of
+top-important features that account for 80\% of the cumulative
+importance on the RF model, if that species was significant (q
+\textless{} 0.2) in linear models, both, or neither. Ranked feature
+importance for taxa in RF models for children between birth and 6 months
+old (C), and for those older than 18 months (D). Taxa that are important
+for RF models both for children under 6 months and for those over 18
+months (E), only for children under 6 months (F), and only for children
+over 18 months (G)
+
+Given that gut microbial profiles, as well as neurocognitive
+development, may partially reflect socioeconomic and demographic
+factors, we assessed the performance of RF regressors where maternal
+education (a proxy of socioeconomic status (SES)), sex, and age were
+included as possible predictors, either alone or in combination with
+microbial taxonomic profiles (Table 3).
+
+\textbf{Table 3. Benchmark metrics for the cognitive assessment score
+prediction models. Confidence intervals are calculated from the
+distribution of metrics from repeated CV at a confidence level of 95\%}
+
+As with linear models, RF models for children under 6 months old were
+less generalizable (mean test-set correlation -0.13, mean RMSE 12.70),
+but RFs were consistently able to learn the relationship between taxa
+and cognitive function scores in children over 18 months of age (mean
+test-set correlation 0.429, mean RMSE 17.29). For both age groups,
+species that were important in RF overlapped with those that were
+significant when testing the relationship with linear models (Figure
+3A-B, Supplementary Table XXA-B), though there were substantial
+differences. For children under 6 months, only \emph{B. wexlerae} was
+significantly associated with cognitive function in linear models, but
+was ranked 20th in importance for RF models. The importance metric
+employed (MDI) is a measure of how well the variable differentiates
+(splits) sets of samples by creating subgroups that reduce the
+intra-group deviations, while also accounting for the amount of samples
+affected by a split. All taxa significantly associated with cognitive
+function score in children over 18 months using LMs belong to the
+top-ranking group responsible for 80\% of the total relative importance,
+except for \emph{Eubacterium ramulus} (Table 3, Figure 3A).
+
+Interestingly, several taxa highly ranked in importance in both age
+groups, including several that were significant in LMs for children over
+18 months old, including \emph{R. gnavus} (0--6 months, rank = 13;
+18--120 months, rank = 7) and \emph{G. pamelaeae} (0--6 months, rank =
+16; 18--120 months, rank = 20), while others such as \emph{Allistipes
+finegoldii} were age-group specific. Several taxa important in RF models
+were not statistically significant when using linear models after
+multiple hypothesis correction. However, these taxa had small nominal
+p-values. For example, \emph{Erysipelatoclostridium ramosum} was the
+most important feature in RF models for children under 6 months old and
+had an LM p-value of 0.04. Subject age was consistently ranked highly in
+feature importance, which could indicate that decision branches based on
+microbial taxa have increased purity when considering the subject's age
+or that age itself is a useful predictor.
+
+\subsubsection{Gut microbial taxonomic profiles predict brain structure differences}
+
+If there are causal effects of microbial metabolism on cognitive
+function, they might be reflected in changes in neuroanatomy. We again
+employed a Random Forest modeling approach to associate gut taxonomic
+profiles with individual brain regions identified in MRI scans,
+normalized to total brain volume. Some brain regions were more readily
+predicted by RF models trained on microbial taxa (Table 4, Supplementary
+Table XX), in particular those that were highly correlated with age.
+These included the L/R lingual gyrus (mean RF correlations, Left =
+0.421, Right = 0.434; relative age importances, Left = 7.5\%, Right =
+8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
+0.200, Right = 0.273; relative age importances, Left = 3.7\%, Right =
+6.3\%). In many cases, however, age was not an important variable in
+high-performing models, such as that for the left accumbens area (mean
+RF correlations = 0.288; relative age importances = 1.2\%).
+
+\textbf{Table 4. Summary statistics for the neuroanatomy prediction
+benchmarks}
+
+We also observed that many brain regions had high accordance across the
+left and right hemispheres in terms of both model performance and
+microbial feature importance. In contrast, other regions had substantial
+differences between the hemispheres. For example, the left accumbens
+area, which plays an important role in reward circuits
+\cite{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012},
+has one of the highest test-set correlations of our
+brain region models (R = 0.288), as compared to the right accumbens
+models which could not adequately generalize, and had a negative mean
+test-set correlation (R = -0.041). Most healthy individuals have a
+rightward asymmetry in the nucleus accumbens, and reduced asymmetry has
+been linked to substance use disorder in young adults
+\cite{caoMappingCorticalSubcortical2021}.
+
+Feature importance for models of the left accumbens area were dominated
+by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.4\%
+relative importance), \emph{B. ovatus} (3.7\% relative importance), and
+\emph{B. uniformis} (3.0\% relative importance).. The accumbens area is
+associated with reward control, and in individuals diagnosed with ADHD,
+it has been shown to have a divergent neuromorphology
+\cite{hoogmanSubcorticalBrainVolume2017}.
+Independently, \emph{B. ovatus}, \emph{B. uniformis} and
+\emph{B. vulgatus} have been linked to ADHD
+\cite{wangGutMicrobiotaDietary2020}.
+
+In fact, there are studies showing that alterations on the striatal
+dopamine transporters can cause effects resembling hyperactivity and
+attention deficit
+\cite{yaelDisinhibitionNucleusAccumbens2019},
+and that \emph{B. uniformis} is gut-microbial modulator of the brain
+dopamine transporter
+\cite{hartstraInfusionDonorFeces2020}.
+
+Many of the taxa identified included taxa also identified in LMs and RF
+models of cognitive function. Interesting to note, while RF models for
+multiple brain regions had many important microbial taxa, others were
+dominated by a small number of taxa. In general, for all segments, a
+consistent number of species between 14 and 22 (median = 20) was
+responsible for one third of the fitness-weighted cumulative importance
+(see Methods), regardless of the model benchmarks. We proceeded to
+select a subset of the most important taxa (averaged over all segments)
+and the segments whose importance was more heavily loaded to further
+analyze the multiple relationships unveiled by the RF models
+(Figure 4 A and B)
+
+\includegraphics[width=6.5in,height=5.19444in]{media/image1.png}
+
+\textbf{Figure 4: Microbial feature importance in predicting brain
+volumes in children over 18 months of age.}
+
+(A) Average test-set correlations for prediction of MRI segmentation
+data from microbiome and demographics on select segments after repeated
+cross validation. (B) Heatmap of average individual relative taxa
+importances on each brain segment. Importances are reported as
+proportions relative to the sum of importances for each model - since
+every model is trained with 132 features, and even distribution of
+importance would be 0.75\% for each feature. Segments and taxa ordered
+by HCA on a select list of species with high ranks on average importances,
+and their respective highest-load segments (C).
+
+Our analysis revealed two major patterns of importance distribution from
+taxa over the brain segments; some species portrayed high contributions
+to multiple different segments, while others contributed modestly to
+just one or two brain segments. Notable cases of the first pattern
+included seven species - \emph{Anaerostipes hadrus}, \emph{Bacteroides
+vulgatus}, \emph{Fusicatenibacter saccharivorans}, \emph{Ruminococcus
+torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
+\emph{Blautia wexlerae} - that combined, account for approximately one
+third of the cumulative relative importances, computed after subsetting
+on the taxa of interest (Figure 4C).
+
+Among these, the most important variable is \emph{Anaerostipes hadrus},
+which is a butyrate-producing anaerobe that has been positively
+associated with cognitive function
+\cite{kantGenomeSequenceButyrateProducing2015,liCorrelationsGutMicrobiota2022}.
+Its importance is, however, heavily loaded on the volume
+of entire major lobes (frontal, parietal), which cannot be readily
+correlated to specific cognitive functions. After accounting for the
+major lobes, it is found to be highly important on the prediction of the
+cerebellar vermal lobules VIII-X, pars opercularis, cuneus and
+precuneus, and anterior-cingulate (Figure 4B).
+
+To better understand this relationship, we performed hierarchical
+cluster analysis, which revealed a well defined cluster of species with
+high importance loadings on close segments of the lower temporal and
+close occipital lobe, to which \emph{A. hadrus} belongs. The other
+species on this cluster are \emph{B. wexlerae}, \emph{R. torques},
+\emph{R. intestinalis, R. bicirculans} and \emph{F. saccahrivorans}, who
+contribute heavily to the entorhinal, fusiform, lingual and
+parahippocampal segments.
+
+This group of taxa and their related segments drew our attention because
+they contained \emph{B. wexlerae}, a highlight from the cognitive
+assessment results. Previous works exploring Parkinson Disease patients
+found out that issues in the cognitive task of confrontation naming were
+positively correlated with thinning in the fusiform gyrus and
+parahippocampal gyrus
+\cite{pagonabarragaPatternRegionalCortical2013}.
+Additionally, the left and right parahippocampal, where \emph{B.
+wexlerae} had the highest importance in both models, are important in
+visual/spatial processing and memory
+\cite{aminoffRoleParahippocampalCortex2013}.
+\emph{B. wexlerae} can also produce acetylcholine in the gut
+\cite{hosomiOralAdministrationBlautia2022},
+and this molecule plays an important role in modulating memory
+function
+\cite{haamCholinergicModulationHippocampal2017}.
+
+Another notable importance cluster contains taxa associated with the
+basal forebrain, especially the cingulate and the accumbens. This
+cluster contains the previously-discussed \emph{B. ovatus} and \emph{B.
+uniformis,} but also \emph{Alistipes finegoldii and Streptococcus
+salivaris.} Reduction in the nucleus accumbens has been associated with
+depression symptoms
+\cite{wackerRoleNucleusAccumbens2009},
+and increased levels of the \emph{Alistipes} genus have been
+observed in patients with depressive disorder
+\cite{jiangAlteredFecalMicrobiota2015}.
+
+While these reports are important, simultaneously probing of the gut
+metagenome and brain structure relationship is novel.
+
+Both \emph{A. equolofaciens} and \emph{A. celatus,} two closely-related
+equol-producing species, are examples of the second contribution
+pattern, and had high importance in predicting the relative volume of
+the right anterior cingulate (respectively, 3.0\% and 2.6\% relative
+importances), which has been linked to social cognition and reward-based
+decision making
+\cite{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002}.
+Equol has a strong
+estrogenic effect
+\cite{setchellSEquolPotentLigand2005},
+and in anterior cingulate, estrogen, has been shown to regulate
+pain-related aversion
+\cite{xiaoEstrogenAnteriorCingulate2013}.
+
+Another example of this pattern is \emph{R. gnavus,} which was the only
+species with significant negative association with cognitive function.
+It is heavily associated with the left pars opercularis (2.7\% relative
+importance), a relationship that may be explained by the emerging
+understanding that this species is increased in individuals with insulin
+resistance and obesity \cite{leyHumanGutMicrobes2006},
+conditions that are known to produce structural abnormalities in the brain
+\cite{opelBrainStructuralAbnormalities2021}.
+Finally, \emph{Coprococcus comes} displays an importance
+distribution that splits almost evenly among the two previously reported
+clusters. Its loading on the prediction of the left posterior cingulate
+is the highest for a microbe in RF models (4.0\% relative importance)
+(only age had higher relative importance in any model), while also being
+one of the most important predictors for neighboring areas of the
+posterior cingulate such as the pars opercularis (relative importances,
+Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
+precentral (2.3\% relative importance) and paracentral lobes (relative
+importances, Left = 2.6\%, Right = 1.8\%).
+
+\section*{Discussion}
+
+The relationship between the gut microbiome and brain function via the
+gut-microbiome-brain axis has gained increasing acceptance largely as a
+result of human epidemiological studies investigating atypical
+neurocognition (eg anxiety and depression, neurodegeneration, attention
+deficit / hyperactivity disorder, and autism) and mechanistic studies in
+animal models. The results from these studies point to the possibility
+that gut microbes and their metabolism may be causally implicated in
+cognitive development, but this study is the first to our knowledge that
+directly investigates microbial species and their genes in relation to
+typical development in young children. Understanding the gut-brain-axis
+in early life is particularly important, since differences or
+interventions in early life can have outsized and longer-term
+consequences than those at later ages. Further, even in the absence of
+causal impacts of microbial metabolism, identifying risk factors that
+could point to other early interventions would also have value.
+
+The use of shotgun metagenomic sequencing enabled us to get
+species-level resolution of microbial taxa. A previous study of
+cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing,
+and showed that genera from the Lachnospiraceae family as well as
+unclassified Clostridiales (now Eubacteriales) were associated with
+higher scores on the Ages and Stages Questionnaire
+\cite{sordilloAssociationInfantGut2019}.
+However, each of these clades encompass dozens of genera with
+diverse functions, each of which may have different effects. Indeed,
+several of the taxa that were positively associated with cognitive
+function in this study, including \emph{B. wexlerae}, the only species
+identified by linear models in children under 6 months, \emph{D.
+longicatena}, \emph{R. faecis}, and \emph{A. finegoldii} are
+Clostridiales, as is \emph{R. gnavus}, which we found was negatively
+associated with cognitive function (Figure 2A-B). This kind of
+species-level resolution is typically not possible with amplicon
+sequencing.
+
+We identified several species in the family Eggerthelaceae that were
+associated with cognitive function, including \emph{Gordonibacter
+pamelaeae, Aldercreutzia equolofaciens, Asaccharobacter celatus}
+(formerly regarded a subspecies of \emph{A. equolofaciens}
+\cite{takahashiCompleteGenomeSequence2021}),
+and \emph{Eggerthella lenta}. Many members of this family are
+known in part due to unique metabolic activities. For example, \emph{A.
+equolofaciens} produces the nonsteroidal estrogen equol from isoflaven
+found in soybeans
+\cite{wangEnantioselectiveSynthesisSEquol2005},
+and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
+(found in pomegranates and some berries) into urolithin, which has been
+shown in some studies to have a neuroprotective effect
+\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
+\emph{E. lenta} has been extensively studied for
+its ability to metabolize drug compounds such as the plant-derived heart
+medication digoxin
+\cite{haiserPredictingManipulatingCardiac2013}.
+The metabolic versatility of this clade, and the large number of
+species that are associated with cognition make these microbes prime
+targets for further mechanistic studies.
+
+In addition to improved species-level resolution, shotgun metagenomic
+sequencing also enables gene-functional insight. We showed here that
+genes for the metabolism of SCFAs, both their degradation and synthesis,
+are associated with cognitive function scores. However, while the
+differential abundance of genes for the metabolism of neuroactive
+compounds like these is suggestive, it is difficult to reason about the
+relationship between levels of these genes and the gut concentrations of
+the molecules their product enzymes act on. For example, while it might
+be intuitive to reason that increased levels of menaquinone synthesis
+genes is indicative of increased menaquinone, it could be the case that
+menaquinone deficiency selects for microbes that can synthesize it. For
+the same reason, increased propionate degradation genes may
+counterintuitively be indicative of high levels of propionate in the gut
+lumen, since high propionate would select for microbes that can
+metabolize it. For this reason, future studies coupling shotgun
+metagenomics with stool metabolomics could improve our understanding of
+the relationship between microbial metabolism and cognitive development.
+Further, strain-level analysis linking specific gene content in species
+of interest could further refine targeted efforts at identifying
+specific metabolic signatures of microbe-brain interactions.
+
+The use of multiple age-appropriate cognitive assessments that could be
+normalized to a common scale enabled us to analyze microbial
+associations across multiple developmental periods, but carries several
+drawbacks. In particular, the test-retest reliability, as well as
+systematic differences between test administrators may introduce
+substantial noise into these observations, particularly in the youngest
+children. In addition, our study period overlapped with the beginning of
+the COVID-19 pandemic, and we and others have observed some reduction in
+measured scores for children that were assessed after the implementation
+of lockdowns. In our subject set for this study, these effects are more
+pronounced in some age groups due to our sampling schedule
+\cite{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
+(Supplementary Figure 3).
+
+This analysis allowed us to establish links between microbial taxa and
+their functional potential with cognition and brain structure. Although
+we cannot test causality or the chemistry behind the interactions
+between gut microbial taxa, gut, and brain, this study provides clear
+and statistically significant associations between the infant and early
+child gut microbiota and neurocognition. Future studies should focus on
+characterizing the early-life microbiome and neurocognitive development
+across different geographic regions and lifestyles such as covering
+traditionally understudied low-resource urban, peri-urban and rural
+communities to obtain the more comprehensive understanding of the
+variability within the different gut microbiomes reflects on
+neurocognition. These studies would also provide us with the wealth of
+data on different strains from the same species to better understand the
+effect of genes and their products. Furthermore, culturing and microbial
+community enrichment studies combined with genetic manipulation and
+genomic approaches to understand microbial metabolism at the molecular
+level is the key, as the metabolic functions shape and influence the
+human host and its health. The discovery of the neuroactive metabolites
+could provide us with biomarkers for early detection or necessary
+medicinally useful molecules that can be applied in intervention.
 
@@
 
-\input{sections/methods}
+\subsubsection{Study Ethics}
 
-\section*{Results}
+All procedures for this study were approved by the local institutional
+review board at Rhode Island Hospital, and all experiments adhered to
+the regulation of the review board. Written informed consent was
+obtained from all parents or legal guardians of enrolled participants.
 
-\input{sections/results}
+\subsubsection{Participants}
+
+Data used in this study were drawn from the ongoing longitudinal
+RESONANCE study of healthy and neurotypical brain and cognitive
+development, based at Brown University in Providence, RI, USA. The
+RESONANCE study is part of the NIH initiative Environmental influences
+on Child Health Outcomes (ECHO)
+\href{https://www.zotero.org/google-docs/?0tZanX}{(Forrest et al., 2018;
+Gillman and Blaisdell, 2018)}, a longitudinal observational study of
+healthy and neurotypical brain development that spans the fetal and
+infant to adolescent life stages, combining neuroimaging (magnetic
+resonance imaging, MRI), neurocognitive assessments, bio-specimen
+analyses, subject genetics, environmental exposures such as lead, and
+rich demographic, socioeconomic, family and medical history information.
+From the RESONANCE cohort, 361 typically-developing children between the
+ages of 2.8 months and 10 years old (median age 2 years, 2 months) were
+selected for analysis in this study.
+
+General participant demographics are provided in Table 1. Children are
+representative of the RI population. Children in the RESONANCE cohort
+were born full-term (\textless{} 37 weeks gestation) with height and
+weight normal for gestational age, and from uncomplicated singleton
+pregnancies. Children with known major risk factors for developmental
+abnormalities at enrollment were excluded. In addition to screening at
+the time of enrollment, on-going screening for worrisome behaviors using
+validated tools was performed to identify at-risk children and remove
+them from subsequent analysis.
+
+Exclusion criteria included: \emph{in utero} exposure to alcohol,
+cigarette or illicit substance exposure; preterm (\textless{} 37 weeks
+gestation) birth; small for gestational age or less than 1500 g; fetal
+ultrasound abnormalities; preeclampsia, high blood pressure, or
+gestational diabetes; 5 minute APGAR scores \textless{} 8; NICU
+admission; neurological disorder (e.g., head injury resulting in loss of
+consciousness, epilepsy); and psychiatric or learning disorder
+(including maternal depression) in the infant, parents, or siblings
+requiring medication in the year prior to pregnancy.
+
+Demographic and other non-biospecimen data such as race and ethnicity,
+parental education and occupation, feeding behavior (breast- and
+formula-feeding), child weight and height, were collected through
+questionnaires or direct examination as appropriate. All data were
+collected at every assessment visit, if possible.
+
+\subsubsection{Cognitive Assessments}
+
+Overall cognitive function was assessed using age-appropriate methods.
+For children from birth to 30 months, we used an Early Learning
+Composite as assessed via the Mullen Scales of Early Learning (MSEL)
+\href{https://www.zotero.org/google-docs/?NSyykd}{(Mullen and others,
+1995)}, a standardized and population-normed tool for assessing fine and
+gross motor, expressive and receptive language, and visual reception
+functioning in children from birth through 68 months of age. The
+Wechsler Intelligence Quotient for Children (WISC)
+\href{https://www.zotero.org/google-docs/?OopdA1}{(Wechsler, 2012)} is
+an individually administered standard intelligence test for children
+aged 6 to 16 years. It derives a full scale intelligence quotient (IQ)
+score, which we used to assess overall cognitive functioning. The fourth
+edition of the Wechsler Preschool and Primary Scale of Intelligence
+(WPPSI-IV) is an individually administered standard intelligence test
+for children aged 2 years 6 months to 7 years 7 months, trying to meet
+the increasing need for the assessment of preschoolers. Just as the
+WISC, it derives a full scale IQ score, which we used to assess overall
+cognitive functioning.
+
+\subsubsection{Stool Sample Collection an
+Sequencing}
+
+Stool samples (n=493) were collected by parents in OMR-200 tubes
+(OMNIgene GUT, DNA Genotek, Ottawa, Ontario, Canada), stored on ice, and
+brought within 24 hrs to the lab in RI where they were immediately
+frozen at -80 ˚C. Stool samples were not collected if the subject had
+taken antibiotics within the last two weeks. DNA extraction was
+performed at Wellesley College (Wellesley, MA). Nucleic acids were
+extracted from stool samples using the RNeasy PowerMicrobiome kit,
+excluding the DNA degradation steps. Briefly, the samples were lysed by
+bead beating using the Powerlyzer 24 Homogenizer (Qiagen, Germantown,
+MD) at 2500 rpm for 45 s and then transferred to the QIAcube (Qiagen,
+Germantown, MD) to complete the extraction protocol. Extracted DNA was
+sequenced at the Integrated Microbiome Resource (IMR, Dalhousie
+University, NS, Canada).
+
+Shotgun metagenomic sequencing was performed on all samples. A pooled
+library (max 96 samples per run) was prepared using the Illumina Nextera
+Flex Kit for MiSeq and NextSeq from 1 ng of each sample. Samples were
+then pooled onto a plate and sequenced on the Illumina NextSeq 550
+platform using 150+150 bp paired-end ``high output'' chemistry,
+generating 400 million raw reads and 120 Gb of sequence per plate.
+
+\subsubsection{Machine learning for cognitiv
+development}
+
+Prediction of cognitive scores was carried out as a set of regression
+experiments targeting real-valued continuous assessment scores.
+Different experiment sets were designed to probe how different
+representations of the gut microbiome (taxonomic profiles, Functional
+profiles encoded as ECs) would behave, with and without the addition of
+demographics (sex and maternal education as a proxy of socioeconomic
+status) on participants from different age groups. Age (in months) was
+provided as a covariate for all models (Table 3).
+
+\textbf{Table 5. Experimental design and input composition for Random
+Forest experiments}
+
+Random Forests (RFs) \href{https://paperpile.com/c/dPbU4e/VDqU}{(Breiman
+1996)} were selected as the prediction engine and processed using the
+DecisionTree.jl
+\href{https://www.zotero.org/google-docs/?dQLyDs}{(Sadeghi et al.,
+2022)} implementation, inside the MLJ.jl
+\href{https://www.zotero.org/google-docs/?y8ywGI}{(Blaom et al., 2020)}
+Machine Learning framework. Independent RFs were trained for each
+experiment, using a set of default regression hyperparameters from
+Breiman and Cutler
+\href{https://www.zotero.org/google-docs/?d3gqFh}{(Breiman, 2001)}, on a
+repeated cross-validation approach with different RNG seeds. One hundred
+repetitions of 3-fold CV with 10 different intra-fold RNG states each
+were employed, for a total of 3000 experiments per input set.
+
+After the training procedures, the root-mean-square error (RMSE) for
+cognitive assessment scores and mean absolute proportional error (MAPE)
+for the brain segmentation data, along with Pearson\textquotesingle s
+correlation coefficient (R) were benchmarked on the validation and train
+sets. MAPE was chosen as the metric for brain segments due to magnitude
+differences between median volumes of each segment, which would hinder
+interpretation of raw error values without additional reference.
+
+To derive biological insight from the models, the covariate variable
+importances for all the input features, measured by mean decrease in
+impurity (MDI, or GINI importance), was also analyzed. Leveraging the
+distribution of results from the extensive repeated cross validation
+experiments, rather than electing a representative model or picking the
+highest validation-set correlation, a measure of model fitness
+(\textbf{Equation 1}) was designed to weight the importances from each
+trained forest. The objective was to give more weight to those with
+higher benchmarks on the validation sets (or higher generalizability),
+while penalizing information from highly overfit models, drawing
+inspiration from the approach used on another work employing repeated CV
+on Random Forests with high-dimensional, low sample size microbiome
+datasets \{***\}. The resulting fitness-weighted importances were used
+to generate the values in Figure 3.
+
+\(fitness =\)
+
+\textbf{Equation 1.} Mathematical expression of the fitness measure used
+to weight feature importances based on model benchmarks
+
+\subsubsection{MRI / segmentation}
+
+MRI data was acquired at a 3T Siemens Trio scanner with the following
+parameters: TE=5.6msec, TR=1400msec, FA=15 degrees, 1.1x1.1x1.1mm
+resolution, 160x160 matrix, with an average of 112 slices. FOV was
+adjusted to infant size. Using a combination of linear and nonlinear
+image registration, we created representative age-specific templates
+using the ANTs package
+\href{https://www.zotero.org/google-docs/?ZlWDfL}{(Avants et al.,
+2014)}. After age specific templates were created, a single flow from
+each age to the 12-month template was estimated and a final warp from
+the 12-month template to standard adult MNI space was performed. For
+each individual infant/ child brain image, we then calculated the warp
+from their native T1w image space to their nearest-in-age template.
+Using the resulting warps (native → nearest age template → 12-month
+template → MNI template), we could move the standard adult brain atlas
+to the space of an individual infant in a single step. In this case, we
+used the Harvard-Oxford brain atlas to provide a coarse-grained
+parcellation of individual brains into subcortical regions (e.g.,
+thalamus, putamen) and total grey and white matter volumes (included as
+part of the FSL package
+\href{https://www.zotero.org/google-docs/?AapPN7}{(Jenkinson et al.,
+2012)}. Total tissue and brainstem volume as well as left and right
+hemisphere volumes were derived for total white and cortical gray
+matter, lateral ventricle, thalamus, caudate, putamen, pallidum,
+hippocampus, amygdala, and accumbens as well as total brainstem volume
+\href{https://www.zotero.org/google-docs/?MCDDB9}{(Bruchhage et al.,
+n.d.)}.
+
+\subsubsection{Computational analysis of metagenomes}
+
+Shotgun metagenomic sequences were analyzed using the bioBakery suite of
+computational tools \href{https://paperpile.com/c/dPbU4e/vDau}{(Beghini
+et al. 2021)}. First, KneadData (v0.7.7) was used to perform quality
+control of raw sequence reads, such as read trimming and removal of
+reads matching a human genome reference. Next, MetaPhlAn (v3.0.7, using
+database mpa\_v30\_CHOCOPhlAn\_201901) was used to generate taxonomic
+profiles by aligning reads to a reference database of marker genes.
+Finally, HUMAnN (v3.0.0a4) was used to functionally profile the
+metagenomes.
+
+\subsubsection{Data and code availability}
+
+Taxonomic and functional microbial profiles, as well as subject
+demographics necessary for statistical analyses and machine learning are
+available on the Open Science Framework
+\href{https://www.zotero.org/google-docs/?61HLlM}{(Bonham et al.,
+2022)}. Data processing, generation of summary statistics, and
+generation of plots was performed using the julia programming language
+\href{https://www.zotero.org/google-docs/?zj3HnY}{(Bezanson et al.,
+2017; Bonham et al., 2021; Danisch and Krumbiegel, 2021)}. All code for
+data analysis and figure generation, as well as scripts for automated
+download of input files are available on github \{\{CITE zenodo\}\}.
+
+\subsection{Research Standards}
+
+\subsection{Acknowledgements}
+
+Research funding: NIH UG3 OD023313 and Wellcome LEAP 1kD.
 

### 2023-02-17 manuscript/main.tex
add figures

@@
 
-\addbibresource{zotero_library}
+\addbibresource{zotero_lib}
 
@@
 \usepackage{graphicx}
+\graphicspath{assets}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{graph_a}
-    \caption{An example graph}
-    \label{fig:x cubed graph}
-\end{figure}
-
-(A) Study design and overview. Stool samples, cognitive assessments and
+    \includegraphics[width=\textwidth]{assets/Figure1.png}
+    \caption{
+        Study design and overview. (\textbf{A}) Stool samples, cognitive assessments and
         neuroimaging were collected from participants in an accelerated
-longitudinal design. (B) Cognitive function scores are assessed by
+        longitudinal design. (\textbf{B}) Cognitive function scores are assessed by
         different instruments, but may be normalized to a common IQ-like scale.
-(C-D) Principal coordinates analysis using Bray-Curtis dissimilarity on
+        (\textbf{C-D}) Principal coordinates analysis using Bray-Curtis dissimilarity on
         taxonomic profiles demonstrates high beta diversity, with much of the
         first axis of variation explained by increasing age and alpha diversity.
-Differences in gene function profiles (E) and neuroimaging (PCA based on
-the Euclidean distance of brain region volumes) (F) are likewise
-dominated by changes as children age. (G) Permutation analysis of
+        Differences in gene function profiles (\textbf{E}) and neuroimaging (PCA based on
+        the Euclidean distance of brain region volumes) (\textbf{F}) are likewise
+        dominated by changes as children age. (\textbf{G}) Permutation analysis of
         variance (PERMANOVA) of taxonomic profiles, functional profiles
@@
         modest but significant percent of the variation in cognitive development
-in children over 18 months of age. (H) Mantel testing of different
+        in children over 18 months of age. (\textbf{H}) Mantel testing of different
         microbial feature matrices, shows overlapping but distinct patterns of
@@
         used as cut-offs in some following models.
+    }
+    \label{fig:Figure 1}
+\end{figure}
 
@@
 prevalence in a given age group (Figure 2A, N = 92 for 0--120 months, N
-= 46 for 0--6 months, N = 97 for 18--120 months). Only \emph{Blautia
-wexlerae} was significantly associated (q value = 0.14, β = 0.0015) with
+= 46 for 0--6 months, N = 97 for 18--120 months). Only \emph{Blautia wexlerae}
+was significantly associated (q value = 0.14, $\beta$ = 0.0015) with
 cognitive function in children under 6 months old after adjusting for
@@
 \cite{ghoshMediterraneanDietIntervention2020}
-(Figure
-2B).
+(Figure 2B).
 
-\includegraphics[width=6.5in,height=4.83333in]{media/image4.png}
-
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{assets/Figure2.png}
+    \caption{
         \textbf{Figure 2 - Taxa and gene functional groups are associated with cognitive function.}
-
         (A) Volcano plot of multivariable linear model results showing the
@@
         median correlation of that gene set.
+    }
+    \label{fig:Figure 2}
+\end{figure}
 
@@
 \cite{valles-colomerNeuroactivePotentialHuman2019}
-and concurrent cognitive function score (Table 2,
-\protect\hyperlink{_i7fd3z19jdvg}{\uline{Figure 2}}C-G) and found that
+and concurrent cognitive function score (Table 2, Figure 2C-G) and found that
 several metabolic pathways were significantly enriched or depleted in
@@
 old (acetate synthesis I, E.S. = -0.194, q = 0.153; acetate synthesis
-II, E.S. = -0.342, q = 0.020; acetate synthesis III, E.S. = -0.31, q =
-0.052). SCFAs are produced by anaerobic fermentation of dietary fiber
+II, E.S. = -0.342, q = 0.020; acetate synthesis III, E.S. = -0.31, q = 0.052).
+SCFAs are produced by anaerobic fermentation of dietary fiber
 and have been linked with immune system regulation as well as directly
-with brain function
-\cite{dalileRoleShortchainFatty2019}.
+with brain function \cite{dalileRoleShortchainFatty2019}.
 
@@
 
-\textbf{Table 2 - Feature set enrichment analysis on neuroactive
-microbial gene sets}
+\textbf{Table 2 - Feature set enrichment analysis on neuroactive microbial gene sets}
 
@@
 
-\includegraphics[width=6.5in,height=5.19444in]{media/image2.png}
-
-\textbf{Figure 3 - Random forest models predict concurrently measured
-cognitive function.}
-
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidt]{assets/Figure3.png}
+    \caption{
+        \textbf{Figure 3 - Random forest models predict concurrently measured cognitive function.}
         Comparison of RF predictor importance versus linear models for children
@@
         over 18 months (G)
+    }
+    \label{fig:Figure 3}
+\end{figure}
 
@@
 
-\textbf{Table 4. Summary statistics for the neuroanatomy prediction
-benchmarks}
+\textbf{Table 4. Summary statistics for the neuroanatomy prediction benchmarks}
 
@@
 
-\includegraphics[width=6.5in,height=5.19444in]{media/image1.png}
-
-\textbf{Figure 4: Microbial feature importance in predicting brain
-volumes in children over 18 months of age.}
-
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{graph_a}
+    \caption{
+        \textbf{Figure 4: Microbial feature importance in predicting brain volumes in children over 18 months of age.}
         (A) Average test-set correlations for prediction of MRI segmentation
@@
         and their respective highest-load segments (C).
+    }
+    \label{fig:x cubed graph}
+\end{figure}
 
@@
 \cite{jiangAlteredFecalMicrobiota2015}.
-
 While these reports are important, simultaneously probing of the gut
@@
 \cite{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002}.
-Equol has a strong
-estrogenic effect
+Equol has a strong estrogenic effect
 \cite{setchellSEquolPotentLigand2005},

### 2023-02-20 manuscript/main.tex
get refs working

@@
 
-\addbibresource{zotero_lib}
+\addbibresource{zotero.bib}
 
@@
 \usepackage{multirow}
+
+% figures
+\usepackage{wrapfig}
+\usepackage{lscape}
+\usepackage{rotating}
 \usepackage{graphicx}
@@
 nerve and through endocrine and immune (cytokine) signaling molecules
-\cite{cerdoEarlyNutritionGut2019,pronovostPerinatalInteractionsMicrobiome2019,sharonCentralNervousSystem2016,togniniGutMicrobiotaPotential2017}
+\citep{cerdoEarlyNutritionGut2019,pronovostPerinatalInteractionsMicrobiome2019,sharonCentralNervousSystem2016,togniniGutMicrobiotaPotential2017}
 In addition, the products of microbial metabolism generated in the gut can
@@
 development, are increasingly being identified
-\cite{spichakMiningMicrobesMental2021}.
+\citep{spichakMiningMicrobesMental2021}.
 Both human epidemiology and animal models point to the effects
 of gut microbes on the development of autism spectrum disorder
-\cite{laueProspectiveAssociationsInfant2020,wanUnderdevelopmentGutMicrobiota2021}
+\citep{laueProspectiveAssociationsInfant2020,wanUnderdevelopmentGutMicrobiota2021}
 and particular microbial taxa have been associated with depression
-\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,valles-colomerNeuroactivePotentialHuman2019}
+\citep{mayneris-perxachsMicrobiotaAlterationsProline2022,valles-colomerNeuroactivePotentialHuman2019}
 and Alzheimer's disease
-\cite{fungInteractionsMicrobiotaImmune2017,kimProbioticSupplementationImproves2021}.
+\citep{fungInteractionsMicrobiotaImmune2017,kimProbioticSupplementationImproves2021}.
 But information about this "microbiome-gut-brain
@@
 microbiome and the brain
-\cite{laueDevelopingMicrobiomeBirth2022}.
+\citep{laueDevelopingMicrobiomeBirth2022}.
 Fetal development \emph{in utero} is believed to be mostly sterile, but
@@
 and other environmental sources such as antibiotics
-\cite{backhedDynamicsStabilizationHuman2015,bokulichAntibioticsBirthMode2016}.
+\citep{backhedDynamicsStabilizationHuman2015,bokulichAntibioticsBirthMode2016}.
 The early microbiome is characterized by low microbial
@@
 Bacteroidetes, especially \emph{Bacteroides}, and Proteobacteria
-\cite{koenigSuccessionMicrobialConsortia2011}.
+\citep{koenigSuccessionMicrobialConsortia2011}.
 Many of these microbes have specialized metabolic capabilities
@@
 and \emph{Bacteroides fragilis}
-\cite{selaGenomeSequenceBifidobacterium2008}.
+\citep{selaGenomeSequenceBifidobacterium2008}.
 Upon the introduction of solid foods, the gut
@@
 adult microbiomes
-\cite{backhedDynamicsStabilizationHuman2015}.
+\citep{backhedDynamicsStabilizationHuman2015}.
 Many prior studies have focused on either infant microbiomes or
@@
 incredibly important
-\cite{tauNormalDevelopmentBrain2010}.
+\citep{tauNormalDevelopmentBrain2010}.
 
@@
 connections has been established
-\cite{silbereisCellularMolecularLandscapes2016}.
+\citep{silbereisCellularMolecularLandscapes2016}.
 Much of this development occurs in discrete windows or
 ``sensitive-'' or ``critical periods'' (CPs)
-\cite{knudsenSensitivePeriodsDevelopment2004}
+\citep{knudsenSensitivePeriodsDevelopment2004}
 when neural plasticity is particularly high,
@@
 environment, including the gut microbiome
-\cite{cowanAnnualResearchReview2020}.
+\citep{cowanAnnualResearchReview2020}.
 In fact, emerging evidence suggests that the timing and duration of CPs
 may be driven in part by cues from the developing gut microbiome
-\cite{callaghanNestedSensitivePeriods2020}.
+\citep{callaghanNestedSensitivePeriods2020}.
 As such,
@@
 an accelerated longitudinal cohort of early development
-\cite{forrestAdvancingScienceChildren2018}
+\citepp{forrestAdvancingScienceChildren2018}
 between the ages of 82 days and 10 years
@@
 for children from birth to 3 years of age
-\cite{mullenMullenScalesEarly1995},
+\citep{mullenMullenScalesEarly1995},
 Wechsler preschool and primary scale of intelligence (WPPSI)
 for 4-5 year-olds
-\cite{wechslerWechslerPreschoolPrimary2012},
+\citep{wechslerWechslerPreschoolPrimary2012},
 and the Wechsler intelligence scale for children (WISC)
 for children 6 years and up
-\cite{wechslerWechslerIntelligenceScale1949}.
+\citep{wechslerWechslerIntelligenceScale1949}.
 
@@
 
-\begin{figure}[h]
+\begin{sidewaysfigure}[!htb]
     \centering
@@
     \label{fig:Figure 1}
-\end{figure}
+\end{sidewaysfigure}
 
@@
 measures of anxiety and depression
-\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022},
+\citep{mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022},
 cognitive flexibility in adults
-\cite{magnussonRelationshipsDietrelatedChanges2015},
+\citep{magnussonRelationshipsDietrelatedChanges2015},
 and atypical neural development
-\cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021}.
+\citep{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021}.
 We, therefore, set out to identify whether specific taxa
@@
 profiles (annotated with clusters of 90\% similarity
-\cite{suzekUniRefComprehensiveNonredundant2007}
+\citep{suzekUniRefComprehensiveNonredundant2007}
 ), both generated from DNA sequencing data, being tightly coupled
@@
 groupings (Enzyme commission level4 - ECs
-\cite{bairochENZYMEDatabase20002000},
+\citep{bairochENZYMEDatabase20002000},
 and
 KEGG orthologues - KOs
-\cite{kanehisaKEGGResourceDeciphering2004}
+\citep{kanehisaKEGGResourceDeciphering2004}
 ) overlapped only slightly with taxonomic profiles in both age
@@
 cognitive function, we fit multivariable linear regression
-\cite{mallickMultivariableAssociationDiscovery2021}
+\citep{mallickMultivariableAssociationDiscovery2021}
 to the relative abundance of each species that had at least 15\%
@@
 previously shown to be depleted in children with diabetes
-\cite{benitez-paezDepletionBlautiaSpecies2020},
+\citep{benitez-paezDepletionBlautiaSpecies2020},
 and that oral administration of \emph{B. wexlerae} partially
@@
 model of T2D
-\cite{hosomiOralAdministrationBlautia2022,liuBlautiaNewFunctional2021}.
+\citep{hosomiOralAdministrationBlautia2022,liuBlautiaNewFunctional2021}.
 In children over 18 months of age,
@@
 neuroprotective metabolite urolithin
-\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014},
+\citep{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014},
 \emph{Asaccharobacter celatus} and
@@
 equol
-\cite{maruoAdlercreutziaEquolifaciensGen2008,thawornkunoBiotransformationDaidzeinEquol2009},
+\citep{maruoAdlercreutziaEquolifaciensGen2008,thawornkunoBiotransformationDaidzeinEquol2009},
 and the SCFA-producing probiotic
@@
 prausnitzii}
-\cite{ghoshMediterraneanDietIntervention2020}
+\citep{ghoshMediterraneanDietIntervention2020}
 (Figure 2B).
 
-\begin{figure}[h]
+\begin{sidewaysfigure}[!htb]
     \centering
@@
     \caption{
-        \textbf{Figure 2 - Taxa and gene functional groups are associated with cognitive function.}
+        Taxa and gene functional groups are associated with cognitive function.
         (A) Volcano plot of multivariable linear model results showing the
@@
     \label{fig:Figure 2}
-\end{figure}
+\end{sidewaysfigure}
 
@@
 of genes with neuroactive potential
-\cite{valles-colomerNeuroactivePotentialHuman2019}
+\citep{valles-colomerNeuroactivePotentialHuman2019}
 and concurrent cognitive function score (Table 2, Figure 2C-G) and found that
@@
 and have been linked with immune system regulation as well as directly
-with brain function \cite{dalileRoleShortchainFatty2019}.
+with brain function \citep{dalileRoleShortchainFatty2019}.
 
@@
 which, MK-4, has been found to be decreased in children diagnosed with ASD
-\cite{dongCorrelationSerumConcentrations2021}
+\citep{dongCorrelationSerumConcentrations2021}
 and is potentially neuroprotective in both rodents and humans
-\cite{elkattawyVitaminK2Menaquinone72022}.
+\citep{elkattawyVitaminK2Menaquinone72022}.
 In children over 18 months old,
@@
 the developing brain
-\cite{cohenkadoshLinkingGABAGlutamate2015,palomo-buitragoGlutamateInteractionsObesity2019}.
+\citep{cohenkadoshLinkingGABAGlutamate2015,palomo-buitragoGlutamateInteractionsObesity2019}.
 Tryptophan metabolism, including
@@
 autism in children
-\cite{hoshinoBloodSerotoninFree1984,xiaoFecalMicrobiomeTransplantation2021}.
+\citep{hoshinoBloodSerotoninFree1984,xiaoFecalMicrobiomeTransplantation2021}.
 Taken together, these results suggest that
@@
 from the domain of genomics
-\cite{amaratungaEnrichedRandomForests2008,brieucPracticalIntroductionRandom2018,chenRandomForestsGenomic2012,franzosaGutMicrobiomeStructure2019,stephanRandomForestApproach2015},
+\citep{amaratungaEnrichedRandomForests2008,brieucPracticalIntroductionRandom2018,chenRandomForestsGenomic2012,franzosaGutMicrobiomeStructure2019,stephanRandomForestApproach2015},
 along with other works where it was used
@@
 different scenarios
-\cite{ardekaniPredictionIncipientAlzheimer2017,velazquezRandomForestModel2021}.
+\citep{ardekaniPredictionIncipientAlzheimer2017,velazquezRandomForestModel2021}.
 Additionally, given the sequential
@@
 
-\begin{figure}[h]
+\begin{sidewaysfigure}[h]
     \centering
-    \includegraphics[width=\textwidt]{assets/Figure3.png}
+    \includegraphics[width=\textwidth]{assets/Figure3.png}
     \caption{
-        \textbf{Figure 3 - Random forest models predict concurrently measured cognitive function.}
+        Random forest models predict concurrently measured cognitive function.
         Comparison of RF predictor importance versus linear models for children
@@
     \label{fig:Figure 3}
-\end{figure}
+\end{sidewaysfigure}
 
@@
 area, which plays an important role in reward circuits
-\cite{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012},
+\citep{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012},
 has one of the highest test-set correlations of our
@@
 been linked to substance use disorder in young adults
-\cite{caoMappingCorticalSubcortical2021}.
+\citep{caoMappingCorticalSubcortical2021}.
 
@@
 it has been shown to have a divergent neuromorphology
-\cite{hoogmanSubcorticalBrainVolume2017}.
+\citep{hoogmanSubcorticalBrainVolume2017}.
 Independently, \emph{B. ovatus}, \emph{B. uniformis} and
 \emph{B. vulgatus} have been linked to ADHD
-\cite{wangGutMicrobiotaDietary2020}.
+\citep{wangGutMicrobiotaDietary2020}.
 
@@
 attention deficit
-\cite{yaelDisinhibitionNucleusAccumbens2019},
+\citep{yaelDisinhibitionNucleusAccumbens2019},
 and that \emph{B. uniformis} is gut-microbial modulator of the brain
 dopamine transporter
-\cite{hartstraInfusionDonorFeces2020}.
+\citep{hartstraInfusionDonorFeces2020}.
 
@@
 
-\begin{figure}[h]
+\begin{sidewaysfigure}[!htb]
     \centering
-    \includegraphics[width=\textwidth]{graph_a}
+    \includegraphics[width=\textwidth]{assets/Figure4.png}
     \caption{
-        \textbf{Figure 4: Microbial feature importance in predicting brain volumes in children over 18 months of age.}
+        Microbial feature importance in predicting brain volumes in children over 18 months of age.
         (A) Average test-set correlations for prediction of MRI segmentation
@@
     }
-    \label{fig:x cubed graph}
-\end{figure}
+    \label{fig:Figure 4}
+\end{sidewaysfigure}
 
@@
 associated with cognitive function
-\cite{kantGenomeSequenceButyrateProducing2015,liCorrelationsGutMicrobiota2022}.
+\citep{kantGenomeSequenceButyrateProducing2015,liCorrelationsGutMicrobiota2022}.
 Its importance is, however, heavily loaded on the volume
@@
 parahippocampal gyrus
-\cite{pagonabarragaPatternRegionalCortical2013}.
+\citep{pagonabarragaPatternRegionalCortical2013}.
 Additionally, the left and right parahippocampal, where \emph{B.
@@
 visual/spatial processing and memory
-\cite{aminoffRoleParahippocampalCortex2013}.
+\citep{aminoffRoleParahippocampalCortex2013}.
 \emph{B. wexlerae} can also produce acetylcholine in the gut
-\cite{hosomiOralAdministrationBlautia2022},
+\citep{hosomiOralAdministrationBlautia2022},
 and this molecule plays an important role in modulating memory
 function
-\cite{haamCholinergicModulationHippocampal2017}.
+\citep{haamCholinergicModulationHippocampal2017}.
 
@@
 depression symptoms
-\cite{wackerRoleNucleusAccumbens2009},
+\citep{wackerRoleNucleusAccumbens2009},
 and increased levels of the \emph{Alistipes} genus have been
 observed in patients with depressive disorder
-\cite{jiangAlteredFecalMicrobiota2015}.
+\citep{jiangAlteredFecalMicrobiota2015}.
 While these reports are important, simultaneously probing of the gut
@@
 decision making
-\cite{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002}.
+\citep{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002}.
 Equol has a strong estrogenic effect
-\cite{setchellSEquolPotentLigand2005},
+\citep{setchellSEquolPotentLigand2005},
 and in anterior cingulate, estrogen, has been shown to regulate
 pain-related aversion
-\cite{xiaoEstrogenAnteriorCingulate2013}.
+\citep{xiaoEstrogenAnteriorCingulate2013}.
 
@@
 understanding that this species is increased in individuals with insulin
-resistance and obesity \cite{leyHumanGutMicrobes2006},
+resistance and obesity \citep{leyHumanGutMicrobes2006},
 conditions that are known to produce structural abnormalities in the brain
-\cite{opelBrainStructuralAbnormalities2021}.
+\citep{opelBrainStructuralAbnormalities2021}.
 Finally, \emph{Coprococcus comes} displays an importance
@@
 higher scores on the Ages and Stages Questionnaire
-\cite{sordilloAssociationInfantGut2019}.
+\citep{sordilloAssociationInfantGut2019}.
 However, each of these clades encompass dozens of genera with
@@
 (formerly regarded a subspecies of \emph{A. equolofaciens}
-\cite{takahashiCompleteGenomeSequence2021}),
+\citep{takahashiCompleteGenomeSequence2021}),
 and \emph{Eggerthella lenta}. Many members of this family are
@@
 found in soybeans
-\cite{wangEnantioselectiveSynthesisSEquol2005},
+\citep{wangEnantioselectiveSynthesisSEquol2005},
 and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
@@
 shown in some studies to have a neuroprotective effect
-\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
+\citep{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
 \emph{E. lenta} has been extensively studied for
@@
 medication digoxin
-\cite{haiserPredictingManipulatingCardiac2013}.
+\citep{haiserPredictingManipulatingCardiac2013}.
 The metabolic versatility of this clade, and the large number of
@@
 pronounced in some age groups due to our sampling schedule
-\cite{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
+\citep{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
 (Supplementary Figure 3).
@@
 on Child Health Outcomes (ECHO)
-\href{https://www.zotero.org/google-docs/?0tZanX}{(Forrest et al., 2018;
-Gillman and Blaisdell, 2018)}, a longitudinal observational study of
+\citep{forrestAdvancingScienceChildren2018,gillmanEnvironmentalInfluencesChild2018},
+ a longitudinal observational study of
 healthy and neurotypical brain development that spans the fetal and
@@
 Composite as assessed via the Mullen Scales of Early Learning (MSEL)
-\href{https://www.zotero.org/google-docs/?NSyykd}{(Mullen and others,
-1995)}, a standardized and population-normed tool for assessing fine and
+\citep{mullenMullenScalesEarly1995},
+ a standardized and population-normed tool for assessing fine and
 gross motor, expressive and receptive language, and visual reception
@@
 Wechsler Intelligence Quotient for Children (WISC)
-\href{https://www.zotero.org/google-docs/?OopdA1}{(Wechsler, 2012)} is
-an individually administered standard intelligence test for children
+\citep{wechslerWechslerPreschoolPrimary2012}
+is an individually administered standard intelligence test for children
 aged 6 to 16 years. It derives a full scale intelligence quotient (IQ)
@@
 
-\subsubsection{Stool Sample Collection an
-Sequencing}
+\subsubsection{Stool Sample Collection and Sequencing}
 
@@
 brought within 24 hrs to the lab in RI where they were immediately
-frozen at -80 ˚C. Stool samples were not collected if the subject had
+frozen at -80$^{\circ}$C. Stool samples were not collected if the subject had
 taken antibiotics within the last two weeks. DNA extraction was
@@
 
-\subsubsection{Machine learning for cognitiv
-development}
+\subsubsection{Machine learning for cognitive development}
 
@@
 
-Random Forests (RFs) \href{https://paperpile.com/c/dPbU4e/VDqU}{(Breiman
-1996)} were selected as the prediction engine and processed using the
-DecisionTree.jl
-\href{https://www.zotero.org/google-docs/?dQLyDs}{(Sadeghi et al.,
-2022)} implementation, inside the MLJ.jl
-\href{https://www.zotero.org/google-docs/?y8ywGI}{(Blaom et al., 2020)}
+Random Forests (RFs)
+\citep{breimanRandomForests2001}
+were selected as the prediction engine and processed using the
+DecisionTree.jl \citep{sadeghiDecisionTreeJlJulia2022}
+implementation, inside the MLJ.jl \citep{blaomMLJJuliaPackage2020}.
+
 Machine Learning framework. Independent RFs were trained for each
 experiment, using a set of default regression hyperparameters from
-Breiman and Cutler
-\href{https://www.zotero.org/google-docs/?d3gqFh}{(Breiman, 2001)}, on a
-repeated cross-validation approach with different RNG seeds. One hundred
+Breiman and Cutler \citep{breimanRandomForests2001},
+on a repeated cross-validation approach with different RNG seeds. One hundred
 repetitions of 3-fold CV with 10 different intra-fold RNG states each
@@
 cognitive assessment scores and mean absolute proportional error (MAPE)
-for the brain segmentation data, along with Pearson\textquotesingle s
+for the brain segmentation data, along with Pearson's
 correlation coefficient (R) were benchmarked on the validation and train
@@
 on Random Forests with high-dimensional, low sample size microbiome
-datasets \{***\}. The resulting fitness-weighted importances were used
+datasets \citep{woodruffInflammationAutoreactivityDefine2022}.
+The resulting fitness-weighted importances were used
 to generate the values in Figure 3.
@@
 image registration, we created representative age-specific templates
-using the ANTs package
-\href{https://www.zotero.org/google-docs/?ZlWDfL}{(Avants et al.,
-2014)}. After age specific templates were created, a single flow from
+using the ANTs package \citep{avantsInsightToolKitImage2014}.
+After age specific templates were created, a single flow from
 each age to the 12-month template was estimated and a final warp from
@@
 thalamus, putamen) and total grey and white matter volumes (included as
-part of the FSL package
-\href{https://www.zotero.org/google-docs/?AapPN7}{(Jenkinson et al.,
-2012)}. Total tissue and brainstem volume as well as left and right
+part of the FSL package \citep{jenkinsonFSL2012}).
+Total tissue and brainstem volume as well as left and right
 hemisphere volumes were derived for total white and cortical gray
@@
 hippocampus, amygdala, and accumbens as well as total brainstem volume
-\href{https://www.zotero.org/google-docs/?MCDDB9}{(Bruchhage et al.,
-n.d.)}.
+\citep{bruchhageLongitudinalBrainCognitive}.
 
@@
 Shotgun metagenomic sequences were analyzed using the bioBakery suite of
-computational tools \href{https://paperpile.com/c/dPbU4e/vDau}{(Beghini
-et al. 2021)}. First, KneadData (v0.7.7) was used to perform quality
+computational tools
+(Beghini
+et al. 2021)
+\citep{}.
+ First, KneadData (v0.7.7) was used to perform quality
 control of raw sequence reads, such as read trimming and removal of
@@
 available on the Open Science Framework
-\href{https://www.zotero.org/google-docs/?61HLlM}{(Bonham et al.,
-2022)}. Data processing, generation of summary statistics, and
+(Bonham et al.,
+2022)
+\citep{}.
+ Data processing, generation of summary statistics, and
 generation of plots was performed using the julia programming language
-\href{https://www.zotero.org/google-docs/?zj3HnY}{(Bezanson et al.,
-2017; Bonham et al., 2021; Danisch and Krumbiegel, 2021)}. All code for
+(Bezanson et al.,
+2017; Bonham et al., 2021; Danisch and Krumbiegel, 2021)
+\citep{}.
+ All code for
 data analysis and figure generation, as well as scripts for automated
@@
 
-\subsection{Research Standards}
-
-\subsection{Acknowledgements}
+\section*{Research Standards}
+
+\section*{Acknowledgements}
 
@@
 
+\section*{References}
+
+\printbibliography
+
 \end{document}

### 2023-02-22 manuscript/main.tex
tweak table 1 and layout

@@
 
-\documentclass[a4paper]{article}
+\documentclass{article}
+
+\usepackage[letterpaper, margin=1in]{geometry}
 
@@
 \usepackage{xurl}
+
+% tables
 \usepackage{longtable,booktabs,array}
@@
 an accelerated longitudinal cohort of early development
-\citepp{forrestAdvancingScienceChildren2018}
+\citep{forrestAdvancingScienceChildren2018}
 between the ages of 82 days and 10 years
@@
 \textbf{Table1: Subject demographics}
+\begin{longtable}{|r|r|r|r|r|}
+    \hline\hline
+    \textbf{group} & \textbf{subgroup} & \textbf{all} & \textbf{under6mo} & \textbf{over18mo} \\\hline
+    \endhead
+    \hline\hline
+    \endfoot
+    \endlastfoot
+    N subjects &  & 361 & 74 & 257 \\  \hline
+    \multirow{3}{*}{Samples} & 1 & 262 (72.6\%) & 74 (100.0\%) & 203 (79.0\%) \\ \cline{2-5}
+         & 2 & 72 (19.9\%) & 0 (0.0\%) & 46 (17.9\%) \\ \cline{2-5}
+         & \textgreater 2 & 27 (7.5\%) & 0 (0.0\%) & 8 (3.1\%) \\ \hline
+    \multirow{3}{*}{Age (months)} & min & 2.76 & 2.76 & 18.03 \\ \cline{2-5}
+         & max & 119.3 & 6.0 & 119.3 \\ \cline{2-5}
+         & median & 25.79 & 3.66 & 45.92 \\ \hline
+    \multirow{2}{*}{Sex} & F & 163 (45.2\%) & 37 (50.0\%) & 114 (44.4\%) \\   \cline{2-5}
+                         & M & 198 (54.8\%) & 37 (50.0\%) & 143 (55.6\%) \\  \hline
+    \multirow{6}{*}{Race} & White & 252 (69.8\%) & 42 (56.8\%) & 191 (74.3\%) \\   \cline{2-5}
+        & Black & 40 (11.1\%) & 16 (21.6\%) & 21 (8.2\%) \\ \cline{2-5}
+        & Indiginous & 1 (0.3\%) & 0 (0.0\%) & 1 (0.4\%) \\ \cline{2-5}
+        & Asian & 7 (1.9\%) & 1 (1.4\%) & 6 (2.3\%) \\ \cline{2-5}
+        & Mixed & 48 (13.3\%) & 12 (16.2\%) & 29 (11.3\%) \\ \cline{2-5}
+        & Other & 2 (0.6\%) & 1 (1.4\%) & 2 (0.8\%) \\ \hline
+    \multirow{6}{*}{Maternal Ed.} & Junior high school & 1 (0.3\%) & 0 (0.0\%) & 1 (0.4\%) \\   \cline{2-5}
+        & Some high school & 6 (1.7\%) & 3 (4.1\%) & 4 (1.6\%) \\ \cline{2-5}
+        & High school grad & 37 (10.2\%) & 13 (17.6\%) & 16 (6.2\%) \\ \cline{2-5}
+        & Some college & 95 (26.3\%) & 25 (33.8\%) & 57 (22.2\%) \\ \cline{2-5}
+        & College grad & 90 (24.9\%) & 17 (23.0\%) & 69 (26.8\%) \\ \cline{2-5}
+        & Grad/professional school & 123 (34.1\%) & 15 (20.3\%) & 104 (40.5\%) \\\hline\hline
+    \label{Table 1}
+  \end{longtable}
 

### 2023-02-22 manuscript/Table2.jl

@@ -47,12 +47,11 @@ end
 
 
 CSV.write(tablefiles("Table2.csv"), fseaout)
-h1 = Highlighter((data, i, j) -> j in (3,5,7) && data[i,j] < 0.2 && data[i,j-1] < 0; bold=true, foreground=:red)
-h2 = Highlighter((data, i, j) -> j in (3,5,7) && data[i,j] < 0.2 && data[i,j-1] > 0; bold=true, foreground=:blue)
+h1 = LatexHighlighter((data, i, j) -> j in (3,5,7) && data[i,j] < 0.2 && data[i,j-1] < 0; ["color{red}","textbf"])
+h2 = LatexHighlighter((data, i, j) -> j in (3,5,7) && data[i,j] < 0.2 && data[i,j-1] > 0; ["color{blue}","textbf"])
 
 pretty_table(fseaout; highlighters=(h1,h2),
     header = (map(n-> replace(replace(n, r"_.+" => ""), "geneset"=>"Age group"), names(fseaout)),
               map(n-> replace(replace(n, r".+_enrichment" => "E.S."), r".+_qvalue"=>"q value"), names(fseaout))
-    )
-    
+    ), backend=Val(:latex)
     )

### 2023-02-23 manuscript/main.tex
tweaks to table 2, figure legends

@@
 \usepackage{multirow}
+\usepackage{afterpage}
 
@@
 
-\subsubsection{Cohort overview and summary data}
+\subsubsection*{Cohort overview and summary data}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{assets/Figure1.png}
+    \includegraphics[width=0.9\textwidth]{assets/Figure1.png}
     \caption{
@@
     }
-    \label{fig:Figure 1}
+    \label{fig:1}
 \end{sidewaysfigure}
 
-\textbf{Table1: Subject demographics}
-\begin{longtable}{|r|r|r|r|r|}
+\begin{longtable}[htb!]{|r|r|r|r|r|}
+    \caption{\label{tab:demo}Subject demographics} \\
     \hline\hline
@@
         & Grad/professional school & 123 (34.1\%) & 15 (20.3\%) & 104 (40.5\%) \\\hline\hline
-    \label{Table 1}
 \end{longtable}
@@
 
-\subsubsection{Microbial species and neuroactive genes are associated with cognitive performance}
+\subsubsection*{Microbial species and neuroactive genes are associated with cognitive performance}
 
@@
         Taxa and gene functional groups are associated with cognitive function.
-        (A) Volcano plot of multivariable linear model results showing the
+        (\textbf{A}) Volcano plot of multivariable linear model results showing the
         relationship between individual taxa and cognitive function score in
@@
         and sequencing depth. Taxa that were significant after FDR correction (q
-        \textless{} 0.2) are colored red. (B) For taxa that were significantly
+        \textless{} 0.2) are colored red. (\textbf{B}) For taxa that were significantly
         associated with cognitive function, heatmaps of prevalence, mean
         relative abundance, and correlation with cognitive function in different
-        age groups. (C-F) Enrichment plots for selected neuroactive gene sets
-        used in feature set enrichment analysis (FSEA). (G) summary of FSEA
+        age groups. (\textbf{C-F}) Enrichment plots for selected neuroactive gene sets
+        used in feature set enrichment analysis (FSEA). (\textbf{G}) summary of FSEA
         results across all samples (left) as well as in the under 6 months
@@
     }
-    \label{fig:Figure 2}
+    \label{fig:2}
 \end{sidewaysfigure}
@@
 
-\textbf{Table 2 - Feature set enrichment analysis on neuroactive microbial gene sets}
-
-\subsubsection{Gut microbial taxonomic and functional profiles predict cognitive function}
+\begin{table}[!h]
+    \begin{center}
+    \begin{tabular}{|r|r|r|r|r|r|r|}
+      \hline\hline
+      \textbf{Age group} & \multicolumn{2}{c}{\textbf{00to120}} & \multicolumn{2}{c}{\textbf{00to06}} & \multicolumn{2}{c}{\textbf{18to120}} \\ \hline
+      \texttt{geneset} & \texttt{E.S.} & \texttt{q value} & \texttt{E.S.} & \texttt{q value} & \texttt{E.S.} & \texttt{q value} \\\hline
+      Acetate syn. I & -0.122 & 0.555 & -0.194 & \color{red}{\textbf{0.153}} & 0.127 & 0.602 \\
+      Acetate syn. II & -0.22 & \color{red}{\textbf{0.066}} & -0.342 & \color{red}{\textbf{0.02}} & -0.095 & 0.92 \\
+      Acetate syn. III & -0.21 & \color{red}{\textbf{0.086}} & -0.31 & \color{red}{\textbf{0.052}} & -0.086 & 0.92 \\
+      Glutamate deg. I & -0.404 & \color{red}{\textbf{0.18}} & -0.284 & 0.816 & -0.249 & 0.602 \\
+      Glutamate syn. II & 0.095 & 0.602 & -0.134 & 0.98 & 0.242 & \color{blue}{\textbf{0.047}} \\
+      Menaquinone syn. I & -0.085 & 0.563 & -0.083 & 0.816 & -0.17 & \color{red}{\textbf{0.182}} \\
+      Propionate deg. I & -0.461 & \color{red}{\textbf{0.13}} & -0.542 & \color{red}{\textbf{0.02}} & -0.674 & \color{red}{\textbf{0.041}} \\
+      Propionate syn. I & -0.334 & 0.555 & -0.32 & 0.571 & -0.676 & \color{red}{\textbf{0.023}} \\
+      Tryptophan syn. & 0.057 & 0.974 & -0.072 & 0.816 & 0.119 & \color{blue}{\textbf{0.041}} \\\hline\hline
+    \end{tabular}
+    \caption{\label{tab:fsea}Feature set enrichment analysis on neuroactive microbial gene sets}
+    \end{center}
+\end{table}
+
+\subsubsection*{Gut microbial taxonomic and functional profiles predict cognitive function}
 
@@
 
-\begin{sidewaysfigure}[h]
+\begin{sidewaysfigure}[!htb]
     \centering
@@
         Comparison of RF predictor importance versus linear models for children
-        between birth and 6 months old (A), and for those older than 18 months
-        (B). Colors represent whether the species belong to the group of
+        between birth and 6 months old (\textbf{A}), and for those older than 18 months
+        (\textbf{B}). Colors represent whether the species belong to the group of
         top-important features that account for 80\% of the cumulative
@@
         importance for taxa in RF models for children between birth and 6 months
-        old (C), and for those older than 18 months (D). Taxa that are important
+        old (\textbf{C}), and for those older than 18 months (\textbf{D}). Taxa that are important
         for RF models both for children under 6 months and for those over 18
-        months (E), only for children under 6 months (F), and only for children
-        over 18 months (G)
+        months (\textbf{E}), only for children under 6 months (\textbf{F}), and only for children
+        over 18 months (\textbf{G})
     }
-    \label{fig:Figure 3}
+    \label{fig:3}
 \end{sidewaysfigure}
@@
 
-\subsubsection{Gut microbial taxonomic profiles predict brain structure differences}
+\subsubsection*{Gut microbial taxonomic profiles predict brain structure differences}
 
@@
         Microbial feature importance in predicting brain volumes in children over 18 months of age.
-        (A) Average test-set correlations for prediction of MRI segmentation
+        (\textbf{A}) Average test-set correlations for prediction of MRI segmentation
         data from microbiome and demographics on select segments after repeated
-        cross validation. (B) Heatmap of average individual relative taxa
+        cross validation. (\textbf{B}) Heatmap of average individual relative taxa
         importances on each brain segment. Importances are reported as
@@
         by HCA on a select list of species with high ranks on average importances,
-        and their respective highest-load segments (C).
+        and their respective highest-load segments (\textbf{C}).
     }
-    \label{fig:Figure 4}
+    \label{fig:4}
 \end{sidewaysfigure}
@@
 
-\subsubsection{Study Ethics}
+\subsubsection*{Study Ethics}
 
@@
 
-\subsubsection{Participants}
+\subsubsection*{Participants}
 
@@
 
-\subsubsection{Cognitive Assessments}
+\subsubsection*{Cognitive Assessments}
 
@@
 
-\subsubsection{Stool Sample Collection and Sequencing}
+\subsubsection*{Stool Sample Collection and Sequencing}
 
@@
 
-\subsubsection{Machine learning for cognitive development}
+\subsubsection*{Machine learning for cognitive development}
 
@@
 
-\subsubsection{MRI / segmentation}
+\subsubsection*{MRI / segmentation}
 
@@
 
-\subsubsection{Computational analysis of metagenomes}
+\subsubsection*{Computational analysis of metagenomes}
 
@@
 
-\subsubsection{Data and code availability}
+\subsubsection*{Data and code availability}
 

### 2023-02-23 manuscript/Table2.jl

@@ -47,10 +47,11 @@ end
 
 
 CSV.write(tablefiles("Table2.csv"), fseaout)
-h1 = LatexHighlighter((data, i, j) -> j in (3,5,7) && data[i,j] < 0.2 && data[i,j-1] < 0; ["color{red}","textbf"])
-h2 = LatexHighlighter((data, i, j) -> j in (3,5,7) && data[i,j] < 0.2 && data[i,j-1] > 0; ["color{blue}","textbf"])
+formatter = (v, i, j) -> v isa AbstractFloat ? round(v, digits = 3) : v
+h1 = LatexHighlighter((data, i, j) -> j in (3,5,7) && data[i,j] < 0.2 && data[i,j-1] < 0, ["color{red}","textbf"])
+h2 = LatexHighlighter((data, i, j) -> j in (3,5,7) && data[i,j] < 0.2 && data[i,j-1] > 0, ["color{blue}","textbf"])
 
-pretty_table(fseaout; highlighters=(h1,h2),
+pretty_table(fseaout; highlighters=(h1,h2), formatters=ft_round(3),
     header = (map(n-> replace(replace(n, r"_.+" => ""), "geneset"=>"Age group"), names(fseaout)),
               map(n-> replace(replace(n, r".+_enrichment" => "E.S."), r".+_qvalue"=>"q value"), names(fseaout))
     ), backend=Val(:latex)

### 2023-02-23 manuscript/main.tex
update title / authors, fix table 1

@@
 \usepackage{lineno}
+\linenumbers
 \usepackage{titlesec}
@@
 backend=biber,
-natbib=true,
-sortcites=true,
-defernumbers=true,
-style=authoryear,
-citestyle=authoryear-comp,
-maxnames=99,
-maxcitenames=2,
-giveninits=true,
-terseinits=true,
-url=false
+style=numeric
 ]{biblatex}
@@
 \graphicspath{assets}
+\usepackage{times}
+
+\newenvironment{sciabstract}{%
+\begin{quote} \bf}
+{\end{quote}}
+
+% Include your paper's title here
 
 \title{Gut microbes and their genes or something else}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%  SET THE LIST OF AUTHORS WITH CORRESPONDING AFFILIATIONS  use \& before last author %%%%%%%%%%%%%%%%%%%%
-\author{Kevin S. Bonham\textsuperscript{1}, \space
-Guilherme Fahur Bottino\textsuperscript{1}, 
-Firstname Lastname\textsuperscript{2}, \&
-Vanja Klepac-Ceraj\textsuperscript{1}
-%\fourthauthor \textsuperscript{i}
-%etc...
+
+% Place the author information here.  Please hand-code the contact
+% information and notecalls; do *not* use \footnote commands.  Let the
+% author contact information appear immediately below the author names
+% as shown.  We would also prefer that you don't change the type-size
+% settings shown here.
+
+\author{%
+    \parbox{\linewidth}{\centering
+        Kevin S. Bongham,$^{1}$
+        Guilherme Fahur Botino,$^{1}$
+        Shelley Hoeft McCann,$^{1}$
+        Jennifer Beauchemin,$^{2}$
+        Elizabeth Weisse,$^{3}$
+        Fatoumata Barry,$^{2}$
+        Rosa Cano Lorente,$^{2}$
+        The RESONANCE Consortium,
+        Jonathan O'Muircheartaigh,$^{3}$
+        Curtis Huttenhower,$^{4}$
+        Muriel Bruchhage,$^{3}$
+        Viren D’Sa,$^{2}$
+        Sean Deoni,$^{2}$
+        Vanja Klepac-Ceraj$^{1\ast}$
     }
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE ABSTRACT %%%%%%%%%%%%%%%%%
+\\
+\normalsize{$^{1}$Department of Biological Sciences, Wellesley College,}\\
+\normalsize{106 Central St, Wellesley, MA \hl{02---}, USA}\\
+\normalsize{$^{2}$Rhode Island Hospital, Providence, RI, \hl{00000} USA}\\
+\normalsize{$^{3}$\hl{Locations for Muriel, Jonathan, Sean}}\\
+\normalsize{$^{4}$\hl{Department of Biostatistics, Harvard T.H. Chan School of Public Health}}\\
+\normalsize{\hl{address of HSPH...}}
+\\
+\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  vklepacc@wellesley.edu.}
+}
+
+% Include the date command, but leave its argument blank.
+
+\date{}
 
 \begin{document}
+
+\baselineskip24pt
+
+% Make the title.
+
+\maketitle
 
@@
 nerve and through endocrine and immune (cytokine) signaling molecules
-\citep{cerdoEarlyNutritionGut2019,pronovostPerinatalInteractionsMicrobiome2019,sharonCentralNervousSystem2016,togniniGutMicrobiotaPotential2017}
+\cite{cerdoEarlyNutritionGut2019,pronovostPerinatalInteractionsMicrobiome2019,sharonCentralNervousSystem2016,togniniGutMicrobiotaPotential2017}
 In addition, the products of microbial metabolism generated in the gut can
@@
 development, are increasingly being identified
-\citep{spichakMiningMicrobesMental2021}.
+\cite{spichakMiningMicrobesMental2021}.
 Both human epidemiology and animal models point to the effects
 of gut microbes on the development of autism spectrum disorder
-\citep{laueProspectiveAssociationsInfant2020,wanUnderdevelopmentGutMicrobiota2021}
+\cite{laueProspectiveAssociationsInfant2020,wanUnderdevelopmentGutMicrobiota2021}
 and particular microbial taxa have been associated with depression
-\citep{mayneris-perxachsMicrobiotaAlterationsProline2022,valles-colomerNeuroactivePotentialHuman2019}
+\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,valles-colomerNeuroactivePotentialHuman2019}
 and Alzheimer's disease
-\citep{fungInteractionsMicrobiotaImmune2017,kimProbioticSupplementationImproves2021}.
+\cite{fungInteractionsMicrobiotaImmune2017,kimProbioticSupplementationImproves2021}.
 But information about this "microbiome-gut-brain
@@
 microbiome and the brain
-\citep{laueDevelopingMicrobiomeBirth2022}.
+\cite{laueDevelopingMicrobiomeBirth2022}.
 Fetal development \emph{in utero} is believed to be mostly sterile, but
@@
 and other environmental sources such as antibiotics
-\citep{backhedDynamicsStabilizationHuman2015,bokulichAntibioticsBirthMode2016}.
+\cite{backhedDynamicsStabilizationHuman2015,bokulichAntibioticsBirthMode2016}.
 The early microbiome is characterized by low microbial
@@
 Bacteroidetes, especially \emph{Bacteroides}, and Proteobacteria
-\citep{koenigSuccessionMicrobialConsortia2011}.
+\cite{koenigSuccessionMicrobialConsortia2011}.
 Many of these microbes have specialized metabolic capabilities
@@
 and \emph{Bacteroides fragilis}
-\citep{selaGenomeSequenceBifidobacterium2008}.
+\cite{selaGenomeSequenceBifidobacterium2008}.
 Upon the introduction of solid foods, the gut
@@
 adult microbiomes
-\citep{backhedDynamicsStabilizationHuman2015}.
+\cite{backhedDynamicsStabilizationHuman2015}.
 Many prior studies have focused on either infant microbiomes or
@@
 incredibly important
-\citep{tauNormalDevelopmentBrain2010}.
+\cite{tauNormalDevelopmentBrain2010}.
 
@@
 connections has been established
-\citep{silbereisCellularMolecularLandscapes2016}.
+\cite{silbereisCellularMolecularLandscapes2016}.
 Much of this development occurs in discrete windows or
 ``sensitive-'' or ``critical periods'' (CPs)
-\citep{knudsenSensitivePeriodsDevelopment2004}
+\cite{knudsenSensitivePeriodsDevelopment2004}
 when neural plasticity is particularly high,
@@
 environment, including the gut microbiome
-\citep{cowanAnnualResearchReview2020}.
+\cite{cowanAnnualResearchReview2020}.
 In fact, emerging evidence suggests that the timing and duration of CPs
 may be driven in part by cues from the developing gut microbiome
-\citep{callaghanNestedSensitivePeriods2020}.
+\cite{callaghanNestedSensitivePeriods2020}.
 As such,
@@
 an accelerated longitudinal cohort of early development
-\citep{forrestAdvancingScienceChildren2018}
+\cite{forrestAdvancingScienceChildren2018}
 between the ages of 82 days and 10 years
@@
 for children from birth to 3 years of age
-\citep{mullenMullenScalesEarly1995},
+\cite{mullenMullenScalesEarly1995},
 Wechsler preschool and primary scale of intelligence (WPPSI)
 for 4-5 year-olds
-\citep{wechslerWechslerPreschoolPrimary2012},
+\cite{wechslerWechslerPreschoolPrimary2012},
 and the Wechsler intelligence scale for children (WISC)
 for children 6 years and up
-\citep{wechslerWechslerIntelligenceScale1949}.
+\cite{wechslerWechslerIntelligenceScale1949}.
 
@@
 
-\begin{longtable}[htb!]{|r|r|r|r|r|}
-    \caption{\label{tab:demo}Subject demographics} \\
+\begin{table}[!htb]
+    \centering
+    \begin{tabular}{|r|r|r|r|r|}
       \hline\hline
       \textbf{group} & \textbf{subgroup} & \textbf{all} & \textbf{under6mo} & \textbf{over18mo} \\\hline
-    \endhead
-    \hline\hline
-    \endfoot
-    \endlastfoot
       N subjects &  & 361 & 74 & 257 \\  \hline
@@
         & Grad/professional school & 123 (34.1\%) & 15 (20.3\%) & 104 (40.5\%) \\\hline\hline
-\end{longtable}
+    \end{tabular}
+    \caption{\label{tab:demo}Subject demographics}
+\end{table}
 
@@
 measures of anxiety and depression
-\citep{mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022},
+\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022},
 cognitive flexibility in adults
-\citep{magnussonRelationshipsDietrelatedChanges2015},
+\cite{magnussonRelationshipsDietrelatedChanges2015},
 and atypical neural development
-\citep{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021}.
+\cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021}.
 We, therefore, set out to identify whether specific taxa
@@
 profiles (annotated with clusters of 90\% similarity
-\citep{suzekUniRefComprehensiveNonredundant2007}
+\cite{suzekUniRefComprehensiveNonredundant2007}
 ), both generated from DNA sequencing data, being tightly coupled
@@
 groupings (Enzyme commission level4 - ECs
-\citep{bairochENZYMEDatabase20002000},
+\cite{bairochENZYMEDatabase20002000},
 and
 KEGG orthologues - KOs
-\citep{kanehisaKEGGResourceDeciphering2004}
+\cite{kanehisaKEGGResourceDeciphering2004}
 ) overlapped only slightly with taxonomic profiles in both age
@@
 cognitive function, we fit multivariable linear regression
-\citep{mallickMultivariableAssociationDiscovery2021}
+\cite{mallickMultivariableAssociationDiscovery2021}
 to the relative abundance of each species that had at least 15\%
@@
 previously shown to be depleted in children with diabetes
-\citep{benitez-paezDepletionBlautiaSpecies2020},
+\cite{benitez-paezDepletionBlautiaSpecies2020},
 and that oral administration of \emph{B. wexlerae} partially
@@
 model of T2D
-\citep{hosomiOralAdministrationBlautia2022,liuBlautiaNewFunctional2021}.
+\cite{hosomiOralAdministrationBlautia2022,liuBlautiaNewFunctional2021}.
 In children over 18 months of age,
@@
 neuroprotective metabolite urolithin
-\citep{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014},
+\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014},
 \emph{Asaccharobacter celatus} and
@@
 equol
-\citep{maruoAdlercreutziaEquolifaciensGen2008,thawornkunoBiotransformationDaidzeinEquol2009},
+\cite{maruoAdlercreutziaEquolifaciensGen2008,thawornkunoBiotransformationDaidzeinEquol2009},
 and the SCFA-producing probiotic
@@
 prausnitzii}
-\citep{ghoshMediterraneanDietIntervention2020}
+\cite{ghoshMediterraneanDietIntervention2020}
 (Figure 2B).
@@
 of genes with neuroactive potential
-\citep{valles-colomerNeuroactivePotentialHuman2019}
+\cite{valles-colomerNeuroactivePotentialHuman2019}
 and concurrent cognitive function score (Table 2, Figure 2C-G) and found that
@@
 and have been linked with immune system regulation as well as directly
-with brain function \citep{dalileRoleShortchainFatty2019}.
+with brain function \cite{dalileRoleShortchainFatty2019}.
 
@@
 which, MK-4, has been found to be decreased in children diagnosed with ASD
-\citep{dongCorrelationSerumConcentrations2021}
+\cite{dongCorrelationSerumConcentrations2021}
 and is potentially neuroprotective in both rodents and humans
-\citep{elkattawyVitaminK2Menaquinone72022}.
+\cite{elkattawyVitaminK2Menaquinone72022}.
 In children over 18 months old,
@@
 the developing brain
-\citep{cohenkadoshLinkingGABAGlutamate2015,palomo-buitragoGlutamateInteractionsObesity2019}.
+\cite{cohenkadoshLinkingGABAGlutamate2015,palomo-buitragoGlutamateInteractionsObesity2019}.
 Tryptophan metabolism, including
@@
 autism in children
-\citep{hoshinoBloodSerotoninFree1984,xiaoFecalMicrobiomeTransplantation2021}.
+\cite{hoshinoBloodSerotoninFree1984,xiaoFecalMicrobiomeTransplantation2021}.
 Taken together, these results suggest that
@@
 from the domain of genomics
-\citep{amaratungaEnrichedRandomForests2008,brieucPracticalIntroductionRandom2018,chenRandomForestsGenomic2012,franzosaGutMicrobiomeStructure2019,stephanRandomForestApproach2015},
+\cite{amaratungaEnrichedRandomForests2008,brieucPracticalIntroductionRandom2018,chenRandomForestsGenomic2012,franzosaGutMicrobiomeStructure2019,stephanRandomForestApproach2015},
 along with other works where it was used
@@
 different scenarios
-\citep{ardekaniPredictionIncipientAlzheimer2017,velazquezRandomForestModel2021}.
+\cite{ardekaniPredictionIncipientAlzheimer2017,velazquezRandomForestModel2021}.
 Additionally, given the sequential
@@
 area, which plays an important role in reward circuits
-\citep{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012},
+\cite{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012},
 has one of the highest test-set correlations of our
@@
 been linked to substance use disorder in young adults
-\citep{caoMappingCorticalSubcortical2021}.
+\cite{caoMappingCorticalSubcortical2021}.
 
@@
 it has been shown to have a divergent neuromorphology
-\citep{hoogmanSubcorticalBrainVolume2017}.
+\cite{hoogmanSubcorticalBrainVolume2017}.
 Independently, \emph{B. ovatus}, \emph{B. uniformis} and
 \emph{B. vulgatus} have been linked to ADHD
-\citep{wangGutMicrobiotaDietary2020}.
+\cite{wangGutMicrobiotaDietary2020}.
 
@@
 attention deficit
-\citep{yaelDisinhibitionNucleusAccumbens2019},
+\cite{yaelDisinhibitionNucleusAccumbens2019},
 and that \emph{B. uniformis} is gut-microbial modulator of the brain
 dopamine transporter
-\citep{hartstraInfusionDonorFeces2020}.
+\cite{hartstraInfusionDonorFeces2020}.
 
@@
 associated with cognitive function
-\citep{kantGenomeSequenceButyrateProducing2015,liCorrelationsGutMicrobiota2022}.
+\cite{kantGenomeSequenceButyrateProducing2015,liCorrelationsGutMicrobiota2022}.
 Its importance is, however, heavily loaded on the volume
@@
 parahippocampal gyrus
-\citep{pagonabarragaPatternRegionalCortical2013}.
+\cite{pagonabarragaPatternRegionalCortical2013}.
 Additionally, the left and right parahippocampal, where \emph{B.
@@
 visual/spatial processing and memory
-\citep{aminoffRoleParahippocampalCortex2013}.
+\cite{aminoffRoleParahippocampalCortex2013}.
 \emph{B. wexlerae} can also produce acetylcholine in the gut
-\citep{hosomiOralAdministrationBlautia2022},
+\cite{hosomiOralAdministrationBlautia2022},
 and this molecule plays an important role in modulating memory
 function
-\citep{haamCholinergicModulationHippocampal2017}.
+\cite{haamCholinergicModulationHippocampal2017}.
 
@@
 depression symptoms
-\citep{wackerRoleNucleusAccumbens2009},
+\cite{wackerRoleNucleusAccumbens2009},
 and increased levels of the \emph{Alistipes} genus have been
 observed in patients with depressive disorder
-\citep{jiangAlteredFecalMicrobiota2015}.
+\cite{jiangAlteredFecalMicrobiota2015}.
 While these reports are important, simultaneously probing of the gut
@@
 decision making
-\citep{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002}.
+\cite{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002}.
 Equol has a strong estrogenic effect
-\citep{setchellSEquolPotentLigand2005},
+\cite{setchellSEquolPotentLigand2005},
 and in anterior cingulate, estrogen, has been shown to regulate
 pain-related aversion
-\citep{xiaoEstrogenAnteriorCingulate2013}.
+\cite{xiaoEstrogenAnteriorCingulate2013}.
 
@@
 understanding that this species is increased in individuals with insulin
-resistance and obesity \citep{leyHumanGutMicrobes2006},
+resistance and obesity \cite{leyHumanGutMicrobes2006},
 conditions that are known to produce structural abnormalities in the brain
-\citep{opelBrainStructuralAbnormalities2021}.
+\cite{opelBrainStructuralAbnormalities2021}.
 Finally, \emph{Coprococcus comes} displays an importance
@@
 higher scores on the Ages and Stages Questionnaire
-\citep{sordilloAssociationInfantGut2019}.
+\cite{sordilloAssociationInfantGut2019}.
 However, each of these clades encompass dozens of genera with
@@
 (formerly regarded a subspecies of \emph{A. equolofaciens}
-\citep{takahashiCompleteGenomeSequence2021}),
+\cite{takahashiCompleteGenomeSequence2021}),
 and \emph{Eggerthella lenta}. Many members of this family are
@@
 found in soybeans
-\citep{wangEnantioselectiveSynthesisSEquol2005},
+\cite{wangEnantioselectiveSynthesisSEquol2005},
 and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
@@
 shown in some studies to have a neuroprotective effect
-\citep{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
+\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
 \emph{E. lenta} has been extensively studied for
@@
 medication digoxin
-\citep{haiserPredictingManipulatingCardiac2013}.
+\cite{haiserPredictingManipulatingCardiac2013}.
 The metabolic versatility of this clade, and the large number of
@@
 pronounced in some age groups due to our sampling schedule
-\citep{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
+\cite{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
 (Supplementary Figure 3).
@@
 on Child Health Outcomes (ECHO)
-\citep{forrestAdvancingScienceChildren2018,gillmanEnvironmentalInfluencesChild2018},
+\cite{forrestAdvancingScienceChildren2018,gillmanEnvironmentalInfluencesChild2018},
  a longitudinal observational study of
@@
 Composite as assessed via the Mullen Scales of Early Learning (MSEL)
-\citep{mullenMullenScalesEarly1995},
+\cite{mullenMullenScalesEarly1995},
  a standardized and population-normed tool for assessing fine and
@@
 Wechsler Intelligence Quotient for Children (WISC)
-\citep{wechslerWechslerPreschoolPrimary2012}
+\cite{wechslerWechslerPreschoolPrimary2012}
 is an individually administered standard intelligence test for children
@@
 Random Forests (RFs)
-\citep{breimanRandomForests2001}
+\cite{breimanRandomForests2001}
 were selected as the prediction engine and processed using the
-DecisionTree.jl \citep{sadeghiDecisionTreeJlJulia2022}
-implementation, inside the MLJ.jl \citep{blaomMLJJuliaPackage2020}.
+DecisionTree.jl \cite{sadeghiDecisionTreeJlJulia2022}
+implementation, inside the MLJ.jl \cite{blaomMLJJuliaPackage2020}.
 
@@
 experiment, using a set of default regression hyperparameters from
-Breiman and Cutler \citep{breimanRandomForests2001},
+Breiman and Cutler \cite{breimanRandomForests2001},
 on a repeated cross-validation approach with different RNG seeds. One hundred
@@
 on Random Forests with high-dimensional, low sample size microbiome
-datasets \citep{woodruffInflammationAutoreactivityDefine2022}.
+datasets \cite{woodruffInflammationAutoreactivityDefine2022}.
 The resulting fitness-weighted importances were used
@@
 image registration, we created representative age-specific templates
-using the ANTs package \citep{avantsInsightToolKitImage2014}.
+using the ANTs package \cite{avantsInsightToolKitImage2014}.
 After age specific templates were created, a single flow from
@@
 thalamus, putamen) and total grey and white matter volumes (included as
-part of the FSL package \citep{jenkinsonFSL2012}).
+part of the FSL package \cite{jenkinsonFSL2012}).
 Total tissue and brainstem volume as well as left and right
@@
 hippocampus, amygdala, and accumbens as well as total brainstem volume
-\citep{bruchhageLongitudinalBrainCognitive}.
+\cite{bruchhageLongitudinalBrainCognitive}.
 
@@
 et al. 2021)
-\citep{}.
+\cite{}.
  First, KneadData (v0.7.7) was used to perform quality
@@
 2022)
-\citep{}.
+\cite{}.
  Data processing, generation of summary statistics, and
@@
 2017; Bonham et al., 2021; Danisch and Krumbiegel, 2021)
-\citep{}.
+\cite{}.
  All code for

### 2023-02-28 manuscript/main.tex
update title / authors, fix table 1
Incorporated tables 3-5 into main.tex

@@
 
-\title{Gut microbes and their genes or something else}
+\title{Gut microbes and their genes are associated with cognition and neuroanatomy in children}
 
@@
 
+\begin{table}
+    \begin{tabular}{rrrrr}
+      \hline\hline
+      \textbf{Subject Ages (months)} & \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set correlation (± C.I.)} & \textbf{Test set Root-mean-square error (± C.I.)} \\\hline
+      0 to 6 & - & + & 13.01 ± 0.05 & -0.14 ± 0.01 \\
+      0 to 6 & taxa & - & 12.66 ± 0.05 & -0.1 ± 0.01 \\
+      0 to 6 & taxa & + & 12.7 ± 0.05 & -0.12 ± 0.01 \\
+      0 to 6 & genes & - & 12.56 ± 0.05 & -0.01 ± 0.01 \\
+      0 to 6 & genes & + & 12.56 ± 0.05 & -0.02 ± 0.01 \\
+      18 to 120 & - & + & 16.27 ± 0.04 & 0.506 ± 0.003 \\
+      18 to 120 & taxa & - & 17.66 ± 0.04 & 0.363 ± 0.003 \\
+      18 to 120 & taxa & + & 17.29 ± 0.04 & 0.429 ± 0.003 \\
+      18 to 120 & genes & - & 18.01 ± 0.04 & 0.303 ± 0.003 \\
+      18 to 120 & genes & + & 17.86 ± 0.04 & 0.333 ± 0.004 \\\hline\hline
+    \end{tabular}
+\end{table}
+
 As with linear models, RF models for children under 6 months old were
@@
 \textbf{Table 4. Summary statistics for the neuroanatomy prediction benchmarks}
+
+\begin{table}
+    \begin{tabular}{rrr}
+      \hline\hline
+      \textbf{Statistic} & \textbf{Mean absolute proportional error (MAPE)} & \textbf{Correlation coefficient (R)} \\\hline
+      Mean & 0.073 & 0.137 \\
+      Standard deviation & 0.028 & 0.137 \\
+      Maximum & 0.206 & 0.605 \\
+      75th percentile & 0.082 & 0.195 \\
+      50th percentile & 0.067 & 0.134 \\
+      25th percentile & 0.057 & 0.062 \\
+      Minimum & 0.035 & -0.147 \\\hline\hline
+    \end{tabular}
+\end{table}
 
@@
 Forest experiments}
+
+\begin{table}
+    \begin{tabular}{rrrr}
+      \hline\hline
+      \textbf{Input set} & \textbf{Age bracket} & \textbf{Microbiome encoding type} & \textbf{Demographics Provided? (sex, education)} \\\hline
+      1 & 0 to 6 months & Not provided & yes \\
+      2 & 0 to 6 months & Taxonomic profile & no \\
+      3 & 0 to 6 months & Taxonomic profile & yes \\
+      4 & 0 to 6 months & Functional Profile (ECs) & no \\
+      5 & 0 to 6 months & Functional Profile (ECs) & yes \\
+      6 & 18 to 120 months & Not provided & yes \\
+      7 & 18 to 120 months & Taxonomic profile & no \\
+      8 & 18 to 120 months & Taxonomic profile & yes \\
+      9 & 18 to 120 months & Functional Profile (ECs) & no \\
+      10 & 18 to 120 months & Functional Profile (ECs) & yes \\\hline\hline
+    \end{tabular}
+\end{table}
 

### 2023-03-01 manuscript/main.tex
fix up tables 3-5

--- 2023-02-23 manuscript/main.tex
@@
 
+\begin{table}
+    \begin{tabular}{|r|r|r|r|r|}
+      \hline\hline
+      \textbf{Subject Ages (months)} & \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set correlation (± C.I.)} & \textbf{Test set RMSE (± C.I.)} \\\hline
+      \multirow{5}{*}{0 to 6} & - & + & 13.01 ± 0.05 & -0.14 ± 0.01 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa} & - & 12.66 ± 0.05 & -0.1 ± 0.01 \\ \cline{3-5}
+            &       & + & 12.7 ± 0.05 & -0.12 ± 0.01 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 12.56 ± 0.05 & -0.01 ± 0.01 \\ \cline{3-5}
+            &       & + & 12.56 ± 0.05 & -0.02 ± 0.01 \\ \cline{1-5}
+      \multirow{5}{*}{18 to 120} & - & + & 16.27 ± 0.04 & 0.506 ± 0.003 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa} & - & 17.66 ± 0.04 & 0.363 ± 0.003 \\ \cline{3-5}
+            &       & + & 17.29 ± 0.04 & 0.429 ± 0.003 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 18.01 ± 0.04 & 0.303 ± 0.003 \\ \cline{3-5}
+            &       & + & 17.86 ± 0.04 & 0.333 ± 0.004 \\\hline\hline
+    \end{tabular}
+\end{table}
+
 As with linear models, RF models for children under 6 months old were
@@
 Forest experiments}
+
+\begin{table}
+    \begin{tabular}{rrrr}
+      \hline\hline
+      \textbf{Input set} & \textbf{Age bracket} & \textbf{Microbiome encoding type} & \textbf{Demographics Provided? (sex, education)} \\\hline
+      1 & \multirow{5}{*}{0 to 6 months} & Not provided & yes \\
+      2 &  & \multirow{2}{*}{Taxonomic profile} & no \\
+      3 &  &  & yes \\
+      4 &  & \multirow{2}{*}{Functional Profile (ECs)} & no \\
+      5 &  &  & yes \\
+      6 & \multirow{5}{*}{18 to 120 months} & Not provided & yes \\
+      7 &  & \multirow{2}{*}{Taxonomic profile} & no \\
+      8 &  &  & yes \\
+      9 &  & \multirow{2}{*}{Functional Profile (ECs)} & no \\
+      10 &  &  & yes \\\hline\hline
+    \end{tabular}
+\end{table}
 

### 2023-03-06 manuscript/main.tex
fix refs

--- 2023-02-23 manuscript/main.tex
@@
 \usepackage[utf8]{inputenc}
+
+\usepackage{scicite}
 % Clean citations with biblatex
@@
 backend=biber,
-style=numeric
+style=science
 ]{biblatex}
@@
 
+\topmargin 0.0cm
+\oddsidemargin 0.2cm
+\textwidth 16cm 
+\textheight 21cm
+\footskip 1.0cm
+
+%The next command sets up an environment for the abstract to your paper.
+
 \newenvironment{sciabstract}{%
@@
         The RESONANCE Consortium,
-        Jonathan O'Muircheartaigh,$^{3}$
         Curtis Huttenhower,$^{4}$
         Muriel Bruchhage,$^{3}$
-        Viren D’Sa,$^{2}$
+        Viren D'Sa,$^{2}$
         Sean Deoni,$^{2}$
@@
 \\
-\normalsize{$^{1}$Department of Biological Sciences, Wellesley College,}\\
-\normalsize{106 Central St, Wellesley, MA \hl{02---}, USA}\\
-\normalsize{$^{2}$Rhode Island Hospital, Providence, RI, \hl{00000} USA}\\
-\normalsize{$^{3}$\hl{Locations for Muriel, Jonathan, Sean}}\\
-\normalsize{$^{4}$\hl{Department of Biostatistics, Harvard T.H. Chan School of Public Health}}\\
-\normalsize{\hl{address of HSPH...}}
+\normalsize{$^{1}$Department of Biological Sciences, Wellesley College, Wellesley, MA, USA}\\
+\normalsize{$^{2}$Rhode Island Hospital, Providence, RI, USA}\\
+\normalsize{$^{3}$Department of Psychology, University of Stavenger, Stavanger, Norway}\\
+\normalsize{$^{4}$Department of Biostatistics, Harvard T.H. Chan School of Public Health, Boston, MA, USA}\\
 \\
@@
 % Make the title.
-
 \maketitle
@@
 overall cognitive function as well as with differences in the size of
-multiple brain regions. Using a combination of multivariate linear
+multiple brain regions.
+Using a combination of multivariate linear
 models and machine learning (ML) models, we showed that many species,
@@
 neuroactive compounds, particularly short-chain fatty acids such as
-glutamate and propionate, were also associated with cognitive function.
+acetate and propionate, were also associated with cognitive function.
 In addition, ML models were able to use microbial taxa to predict the
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%% Text of the preprint %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% use \emph{} for italics and \parencite{} to cite a reference and \ref{XX} to cite the \label{XX}
-% use \\ and a blank line to mark the end of paragraph and starting of a new one
+% Teaser
+% This one-sentence summary provides a snapshot of your research for
+% non-specialist readers and will be used to promote your article if accepted for publication.
+% These should complement rather than repeat the title.
+
+% The microorganisms that colonize the gut of humans can influence
+% health and disease in numerous
 
@@
     \caption{
-        Study design and overview. (\textbf{A}) Stool samples, cognitive assessments and
+        \label{fig:1} Study design and overview. (\textbf{A}) Stool samples, cognitive assessments and
         neuroimaging were collected from participants in an accelerated
@@
     }
-    \label{fig:1}
+
 \end{sidewaysfigure}
@@
 function in children under 6 months, though this may be due to the low
-taxonomic diversity, and broad lack of overlap between taxa in infants.
+taxonomic diversity and broad lack of overlap between taxa in infants.
 As expected, age was significantly associated with microbial beta
-diversity (taxa R\textsuperscript{2} = 0.0215, and UniRef90s
+diversity (taxa R\textsuperscript{2} = 0.0215, and  gene functions
+annotated with UniRef90s - clusters of 90\% similarity
+\cite{suzekUniRefComprehensiveNonredundant2007} -
 R\textsuperscript{2} = 0.0245, q \textless{} 0.001), and very strongly
-associated with neuroimaging profiles (R\textsuperscript{2} = 0.256, q
+associated with variation in neuroimaging profiles (R\textsuperscript{2} = 0.256, q
 \textless{} 0.001).
@@
 captured overlapping variation, with species profiles and gene function
-profiles (annotated with clusters of 90\% similarity
-\cite{suzekUniRefComprehensiveNonredundant2007}
-), both generated from DNA sequencing data, being tightly coupled
+profiles, both generated from DNA sequencing data, being tightly coupled
 (Figure 1H, p \textless{} 0.001). Interestingly, other functional
@@
 To assess whether individual microbial species were associated with
-cognitive function, we fit multivariable linear regression
+cognitive function, we fit multivariable linear regression models
 \cite{mallickMultivariableAssociationDiscovery2021}
-to the relative abundance of each species that had at least 15\%
+for the relative abundance of each species that had at least 15\%
 prevalence in a given age group (Figure 2A, N = 92 for 0--120 months, N
@@
 considering all age groups together, though the enrichment of most
-pathways was more pronounced in children under 6 months or over 18
+pathways was more pronounced either in children under 6 months or those over 18
 months. For example, genes for degrading the 3-carbon SCFA propionate
@@
 
-\textbf{Table 3. Benchmark metrics for the cognitive assessment score
+\begin{table}[!h]
+    \begin{center}
+    \begin{tabular}{|r|r|r|r|r|}
+      \hline\hline
+      \textbf{Subject Ages (months)} & \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
+      \multirow{5}{*}{0 to 6} & - & + & 13.01 ± 0.05 & -0.14 ± 0.01 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa} & - & 12.66 ± 0.05 & -0.1 ± 0.01 \\ \cline{3-5}
+            &       & + & 12.7 ± 0.05 & -0.12 ± 0.01 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 12.56 ± 0.05 & -0.01 ± 0.01 \\ \cline{3-5}
+            &       & + & 12.56 ± 0.05 & -0.02 ± 0.01 \\ \cline{1-5}
+      \multirow{5}{*}{18 to 120} & - & + & 16.27 ± 0.04 & 0.506 ± 0.003 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa} & - & 17.66 ± 0.04 & 0.363 ± 0.003 \\ \cline{3-5}
+            &       & + & 17.29 ± 0.04 & 0.429 ± 0.003 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 18.01 ± 0.04 & 0.303 ± 0.003 \\ \cline{3-5}
+            &       & + & 17.86 ± 0.04 & 0.333 ± 0.004 \\\hline\hline
+    \end{tabular}
+    \caption{\label{tab:rfbench}Benchmark metrics for the cognitive assessment score
     prediction models. Confidence intervals are calculated from the
-distribution of metrics from repeated CV at a confidence level of 95\%}
+    distribution of metrics from repeated CV at a confidence level of 95.}
+    \end{center}
+\end{table}
 
@@
 significant when testing the relationship with linear models (Figure
-3A-B, Supplementary Table XXA-B), though there were substantial
+3A-B, Supplementary Table \hl{XX}A-B), though there were substantial
 differences. For children under 6 months, only \emph{B. wexlerae} was
@@
 predicted by RF models trained on microbial taxa (Table 4, Supplementary
-Table XX), in particular those that were highly correlated with age.
+Table \hl{XX}), in particular those that were highly correlated with age.
 These included the L/R lingual gyrus (mean RF correlations, Left =
@@
 
-\textbf{Table 4. Summary statistics for the neuroanatomy prediction benchmarks}
+\begin{table}[!h]
+    \begin{center}
+    \begin{tabular}{|r|r|r|}
+      \hline\hline
+      \textbf{Statistic} & \textbf{(MAPE)} & \textbf{R} \\\hline
+      Mean & 0.073 & 0.137 \\
+      Standard deviation & 0.028 & 0.137 \\
+      Maximum & 0.206 & 0.605 \\
+      75th percentile & 0.082 & 0.195 \\
+      50th percentile & 0.067 & 0.134 \\
+      25th percentile & 0.057 & 0.062 \\
+      Minimum & 0.035 & -0.147 \\\hline\hline
+    \end{tabular}
+    \caption{\label{tab:neuro}Summary statistics for the neuroanatomy prediction benchmarks.
+    Mean absolute proportional error (MAPE) and correlation coeficient (R)
+    of test-set preditions are reported accross models for all brain regions.}
+    \end{center}
+\end{table}
 
@@
 \cite{wangGutMicrobiotaDietary2020}.
-
 In fact, there are studies showing that alterations on the striatal
@@
 
-Many of the taxa identified included taxa also identified in LMs and RF
-models of cognitive function. Interesting to note, while RF models for
+Many of the taxa identified as important for models predicting brain structure
+were also identified in LMs and RF models of cognitive function. 
+Interestingly, while RF models for
 multiple brain regions had many important microbial taxa, others were
@@
 contribute heavily to the entorhinal, fusiform, lingual and
-parahippocampal segments.
-
-This group of taxa and their related segments drew our attention because
+parahippocampal segments. This group of taxa and their related segments drew our attention because
 they contained \emph{B. wexlerae}, a highlight from the cognitive
@@
 \cite{pagonabarragaPatternRegionalCortical2013}.
-Additionally, the left and right parahippocampal, where \emph{B.
+Additionally, the left and right parahippocampus, where \emph{B.
 wexlerae} had the highest importance in both models, are important in
@@
 \cite{hosomiOralAdministrationBlautia2022},
-and this molecule plays an important role in modulating memory
-function
+and this molecule plays an important role in modulating memory function
 \cite{haamCholinergicModulationHippocampal2017}.
@@
 salivaris.} Reduction in the nucleus accumbens has been associated with
-depression symptoms
-\cite{wackerRoleNucleusAccumbens2009},
+depression symptoms \cite{wackerRoleNucleusAccumbens2009},
 and increased levels of the \emph{Alistipes} genus have been
@@
 metagenome and brain structure relationship is novel.
-
 Both \emph{A. equolofaciens} and \emph{A. celatus,} two closely-related
@@
 
+\begin{table}
+    \begin{tabular}{|r|r|r|r|}
+      \hline\hline
+      \textbf{Input set} & \textbf{Age bracket} & \textbf{Microbiome encoding type} & \textbf{Demographics Provided? (sex, education)} \\\hline
+      1 & \multirow{5}{*}{0 to 6 months} & Not provided & yes \\ \cline{3-4}
+      2 &  & \multirow{2}{*}{Taxonomic profile} & no \\ \cline{4-4}
+      3 &  &  & yes \\ \cline{3-4}
+      4 &  & \multirow{2}{*}{Functional Profile (ECs)} & no \\ \cline{4-4}
+      5 &  &  & yes \\ \cline{2-4}
+      6 & \multirow{5}{*}{18 to 120 months} & Not provided & yes \\ \cline{3-4}
+      7 &  & \multirow{2}{*}{Taxonomic profile} & no \\ \cline{4-4}
+      8 &  &  & yes \\ \cline{3-4}
+      9 &  & \multirow{2}{*}{Functional Profile (ECs)} & no \\ \cline{4-4}
+      10 &  &  & yes \\\hline\hline
+    \end{tabular}
+\end{table}
+
 Random Forests (RFs)
@@
 
-\(fitness =\)
-
-\textbf{Equation 1.} Mathematical expression of the fitness measure used
-to weight feature importances based on model benchmarks
+\begin{equation}
+    fitness = \sqrt{max(r_{train}, 0) \cdot max(r_{test}, 0)}
+\end{equation}
 
@@
 Shotgun metagenomic sequences were analyzed using the bioBakery suite of
-computational tools
-(Beghini
-et al. 2021)
-\cite{}.
+computational tools \cite{beghiniIntegratingTaxonomicFunctional2021}.
 First, KneadData (v0.7.7) was used to perform quality
@@
 available on the Open Science Framework
-(Bonham et al.,
-2022)
-\cite{}.
+\cite{bonhamECHORESONANCEMicrobiome2022}.
  Data processing, generation of summary statistics, and
 generation of plots was performed using the julia programming language
-(Bezanson et al.,
-2017; Bonham et al., 2021; Danisch and Krumbiegel, 2021)
-\cite{}.
+\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021}.
  All code for
 data analysis and figure generation, as well as scripts for automated
-download of input files are available on github \{\{CITE zenodo\}\}.
-
-\section*{Research Standards}
+download of input files are available on github \hl{\{\{CITE zenodo\}\}}.
 

### 2023-03-10 manuscript/main.tex

@@
 species such as \emph{Ruminococcus gnavus} were more commonly found in
-children with low cognitive scores regardless of age or maternal
-education. Microbial genes for enzymes involved in the metabolism of
+children with low cognitive scores after controlling for sociodemographic factors.
+Microbial genes for enzymes involved in the metabolism of
 neuroactive compounds, particularly short-chain fatty acids such as
@@
 In addition, ML models were able to use microbial taxa to predict the
-volume of brain regions, with particular taxa often dominating the
-feature importance metric for individual brain regions, such as \emph{B.
-wexlerae} being the most dominant for the parahippocampal region or
-GABA-producing \emph{Bacteroidetes ovatus} for left accumbens. These
-findings provide potential biomarkers of neurocognition and may lead to
-the future development of targets for early detection and early
+volume of brain regions, with particular taxa, many that were identified
+as important in predicting cognitive function, dominating the
+feature importance metric for individual brain regions. For example, \emph{B.
+wexlerae} was dominant for the predicting the size of the parahippocampal region
+in both the left and right hemispheres, while several \textit{Bacteroitetes} species,
+including GABA-producing \emph{B. ovatus}, were important for predicting
+the size of the left accumbens area, but not the right. These
+findings provide potential biomarkers of neurocognition and brain development
+and may lead to the future development of targets for early detection and early
 intervention.
@@
 nerve and through endocrine and immune (cytokine) signaling molecules
-\cite{cerdoEarlyNutritionGut2019,pronovostPerinatalInteractionsMicrobiome2019,sharonCentralNervousSystem2016,togniniGutMicrobiotaPotential2017}
+\cite{cerdoEarlyNutritionGut2019,pronovostPerinatalInteractionsMicrobiome2019,sharonCentralNervousSystem2016,togniniGutMicrobiotaPotential2017}.
 In addition, the products of microbial metabolism generated in the gut can
@@
 is rapidly seeded at birth through contact with the birth canal (if
-birthed vaginally), caregivers, food sources (breastmilk or formula),
-and other environmental sources such as antibiotics
+birthed vaginally), caregivers, food sources (breast milk or formula),
+and other environmental sources
 \cite{backhedDynamicsStabilizationHuman2015,bokulichAntibioticsBirthMode2016}.
@@
 \cite{silbereisCellularMolecularLandscapes2016}.
-Much of this development occurs in discrete windows or
-``sensitive-'' or ``critical periods'' (CPs)
+Much of this development occurs in discrete windows called
+"sensitive-" or "critical periods" (CPs)
 \cite{knudsenSensitivePeriodsDevelopment2004}
@@
 and particular modes of learning
-and skill development are preferred. The timing of these sensitive
-periods is driven in part by genetics, but can also be affected by the
+and skill development are preferred. The timing of these CPs
+is driven in part by genetics, but can also be affected by the
 environment, including the gut microbiome
@@
 To begin to address this need, we investigated the gut microbiome and
-neurocognitive development of children from infancy through 10 years of
+neurocognitive development of healthy, neurotypical children from infancy through 10 years of
 age in an accelerated longitudinal study design. Gut microbial
@@
 \cite{bairochENZYMEDatabase20002000},
-and
-KEGG orthologues - KOs
+and KEGG orthologues - KOs
 \cite{kanehisaKEGGResourceDeciphering2004}
@@
 for the relative abundance of each species that had at least 15\%
-prevalence in a given age group (Figure 2A, N = 92 for 0--120 months, N
-= 46 for 0--6 months, N = 97 for 18--120 months). Only \emph{Blautia wexlerae}
-was significantly associated (q value = 0.14, $\beta$ = 0.0015) with
+prevalence in a given age group (Figure 2A, N\textsubscript{species} = 92 for 0--120 months, 
+N\textsubscript{species} = 46 for 0--6 months, N\textsubscript{species} = 97 for 18--120 months).
+Only \emph{Blautia wexlerae} was significantly associated (q value = 0.14, $\beta$ = 0.0015) with
 cognitive function in children under 6 months old after adjusting for
@@
 \emph{Asaccharobacter celatus} and
-\emph{Adelcreutzia equolifaciens}, which produce phytoestrogen-derived
-equol
+\emph{Adelcreutzia equolifaciens}, which produce phytoestrogen-derived equol
 \cite{maruoAdlercreutziaEquolifaciensGen2008,thawornkunoBiotransformationDaidzeinEquol2009},
-and the SCFA-producing probiotic
+and SCFA-producing probiotic
 species such as \emph{Eubacterium eligens} and \emph{Faecalibacterium
@@
 genes. However, because the relationships between individual taxa are
-still largely unknown, we turned to Random Forest (RF) models, an
+still largely unknown, we turned to Random Forest (RF) models, a family of 
 unsupervised non-parametric machine learning (ML) method that enables
@@
 interaction terms.
+Given that gut microbial profiles, as well as neurocognitive
+development, may partially reflect socioeconomic and demographic
+factors, we assessed the performance of RF regressors where maternal
+education (a proxy of socioeconomic status (SES)), sex, and age were
+included as possible predictors, either alone or in combination with
+microbial taxonomic profiles (Table 3).
 
@@
 \end{sidewaysfigure}
-
-Given that gut microbial profiles, as well as neurocognitive
-development, may partially reflect socioeconomic and demographic
-factors, we assessed the performance of RF regressors where maternal
-education (a proxy of socioeconomic status (SES)), sex, and age were
-included as possible predictors, either alone or in combination with
-microbial taxonomic profiles (Table 3).
 
@@
 
-As with linear models, RF models for children under 6 months old were
-less generalizable (mean test-set correlation -0.13, mean RMSE 12.70),
+As with linear models, RF models for children under 6 months old
+performed poorly (mean test-set correlation -0.13, mean RMSE 12.70),
 but RFs were consistently able to learn the relationship between taxa
@@
 employed a Random Forest modeling approach to associate gut taxonomic
-profiles with individual brain regions identified in MRI scans,
-normalized to total brain volume. Some brain regions were more readily
+profiles with individual brain regions identified in MRI scans.
+Some brain regions were more readily
 predicted by RF models trained on microbial taxa (Table 4, Supplementary
@@
 
-We also observed that many brain regions had high accordance across the
+While models for many regions were highly concordant across the
 left and right hemispheres in terms of both model performance and
-microbial feature importance. In contrast, other regions had substantial
+microbial feature importance, other regions had substantial
 differences between the hemispheres. For example, the left accumbens
@@
 brain region models (R = 0.288), as compared to the right accumbens
-models which could not adequately generalize, and had a negative mean
-test-set correlation (R = -0.041). Most healthy individuals have a
+models which had a negative mean test-set correlation (R = -0.041)
+indicated that generalizable models could not be identified.
+Most healthy individuals have a
 rightward asymmetry in the nucleus accumbens, and reduced asymmetry has
 been linked to substance use disorder in young adults
-\cite{caoMappingCorticalSubcortical2021}.
-
+\cite{caoMappingCorticalSubcortical2021}. The accumbens area is also
+associated with reward control, and in individuals diagnosed with ADHD,
+it has been shown to have a divergent neuromorphology
+\cite{hoogmanSubcorticalBrainVolume2017}.
 Feature importance for models of the left accumbens area were dominated
@@
 relative importance), \emph{B. ovatus} (3.7\% relative importance), and
-\emph{B. uniformis} (3.0\% relative importance).. The accumbens area is
-associated with reward control, and in individuals diagnosed with ADHD,
-it has been shown to have a divergent neuromorphology
-\cite{hoogmanSubcorticalBrainVolume2017}.
-Independently, \emph{B. ovatus}, \emph{B. uniformis} and
-\emph{B. vulgatus} have been linked to ADHD
+\emph{B. uniformis} (3.0\% relative importance). Each of these species
+have been independently linked to ADHD
 \cite{wangGutMicrobiotaDietary2020}.
-In fact, there are studies showing that alterations on the striatal
-dopamine transporters can cause effects resembling hyperactivity and
-attention deficit
-\cite{yaelDisinhibitionNucleusAccumbens2019},
-and that \emph{B. uniformis} is gut-microbial modulator of the brain
-dopamine transporter
-\cite{hartstraInfusionDonorFeces2020}.
-
-Many of the taxa identified as important for models predicting brain structure
-were also identified in LMs and RF models of cognitive function. 
+
 Interestingly, while RF models for
 multiple brain regions had many important microbial taxa, others were
-dominated by a small number of taxa. In general, for all segments, a
+dominated by a small number of taxa, many of which
+were also identified in LMs and RF models of cognitive function. 
+In general, for all segments, a
 consistent number of species between 14 and 22 (median = 20) was
-responsible for one third of the fitness-weighted cumulative importance
-(see Methods), regardless of the model benchmarks. We proceeded to
-select a subset of the most important taxa (averaged over all segments)
-and the segments whose importance was more heavily loaded to further
-analyze the multiple relationships unveiled by the RF models
-(Figure 4 A and B)
+responsible for one third of the fitness-weighted cumulative importance,
+regardless of model performance (\hl{Suplementary Table XX}, Figure 4 A and B).
+We observed two major patterns of importance distribution from
+taxa over the brain segments; some species portrayed high contributions
+to multiple different segments, while others contributed modestly to
+just one or two brain segments. Notable cases of the first pattern
+included seven species - \emph{Anaerostipes hadrus}, \emph{Bacteroides
+vulgatus}, \emph{Fusicatenibacter saccharivorans}, \emph{Ruminococcus
+torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
+\emph{Blautia wexlerae} - that combined, account for approximately one
+third of the cumulative relative importances, computed after subsetting
+on the taxa of interest (Figure 4C).
 
@@
 
-Our analysis revealed two major patterns of importance distribution from
-taxa over the brain segments; some species portrayed high contributions
-to multiple different segments, while others contributed modestly to
-just one or two brain segments. Notable cases of the first pattern
-included seven species - \emph{Anaerostipes hadrus}, \emph{Bacteroides
-vulgatus}, \emph{Fusicatenibacter saccharivorans}, \emph{Ruminococcus
-torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
-\emph{Blautia wexlerae} - that combined, account for approximately one
-third of the cumulative relative importances, computed after subsetting
-on the taxa of interest (Figure 4C).
-
 Among these, the most important variable is \emph{Anaerostipes hadrus},
@@
 of entire major lobes (frontal, parietal), which cannot be readily
-correlated to specific cognitive functions. After accounting for the
-major lobes, it is found to be highly important on the prediction of the
-cerebellar vermal lobules VIII-X, pars opercularis, cuneus and
-precuneus, and anterior-cingulate (Figure 4B).
-
-To better understand this relationship, we performed hierarchical
-cluster analysis, which revealed a well defined cluster of species with
-high importance loadings on close segments of the lower temporal and
-close occipital lobe, to which \emph{A. hadrus} belongs. The other
-species on this cluster are \emph{B. wexlerae}, \emph{R. torques},
-\emph{R. intestinalis, R. bicirculans} and \emph{F. saccahrivorans}, who
-contribute heavily to the entorhinal, fusiform, lingual and
-parahippocampal segments. This group of taxa and their related segments drew our attention because
-they contained \emph{B. wexlerae}, a highlight from the cognitive
-assessment results. Previous works exploring Parkinson Disease patients
+correlated to specific cognitive functions, (Figure 4B).
+\emph{B. wexlerae}, which was important in RF models of cognitive function,
+and was the only species to be significantly associated
+with cogntive function in young children, 
+was also important in models of the entorhinal, fusiform, lingual and
+parahippocampal regions. Previous works exploring Parkinson Disease patients
 found out that issues in the cognitive task of confrontation naming were
@@
 \cite{haamCholinergicModulationHippocampal2017}.
-
-Another notable importance cluster contains taxa associated with the
+Another cluster of model importance contains taxa associated with the
 basal forebrain, especially the cingulate and the accumbens. This
 cluster contains the previously-discussed \emph{B. ovatus} and \emph{B.
-uniformis,} but also \emph{Alistipes finegoldii and Streptococcus
-salivaris.} Reduction in the nucleus accumbens has been associated with
+uniformis}, but also \emph{Alistipes finegoldii and Streptococcus
+salivaris}. Reduction in the nucleus accumbens has been associated with
 depression symptoms \cite{wackerRoleNucleusAccumbens2009},
@@
 \cite{jiangAlteredFecalMicrobiota2015}.
-While these reports are important, simultaneously probing of the gut
-metagenome and brain structure relationship is novel.
+
 Both \emph{A. equolofaciens} and \emph{A. celatus,} two closely-related
 equol-producing species, are examples of the second contribution
-pattern, and had high importance in predicting the relative volume of
+pattern, where a model is dominated by a small number of taxa.
+Each of these species were highly important in predicting the relative volume of
 the right anterior cingulate (respectively, 3.0\% and 2.6\% relative
@@
 \cite{setchellSEquolPotentLigand2005},
-and in anterior cingulate, estrogen, has been shown to regulate
+and in the anterior cingulate, estrogen has been shown to regulate
 pain-related aversion
 \cite{xiaoEstrogenAnteriorCingulate2013}.
-
-Another example of this pattern is \emph{R. gnavus,} which was the only
-species with significant negative association with cognitive function.
+Another example of this pattern is \emph{R. gnavus}, which was the only
+species with significant negative association with cognitive function in linear models.
 It is heavily associated with the left pars opercularis (2.7\% relative
-importance), a relationship that may be explained by the emerging
-understanding that this species is increased in individuals with insulin
+importance). \textit{R. gnavus} is increased in individuals with insulin
 resistance and obesity \cite{leyHumanGutMicrobes2006},
@@
 clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (4.0\% relative importance)
-(only age had higher relative importance in any model), while also being
-one of the most important predictors for neighboring areas of the
+is the highest for a microbe in RF models (4.0\% relative importance) -
+only age had higher relative importance in any model. At the same time,
+\textit{C. comes} is one of the most important predictors for neighboring areas of the
 posterior cingulate such as the pars opercularis (relative importances,
@@
 neurocognition (eg anxiety and depression, neurodegeneration, attention
-deficit / hyperactivity disorder, and autism) and mechanistic studies in
-animal models. The results from these studies point to the possibility
+deficit / hyperactivity disorder, and autism \cite{romanGutBrainAxis2018,fosterGutBrainAxis2013})
+and mechanistic studies in animal models
+\cite{hsiaoMicrobiotaModulateBehavioral2013,needhamGutderivedMetaboliteAlters2022}.
+The results from these studies point to the possibility
 that gut microbes and their metabolism may be causally implicated in
@@
 equolofaciens} produces the nonsteroidal estrogen equol from isoflaven
-found in soybeans
-\cite{wangEnantioselectiveSynthesisSEquol2005},
+found in soybeans \cite{wangEnantioselectiveSynthesisSEquol2005},
 and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
@@
 its ability to metabolize drug compounds such as the plant-derived heart
-medication digoxin
-\cite{haiserPredictingManipulatingCardiac2013}.
+medication digoxin \cite{haiserPredictingManipulatingCardiac2013}.
 The metabolic versatility of this clade, and the large number of
@@
 of lockdowns. In our subject set for this study, these effects are more
-pronounced in some age groups due to our sampling schedule
+pronounced in some age groups due to the period when active recruitment for
+the study was occurring.
 \cite{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
 (Supplementary Figure 3).
+
+% and alterations on the striatal
+% dopamine transporters can cause effects resembling hyperactivity and
+% attention deficit
+% \cite{yaelDisinhibitionNucleusAccumbens2019},
+% and that \emph{B. uniformis} is gut-microbial modulator of the brain
+% dopamine transporter
+% \cite{hartstraInfusionDonorFeces2020}.
 
@@
 \cite{bruchhageLongitudinalBrainCognitive}.
+Brain region volumes used in RF models were normalized to total brain volume
+(the sum of white and gray matter).
 

### 2023-03-12 manuscript/main.tex
through figures 1-3 sean comments

@@
 \begin{abstract}
-The gastrointestinal tract, its resident microorganisms and the central
+The gastrointestinal tract, its resident microorganisms, and the central
 nervous system are connected by biochemical signaling, also known as
@@
 \cite{laueDevelopingMicrobiomeBirth2022}.
-Fetal development \emph{in utero} is believed to be mostly sterile, but
+Development \emph{in utero} is believed to be mostly sterile, but
 is rapidly seeded at birth through contact with the birth canal (if
@@
 \cite{backhedDynamicsStabilizationHuman2015}.
-Many prior studies have focused on either infant microbiomes or
+Prior studies have typically focused on either infant microbiomes or
 adult microbiomes, since performing statistical analyses across this
@@
 transition coincides with critical neural developmental windows and
-neural synaptogenesis, investigation across this solid-food boundary is
-incredibly important
+associated neurodevelopmental processes including myelination, neurogenesis, and synaptic pruning,
+investigation across this solid-food boundary is important
 \cite{tauNormalDevelopmentBrain2010}.
@@
 Much of this development occurs in discrete windows called
-"sensitive-" or "critical periods" (CPs)
+sensitive periods (SPs)
 \cite{knudsenSensitivePeriodsDevelopment2004}
-when neural plasticity is particularly high,
-and particular modes of learning
-and skill development are preferred. The timing of these CPs
-is driven in part by genetics, but can also be affected by the
-environment, including the gut microbiome
+during which neural plasticity is particularly high. The timing of these SPs
+is driven by both genetic and environmental cues, including the gut microbiome
 \cite{cowanAnnualResearchReview2020}.
-In fact, emerging evidence suggests that the timing and duration of CPs
+Emerging evidence suggests that the timing and duration of SPs
 may be driven in part by cues from the developing gut microbiome
@@
 
-To begin to address this need, we investigated the gut microbiome and
-neurocognitive development of healthy, neurotypical children from infancy through 10 years of
-age in an accelerated longitudinal study design. Gut microbial
+To begin to address this need, we investigated trends and associations between
+the gut microbiome and
+neurocognitive development in a large cohort of healthy and neurotypically-developing
+children from infancy through 10 years of age. Gut microbial
 communities were assessed using shotgun metagenomic sequencing, enabling
-profiling at both the taxonomic and gene-functional level, and
-neurocognitive development was measured using expert-assessment of
-cognitive function and neuroimaging using magnetic resonance imaging
-(MRI). Using a combination of classical statistical analysis and machine
-learning, we showed that the development of the gut microbiome, the
-brain, and children's cognitive abilities are intimately linked, with
+profiling at both the taxonomic and gene-functional level.
+Cognitive skills and abilities were measured using age-appropriate
+psychometric assessments of cognitive function s, i.e.,
+the Mullen Scales of Early Learning (MSEL) \cite{mullenMullenScalesEarly1995},
+Weschler Preschool and Primary Scale of Intelligence, 4th Edition (WPPSI-4) 
+\cite{wechslerWechslerPreschoolPrimary2012}
+and Weschler Intelligence Scale for Children, 5th Edition (WISC-V)
+\cite{wechslerWechslerIntelligenceScale1949}.
+Finally, we also assessed emerging brain structure magnetic resonance imaging
+(MRI). Through a combination of classical statistical analysis and machine
+learning, we find that the development of the gut microbiome,
+children's cognitive abilities, and brain structure are intimately linked, with
 both microbial taxa and gene functions able to predict cognitive
@@
 We investigated the co-development of the brain and the microbiome in
-early childhood in a cohort of typically-developing children in their
-first years of life using a variety of orthogonal microbial and
-neurocognitive assessments, including shotgun metagenomic sequencing,
-cognitive and behavioral assessments, and neuroimaging. In all, 361
-children from RESONANCE,
-an accelerated longitudinal cohort of early development
-\cite{forrestAdvancingScienceChildren2018}
-between the ages of 82 days and 10 years
-old (median age 2.21 years, Supplementary Figure 1) were included in
-this study (Table 1, Figure 1A). To measure cognitive function, we used
-age-appropriate assessments that can be normalized to a common, IQ-like
-scale (Figure 1B), including the Mullen scale of early learning (MSEL)
-for children from birth to 3 years of age
-\cite{mullenMullenScalesEarly1995},
-Wechsler preschool and primary scale of intelligence (WPPSI)
-for 4-5 year-olds
-\cite{wechslerWechslerPreschoolPrimary2012},
-and the Wechsler intelligence scale for children (WISC)
-for children 6 years and up
-\cite{wechslerWechslerIntelligenceScale1949}.
-
-As expected, the greatest differences in microbial taxa were driven by
+361 healthy and neurotypically-developing children (163 female) between 82 days and 10 years of age
+(Table 1,  Figure 1A, Supplementary Figure 1)\cite{forrestAdvancingScienceChildren2018}
+using a variety of orthogonal microbial and neurocognitive measures.
+These included shotgun metagenomic sequencing,
+age-appropriate cognitive and behavioral assessments
+using full-scale composite scores from the MSEL, WPPSI-4 and WISC-V,
+and neuroimaging measures of cortical and sub-cortical morphology. 
+To measure cognitive function, we used
+age-appropriate assessments that can be 
+Different assessment instruments were used depending on the child's age
+due to the unique age-range and psychometric properties measured by each
+(i.e., the MSEL for children under 4 years of age, The WPPSI-3 for 3-6 year-olds,
+and the WISC-V for children older than 6), but were normalized to a common scale (Figure 1B),
+
+As expected, the greatest differences in microbial taxa were observed across
 age (Figure 1B-C, PCoA axis 1), while older children were primarily

### 2023-03-12 manuscript/main.tex
through figure 4

@@
 As expected, the greatest differences in microbial taxa were observed across
-age (Figure 1B-C, PCoA axis 1), while older children were primarily
+child age (Figure 1B-C, PCoA axis 1), while older children were primarily
 stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
@@
 
-Several studies have demonstrated links between specific taxa and
-measures of anxiety and depression
-\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022},
-cognitive flexibility in adults
-\cite{magnussonRelationshipsDietrelatedChanges2015},
-and atypical neural development
-\cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021}.
-We, therefore, set out to identify whether specific taxa
-or gene functions were linked with normal cognitive development in
-children. In order to assess whether variations in gut microbial taxa,
-their genes, or their metabolism are linked with neurocognitive
-development, we tested whether the beta diversity of microbial taxa and
-gene functions, as well as variation in brain development as assessed by
-neuroimaging, were associated with these contemporaneous measures of
-cognitive function using permutation analysis of variance (PERMANOVA).
-Due to the large ecological shift in the microbiome that occurs upon the
-introduction of solid food, and a relatively wide range of ages when
-infants are transitioned to solid food, we considered the pre-transition
-(less than 6 months old) and post-transition (over 18 months old)
-microbiomes separately. Even still, age was a major driver of variation
-in gut microbiomes in children over 18 months old (Figure 1G). We also
-found that overall variation in microbial species in children over 18
-months old was associated with a small but significant variation in
-cognitive function score (Figure 1G, R\textsuperscript{2} = 0.0124, q
-\textless{} 0.001), as was variation in microbial gene functions
-(R\textsuperscript{2} = 0.0119, q \textless{} 0.001). Variation in
-microbial taxa and genes was not significantly associated with cognitive
+As several prior studies have demonstrated links between specific gut taxa
+with ASD \cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021},
+adult cognitive skill (e.g., executive functioning)\cite{magnussonRelationshipsDietrelatedChanges2015},
+and adult mental health (both anxiety and depressive symptioms)\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022}
+we sought to determine if specific taxa or gene functions were linked with normal cognitive development in children.
+To test if variations in gut microbial taxa,
+their genes, or their metabolism are associated with neurocognitive
+development, we used permutation analysis of variance (PERMANOVA) 
+that included the @b diversity of microbial taxa and gene functions,
+neuroimaging-derived volume measures of cortical and subcortical structures,
+and general cognitive ability.
+Given the large ecological shift in the microbiome that occurs upon the introduction of solid food,
+and the relatively wide range of ages when over which infant transition from milk to solid foods,
+we considered the pre-transition (prior to 6 months of age old)
+and post-transition (over 18 months of age) microbiomes separately.
+However, even with this categorization, we find increasing variation
+in gut microbiomes in children over 18 months old (Figure 1G).
+We also found that overall variation in microbial species in children over 18 months old
+was associated with a small but significant variation in cognitive function score
+(Figure 1G, R\textsuperscript{2} = 0.0124, q \textless{} 0.001),
+as was variation in microbial gene functions (R\textsuperscript{2} = 0.0119, q \textless{} 0.001).
+Variation in microbial taxa and genes was not significantly associated with cognitive
 function in children under 6 months, though this may be due to the low
@@
 \cite{valles-colomerNeuroactivePotentialHuman2019}
-and concurrent cognitive function score (Table 2, Figure 2C-G) and found that
+and concurrent cognitive ability score (Table 2, Figure 2C-G) and found that
 several metabolic pathways were significantly enriched or depleted in
-children with higher cognitive function scores. This was true both when
-considering all age groups together, though the enrichment of most
+children with higher cognitive ability scores. This was true both when
+considering all ages together, though the enrichment of most
 pathways was more pronounced either in children under 6 months or those over 18
 months. For example, genes for degrading the 3-carbon SCFA propionate
-were significantly depleted in children with higher cognitive function
+were significantly depleted in children with higher cognitive ability
 scores across all age groups tested (Table 2; propionate degradation I,
@@
 
-Synthesis of menaquinone (vitamin K) was also negatively associated with
+We found that synthesis of menaquinone (vitamin K) was negatively associated with
 cognitive function score in older children (menaquinone synthesis I,
@@
 microbial feature importance, other regions had substantial
-differences between the hemispheres. For example, the left accumbens
-area, which plays an important role in reward circuits
-\cite{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012},
+bilateral differences. For example, the left accumbens area
 has one of the highest test-set correlations of our
-brain region models (R = 0.288), as compared to the right accumbens
-models which had a negative mean test-set correlation (R = -0.041)
-indicated that generalizable models could not be identified.
+brain region models (R = 0.288), as compared to the right which
+had a negative mean test-set correlation (R = -0.041)
+indicated that generalizable models were not identified.
 Most healthy individuals have a
-rightward asymmetry in the nucleus accumbens, and reduced asymmetry has
-been linked to substance use disorder in young adults
+rightward asymmetry in the nucleus accumbens,
+which is involved in reward learning and risk taking behaviours
+\cite{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012}.
+Reduced asymmetry between right and left nucleus accumbens
+has been associated with substance use disorder in young adults
 \cite{caoMappingCorticalSubcortical2021}. The accumbens area is also

### 2023-03-13 manuscript/main.tex
sean methods

@@
 
-Data used in this study were drawn from the ongoing longitudinal
-RESONANCE study of healthy and neurotypical brain and cognitive
-development, based at Brown University in Providence, RI, USA. The
-RESONANCE study is part of the NIH initiative Environmental influences
-on Child Health Outcomes (ECHO)
+Data used in this study were drawn from an ongoing longitudinal
+brain and cognitive development, termed RESONANCE, based in Providence, RI, USA,
+and spanning fetal through adolescent development. The
+RESONANCE cohort is part of the NIH Environmental influences
+on Child Health Outcomes (ECHO) study
 \cite{forrestAdvancingScienceChildren2018,gillmanEnvironmentalInfluencesChild2018},
- a longitudinal observational study of
-healthy and neurotypical brain development that spans the fetal and
-infant to adolescent life stages, combining neuroimaging (magnetic
+and includes neuroimaging (magnetic
 resonance imaging, MRI), neurocognitive assessments, bio-specimen
-analyses, subject genetics, environmental exposures such as lead, and
+analyses, subject genetics, broad environmental exposure data
+(such as chemical, water quality, nutrition, etc.), and
 rich demographic, socioeconomic, family and medical history information.
-From the RESONANCE cohort, 361 typically-developing children between the
-ages of 2.8 months and 10 years old (median age 2 years, 2 months) were
-selected for analysis in this study.
-
-General participant demographics are provided in Table 1. Children are
-representative of the RI population. Children in the RESONANCE cohort
-were born full-term (\textless{} 37 weeks gestation) with height and
-weight normal for gestational age, and from uncomplicated singleton
-pregnancies. Children with known major risk factors for developmental
-abnormalities at enrollment were excluded. In addition to screening at
+
+As a study of neurotypical development, mothers and their children
+were devoid of major risk factors for abnormal or atypical development.
+These included preterm birth (< 37 weeks gestation) low birth weight (<1500g),
+in utero exposure to alcohol, cigarette or illicit substance exposure;
+abnormalities on fetal ultrasound; non-singleton or complicated pregnancy,
+including preeclampsia, high blood pressure, or gestational diabetes;
+complicated vaginal or cesarian birth; 5 minute APGAR scores < 8;
+NICU admission; neurological disorder in the child
+(e.g., head injury resulting in loss of consciousness, epilepsy);
+and psychiatric or learning disorder (including maternal depression)
+in the infant, parents, or siblings requiring medication in the year prior to pregnancy.
+In addition to screening at
 the time of enrollment, on-going screening for worrisome behaviors using
@@
 them from subsequent analysis.
-
-Exclusion criteria included: \emph{in utero} exposure to alcohol,
-cigarette or illicit substance exposure; preterm (\textless{} 37 weeks
-gestation) birth; small for gestational age or less than 1500 g; fetal
-ultrasound abnormalities; preeclampsia, high blood pressure, or
-gestational diabetes; 5 minute APGAR scores \textless{} 8; NICU
-admission; neurological disorder (e.g., head injury resulting in loss of
-consciousness, epilepsy); and psychiatric or learning disorder
-(including maternal depression) in the infant, parents, or siblings
-requiring medication in the year prior to pregnancy.
-
-Demographic and other non-biospecimen data such as race and ethnicity,
-parental education and occupation, feeding behavior (breast- and
-formula-feeding), child weight and height, were collected through
-questionnaires or direct examination as appropriate. All data were
-collected at every assessment visit, if possible.
+For this study, 361 typically-developing children between the
+ages of 2.8 months and 10 years old (median age 2 years, 2 months) were
+selected for analysis based on having provided at least one stool sample
+in the same week as being evaluated using one of the cogntitive ability
+assessments (MSEL, WPPSI, WISC-V).
 
@@
 
-Overall cognitive function was assessed using age-appropriate methods.
-For children from birth to 30 months, we used an Early Learning
-Composite as assessed via the Mullen Scales of Early Learning (MSEL)
-\cite{mullenMullenScalesEarly1995},
- a standardized and population-normed tool for assessing fine and
-gross motor, expressive and receptive language, and visual reception
-functioning in children from birth through 68 months of age. The
-Wechsler Intelligence Quotient for Children (WISC)
-\cite{wechslerWechslerPreschoolPrimary2012}
-is an individually administered standard intelligence test for children
-aged 6 to 16 years. It derives a full scale intelligence quotient (IQ)
-score, which we used to assess overall cognitive functioning. The fourth
-edition of the Wechsler Preschool and Primary Scale of Intelligence
-(WPPSI-IV) is an individually administered standard intelligence test
-for children aged 2 years 6 months to 7 years 7 months, trying to meet
-the increasing need for the assessment of preschoolers. Just as the
-WISC, it derives a full scale IQ score, which we used to assess overall
-cognitive functioning.
+General cognitive ability was assessed using age-appropriate performance-based measures.
+For children under 3 years of age, we used the Early Learning
+Composite (ELC) from the Mullen Scales of Early Learning (MSEL)
+\cite{mullenMullenScalesEarly1995}.
+The MSEL isa standardized and population-normed tool that assesses 
+function in 5 major domains: fine and gross motor, expressive and receptive language,  
+and visual functioning.
+The ELC is an age-normalized composite derived from these individual domains
+with a mean of 100 and standard deviation of 15. For older children, full-scale IQ (FSIQ)
+was assessed using the Wechsler Preschool and Primary Scale of Intelligence 4th Edition
+(WPPSI-IV, children 3 to 6)\cite{wechslerWechslerPreschoolPrimary2012},
+and the Wechsler Intelligence Scale for Children 5th Edition (WISC-V)
+for children older than 6\cite{wechslerWechslerIntelligenceScale1949}. 
+Like the ELC, FSIQ is normalized to a mean of 100 and standard deviation of 15.
 
@@
 
-MRI data was acquired at a 3T Siemens Trio scanner with the following
-parameters: TE=5.6msec, TR=1400msec, FA=15 degrees, 1.1x1.1x1.1mm
-resolution, 160x160 matrix, with an average of 112 slices. FOV was
-adjusted to infant size. Using a combination of linear and nonlinear
-image registration, we created representative age-specific templates
-using the ANTs package \cite{avantsInsightToolKitImage2014}.
-After age specific templates were created, a single flow from
-each age to the 12-month template was estimated and a final warp from
-the 12-month template to standard adult MNI space was performed. For
-each individual infant/ child brain image, we then calculated the warp
-from their native T1w image space to their nearest-in-age template.
-Using the resulting warps (native → nearest age template → 12-month
-template → MNI template), we could move the standard adult brain atlas
-to the space of an individual infant in a single step. In this case, we
-used the Harvard-Oxford brain atlas to provide a coarse-grained
-parcellation of individual brains into subcortical regions (e.g.,
-thalamus, putamen) and total grey and white matter volumes (included as
-part of the FSL package \cite{jenkinsonFSL2012}).
-Total tissue and brainstem volume as well as left and right
-hemisphere volumes were derived for total white and cortical gray
-matter, lateral ventricle, thalamus, caudate, putamen, pallidum,
-hippocampus, amygdala, and accumbens as well as total brainstem volume
-\cite{bruchhageLongitudinalBrainCognitive}.
+T1-weighted anatomical MRI data were acquired using a 3-Dimensional MP-RAGE sequence
+on 3T Siemens Trio scanner with the following parameters:
+TE=5.6msec, TR=1400msec, TI = 950msec, FA=15 degrees, and 1.1x1.1x1.1mm resolution.
+Image matrix and field-of-view were adjusted by infant size to achieve the desired image resolution.
+For children under ~4 years of age, MRI was performed during natural non-sedated sleep using previously described methods \hl{(Doug's paper ref)}.
+Older children were scanned awake whilst watching a movie or favorite TV show.
+To help minimize motion, pediatric (MedVac) immobilizers were used along with foam cushions and padding.
+Image quality was monitored during and following scanning.
+If motion artifacts were observed (e.g., ghosting, blurring, or other signal variations), scans were repeated or removed from analysis.
+
+Following data acquisition, brain region volumes corresponding to total white and gray matter,
+cortical gray matter, brainstem, and left and right hemisphere lateral ventricles,
+thalamus, caudate, putamen, pallidum, hippocampus, amygdala,
+and nucleus accumbens were calculated using an atlas matching approach\cite{bruchhageLongitudinalBrainCognitive}.
+Individual images were first nonlinearly aligned to MNI-space using
+a multi-step registration approach (ANTs 2.2 toolbox, \cite{avantsInsightToolKitImage2014}).
+The MNI-aligned Harvard-Oxford brain atlas was then aligned and superimposed onto the individual child
+by applying the inverse transformation, and the brain regions of interest delinated (segmented)
+and their volune calculated by summing the number of image voxels within each\cite{jenkinsonFSL2012}.
+
 Brain region volumes used in RF models were normalized to total brain volume

### 2023-03-13 manuscript/main.tex
begin results stripping

@@
 and neuroimaging measures of cortical and sub-cortical morphology. 
-To measure cognitive function, we used
-age-appropriate assessments that can be 
 Different assessment instruments were used depending on the child's age
@@
 and the WISC-V for children older than 6), but were normalized to a common scale (Figure 1B),
-
 As expected, the greatest differences in microbial taxa were observed across
-child age (Figure 1B-C, PCoA axis 1), while older children were primarily
+child age (Figure 1C-D, PCoA axis 1), while older children were primarily
 stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
@@
 in gut microbial genes and also in brain volume profiles was similarly
-driven largely by subject age (Figure 1C-F).
+driven largely by subject age (Figure 1E-F).
 
@@
 development, we used permutation analysis of variance (PERMANOVA) 
-that included the @b diversity of microbial taxa and gene functions,
+that included the $\beta$ diversity of microbial taxa and gene functions,
 neuroimaging-derived volume measures of cortical and subcortical structures,
@@
 and the relatively wide range of ages when over which infant transition from milk to solid foods,
-we considered the pre-transition (prior to 6 months of age old)
-and post-transition (over 18 months of age) microbiomes separately.
+we considered ages that are generally pre-transition (prior to 6 months of age old)
+and those that are generally post-transition (over 18 months of age) separately.
 However, even with this categorization, we find increasing variation
@@
 profiles, both generated from DNA sequencing data, being tightly coupled
-(Figure 1H, p \textless{} 0.001). Interestingly, other functional
+(Figure 1H, p \textless{} 0.001). Other functional
 groupings (Enzyme commission level4 - ECs
@@
 and KEGG orthologues - KOs
-\cite{kanehisaKEGGResourceDeciphering2004}
-) overlapped only slightly with taxonomic profiles in both age
+\cite{kanehisaKEGGResourceDeciphering2004})
+overlapped only slightly with taxonomic profiles in both age
 cohorts, despite being derived from UniRef90 labels. In children over 18

### 2023-03-13 manuscript/main.tex
remove discussion from Figure 2A-B

@@
 cognitive function in children under 6 months old after adjusting for
-age and maternal education (Figure 2B). \emph{B. wexlerae} was
-previously shown to be depleted in children with diabetes
-\cite{benitez-paezDepletionBlautiaSpecies2020},
-and that oral administration of \emph{B. wexlerae} partially
-ameliorated weight gain and inflammation from a high-fat diet in a mouse
-model of T2D
-\cite{hosomiOralAdministrationBlautia2022,liuBlautiaNewFunctional2021}.
-In children over 18 months of age,
-several microbial species were significantly enriched (q-value
+age and maternal education (Figure 2B).
+By contrast, in children over 18 months of age,
+several mirobial species were significantly enriched (q-value
 \textless{} 0.20) in children with higher cognitive function scores,
-including \emph{Gordonibacter pamelaeae}, which produces the
-neuroprotective metabolite urolithin
-\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014},
-\emph{Asaccharobacter celatus} and
-\emph{Adelcreutzia equolifaciens}, which produce phytoestrogen-derived equol
-\cite{maruoAdlercreutziaEquolifaciensGen2008,thawornkunoBiotransformationDaidzeinEquol2009},
+including \emph{Gordonibacter pamelaeae}
+\emph{Asaccharobacter celatus} and \emph{Adelcreutzia equolifaciens}
+(also known as \textit{Adlercreutzia equolifaciens} subsp. \textit{celatus}
+and \textit{Adlercreutzia equolifaciens} subsp. \textit{equolifaciens} \cite{maruoAdlercreutziaEquolifaciensGen2008}),
 and SCFA-producing probiotic

### 2023-03-13 manuscript/main.tex
remove discussion from Figure 2C-G

@@
 \textless{} 0.20) in children with higher cognitive function scores,
-including \emph{Gordonibacter pamelaeae}
+including \emph{Gordonibacter pamelaeae},
 \emph{Asaccharobacter celatus} and \emph{Adelcreutzia equolifaciens}
@@
 II, E.S. = -0.342, q = 0.020; acetate synthesis III, E.S. = -0.31, q = 0.052).
-SCFAs are produced by anaerobic fermentation of dietary fiber
-and have been linked with immune system regulation as well as directly
-with brain function \cite{dalileRoleShortchainFatty2019}.
-
-We found that synthesis of menaquinone (vitamin K) was negatively associated with
-cognitive function score in older children (menaquinone synthesis I,
-E.S. = -0.170, q = 0.0183). Menaquinone has several isoforms, one of
-which, MK-4, has been found to be decreased in children diagnosed with ASD
-\cite{dongCorrelationSerumConcentrations2021}
-and is potentially neuroprotective in both rodents and humans
-\cite{elkattawyVitaminK2Menaquinone72022}.
-In children over 18 months old,
+In children over 18 months old, synthesis of menaquinone (vitamin K)
+was also negatively associated with cognitive function score (menaquinone synthesis I,
+E.S. = -0.170, q = 0.0183), while
 genes for the synthesis of the amino acids
 glutamate and tryptophan were significantly enriched in children with
-higher cognitive function scores (Glutamate synthesis I, E.S. = 0.242, q
-= 0.047; Tryptophan synthesis, E.S. = 0.119, q = 0.041). Glutamate is a
-critical neurotransmitter controlling neuronal excitatory/inhibitory
-signaling along with gamma-aminobutyric acid (GABA), and their balance
-in the brain controls neural plasticity and learning, particularly in
-the developing brain
-\cite{cohenkadoshLinkingGABAGlutamate2015,palomo-buitragoGlutamateInteractionsObesity2019}.
-Tryptophan metabolism, including
-microbial metabolism of tryptophan, has previously been linked with
-autism in children
-\cite{hoshinoBloodSerotoninFree1984,xiaoFecalMicrobiomeTransplantation2021}.
+higher cognitive function scores (Glutamate synthesis I, E.S. = 0.242, 
+q = 0.047; Tryptophan synthesis, E.S. = 0.119, q = 0.041).
 Taken together, these results suggest that

### 2023-03-13 manuscript/main.tex
remove discussion from Figure 3

@@
 To assess whether individual microbial species were associated with
-cognitive function, we fit multivariable linear regression models
+cognitive function, we fit multivariable linear regression models (LMs)
 \cite{mallickMultivariableAssociationDiscovery2021}
@@
 to predict cognitive conditions related to Alzheimer's disease in
-different scenarios
-\cite{ardekaniPredictionIncipientAlzheimer2017,velazquezRandomForestModel2021}.
+different scenarios \cite{ardekaniPredictionIncipientAlzheimer2017,velazquezRandomForestModel2021}.
 Additionally, given the sequential
@@
 microbiome-wide study, without the necessity to explicitly compute
-interaction terms.
-Given that gut microbial profiles, as well as neurocognitive
+interaction terms. Given that gut microbial profiles, as well as neurocognitive
 development, may partially reflect socioeconomic and demographic
@@
 
-As with linear models, RF models for children under 6 months old
+As with LMs, RF models for children under 6 months old
 performed poorly (mean test-set correlation -0.13, mean RMSE 12.70),
@@
 species that were important in RF overlapped with those that were
-significant when testing the relationship with linear models (Figure
-3A-B, Supplementary Table \hl{XX}A-B), though there were substantial
-differences. For children under 6 months, only \emph{B. wexlerae} was
-significantly associated with cognitive function in linear models, but
-was ranked 20th in importance for RF models. The importance metric
-employed (MDI) is a measure of how well the variable differentiates
-(splits) sets of samples by creating subgroups that reduce the
-intra-group deviations, while also accounting for the amount of samples
-affected by a split. All taxa significantly associated with cognitive
-function score in children over 18 months using LMs belong to the
-top-ranking group responsible for 80\% of the total relative importance,
-except for \emph{Eubacterium ramulus} (Table 3, Figure 3A).
-
-Interestingly, several taxa highly ranked in importance in both age
-groups, including several that were significant in LMs for children over
+significant when testing the relationship with LMs (Figure
+3A-B, Supplementary Table \hl{XX}A-B); aside from except for \emph{Eubacterium ramulus}
+all taxa that were significant using LMs belonged to a group
+responsible for 80\% of the total relative importance in RF models
+(Table 3, Figure 3A).
+However, the rank-ordering of LM significance and RF importance
+were had substantive differences. For example, in children under 6 months,
+only \emph{B. wexlerae} was significantly associated with cognitive function in LMs, but
+was ranked 20th in importance for RF models.
+
+Many taxa highly ranked in importance in RF models in both age
+groups, including several that were identified in LMs for children over
 18 months old, including \emph{R. gnavus} (0--6 months, rank = 13;
@@
 16; 18--120 months, rank = 20), while others such as \emph{Allistipes
-finegoldii} were age-group specific. Several taxa important in RF models
-were not statistically significant when using linear models after
-multiple hypothesis correction. However, these taxa had small nominal
+finegoldii} were age-group specific. A number of taxa that were important in RF models
+were not statistically significant when using LMs after
+multiple hypothesis correction. However, these taxa often had small nominal
 p-values. For example, \emph{Erysipelatoclostridium ramosum} was the

### 2023-03-13 manuscript/main.tex
remove discussion from Figure 4

@@
 indicated that generalizable models were not identified.
-Most healthy individuals have a
-rightward asymmetry in the nucleus accumbens,
-which is involved in reward learning and risk taking behaviours
-\cite{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012}.
-Reduced asymmetry between right and left nucleus accumbens
-has been associated with substance use disorder in young adults
-\cite{caoMappingCorticalSubcortical2021}. The accumbens area is also
-associated with reward control, and in individuals diagnosed with ADHD,
-it has been shown to have a divergent neuromorphology
-\cite{hoogmanSubcorticalBrainVolume2017}.
 Feature importance for models of the left accumbens area were dominated
@@
 relative importance), \emph{B. ovatus} (3.7\% relative importance), and
-\emph{B. uniformis} (3.0\% relative importance). Each of these species
-have been independently linked to ADHD
-\cite{wangGutMicrobiotaDietary2020}.
-
-Interestingly, while RF models for
-multiple brain regions had many important microbial taxa, others were
+\emph{B. uniformis} (3.0\% relative importance).
+
+While RF models for multiple brain regions had many important microbial taxa, others were
 dominated by a small number of taxa, many of which
@@
 We observed two major patterns of importance distribution from
-taxa over the brain segments; some species portrayed high contributions
+species over the brain segments; some species portrayed high contributions
 to multiple different segments, while others contributed modestly to
@@
 
-Among these, the most important variable is \emph{Anaerostipes hadrus},
-which is a butyrate-producing anaerobe that has been positively
-associated with cognitive function
-\cite{kantGenomeSequenceButyrateProducing2015,liCorrelationsGutMicrobiota2022}.
-Its importance is, however, heavily loaded on the volume
-of entire major lobes (frontal, parietal), which cannot be readily
-correlated to specific cognitive functions, (Figure 4B).
+Among these, the most important variable is \emph{Anaerostipes hadrus}
+\hl{!!importance stats!!} (Figure 4B).
 \emph{B. wexlerae}, which was important in RF models of cognitive function,
 and was the only species to be significantly associated
-with cogntive function in young children, 
-was also important in models of the entorhinal, fusiform, lingual and
-parahippocampal regions. Previous works exploring Parkinson Disease patients
-found out that issues in the cognitive task of confrontation naming were
-positively correlated with thinning in the fusiform gyrus and
-parahippocampal gyrus
-\cite{pagonabarragaPatternRegionalCortical2013}.
-Additionally, the left and right parahippocampus, where \emph{B.
-wexlerae} had the highest importance in both models, are important in
-visual/spatial processing and memory
-\cite{aminoffRoleParahippocampalCortex2013}.
-\emph{B. wexlerae} can also produce acetylcholine in the gut
-\cite{hosomiOralAdministrationBlautia2022},
-and this molecule plays an important role in modulating memory function
-\cite{haamCholinergicModulationHippocampal2017}.
-Another cluster of model importance contains taxa associated with the
+with cogntive function in young children using LMs, 
+was also important in models of the entorhinal, fusiform and lingual areas
+and had the highest importance in models predicting the size of
+both the left and right right parahippocampus.
+Another cluster of model importance contained taxa associated with the
 basal forebrain, especially the cingulate and the accumbens. This
-cluster contains the previously-discussed \emph{B. ovatus} and \emph{B.
-uniformis}, but also \emph{Alistipes finegoldii and Streptococcus
-salivaris}. Reduction in the nucleus accumbens has been associated with
-depression symptoms \cite{wackerRoleNucleusAccumbens2009},
-and increased levels of the \emph{Alistipes} genus have been
-observed in patients with depressive disorder
-\cite{jiangAlteredFecalMicrobiota2015}.
+cluster contains \emph{B. ovatus} and \emph{B. uniformis},
+but also \emph{Alistipes finegoldii and Streptococcus salivaris}.
 
@@
 the right anterior cingulate (respectively, 3.0\% and 2.6\% relative
-importances), which has been linked to social cognition and reward-based
-decision making
-\cite{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002}.
-Equol has a strong estrogenic effect
-\cite{setchellSEquolPotentLigand2005},
-and in the anterior cingulate, estrogen has been shown to regulate
-pain-related aversion
-\cite{xiaoEstrogenAnteriorCingulate2013}.
-Another example of this pattern is \emph{R. gnavus}, which was the only
+importances). Another example of this pattern is \emph{R. gnavus}, which was the only
 species with significant negative association with cognitive function in linear models.
-It is heavily associated with the left pars opercularis (2.7\% relative
-importance). \textit{R. gnavus} is increased in individuals with insulin
-resistance and obesity \cite{leyHumanGutMicrobes2006},
-conditions that are known to produce structural abnormalities in the brain
-\cite{opelBrainStructuralAbnormalities2021}.
+It is heavily associated with the left pars opercularis (2.7\% relative importance). 
 Finally, \emph{Coprococcus comes} displays an importance

### 2023-03-15 manuscript/main.tex
fix author spelling

@@
     \parbox{\linewidth}{\centering
-        Kevin S. Bongham,$^{1}$
-        Guilherme Fahur Botino,$^{1}$
+        Kevin S. Bonham,$^{1}$
+        Guilherme Fahur Bottino,$^{1}$
         Shelley Hoeft McCann,$^{1}$

### 2023-03-15 manuscript/main.tex
begin Vanja edits again

@@
 wexlerae} was dominant for the predicting the size of the parahippocampal region
-in both the left and right hemispheres, while several \textit{Bacteroitetes} species,
-including GABA-producing \emph{B. ovatus}, were important for predicting
+in both the left and right hemispheres, while several species from
+the phylum \textit{Bacteroidetes}, including GABA-producing
+\emph{B. ovatus}, were important for predicting
 the size of the left accumbens area, but not the right. These
@@
 of gut microbes on the development of autism spectrum disorder
-\cite{laueProspectiveAssociationsInfant2020,wanUnderdevelopmentGutMicrobiota2021}
-and particular microbial taxa have been associated with depression
+\cite{laueProspectiveAssociationsInfant2020,wanUnderdevelopmentGutMicrobiota2021},
+and specific microbial taxa have been associated with depression
 \cite{mayneris-perxachsMicrobiotaAlterationsProline2022,valles-colomerNeuroactivePotentialHuman2019}
@@
 \cite{laueDevelopingMicrobiomeBirth2022}.
-Development \emph{in utero} is believed to be mostly sterile, but
-is rapidly seeded at birth through contact with the birth canal (if
-birthed vaginally), caregivers, food sources (breast milk or formula),
+Fetal development is believed to be mostly sterile, but newborns
+are rapidly seeded at birth through contact with the birth canal (if
+birthed vaginally), caregivers, food (breast milk or formula),
 and other environmental sources
@@
 \cite{koenigSuccessionMicrobialConsortia2011}.
-Many of these microbes have specialized metabolic capabilities
-for digesting human breast milk, such as \emph{Bifidobacterium infantis}
-and \emph{Bacteroides fragilis}
+Many of these bacteria have specialized metabolic capabilities
+for digesting human breast milk, such as \emph{Bifidobacterium longum}
+subsp. \textit{infantis} and \emph{Bacteroides fragilis}
 \cite{selaGenomeSequenceBifidobacterium2008}.
 Upon the introduction of solid foods, the gut
-microbiome undergoes another categorical transformation, with most taxa
-of the infant microbiome being replaced by taxa more reminiscent of
-adult microbiomes
+microbiome undergoes another categorical transformation;
+its diversity increases, and most taxa of the infant microbiome
+are replaced by taxa more reminiscent of adult microbiomes
 \cite{backhedDynamicsStabilizationHuman2015}.
@@
 \cite{knudsenSensitivePeriodsDevelopment2004}
-during which neural plasticity is particularly high. The timing of these SPs
-is driven by both genetic and environmental cues, including the gut microbiome
-\cite{cowanAnnualResearchReview2020}.
+during which neural plasticity is particularly high.
 Emerging evidence suggests that the timing and duration of SPs
 may be driven in part by cues from the developing gut microbiome
-\cite{callaghanNestedSensitivePeriods2020}.
-As such,
-understanding the normal spectrum of healthy microbiome
+\cite{callaghanNestedSensitivePeriods2020,cowanAnnualResearchReview2020}.
+As such, understanding the normal spectrum of healthy microbiome
 development and how it relates to normal neurocognitive development may

### 2023-03-15 manuscript/main.tex
through figure 2

@@
 \cite{wechslerWechslerIntelligenceScale1949}.
-Finally, we also assessed emerging brain structure magnetic resonance imaging
+Finally, we also assessed emerging brain structure using magnetic resonance imaging
 (MRI). Through a combination of classical statistical analysis and machine
@@
 361 healthy and neurotypically-developing children (163 female) between 82 days and 10 years of age
-(Table 1,  Figure 1A, Supplementary Figure 1)\cite{forrestAdvancingScienceChildren2018}
+(Table 1,  Figure 1A, Supplementary Figure 1)
 using a variety of orthogonal microbial and neurocognitive measures.
@@
 stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
-high-\emph{Prevotella copri} (Supplementary Figure 2). Overall variation
-in gut microbial genes and also in brain volume profiles was similarly
-driven largely by subject age (Figure 1E-F).
+high-abundance \emph{Prevotella copri} (Supplementary Figure 2).
+Overall subject-to-subject variation in gut microbial genes
+and brain volume profiles was similarly driven largely by subject age (Figure 1E-F).
 
@@
     \end{tabular}
-    \caption{\label{tab:demo}Subject demographics}
+    \caption{\label{tab:demo}Subject demographics of the ECHO RESONANCE cohort.}
 \end{table}
@@
 with ASD \cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021},
-adult cognitive skill (e.g., executive functioning)\cite{magnussonRelationshipsDietrelatedChanges2015},
-and adult mental health (both anxiety and depressive symptioms)\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022}
-we sought to determine if specific taxa or gene functions were linked with normal cognitive development in children.
+adult cognitive skill (e.g., executive functioning)
+\cite{magnussonRelationshipsDietrelatedChanges2015},
+and adult mental health (both anxiety and depressive symptioms)
+\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022},
+we sought to determine if specific taxa or gene functions were linked
+with normal cognitive development in children.
 To test if variations in gut microbial taxa,
-their genes, or their metabolism are associated with neurocognitive
+their genes, or their metabolism were associated with neurocognitive
 development, we used permutation analysis of variance (PERMANOVA) 
@@
 function in children under 6 months, though this may be due to the low
-taxonomic diversity and broad lack of overlap between taxa in infants.
+taxonomic diversity and broad lack of overlap between taxa in infants (Figure 1G).
 As expected, age was significantly associated with microbial beta
@@
 and concurrent cognitive ability score (Table 2, Figure 2C-G) and found that
-several metabolic pathways were significantly enriched or depleted in
+several metabolic pathways were either significantly enriched or depleted in
 children with higher cognitive ability scores. This was true both when
-considering all ages together, though the enrichment of most
+considering all ages (0-120 months) together, though the enrichment of most
 pathways was more pronounced either in children under 6 months or those over 18
-months. For example, genes for degrading the 3-carbon SCFA propionate
+months. For example, genes for degrading the 3-carbon SCFA, propionate,
 were significantly depleted in children with higher cognitive ability
@@
 unsupervised non-parametric machine learning (ML) method that enables
-the identification of underlying patterns in large numbers of individual
-features (here, microbial species). Previous studies have reported
+the identification of underlying patterns in large numbers of individual features
+(here, microbial species).
+Previous studies have reported
 successful use of RFs for processing highly-dimensional and sparse data

### 2023-03-15 manuscript/main.tex
more vanja

@@
 factors, we assessed the performance of RF regressors where maternal
-education (a proxy of socioeconomic status (SES)), sex, and age were
-included as possible predictors, either alone or in combination with
-microbial taxonomic profiles (Table 3).
+education (here used as a proxy of socioeconomic status (SES)),
+sex, and age were included as possible predictors, either alone
+or in combination with microbial taxonomic profiles (Table 3).
 

### 2023-03-15 manuscript/main.tex
finish Vanja review

@@
 function, they might be reflected in changes in neuroanatomy. We again
-employed a Random Forest modeling approach to associate gut taxonomic
+employed a RF modeling approach to associate gut taxonomic
 profiles with individual brain regions identified in MRI scans.
@@
 
-While RF models for multiple brain regions had many important microbial taxa, others were
-dominated by a small number of taxa, many of which
-were also identified in LMs and RF models of cognitive function. 
+While RF models for multiple brain regions had many important microbial taxa,
+others were dominated by a small number of taxa, many of which
+were also identified in LMs and RF models of cognitive function (Figures 2 and 3). 
 In general, for all segments, a
@@
 Among these, the most important variable is \emph{Anaerostipes hadrus}
-\hl{!!importance stats!!} (Figure 4B).
+\hl{!!importance stats!!} (Figure 4B and C).
 \emph{B. wexlerae}, which was important in RF models of cognitive function,
@@
 clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (4.0\% relative importance) -
+is the highest for a microbe in RF models (4.0\% relative importance);
 only age had higher relative importance in any model. At the same time,
@@
 then pooled onto a plate and sequenced on the Illumina NextSeq 550
-platform using 150+150 bp paired-end ``high output'' chemistry,
+platform using 150+150 bp paired-end "high output";' chemistry,
 generating 400 million raw reads and 120 Gb of sequence per plate.
+
+\subsubsection*{Computational analysis of metagenomes}
+
+Shotgun metagenomic sequences were analyzed using the bioBakery suite of
+computational tools \cite{beghiniIntegratingTaxonomicFunctional2021}.
+First, KneadData (v0.7.7) was used to perform quality
+control of raw sequence reads, such as read trimming and removal of
+reads matching a human genome reference. Next, MetaPhlAn (v3.0.7, using
+database mpa\_v30\_CHOCOPhlAn\_201901) was used to generate taxonomic
+profiles by aligning reads to a reference database of marker genes.
+Finally, HUMAnN (v3.0.0a4) was used to functionally profile the
+metagenomes.
 
@@
 status) on participants from different age groups. Age (in months) was
-provided as a covariate for all models (Table 3).
+provided as a covariate for all models (Table 5).
 
@@
 
-\subsubsection*{Computational analysis of metagenomes}
-
-Shotgun metagenomic sequences were analyzed using the bioBakery suite of
-computational tools \cite{beghiniIntegratingTaxonomicFunctional2021}.
-First, KneadData (v0.7.7) was used to perform quality
-control of raw sequence reads, such as read trimming and removal of
-reads matching a human genome reference. Next, MetaPhlAn (v3.0.7, using
-database mpa\_v30\_CHOCOPhlAn\_201901) was used to generate taxonomic
-profiles by aligning reads to a reference database of marker genes.
-Finally, HUMAnN (v3.0.0a4) was used to functionally profile the
-metagenomes.
-
 \subsubsection*{Data and code availability}

### 2023-03-15 manuscript/main.tex
add discussion section with deleted lines

@@
 
+\subsection*{Deleted}
+
+SCFAs are produced by anaerobic fermentation of dietary fiber
+and have been linked with immune system regulation as well as directly
+with brain function \cite{dalileRoleShortchainFatty2019}
+
+Menaquinone has several isoforms, one of
+which, MK-4, has been found to be decreased in children diagnosed with ASD
+\cite{dongCorrelationSerumConcentrations2021}
+and is potentially neuroprotective in both rodents and humans
+\cite{elkattawyVitaminK2Menaquinone72022}.
+
+Glutamate is a
+critical neurotransmitter controlling neuronal excitatory/inhibitory
+signaling along with gamma-aminobutyric acid (GABA), and their balance
+in the brain controls neural plasticity and learning, particularly in
+the developing brain
+\cite{cohenkadoshLinkingGABAGlutamate2015,palomo-buitragoGlutamateInteractionsObesity2019}.
+
+Tryptophan metabolism, including
+microbial metabolism of tryptophan, has previously been linked with
+autism in children
+\cite{hoshinoBloodSerotoninFree1984,xiaoFecalMicrobiomeTransplantation2021}.
+
+For children under 6 months, only \emph{B. wexlerae} was
+significantly associated with cognitive function in linear models, but
+was ranked 20th in importance for RF models.
+
+The importance metric
+employed (MDI) is a measure of how well the variable differentiates
+(splits) sets of samples by creating subgroups that reduce the
+intra-group deviations, while also accounting for the amount of samples
+affected by a split.
+
+Most healthy individuals have a
+rightward asymmetry in the nucleus accumbens,
+which is involved in reward learning and risk taking behaviours
+\cite{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012}.
+Reduced asymmetry between right and left nucleus accumbens
+has been associated with substance use disorder in young adults
+\cite{caoMappingCorticalSubcortical2021}. The accumbens area is also
+associated with reward control, and in individuals diagnosed with ADHD,
+it has been shown to have a divergent neuromorphology
+\cite{hoogmanSubcorticalBrainVolume2017}.
+
+[\emph{Anaerostipes hadrus}] is a butyrate-producing anaerobe that has been positively
+associated with cognitive function
+\cite{kantGenomeSequenceButyrateProducing2015,liCorrelationsGutMicrobiota2022}.
+Its importance is, however, heavily loaded on the volume
+of entire major lobes (frontal, parietal), which cannot be readily
+correlated to specific cognitive functions
+
+Reduction in the nucleus accumbens has been associated with
+depression symptoms \cite{wackerRoleNucleusAccumbens2009},
+and increased levels of the \emph{Alistipes} genus have been
+observed in patients with depressive disorder
+\cite{jiangAlteredFecalMicrobiota2015}.
+
+[The right anterior cingulate] has been linked to social cognition and reward-based
+decision making
+\cite{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002}.
+Equol has a strong estrogenic effect
+\cite{setchellSEquolPotentLigand2005},
+and in the anterior cingulate, estrogen has been shown to regulate
+pain-related aversion
+\cite{xiaoEstrogenAnteriorCingulate2013}.
+
+\textit{R. gnavus} is increased in individuals with insulin
+resistance and obesity \cite{leyHumanGutMicrobes2006},
+conditions that are known to produce structural abnormalities in the brain
+\cite{opelBrainStructuralAbnormalities2021}.
+
+\subsection*{Main}
+
 The relationship between the gut microbiome and brain function via the

### 2023-03-22 manuscript/main.tex
adding some stuff to discussion

@@
 
-Both \emph{A. equolofaciens} and \emph{A. celatus,} two closely-related
+Both \emph{A. equolofaciens} and \emph{A. celatus}, two closely-related
 equol-producing species, are examples of the second contribution
@@
 
-Most healthy individuals have a
-rightward asymmetry in the nucleus accumbens,
-which is involved in reward learning and risk taking behaviours
-\cite{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012}.
-Reduced asymmetry between right and left nucleus accumbens
-has been associated with substance use disorder in young adults
-\cite{caoMappingCorticalSubcortical2021}. The accumbens area is also
-associated with reward control, and in individuals diagnosed with ADHD,
-it has been shown to have a divergent neuromorphology
-\cite{hoogmanSubcorticalBrainVolume2017}.
-
 [\emph{Anaerostipes hadrus}] is a butyrate-producing anaerobe that has been positively
@@
 \cite{jiangAlteredFecalMicrobiota2015}.
-
-[The right anterior cingulate] has been linked to social cognition and reward-based
-decision making
-\cite{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002}.
-Equol has a strong estrogenic effect
-\cite{setchellSEquolPotentLigand2005},
-and in the anterior cingulate, estrogen has been shown to regulate
-pain-related aversion
-\cite{xiaoEstrogenAnteriorCingulate2013}.
 
@@
 
+We identified several species in the family Eggerthelaceae that were
+associated with cognitive function, including \emph{Gordonibacter
+pamelaeae, Aldercreutzia equolofaciens, Asaccharobacter celatus}
+(formerly regarded a subspecies of \emph{A. equolofaciens}
+\cite{takahashiCompleteGenomeSequence2021}),
+and \emph{Eggerthella lenta}. Many members of this family are
+known in part due to unique metabolic activities. For example, \emph{A.
+equolofaciens} produces the nonsteroidal estrogen equol from isoflaven
+found in soybeans \cite{wangEnantioselectiveSynthesisSEquol2005}.
+This is particularly intriguing, since both \emph{A. equolofaciens} and \emph{A. celatus}
+were also the most important features in RF models of the
+right anterior cingulate, which has been linked to social cognition
+and reward-based decision making\cite{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002},
+and estrogen has previously been shown to have activity 
+in this brain region\cite{xiaoEstrogenAnteriorCingulate2013}.
+\emph{G. pamelaeae} also has a unique metabolism that may impact the brain,
+as it can metabolize the polyphenol ellagic acid
+(found in pomegranates and some berries) into urolithin, which has been
+shown in some studies to have a neuroprotective effect
+\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
+While we could not find evidence of neurological roles for \emph{E. lenta},
+it has been extensively studied for
+its ability to metabolize drug compounds such as the plant-derived heart
+medication digoxin \cite{haiserPredictingManipulatingCardiac2013}.
+The metabolic versatility of this clade, and the large number of
+species that are associated with cognition make these microbes prime
+targets for further mechanistic studies.
+
+Other bacteria that were associated with both cognitive ability and
+brain structure had \hl{notable hints in the literature}. % This is stupid - find a better way to say this
+For example, RF models predicting the size of the left nucleus accumbens
+(but not the right) were dominated by \textit{Bacteroides} species,
+especially \textit{B. vulgatus}, \textit{B. uniformis}, and \textit{B. ovatus}.
+\textit{B. ovatus} has previously been shown to produce
+a number of neuroactive compounds including SCFAs and neurotransmitters
+such as GABA\cite{horvathBacteroidesOvatusColonization2022}.
+Another \textit{Bacteroides} species, \textit{B. fragilis} was
+able to ameliorate ASD-like symptoms in a mouse model
+by consuming a microbially-dependent metabolite 4-ethylphenylsulfate (4EPS)
+\cite{hsiaoMicrobiotaModulateBehavioral2013},
+while \textit{B. ovatus} is the primary producer of a precursor to 4EPS
+\cite{needhamGutderivedMetaboliteAlters2022} and increased anxiety
+behaviors
+
+Most healthy individuals have a
+rightward asymmetry in the nucleus accumbens,
+which is involved in reward learning and risk taking behaviours
+\cite{ernstAmygdalaNucleusAccumbens2005,yauNucleusAccumbensResponse2012}.
+Reduced asymmetry between right and left nucleus accumbens
+has been associated with substance use disorder in young adults
+\cite{caoMappingCorticalSubcortical2021}. The accumbens area is also
+associated with reward control, and in individuals diagnosed with ADHD,
+it has been shown to have a divergent neuromorphology
+\cite{hoogmanSubcorticalBrainVolume2017}.
+
 The use of shotgun metagenomic sequencing enabled us to get
@@
 several of the taxa that were positively associated with cognitive
-function in this study, including \emph{B. wexlerae}, the only species
-identified by linear models in children under 6 months, \emph{D.
+function in this study, including \emph{B. wexlerae}, \emph{D.
 longicatena}, \emph{R. faecis}, and \emph{A. finegoldii} are
@@
 sequencing.
-
-We identified several species in the family Eggerthelaceae that were
-associated with cognitive function, including \emph{Gordonibacter
-pamelaeae, Aldercreutzia equolofaciens, Asaccharobacter celatus}
-(formerly regarded a subspecies of \emph{A. equolofaciens}
-\cite{takahashiCompleteGenomeSequence2021}),
-and \emph{Eggerthella lenta}. Many members of this family are
-known in part due to unique metabolic activities. For example, \emph{A.
-equolofaciens} produces the nonsteroidal estrogen equol from isoflaven
-found in soybeans \cite{wangEnantioselectiveSynthesisSEquol2005},
-and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
-(found in pomegranates and some berries) into urolithin, which has been
-shown in some studies to have a neuroprotective effect
-\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
-\emph{E. lenta} has been extensively studied for
-its ability to metabolize drug compounds such as the plant-derived heart
-medication digoxin \cite{haiserPredictingManipulatingCardiac2013}.
-The metabolic versatility of this clade, and the large number of
-species that are associated with cognition make these microbes prime
-targets for further mechanistic studies.
 
@@
 Image matrix and field-of-view were adjusted by infant size to achieve the desired image resolution.
-For children under ~4 years of age, MRI was performed during natural non-sedated sleep using previously described methods \hl{(Doug's paper ref)}.
+For children under ~4 years of age, MRI was performed during natural non-sedated sleep using previously described methods\cite{deanPediatricNeuroimagingUsing2014}.
 Older children were scanned awake whilst watching a movie or favorite TV show.

### 2023-03-22 manuscript/main.tex
final changes to discussion

@@
 
-\subsection*{Deleted}
-
-SCFAs are produced by anaerobic fermentation of dietary fiber
-and have been linked with immune system regulation as well as directly
-with brain function \cite{dalileRoleShortchainFatty2019}
-
-Menaquinone has several isoforms, one of
-which, MK-4, has been found to be decreased in children diagnosed with ASD
-\cite{dongCorrelationSerumConcentrations2021}
-and is potentially neuroprotective in both rodents and humans
-\cite{elkattawyVitaminK2Menaquinone72022}.
-
-Glutamate is a
-critical neurotransmitter controlling neuronal excitatory/inhibitory
-signaling along with gamma-aminobutyric acid (GABA), and their balance
-in the brain controls neural plasticity and learning, particularly in
-the developing brain
-\cite{cohenkadoshLinkingGABAGlutamate2015,palomo-buitragoGlutamateInteractionsObesity2019}.
-
-Tryptophan metabolism, including
-microbial metabolism of tryptophan, has previously been linked with
-autism in children
-\cite{hoshinoBloodSerotoninFree1984,xiaoFecalMicrobiomeTransplantation2021}.
-
-For children under 6 months, only \emph{B. wexlerae} was
-significantly associated with cognitive function in linear models, but
-was ranked 20th in importance for RF models.
-
-The importance metric
-employed (MDI) is a measure of how well the variable differentiates
-(splits) sets of samples by creating subgroups that reduce the
-intra-group deviations, while also accounting for the amount of samples
-affected by a split.
-
-[\emph{Anaerostipes hadrus}] is a butyrate-producing anaerobe that has been positively
-associated with cognitive function
-\cite{kantGenomeSequenceButyrateProducing2015,liCorrelationsGutMicrobiota2022}.
-Its importance is, however, heavily loaded on the volume
-of entire major lobes (frontal, parietal), which cannot be readily
-correlated to specific cognitive functions
-
-Reduction in the nucleus accumbens has been associated with
-depression symptoms \cite{wackerRoleNucleusAccumbens2009},
-and increased levels of the \emph{Alistipes} genus have been
-observed in patients with depressive disorder
-\cite{jiangAlteredFecalMicrobiota2015}.
-
-\textit{R. gnavus} is increased in individuals with insulin
-resistance and obesity \cite{leyHumanGutMicrobes2006},
-conditions that are known to produce structural abnormalities in the brain
-\cite{opelBrainStructuralAbnormalities2021}.
-
-\subsection*{Main}
-
 The relationship between the gut microbiome and brain function via the
@@
 directly investigates microbial species and their genes in relation to
-typical development in young children. Understanding the gut-brain-axis
+typical development in young children. Understanding the gut-brain-microbiome axis
 in early life is particularly important, since differences or
@@
 \cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
-While we could not find evidence of neurological roles for \emph{E. lenta},
+While we could not find published evidence of neurological roles for \emph{E. lenta},
 it has been extensively studied for
@@
 medication digoxin \cite{haiserPredictingManipulatingCardiac2013}.
-The metabolic versatility of this clade, and the large number of
+The metabolic versatility of this clade and the large number of
 species that are associated with cognition make these microbes prime
@@
 Other bacteria that were associated with both cognitive ability and
-brain structure had \hl{notable hints in the literature}. % This is stupid - find a better way to say this
+brain structure had also have intriguing connections in previous studies.
 For example, RF models predicting the size of the left nucleus accumbens
-(but not the right) were dominated by \textit{Bacteroides} species,
+were dominated by \textit{Bacteroides} species,
 especially \textit{B. vulgatus}, \textit{B. uniformis}, and \textit{B. ovatus}.
@@
 \cite{hsiaoMicrobiotaModulateBehavioral2013},
-while \textit{B. ovatus} is the primary producer of a precursor to 4EPS
+while \textit{B. ovatus} is a primary producer of a precursor to 4EPS
 \cite{needhamGutderivedMetaboliteAlters2022} and increased anxiety
-behaviors
-
-Most healthy individuals have a
+behaviors in mice.
+The fact that the importance of these species in RF models,
+and the high performance of RF models in predicting this region
+were unilateral (affecting the left but not the right accumbens area)
+was also notable. Most healthy individuals have a
 rightward asymmetry in the nucleus accumbens,
@@
 genes for the metabolism of SCFAs, both their degradation and synthesis,
-are associated with cognitive function scores. However, while the
+are associated with cognitive function scores (Figure 2). 
+SCFAs are produced by anaerobic fermentation of dietary fiber
+and have been linked with immune system regulation as well as directly
+with brain function \cite{dalileRoleShortchainFatty2019}.
+Microbial metabolism of other potentially neuroactive molecules,
+were also associated with cognitive ability. We were particularly interested
+in genes for metabolizing glutamate, which is a
+critical neurotransmitter controlling neuronal excitatory/inhibitory
+signaling along with gamma-aminobutyric acid (GABA), and their balance
+in the brain controls neural plasticity and learning, particularly in
+the developing brain \cite{cohenkadoshLinkingGABAGlutamate2015,palomo-buitragoGlutamateInteractionsObesity2019}.
+However, while the
 differential abundance of genes for the metabolism of neuroactive
@@
 
-% and alterations on the striatal
-% dopamine transporters can cause effects resembling hyperactivity and
-% attention deficit
-% \cite{yaelDisinhibitionNucleusAccumbens2019},
-% and that \emph{B. uniformis} is gut-microbial modulator of the brain
-% dopamine transporter
-% \cite{hartstraInfusionDonorFeces2020}.
-
 This analysis allowed us to establish links between microbial taxa and
@@
 we cannot test causality or the chemistry behind the interactions
-between gut microbial taxa, gut, and brain, this study provides clear
-and statistically significant associations between the infant and early
-child gut microbiota and neurocognition. Future studies should focus on
-characterizing the early-life microbiome and neurocognitive development
-across different geographic regions and lifestyles such as covering
-traditionally understudied low-resource urban, peri-urban and rural
-communities to obtain the more comprehensive understanding of the
+between gut microbial taxa and their genes, the gut, and the brain,
+this study provides clear and statistically significant associations
+between the infant and early child gut microbiota and neurocognition.
+Future studies should focus on characterizing the early-life microbiome
+and neurocognitive development across different geographic regions
+and lifestyles such as covering traditionally understudied populations,
+such as low-resource urban, peri-urban and rural communities
+to obtain the more comprehensive understanding of the
 variability within the different gut microbiomes reflects on
@@
 genomic approaches to understand microbial metabolism at the molecular
-level is the key, as the metabolic functions shape and influence the
+level is key, as the metabolic functions shape and influence the
 human host and its health. The discovery of the neuroactive metabolites

### 2023-03-22 other/anpan.qmd
work on anpan

---
title: "Running Anpan"
format: html
---

## Installation

```{r}
#| eval: false
install.packages(c("ape", 
                   "data.table",
                   "dplyr", 
                   "fastglm",
                   "furrr", 
                   "ggdendro",
                   "ggnewscale",
                   "ggplot2",
                   "loo",
                   "patchwork",
                   "phylogram",
                   "posterior",
                   "progressr",
                   "purrr",
                   "R.utils",
                   "remotes",
                   "stringr",
                   "tibble",
                   "tidyselect"), Ncpus = 8) # add Ncpus = 4 to go faster

install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
```

```{r}
#| eval: false
library(cmdstanr)
check_cmdstan_toolchain()
install_cmdstan(cores = 4)
```

```{r}
#| eval: false
remotes::install_github("biobakery/anpan")
```

## Running

After identifying samples with *B. wexlerae* in our sample set,
I generated a table with stratified gene families by running

```fish
echo -e "sample\tfeature\tabundance" > data/bwexlerae_genefamilies.tsv
for s in (cat has_bwex.txt)
    fd "$s"_genefamilies /grace/sequencing/processed/mgx/humann/main/ -x rg s__Blautia_wexlerae | sd '^q' "$s\t" >> data/bwexlerae_genefamilies.tsv
end

sd '^.+\t(.+\t.+\t.+)$' '$1' data/bwexlerae_genefamilies.tsv
sd '^FG\d+_S\d+\t$' '' data/bwexlerae_genefamilies.tsv

```

and then added a header manually.

```{r}
library(data.table)

df <- fread("data/bwexlerae_genefamilies.tsv")
df2 <- dcast(df[, .(feature, sample, abundance)],
                 feature ~ sample,
                 value.var = "abundance",
                 fun.aggregate = sum)
names(df2)[1] = "# Gene Family"
```

```{r}
library(anpan)
future::plan(future::multicore, workers=16)
```

```{r}
bwex_results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
                        bug_file          = "./data/bwexlerae_genefamilies_wide.tsv",
                        meta_file         = "./data/manuscript_meta.csv",
                        out_dir           = "./data/anpan",
                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
                        filtering_method  = "kmeans",
                        model_type        = "fastglm",
                        covariates        = c("ageMonths", "sex", "education"),
                        outcome           = "cogScore",
                        plot_ext          = "pdf",
                        save_filter_stats = TRUE)
```

```{r}
input_path = file.path("data", "anpan", "filter_stats",
                       "filtered_bwexlerae_genefamilies_wide.tsv")

model_input = fread(input_path)

plot_results(res         = bwex_results,
             model_input = model_input,
             covariates  = c("ageMonths", "sex", "education"), 
             outcome     = "cogScore",
             bug_name    = "Blautia wexlerae",
             cluster     = "both",
             show_trees  = TRUE)

```

### 2023-03-23 manuscript/main.tex
add author contributions

@@
 
+All data needed to evaluate the conclusions in the paper
+are present in the paper and/or the Supplementary Materials.
 Taxonomic and functional microbial profiles, as well as subject
@@
 \cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021}.
- All code for
-data analysis and figure generation, as well as scripts for automated
+All code for data analysis and figure generation, as well as scripts for automated
 download of input files are available on github \hl{\{\{CITE zenodo\}\}}.
 
+\section*{References}
+
+\printbibliography
+
 \section*{Acknowledgements}
-
-Research funding: NIH UG3 OD023313 and Wellcome LEAP 1kD.
-
-\section*{References}
-
-\printbibliography
+% https://www.science.org/content/page/science-advances-information-authors#acknowledgments
+
+This research was funded by NIH UG3 OD023313 and Wellcome LEAP 1kD.
+The authors declare that they have no competing interests.
+
+% https://credit.niso.org/
+
+KSB, VKC, CH, and SD were responsible for conceptualization,
+methodology and project administration.
+VKC, SD, and VD were responsible for funding aqusition.
+KSB, MB, FB, RCL, and JB contributed to data curation.
+Formal analyses were conducted by KSB, GFB, and MB.
+Investigation was performed by KSB, GFB, SHM, EW, FB, RCL, and MB.
+Resources were provided by SHM, EW, FB, RCL, JB, CH, and MB.
+Software and data visualizations were produced by KSB and GFB.
+KSB, MB, SD, VD, and VKC were responsible for supervision.
+Validation was performed by KSB, SHM, JB, and MB.
+KSB and GFB wrote the initial draft, 
+and review and editing were provided by SD, CH, and VKC.
 

### 2023-03-23 src/Resonance.jl
add more package refs

@@ -116,7 +116,6 @@ using ReTest
 
 using Airtable
 using Arrow
-using AlgebraOfGraphics
 using CairoMakie
 using CategoricalArrays
 using CodecZlib

### 2023-03-23 src/microbiomejl_interface.jl
fixed typo on microbiome_interface.jl

@@ -56,5 +56,5 @@ end
 function comm2wide(cmp::CommunityProfile; feature_filter=identity, sample_filter=identity, feature_funct=name)
     cmp = cmp[feature_filter(features(cmp)), sample_filter(samples(cmp))]
     md = DataFrame(get(cmp))
-    return hcat(md, DataFrame(collect(abundances(cmp)'), feature_func.(features(cmp)), makeunique=true))
+    return hcat(md, DataFrame(collect(abundances(cmp)'), feature_funct.(features(cmp)), makeunique=true))
 end

### 2023-03-23 manuscript/TableS1.jl
Fixed path to Table S1A and S1B output

@@ -65,11 +65,11 @@ end
 
 supptblA = suppl_table(regression_currentCogScores_00to06mo_onlytaxa)
 tblformat_suppl_table!(supptblA)
-CSV.write("TableS1A.csv", supptblA)
+CSV.write("manuscript/assets/TableS1A.csv", supptblA)
 
 supptblB = suppl_table(regression_currentCogScores_18to120mo_onlytaxa)
 tblformat_suppl_table!(supptblB)
-CSV.write("TableS1B.csv", supptblB)
+CSV.write("manuscript/assets/TableS1A.csv", supptblB)
 
 ## 4. LaTeX outputs
 

### 2023-03-23 manuscript/TableS1.jl
Removed deprecated call to RandomForestClassifier on TableS1.jl

@@ -14,7 +14,6 @@ isdir(tablefiles()) || mkpath(tablefiles())
 
 ## 1. Loading the model files
 
-RandomForestClassifier = MLJ.@load RandomForestClassifier pkg=DecisionTree
 RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
 # concurrent cogScore reression from taxonomic profiles
 JLD2.@load "models/2023-02-15/regression_currentCogScores_00to06mo_onlytaxa.jld"

### 2023-03-23 manuscript/TableS2.jl
Fixed path to Table S2 output

@@ -64,7 +64,7 @@ tableS2 = reduce(
 
 tableS2.variable = uppercasefirst.(map(x -> replace(x, "-"=>" "), tableS2.variable))
 
-CSV.write("TableS2.csv", tableS2)
+CSV.write("manuscript/assets/TableS2.csv", tableS2)
 
 pretty_table(tableS2;
     header = [ "Segment", "Mean absolute proportional error (MAPE)", "Correlation coefficient (R)" ],

### 2023-03-23 manuscript/TableS2.jl
Substituted comprehension by mapreduce+do block on Table S2

@@ -62,6 +62,11 @@ tableS2 = reduce(
     [ DataFrame(:variable => ordered_brain_segments_list[i], :Test_MAPE => mean(brain_models[ordered_brain_segments_list[i]].merits.Test_MAPE), :Test_Cor => mean(brain_models[ordered_brain_segments_list[i]].merits.Test_Cor)) for i in eachindex(ordered_brain_segments_list) ]
 )
 
+tableS2 = mapreduce( vcat, eachindex(ordered_brain_segments_list)) do i
+    DataFrame(:variable => ordered_brain_segments_list[i],
+              :Test_MAPE => mean(brain_models[ordered_brain_segments_list[i]].merits.Test_MAPE), 
+              :Test_Cor => mean(brain_models[ordered_brain_segments_list[i]].merits.Test_Cor))
+end
 tableS2.variable = uppercasefirst.(map(x -> replace(x, "-"=>" "), tableS2.variable))
 
 CSV.write("manuscript/assets/TableS2.csv", tableS2)

### 2023-03-23 manuscript/main.tex
add more package refs
add zenodo ref

@@
 generation of plots was performed using the julia programming language
-\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021}.
+\cite{bezansonJuliaFreshApproach2017,
+    bonhamMicrobiomeJlBiobakeryUtils2021,
+    danischMakieJlFlexible2021,
+    blaomMLJJuliaPackage2020,
+    ben_sadeghi_2022_7359268,
+    dahua_lin_2023_7695673,
+    douglas_bates_2023_7734970,
+    douglas_bates_2023_7734970,
+    kevin_bonham_2023_7647510}.
 All code for data analysis and figure generation, as well as scripts for automated
-download of input files are available on github \hl{\{\{CITE zenodo\}\}}.
+download of input files are available on github \cite{kevin_bonham_2023_7647510}.
 

### 2023-03-24 manuscript/main.tex

@@
     \caption{
-        \label{fig:1} Study design and overview. (\textbf{A}) Stool samples, cognitive assessments and
-        neuroimaging were collected from participants in an accelerated
-        longitudinal design. (\textbf{B}) Cognitive function scores are assessed by
-        different instruments, but may be normalized to a common IQ-like scale.
+        \label{fig:1} Study design and overview. (\textbf{A}) Stool samples,
+        cognitive assessments and neuroimaging were collected from participants
+        at different ages throughout the first years of life.
+        (\textbf{B}) Cognitive function scores are assessed by
+        different instruments, but may be normalized using full-scale composite scores.
         (\textbf{C-D}) Principal coordinates analysis using Bray-Curtis dissimilarity on

### 2023-03-25 manuscript/main.tex
add supplementary table numbers

@@
 361 healthy and neurotypically-developing children (163 female) between 82 days and 10 years of age
-(Table 1,  Figure 1A, Supplementary Figure 1)
+(Table 1,  Figure 1A, Figure S1)
 using a variety of orthogonal microbial and neurocognitive measures.
@@
 stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
-high-abundance \emph{Prevotella copri} (Supplementary Figure 2).
+high-abundance \emph{Prevotella copri} (Figure S2).
 Overall subject-to-subject variation in gut microbial genes
@@
 significant when testing the relationship with LMs (Figure
-3A-B, Supplementary Table \hl{XX}A-B); aside from except for \emph{Eubacterium ramulus}
+3A-B, Table S1A-B); aside from except for \emph{Eubacterium ramulus}
 all taxa that were significant using LMs belonged to a group
@@
 Some brain regions were more readily
-predicted by RF models trained on microbial taxa (Table 4, Supplementary
-Table \hl{XX}), in particular those that were highly correlated with age.
+predicted by RF models trained on microbial taxa (Table 4, Table S2), in particular those that were highly correlated with age.
 These included the L/R lingual gyrus (mean RF correlations, Left =
@@
 responsible for one third of the fitness-weighted cumulative importance,
-regardless of model performance (\hl{Suplementary Table XX}, Figure 4 A and B).
+regardless of model performance (Figure 4A-B, Table S2).
 We observed two major patterns of importance distribution from
@@
 \cite{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
-(Supplementary Figure 3).
+(Figure S3).
 

### 2023-03-25 manuscript/TableS3.jl
move T5 -> TS3

@@ -14,16 +14,16 @@ isdir(tablefiles()) || mkpath(tablefiles())
 
 ## Building the Table
 
-table5 = DataFrame(
+tableS3 = DataFrame(
     :input_set => 1:10,
     :age_bracket => vcat( repeat([ "0 to 6 months" ], 5), repeat( [ "18 to 120 months" ], 5)),
     :microbial_feature => repeat([ "Not provided", "Taxonomic profile", "Taxonomic profile", "Functional Profile (ECs)", "Functional Profile (ECs)" ], 2),
     :demo_provided => repeat([ "yes", "no", "yes", "no", "yes" ], 2),
 )
 
-CSV.write(tablefiles("Table5.csv"), table5)
+CSV.write(joinpath("manuscript", "assets", "TableS3.csv"), tableS3)
 
-pretty_table(table5;
+pretty_table(tableS3;
     header = [ "Input set", "Age bracket", "Microbiome encoding type", "Demographics Provided? (sex, education)" ],
     backend = Val(:latex)
 )

### 2023-03-25 manuscript/main.tex

@@
 status) on participants from different age groups. Age (in months) was
-provided as a covariate for all models (Table 5).
-
-\textbf{Table 5. Experimental design and input composition for Random
-Forest experiments}
-
-\begin{table}
-    \begin{tabular}{|r|r|r|r|}
-      \hline\hline
-      \textbf{Input set} & \textbf{Age bracket} & \textbf{Microbiome encoding type} & \textbf{Demographics Provided? (sex, education)} \\\hline
-      1 & \multirow{5}{*}{0 to 6 months} & Not provided & yes \\ \cline{3-4}
-      2 &  & \multirow{2}{*}{Taxonomic profile} & no \\ \cline{4-4}
-      3 &  &  & yes \\ \cline{3-4}
-      4 &  & \multirow{2}{*}{Functional Profile (ECs)} & no \\ \cline{4-4}
-      5 &  &  & yes \\ \cline{2-4}
-      6 & \multirow{5}{*}{18 to 120 months} & Not provided & yes \\ \cline{3-4}
-      7 &  & \multirow{2}{*}{Taxonomic profile} & no \\ \cline{4-4}
-      8 &  &  & yes \\ \cline{3-4}
-      9 &  & \multirow{2}{*}{Functional Profile (ECs)} & no \\ \cline{4-4}
-      10 &  &  & yes \\\hline\hline
-    \end{tabular}
-\end{table}
+provided as a covariate for all models (Table S3).
 

### 2023-03-27 src/prediction.jl
Cleanup/update struct block of source prediction.jl

@@ -2,78 +2,20 @@
 # 0. Hierarchical Structures and Types #
 ########################################
 
-struct CustomRangeNormalizer
-    rmin::Float64
-    rmax::Float64
-    tmin::Float64
-    tmax::Float64
-end
-
 abstract type Prediction end
 struct Classification <: Prediction end
 struct Regression <: Prediction end
 
-abstract type Predictor end
-abstract type UnivariatePredictor <: Resonance.Predictor end
-abstract type MultivariatePredictor <: Resonance.Predictor end
-
-mutable struct ExpandedRandomForestClassifier
-    model::Machine
-    train_accuracy::Float64
-    test_accuracy::Float64
-    fitness::Float64
-end
-
-mutable struct UnivariateRandomForestClassifier <: Resonance.UnivariatePredictor
-    name::String
-    original_data::DataFrame
-    inputs_outputs::Tuple{DataFrame, CategoricalArrays.CategoricalVector{Bool, UInt32, Bool, CategoricalArrays.CategoricalValue{Bool, UInt32}, Union{}}}
-    n_splits::Int64
-    dataset_partitions::Vector{Tuple{Vector{Int64}, Vector{Int64}}}
-    models:: Vector{Vector{ExpandedRandomForestClassifier}}
-    selected_split::Tuple{Float64, Int64}
-    train_accuracies::Vector{Float64}
-    test_accuracies::Vector{Float64}
-end
-
-mutable struct UnivariateRandomForestRegressor <: Resonance.UnivariatePredictor
-    name::String
-    original_data::DataFrame
-    inputs_outputs::Tuple{DataFrame, Vector{Float64}}
-    n_splits::Int64
-    dataset_partitions::Vector{Tuple{Vector{Int64}, Vector{Int64}}}
-    models::Vector{Machine}
-    scale_corrections::Vector{CustomRangeNormalizer}
-    selected_split::Tuple{Float64, Int64}
-    train_rmses::Vector{Float64}
-    test_rmses::Vector{Float64}
-    train_cors::Vector{Float64}
-    test_cors::Vector{Float64}
-end
-
-mutable struct ExpandedRandomForestRegressor <: Resonance.UnivariatePredictor
-    model::Machine
-    scale_correction::CustomRangeNormalizer
-    train_rmse::Float64
-    test_rmse::Float64
-    train_cor::Float64
-    test_cor::Float64
-    fitness::Float64
-end
-
-mutable struct UnivariatePredictorEnsemble
[…]

### 2023-03-27 src/prediction.jl
Cleanup/update processing block of source prediction.jl

@@ -22,177 +22,45 @@ end
 # 1. Preprocessing/Wrangling Functions #
 ########################################
 
-## 1.1. Helper/utility functions
-dropnan(vv) = vv[.!(isnan.(vv))]
+function filter_prevalences(
+    mdata_profile_df::DataFrame,
 
-## 1.1.1. \xor used to aggregate columns that contain parts of the same original data
-function myxor(a::Float64, b::Float64)
-    a == b && (return a)
-    a > b && (return a)
-    a < b && (return b)
-end
+    label_col::Symbol,
+    metadata_cols::Vector{Symbol},
+    cols_to_delete::Vector{Symbol};
+    lbound = 0.1,
+    ubound = 1.0,
+    education2int = true,
+    sex2int = true
+    )
 
-## 1.2. Raw data processing functions
+    ## Step 1. Remove unnecessary stuff
+    profile_df = select(mdata_profile_df, Not(vcat(label_col, metadata_cols, cols_to_delete)))
 
-### 1.2.1. Age bracket filtering for concurrent prediction
-function filter_age_bracket(df, min_age, max_age)
-    prediction_df = @chain df begin
-        subset(:ageMonths => x -> min_age .<= x .< max_age)    
-    end
-    return prediction_df
-end
+    ## Step 2. Calculate prevalences
+    prevalences = map( x -> mean(x .> 0.0), eachcol(profile_df) )
+    columns_to_keep = (prevalences .>= lbound) .& (prevalences .<= ubound)
 
-### 1.2.2. Build a DataFrame for prediction of future target variables from mdata
-function build_metadata_prediction_df(base_df::DataFrame, inputs::Vector{Symbol}, targets::Vector{Symbol}; keep_demographic = true)
+    ## Step 3. update profile_df
+    filtered_profile_df = profile_df[:, columns_to_keep]
 
-    subjects = unique(base_df.subject)
-    has_stool = findall(.!(ismissing.(base_df.sample)))
-    has_cogscore = findall(.!(ismissing.(base_df.cogScorePercentile)))
-
-    subjects_stool_idx = [ intersect(findall(base_df.subject .== el), has_stool) for el in subjects ]
-    subjects_cogscore_idx = [ intersect(findall(base_df.subject .== el), has_cogscore) for el in subjects ]
-
-    line_combinations = vcat( [ vec(collect(Base.product(subjects_stool_idx[i], subjects_cogscore_idx[i]))) for i in 1:length(subjects) ]... )
[…]

### 2023-03-27 src/prediction.jl
Cleanup/update postprocessing block of source prediction.jl

@@ -36,7 +36,7 @@ function filter_prevalences(
 
     ## Step 1. Remove unnecessary stuff
     profile_df = select(mdata_profile_df, Not(vcat(label_col, metadata_cols, cols_to_delete)))
-
+    
     ## Step 2. Calculate prevalences
     prevalences = map( x -> mean(x .> 0.0), eachcol(profile_df) )
     columns_to_keep = (prevalences .>= lbound) .& (prevalences .<= ubound)
@@ -48,11 +48,11 @@ function filter_prevalences(
     filtered_profile_df = hcat(
         mdata_profile_df[:, vcat(label_col, metadata_cols)],
         filtered_profile_df
-        )
-    
+    )
+
     if education2int
         filtered_profile_df.education = coerce(int.(skipmissing(filtered_profile_df.education), type = Int), OrderedFactor)
-end
+    end
 
     if sex2int
         filtered_profile_df.sex = coerce(int.(skipmissing(filtered_profile_df.sex), type = Int), OrderedFactor)
@@ -144,29 +144,29 @@ function probe_prod_randomforest(
                 train, test = dataset_partitions[fold_idx]
 
                 for rng_idx in 1:n_rngs
-
+    
                     Random.seed!(train_rng, rng_idx-1)
-
-            rf_model = RandomForestRegressor(
+    
+                    rf_model = RandomForestRegressor(
                         max_depth = tuning_grid[hp_idx][1],
                         min_samples_leaf = tuning_grid[hp_idx][2],
                         sampling_fraction = tuning_grid[hp_idx][3],
                         n_subfeatures = tuning_grid[hp_idx][4],
                         n_trees = tuning_grid[hp_idx][5],
                         rng = train_rng
-            )
-
-            rf_machine = machine(rf_model, X[train, :], y[train])
-            MLJ.fit!(rf_machine, verbosity=0)
+                    )
+    
+                    rf_machine = machine(rf_model, X[train, :], y[train])
+                    MLJ.fit!(rf_machine, verbosity=0)
             
-            train_y_hat = MLJ.predict(rf_machine, X[train, :])
+                    train_y_hat = MLJ.predict(rf_machine, X[train, :]) 
                     train_rmse = rmse(train_y_hat, y[train])
                     train_mape = mean(abs.(train_y_hat .- y[train]) ./ y[train])
-            train_cor = Statistics.cor(train_y_hat, y[train])
[…]

### 2023-03-27 manuscript/Figure3-calculations.jl
update header

@@ -1,6 +1,6 @@
-#####
-# Cognitive Assessment Score calculation notebook
-#####
+##########
+# Concurrent cognitive assessment Score regression notebook
+##########
 
 #####
 # 0. Loading libraries

### 2023-03-27 manuscript/Figure3-calculations.jl
Remove block used for LEAP calculation (moved back to LEAP repo)

@@ -22,39 +22,11 @@ ml_rng = StableRNG(0)
 # Loading Data
 #####
 
-## 2. Taxonomic Profiles
+## 1. Taxonomic Profiles
 taxa = Resonance.load(TaxonomicProfiles())
 mdata_taxa_df = sort(Resonance.comm2wide(taxa), [ :subject, :timepoint ]);
 
-open("taxa_present_echo_all.txt", "w") do io
-    for i in [ f.name for f in features(taxa) ]
-        println(io, i)
-    end
-end
-
-open("taxa_present_echo_6mo.txt", "w") do io
-    taxa_6mo = @chain mdata_taxa_df begin
-        subset(:ageMonths => x -> x .< 6.0)
-        select( _ , names( _ )[15:end])
-        select( _ , names( _ )[[ sum(coll) > 0.0 for coll in eachcol(_) ]])
-    end
-    for i in names(taxa_6mo)
-        println(io, i)
-    end
-end
-
-open("taxa_present_echo_18mo.txt", "w") do io
-    taxa_6mo = @chain mdata_taxa_df begin
-        subset(:ageMonths => x -> 18.0 .< x .< 120.0)
-        select( _ , names( _ )[15:end])
-        select( _ , names( _ )[[ sum(coll) > 0.0 for coll in eachcol(_) ]])
-    end
-    for i in names(taxa_6mo)
-        println(io, i)
-    end
-end
-
-## 3. Functional Profiles
+## 2. Functional Profiles
 ecs = Resonance.load(ECProfiles())
 mdata_ecs_df = sort(Resonance.comm2wide(ecs), [ :subject, :timepoint ]);
 

### 2023-03-27 manuscript/Figure3-calculations.jl
Moved filtering function to source

@@ -58,46 +58,6 @@ taxa_prevalence_threshold_upper = 1.00
 ecs_prevalence_threshold_lower = 0.25
 ecs_prevalence_threshold_upper = 0.85
 
-function filter_prevalences(
-    mdata_profile_df::DataFrame,
-
-    label_col::Symbol,
-    metadata_cols::Vector{Symbol},
-    cols_to_delete::Vector{Symbol};
-    lbound = 0.1,
-    ubound = 1.0,
-    education2int = true,
-    sex2int = true
-    )
-
-    ## Step 1. Remove unnecessary stuff
-    profile_df = select(mdata_profile_df, Not(vcat(label_col, metadata_cols, cols_to_delete)))
-    
-    ## Step 2. Calculate prevalences
-    prevalences = map( x -> mean(x .> 0.0), eachcol(profile_df) )
-    columns_to_keep = (prevalences .>= lbound) .& (prevalences .<= ubound)
-
-    ## Step 3. update profile_df
-    filtered_profile_df = profile_df[:, columns_to_keep]
-
-    ## Step 4. append label and metadata to the beginning of the dataFrame
-    filtered_profile_df = hcat(
-        mdata_profile_df[:, vcat(label_col, metadata_cols)],
-        filtered_profile_df
-    )
-
-    if education2int
-        filtered_profile_df.education = coerce(int.(skipmissing(filtered_profile_df.education), type = Int), OrderedFactor)
-    end
-
-    if sex2int
-        filtered_profile_df.sex = coerce(int.(skipmissing(filtered_profile_df.sex), type = Int), OrderedFactor)
-    end
-
-    return (filtered_profile_df)
-
-end
-
 ## 00 to 06 months
 
 filtered_mdata_taxa_df_00to06 = filter_prevalences(

### 2023-03-27 manuscript/Figure3-calculations.jl
added metadata columns to prediction DFs for sample referencing

@@ -63,8 +63,8 @@ ecs_prevalence_threshold_upper = 0.85
 filtered_mdata_taxa_df_00to06 = filter_prevalences(
     dropmissing(mdata_taxa_df[mdata_taxa_df.filter_00to06, :], Not([:race, :read_depth])),
     :cogScore,
-    [:sex, :education, :ageMonths],
-    [:sample, :sample_base, :subject, :timepoint, :race, :date, :read_depth, :filter_00to120, :filter_00to06, :filter_18to120];
+    [:subject, :timepoint, :sex, :education, :ageMonths],
+    [:sample, :sample_base, :race, :date, :read_depth, :filter_00to120, :filter_00to06, :filter_18to120];
     lbound = taxa_prevalence_threshold_lower,
     ubound = taxa_prevalence_threshold_upper
 )
@@ -72,8 +72,8 @@ filtered_mdata_taxa_df_00to06 = filter_prevalences(
 filtered_mdata_ecs_df_00to06 = filter_prevalences(
     dropmissing(mdata_ecs_df[mdata_ecs_df.filter_00to06, :], Not([:race, :read_depth])),
     :cogScore,
-    [:sex, :education, :ageMonths],
-    [:sample, :sample_base, :subject, :timepoint, :race, :date, :read_depth, :filter_00to120, :filter_00to06, :filter_18to120];
+    [:subject, :timepoint, :sex, :education, :ageMonths],
+    [:sample, :sample_base, :race, :date, :read_depth, :filter_00to120, :filter_00to06, :filter_18to120];
     lbound = ecs_prevalence_threshold_lower,
     ubound = ecs_prevalence_threshold_upper
 )
@@ -81,8 +81,8 @@ filtered_mdata_ecs_df_00to06 = filter_prevalences(
 filtered_mdata_taxa_df_18to120 = filter_prevalences(
     dropmissing(mdata_taxa_df[mdata_taxa_df.filter_18to120, :], Not([:race, :read_depth])),
     :cogScore,
-    [:sex, :education, :ageMonths],
-    [:sample, :sample_base, :subject, :timepoint, :race, :date, :read_depth, :filter_00to120, :filter_00to06, :filter_18to120];
+    [:subject, :timepoint, :sex, :education, :ageMonths],
+    [:sample, :sample_base, :race, :date, :read_depth, :filter_00to120, :filter_00to06, :filter_18to120];
     lbound = taxa_prevalence_threshold_lower,
[…]

### 2023-03-27 manuscript/Figure3-calculations.jl
moved RF function to source

@@ -132,159 +132,6 @@ ecs_tuning_space = (
     ntrees_range = [ 500 ]
 )
 
-# 1000 of each for onlydemo, onlytaxa, demoplustaxa
-
-#####
-# Functions
-#####
-
-struct ProbeData
-    name::String
-    n_folds::Int64
-    n_replicas::Int64
-    n_rngs::Int64
-    merits::DataFrame
-    importances::DataFrame
-end
-
-function probe_prod_randomforest(
-    type::Resonance.Regression,
-    name::String,
-    original_df::DataFrame,
-    data_preparation_fun::Function,
-    input_cols,
-    output_col;
-    n_folds = 10,
-    n_replicas = 10,
-    n_rngs = 100,
-    tuning_space = (;
-        maxnodes_range = [ -1 ],
-        nodesize_range = [ 0 ],
-        sampsize_range = [ 0.7 ],
-        mtry_range = [ 5 ],
-        ntrees_range = [ 10 ]
-    ),
-    train_rng=Random.GLOBAL_RNG
-    )
-
-    # ## 1. Subsetting Data and shuffling DataFrame
-    prediction_df = data_preparation_fun(original_df)
-
-    # ## 3. Building hyperparameter tuning grid
-    tuning_grid = vec(collect(Base.product(
-        tuning_space.maxnodes_range,
-        tuning_space.nodesize_range,
-        tuning_space.sampsize_range,
-        tuning_space.mtry_range,
-        tuning_space.ntrees_range
-    )))
-
-    dataset_partitions = Vector{Tuple{Vector{Int64}, Vector{Int64}}}()
-    for fold_idx in 1:n_folds
-        train, test = Resonance.partitionvec(collect(1:nrow(prediction_df)),floor(Int64, nrow(prediction_df)/n_folds),fold_idx, true)
-        push!(dataset_partitions, (train, test))
-    end
-
-    ys = Vector{Vector{Any}}()
-    hyperpar_idxes = Vector{Int64}()
-    replica_idxes = Vector{Int64}()
-    fold_idxes = Vector{Int64}()
-    rng_idxes = Vector{Int64}()
-    train_rmses = Vector{Float64}()
-    test_rmses = Vector{Float64}()
-    train_mapes = Vector{Float64}()
-    test_mapes = Vector{Float64}()
-    train_cors = Vector{Float64}()
-    test_cors = Vector{Float64}()
-    importances = DataFrame(:variable => [])
-
[…]

### 2023-03-27 manuscript/Figure3-calculations.jl
Fixing comments and calls to probe function due to incorporation of metedata columns on prediction DFs

@@ -30,20 +30,17 @@ mdata_taxa_df = sort(Resonance.comm2wide(taxa), [ :subject, :timepoint ]);
 ecs = Resonance.load(ECProfiles())
 mdata_ecs_df = sort(Resonance.comm2wide(ecs), [ :subject, :timepoint ]);
 
+### 2.1. Hashing EC names with fixed-length alphabetic hashes due to length and special characters; storing text files for remapping.
 oldnames = names(mdata_ecs_df)[15:2414]
-
 for oldname in names(mdata_ecs_df)[15:2414]
     rename!(mdata_ecs_df, oldname => randstring(['A':'Z'; 'a':'z'], 12))
 end
-
 newnames = names(mdata_ecs_df)[15:2414]
-
 open("longnames.txt", "w") do io
     for i in oldnames
         println(io, i)
     end
 end
-
 open("hashnames.txt", "w") do io
     for i in newnames
         println(io, i)
@@ -137,13 +134,12 @@ ecs_tuning_space = (
 #####
 
 ## 1. Only SES
-
 regression_currentCogScores_00to06mo_onlydemo = probe_prod_randomforest(
     Resonance.Regression(),
     "regression_currentCogScores_00to06mo_onlydemo",
     filtered_mdata_taxa_df_00to06,
     identity,
-    [2, 3, 4],
+    [4, 5, 6],
     :cogScore;
     n_folds = 3,
     n_replicas = 100,
@@ -152,16 +148,15 @@ regression_currentCogScores_00to06mo_onlydemo = probe_prod_randomforest(
     train_rng=ml_rng
 )
     
-JLD2.@save "models/aregression_currentCogScores_00to06mo_onlydemo.jld" regression_currentCogScores_00to06mo_onlydemo
+JLD2.@save "models/regression_currentCogScores_00to06mo_onlydemo.jld" regression_currentCogScores_00to06mo_onlydemo
 
 ## 2. Only taxonomic profiles
-
 regression_currentCogScores_00to06mo_onlytaxa = probe_prod_randomforest(
     Resonance.Regression(),
     "regression_currentCogScores_00to06mo_onlytaxa",
     filtered_mdata_taxa_df_00to06,
     identity,
-    collect(4:ncol(filtered_mdata_taxa_df_00to06)), # 4 to include age
+    collect(6:ncol(filtered_mdata_taxa_df_00to06)), # 4 to include age
     :cogScore;
     n_folds = 3,
[…]

### 2023-03-27 src/Resonance.jl
update exports to reflect changes to source prediciton.jl

@@ -39,67 +39,15 @@ export  # 0. structs/types
         Prediction,
         Classification,
         Regression,
-        UnivariateRandomForestClassifier,
-        UnivariateRandomForestRegressor,
-        ExpandedRandomForestClassifier,
-        ExpandedRandomForestRegressor,
-        UnivariatePredictorEnsemble,
         ProbeData,
         # 1. preprocessing functions
-        dropmissing,
-        dropnan,
-        myxor,
-        filter_age_bracket,
-        build_metadata_prediction_df,
-        prepare_future_prediction_df,
-        meanclass,
-        compute_tietjenmoore,
-        test_tietjenmoore,
-        univariate_tietjenmoore,
-        try_outliers,
+        filter_prevalences,
         # 2. training functions
         partitionvec,
-        train_randomforest,
-        try_stress_hyperparameters,
-        expand_pretrained_model,
+        probe_prod_randomforest,
         # postprocessing functions
         CustomRangeNormalizer,
-        compute_custom_scale,
-        normalize_number,
-        scale_normalization,
-        predict_proba,
-        predict,
-        report_merits,
-        singlemodel_singlesplit_importance,
-        singlemodel_allsplits_importances,
-        singlemodel_summary_importances,
-        multimodel_individual_summaryimportances,
-        multimodel_aggregate_summaryimportances,
-        singlemodel_binarytopn_importances,
-        multimodel_individual_binarytopns,
-        multimodel_aggregate_binarytopns,
-        descript_inputs,
-        singlemodel_summary_prevalences,
-        singlemodel_summary_abundances,
-        singlemodel_summary_sdevs,
-        multimodel_individual_prevalences,
-        multimodel_individual_abundances,
-        biserial_correlation,
-        feature_split_correlation_analysis,
-        singlemodel_singlesplit_correlations,
-        singlemodel_allsplits_correlations,
-        singlemodel_allsplits_correlationsummary,
-        multimodel_individual_correlations,
-        # plotting functions
-        build_confusion_matrix,
-        average_confusion_matrix,
-        confmatrix2barplot,
-        singlemodel_merit_barplot!,
-        singlemodel_merit_scatterplot!,
-        singlemodel_avgimportance_barplot!,
-        multimodel_avgimportance_barplot!,
-        singlemodel_logistic_regression!,
-        multimodel_logistic_regression!
+        compute_custom_scale
 
 # Percentiles
 export  AgeBracketPercentiles,

### 2023-03-27 manuscript/main.tex
unify formatting of figure ref
one more read-through

@@
 \graphicspath{assets}
+\usepackage{float}
+
 \usepackage{times}
@@
 "microbiome-gut-brain-axis." Both the human brain and the gut microbiome
-have critical developmental windows in the first three years of life,
+have critical developmental windows in the first years of life,
 raising the possibility that their development is co-occurring and
@@
 diversity (taxa R\textsuperscript{2} = 0.0215, and  gene functions
-annotated with UniRef90s - clusters of 90\% similarity
+annotated with UniRef90s clusters of 90\% similarity
 \cite{suzekUniRefComprehensiveNonredundant2007} -
@@
 (also known as \textit{Adlercreutzia equolifaciens} subsp. \textit{celatus}
-and \textit{Adlercreutzia equolifaciens} subsp. \textit{equolifaciens} \cite{maruoAdlercreutziaEquolifaciensGen2008}),
+and \textit{Adlercreutzia equolifaciens} subsp. \textit{equolifaciens} respectively
+\cite{maruoAdlercreutziaEquolifaciensGen2008}),
 and SCFA-producing probiotic
@@
 
-\begin{table}[!h]
+\begin{table}[!htb]
     \begin{center}
@@
 significant when testing the relationship with LMs (Figure
-3A-B, Table S1A-B); aside from except for \emph{Eubacterium ramulus}
+3A-B, Table S1A-B); except for \emph{Eubacterium ramulus}
 all taxa that were significant using LMs belonged to a group
@@
 However, the rank-ordering of LM significance and RF importance
-were had substantive differences. For example, in children under 6 months,
+had substantive differences. For example, in children under 6 months,
 only \emph{B. wexlerae} was significantly associated with cognitive function in LMs, but
@@
       \hline\hline
-      \textbf{Statistic} & \textbf{(MAPE)} & \textbf{R} \\\hline
+      \textbf{Statistic} & \textbf{MAPE} & \textbf{R} \\\hline
       Mean & 0.073 & 0.137 \\
@@
 brain region models (R = 0.288), as compared to the right which
-had a negative mean test-set correlation (R = -0.041)
-indicated that generalizable models were not identified.
+had a negative mean test-set correlation (R = -0.041),
+indicating that generalizable models were not identified.
 Feature importance for models of the left accumbens area were dominated
-by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.4\%
+by three species of \emph{Bacteroides}: \emph{B. vulgatus} (3.4\%
 relative importance), \emph{B. ovatus} (3.7\% relative importance), and
@@
 Among these, the most important variable is \emph{Anaerostipes hadrus}
-\hl{!!importance stats!!} (Figure 4B and C).
+\hl{!!importance stats!!} (Figure 4B-C).
 \emph{B. wexlerae}, which was important in RF models of cognitive function,
@@
 The use of shotgun metagenomic sequencing enabled us to get
-species-level resolution of microbial taxa. A previous study of
-cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing,
+species-level resolution of microbial taxa. There are many examples
+in which different members of the same genus may play opposing roles,
+as with \textit{B. fragilis} and \textit{B. ovatus} mentioned above.
+A previous study of
+cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing
 and showed that genera from the Lachnospiraceae family as well as

### 2023-03-27 src/prediction.jl
Moved figure3.md functions to source

@@ -258,4 +258,155 @@ function scale_normalization(inputs::Vector{Float64}, scale::CustomRangeNormaliz
     return map(x -> normalize_number(x, scale), inputs)
 end
 
-## 3.1. Generate Predictions (method predict)
+## 3.1. Importance analysis
+function calculate_fitness(train_cor::Vector{Float64}, test_cor::Vector{Float64})
+
+    positive_train_cor = map( x -> maximum([x, 0.0]), train_cor)
+    positive_test_cor = map( x -> maximum([x, 0.0]), test_cor)
+
+    return positive_train_cor .* positive_test_cor
+
+end
+
+function weighted_hpimportances(m, hp = 1; change_hashnames=false, hashnamestable::Union{Nothing, DataFrame} = nothing, normalize_importances=true)
+    merits = m.merits
+    importances = m.importances
+    fitnesses = calculate_fitness(merits.Train_Cor, merits.Test_Cor)
+    fitness_subset_idx = findall(merits.Hyperpar_Idx .== hp)
+
+    # Option 1: divide by amount of positive-only fitnesses
+    # mean_importances = map(x -> sum(x .* fitnesses[fitness_subset_idx])/sum(fitnesses[fitness_subset_idx] .> 0.0), collect(eachrow(Matrix(importances[:, fitness_subset_idx .+ 1]))))
+    # Option 2: divide importances by total number of models
+    mean_importances = map(x -> sum(x .* fitnesses[fitness_subset_idx])/length(fitness_subset_idx), collect(eachrow(Matrix(importances[:, fitness_subset_idx .+ 1]))))
+    mean_importances_df = sort( DataFrame(:variable => importances[:,1], :weightedImportance => mean_importances), :weightedImportance; rev=true)
+
+    if normalize_importances
+        mean_importances_df.weightedImportance = mean_importances_df.weightedImportance ./ sum(mean_importances_df.weightedImportance)
+    end
+
+    if change_hashnames
+        newnames = leftjoin(mean_importances_df, hashnamestable, on = :variable => :hashname).longname
+        mean_importances_df.variable = newnames
+    end
+
+    return mean_importances_df
+end
+
[…]

### 2023-03-28 manuscript/main.tex
Edited to match number on Table

@@
 As with LMs, RF models for children under 6 months old
-performed poorly (mean test-set correlation -0.13, mean RMSE 12.70),
+performed poorly (mean test-set correlation -0.12, mean RMSE 12.70),
 but RFs were consistently able to learn the relationship between taxa

### 2023-03-28 manuscript/main.tex
specify that the percentage is of the fitness-weighted importance

@@
 all taxa that were significant using LMs belonged to a group
-responsible for 80\% of the total relative importance in RF models
+responsible for 80\% of the total fitness-weighted relative importance in RF models
 (Table 3, Figure 3A).

### 2023-03-28 manuscript/main.tex
added "weighted"

@@
 These included the L/R lingual gyrus (mean RF correlations, Left =
-0.421, Right = 0.434; relative age importances, Left = 7.5\%, Right =
+0.421, Right = 0.434; relative weighted importance for age, Left = 7.5\%, Right =
 8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =

### 2023-03-28 manuscript/main.tex
added "weighted" and corrected numbers

--- 2023-03-28 manuscript/main.tex
@@
 These included the L/R lingual gyrus (mean RF correlations, Left =
-0.421, Right = 0.434; relative age importances, Left = 7.5\%, Right =
+0.421, Right = 0.434; relative weighted importance for age, Left = 7.5\%, Right =
 8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
-0.200, Right = 0.273; relative age importances, Left = 3.7\%, Right =
-6.3\%). In many cases, however, age was not an important variable in
+0.200, Right = 0.273; relative weighted importance for age, Left = 3.6\%, Right = 6.3\%).
+In many cases, however, age was not an important variable in
 high-performing models, such as that for the left accumbens area (mean

### 2023-03-28 manuscript/main.tex
added "weighted" and corrected numbers
Changed text to reflect the proportion before filtering by improtant taxa

@@
 Feature importance for models of the left accumbens area were dominated
-by three species of \emph{Bacteroides}: \emph{B. vulgatus} (3.4\%
-relative importance), \emph{B. ovatus} (3.7\% relative importance), and
-\emph{B. uniformis} (3.0\% relative importance).
+by three species of \emph{Bacteroides}: \emph{B. vulgatus} (3.3\%
+relative weighted importance), \emph{B. ovatus} (3.7\% relative weighted importance), and
+\emph{B. uniformis} (2.9\% relative weighted importance).
 
@@
 were also identified in LMs and RF models of cognitive function (Figures 2 and 3). 
-In general, for all segments, a
-consistent number of species between 14 and 22 (median = 20) was
-responsible for one third of the fitness-weighted cumulative importance,
+In general, for all segments, a consistent number of species between 31 and 45 (median = 37, less than 30\% of the input features)
+was responsible for half of the fitness-weighted cumulative importance,
 regardless of model performance (Figure 4A-B, Table S2).

### 2023-03-28 manuscript/main.tex
changed text to reflect not filtering by important taxa

@@
 torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
-\emph{Blautia wexlerae} - that combined, account for approximately one
-third of the cumulative relative importances, computed after subsetting
-on the taxa of interest (Figure 4C).
+\emph{Blautia wexlerae} - that combined, account for approximately 10\% of the cumulative relative importances,
+computed after fitness-weighting on the reported brain segments (Figure 4A-B).
 

### 2023-03-28 manuscript/main.tex
added stats for A.hadrus
corrected pattern two definition

@@
 
-Among these, the most important variable is \emph{Anaerostipes hadrus}
-\hl{!!importance stats!!} (Figure 4B-C).
+Among these, \emph{Anaerostipes hadrus} is the variable with 
+highest average relative importance along all reported 
+(mean = 1.7\%; min = 1.0\%; max = 3.8\%, Figure 4B-C).
 \emph{B. wexlerae}, which was important in RF models of cognitive function,
@@
 equol-producing species, are examples of the second contribution
-pattern, where a model is dominated by a small number of taxa.
+pattern, where a taxon only contributes significantly to a very small number of segments. 
 Each of these species were highly important in predicting the relative volume of

### 2023-03-28 manuscript/main.tex
added "weighted" and corrected numebrs

@@
 Each of these species were highly important in predicting the relative volume of
-the right anterior cingulate (respectively, 3.0\% and 2.6\% relative
+the right anterior cingulate (respectively, 3.0\% and 2.6\% relative weighted
 importances). Another example of this pattern is \emph{R. gnavus}, which was the only
 species with significant negative association with cognitive function in linear models.
-It is heavily associated with the left pars opercularis (2.7\% relative importance). 
+It is heavily associated with the left pars opercularis (2.7\% relative weighted importance). 
 Finally, \emph{Coprococcus comes} displays an importance
@@
 clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (4.0\% relative importance);
+is the highest for a microbe in RF models (4.0\% relative weighted importance);
 only age had higher relative importance in any model. At the same time,
 \textit{C. comes} is one of the most important predictors for neighboring areas of the
-posterior cingulate such as the pars opercularis (relative importances,
+posterior cingulate such as the pars opercularis (relative weighted importances,
 Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
-precentral (2.3\% relative importance) and paracentral lobes (relative
-importances, Left = 2.6\%, Right = 1.8\%).
+precentral (2.3\% relative weighted importance) and paracentral lobes (relative weighted
+importances, Left = 2.6\%, Right = 1.9\%).
 

### 2023-03-28 manuscript/main.tex
incorporate Vanja feedback

--- 2023-03-28 manuscript/main.tex
@@
 
-\title{Gut microbes and their genes are associated with cognition and neuroanatomy in children}
+\title{Gut-resident microbial species and their genes are associated with cognition and neuroanatomy in children}
 
@@
 differences in gut microbial taxa and gene functions are associated with
-overall cognitive function as well as with differences in the size of
+overall cognitive function and with differences in the size of
 multiple brain regions.
-Using a combination of multivariate linear
-models and machine learning (ML) models, we showed that many species,
+Using a combination of multivariate linear and machine learning (ML) models,
+we showed that many species,
 including \emph{Gordonibacter pamelae} and \emph{Blautia wexlerae}, were
@@
 In addition, ML models were able to use microbial taxa to predict the
-volume of brain regions, with particular taxa, many that were identified
-as important in predicting cognitive function, dominating the
+volume of brain regions, and many taxa that were identified
+as important in predicting cognitive function also dominated the
 feature importance metric for individual brain regions. For example, \emph{B.
-wexlerae} was dominant for the predicting the size of the parahippocampal region
+wexlerae} was the most important species in models predicting the size of the parahippocampal region
 in both the left and right hemispheres, while several species from
@@
 
-We investigated the co-development of the brain and the microbiome in
+We investigated the co-development of the brain and microbiome in
 361 healthy and neurotypically-developing children (163 female) between 82 days and 10 years of age
@@
 As expected, the greatest differences in microbial taxa were observed across
-child age (Figure 1C-D, PCoA axis 1), while older children were primarily
+child age (Figure 1C-D, PCoA axis 1), with older children primarily
 stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
@@
       \hline\hline
-      \textbf{group} & \textbf{subgroup} & \textbf{all} & \textbf{under6mo} & \textbf{over18mo} \\\hline
+      \textbf{Group} & \textbf{Subgroup} & \textbf{All} & \textbf{Under 6 months} & \textbf{Over 18 months} \\\hline
       N subjects &  & 361 & 74 & 257 \\  \hline
@@
 
-As several prior studies have demonstrated links between specific gut taxa
-with ASD \cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021},
-adult cognitive skill (e.g., executive functioning)
-\cite{magnussonRelationshipsDietrelatedChanges2015},
-and adult mental health (both anxiety and depressive symptioms)
-\cite{mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022},
-we sought to determine if specific taxa or gene functions were linked
+As several prior studies have demonstrated associations between specific gut taxa
+and atypical neurocognition and neurological disorders
+\cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021,
+      magnussonRelationshipsDietrelatedChanges2015,mayneris-perxachsMicrobiotaAlterationsProline2022,
+      needhamGutderivedMetaboliteAlters2022},
+we sought to determine if specific taxa or gene functions were associated
 with normal cognitive development in children.
@@
 and the relatively wide range of ages when over which infant transition from milk to solid foods,
-we considered ages that are generally pre-transition (prior to 6 months of age old)
-and those that are generally post-transition (over 18 months of age) separately.
+we separately considered ages that are generally pre-transition (prior to 6 months of age old)
+and those that are generally post-transition (over 18 months of age).
 However, even with this categorization, we find increasing variation
@@
 We also found that overall variation in microbial species in children over 18 months old
-was associated with a small but significant variation in cognitive function score
+was significantly associated with variation in cognitive function score
 (Figure 1G, R\textsuperscript{2} = 0.0124, q \textless{} 0.001),
@@
 captured overlapping variation, with species profiles and gene function
-profiles, both generated from DNA sequencing data, being tightly coupled
+profiles, both generated from metagenomic sequences, being tightly coupled
 (Figure 1H, p \textless{} 0.001). Other functional
@@
 \cite{ghoshMediterraneanDietIntervention2020}
-(Figure 2B).
+(Figure 2B). \textit{Ruminococcus gnavus} was the only microbial species
+that was significantly negatively associated with cognitive function score (q value = 0.18).
 
@@
 glutamate and tryptophan were significantly enriched in children with
-higher cognitive function scores (Glutamate synthesis I, E.S. = 0.242, 
-q = 0.047; Tryptophan synthesis, E.S. = 0.119, q = 0.041).
+higher cognitive function scores (glutamate synthesis I, E.S. = 0.242, 
+q = 0.047; tryptophan synthesis, E.S. = 0.119, q = 0.041).
 Taken together, these results suggest that
@@
       \hline\hline
-      \textbf{Age group} & \multicolumn{2}{c}{\textbf{00to120}} & \multicolumn{2}{c}{\textbf{00to06}} & \multicolumn{2}{c}{\textbf{18to120}} \\ \hline
-      \texttt{geneset} & \texttt{E.S.} & \texttt{q value} & \texttt{E.S.} & \texttt{q value} & \texttt{E.S.} & \texttt{q value} \\\hline
+      \textbf{Age group} & \multicolumn{2}{c}{\textbf{0 to 120 months}} & \multicolumn{2}{c}{\textbf{0 to 6 months}} & \multicolumn{2}{c}{\textbf{18 to 120 months}} \\ \hline
+      geneset & E.S. & q value & E.S. & q value & E.S. & q value \\\hline
       Acetate syn. I & -0.122 & 0.555 & -0.194 & \color{red}{\textbf{0.153}} & 0.127 & 0.602 \\
@@
 different scenarios \cite{ardekaniPredictionIncipientAlzheimer2017,velazquezRandomForestModel2021}.
-Additionally, given the sequential
-nature of variable consideration in each tree, RFs are naturally able to
-work out complex input feature interactions, such as those present in a
-microbiome-wide study, without the necessity to explicitly compute
-interaction terms. Given that gut microbial profiles, as well as neurocognitive
+Given that gut microbial profiles, as well as neurocognitive
 development, may partially reflect socioeconomic and demographic
@@
     \caption{\label{tab:rfbench}Benchmark metrics for the cognitive assessment score
-    prediction models. Confidence intervals are calculated from the
-    distribution of metrics from repeated CV at a confidence level of 95.}
+    prediction models. Confidence intervals (C.I.) are calculated from the
+    distribution of metrics from repeated CV at a confidence level of 95.
+    Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
+    Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.}
     \end{center}
@@
 As with LMs, RF models for children under 6 months old
-performed poorly (mean test-set correlation -0.12, mean RMSE 12.70),
+performed poorly (mean test-set correlation -0.13, mean RMSE 12.70),
 but RFs were consistently able to learn the relationship between taxa
@@
 all taxa that were significant using LMs belonged to a group
-responsible for 80\% of the total fitness-weighted relative importance in RF models
+responsible for 80\% of the total relative importance in RF models
 (Table 3, Figure 3A).
@@
 function, they might be reflected in changes in neuroanatomy. We again
-employed a RF modeling approach to associate gut taxonomic
+used RF modeling to associate gut taxonomic
 profiles with individual brain regions identified in MRI scans.
@@
 These included the L/R lingual gyrus (mean RF correlations, Left =
-0.421, Right = 0.434; relative weighted importance for age, Left = 7.5\%, Right =
+0.421, Right = 0.434; relative age importances, Left = 7.5\%, Right =
 8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
-0.200, Right = 0.273; relative weighted importance for age, Left = 3.6\%, Right = 6.3\%).
-In many cases, however, age was not an important variable in
+0.200, Right = 0.273; relative age importances, Left = 3.7\%, Right =
+6.3\%). In many cases, however, age was not an important variable in
 high-performing models, such as that for the left accumbens area (mean
@@
 Feature importance for models of the left accumbens area were dominated
-by three species of \emph{Bacteroides}: \emph{B. vulgatus} (3.3\%
-relative weighted importance), \emph{B. ovatus} (3.7\% relative weighted importance), and
-\emph{B. uniformis} (2.9\% relative weighted importance).
+by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.4\%
+relative importance), \emph{B. ovatus} (3.7\% relative importance), and
+\emph{B. uniformis} (3.0\% relative importance).
 
@@
 were also identified in LMs and RF models of cognitive function (Figures 2 and 3). 
-In general, for all segments, a consistent number of species between 31 and 45 (median = 37, less than 30\% of the input features)
-was responsible for half of the fitness-weighted cumulative importance,
+In general, for all segments, a
+consistent number of species between 14 and 22 (median = 20) was
+responsible for one third of the fitness-weighted cumulative importance,
 regardless of model performance (Figure 4A-B, Table S2).
@@
 just one or two brain segments. Notable cases of the first pattern
-included seven species - \emph{Anaerostipes hadrus}, \emph{Bacteroides
+included seven species - \emph{Anaerostipes hadrus}, \emph{B.
 vulgatus}, \emph{Fusicatenibacter saccharivorans}, \emph{Ruminococcus
 torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
-\emph{Blautia wexlerae} - that combined, account for approximately 10\% of the cumulative relative importances,
-computed after fitness-weighting on the reported brain segments (Figure 4A-B).
+\emph{B. wexlerae} - that combined, account for approximately one
+third of the cumulative relative importances, computed after subsetting
+on the taxa of interest (Figure 4C).
 
@@
 
-Among these, the most important variable is \emph{Anaerostipes hadrus}
+Among these, the most important variable is \emph{A. hadrus}
 \hl{!!importance stats!!} (Figure 4B-C).
@@
 It is heavily associated with the left pars opercularis (2.7\% relative importance). 
-Finally, \emph{Coprococcus comes} displays an importance
+Finally, \emph{C.comes} displays an importance
 distribution that splits almost evenly among the two previously reported
@@
 directly investigates microbial species and their genes in relation to
-typical development in young children. Understanding the gut-brain-microbiome axis
+typical development in young, healthy children. Understanding the gut-brain-microbiome axis
 in early life is particularly important, since differences or
 interventions in early life can have outsized and longer-term
-consequences than those at later ages. Further, even in the absence of
+consequences than those at later ages
+due to the dynamic and plastic nature of both the gut microbiome and the brain.
+Further, even in the absence of
 causal impacts of microbial metabolism, identifying risk factors that
@@
 Other bacteria that were associated with both cognitive ability and
-brain structure had also have intriguing connections in previous studies.
+brain structure had also have possible connections to previous studies.
 For example, RF models predicting the size of the left nucleus accumbens
@@
 especially \textit{B. vulgatus}, \textit{B. uniformis}, and \textit{B. ovatus}.
-\textit{B. ovatus} has previously been shown to produce
+\textit{B. ovatus} has been shown to produce
 a number of neuroactive compounds including SCFAs and neurotransmitters
@@
 
-The use of shotgun metagenomic sequencing enabled us to get
-species-level resolution of microbial taxa. There are many examples
-in which different members of the same genus may play opposing roles,
-as with \textit{B. fragilis} and \textit{B. ovatus} mentioned above.
+As is made apparent by the potentially opposing roles of 
+\textit{B. fragilis} and \textit{B. ovatus},
+individual species from the same genus may play different
+or even opposing roles. As such, the use of shotgun metagenomic sequencing
+may be crucial to identifying microbial influences on the brain,
+since it enables species-level resolution of microbial taxa.
 A previous study of
@@
 In addition to improved species-level resolution, shotgun metagenomic
-sequencing also enables gene-functional insight. We showed here that
+sequencing also enables gene-functional insight. We showed that
 genes for the metabolism of SCFAs, both their degradation and synthesis,
@@
 with brain function \cite{dalileRoleShortchainFatty2019}.
-Microbial metabolism of other potentially neuroactive molecules,
-were also associated with cognitive ability. We were particularly interested
-in genes for metabolizing glutamate, which is a
+Microbial metabolism of other potentially neuroactive molecules
+were also associated with cognitive ability. For example, we were
+particularly interested in genes for metabolizing glutamate, which is a
 critical neurotransmitter controlling neuronal excitatory/inhibitory
@@
 relationship between levels of these genes and the gut concentrations of
-the molecules their product enzymes act on. For example, while it might
-be intuitive to reason that increased levels of menaquinone synthesis
-genes is indicative of increased menaquinone, it could be the case that
-menaquinone deficiency selects for microbes that can synthesize it. For
-the same reason, increased propionate degradation genes may
-counterintuitively be indicative of high levels of propionate in the gut
-lumen, since high propionate would select for microbes that can
-metabolize it. For this reason, future studies coupling shotgun
+the molecules their product enzymes act on since the presence or absence
+of metabolites may select for microbes with the ability to degrade
+or produce that metabolite respectively.
+For this reason, future studies coupling shotgun
 metagenomics with stool metabolomics could improve our understanding of

### 2023-03-28 manuscript/main.tex

@@
 Feature importance for models of the left accumbens area were dominated
-by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.4\%
-relative importance), \emph{B. ovatus} (3.7\% relative importance), and
-\emph{B. uniformis} (3.0\% relative importance).
+by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.3\%
+relative weighted importance), \emph{B. ovatus} (3.7\% relative weighted importance), and
+\emph{B. uniformis} (2.9\% relative weighted importance).
 
@@
 torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
-\emph{B. wexlerae} - that combined, account for approximately one
-third of the cumulative relative importances, computed after subsetting
-on the taxa of interest (Figure 4C).
+\emph{B. wexlerae}, which combined, account for approximately 10\% of the cumulative relative importances,
+computed after fitness-weighting on the reported brain segments (Figure 4A-B).
 
@@
 
-Among these, the most important variable is \emph{A. hadrus}
-\hl{!!importance stats!!} (Figure 4B-C).
+Among these, \emph{A. hadrus} is the variable with 
+highest average relative importance along all reported 
+(mean = 1.7\%; min = 1.0\%; max = 3.8\%, Figure 4B-C).
 \emph{B. wexlerae}, which was important in RF models of cognitive function,
@@
 equol-producing species, are examples of the second contribution
-pattern, where a model is dominated by a small number of taxa.
+pattern, where a taxon only contributes significantly to a very small number of segments. 
 Each of these species were highly important in predicting the relative volume of
-the right anterior cingulate (respectively, 3.0\% and 2.6\% relative
+the right anterior cingulate (respectively, 3.0\% and 2.6\% relative weighted
 importances). Another example of this pattern is \emph{R. gnavus}, which was the only
 species with significant negative association with cognitive function in linear models.
-It is heavily associated with the left pars opercularis (2.7\% relative importance). 
-Finally, \emph{C.comes} displays an importance
+It is heavily associated with the left pars opercularis (2.7\% relative weighted importance). 
+Finally, \emph{C. comes} displays an importance
 distribution that splits almost evenly among the two previously reported
 clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (4.0\% relative importance);
+is the highest for a microbe in RF models (4.0\% relative weighted importance);
 only age had higher relative importance in any model. At the same time,
 \textit{C. comes} is one of the most important predictors for neighboring areas of the
-posterior cingulate such as the pars opercularis (relative importances,
+posterior cingulate such as the pars opercularis (relative weighted importances,
 Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
-precentral (2.3\% relative importance) and paracentral lobes (relative
-importances, Left = 2.6\%, Right = 1.8\%).
+precentral (2.3\% relative weighted importance) and paracentral lobes (relative weighted
+importances, Left = 2.6\%, Right = 1.9\%).
 

### 2023-03-28 manuscript/TableS3.jl
Delete rogue Table S3 file

@@ -1,29 +1,0 @@
-#####
-# Table 5. Experimental design and input composition for Random Forest experiments
-#####
-
-using Resonance
-using CategoricalArrays
-using Statistics
-using Chain
-using JLD2
-using MLJ
-using DataFrames.PrettyTables
-
-isdir(tablefiles()) || mkpath(tablefiles())
-
-## Building the Table
-
-tableS3 = DataFrame(
-    :input_set => 1:10,
-    :age_bracket => vcat( repeat([ "0 to 6 months" ], 5), repeat( [ "18 to 120 months" ], 5)),
-    :microbial_feature => repeat([ "Not provided", "Taxonomic profile", "Taxonomic profile", "Functional Profile (ECs)", "Functional Profile (ECs)" ], 2),
-    :demo_provided => repeat([ "yes", "no", "yes", "no", "yes" ], 2),
-)
-
-CSV.write(joinpath("manuscript", "assets", "TableS3.csv"), tableS3)
-
-pretty_table(tableS3;
-    header = [ "Input set", "Age bracket", "Microbiome encoding type", "Demographics Provided? (sex, education)" ],
-    backend = Val(:latex)
-)

### 2023-03-28 manuscript/TableS1.jl
remove some blank lines

@@ -1,7 +1,6 @@
 #####
 # Supplementary Table S1A-B. List of most important taxa for Random Forest models on the (A) 00-06 and (B) 18-120 months age bracket, MICROBIOME ONLY, ordered by relative weighted importance rank
 #####
-
 using Resonance
 using CategoricalArrays
 using Statistics
@@ -13,7 +12,6 @@ using DataFrames.PrettyTables
 isdir(tablefiles()) || mkpath(tablefiles())
 
 ## 1. Loading the model files
-
 RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
 # concurrent cogScore reression from taxonomic profiles
 regression_currentCogScores_00to06mo_onlytaxa = JLD2.load(modelfiles("2023-02-15", "regression_currentCogScores_00to06mo_onlytaxa.jld"))["regression_currentCogScores_00to06mo_onlytaxa"]

### 2023-03-28 manuscript/TableS1.jl
reference manuscript models

@@ -13,9 +13,8 @@ isdir(tablefiles()) || mkpath(tablefiles())
 
 ## 1. Loading the model files
 RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
-# concurrent cogScore reression from taxonomic profiles
-regression_currentCogScores_00to06mo_onlytaxa = JLD2.load(modelfiles("2023-02-15", "regression_currentCogScores_00to06mo_onlytaxa.jld"))["regression_currentCogScores_00to06mo_onlytaxa"]
-regression_currentCogScores_18to120mo_onlytaxa = JLD2.load(modelfiles("2023-02-15", "regression_currentCogScores_18to120mo_onlytaxa.jld"))["regression_currentCogScores_18to120mo_onlytaxa"]
+regression_currentCogScores_00to06mo_onlytaxa = JLD2.load(modelfiles("manuscript", "regression_currentCogScores_00to06mo_onlytaxa.jld"))["regression_currentCogScores_00to06mo_onlytaxa"]
+regression_currentCogScores_18to120mo_onlytaxa = JLD2.load(modelfiles("manuscript", "regression_currentCogScores_18to120mo_onlytaxa.jld"))["regression_currentCogScores_18to120mo_onlytaxa"]
 
 ## 2. Relevant functions
 

### 2023-03-28 manuscript/TableS1.jl
moved functions to src, renamed formatting functions, made small changes to rounding

@@ -16,41 +16,18 @@ RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
 regression_currentCogScores_00to06mo_onlytaxa = JLD2.load(modelfiles("manuscript", "regression_currentCogScores_00to06mo_onlytaxa.jld"))["regression_currentCogScores_00to06mo_onlytaxa"]
 regression_currentCogScores_18to120mo_onlytaxa = JLD2.load(modelfiles("manuscript", "regression_currentCogScores_18to120mo_onlytaxa.jld"))["regression_currentCogScores_18to120mo_onlytaxa"]
 
-## 2. Relevant functions
-
-function calculate_fitness(train_cor::Vector{Float64}, test_cor::Vector{Float64})
-    positive_train_cor = map( x -> maximum([x, 0.0]), train_cor)
-    positive_test_cor = map( x -> maximum([x, 0.0]), test_cor)
-    return positive_train_cor .* positive_test_cor
-end
-
-function weighted_hpimportances(m, hp = 1; change_hashnames=false, hashnamestable::Union{Nothing, DataFrame} = nothing)
-    merits = m.merits
-    importances = m.importances
-    fitnesses = calculate_fitness(merits.Train_Cor, merits.Test_Cor)
-    fitness_subset_idx = findall(merits.Hyperpar_Idx .== hp)
-
-    mean_importances = map(x -> sum(x .* fitnesses[fitness_subset_idx])/sum(fitnesses[fitness_subset_idx] .> 0.0), collect(eachrow(Matrix(importances[:, fitness_subset_idx .+ 1]))))
-    mean_importances_df = sort( DataFrame(:variable => importances[:,1], :weightedImportance => mean_importances), :weightedImportance; rev=true)
-
-    if change_hashnames
-        newnames = leftjoin(mean_importances_df, hashnamestable, on = :variable => :hashname).longname
-        mean_importances_df.variable = newnames
-    end
-
-    return mean_importances_df
-end
-
-function suppl_table(rf_model)
-    rf_model_importances = weighted_hpimportances(rf_model; change_hashnames=false)
+## 2. Defining Suppl Table Functions
+function singlemodel_importances_suppltable(rf_model)
+    rf_model_importances = weighted_hpimportances(rf_model; normalize_importances=false, change_hashnames=false)
     rf_model_importances.relativeWeightedImportance = rf_model_importances.weightedImportance ./ sum(skipmissing(rf_model_importances.weightedImportance))
[…]

### 2023-03-28 manuscript/TableS1.jl
changed function names

@@ -35,17 +35,15 @@ function prettyformat_suppltable!(suptbl)
 end
 
 ## 3. Building the tables
-
-supptblA = suppl_table(regression_currentCogScores_00to06mo_onlytaxa)
-tblformat_suppl_table!(supptblA)
+supptblA = singlemodel_importances_suppltable(regression_currentCogScores_00to06mo_onlytaxa)
+prettyformat_suppltable!(supptblA)
 CSV.write(joinpath("manuscript", "assets", "TableS1A.csv"), supptblA)
 
-supptblB = suppl_table(regression_currentCogScores_18to120mo_onlytaxa)
-tblformat_suppl_table!(supptblB)
+supptblB = singlemodel_importances_suppltable(regression_currentCogScores_18to120mo_onlytaxa)
+prettyformat_suppltable!(supptblB)
 CSV.write(joinpath("manuscript", "assets", "TableS1B.csv"), supptblB)
 
 ## 4. LaTeX outputs
-
 spec_hl = LatexHighlighter((val, i, j) -> j == 2, ["textit"])
 spec_ft = (val, i, j) -> j == 2 ? replace(val, "_"=> " ") : val
 

### 2023-03-28 manuscript/TableS2.jl
removed blank space

@@ -1,7 +1,6 @@
 #####
-# Table 4. Summary statistics for the neuroanatomy prediction benchmarks
+# Supplementary Table 2. Summary statistics for the neuroanatomy prediction benchmarks
 #####
-
 using Resonance
 using CategoricalArrays
 using Statistics
@@ -13,7 +12,6 @@ using DataFrames.PrettyTables
 isdir(tablefiles()) || mkpath(tablefiles())
 
 ## 1. Loading the model files
-
 RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
 # concurrent brain volumes regression from taxonomic profiles
 brain_models = JLD2.load(modelfiles("2023-02-15", "brain_models.jld"))["brain_models"]

### 2023-03-28 manuscript/TableS2.jl
moved long segment vector declaration to separate file

@@ -13,49 +13,10 @@ isdir(tablefiles()) || mkpath(tablefiles())
 
 ## 1. Loading the model files
 RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
-# concurrent brain volumes regression from taxonomic profiles
-brain_models = JLD2.load(modelfiles("2023-02-15", "brain_models.jld"))["brain_models"]
-
-ordered_brain_segments_list = [
-    "left-temporal", "right-temporal",
-    "left-orbitofrontal", "right-orbitofrontal",
-    "left-parietal", "right-parietal",
-    "left-middle-frontal", "right-middle-frontal",
-    "left-anterior-cingulate", "right-anterior-cingulate",
-    "left-lateral-occipital", "right-lateral-occipital",
-    "left-cerebellum-white-matter", "right-cerebellum-white-matter",
-    "left-thalamus-proper", "right-thalamus-proper",
-    "left-caudate", "right-caudate",
-    "left-putamen", "right-putamen",
-    "left-pallidum", "right-pallidum",
-    "left-hippocampus", "right-hippocampus",
-    "left-amygdala", "right-amygdala",
-    "left-accumbens-area", "right-accumbens-area",
-    "left-basal-forebrain", "right-basal-forebrain", 
-    "left-cuneus", "right-cuneus", 
-    "left-entorhinal", "right-entorhinal",
-    "left-fusiform", "right-fusiform",
-    "left-isthmus-cingulate", "right-isthmus-cingulate",
-    "left-lingual", "right-lingual",
-    "left-parahippocampal", "right-parahippocampal",
-    "left-paracentral", "right-paracentral",
-    "left-pars-opercularis", "right-pars-opercularis",
-    "left-pars-orbitalis", "right-pars-orbitalis",
-    "left-pars-triangularis", "right-pars-triangularis",
-    "left-pericalcarine", "right-pericalcarine",
-    "left-postcentral", "right-postcentral",
-    "left-posterior-cingulate", "right-posterior-cingulate",
-    "left-precentral", "right-precentral",
-    "left-precuneus", "right-precuneus",
-    "left-superior-frontal", "right-superior-frontal",
-    "left-supramarginal", "right-supramarginal",
-    "left-insula", "right-insula",
[…]

### 2023-03-28 manuscript/TableS2.jl
added rounding

@@ -19,8 +19,8 @@ include("manuscript/Figure4-definitions.jl")
 ## 2. Calculating the Figures of Merit
 tableS2 = mapreduce( vcat, eachindex(ordered_brain_segments_list)) do i
     DataFrame(:variable => ordered_brain_segments_list[i],
-              :Test_MAPE => mean(brain_models[ordered_brain_segments_list[i]].merits.Test_MAPE), 
-              :Test_Cor => mean(brain_models[ordered_brain_segments_list[i]].merits.Test_Cor))
+              :Test_MAPE => round(mean(brain_models[ordered_brain_segments_list[i]].merits.Test_MAPE); digits = 4), 
+              :Test_Cor => round(mean(brain_models[ordered_brain_segments_list[i]].merits.Test_Cor); digits = 4))
 end
 tableS2.variable = uppercasefirst.(map(x -> replace(x, "-"=>" "), tableS2.variable))
 

### 2023-03-28 manuscript/TableS1.jl
fixed percentage typo

@@ -27,7 +27,7 @@ end
 function prettyformat_suppltable!(suptbl)
     suptbl = @chain suptbl begin
         insertcols!(1, :rank => 1:nrow(suptbl))
-        transform!(:weightedImportance => (x -> string.(round.(x; digits = 5)).*" %") => :weightedImportance, renamecols = false)
+        transform!(:weightedImportance => (x -> string.(round.(x; digits = 5))) => :weightedImportance, renamecols = false)
         transform!(:relativeWeightedImportance => (x -> string.(round.(100*x; digits = 2)).*" %"), renamecols = false)
         transform!(:cumulativeWeightedImportance => (x -> string.(round.(100*x; digits = 2)).*" %"), renamecols = false)
     end    

### 2023-03-28 manuscript/main.tex
Just adding the reference to S2 here.

--- 2023-03-28 manuscript/main.tex
@@
 \graphicspath{assets}
-\usepackage{float}
-
 \usepackage{times}
@@
 "microbiome-gut-brain-axis." Both the human brain and the gut microbiome
-have critical developmental windows in the first years of life,
+have critical developmental windows in the first three years of life,
 raising the possibility that their development is co-occurring and
@@
 diversity (taxa R\textsuperscript{2} = 0.0215, and  gene functions
-annotated with UniRef90s clusters of 90\% similarity
+annotated with UniRef90s - clusters of 90\% similarity
 \cite{suzekUniRefComprehensiveNonredundant2007} -
@@
 (also known as \textit{Adlercreutzia equolifaciens} subsp. \textit{celatus}
-and \textit{Adlercreutzia equolifaciens} subsp. \textit{equolifaciens} respectively
-\cite{maruoAdlercreutziaEquolifaciensGen2008}),
+and \textit{Adlercreutzia equolifaciens} subsp. \textit{equolifaciens} \cite{maruoAdlercreutziaEquolifaciensGen2008}),
 and SCFA-producing probiotic
@@
 
-\begin{table}[!htb]
+\begin{table}[!h]
     \begin{center}
@@
 significant when testing the relationship with LMs (Figure
-3A-B, Table S1A-B); except for \emph{Eubacterium ramulus}
+3A-B, Table S1A-B); aside from except for \emph{Eubacterium ramulus}
 all taxa that were significant using LMs belonged to a group
-responsible for 80\% of the total fitness-weighted relative importance in RF models
+responsible for 80\% of the total relative importance in RF models
 (Table 3, Figure 3A).
 However, the rank-ordering of LM significance and RF importance
-had substantive differences. For example, in children under 6 months,
+were had substantive differences. For example, in children under 6 months,
 only \emph{B. wexlerae} was significantly associated with cognitive function in LMs, but
@@
       \hline\hline
-      \textbf{Statistic} & \textbf{MAPE} & \textbf{R} \\\hline
+      \textbf{Statistic} & \textbf{(MAPE)} & \textbf{R} \\\hline
       Mean & 0.073 & 0.137 \\
@@
     Mean absolute proportional error (MAPE) and correlation coeficient (R)
-    of test-set preditions are reported accross models for all brain regions.}
+    of test-set preditions are reported accross models for all brain regions.
+    Complete list of values can be found on Table S2.}
     \end{center}
@@
 brain region models (R = 0.288), as compared to the right which
-had a negative mean test-set correlation (R = -0.041),
-indicating that generalizable models were not identified.
+had a negative mean test-set correlation (R = -0.041)
+indicated that generalizable models were not identified.
 Feature importance for models of the left accumbens area were dominated
-by three species of \emph{Bacteroides}: \emph{B. vulgatus} (3.3\%
-relative weighted importance), \emph{B. ovatus} (3.7\% relative weighted importance), and
-\emph{B. uniformis} (2.9\% relative weighted importance).
+by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.4\%
+relative importance), \emph{B. ovatus} (3.7\% relative importance), and
+\emph{B. uniformis} (3.0\% relative importance).
 
@@
 torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
-\emph{Blautia wexlerae} - that combined, account for approximately 10\% of the cumulative relative importances,
-computed after fitness-weighting on the reported brain segments (Figure 4A-B).
+\emph{Blautia wexlerae} - that combined, account for approximately one
+third of the cumulative relative importances, computed after subsetting
+on the taxa of interest (Figure 4C).
 
@@
 
-Among these, \emph{Anaerostipes hadrus} is the variable with 
-highest average relative importance along all reported 
-(mean = 1.7\%; min = 1.0\%; max = 3.8\%, Figure 4B-C).
+Among these, the most important variable is \emph{Anaerostipes hadrus}
+\hl{!!importance stats!!} (Figure 4B and C).
 \emph{B. wexlerae}, which was important in RF models of cognitive function,
@@
 equol-producing species, are examples of the second contribution
-pattern, where a taxon only contributes significantly to a very small number of segments. 
+pattern, where a model is dominated by a small number of taxa.
 Each of these species were highly important in predicting the relative volume of
-the right anterior cingulate (respectively, 3.0\% and 2.6\% relative weighted
+the right anterior cingulate (respectively, 3.0\% and 2.6\% relative
 importances). Another example of this pattern is \emph{R. gnavus}, which was the only
 species with significant negative association with cognitive function in linear models.
-It is heavily associated with the left pars opercularis (2.7\% relative weighted importance). 
+It is heavily associated with the left pars opercularis (2.7\% relative importance). 
 Finally, \emph{Coprococcus comes} displays an importance
@@
 clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (4.0\% relative weighted importance);
+is the highest for a microbe in RF models (4.0\% relative importance);
 only age had higher relative importance in any model. At the same time,
 \textit{C. comes} is one of the most important predictors for neighboring areas of the
-posterior cingulate such as the pars opercularis (relative weighted importances,
+posterior cingulate such as the pars opercularis (relative importances,
 Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
-precentral (2.3\% relative weighted importance) and paracentral lobes (relative weighted
-importances, Left = 2.6\%, Right = 1.9\%).
+precentral (2.3\% relative importance) and paracentral lobes (relative
+importances, Left = 2.6\%, Right = 1.8\%).
 
@@
 The use of shotgun metagenomic sequencing enabled us to get
-species-level resolution of microbial taxa. There are many examples
-in which different members of the same genus may play opposing roles,
-as with \textit{B. fragilis} and \textit{B. ovatus} mentioned above.
-A previous study of
-cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing
+species-level resolution of microbial taxa. A previous study of
+cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing,
 and showed that genera from the Lachnospiraceae family as well as

### 2023-03-28 manuscript/main.tex

--- 2023-03-28 manuscript/main.tex
@@
 all taxa that were significant using LMs belonged to a group
-responsible for 80\% of the total fitness-weighted relative importance in RF models
+responsible for 80\% of the total importance in RF models
 (Table 3, Figure 3A).
@@
 These included the L/R lingual gyrus (mean RF correlations, Left =
-0.421, Right = 0.434; relative weighted importance for age, Left = 7.5\%, Right =
+0.421, Right = 0.434; importance for age, Left = 7.5\%, Right =
 8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
-0.200, Right = 0.273; relative weighted importance for age, Left = 3.6\%, Right = 6.3\%).
+0.200, Right = 0.273; importance for age, Left = 3.6\%, Right = 6.3\%).
 In many cases, however, age was not an important variable in
@@
 Feature importance for models of the left accumbens area were dominated
-by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.3\%
-relative weighted importance), \emph{B. ovatus} (3.7\% relative weighted importance), and
-\emph{B. uniformis} (2.9\% relative weighted importance).
+by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.3\% importance),
+\emph{B. ovatus} (3.7\% importance), and
+\emph{B. uniformis} (2.9\% importance).
 
@@
 In general, for all segments, a consistent number of species between 31 and 45 (median = 37, less than 30\% of the input features)
-was responsible for half of the fitness-weighted cumulative importance,
+was responsible for half of the cumulative importance,
 regardless of model performance (Figure 4A-B, Table S2).
@@
 Each of these species were highly important in predicting the relative volume of
-the right anterior cingulate (respectively, 3.0\% and 2.6\% relative weighted
-importances). Another example of this pattern is \emph{R. gnavus}, which was the only
+the right anterior cingulate (respectively, 3.0\% and 2.6\% importances).
+Another example of this pattern is \emph{R. gnavus}, which was the only
 species with significant negative association with cognitive function in linear models.
-It is heavily associated with the left pars opercularis (2.7\% relative weighted importance). 
-Finally, \emph{C. comes} displays an importance
+It is heavily associated with the left pars opercularis (2.7\% importance). 
+Finally, \emph{C. comes} has an importance
 distribution that splits almost evenly among the two previously reported
 clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (4.0\% relative weighted importance);
+is the highest for a microbe in RF models (4.0\% importance);
 only age had higher relative importance in any model. At the same time,
 \textit{C. comes} is one of the most important predictors for neighboring areas of the
-posterior cingulate such as the pars opercularis (relative weighted importances,
+posterior cingulate such as the pars opercularis (importances,
 Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
-precentral (2.3\% relative weighted importance) and paracentral lobes (relative weighted
-importances, Left = 2.6\%, Right = 1.9\%).
+precentral (2.3\% importance) and paracentral lobes (importances, Left = 2.6\%, Right = 1.9\%).
 
@@
 (\textbf{Equation 1}) was designed to weight the importances from each
-trained forest. The objective was to give more weight to those with
+trained forest, and all "importance" values reported refer to 
+the average fitness-weighted importance accross all models.
+The objective was to give more weight to those with
 higher benchmarks on the validation sets (or higher generalizability),
@@
 inspiration from the approach used on another work employing repeated CV
-on Random Forests with high-dimensional, low sample size microbiome
+on Random Forests with high-dimensional, low sample size
 datasets \cite{woodruffInflammationAutoreactivityDefine2022}.
 The resulting fitness-weighted importances were used
-to generate the values in Figure 3.
+to generate the values in Figures 3-4.
 

### 2023-03-29 manuscript/main.tex
fix colors in table 2

@@
 
-\begin{sidewaysfigure}[!htb]
+\begin{sidewaysfigure}[!h]
     \centering
@@
 
-\begin{table}[!htb]
+\begin{table}[!h]
     \centering
@@
       geneset & E.S. & q value & E.S. & q value & E.S. & q value \\\hline
-      Acetate syn. I & -0.122 & 0.555 & -0.194 & \color{red}{\textbf{0.153}} & 0.127 & 0.602 \\
-      Acetate syn. II & -0.22 & \color{red}{\textbf{0.066}} & -0.342 & \color{red}{\textbf{0.02}} & -0.095 & 0.92 \\
-      Acetate syn. III & -0.21 & \color{red}{\textbf{0.086}} & -0.31 & \color{red}{\textbf{0.052}} & -0.086 & 0.92 \\
-      Glutamate deg. I & -0.404 & \color{red}{\textbf{0.18}} & -0.284 & 0.816 & -0.249 & 0.602 \\
-      Glutamate syn. II & 0.095 & 0.602 & -0.134 & 0.98 & 0.242 & \color{blue}{\textbf{0.047}} \\
-      Menaquinone syn. I & -0.085 & 0.563 & -0.083 & 0.816 & -0.17 & \color{red}{\textbf{0.182}} \\
-      Propionate deg. I & -0.461 & \color{red}{\textbf{0.13}} & -0.542 & \color{red}{\textbf{0.02}} & -0.674 & \color{red}{\textbf{0.041}} \\
-      Propionate syn. I & -0.334 & 0.555 & -0.32 & 0.571 & -0.676 & \color{red}{\textbf{0.023}} \\
-      Tryptophan syn. & 0.057 & 0.974 & -0.072 & 0.816 & 0.119 & \color{blue}{\textbf{0.041}} \\\hline\hline
+      Acetate syn. I & -0.122 & 0.555 & -0.194 & \color{blue}{\textbf{0.153}} & 0.127 & 0.602 \\
+      Acetate syn. II & -0.22 & \color{blue}{\textbf{0.066}} & -0.342 & \color{blue}{\textbf{0.02}} & -0.095 & 0.92 \\
+      Acetate syn. III & -0.21 & \color{blue}{\textbf{0.086}} & -0.31 & \color{blue}{\textbf{0.052}} & -0.086 & 0.92 \\
+      Glutamate deg. I & -0.404 & \color{blue}{\textbf{0.18}} & -0.284 & 0.816 & -0.249 & 0.602 \\
+      Glutamate syn. II & 0.095 & 0.602 & -0.134 & 0.98 & 0.242 & \color{red}{\textbf{0.047}} \\
+      Menaquinone syn. I & -0.085 & 0.563 & -0.083 & 0.816 & -0.17 & \color{blue}{\textbf{0.182}} \\
+      Propionate deg. I & -0.461 & \color{blue}{\textbf{0.13}} & -0.542 & \color{blue}{\textbf{0.02}} & -0.674 & \color{blue}{\textbf{0.041}} \\
+      Propionate syn. I & -0.334 & 0.555 & -0.32 & 0.571 & -0.676 & \color{blue}{\textbf{0.023}} \\
+      Tryptophan syn. & 0.057 & 0.974 & -0.072 & 0.816 & 0.119 & \color{red}{\textbf{0.041}} \\\hline\hline
     \end{tabular}
-    \caption{\label{tab:fsea}Feature set enrichment analysis on neuroactive microbial gene sets}
+    \caption{\label{tab:fsea}Feature set enrichment analysis on neuroactive microbial gene sets.
+    Red and blue text indicates significant (q value \textless 0.2) genesets with a positive or negative
+    pseudomedian respectively.}
     \end{center}

### 2023-03-29 manuscript/FigureSX1.jl
Initial draft of Pareto plot of relative importances

@@ -1,0 +1,144 @@
+#####
+# Supplementary Figure: Pareto plot for taxa importances from FIgure 3
+#####
+using Resonance
+using CategoricalArrays
+using Statistics
+using Chain
+using JLD2
+using MLJ
+using CairoMakie
+
+isdir(tablefiles()) || mkpath(tablefiles())
+
+## 1. Loading the model files
+RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
+regression_currentCogScores_00to06mo_onlytaxa = JLD2.load(modelfiles("manuscript", "regression_currentCogScores_00to06mo_onlytaxa.jld"))["regression_currentCogScores_00to06mo_onlytaxa"]
+regression_currentCogScores_18to120mo_onlytaxa = JLD2.load(modelfiles("manuscript", "regression_currentCogScores_18to120mo_onlytaxa.jld"))["regression_currentCogScores_18to120mo_onlytaxa"]
+
+## 2. Defining Suppl Table Functions
+function singlemodel_importances_suppltable(rf_model)
+    rf_model_importances = weighted_hpimportances(rf_model; normalize_importances=false, change_hashnames=false)
+    rf_model_importances.relativeWeightedImportance = rf_model_importances.weightedImportance ./ sum(skipmissing(rf_model_importances.weightedImportance))
+    rf_model_importances.cumulativeWeightedImportance = cumsum(rf_model_importances.relativeWeightedImportance)
+    return rf_model_importances
+end
+
+## 3. Building the tables
+supptblA = singlemodel_importances_suppltable(regression_currentCogScores_00to06mo_onlytaxa)
+supptblB = singlemodel_importances_suppltable(regression_currentCogScores_18to120mo_onlytaxa)
+
+## 4. Building the Figure
+figure = Figure(resolution = (1200, 1200))
+barcolor = :lightblue
+curvecolor = :orange
+
+ax1L = Axis(
+    figure[1,1],
+    xlabel = "Feature",
+    xticks = (collect(1:nrow(supptblA)), replace.(supptblA.variable, "_"=>" ")),
+    xticklabelsize=8,
+    xticklabelrotation= pi/4,
+    xticklabelfont="TeX Gyre Heros Makie Italic",
[…]

### 2023-03-29 manuscript/main.tex
updates to acknowledgements

@@
 generation of plots was performed using the julia programming language
+and 3rd party packages
 \cite{bezansonJuliaFreshApproach2017,
@@
     dahua_lin_2023_7695673,
-    douglas_bates_2023_7734970,
-    douglas_bates_2023_7734970,
-    kevin_bonham_2023_7647510}.
-All code for data analysis and figure generation, as well as scripts for automated
+    douglas_bates_2023_7734970}.
+Information for replicating the package environment
+as well as code for data analysis and figure generation, as well as scripts for automated
 download of input files are available on github \cite{kevin_bonham_2023_7647510}.
@@
 VKC, SD, and VD were responsible for funding aqusition.
-KSB, MB, FB, RCL, and JB contributed to data curation.
+KSB, SHM, MB, FB, RCL, and JB contributed to data curation.
 Formal analyses were conducted by KSB, GFB, and MB.

### 2023-03-29 other/anpan.qmd
add anpan for pathways

@@
 for s in (cat has_bwex.txt)
-    fd "$s"_genefamilies /grace/sequencing/processed/mgx/humann/main/ -x rg s__Blautia_wexlerae | sd '^q' "$s\t" >> data/bwexlerae_genefamilies.tsv
+    fd "$s"_genefamilies /grace/sequencing/processed/mgx/humann/main/ -x rg s__Blautia_wexlerae | sd '^' "$s\t" >> data/bwexlerae_genefamilies.tsv
 end
@@
 ```
-
-and then added a header manually.
 
@@
 ```
+
+### Using Pathways
+
+```fish
+echo -e "sample\tfeature\tabundance" > data/bwexlerae_pathabundance.tsv
+for s in (cat has_bwex.txt)
+    fd "$s"_pathabundance /grace/sequencing/processed/mgx/humann/main/ -x rg s__Blautia_wexlerae | sd '^' "$s\t" >> data/bwexlerae_pathabundance.tsv
+end
+
+sd '^.+\t(.+\t.+\t.+)$' '$1' data/bwexlerae_pathabundance.tsv
+sd '^FG\d+_S\d+\t$' '' data/bwexlerae_pathabundance.tsv
+sd ', ' ' ' data/bwexlerae_pathabundance.tsv
+```
+
+```{r}
+library(data.table)
+
+df <- fread("data/bwexlerae_pathabundance.tsv")
+df2 <- dcast(df[, .(feature, sample, abundance)],
+                 feature ~ sample,
+                 value.var = "abundance",
+                 fun.aggregate = sum)
+names(df2)[1] = "# Gene Family"
+fwrite(df2, "data/bwexlerae_pathabundance_wide.tsv")
+```
+
+```{r}
+library(anpan)
+future::plan(future::multicore, workers=16)
+```
+
+```{r}
+bwex_results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
+                        bug_file          = "./data/bwexlerae_pathabundance_wide.tsv",
+                        meta_file         = "./data/manuscript_meta.csv",
+                        out_dir           = "./data/anpan",
+                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
+                        filtering_method  = "kmeans",
+                        model_type        = "fastglm",
+                        covariates        = c("ageMonths", "sex", "education"),
+                        outcome           = "cogScore",
+                        plot_ext          = "pdf",
+                        save_filter_stats = TRUE)
+```
+
+```{r}
+input_path = file.path("data", "anpan", "filter_stats",
+                       "filtered_bwexlerae_pathabundance_wide.tsv.gz")
+
+model_input = fread(input_path)
+
+plot_results(res         = bwex_results,
+             model_input = model_input,
+             covariates  = c("ageMonths", "sex", "education"), 
+             outcome     = "cogScore",
+             bug_name    = "Blautia wexlerae",
+             cluster     = "both",
+             show_trees  = TRUE)
+
+```

### 2023-03-30 other/anpan.qmd

--- 2023-03-22 other/anpan.qmd
@@
 
+## Identifying taxa
+
+```{julia}
+using Resonance
+
+species = Resonance.load(TaxonomicProfiles())
+
+bugs = [
+    "Blautia_wexlerae",
+    "Ruminococcus_gnavus",
+    "Erysipelatoclostridium_ramosum",
+    "Bifidobacterium_longum"
+]
+
+for bug in bugs
+    ss = filter(s-> species[Taxon(bug, :species), s] > 0, samples(species))
+    open(joinpath("data", "anpan", "has_" * lowercase(replace(bug, r"([A-Z])[a-z]+_([a-z]{3}).+"=> s"\1\2")) * ".txt"), "w") do io
+        println.(io, name.(ss))
+    end
+end
+```
+
+```{julia}
+mdata = Resonance.load(Metadata())
+pathways = Resonance.load_raw_humann(; kind="pathabundance", sample_filter=mdata.sample, stratified = true)
+
+for bug in bugs
+    comm = pathways[findall(f-> !ismissing(taxon(f)) && replace(name(taxon(f)), r".+s__(.+)"=> s"\1") == bug, features(pathways)), :]
+    df = DataFrame(comm)
+    df.features = string.(df.features)
+    rename!(df, "features"=> "# Gene Family")
+    df = df[!, findall(col-> col == "# Gene Family" || any(>(0), df[!, col]), names(df))]
+    CSV.write(joinpath("data", "anpan", bug * "_pathabundance.tsv"))
+end
+```
+
 ## Running
 
-After identifying samples with *B. wexlerae* in our sample set,
-I generated a table with stratified gene families by running
-
-```fish
-echo -e "sample\tfeature\tabundance" > data/bwexlerae_genefamilies.tsv
-for s in (cat has_bwex.txt)
-    fd "$s"_genefamilies /grace/sequencing/processed/mgx/humann/main/ -x rg s__Blautia_wexlerae | sd '^q' "$s\t" >> data/bwexlerae_genefamilies.tsv
-end
-
-sd '^.+\t(.+\t.+\t.+)$' '$1' data/bwexlerae_genefamilies.tsv
-sd '^FG\d+_S\d+\t$' '' data/bwexlerae_genefamilies.tsv
-
-```
-
-and then added a header manually.
+After identifying samples with *B. wexlerae* in our sample set
 
@@
 library(data.table)
-
-df <- fread("data/bwexlerae_genefamilies.tsv")
-df2 <- dcast(df[, .(feature, sample, abundance)],
-                 feature ~ sample,
-                 value.var = "abundance",
-                 fun.aggregate = sum)
-names(df2)[1] = "# Gene Family"
 ```
@@
 
+## Blautia_wexlerae
+
 ```{r}
 bwex_results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
-                        bug_file          = "./data/bwexlerae_genefamilies_wide.tsv",
-                        meta_file         = "./data/manuscript_meta.csv",
-                        out_dir           = "./data/anpan",
+                        bug_file          = "./data/anpan/Blautia_wexlerae_pathabundance.tsv",
+                        meta_file         = "./data/anpan/manuscript_meta.csv",
+                        out_dir           = "./data/anpan/Blautia_wexlerae",
                         # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
@@
 ```{r}
-input_path = file.path("data", "anpan", "filter_stats",
-                       "filtered_bwexlerae_genefamilies_wide.tsv")
+input_path = file.path("data", "anpan", "Blautia_wexlerae", "filter_stats",
+                       "filtered_Blautia_wexlerae_pathabundance.tsv.gz")
 
@@
 ```
+
+## Ruminococcus_gnavus
+
+```{r}
+results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
+                        bug_file          = "./data/anpan/Ruminococcus_gnavus_pathabundance.tsv",
+                        meta_file         = "./data/anpan/manuscript_meta.csv",
+                        out_dir           = "./data/anpan/Ruminococcus_gnavus",
+                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
+                        filtering_method  = "kmeans",
+                        model_type        = "fastglm",
+                        covariates        = c("ageMonths", "sex", "education"),
+                        outcome           = "cogScore",
+                        plot_ext          = "pdf",
+                        save_filter_stats = TRUE)
+
+input_path = file.path("data", "anpan", "Ruminococcus_gnavus", "filter_stats",
+                       "filtered_Ruminococcus_gnavus_pathabundance.tsv.gz")
+
+model_input = fread(input_path)
+
+plot_results(res         = results,
+             model_input = model_input,
+             covariates  = c("ageMonths", "sex", "education"), 
+             outcome     = "cogScore",
+             bug_name    = "R. gnavus",
+             cluster     = "both",
+             show_trees  = TRUE)
+
+```
+
+## Erysipelatoclostridium_ramosum
+
+```{r}
+results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
+                        bug_file          = "./data/anpan/Erysipelatoclostridium_ramosum_pathabundance.tsv",
+                        meta_file         = "./data/anpan/manuscript_meta.csv",
+                        out_dir           = "./data/anpan/Erysipelatoclostridium_ramosum",
+                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
+                        filtering_method  = "kmeans",
+                        model_type        = "fastglm",
+                        covariates        = c("ageMonths", "sex", "education"),
+                        outcome           = "cogScore",
+                        plot_ext          = "pdf",
+                        save_filter_stats = TRUE)
+
+input_path = file.path("data", "anpan", "Erysipelatoclostridium_ramosum", "filter_stats",
+                       "filtered_Erysipelatoclostridium_ramosum_pathabundance.tsv.gz")
+
+model_input = fread(input_path)
+
+plot_results(res         = results,
+             model_input = model_input,
+             covariates  = c("ageMonths", "sex", "education"), 
+             outcome     = "cogScore",
+             bug_name    = "E. ramosum",
+             cluster     = "both",
+             show_trees  = TRUE)
+
+```
+
+## Bifidobacterium_longum
+
+```{r}
+results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
+                        bug_file          = "./data/anpan/Bifidobacterium_longum_pathabundance.tsv",
+                        meta_file         = "./data/anpan/manuscript_meta.csv",
+                        out_dir           = "./data/anpan/Bifidobacterium_longum",
+                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
+                        filtering_method  = "kmeans",
+                        model_type        = "fastglm",
+                        covariates        = c("ageMonths", "sex", "education"),
+                        outcome           = "cogScore",
+                        plot_ext          = "pdf",
+                        save_filter_stats = TRUE)
+
+input_path = file.path("data", "anpan", "Bifidobacterium_longum", "filter_stats",
+                       "filtered_Bifidobacterium_longum_pathabundance.tsv.gz")
+
+model_input = fread(input_path)
+
+plot_results(res         = results,
+             model_input = model_input,
+             covariates  = c("ageMonths", "sex", "education"), 
+             outcome     = "cogScore",
+             bug_name    = "B. longum",
+             cluster     = "both",
+             show_trees  = TRUE)
+
+```

### 2023-03-31 manuscript/main.tex
change title
vanja comments 1

@@
 
-\title{Gut-resident microbial species and their genes are associated with cognition and neuroanatomy in children}
+\title{Gut-resident microorganisms and their genes are associated with cognition and neuroanatomy in children}
 
@@
 \cite{fungInteractionsMicrobiotaImmune2017,kimProbioticSupplementationImproves2021}.
-But information about this "microbiome-gut-brain
+However, information about this "microbiome-gut-brain
 axis" in normal neurocognitive development remains lacking,
@@
 \cite{laueDevelopingMicrobiomeBirth2022}.
-Fetal development is believed to be mostly sterile, but newborns
+Fetal development is believed to occur in a sterile environment, but newborns
 are rapidly seeded at birth through contact with the birth canal (if
@@
 
-To begin to address this need, we investigated trends and associations between
+To begin to address this need, we investigated relationship between
 the gut microbiome and
@@
 Cognitive skills and abilities were measured using age-appropriate
-psychometric assessments of cognitive function s, i.e.,
+psychometric assessments of cognitive function, i.e.,
 the Mullen Scales of Early Learning (MSEL) \cite{mullenMullenScalesEarly1995},
@@
 Image quality was monitored during and following scanning.
-If motion artifacts were observed (e.g., ghosting, blurring, or other signal variations), scans were repeated or removed from analysis.
+If motion artifacts were observed (e.g., ghosting, blurring, or other signal variations),
+scans were repeated or removed from analysis.
 

### 2023-03-31 manuscript/main.tex
vanja comments 2

@@
-%%%%%%%%%%TO FORMAT A PREPRINT FOR A PCI%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \documentclass{article}
@@
 \usepackage{float}
+\floatstyle{plaintop}
+\restylefloat{table}
 
@@
     Red and blue text indicates significant (q value \textless 0.2) genesets with a positive or negative
-    pseudomedian respectively.}
+    pseudomedian, respectively.}
     \end{center}
@@
 significant when testing the relationship with LMs (Figure
-3A-B, Table S1A-B); except for \emph{Eubacterium ramulus}
+3A-B, Table S1A-B); except for \emph{Eubacterium ramulus},
 all taxa that were significant using LMs belonged to a group

### 2023-03-31 manuscript/main.tex
vanja comments 3

@@
 
-\subsubsection*{Gut microbial taxonomic and functional profiles predict cognitive function}
+\subsubsection*{Gut microbial taxonomice profiles predict cognitive function}
 
@@
 We identified several species in the family Eggerthelaceae that were
-associated with cognitive function, including \emph{Gordonibacter
-pamelaeae, Aldercreutzia equolofaciens, Asaccharobacter celatus}
+associated with cognitive function, including \emph{G.pamelaeae, A. equolofaciens, A. celatus}
 (formerly regarded a subspecies of \emph{A. equolofaciens}
@@
 right anterior cingulate, which has been linked to social cognition
-and reward-based decision making\cite{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002},
+and reward-based decision making \cite{appsAnteriorCingulateGyrus2016,boesRightAnteriorCingulate2008,bushDorsalAnteriorCingulate2002},
 and estrogen has previously been shown to have activity 
-in this brain region\cite{xiaoEstrogenAnteriorCingulate2013}.
+in this brain region \cite{xiaoEstrogenAnteriorCingulate2013}.
 \emph{G. pamelaeae} also has a unique metabolism that may impact the brain,
@@
 a number of neuroactive compounds including SCFAs and neurotransmitters
-such as GABA\cite{horvathBacteroidesOvatusColonization2022}.
+such as GABA \cite{horvathBacteroidesOvatusColonization2022}.
 Another \textit{Bacteroides} species, \textit{B. fragilis} was
@@
 \textit{B. fragilis} and \textit{B. ovatus},
-individual species from the same genus may play different
-or even opposing roles. As such, the use of shotgun metagenomic sequencing
+individual species from the same genus may play different roles.
+As such, the use of shotgun metagenomic sequencing
 may be crucial to identifying microbial influences on the brain,
@@
 pronounced in some age groups due to the period when active recruitment for
-the study was occurring.
+the study was occurring
 \cite{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
@@
 were devoid of major risk factors for abnormal or atypical development.
-These included preterm birth (< 37 weeks gestation) low birth weight (<1500g),
+These included preterm birth (\textless 37 weeks gestation) low birth weight (\textless 1500g),
 in utero exposure to alcohol, cigarette or illicit substance exposure;
@@
 including preeclampsia, high blood pressure, or gestational diabetes;
-complicated vaginal or cesarian birth; 5 minute APGAR scores < 8;
+complicated vaginal or cesarian birth; 5 minute APGAR scores \textless 8;
 NICU admission; neurological disorder in the child
@@
 was assessed using the Wechsler Preschool and Primary Scale of Intelligence 4th Edition
-(WPPSI-IV, children 3 to 6)\cite{wechslerWechslerPreschoolPrimary2012},
+(WPPSI-IV, children 3 to 6) \cite{wechslerWechslerPreschoolPrimary2012},
 and the Wechsler Intelligence Scale for Children 5th Edition (WISC-V)
-for children older than 6\cite{wechslerWechslerIntelligenceScale1949}. 
+for children older than 6 \cite{wechslerWechslerIntelligenceScale1949}. 
 Like the ELC, FSIQ is normalized to a mean of 100 and standard deviation of 15.

### 2023-03-31 manuscript/main.tex
vanja comments 4

@@
 
-Stool samples (n=493) were collected by parents in OMR-200 tubes
+Stool samples (n = 493) were collected by parents in OMR-200 tubes
 (OMNIgene GUT, DNA Genotek, Ottawa, Ontario, Canada), stored on ice, and
@@
 Different experiment sets were designed to probe how different
-representations of the gut microbiome (taxonomic profiles, Functional
+representations of the gut microbiome (taxonomic profiles, functional
 profiles encoded as ECs) would behave, with and without the addition of
@@
 implementation, inside the MLJ.jl \cite{blaomMLJJuliaPackage2020}.
-
-Machine Learning framework. Independent RFs were trained for each
+Independent RFs were trained for each
 experiment, using a set of default regression hyperparameters from
@@
 Image matrix and field-of-view were adjusted by infant size to achieve the desired image resolution.
-For children under ~4 years of age, MRI was performed during natural non-sedated sleep using previously described methods\cite{deanPediatricNeuroimagingUsing2014}.
+For children under 4 years of age, MRI was performed during natural non-sedated sleep using previously described methods \cite{deanPediatricNeuroimagingUsing2014}.
 Older children were scanned awake whilst watching a movie or favorite TV show.
@@
 thalamus, caudate, putamen, pallidum, hippocampus, amygdala,
-and nucleus accumbens were calculated using an atlas matching approach\cite{bruchhageLongitudinalBrainCognitive}.
+and nucleus accumbens were calculated using an atlas matching approach \cite{bruchhageLongitudinalBrainCognitive}.
 Individual images were first nonlinearly aligned to MNI-space using
-a multi-step registration approach (ANTs 2.2 toolbox, \cite{avantsInsightToolKitImage2014}).
+a multi-step registration approach (ANTs 2.2 toolbox \cite{avantsInsightToolKitImage2014}).
 The MNI-aligned Harvard-Oxford brain atlas was then aligned and superimposed onto the individual child
 by applying the inverse transformation, and the brain regions of interest delinated (segmented)
-and their volune calculated by summing the number of image voxels within each\cite{jenkinsonFSL2012}.
-
+and their volume calculated by summing the number of image voxels within each \cite{jenkinsonFSL2012}.
 Brain region volumes used in RF models were normalized to total brain volume
@@
 as well as code for data analysis and figure generation, as well as scripts for automated
-download of input files are available on github \cite{kevin_bonham_2023_7647510}.
+download of input files are available on GitHub \cite{kevin_bonham_2023_7647510}.
 
@@
 
-\printbibliography
+\printbibliography[heading=none]
 

### 2023-03-31 manuscript/main.tex
Guilherme comments

@@
         (\textbf{B}). Colors represent whether the species belong to the group of
-        top-important features that account for 80\% of the cumulative
+        top-important features that account for 60\% of the cumulative
         importance on the RF model, if that species was significant (q
@@
 all taxa that were significant using LMs belonged to a group
-responsible for 80\% of the total importance in RF models
+responsible for 60\% of the total importance in RF models
 (Table 3, Figure 3A).

### 2023-03-31 other/anpan.qmd
sdd some other bugs

@@
 
+```{r}
+bwex_results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
+                        bug_file          = "./data/anpan/Blautia_wexlerae_pathabundance.tsv",
+                        meta_file         = "./data/anpan/manuscript_meta.csv",
+                        out_dir           = "./data/anpan/Blautia_wexlerae",
+                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
+                        filtering_method  = "kmeans",
+                        model_type        = "fastglm",
+                        covariates        = c("sex", "education"),
+                        outcome           = "ageMonths",
+                        plot_ext          = "pdf",
+                        save_filter_stats = TRUE)
+
+input_path = file.path("data", "anpan", "Blautia_wexlerae", "filter_stats",
+                       "filtered_Blautia_wexlerae_pathabundance.tsv.gz")
+
+model_input = fread(input_path)
+
+plot_results(res         = bwex_results,
+             model_input = model_input,
+             covariates        = c("sex", "education"),
+            outcome           = "ageMonths",
+             bug_name    = "Blautia wexlerae",
+             cluster     = "both",
+             show_trees  = TRUE)
+
+```
+
 ## Ruminococcus_gnavus

### 2023-03-31 manuscript/main.tex
clarify cohort for MRI
revert removal of 'functional' from section 3

@@
 
-\subsubsection*{Gut microbial taxonomice profiles predict cognitive function}
+\subsubsection*{Gut microbial taxonomic and functional profiles predict cognitive function}
 
@@
 used RF modeling to associate gut taxonomic
-profiles with individual brain regions identified in MRI scans.
+profiles with individual brain regions identified in MRI scans
+(children over 18 months old, N = 121).
 Some brain regions were more readily

### 2023-03-31 other/pcresol.jl
check on some genes (not significant)

@@ -1,0 +1,25 @@
+using Resonance
+using CairoMakie
+
+mdata = Resonance.load(Metadata())
+ecs = Resonance.load(ECProfiles(); timepoint_metadata = mdata)
+relativeabundance!(ecs)
+
+dihyrofolate = [
+    "6.3.2.12",
+    "1.5.1.3"
+]
+
+pcresol = "1.17.9.1"
+
+dhf1 = vec(abundances(ecs[r"6\.3\.2\.12", :]))
+dhf2 = vec(abundances(ecs[r"1\.5\.1\.3:", :]))
+# ecs[r"1\.17\.9\.1", :] # not found
+
+scatter(dhf1, get(ecs, :cogScore); axis = (; xlabel = "abundance dihydrofolate synthetase (6.3.2.12)", ylabel="cogScore"))
+scatter(dhf2, get(ecs, :cogScore); axis = (; xlabel = "abundance dihydrofolate reductase (1.5.1.3)", ylabel="cogScore"))
+
+
+ecdf = comm2wide(ecs[r"(6\.3\.2\.12:|1\.5\.1\.3:)", :])
+
+runlms(ecdf[ecdf.filter_00to120, :], tablefiles("lms_ecs_dihydrofolate.csv"), last(names(ecdf), 2))

### 2023-03-31 manuscript/main.tex
errant 'and'

@@
 The gastrointestinal tract, its resident microorganisms, and the central
-nervous system are connected by biochemical signaling, also known as
+nervous system are connected by biochemical signaling, also known as the
 "microbiome-gut-brain-axis." Both the human brain and the gut microbiome

### 2023-08-09 notes/mbx.jl
load up longform of metabolites

@@ -1,6 +1,7 @@
 using Resonance
 using XLSX
 using ThreadsX
+using Chain
 
 biospecimens = CSV.read("input/Biospecimens-ECHO.csv", DataFrame)
 olddb = CSV.read("input/Samples-olddb.csv", DataFrame)
@@ -73,3 +74,110 @@ subset(mdata, "mbxsample" => ByRow(!ismissing)) |> size
 unique(subset(mdata, "mbxsample" => ByRow(!ismissing)), "subject") |> size
 
 subset(mdata, "mbxsample" => ByRow(!ismissing)).ageMonths |> describe
+
+#-
+
+fixcol(col, n::String)                  = fixcol(col, Val{Symbol(n)}())
+fixcol(col, n::Val)                     = coalesce.(col, 0)
+fixcol(col, n::Val{:MZ})                = convert(Vector{Float64}, col)
+fixcol(col, n::Val{:RT})                = convert(Vector{Float64}, col)
+fixcol(col, n::Val{:Method})            = convert(Vector{String}, col)
+fixcol(col, n::Val{:Compound_ID})       = convert(Vector{String}, col)
+fixcol(col, n::Val{:HMDB_ID})           = convert(Vector{Union{Missing,String}}, col)
+fixcol(col, n::Val{:Metabolite})        = convert(Vector{Union{Missing,String}}, col)
+fixcol(col, n::Val{:HMDB_ID_Certainty}) = map(col) do c
+    c isa Int && return c
+    ismissing(c) && return c
+    c == "NA" && return 0
+end
+
+#-
+
+mbx_xlsx_files = filter(f-> endswith(f, ".xlsx") && !contains(f, r"centric"i), readdir("input/", join=true))
+
+mbx_long = mapreduce(vcat, mbx_xlsx_files) do file
+    xf = XLSX.readxlsx(file)
+    sh = xf[first(XLSX.sheetnames(xf))]
+    mat = sh[:]
+    startrow = findfirst(!ismissing, mat[:,1])
+    endcol = let i = findfirst(ismissing, mat[startrow, :])
+        isnothing(i) ? size(mat, 2) : i - 1
+    end
+
+    df = DataFrame(mat[startrow+1:size(mat, 1), 1:endcol], vec(mat[startrow, 1:endcol]); makeunique=true)
+    rename!(df, first(names(df, r"HMDB_ID_")) => "HMDB_ID_Certainty")
+    for n in names(df)
[…]

### 2023-08-09 manuscript/main.tex
begin R1 response

@@
 This analysis allowed us to establish links between microbial taxa and
-their functional potential with cognition and brain structure. Although
-we cannot test causality or the chemistry behind the interactions
+their functional potential with cognition and brain structure. 
+Although we did not directly test the causal relationships
 between gut microbial taxa and their genes, the gut, and the brain,
 this study provides clear and statistically significant associations
-between the infant and early child gut microbiota and neurocognition.
-Future studies should focus on characterizing the early-life microbiome
+that could serve as targets for future efforts in pre-clinical models.
+For example, mouse models of brain development, learning, and executive function
+could be assessed in germ-free animals with the addition of
+metabolically characterized and manipulated microbes.
+Future studies should also focus on characterizing the early-life microbiome
 and neurocognitive development across different geographic regions
@@
 The authors declare that they have no competing interests.
+We thank Libusha Kelly (Albert Einstein School of Medicine) and Daphne Koinis Mitchell (Brown University)
+for helpful feedback on the manuscript. The research team also thanks the participants and families
+of the participants who generously shared their time and samples for this project.
 

### 2023-08-09 manuscript/main.tex
R1 point 2

@@
 
+While we did not identify any statistical associations between individual species
+and cognitive performance in our youngest subset (0 to 6 months),
+it may be because brain developmental trajectories begin \textit{in utero},
+and microbial influences (that can only begin at birth)
+take additional time to manifest.
+Future studies should attempt to link early-life microbial exposures
+to future cognitive performance or brain development.
+
 We identified several species in the family Eggerthelaceae that were

### 2023-08-14 notes/mbx.jl
plot some metabolomics stuff

@@ -103,21 +103,28 @@ function mbx_xlsx_to_df(file)
         isnothing(i) ? size(mat, 2) : i - 1
     end
 
-    df = DataFrame(mat[startrow+1:size(mat, 1), 1:endcol], vec(mat[startrow, 1:endcol]); makeunique=true)
+    nms = vec(mat[startrow, 1:endcol])
+    dups = filter(x-> count(==(x), nms) > 1, nms)
+    if !isempty(dups)
+        @info file unique(dups)
+        ixs = unique(map(x-> findfirst(==(x), nms), nms))
+    else
+        ixs = collect(eachindex(nms))
+    end
+
+
+    df = DataFrame(mat[startrow+1:size(mat, 1), ixs], vec(mat[startrow, ixs]))
 end
 
-dfs = collect((mbx_xlsx_to_df(file) for file in sort(mbx_xlsx_files)))
+# until we can reconcile batches...
+# dfs = collect((mbx_xlsx_to_df(file) for file in sort(mbx_xlsx_files)))
 
-
-#-
-
-
+dfs = collect((mbx_xlsx_to_df(file) for file in sort(mbx_xlsx_files[5:end])))
 
 #-
 
 
 mbx_long = mapreduce(vcat, dfs) do df
-
     rename!(df, first(names(df, r"HMDB_ID_")) => "HMDB_ID_Certainty")
     for n in names(df)
         df[!, n] = fixcol(df[!, n], n)
@@ -170,8 +177,7 @@ using XLSX
 using ThreadsX
 using Chain
 using Arrow
-
-VKCComputing.set_readonly_pat!()
+using Statistics
 
 base = LocalBase()
 
@@ -216,8 +222,84 @@ end) => ["subject", "timepoint", "biospecimen"])
 #-
 
 newmeta = CSV.read("input/ECHO_metabolomics_preselectedinputs.csv", DataFrame)
+fmp_tp = CSV.read("input/fmp_alltp.csv", DataFrame)
+
+transform!(fmp_tp, AsTable(["ageMonths", "assessmentAgeDays", "assessmentAgeMonths", "scanAgeDays", "scanAgeMonths"]) => ByRow(row-> begin
+        am, aad, aam, sad, sam = row
+        calcage = coalesce(aad / 365 * 12 + aam, sad / 365 * 12 + sam)
+        if !ismissing(am)
[…]

### 2023-08-14 manuscript/main.tex
add IRB numbers

@@
 All procedures for this study were approved by the local institutional
-review board at Rhode Island Hospital, and all experiments adhered to
+review board at Rhode Island Hospital, 
+(Investigating Early Childhood Neurodevelopment 1500991,
+and Prenatal Influences on Fetal and Infant Brain Development 1501119;
+most recent approval for both: 2020-01-15)
+and all experiments adhered to
 the regulation of the review board. Written informed consent was

### 2023-08-26 manuscript/main.tex
add note on serum metabolomics
swap g pam for a obesi in abstract

@@
 we showed that many species,
-including \emph{Gordonibacter pamelae} and \emph{Blautia wexlerae}, were
-significantly associated with higher cognitive function, while some
+including \emph{Alistipes obesi} and \emph{Blautia wexlerae}, were
+associated with higher cognitive function, while some
 species such as \emph{Ruminococcus gnavus} were more commonly found in
@@
 the relationship between microbial metabolism and cognitive development.
+Serum metabolomic analysis may be even more informative,
+since the metabolites measured in stool are not necessarily indicative
+of exposure, but drawing sufficient blood from children,
+especially very young children, is typically not permitted.
 Further, strain-level analysis linking specific gene content in species

### 2023-08-26 manuscript/main.tex
add some todos to abstract

@@
 we showed that many species,
-including \emph{Alistipes obesi} and \emph{Blautia wexlerae}, were
+including \emph{Alistipes obesi} and \emph{Blautia wexlerae}, were % check RF models for direction of B. wex
 associated with higher cognitive function, while some
-species such as \emph{Ruminococcus gnavus} were more commonly found in
+species such as \emph{Ruminococcus gnavus} were more commonly found in % check RF models for direction of R. gnav
 children with low cognitive scores after controlling for sociodemographic factors.
@@
 as important in predicting cognitive function also dominated the
-feature importance metric for individual brain regions. For example, \emph{B.
-wexlerae} was the most important species in models predicting the size of the parahippocampal region
+feature importance metric for individual brain regions.
+For example, \emph{B. wexlerae} % is this still true?
+was the most important species in models predicting the size of the parahippocampal region
 in both the left and right hemispheres, while several species from
-the phylum \textit{Bacteroidetes}, including GABA-producing
+the phylum \textit{Bacteroidetes}, including GABA-producing % is this still true?
 \emph{B. ovatus}, were important for predicting

### 2023-08-26 manuscript/Table1.jl
fix table 1 generation code to account for rows without stool samples

@@ -53,7 +53,7 @@ Table1 = DataFrame(
 Table1.all = let df = mdata; ufilt = df.filter_00to120
     nsub = [length(unique(df.subject))]
     
-    ns = DataFrames.combine(groupby(df, :subject), "timepoint"=> length => "n sample")."n sample"
+    ns = DataFrames.combine(groupby(subset(df, "sample"=> ByRow(!ismissing)), :subject), "timepoint"=> length => "n sample")."n sample"
     ss = [count(==(1), ns), count(==(2), ns), count(>(2), ns)]
     ss = ["$s ($(round(s / length(ns) * 100; digits = 1))%)" for s in ss]
 
@@ -75,7 +75,7 @@ end
 Table1.under6mo = let df = subset(mdata, "ageMonths"=> ByRow(<=(6))); ufilt = df.filter_00to06
     nsub = [length(unique(df.subject))]
     
-    ns = DataFrames.combine(groupby(df, :subject), "timepoint"=> length => "n sample")."n sample"
+    ns = DataFrames.combine(groupby(subset(df, "sample"=> ByRow(!ismissing)), :subject), "timepoint"=> length => "n sample")."n sample"
     ss = [count(==(1), ns), count(==(2), ns), count(>(2), ns)]
     ss = ["$s ($(round(s / length(ns) * 100; digits = 1))%)" for s in ss]
 
@@ -97,7 +97,7 @@ end
 Table1.over18mo = let df = subset(mdata, "ageMonths"=> ByRow(>(18))); ufilt = df.filter_18to120
     nsub = [length(unique(df.subject))]
     
-    ns = DataFrames.combine(groupby(df, :subject), "timepoint"=> length => "n sample")."n sample"
+    ns = DataFrames.combine(groupby(subset(df, "sample"=> ByRow(!ismissing)), :subject), "timepoint"=> length => "n sample")."n sample"
     ss = [count(==(1), ns), count(==(2), ns), count(>(2), ns)]
     ss = ["$s ($(round(s / length(ns) * 100; digits = 1))%)" for s in ss]
 

### 2023-08-26 manuscript/main.tex

@@
       \textbf{Group} & \textbf{Subgroup} & \textbf{All} & \textbf{Under 6 months} & \textbf{Over 18 months} \\\hline
-      N subjects &  & 361 & 74 & 257 \\  \hline
-    \multirow{3}{*}{Samples} & 1 & 262 (72.6\%) & 74 (100.0\%) & 203 (79.0\%) \\ \cline{2-5}
-         & 2 & 72 (19.9\%) & 0 (0.0\%) & 46 (17.9\%) \\ \cline{2-5}
-         & \textgreater 2 & 27 (7.5\%) & 0 (0.0\%) & 8 (3.1\%) \\ \hline
-    \multirow{3}{*}{Age (months)} & min & 2.76 & 2.76 & 18.03 \\ \cline{2-5}
-         & max & 119.3 & 6.0 & 119.3 \\ \cline{2-5}
-         & median & 25.79 & 3.66 & 45.92 \\ \hline
-    \multirow{2}{*}{Sex} & F & 163 (45.2\%) & 37 (50.0\%) & 114 (44.4\%) \\   \cline{2-5}
-                         & M & 198 (54.8\%) & 37 (50.0\%) & 143 (55.6\%) \\  \hline
-    \multirow{6}{*}{Race} & White & 252 (69.8\%) & 42 (56.8\%) & 191 (74.3\%) \\   \cline{2-5}
-        & Black & 40 (11.1\%) & 16 (21.6\%) & 21 (8.2\%) \\ \cline{2-5}
-        & Indiginous & 1 (0.3\%) & 0 (0.0\%) & 1 (0.4\%) \\ \cline{2-5}
-        & Asian & 7 (1.9\%) & 1 (1.4\%) & 6 (2.3\%) \\ \cline{2-5}
-        & Mixed & 48 (13.3\%) & 12 (16.2\%) & 29 (11.3\%) \\ \cline{2-5}
-        & Other & 2 (0.6\%) & 1 (1.4\%) & 2 (0.8\%) \\ \hline
-    \multirow{6}{*}{Maternal Ed.} & Junior high school & 1 (0.3\%) & 0 (0.0\%) & 1 (0.4\%) \\   \cline{2-5}
-        & Some high school & 6 (1.7\%) & 3 (4.1\%) & 4 (1.6\%) \\ \cline{2-5}
-        & High school grad & 37 (10.2\%) & 13 (17.6\%) & 16 (6.2\%) \\ \cline{2-5}
-        & Some college & 95 (26.3\%) & 25 (33.8\%) & 57 (22.2\%) \\ \cline{2-5}
-        & College grad & 90 (24.9\%) & 17 (23.0\%) & 69 (26.8\%) \\ \cline{2-5}
-        & Grad/professional school & 123 (34.1\%) & 15 (20.3\%) & 104 (40.5\%) \\\hline\hline
+      N subjects &  & 381 & 149 & 328 \\  \hline
+    \multirow{3}{*}{Samples} & 1 & 77 (20.2\%) & 125 (83.9\%) & 92 (28.0\%) \\ \cline{2-5}
+         & 2 & 65 (17.1\%) & 20 (13.4\%) & 80 (24.4\%) \\ \cline{2-5}
+         & \textgreater 2 & 239 (62.7\%) & 4 (2.7\%) & 156 (47.6\%) \\ \hline
+    \multirow{3}{*}{Age (months)} & min & 2.89 & 2.89 & 18.03 \\ \cline{2-5}
+         & max & 119.53 & 5.95 & 119.53 \\ \cline{2-5}
+         & median & 25.85 & 3.64 & 41.59 \\ \hline
+    \multirow{2}{*}{Sex} & F & 172 (45.1\%) & 42 (51.2\%) & 122 (44.0\%) \\   \cline{2-5}
+                         & M & 209 (54.9\%) & 40 (48.8\%) & 155 (56.0\%) \\  \hline
+    \multirow{6}{*}{Race} & White & 265 (69.6\%) & 48 (58.5\%) & 201 (72.6\%) \\   \cline{2-5}
+        & Black & 40 (10.5\%) & 15 (18.3\%) & 21 (7.6\%) \\ \cline{2-5}
+        & Indigenous & 1 (0.3\%) & 0 (0.0\%) & 1 (0.4\%) \\ \cline{2-5}
+        & Asian & 8 (2.1\%) & 2 (2.4\%) & 7 (2.5\%) \\ \cline{2-5}
+        & Mixed & 54 (14.2\%) & 14 (17.1\%) & 38 (13.7\%) \\ \cline{2-5}
+        & Other & 3 (0.8\%) & 1 (1.2\%) & 3 (1.1\%) \\ \hline
+    \multirow{6}{*}{Maternal Ed.} & Junior high school & 2 (0.5\%) & 0 (0.0\%) & 2 (0.7\%) \\   \cline{2-5}
+        & Some high school & 6 (1.6\%) & 3 (3.7\%) & 4 (1.4\%) \\ \cline{2-5}
+        & High school grad & 39 (10.2\%) & 13 (15.9\%) & 22 (7.9\%) \\ \cline{2-5}
+        & Some college & 100 (26.2\%) & 28 (34.1\%) & 60 (21.7\%) \\ \cline{2-5}
+        & College grad & 95 (24.9\%) & 19 (23.2\%) & 75 (27.1\%) \\ \cline{2-5}
+        & Grad/professional school & 126 (33.1\%) & 16 (19.5\%) & 106 (38.3\%) \\\hline\hline
     \end{tabular}

### 2023-08-26 manuscript/main.tex
update table 1

@@
 We investigated the co-development of the brain and microbiome in
-361 healthy and neurotypically-developing children (163 female) between 82 days and 10 years of age
+381 healthy and neurotypically-developing children (163 female) between 82 days and 10 years of age
 (Table 1,  Figure 1A, Figure S1)
@@
       N subjects &  & 381 & 149 & 328 \\  \hline
-    \multirow{3}{*}{Samples} & 1 & 77 (20.2\%) & 125 (83.9\%) & 92 (28.0\%) \\ \cline{2-5}
-         & 2 & 65 (17.1\%) & 20 (13.4\%) & 80 (24.4\%) \\ \cline{2-5}
-         & \textgreater 2 & 239 (62.7\%) & 4 (2.7\%) & 156 (47.6\%) \\ \hline
+    \multirow{3}{*}{Samples} & 1 & 214 (56.2\%) & 91 (94.8\%) & 172 (61.9\%) \\ \cline{2-5}
+         & 2 & 90 (23.6\%) & 5 (5.2\%) & 60 (21.6\%) \\ \cline{2-5}
+         & \textgreater 2 & 77 (20.2\%) & 0 (0.0\%) & 46 (16.5\%) \\ \hline
     \multirow{3}{*}{Age (months)} & min & 2.89 & 2.89 & 18.03 \\ \cline{2-5}

### 2023-08-26 manuscript/main.tex
update figure 2 legend

@@
 was significantly associated with variation in cognitive function score
-(Figure 1G, R\textsuperscript{2} = 0.0124, q \textless{} 0.001),
+(Figure 1G, R\textsuperscript{2} = 0.0099, q \textless{} 0.01),
 as was variation in microbial gene functions (R\textsuperscript{2} = 0.0119, q \textless{} 0.001).
@@
 As expected, age was significantly associated with microbial beta
-diversity (taxa R\textsuperscript{2} = 0.0215, and  gene functions
+diversity (taxa R\textsuperscript{2} = 0.0207, and  gene functions
 annotated with UniRef90s clusters of 90\% similarity
 \cite{suzekUniRefComprehensiveNonredundant2007} -
-R\textsuperscript{2} = 0.0245, q \textless{} 0.001), and very strongly
-associated with variation in neuroimaging profiles (R\textsuperscript{2} = 0.256, q
+R\textsuperscript{2} = 0.0263, q \textless{} 0.001), and very strongly
+associated with variation in neuroimaging profiles (R\textsuperscript{2} = 0.258, q
 \textless{} 0.001).
@@
 cohorts, despite being derived from UniRef90 labels. In children over 18
-months, some variation (15.9\%, p \textless{} 0.01) in neuroimaging
+months, some variation (taxa - 20.1\%, p \textless{} 0.01;
+gene functions - 15.9\%, p \textless{} 0.05) in neuroimaging
 overlapped with microbial measures, though this may be due to the
@@
 for the relative abundance of each species that had at least 15\%
-prevalence in a given age group (Figure 2A, N\textsubscript{species} = 92 for 0--120 months, 
-N\textsubscript{species} = 46 for 0--6 months, N\textsubscript{species} = 97 for 18--120 months).
-Only \emph{Blautia wexlerae} was significantly associated (q value = 0.14, $\beta$ = 0.0015) with
+prevalence in a given age group (Figure 2A, N\textsubscript{species} = 116 for 0--120 months, 
+N\textsubscript{species} = 54 for 0--6 months, N\textsubscript{species} = 136 for 18--120 months).
+No species were significantly associated (q \textless{} 0.20) with
 cognitive function in children under 6 months old after adjusting for
-age and maternal education (Figure 2B).
+age and maternal education.
 By contrast, in children over 18 months of age,
-several mirobial species were significantly enriched (q-value
-\textless{} 0.20) in children with higher cognitive function scores,
-including \emph{Gordonibacter pamelaeae},
-\emph{Asaccharobacter celatus} and \emph{Adelcreutzia equolifaciens}
-(also known as \textit{Adlercreutzia equolifaciens} subsp. \textit{celatus}
-and \textit{Adlercreutzia equolifaciens} subsp. \textit{equolifaciens} respectively
+several mirobial species were significantly enriched
+in children with higher cognitive function scores (Figure 2B),
+including \emph{Alistipes obesi},
+\emph{Asaccharobacter celatus} (also known as \textit{Adlercreutzia equolifaciens} subsp. \textit{celatus}
 \cite{maruoAdlercreutziaEquolifaciensGen2008}),
@@
 \cite{ghoshMediterraneanDietIntervention2020}
-(Figure 2B). \textit{Ruminococcus gnavus} was the only microbial species
-that was significantly negatively associated with cognitive function score (q value = 0.18).
+(Figure 2B). \textit{Sutterella wadsworthensis} was the only microbial species
+that was significantly negatively associated with cognitive function score (q value = 0.165).
 
@@
         relationship between individual taxa and cognitive function score in
-        children over 18 months of age, controlling for age, maternal education,
-        and sequencing depth. Taxa that were significant after FDR correction (q
+        children over 18 months of age (left) or across all ages (right),
+        controlling for age, maternal education, and sequencing depth.
+        Taxa that were significant after FDR correction (q
         \textless{} 0.2) are colored red. (\textbf{B}) For taxa that were significantly
-        associated with cognitive function, heatmaps of prevalence, mean
-        relative abundance, and correlation with cognitive function in different
-        age groups. (\textbf{C-F}) Enrichment plots for selected neuroactive gene sets
-        used in feature set enrichment analysis (FSEA). (\textbf{G}) summary of FSEA
-        results across all samples (left) as well as in the under 6 months
-        (middle) and over 18 months (right) subsets, colored based on the
+        associated with cognitive function,
+        heatmaps of correlation with cognitive function in different age groups.
+        (\textbf{C}) Multivariate linear models showing the relationship
+        between individual taxa and domain-specific skill assessments
+        from the MSEL for children under 6m (left), 18m to 48m (right),
+        or over across all ages (right)
+        controlling for age, maternal education, and sequencing depth.
+        Asterisks in (A) and (C) indicate significance (q \textless{} 0.2)
+        after multiple hypothesis correction.
+        (\textbf{D-E}) Enrichment plots for proprionate degradation I gene set
+        used in feature set enrichment analysis (FSEA) in children under 6m (D)
+        or over 18m (E). (\textbf{G}) summary of FSEA
+        results in children under 6m of age (left)
+        over 18 months (middle) or across all ages (right), colored based on the
         significance of the association. Markers indicate the individual
-        correlation of genes within a gene set, and vertical bars represent the
+        correlation of genes within a gene set,
+        and vertical bars colored bars represent the
         median correlation of that gene set.
+        Vertical dotted lines within each panel represent the median T statistic
+        for all genes in that age group.
     }

### 2023-08-29 manuscript/main.tex
update table 2
update table 3
add future prediction as table 4

@@
       \hline\hline
-      \textbf{Age group} & \multicolumn{2}{c}{\textbf{0 to 120 months}} & \multicolumn{2}{c}{\textbf{0 to 6 months}} & \multicolumn{2}{c}{\textbf{18 to 120 months}} \\ \hline
-      geneset & E.S. & q value & E.S. & q value & E.S. & q value \\\hline
-      Acetate syn. I & -0.122 & 0.555 & -0.194 & \color{blue}{\textbf{0.153}} & 0.127 & 0.602 \\
-      Acetate syn. II & -0.22 & \color{blue}{\textbf{0.066}} & -0.342 & \color{blue}{\textbf{0.02}} & -0.095 & 0.92 \\
-      Acetate syn. III & -0.21 & \color{blue}{\textbf{0.086}} & -0.31 & \color{blue}{\textbf{0.052}} & -0.086 & 0.92 \\
-      Glutamate deg. I & -0.404 & \color{blue}{\textbf{0.18}} & -0.284 & 0.816 & -0.249 & 0.602 \\
-      Glutamate syn. II & 0.095 & 0.602 & -0.134 & 0.98 & 0.242 & \color{red}{\textbf{0.047}} \\
-      Menaquinone syn. I & -0.085 & 0.563 & -0.083 & 0.816 & -0.17 & \color{blue}{\textbf{0.182}} \\
-      Propionate deg. I & -0.461 & \color{blue}{\textbf{0.13}} & -0.542 & \color{blue}{\textbf{0.02}} & -0.674 & \color{blue}{\textbf{0.041}} \\
-      Propionate syn. I & -0.334 & 0.555 & -0.32 & 0.571 & -0.676 & \color{blue}{\textbf{0.023}} \\
-      Tryptophan syn. & 0.057 & 0.974 & -0.072 & 0.816 & 0.119 & \color{red}{\textbf{0.041}} \\\hline\hline
+      \textbf{Age group} & \textbf{under 6m} & \textbf{under 6m} & \textbf{over 18m} & \textbf{over 18m} \\
+      \texttt{geneset} & \texttt{E.S.} & \texttt{q value} & \texttt{E.S.} & \texttt{q value} \\\hline
+      Acetate syn. I & -0.21 & \color{red}{\textbf{0.068}} & 0.123 & 0.403 \\
+      Acetate syn. II & -0.322 & \color{red}{\textbf{0.046}} & 0.113 & 0.866 \\
+      Acetate syn. III & -0.291 & \color{red}{\textbf{0.089}} & 0.109 & 0.866 \\
+      Glutamate syn. II & -0.152 & 0.882 & 0.293 & \color{blue}{\textbf{0.001}} \\
+      Propionate deg. I & -0.639 & \color{red}{\textbf{0.006}} & -0.495 & 0.446 \\\hline\hline
     \end{tabular}
@@
       \textbf{Subject Ages (months)} & \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
-      \multirow{5}{*}{0 to 6} & - & + & 13.01 ± 0.05 & -0.14 ± 0.01 \\ \cline{2-5}
-            & \multirow{2}{*}{taxa} & - & 12.66 ± 0.05 & -0.1 ± 0.01 \\ \cline{3-5}
-            &       & + & 12.7 ± 0.05 & -0.12 ± 0.01 \\ \cline{2-5}
-            & \multirow{2}{*}{genes} & - & 12.56 ± 0.05 & -0.01 ± 0.01 \\ \cline{3-5}
-            &       & + & 12.56 ± 0.05 & -0.02 ± 0.01 \\ \cline{1-5}
-      \multirow{5}{*}{18 to 120} & - & + & 16.27 ± 0.04 & 0.506 ± 0.003 \\ \cline{2-5}
-            & \multirow{2}{*}{taxa} & - & 17.66 ± 0.04 & 0.363 ± 0.003 \\ \cline{3-5}
-            &       & + & 17.29 ± 0.04 & 0.429 ± 0.003 \\ \cline{2-5}
-            & \multirow{2}{*}{genes} & - & 18.01 ± 0.04 & 0.303 ± 0.003 \\ \cline{3-5}
-            &       & + & 17.86 ± 0.04 & 0.333 ± 0.004 \\\hline\hline
+      \multirow{5}{*}{0 to 6} & -    & + & 13.37 ± 0.05 & -0.16 ± 0.01 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa}  & - & 12.99 ± 0.04 & -0.11 ± 0.01 \\ \cline{3-5}
+            &                        & + & 13.01 ± 0.04 & -0.13 ± 0.01 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 12.95 ± 0.04 & -0.05 ± 0.01 \\ \cline{3-5}
+            &                        & + & 12.95 ± 0.04 & -0.05 ± 0.01 \\ \cline{1-5}
+      \multirow{5}{*}{18 to 120} & - & + & 17.06 ± 0.04 & 0.511 ± 0.003 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa}  & - & 18.71 ± 0.04 & 0.347 ± 0.003 \\ \cline{3-5}
+            &                        & + & 18.25 ± 0.04 & 0.429 ± 0.003 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 18.96 ± 0.04 & 0.299 ± 0.003 \\ \cline{3-5}
+            &                        & + & 18.74 ± 0.04 & 0.341 ± 0.003 \\ \hline\hline
     \end{tabular}
@@
     prediction models. Confidence intervals (C.I.) are calculated from the
+    distribution of metrics from repeated CV at a confidence level of 95.
+    Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
+    Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.}
+    \end{center}
+\end{table}
+
+\begin{table}[!h]
+    \begin{center}
+    \begin{tabular}{|r|r|r|r|r}
+        \hline
+        \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set correlation (± C.I.)} & \textbf{Test set Root-mean-square error (± C.I.)} \\\hline
+        - & + & 20.25 ± 0.07 & 0.43 ± 0.0 \\ \cline{1-4}
+        \multirow{2}{*}{taxa} & - & 22.32 ± 0.08 & 0.07 ± 0.0 \\ \cline{2-4}
+            & + & 21.53 ± 0.08 & 0.27 ± 0.0 \\ \cline{1-4}
+        \multirow{2}{*}{genes} & - & 22.21 ± 0.08 & 0.12 ± 0.0 \\ \cline{2-4}
+            & + & 22.08 ± 0.08 & 0.15 ± 0.0 \\\hline\hline
+    \end{tabular}
+    \caption{\label{tab:rfbench}Benchmark metrics for the cognitive assessment score
+    future prediction models. Confidence intervals (C.I.) are calculated from the
     distribution of metrics from repeated CV at a confidence level of 95.

### 2023-08-30 manuscript/main.tex
add subsection on future pred

@@
 As with LMs, RF models for children under 6 months old
-performed poorly (mean test-set correlation -0.12, mean RMSE 12.70),
+performed poorly (mean test-set correlation -0.13, mean RMSE 13.01),
 but RFs were consistently able to learn the relationship between taxa
 and cognitive function scores in children over 18 months of age (mean
-test-set correlation 0.429, mean RMSE 17.29). For both age groups,
-species that were important in RF overlapped with those that were
+test-set correlation 0.429, mean RMSE 18.25).[]
+Species that were important in RF overlapped with those that were
 significant when testing the relationship with LMs (Figure
-3A-B, Table S1A-B); except for \emph{Eubacterium ramulus},
-all taxa that were significant using LMs belonged to a group
-responsible for 60\% of the total importance in RF models
-(Table 3, Figure 3A).
-However, the rank-ordering of LM significance and RF importance
-had substantive differences. For example, in children under 6 months,
-only \emph{B. wexlerae} was significantly associated with cognitive function in LMs, but
-was ranked 20th in importance for RF models.
-
-Many taxa highly ranked in importance in RF models in both age
-groups, including several that were identified in LMs for children over
-18 months old, including \emph{R. gnavus} (0--6 months, rank = 13;
-18--120 months, rank = 7) and \emph{G. pamelaeae} (0--6 months, rank =
-16; 18--120 months, rank = 20), while others such as \emph{Allistipes
-finegoldii} were age-group specific. A number of taxa that were important in RF models
-were not statistically significant when using LMs after
-multiple hypothesis correction. However, these taxa often had small nominal
-p-values. For example, \emph{Erysipelatoclostridium ramosum} was the
-most important feature in RF models for children under 6 months old and
-had an LM p-value of 0.04. Subject age was consistently ranked highly in
+3A-B, Table S1A-B); Both analysis identified 
+\textit{F. prausnitzii}, % check if we've mentioned these before, or use the full name
+\textit{E. eligens}, 
+\textit{P. excrementihominis}, and
+\textit{Alistipes obesi} as important
+(q \textless 0.2 for LMs, cumulative importance \textless 60\% for RFs)
+but RF models did not identify \textit{Eggerthella} sp. CAG 298
+nor \textit{B. intestinihominis}, and \textit{E. ramulus}
+was just outside our cutoff (64\% cumulative importance).
+RF models also identified an additional 48 species
+that were important for discriminating cognitive function scores
+(Table 3, Figure 3A). Subject age was consistently ranked highly in
 feature importance, which could indicate that decision branches based on
@@
 or that age itself is a useful predictor.
+
+\subsection*{Gut microbial profiles in the first year of life predictict future cogntitive function}
+
+While LMs did not identify any taxa in the 0-6 month range
+that were associated with concurrently measured cognitive function,
+we reasoned that developmental trajectories that begin
+\textit{in utero} may not be immediately altered by
+microbial exposures, such that microbial influences early in life
+may take time to manifest. To test this hypothesis,
+we asked whether RF models trained on taxonomic profiles
+in the first year of life
+could predict \textit{future} cognitive performance (Figure 3B, D).
+Models were consistently able to learn microbial species that are associated
+with cognitive performance at a future visit
+(mean time between visits: XX, %TODO
+mean test set correlation: YY).
+
+Nearly all important taxa in RF models that predicted cognitive function
+in children older than 18 months were also important in
+models that predicted future cognitive function, if they were present (Figure 3B).
+At the same time, many additional species were important for future prediction models
+that were not important in the models measuring concurrent cognitive function.
+In particular, feature importance for future prediction models
+were dominated by two species of \textit{Klebsiella},
+\textit{K. pneumoniae} and \textit{K. varicola} (Figure 3C).
+Many of the other species that ranked highest in feature importance
+were also oportunistic pathogens, or previously reported
+to be linked with disease (eg. \textit{R. gnavus}, \textit{E. coli}).
 

### 2023-09-01 manuscript/main.tex
rm A equol.
align table 1
add future column

@@
     \centering
-    \begin{tabular}{|r|r|r|r|r|}
+    \begin{tabular}{|r|r|r|r|r|r|}
       \hline\hline
-      \textbf{Group} & \textbf{Subgroup} & \textbf{All} & \textbf{Under 6 months} & \textbf{Over 18 months} \\\hline
-      N subjects &  & 381 & 149 & 328 \\  \hline
-    \multirow{3}{*}{Samples} & 1 & 214 (56.2\%) & 91 (94.8\%) & 172 (61.9\%) \\ \cline{2-5}
-         & 2 & 90 (23.6\%) & 5 (5.2\%) & 60 (21.6\%) \\ \cline{2-5}
-         & \textgreater 2 & 77 (20.2\%) & 0 (0.0\%) & 46 (16.5\%) \\ \hline
-    \multirow{3}{*}{Age (months)} & min & 2.89 & 2.89 & 18.03 \\ \cline{2-5}
-         & max & 119.53 & 5.95 & 119.53 \\ \cline{2-5}
-         & median & 25.85 & 3.64 & 41.59 \\ \hline
-    \multirow{2}{*}{Sex} & F & 172 (45.1\%) & 42 (51.2\%) & 122 (44.0\%) \\   \cline{2-5}
-                         & M & 209 (54.9\%) & 40 (48.8\%) & 155 (56.0\%) \\  \hline
-    \multirow{6}{*}{Race} & White & 265 (69.6\%) & 48 (58.5\%) & 201 (72.6\%) \\   \cline{2-5}
-        & Black & 40 (10.5\%) & 15 (18.3\%) & 21 (7.6\%) \\ \cline{2-5}
-        & Indigenous & 1 (0.3\%) & 0 (0.0\%) & 1 (0.4\%) \\ \cline{2-5}
-        & Asian & 8 (2.1\%) & 2 (2.4\%) & 7 (2.5\%) \\ \cline{2-5}
-        & Mixed & 54 (14.2\%) & 14 (17.1\%) & 38 (13.7\%) \\ \cline{2-5}
-        & Other & 3 (0.8\%) & 1 (1.2\%) & 3 (1.1\%) \\ \hline
-    \multirow{6}{*}{Maternal Ed.} & Junior high school & 2 (0.5\%) & 0 (0.0\%) & 2 (0.7\%) \\   \cline{2-5}
-        & Some high school & 6 (1.6\%) & 3 (3.7\%) & 4 (1.4\%) \\ \cline{2-5}
-        & High school grad & 39 (10.2\%) & 13 (15.9\%) & 22 (7.9\%) \\ \cline{2-5}
-        & Some college & 100 (26.2\%) & 28 (34.1\%) & 60 (21.7\%) \\ \cline{2-5}
-        & College grad & 95 (24.9\%) & 19 (23.2\%) & 75 (27.1\%) \\ \cline{2-5}
-        & Grad/professional school & 126 (33.1\%) & 16 (19.5\%) & 106 (38.3\%) \\\hline\hline
+      \textbf{Group}              & \textbf{Subgroup}        & \textbf{All} & \textbf{Under 6 months} & \textbf{Over 18 months} & \textbf{Future} \\\hline
+      N subjects                  &                          & 381 & 82 & 277 & 104 \\  \hline
+    \multirow{3}{*}{Samples}      & 1                        & 314 (82.4\%) &  &  &  \\ \cline{2-6}
+                                  & 2                        & 55 (14.4\%) &  &  &  \\ \cline{2-6}
+                                  & \textgreater 2           & 12 (3.1\%) &  &  &  \\ \hline
+    \multirow{3}{*}{Age (months)} & min                      & 1.3 & 2.89 & 18.03 & 1.3 \\ \cline{2-6}
+                                  & max                      & 119.53 & 5.95 & 119.53 & 11.95 \\ \cline{2-6}
+                                  & median                   & 24.62 & 3.64 & 41.59 & 6.08 \\ \hline
+    \multirow{2}{*}{Sex}          & F                        & 172 (45.1\%) & 42 (51.2\%) & 122 (44.0\%) & 47 (45.2\%) \\   \cline{2-6}
+                                  & M                        & 209 (54.9\%) & 40 (48.8\%) & 155 (56.0\%) & 57 (54.8\%) \\  \hline
+    \multirow{6}{*}{Race}         & White                    & 265 (69.6\%) & 48 (58.5\%) & 201 (72.6\%) & 62 (59.6\%) \\   \cline{2-6}
+                                  & Black                    & 40 (10.5\%) & 15 (18.3\%) & 21 (7.6\%) & 17 (16.3\%) \\ \cline{2-6}
+                                  & Indigenous               & 1 (0.3\%) & 0 (0.0\%) & 1 (0.4\%) & 0 (0.0\%) \\ \cline{2-6}
+                                  & Asian                    & 8 (2.1\%) & 2 (2.4\%) & 7 (2.5\%) & 3 (2.9\%) \\ \cline{2-6}
+                                  & Mixed                    & 54 (14.2\%) & 14 (17.1\%) & 38 (13.7\%) & 16 (15.4\%) \\ \cline{2-6}
+                                  & Other                    & 3 (0.8\%) & 1 (1.2\%) & 3 (1.1\%) & 2 (1.9\%) \\ \hline
+    \multirow{6}{*}{Maternal Ed.} & Junior high school       & 2 (0.5\%) & 0 (0.0\%) & 2 (0.7\%) & 0 (0.0\%) \\   \cline{2-6}
+                                  & Some high school         & 6 (1.6\%) & 3 (3.7\%) & 4 (1.4\%) & 2 (1.9\%) \\ \cline{2-6}
+                                  & High school grad         & 39 (10.2\%) & 13 (15.9\%) & 22 (7.9\%) & 17 (16.3\%) \\ \cline{2-6}
+                                  & Some college             & 100 (26.2\%) & 28 (34.1\%) & 60 (21.7\%) & 30 (28.8\%) \\ \cline{2-6}
+                                  & College grad             & 95 (24.9\%) & 19 (23.2\%) & 75 (27.1\%) & 30 (28.8\%) \\ \cline{2-6}
+                                  & Grad/professional school & 126 (33.1\%) & 16 (19.5\%) & 106 (38.3\%) & 23 (22.1\%) \\ \hline\hline
     \end{tabular}
@@
 We identified several species in the family Eggerthelaceae that were
-associated with cognitive function, including \emph{G.pamelaeae, A. equolofaciens, A. celatus}
+associated with cognitive function, including \emph{G.pamelaeae, A. celatus}
 (formerly regarded a subspecies of \emph{A. equolofaciens}

### 2023-09-01 manuscript/main.tex
write up fig 2 subscales

@@
 
+Each of the cognitive assessments (MSEL, WPPSI, WISC)
+are composites of multiple subscales that assess different cognitive domains.
+While the composite scores are broadly analagous accross ages,
+the subscales assessed in each test are different.
+In order to assess whether microbial taxa were associated
+
 Given that different microbial species might occupy the same metabolic
@@
 \cite{valles-colomerNeuroactivePotentialHuman2019}
-and concurrent cognitive ability score (Table 2, Figure 2C-G) and found that
+and concurrent cognitive ability score (Table 2, Figure 2D-F) and found that
 several metabolic pathways were either significantly enriched or depleted in

### 2023-09-01 manuscript/main.tex
update figure 2 results

@@
 the subscales assessed in each test are different.
-In order to assess whether microbial taxa were associated
+In order to assess whether microbial taxa were associated with
+the development of specific cognitive domains,
+we fit multivariate linear models for each species present in at least 10\%
+of samples for children under 6 months, 18 to 48 months, or all ages
+with a concurrent MSEL score (Figure 2C).
+Several of the taxa that were associated with composite scores (Figure 2B)
+were also associated with one or more subscores, including
+\textit{F. prausnitzii} and \textit{A. obesi}.
+Interestingly, while no taxa were significantly associated
+with composite scores in chilren under 6 months old,
+two species from the genus \textit{Streptococcus}
+(\textit{S. peroris} and \textit{S. mitis}), as well as
+\textit{Bacteroides fragilis}, were positively associated
+with expressive language in this age group.
+\textit{S. mitis} was also associated with expressive language
+when looking at all ages, while \textit{Eisenbergiella massiliensis},
+\textit{Anaerotruncus colihominis}, and an unclassified \textit{Blautia}
+species were negatively associated. % Add more here if they match brain?
 
@@
 several metabolic pathways were either significantly enriched or depleted in
-children with higher cognitive ability scores. This was true both when
-considering all ages (0-120 months) together, though the enrichment of most
-pathways was more pronounced either in children under 6 months or those over 18
-months. For example, genes for degrading the 3-carbon SCFA, propionate,
+children with higher cognitive ability scores.
+For example, genes for degrading the 3-carbon SCFA, propionate,
 were significantly depleted in children with higher cognitive ability
-scores across all age groups tested (Table 2; propionate degradation I,
-under 6 months, enrichment score (E.S.) = -0.542, corrected p-value (q)
-= 0.020; over 18 months, E.S. = -0.674, q = 0.041). Interestingly, genes
-for propionate synthesis were also significantly depleted in higher
-scoring children over 18 months (E.S. -0.676, q = 0.023), as were genes
-for synthesizing the 2 carbon SCFA acetate in children under 6 months
-old (acetate synthesis I, E.S. = -0.194, q = 0.153; acetate synthesis
-II, E.S. = -0.342, q = 0.020; acetate synthesis III, E.S. = -0.31, q = 0.052).
-In children over 18 months old, synthesis of menaquinone (vitamin K)
-was also negatively associated with cognitive function score (menaquinone synthesis I,
-E.S. = -0.170, q = 0.0183), while
-genes for the synthesis of the amino acids
-glutamate and tryptophan were significantly enriched in children with
-higher cognitive function scores (glutamate synthesis I, E.S. = 0.242, 
-q = 0.047; tryptophan synthesis, E.S. = 0.119, q = 0.041).
-Taken together, these results suggest that
+scores in children under 6m (Table 2; propionate degradation I,
+enrichment score (E.S.) = -0.64, corrected p-value (q)
+= 0.006), as were genes for synthesizing the 2 carbon SCFA acetate
+(acetate synthesis I, E.S. = -0.210, q = 0.07; acetate synthesis
+II, E.S. = -0.342, q = 0.049; acetate synthesis III, E.S. = -0.291, q = 0.09).
+In children over 18 months old, synthesis of the amino acid
+glutamate was significantly enriched in children with
+higher cognitive function scores (glutamate synthesis II, E.S. = 0.293, 
+q = 0.0008). Taken together, these results suggest that
 microbial metabolic activity, particularly the metabolism (synthesis and

### 2023-10-03 manuscript/TableS1.jl
some modes to supp table 1

@@ -7,7 +7,7 @@ using Statistics
 using Chain
 using JLD2
 using MLJ
-using DataFrames.PrettyTables
+using DataFrames.PrettyTables: LatexHighlighter, pretty_table
 
 isdir(tablefiles()) || mkpath(tablefiles())
 
@@ -90,14 +90,15 @@ models_to_report = [
 ## 4. LaTeX outputs
 spec_hl = LatexHighlighter((val, i, j) -> j == 2, ["textit"])
 spec_ft = (val, i, j) -> j == 2 ? replace(val, "_"=> " ") : val
+isdir(tablefiles("figure3_merits_importances")) || mkpath(tablefiles("figure3_merits_importances"))
 
 for m in models_to_report
-
+    @info m.name
     supptbl = singlemodel_importances_suppltable(m)
     prettyformat_suppltable!(supptbl)
     CSV.write(tablefiles("figure3_merits_importances", "importances_"*m.name*".csv"), supptbl)
 
-    @show latextable = pretty_table(supptbl; highlighters=spec_hl, formatters=spec_ft,
+    pretty_table(supptbl; highlighters=spec_hl, formatters=spec_ft,
         header = [ "Rank", "Variable", "Average fitness-weighted importance", "Relative fitness-weighted importance", "Rank-cumulative relative importance " ],
         backend = Val(:latex) )
     # write(tablefiles("figure3_merits_importances", "latex_importances_"*m.name*".txt"), latextable) # @Kevin I don't know how to save that.

### 2023-10-03 manuscript/main.tex
Figure 3 legend

@@
     \caption{
-        Random forest models predict concurrently measured cognitive function.
+        Random forest models predict both concurrent and future measured cognitive function.
         Comparison of RF predictor importance versus linear models for children
-        between birth and 6 months old (\textbf{A}), and for those older than 18 months
-        (\textbf{B}). Colors represent whether the species belong to the group of
+        over 18 months (\textbf{A}).
+        Colors represent whether the species belong to the group of
         top-important features that account for 60\% of the cumulative
         importance on the RF model, if that species was significant (q
-        \textless{} 0.2) in linear models, both, or neither. Ranked feature
-        importance for taxa in RF models for children between birth and 6 months
-        old (\textbf{C}), and for those older than 18 months (\textbf{D}). Taxa that are important
-        for RF models both for children under 6 months and for those over 18
-        months (\textbf{E}), only for children under 6 months (\textbf{F}), and only for children
-        over 18 months (\textbf{G})
+        \textless{} 0.2) in linear models, both, or neither.
+        Comparison of RF predictor importance for microbial features
+        between concurrent and future cognitive function measure
+        (\textbf{B}).
+        Colors represent whether the species belong to the group of
+        top-important features that account for 60\% of the cumulative
+        importance on the RF model only for concurrent prediction,
+        only for the future prediction, or for both models.
+        Features absent of one model are shown on the origin-aligned sideplots.
+        Distribution of ages for each microbiome sampling - cognitive assesment dyad (\textbf{C}).
+        Blue violin plot shows distribution of sampling age,
+        while red violin plot shows distribution of asessment age.
+        Black lines connect ages for each individual subject.      
+        Ranked feature importance for taxa in RF models for
+        both concurrent (\textbf{D}) and future (\textbf{E})
+        cognitive function measurement prediction.
+        Red dots indicate how relative importances change (among taxa)
+        when demographics (sex, SES) are arred as covariables.
+        Abundances and prevalences of taxa that are important
+        for RF models both for concurrent (\textbf{E}) and future (\textbf{G}) assesment prediction.
         }

### 2023-10-03 manuscript/main.tex
fixed typo on Fig 3 caption
checked previous mentions of bugs

@@
         Abundances and prevalences of taxa that are important
-        for RF models both for concurrent (\textbf{E}) and future (\textbf{G}) assesment prediction.
+        for RF models both for concurrent (\textbf{F}) and future (\textbf{G}) assesment prediction.
         }
@@
 3A-B, Table S1A-B); Both analysis identified 
-\textit{F. prausnitzii}, % check if we've mentioned these before, or use the full name
-\textit{E. eligens}, 
-\textit{P. excrementihominis}, and
-\textit{Alistipes obesi} as important
+\textit{F. prausnitzii}, % F. prausz was mentioned before, ~ line 395
+\textit{E. eligens}, % E. eligens was mentioned before, ~ line 395
+\textit{Parasutterella excrementihominis}, and % Not mentioned before
+\textit{Alistipes obesi} as important % A. obesi was mentioned before, ~ lines 120, 390
 (q \textless 0.2 for LMs, cumulative importance \textless 60\% for RFs)

### 2023-10-03 manuscript/main.tex
added numbers to future experiment

@@
 with cognitive performance at a future visit
-(mean time between visits: XX, %TODO
-mean test set correlation: YY).
+(mean time between visits: 21.5 months [sd = 9.9 months] (Figure 3C),
+mean test set correlation: 0.27).
 

### 2023-10-03 manuscript/main.tex
changed figure reference
Added methods for future (fig3) and subscales (fig 4)

@@
 were dominated by two species of \textit{Klebsiella},
-\textit{K. pneumoniae} and \textit{K. varicola} (Figure 3C).
+\textit{K. pneumoniae} and \textit{K. varicola} (Figure 3D, E).
 Many of the other species that ranked highest in feature importance
@@
 demographics (sex and maternal education as a proxy of socioeconomic
-status) on participants from different age groups. Age (in months) was
-provided as a covariate for all models (Table S3).
+status) on participants from different age groups.
+
+For the concurrent cognitive assessment score preditions,
+all viable datapoints containing 
+age at data collection, demographics, stool metagenomics and cognitive assessment were employed.
+The cognitive assessments were collected as either composite scores
+(for the experiments in Figure \textbf{3}) or, when available, segregate scores on the
+Mullen Scales of Early Learning ( Figure \textbf{4 A-B}: Expressive Language, Gross Motor and Visual Reception).
+For the concurrent prediction experiments, when multiple viable longitudinal samples were present,
+the single earliest viable datapoint inside the age range of the experiment was considered.
+For the future prediction experiments, every demographics-complete subject
+contributed a single datapoint built from the combination of
+the latest metagenome collected within the stool collection time range and
+the latest future cognitive assessment on the assesment time range.
+Age (in months) was provided as a covariate for all models (Table S3).
 

### 2023-10-03 manuscript/Figure3-futurecogscore-calculations.jl
separated hashnames.txt for future prediction experiment

@@ -77,12 +77,12 @@ for oldname in names(mdata_ecs_df)[20:end]
     rename!(mdata_ecs_df, oldname => randstring(['A':'Z'; 'a':'z'], 12))
 end
 newnames = names(mdata_ecs_df)[20:end]
-open(scratchfiles("longnames.txt"), "w") do io
+open(scratchfiles("future_longnames.txt"), "w") do io
     for i in oldnames
         println(io, i)
     end
 end
-open(scratchfiles("hashnames.txt"), "w") do io
+open(scratchfiles("future_hashnames.txt"), "w") do io
     for i in newnames
         println(io, i)
     end

### 2023-10-04 manuscript/main.tex
Subscales results

@@
 
+\subsubsection*{Gut microbial taxa relate diffrentially to distinct cognitive subscales} %TODO: add numbers!
+
+Is it possible to associate the predictive species with different components of cognition? 
+After probing the aforedescribed importances of potential of taxa and ECs 
+towards composite cognitive scores, we proceeded to verify if and how this relationship 
+would hold for different subcomponents of the cognitive function assessment. 
+We collected those in the form of the 
+Mullen Scales of Early Learning (MSEL - see \textit{see Methods}). 
+Taxonomic profiles coupled with demographics exhibited predictive potential towards three probed subscales, 
+namely Expressive Language (EL; test-set RMSE XXX, test-set correlation YYYY), 
+Gross Motor (GM; test-set RMSE XXX, test-set correlation YYYY) and 
+Visual Reception (VR; test-set RMSE XXX, test-set correlation YYYY).
+Between the subscales, not only the predictability (measured by the test-set figures of merit) is different,
+but also the relative importances of several taxa.
+Although certain species such as \textit{B. longum}, \textit{F.prausznitzii} 
+have similar relative importance across all subscales, others are substantially different.
+Notably, \textit{Bifidobacterium pseudocatenulatum}, \textit{B. wexlerae} and \textit{E. eligens} are
+substantially more important for the prediction of EL, while
+\textit{R. faecis}, \textit{S. salivarius} and \textit{F. saccharivorans} are
+distinctly predictive of GM, and
+\textit{Clostridium innocuum}, \textit{B. vulgatus} 
+stand out on models targeting VR.
+
 \subsubsection*{Gut microbial taxonomic profiles predict brain structure differences}

### 2023-10-04 manuscript/TableS2.jl
Adapted table S2 toa ccount for interesting segments only

@@ -13,7 +13,7 @@ isdir(tablefiles()) || mkpath(tablefiles())
 
 ## 1. Loading the model files
 RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
-brain_models = JLD2.load(modelfiles("manuscript", "brain_models.jld"))["brain_models"]
+brain_models = JLD2.load(modelfiles("brain_models.jld"))["brain_models"]
 include("manuscript/Figure4-definitions.jl")
 
 ## 2. Calculating the Figures of Merit
@@ -22,6 +22,7 @@ tableS2 = mapreduce( vcat, eachindex(ordered_brain_segments_list)) do i
               :Test_MAPE => round(mean(brain_models[ordered_brain_segments_list[i]].merits.Test_MAPE); digits = 4), 
               :Test_Cor => round(mean(brain_models[ordered_brain_segments_list[i]].merits.Test_Cor); digits = 4))
 end
+tableS2 = subset(tableS2, :variable => x -> x .∈ Ref(interesting_segments) )
 tableS2.variable = uppercasefirst.(map(x -> replace(x, "-"=>" "), tableS2.variable))
 
 CSV.write(joinpath("manuscript", "assets", "TableS2.csv"), tableS2)

### 2023-10-04 manuscript/main.tex
fixed figure refs and typos on paragraph

@@
 We collected those in the form of the 
-Mullen Scales of Early Learning (MSEL - see \textit{see Methods}). 
-Taxonomic profiles coupled with demographics exhibited predictive potential towards three probed subscales, 
+Mullen Scales of Early Learning (MSEL - \textit{see Methods}). 
+Taxonomic profiles coupled with demographics exhibited predictive potential towards three probed subscales (Figure 4A), 
 namely Expressive Language (EL; test-set RMSE XXX, test-set correlation YYYY), 
@@
 Between the subscales, not only the predictability (measured by the test-set figures of merit) is different,
-but also the relative importances of several taxa.
-Although certain species such as \textit{B. longum}, \textit{F.prausznitzii} 
+but also the relative importances of several taxa (Figure 4B).
+Although certain species such as \textit{B. longum}, \textit{F. prausznitzii} 
 have similar relative importance across all subscales, others are substantially different.

### 2023-10-04 manuscript/main.tex
updated numbers

@@
 profiles with individual brain regions identified in MRI scans
-(children over 18 months old, N = 121).
+(children over 18 months old, N = 124).
 Some brain regions were more readily
@@
 These included the L/R lingual gyrus (mean RF correlations, Left =
-0.421, Right = 0.434; importance for age, Left = 7.5\%, Right =
+0.444, Right = 0.441; importance for age, Left = 7.5\%, Right =
 8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
-0.200, Right = 0.273; importance for age, Left = 3.6\%, Right = 6.3\%).
+0.267, Right = 0.315; importance for age, Left = 3.6\%, Right = 6.3\%).
 In many cases, however, age was not an important variable in
 high-performing models, such as that for the left accumbens area (mean
-RF correlations = 0.288; relative age importances = 1.2\%).
+RF correlations = 0.275; relative age importances = 1.2\%).
 

### 2023-10-04 manuscript/main.tex
updated table 5

@@
       \textbf{Statistic} & \textbf{MAPE} & \textbf{R} \\\hline
-      Mean & 0.073 & 0.137 \\
-      Standard deviation & 0.028 & 0.137 \\
-      Maximum & 0.206 & 0.605 \\
-      75th percentile & 0.082 & 0.195 \\
-      50th percentile & 0.067 & 0.134 \\
-      25th percentile & 0.057 & 0.062 \\
-      Minimum & 0.035 & -0.147 \\\hline\hline
+      Mean & 0.069 & 0.206 \\
+      Standard deviation & 0.015 & 0.099 \\
+      Maximum & 0.11 & 0.444 \\
+      75th percentile & 0.076 & 0.262 \\
+      50th percentile & 0.066 & 0.206 \\
+      25th percentile & 0.057 & 0.16 \\
+      Minimum & 0.046 & 0.005 \\\hline
     \end{tabular}

### 2023-10-04 manuscript/main.tex
Apply suggestions from code review
First review of PR by @kescobo

Co-authored-by: Kevin Bonham <kbonham@wellesley.edu>

@@
 
-\subsubsection*{Gut microbial taxa relate diffrentially to distinct cognitive subscales} %TODO: add numbers!
-
-Is it possible to associate the predictive species with different components of cognition? 
-After probing the aforedescribed importances of potential of taxa and ECs 
-towards composite cognitive scores, we proceeded to verify if and how this relationship 
-would hold for different subcomponents of the cognitive function assessment. 
-We collected those in the form of the 
+\subsubsection*{Gut microbial taxa are associated with specific cognitive functions} %TODO: add numbers!
+
+Given that the development of different domains of cognitive function
+(eg gross motor development vs visual perception) may develop at different times
+and be affected by different molecular signals,
+we sought to determine whether the associations of microbial taxa
+with cognitive function are general across all cognitive subdomains,
+or specific.
+To accomplish this, we assessed individual subscales of the  
 Mullen Scales of Early Learning (MSEL - \textit{see Methods}). 
@@
 
-If there are causal effects of microbial metabolism on cognitive
+Given that individual microbes could be associated with particular domains
+of cognitive function and not others, if there are causal effects of microbial metabolism on cognitive
 function, they might be reflected in changes in neuroanatomy. We again

### 2023-10-04 manuscript/main.tex
Apply suggestions from code review
First review of PR by @kescobo

Co-authored-by: Kevin Bonham <kbonham@wellesley.edu>

--- 2023-10-04 manuscript/main.tex
@@
 \textit{Parasutterella excrementihominis}, and % Not mentioned before
-\textit{Alistipes obesi} as important % A. obesi was mentioned before, ~ lines 120, 390
+\textit{A. obesi} as important % A. obesi was mentioned before, ~ lines 120, 390
 (q \textless 0.2 for LMs, cumulative importance \textless 60\% for RFs)
@@
 
-\subsubsection*{Gut microbial taxa relate diffrentially to distinct cognitive subscales} %TODO: add numbers!
-
-Is it possible to associate the predictive species with different components of cognition? 
-After probing the aforedescribed importances of potential of taxa and ECs 
-towards composite cognitive scores, we proceeded to verify if and how this relationship 
-would hold for different subcomponents of the cognitive function assessment. 
-We collected those in the form of the 
-Mullen Scales of Early Learning (MSEL - \textit{see Methods}). 
-Taxonomic profiles coupled with demographics exhibited predictive potential towards three probed subscales (Figure 4A), 
-namely Expressive Language (EL; test-set RMSE XXX, test-set correlation YYYY), 
-Gross Motor (GM; test-set RMSE XXX, test-set correlation YYYY) and 
-Visual Reception (VR; test-set RMSE XXX, test-set correlation YYYY).
-Between the subscales, not only the predictability (measured by the test-set figures of merit) is different,
-but also the relative importances of several taxa (Figure 4B).
-Although certain species such as \textit{B. longum}, \textit{F. prausznitzii} 
-have similar relative importance across all subscales, others are substantially different.
-Notably, \textit{Bifidobacterium pseudocatenulatum}, \textit{B. wexlerae} and \textit{E. eligens} are
-substantially more important for the prediction of EL, while
-\textit{R. faecis}, \textit{S. salivarius} and \textit{F. saccharivorans} are
-distinctly predictive of GM, and
-\textit{Clostridium innocuum}, \textit{B. vulgatus} 
-stand out on models targeting VR.
-
 \subsubsection*{Gut microbial taxonomic profiles predict brain structure differences}
@@
 profiles with individual brain regions identified in MRI scans
-(children over 18 months old, N = 124).
+(children over 18 months old, N = 121).
 Some brain regions were more readily
@@
 These included the L/R lingual gyrus (mean RF correlations, Left =
-0.444, Right = 0.441; importance for age, Left = 7.5\%, Right =
+0.421, Right = 0.434; importance for age, Left = 7.5\%, Right =
 8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
-0.267, Right = 0.315; importance for age, Left = 3.6\%, Right = 6.3\%).
+0.200, Right = 0.273; importance for age, Left = 3.6\%, Right = 6.3\%).
 In many cases, however, age was not an important variable in
 high-performing models, such as that for the left accumbens area (mean
-RF correlations = 0.275; relative age importances = 1.2\%).
+RF correlations = 0.288; relative age importances = 1.2\%).
 
@@
 The cognitive assessments were collected as either composite scores
-(for the experiments in Figure \textbf{3}) or, when available, segregate scores on the
+(for the experiments in Figure \textbf{3}) or, when available, subscores on the
 Mullen Scales of Early Learning ( Figure \textbf{4 A-B}: Expressive Language, Gross Motor and Visual Reception).

### 2023-10-06 manuscript/main.tex
fix table 2

@@
 \addbibresource{zotero.bib}
+\addbibresource{addons.bib}
 
@@
     \begin{tabular}{|r|r|r|r|r|r|r|}
-      \hline\hline
-      \textbf{Age group} & \textbf{under 6m} & \textbf{under 6m} & \textbf{over 18m} & \textbf{over 18m} \\
+        \hline
+        \textbf{Age group}& \multicolumn{2}{c|}{\textbf{0 to 6m}} & \multicolumn{2}{c|}{\textbf{18 to 120m}} \\\hline
         \texttt{geneset} & \texttt{E.S.} & \texttt{q value} & \texttt{E.S.} & \texttt{q value} \\\hline
-      Acetate syn. I & -0.21 & \color{red}{\textbf{0.068}} & 0.123 & 0.403 \\
-      Acetate syn. II & -0.322 & \color{red}{\textbf{0.046}} & 0.113 & 0.866 \\
-      Acetate syn. III & -0.291 & \color{red}{\textbf{0.089}} & 0.109 & 0.866 \\
-      Glutamate syn. II & -0.152 & 0.882 & 0.293 & \color{blue}{\textbf{0.001}} \\
-      Propionate deg. I & -0.639 & \color{red}{\textbf{0.006}} & -0.495 & 0.446 \\\hline\hline
+        Acetate syn. I & -0.165 & 0.284 & 0.192 & \color{red}{\textbf{0.145}} \\
+        Acetate syn. II & -0.491 & \color{blue}{\textbf{0.059}} & 0.248 & \color{red}{\textbf{0.15}} \\
+        Acetate syn. III & -0.409 & \color{blue}{\textbf{0.133}} & 0.215 & 0.209 \\
+        Butyrate syn. II & -0.357 & 0.284 & 0.414 & \color{red}{\textbf{0.178}} \\
+        Glutamate syn. II & 0.343 & 0.284 & 0.355 & \color{red}{\textbf{0.023}} \\
+        Propionate deg. I & -0.659 & \color{blue}{\textbf{0.029}} & 0.154 & \color{red}{\textbf{0.178}} \\
+        Tryptophan syn. & 0.143 & 0.284 &  & \\\hline
     \end{tabular}
@@
 substantially more important for the prediction of EL, while
-\textit{R. faecis}, \textit{S. salivarius} and \textit{F. saccharivorans} are
+\textit{R. faecis}, \textit{Streptococcus salivarius} and \textit{F. saccharivorans} are
 distinctly predictive of GM, and
@@
 These included the L/R lingual gyrus (mean RF correlations, Left =
-0.444, Right = 0.441; importance for age, Left = 7.5\%, Right =
-8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
-0.267, Right = 0.315; importance for age, Left = 3.6\%, Right = 6.3\%).
+0.444, Right = 0.441; importance for age, Left = 6.8\%, Right =
+7.7\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
+0.267, Right = 0.315; importance for age, Left = 3.6\%, Right = 6.2\%).
 In many cases, however, age was not an important variable in
 high-performing models, such as that for the left accumbens area (mean
-RF correlations = 0.275; relative age importances = 1.2\%).
+RF correlations = 0.275; relative age importances = 1.1\%).
 
@@
 brain region models (R = 0.288), as compared to the right which
-had a negative mean test-set correlation (R = -0.041),
+had a negligible test-set correlation (R = 0.005),
 indicating that generalizable models were not identified.
@@
 by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.3\% importance),
-\emph{B. ovatus} (3.7\% importance), and
-\emph{B. uniformis} (2.9\% importance).
+\emph{B. ovatus} (3.2\% importance), and
+\emph{B. uniformis} (2.8\% importance).
 
@@
 were also identified in LMs and RF models of cognitive function (Figures 2 and 3). 
-In general, for all segments, a consistent number of species between 31 and 45 (median = 37, less than 30\% of the input features)
+In general, for all segments, a consistent number of species between 28 and 42 (median = 38, less than 30\% of the input features)
 was responsible for half of the cumulative importance,
-regardless of model performance (Figure 4A-B, Table S2).
+regardless of model performance (Figure 4C-E, Table S2).
 We observed two major patterns of importance distribution from
@@
 just one or two brain segments. Notable cases of the first pattern
-included seven species - \emph{Anaerostipes hadrus}, \emph{B.
-vulgatus}, \emph{Fusicatenibacter saccharivorans}, \emph{Ruminococcus
-torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
+included nine species - 
+\emph{Anaerostipes hadrus},
+\emph{B. vulgatus},
+\emph{B. ovatus},
+\emph{Fusicatenibacter saccharivorans},
+\emph{Ruminococcus torques},
+\emph{Eubacterium rectale},
+\emph{Coprococcus comes},
+\emph{Asaccharobacter celatus} and
 \emph{B. wexlerae}, which combined, account for approximately 10\% of the cumulative relative importances,
-computed after fitness-weighting on the reported brain segments (Figure 4A-B).
+computed after fitness-weighting on the reported brain segments (Figure 4D-E).
 
@@
 highest average relative importance along all reported 
-(mean = 1.7\%; min = 1.0\%; max = 3.8\%, Figure 4B-C).
+(mean = 1.7\%; min = 1.0\%; max = 3.8\%, Figure 4D-E).
 \emph{B. wexlerae}, which was important in RF models of cognitive function,
@@
 cluster contains \emph{B. ovatus} and \emph{B. uniformis},
-but also \emph{Alistipes finegoldii and Streptococcus salivaris}.
+but also \emph{Alistipes finegoldii and S. salivarius}.
 
@@
 
-While we did not identify any statistical associations between individual species
-and cognitive performance in our youngest subset (0 to 6 months),
-it may be because brain developmental trajectories begin \textit{in utero},
-and microbial influences (that can only begin at birth)
-take additional time to manifest.
-Future studies should attempt to link early-life microbial exposures
-to future cognitive performance or brain development.
-
+--- Start of Guilherme flag for review of section:
 We identified several species in the family Eggerthelaceae that were
@@
 targets for further mechanistic studies.
+--- End of Guilherme flag for review of section
+
+While we did not identify any statistical associations between individual species
+and cognitive performance in our youngest subset (0 to 6 months),
+it may be because brain developmental trajectories begin \textit{in utero},
+and microbial influences (that can only begin at birth)
+take additional time to manifest.
+In line with this hypothesis, RF models that were able to learn associations
+between early-life microbiomes and future cognitive function scores (Figure 3).
+The top-ranked species for these models included several 
+taxa that were important for predicting concurrent cognitive assessment scores after 18 months, notably
+pathobiontic \textit{Enterobacteriaceae} such as
+\textit{Escherichia} and \textit{Klebsiella}. 
+Representatives of these genera are known to be early and long-term persistent
+colonizers of infant intestine that can, in some settings,
+promote inflammatory response \cite{popeEnterobacEarlyGut2019}.
+By contrast, several traditionally inflammation-reducing representatives
+of the \textit{Bifidobacterium} genus,
+from the early HMO-metabolizer \textit{B. longum} to the
+later complex carbohydrate relying \textit{B. pseudocatenulatum} \cite{chungBifidoPsudocatenulatum2021},
+commonly associated with weaning to solid food in early age,
+also showed high relative importance.
+The involvement of most important taxa in different aspects of inflammatory response
+resonate with previous findings that associate chronic inflammation with
+changes to neural responses in some types of stimulae, notably visual\cite{xieChronicInfNeuro2019}.
 
@@
 \cite{hoogmanSubcorticalBrainVolume2017}.
+
+Our ability to probe individual cognitive subscales provided
+results that also endorse previous findings and provide insight to
+potential gut-microbial connections between cognition and neuroanatomy.
+For example, a longitudinal neurocognitive development study pointed at bilateral
+associations between the thalamus and the EL component of the Mullen scales\cite{chenLongitudinalBrainMullen2023}.
+In line with this finding, we found that taxa such as 
+\textit{B. wexlerae} and \textit{A. hadrus} to be among the top 15\% important predictors
+on both left and right thalamus and the EL subscale.
+Another instance of those findings is a previously reported association between
+the central opercular cortex (COC) and Gross Motor development\cite{jiangBrainNetworkCognition2023}. In this case, we identified that
+butyrate-producing \textit{Fusicatenibacter saccharivorans} is among the 15\% most important predictors
+for both the GM subscale and three bilateral components of the COC
+(left and right pars orbitalis, pars opercularis and pars triangularis).
+Previous results also link to relationships between the caudate and the cerebellum,
+and the Visual Reception score\cite{deoniMyelinationMullenPreprint2022}.
+\textit{S. salivarius} was consistently found among the top 20\% predictors for this subscale, along with
+both sides of the thalamus and cerebellum white matter, along with cerebellar vermal lobules I-X.
 

### 2023-10-06 manuscript/main.tex
added `addons.bib` to TeX header
new discussion paragraph

--- 2023-10-06 manuscript/main.tex
@@
 
-While we did not identify any statistical associations between individual species
-and cognitive performance in our youngest subset (0 to 6 months),
-it may be because brain developmental trajectories begin \textit{in utero},
-and microbial influences (that can only begin at birth)
-take additional time to manifest.
-Future studies should attempt to link early-life microbial exposures
-to future cognitive performance or brain development.
-
 We identified several species in the family Eggerthelaceae that were
@@
 targets for further mechanistic studies.
+
+While we did not identify any statistical associations between individual species
+and cognitive performance in our youngest subset (0 to 6 months),
+it may be because brain developmental trajectories begin \textit{in utero},
+and microbial influences (that can only begin at birth)
+take additional time to manifest.
+A prediction experiment aligned to that hypothesis,
+targeting future cognitive assessment scores using metagenomes acquired prior to 1 year of age,
+revealed a list of most predictive taxa for that task.
+The top-ranked species on that list include several 
+taxa that were important for predicting concurrent cognitive assessment scores after 18 months, notably
+pathobiontic \textit{Enterobacteriaceae} such as
+\textit{Escherichia} and \textit{Klebsiella}. 
+Representatives of this genus are known to be early and long-term persistent
+colonizers of infant intestine that can, in some settings,
+promote inflammatory response \cite{popeEnterobacEarlyGut2019}.
+Parallel to those species, several tradinionally health-promoting and inflammation-mudulating representatives
+of the \textit{Bifidobacterium} genus,
+from the early HMO-metabolizer \textit{B. longum} to the
+later complex carbohydrate relying \textit{B. pseudocatenulatum} \cite{chungBifidoPsudocatenulatum2021},
+commonly associated with weaning to solid food in early age,
+have also shown high relative importance.
+The involvement of most important taxa in different aspects of inflammatory response
+resonate with previous findings that associate chronic inflammation with
+changes to neural responses in some types of stimula, notably visual\cite{xieChronicInfNeuro2019}.
 

### 2023-10-06 manuscript/Table2.jl
include all samples

@@ -7,11 +7,11 @@ using DataFrames.PrettyTables
 isdir(tablefiles()) || mkpath(tablefiles())
 
 fsea = let
-    fsea_all = CSV.read(scratchfiles("figure2", "fsea_all_00to120.csv"), DataFrame)
+    fsea_all = CSV.read(tablefiles("figure2", "fsea_all_00to120.csv"), DataFrame)
     fsea_all.group .= "00to120"
-    fsea_u6 = CSV.read(scratchfiles("figure2", "fsea_all_00to06.csv"), DataFrame)
+    fsea_u6 = CSV.read(tablefiles("figure2", "fsea_all_00to06.csv"), DataFrame)
     fsea_u6.group .= "00to06"
-    fsea_o18 = CSV.read(scratchfiles("figure2", "fsea_all_18to120.csv"), DataFrame)
+    fsea_o18 = CSV.read(tablefiles("figure2", "fsea_all_18to120.csv"), DataFrame)
     fsea_o18.group .= "18to120"
     vcat(fsea_all, fsea_u6, fsea_o18)
 end
@@ -40,16 +40,18 @@ end
 
 fseaout = DataFrame(geneset=unique(fsea.geneset))
 for g in levels(fsea.group)
-    df = select(subset(fsea, "group"=> ByRow(==(g))), "enrichment", "qvalue")
+    df = select(leftjoin(select(fseaout, "geneset"), subset(fsea, "group"=> ByRow(==(g))); on="geneset"), "enrichment", "qvalue")
     rename!(df, ["$(g)_enrichment", "$(g)_qvalue"])
     fseaout = hcat(fseaout, df)
 end
-
+for col in names(fseaout, Not("geneset"))
+    fseaout[!, col] = coalesce.(fseaout[!, col], NaN)
+end
 
 CSV.write(tablefiles("Table2.csv"), fseaout)
 formatter = (v, i, j) -> v isa AbstractFloat ? round(v, digits = 3) : v
-h1 = LatexHighlighter((data, i, j) -> j in (3,5,7) && data[i,j] < 0.2 && data[i,j-1] < 0, ["color{red}","textbf"])
-h2 = LatexHighlighter((data, i, j) -> j in (3,5,7) && data[i,j] < 0.2 && data[i,j-1] > 0, ["color{blue}","textbf"])
[…]

### 2023-10-06 manuscript/main.tex
fixed previous reference to S salivarius
cog-brain-taxa discussion paragraph

@@
 \cite{hoogmanSubcorticalBrainVolume2017}.
+
+Our ability to probe individual cognitive subscales provided
+results that also endorse previous findings and lay insight to
+potential gut-microbial connections between cognition and neuroanatomy.
+For example, a longitudinal neurocognitive development study pointed at bilateral
+associations between the thalamus and the EL component of the Mullen scales.
+Linking to this observation, we found previously-mentioned taxa such as 
+\textit{B. wexlerae} and \textit{A. hadrus} to be among the top 15\% important predictors
+on both left and right thalamus and the EL subscale.
+Another instance of those findings is a previously reported association between
+the central opercular cortex (COC) and Gross Motor development. In this case, we identified that
+butyrate-producing \textit{Fusicatenibacter saccharivorans} is among the 15\% most important predictors
+for both the GM subscale and three bilateral components of the COC
+(left and right pars orbitalis, pars opercularis and pars triangularis).
+Previous results also link to relationships between the caudate and the cerebellum,
+and the Visual Reception score.
+\textit{S. salivarius} was consistently found among the top 20\% predictors for this subscale, along with
+both sides of the thalamus and cerebellum white matter, along with cerebellar vermal lobules I-X.
 

### 2023-10-06 manuscript/main.tex
added citations

@@
 For example, a longitudinal neurocognitive development study pointed at bilateral
-associations between the thalamus and the EL component of the Mullen scales.
+associations between the thalamus and the EL component of the Mullen scales\cite{chenLongitudinalBrainMullen2023}.
 Linking to this observation, we found previously-mentioned taxa such as 
@@
 Another instance of those findings is a previously reported association between
-the central opercular cortex (COC) and Gross Motor development. In this case, we identified that
+the central opercular cortex (COC) and Gross Motor development\cite{jiangBrainNetworkCognition2023}. In this case, we identified that
 butyrate-producing \textit{Fusicatenibacter saccharivorans} is among the 15\% most important predictors
@@
 Previous results also link to relationships between the caudate and the cerebellum,
-and the Visual Reception score.
+and the Visual Reception score\cite{deoniMyelinationMullenPreprint2022}.
 \textit{S. salivarius} was consistently found among the top 20\% predictors for this subscale, along with

### 2023-10-06 manuscript/main.tex
flag section for review/rewrite with senior lab members

@@
 
+--- Start of Guilherme flag for review of section:
 We identified several species in the family Eggerthelaceae that were
@@
 targets for further mechanistic studies.
+--- End of Guilherme flag for review of section
 

### 2023-10-07 manuscript/main.tex

@@
 take additional time to manifest.
-A prediction experiment aligned to that hypothesis,
-targeting future cognitive assessment scores using metagenomes acquired prior to 1 year of age,
-revealed a list of most predictive taxa for that task.
-The top-ranked species on that list include several 
+In line with this hypothesis, RF models that were able to learn associations
+between early-life microbiomes and future cognitive function scores (Figure 3).
+The top-ranked species for these models included several 
 taxa that were important for predicting concurrent cognitive assessment scores after 18 months, notably
@@
 \textit{Escherichia} and \textit{Klebsiella}. 
-Representatives of this genus are known to be early and long-term persistent
+Representatives of these genera are known to be early and long-term persistent
 colonizers of infant intestine that can, in some settings,
 promote inflammatory response \cite{popeEnterobacEarlyGut2019}.
-Parallel to those species, several tradinionally health-promoting and inflammation-mudulating representatives
+By contrast, several traditionally inflammation-reducing representatives
 of the \textit{Bifidobacterium} genus,
@@
 commonly associated with weaning to solid food in early age,
-have also shown high relative importance.
+also showed high relative importance.
 The involvement of most important taxa in different aspects of inflammatory response
 resonate with previous findings that associate chronic inflammation with
-changes to neural responses in some types of stimula, notably visual\cite{xieChronicInfNeuro2019}.
+changes to neural responses in some types of stimulae, notably visual\cite{xieChronicInfNeuro2019}.
 
@@
 Our ability to probe individual cognitive subscales provided
-results that also endorse previous findings and lay insight to
+results that also endorse previous findings and provide insight to
 potential gut-microbial connections between cognition and neuroanatomy.
@@
 associations between the thalamus and the EL component of the Mullen scales\cite{chenLongitudinalBrainMullen2023}.
-Linking to this observation, we found previously-mentioned taxa such as 
+In line with this finding, we found that taxa such as 
 \textit{B. wexlerae} and \textit{A. hadrus} to be among the top 15\% important predictors

### 2023-10-07 manuscript/main.tex

@@
 We investigated the relationship of the microbiome with the neuroanatomy
-and cognitive function of 361 healthy children, demonstrating that
+and cognitive function of 381 healthy children, demonstrating that
 differences in gut microbial taxa and gene functions are associated with
@@
 We investigated the co-development of the brain and microbiome in
-381 healthy and neurotypically-developing children (163 female) between 82 days and 10 years of age
+381 healthy and neurotypically-developing children (172 female) between 40 days and 10 years of age
 (Table 1,  Figure 1A, Figure S1)
@@
       N subjects                  &                          & 381 & 82 & 277 & 104 \\  \hline
-    \multirow{3}{*}{Samples}      & 1                        & 314 (82.4\%) &  &  &  \\ \cline{2-6}
-                                  & 2                        & 55 (14.4\%) &  &  &  \\ \cline{2-6}
-                                  & \textgreater 2           & 12 (3.1\%) &  &  &  \\ \hline
+    \multirow{3}{*}{Samples}      & 1                        & 214 (56.2\%) &  &  &  \\ \cline{2-6}
+                                  & 2                        & 90 (23.6\%) &  &  &  \\ \cline{2-6}
+                                  & \textgreater 2           & 77 (20.2\%) &  &  &  \\ \hline
     \multirow{3}{*}{Age (months)} & min                      & 1.3 & 2.89 & 18.03 & 1.3 \\ \cline{2-6}
@@
     \begin{tabular}{|r|r|r|r|r|r|r|}
-      \hline\hline
-      \textbf{Age group} & \textbf{under 6m} & \textbf{under 6m} & \textbf{over 18m} & \textbf{over 18m} \\
+        \hline
+        \textbf{Age group}& \multicolumn{2}{c|}{\textbf{0 to 6m}} & \multicolumn{2}{c|}{\textbf{18 to 120m}} \\\hline
         \texttt{geneset} & \texttt{E.S.} & \texttt{q value} & \texttt{E.S.} & \texttt{q value} \\\hline
-      Acetate syn. I & -0.21 & \color{red}{\textbf{0.068}} & 0.123 & 0.403 \\
-      Acetate syn. II & -0.322 & \color{red}{\textbf{0.046}} & 0.113 & 0.866 \\
-      Acetate syn. III & -0.291 & \color{red}{\textbf{0.089}} & 0.109 & 0.866 \\
-      Glutamate syn. II & -0.152 & 0.882 & 0.293 & \color{blue}{\textbf{0.001}} \\
-      Propionate deg. I & -0.639 & \color{red}{\textbf{0.006}} & -0.495 & 0.446 \\\hline\hline
+        Acetate syn. I & -0.165 & 0.284 & 0.192 & \color{red}{\textbf{0.145}} \\
+        Acetate syn. II & -0.491 & \color{blue}{\textbf{0.059}} & 0.248 & \color{red}{\textbf{0.15}} \\
+        Acetate syn. III & -0.409 & \color{blue}{\textbf{0.133}} & 0.215 & 0.209 \\
+        Butyrate syn. II & -0.357 & 0.284 & 0.414 & \color{red}{\textbf{0.178}} \\
+        Glutamate syn. II & 0.343 & 0.284 & 0.355 & \color{red}{\textbf{0.023}} \\
+        Propionate deg. I & -0.659 & \color{blue}{\textbf{0.029}} & 0.154 & \color{red}{\textbf{0.178}} \\\hline
     \end{tabular}

### 2023-10-07 manuscript/main.tex
fix some FSEA

@@
 scores in children under 6m (Table 2; propionate degradation I,
-enrichment score (E.S.) = -0.64, corrected p-value (q)
-= 0.006), as were genes for synthesizing the 2 carbon SCFA acetate
-(acetate synthesis I, E.S. = -0.210, q = 0.07; acetate synthesis
-II, E.S. = -0.342, q = 0.049; acetate synthesis III, E.S. = -0.291, q = 0.09).
-In children over 18 months old, synthesis of the amino acid
-glutamate was significantly enriched in children with
-higher cognitive function scores (glutamate synthesis II, E.S. = 0.293, 
-q = 0.0008). Taken together, these results suggest that
+enrichment score (E.S.) = -0.66, corrected p-value (q)
+= 0.029), as were genes for synthesizing the 2 carbon SCFA acetate
+(acetate synthesis II, E.S. = -0.409, q = 0.059; acetate synthesis
+III, E.S. = -0.409, q = 0.133).
+Interstingly, in children over 18m,
+these metabolic associations were reversed, with both acetate synthesis
+and propionate degredation being enriched in children
+with higher cognitive scores
+(acetate synthesis I, E.S. = 0.192, q = 0.145;
+acetate synthesis II, E.S. = 0.248, q = 0.15;
+propionate degredation I, E.S. = 0.154, q = 0.178),
+as was synthesis of the 4-carbon SCFA butyrate
+(butyrate synthesis II, E.S. = 0.414, q = 0.178).
+Synthesis of the amino acid
+glutamate was also significantly enriched in children with
+higher cognitive function scores in this age group (glutamate synthesis II, E.S. = 0.154, 
+q = 0.178). Taken together, these results suggest that
 microbial metabolic activity, particularly the metabolism (synthesis and

### 2023-10-10 manuscript/main.tex
not possible -> challenging

@@
 of exposure, but drawing sufficient blood from children,
-especially very young children, is typically not permitted.
+especially very young children, is challenging.
 Further, strain-level analysis linking specific gene content in species

### 2023-10-10 manuscript/main.tex
begin addressing Vanja review

@@
 Many of these bacteria have specialized metabolic capabilities
-for digesting human breast milk, such as \emph{Bifidobacterium longum}
+for digesting human milk, such as \emph{Bifidobacterium longum}
 subsp. \textit{infantis} and \emph{Bacteroides fragilis}
@@
     \end{tabular}
-    \caption{\label{tab:demo}Subject demographics of the ECHO RESONANCE cohort.}
+    \caption{\label{tab:demo}Subject demographics of the ECHO RESONANCE cohort.
+            "All" refers to the entire cohort.
+            "Under 6 months", "Over 18 months", and "Future" refer specifically
+            to samples that were used included in subgroup analyses.
+    }
 \end{table}

### 2023-10-10 manuscript/main.tex
through p9

@@
             "Under 6 months", "Over 18 months", and "Future" refer specifically
-            to samples that were used included in subgroup analyses.
+            to samples that were used included in the respective subgroup analyses.
     }
@@
 
-Each of the cognitive assessments (MSEL, WPPSI, WISC)
+Cognitive assessments (MSEL, WPPSI, WISC)
 are composites of multiple subscales that assess different cognitive domains.
-While the composite scores are broadly analagous accross ages,
+While the composite scores are broadly analagous accross ages and test modality,
 the subscales assessed in each test are different.
-In order to assess whether microbial taxa were associated with
+To assess whether microbial taxa were associated with
 the development of specific cognitive domains,
@@
 with a concurrent MSEL score (Figure 2C).
-Several of the taxa that were associated with composite scores (Figure 2B)
+Several taxa associated with composite scores (Figure 2B)
 were also associated with one or more subscores, including
@@
 \textit{A. obesi} as important % A. obesi was mentioned before, ~ lines 120, 390
-(q \textless 0.2 for LMs, cumulative importance \textless 60\% for RFs)
-but RF models did not identify \textit{Eggerthella} sp. CAG 298
-nor \textit{B. intestinihominis}, and \textit{E. ramulus}
+(q \textless 0.2 for LMs, cumulative importance \textless 60\% for RFs).
+Other taxa that were significant in LMs were not identified by RF models
+including \textit{Eggerthella} sp. CAG 298,
+and \textit{B. intestinihominis}. \textit{E. ramulus}
 was just outside our cutoff (64\% cumulative importance).
@@
 
-\subsection*{Gut microbial profiles in the first year of life predictict future cogntitive function}
+\subsection*{Gut microbial profiles in the first year of life predict future cogntitive function}
 

### 2023-10-10 manuscript/main.tex
start p10

@@
 Given that the development of different domains of cognitive function
-(eg gross motor development vs visual perception) may develop at different times
+(e.g., gross motor development vs visual perception) may develop at different times
 and be affected by different molecular signals,
 we sought to determine whether the associations of microbial taxa
-with cognitive function are general across all cognitive subdomains,
-or specific.
+with cognitive function are general across all cognitive subdomains.
 To accomplish this, we assessed individual subscales of the  
-Mullen Scales of Early Learning (MSEL - \textit{see Methods}). 
-Taxonomic profiles coupled with demographics exhibited predictive potential towards three probed subscales (Figure 4A), 
+Mullen Scales of Early Learning (MSEL, birth to 36 months - \textit{see Methods}). 
+Taxonomic profiles coupled with demographics exhibited predictive potential towards three
+of five probed subscales (Figure 4A), 
 namely Expressive Language (EL; test-set RMSE XXX, test-set correlation YYYY), 

### 2023-10-10 manuscript/main.tex
add highlights

@@
 of five probed subscales (Figure 4A), 
-namely Expressive Language (EL; test-set RMSE XXX, test-set correlation YYYY), 
-Gross Motor (GM; test-set RMSE XXX, test-set correlation YYYY) and 
-Visual Reception (VR; test-set RMSE XXX, test-set correlation YYYY).
+namely Expressive Language (EL; test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}), 
+Gross Motor (GM; test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}) and 
+Visual Reception (VR; test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}).
 Between the subscales, not only the predictability (measured by the test-set figures of merit) is different,

### 2023-10-10 manuscript/main.tex

@@
 brain region models (R = 0.288), as compared to the right which
-had a negligible test-set correlation (R = 0.005),
-indicating that generalizable models were not identified.
+had a negligible test-set correlation (R = 0.005).
 Feature importance for models of the left accumbens area were dominated

### 2023-10-10 manuscript/Figure2-calculations_beforediscard.jl
updated the hashname generator to differentiate between models

@@ -1,0 +1,355 @@
+using Resonance
+using CairoMakie
+using Statistics
+using HypothesisTests
+using MultipleTesting
+# using KernelDensity
+using MultivariateStats
+using Distributions
+using CategoricalArrays
+using ThreadsX
+using ColorSchemes
+using GLM
+using Dates
+using JLD2
+using MultipleTesting
+
+const adjust = MultipleTesting.adjust
+
+isdir(tablefiles("figure2")) || mkpath(tablefiles("figure2"))
+
+## Data Loading
+
+mdata = Resonance.load(Metadata())
+seqs = subset(mdata, "sample"=> ByRow(!ismissing)) 
+
+seqs.sample = [s for s in seqs.sample]
+seqs.edfloat = map(x-> ismissing(x) ? missing : Float64(levelcode(x)), seqs.education)
+taxa = Resonance.load(TaxonomicProfiles(); timepoint_metadata = seqs) # this can take a bit
+species = filter(f-> taxrank(f) == :species, taxa)
+relativeabundance!(species)
+unirefs = Resonance.load(UnirefProfiles(); timepoint_metadata = seqs) # this can take a bit
+ecs = Resonance.load(ECProfiles(); timepoint_metadata = seqs)
+kos = Resonance.load(KOProfiles(); timepoint_metadata = seqs)
+
+# metabolites = Resonance.load(MetabolicProfiles(); timepoint_metadata = mdata)
+brain = Resonance.load(Neuroimaging(), timepoint_metadata = seqs, samplefield="sample")
+
+
+# @assert all(samplenames(species) .== samplenames(unirefs))
+@assert all(samplenames(species) .== samplenames(ecs) .== samplenames(kos) .== samplenames(unirefs))
+
+brain_roi = [
+    "right-lateral-occipital",
+    "left-lateral-occipital",
+    "right-inferior-parietal",
+    "left-inferior-parietal",
+    "right-middle-temporal",
+    "left-middle-temporal",
+    "right-cerebellum-white-matter",
+    "left-cerebellum-white-matter",
+    "right-thalamus-proper",
+    "left-thalamus-proper",
+    "right-insula",
+    "left-insula",
+    "Gray-matter",
+    "White-matter"
+]
+brainmeta = let
[…]

### 2023-10-10 manuscript/Figure3-currentvisualreception-calculations.jl

@@ -53,12 +53,12 @@ for oldname in names(mdata_ecs_df)[22:end]
     rename!(mdata_ecs_df, oldname => randstring(['A':'Z'; 'a':'z'], 12))
 end
 newnames = names(mdata_ecs_df)[22:end]
-open(scratchfiles("longnames.txt"), "w") do io
+open(scratchfiles("concurrentVR_longnames.txt"), "w") do io
     for i in oldnames
         println(io, i)
     end
 end
-open(scratchfiles("hashnames.txt"), "w") do io
+open(scratchfiles("concurrentVR_hashnames.txt"), "w") do io
     for i in newnames
         println(io, i)
     end

### 2023-10-10 manuscript/Figure3-futurecogscore-calculations.jl

@@ -55,10 +55,6 @@ end
 # Loading Data
 #####
 
-## Printing quantiles for paper results
-@show quantile(future_mdata.future_ageMonths .- future_mdata.present_ageMonths, [0.05, 0.5, 0.95])
-@show quantile(mdata_taxa_df.future_ageMonths .- mdata_taxa_df.present_ageMonths, [0.05, 0.5, 0.95])
-
 ## 1. Metadata
 mdata = Resonance.load(Metadata())
 future_mdata = vcat( [ build_future_row(df) for df in groupby(mdata[ mdata.filter_future_omni .| mdata.filter_future_cog, : ], :subject) ]... )
@@ -70,6 +66,10 @@ seqs.edfloat = map(x-> ismissing(x) ? missing : Float64(levelcode(x)), seqs.educ
 taxa = Resonance.load(TaxonomicProfiles(); timepoint_metadata = seqs) # this can take a bit
 species = filter(f-> taxrank(f) == :species, taxa)
 mdata_taxa_df = sort(Resonance.comm2wide(species), [ :subject ]);
+
+## Printing quantiles for paper results
+@show quantile(future_mdata.future_ageMonths .- future_mdata.present_ageMonths, [0.05, 0.5, 0.95])
+@show quantile(mdata_taxa_df.future_ageMonths .- mdata_taxa_df.present_ageMonths, [0.05, 0.5, 0.95])
 
 ## 3. Functional Profiles
 ecs = Resonance.load(ECProfiles(); timepoint_metadata = seqs)

### 2023-10-10 manuscript/TableS1-5.jl
moved Suppl table generation to unified code

@@ -10,10 +10,28 @@ using MLJ
 using DataFrames.PrettyTables
 
 isdir(tablefiles()) || mkpath(tablefiles())
-
 RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
 
 ## 1. Loading the model files
+
+### 1.1. Concurrent composite scores
+regression_currentCogScores_00to06mo_onlydemo = JLD2.load(modelfiles("regression_currentCogScores_00to06mo_onlydemo.jld"))["regression_currentCogScores_00to06mo_onlydemo"]
+regression_currentCogScores_00to06mo_onlytaxa = JLD2.load(modelfiles("regression_currentCogScores_00to06mo_onlytaxa.jld"))["regression_currentCogScores_00to06mo_onlytaxa"]
+regression_currentCogScores_00to06mo_demoplustaxa = JLD2.load(modelfiles("regression_currentCogScores_00to06mo_demoplustaxa.jld"))["regression_currentCogScores_00to06mo_demoplustaxa"]
+regression_currentCogScores_00to06mo_onlyecs = JLD2.load(modelfiles("regression_currentCogScores_00to06mo_onlyecs.jld"))["regression_currentCogScores_00to06mo_onlyecs"]
+regression_currentCogScores_00to06mo_demoplusecs = JLD2.load(modelfiles("regression_currentCogScores_00to06mo_demoplusecs.jld"))["regression_currentCogScores_00to06mo_demoplusecs"]
+regression_currentCogScores_18to120mo_onlydemo = JLD2.load(modelfiles("regression_currentCogScores_18to120mo_onlydemo.jld"))["regression_currentCogScores_18to120mo_onlydemo"]
+regression_currentCogScores_18to120mo_onlytaxa = JLD2.load(modelfiles("regression_currentCogScores_18to120mo_onlytaxa.jld"))["regression_currentCogScores_18to120mo_onlytaxa"]
+regression_currentCogScores_18to120mo_demoplustaxa = JLD2.load(modelfiles("regression_currentCogScores_18to120mo_demoplustaxa.jld"))["regression_currentCogScores_18to120mo_demoplustaxa"]
+regression_currentCogScores_18to120mo_onlyecs = JLD2.load(modelfiles("regression_currentCogScores_18to120mo_onlyecs.jld"))["regression_currentCogScores_18to120mo_onlyecs"]
[…]

### 2023-10-10 manuscript/TableS1.jl

@@ -1,105 +1,0 @@
-#####
-# Supplementary Table S1A-B. List of most important taxa for Random Forest models on the (A) 00-06 and (B) 18-120 months age bracket, MICROBIOME ONLY, ordered by relative weighted importance rank
-#####
-using Resonance
-using CategoricalArrays
-using Statistics
-using Chain
-using JLD2
-using MLJ
-using DataFrames.PrettyTables
-
-isdir(tablefiles()) || mkpath(tablefiles())
-
-RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
-
-# function weighted_hpshapvalues(m, hp = 1; change_hashnames=false, hashnamestable::Union{Nothing, DataFrame} = nothing) ### REMOVED
-#     merits = m.merits
-#     importances = m.shapley
-#     fitnesses = calculate_fitness(merits.Train_Cor, merits.Test_Cor)
-#     fitness_subset_idx = findall(merits.Hyperpar_Idx .== hp)
-#     # Option 1: divide by amount of positive-only fitnesses
-#     # mean_importances = map(x -> sum(x .* fitnesses[fitness_subset_idx])/sum(fitnesses[fitness_subset_idx] .> 0.0), collect(eachrow(Matrix(importances[:, fitness_subset_idx .+ 1]))))
-#     # Option 2: divide importances by total number of models
-#     mean_importances = map(x -> sum(x .* fitnesses[fitness_subset_idx])/length(fitness_subset_idx), collect(eachrow(Matrix(importances[:, fitness_subset_idx .+ 1]))))
-#     mean_importances_df = sort( DataFrame(:variable => string.(importances[:,1]), :weightedImportance => mean_importances), :weightedImportance; rev=true)
-#     if change_hashnames
-#         newnames = leftjoin(mean_importances_df, hashnamestable, on = :variable => :hashname).longname
-#         mean_importances_df.variable = newnames
-#     end
-#     return mean_importances_df
-# end
-# function singlemodel_shapley_suppltable(rf_model)
-#     rf_model_shapvalues = weighted_hpshapvalues(rf_model; change_hashnames=false)
-#     return rf_model_shapvalues
-# end  ### REMOVED
-
-## 1. Loading the model files
[…]

### 2023-10-10 manuscript/main.tex
updates to results and discussion

@@
 R\textsuperscript{2} = 0.0263, q \textless{} 0.001), and very strongly
-associated with variation in neuroimaging profiles (R\textsuperscript{2} = 0.258, q
+associated with variation in neuroimaging (MRI) profiles (R\textsuperscript{2} = 0.258, q
 \textless{} 0.001).
@@
         \textbf{Age group}& \multicolumn{2}{c|}{\textbf{0 to 6m}} & \multicolumn{2}{c|}{\textbf{18 to 120m}} \\\hline
-        \texttt{geneset} & \texttt{E.S.} & \texttt{q value} & \texttt{E.S.} & \texttt{q value} \\\hline
+        geneset & E.S. & q value & E.S. & q value \\\hline
         Acetate syn. I & -0.165 & 0.284 & 0.192 & \color{red}{\textbf{0.145}} \\
@@
 
-\subsubsection*{Gut microbial taxonomic and functional profiles predict cognitive function}
+\subsubsection*{Gut microbial taxonomic and functional profiles predict concurrently measured cognitive function}
 
@@
     \end{tabular}
-    \caption{\label{tab:rfbench}Benchmark metrics for the cognitive assessment score
+    \caption{\label{tab:rfbench2}Benchmark metrics for the cognitive assessment score
     future prediction models. Confidence intervals (C.I.) are calculated from the
@@
 and cognitive function scores in children over 18 months of age (mean
-test-set correlation 0.429, mean RMSE 18.25).[]
+test-set correlation 0.429, mean RMSE 18.25).
 Species that were important in RF overlapped with those that were
@@
 
-\subsection*{Gut microbial profiles in the first year of life predict future cogntitive function}
+\subsubsection*{Gut microbial profiles in the first year of life predict future cogntitive function}
 
@@
 of five probed subscales (Figure 4A), 
-namely Expressive Language (EL; test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}), 
-Gross Motor (GM; test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}) and 
-Visual Reception (VR; test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}).
+namely Expressive Language ( test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}), 
+Gross Motor (test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}) and 
+Visual Reception (test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}).
 Between the subscales, not only the predictability (measured by the test-set figures of merit) is different,
@@
 \textit{Clostridium innocuum}, \textit{B. vulgatus} 
-stand out on models targeting VR.
+stand out on models targeting visual reception.
 
@@
 
-Both \emph{A. equolofaciens} and \emph{A. celatus}, two closely-related
-equol-producing species, are examples of the second contribution
+\emph{A. celatus}, an equol-producing species, is an examples of the second contribution
 pattern, where a taxon only contributes significantly to a very small number of segments. 
-Each of these species were highly important in predicting the relative volume of
-the right anterior cingulate (respectively, 3.0\% and 2.6\% importances).
-Another example of this pattern is \emph{R. gnavus}, which was the only
-species with significant negative association with cognitive function in linear models.
-It is heavily associated with the left pars opercularis (2.7\% importance). 
+This species was highly important in predicting the relative volume of
+the right anterior cingulate (respectively, \hl{3.0\% and 2.6\%} importances).
+Another example of this pattern is \emph{Coprococcus eutactus}
+which is heavily associated with the left precentral area (\hl{2.7\%} importance). 
 Finally, \emph{C. comes} has an importance
@@
 clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (4.0\% importance);
+is the highest for a microbe in RF models (\hl{4.0\%} importance);
 only age had higher relative importance in any model. At the same time,
@@
 posterior cingulate such as the pars opercularis (importances,
-Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
-precentral (2.3\% importance) and paracentral lobes (importances, Left = 2.6\%, Right = 1.9\%).
+\hl{Left = 2.3\%, Right = 2.8\%}), along with upper regions like the left
+precentral (2.3\% importance) and paracentral lobes (importances, \hl{Left = 2.6\%, Right = 1.9\%}).
 
@@
 
---- Start of Guilherme flag for review of section:
 We identified several species in the family Eggerthelaceae that were
-associated with cognitive function, including \emph{G.pamelaeae, A. celatus}
-(formerly regarded a subspecies of \emph{A. equolofaciens}
+associated with cognitive function, including \emph{A. celatus}
+(a subspecies of \emph{A. equolofaciens}
 \cite{takahashiCompleteGenomeSequence2021}),
@@
 found in soybeans \cite{wangEnantioselectiveSynthesisSEquol2005}.
-This is particularly intriguing, since both \emph{A. equolofaciens} and \emph{A. celatus}
-were also the most important features in RF models of the
+This is particularly intriguing, since \emph{A. celatus}
+was also the most important features in RF models of the
 right anterior cingulate, which has been linked to social cognition
@@
 in this brain region \cite{xiaoEstrogenAnteriorCingulate2013}.
-\emph{G. pamelaeae} also has a unique metabolism that may impact the brain,
+Another Eggerthelaceae, \emph{Gordonibacter pamelaeae},
+also has a unique metabolism that may impact the brain,
 as it can metabolize the polyphenol ellagic acid
@@
 \cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
+Though \textit{G. pamelaeae} was not identified
+as an important feature in RF models,
+it was also relatively low prevalence in our dataset,
+occuring in only ~5\% of samples at greater than 2\% abundance
+(in contrast to \textit{A. celatus}, which was in 32\% of samples)
 While we could not find published evidence of neurological roles for \emph{E. lenta},
@@
 targets for further mechanistic studies.
---- End of Guilherme flag for review of section
-
-While we did not identify any statistical associations between individual species
-and cognitive performance in our youngest subset (0 to 6 months),
-it may be because brain developmental trajectories begin \textit{in utero},
-and microbial influences (that can only begin at birth)
+
+We did not identify any statistical associations between individual species
+and cognitive performance in our youngest subset (0 to 6 months).
+We hypothesize that this may be because brain developmental trajectories
+begin \textit{in utero}, and microbial influences (that can only begin at birth)
 take additional time to manifest.
@@
 promote inflammatory response \cite{popeEnterobacEarlyGut2019}.
-By contrast, several traditionally inflammation-reducing representatives
+By contrast, several inflammation-reducing representatives
 of the \textit{Bifidobacterium} genus,
@@
 Our ability to probe individual cognitive subscales provided
-results that also endorse previous findings and provide insight to
+results that also align with previous findings and provide insight to
 potential gut-microbial connections between cognition and neuroanatomy.
 For example, a longitudinal neurocognitive development study pointed at bilateral
-associations between the thalamus and the EL component of the Mullen scales\cite{chenLongitudinalBrainMullen2023}.
+associations between the thalamus and the expressive language
+component of the Mullen scales \cite{chenLongitudinalBrainMullen2023}.
 In line with this finding, we found that taxa such as 
@@
 on both left and right thalamus and the EL subscale.
-Another instance of those findings is a previously reported association between
-the central opercular cortex (COC) and Gross Motor development\cite{jiangBrainNetworkCognition2023}. In this case, we identified that
+Another example of those findings is a previously reported association between
+the central opercular cortex (COC) and gross motor development \cite{jiangBrainNetworkCognition2023}. In this case, we identified that
 butyrate-producing \textit{Fusicatenibacter saccharivorans} is among the 15\% most important predictors
-for both the GM subscale and three bilateral components of the COC
+for both the gross motor subscale and three bilateral components of the COC
 (left and right pars orbitalis, pars opercularis and pars triangularis).
 Previous results also link to relationships between the caudate and the cerebellum,
-and the Visual Reception score\cite{deoniMyelinationMullenPreprint2022}.
+and the Visual Reception score \cite{deoniMyelinationMullenPreprint2022}.
 \textit{S. salivarius} was consistently found among the top 20\% predictors for this subscale, along with
@@
 These included preterm birth (\textless 37 weeks gestation) low birth weight (\textless 1500g),
-in utero exposure to alcohol, cigarette or illicit substance exposure;
+\textit{in utero} exposure to alcohol, cigarette or illicit substance exposure;
 abnormalities on fetal ultrasound; non-singleton or complicated pregnancy,
@@
 The cognitive assessments were collected as either composite scores
-(for the experiments in Figure \textbf{3}) or, when available, subscores on the
-Mullen Scales of Early Learning ( Figure \textbf{4 A-B}: Expressive Language, Gross Motor and Visual Reception).
+(for the experiments in Figure 3) or, when available, subscores on the
+Mullen Scales of Early Learning (Figure 4 A-B: Expressive Language, Gross Motor and Visual Reception).
 For the concurrent prediction experiments, when multiple viable longitudinal samples were present,

### 2023-10-10 manuscript/TableS1-5.jl
renaming supp table notebooks

@@ -1,5 +1,5 @@
 #####
-# Supplementary Table S1A-B. List of most important taxa for Random Forest models on the (A) 00-06 and (B) 18-120 months age bracket, MICROBIOME ONLY, ordered by relative weighted importance rank
+# Supplementary Table S1 thourgh S5. List of most important taxa for Random Forest models, MICROBIOME ONLY, ordered by relative weighted importance rank
 #####
 using Resonance
 using CategoricalArrays

### 2023-10-10 manuscript/TableS6.jl

@@ -1,5 +1,5 @@
 #####
-# Supplementary Table 2. Summary statistics for the neuroanatomy prediction benchmarks
+# Supplementary Table 6. Summary statistics for the neuroanatomy prediction benchmarks
 #####
 using Resonance
 using CategoricalArrays
@@ -25,7 +25,7 @@ end
 tableS2 = subset(tableS2, :variable => x -> x .∈ Ref(interesting_segments) )
 tableS2.variable = uppercasefirst.(map(x -> replace(x, "-"=>" "), tableS2.variable))
 
-CSV.write(joinpath("manuscript", "assets", "TableS2.csv"), tableS2)
+CSV.write(joinpath("manuscript", "assets", "TableS6.csv"), tableS2)
 
 pretty_table(tableS2;
     header = [ "Segment", "Mean absolute proportional error (MAPE)", "Correlation coefficient (R)" ],

### 2023-10-10 manuscript/main.tex
fix t3 header

@@
       \hline\hline
-      \textbf{Subject Ages (months)} & \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
+      \textbf{Subject Ages (months)} & \textbf{Microbial feature} & \textbf{Demo.} &  \textbf{Test set correlation (± C.I.)} & \textbf{Test set RMSE (± C.I.)} \\\hline
       \multirow{5}{*}{0 to 6} & -    & + & 13.37 ± 0.05 & -0.16 ± 0.01 \\ \cline{2-5}

### 2023-10-10 manuscript/main.tex

@@
     \caption{
-        Microbial feature importance in predicting brain volumes in children over 18 months of age.
-        (\textbf{A}) Average test-set correlations for prediction of MRI segmentation
+        Microbial feature importance in predicting MSEL subscores and brain volumes in children over 18 months of age.
+        (\textbf{A}) Test set correlations for RF models predicting MSEL subscale scores,
+        with or without demographics included as predictors.
+        (\textbf{B}) Important taxa in models predicting MSEL subscale scores.
+        (\textbf{C}) Average test-set correlations for prediction of MRI segmentation
         data from microbiome and demographics on select segments after repeated
-        cross validation. (\textbf{B}) Heatmap of average individual relative taxa
+        cross validation. (\textbf{D}) Heatmap of average individual relative taxa
         importances on each brain segment. Importances are reported as
@@
         by HCA on a select list of species with high ranks on average importances,
-        and their respective highest-load segments (\textbf{C}).
+        and their respective highest-load segments (\textbf{E}).
     }

### 2023-10-10 manuscript/main.tex
flag importance percentage
added numbers to subscale merits

--- 2023-10-10 manuscript/main.tex
@@
 and \textit{B. intestinihominis}. \textit{E. ramulus}
-was just outside our cutoff (64\% cumulative importance).
+was just outside our cutoff (64\% cumulative importance). % Flag - 64%?
 RF models also identified an additional 48 species
@@
 of five probed subscales (Figure 4A), 
-namely Expressive Language ( test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}), 
-Gross Motor (test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}) and 
-Visual Reception (test-set RMSE \hl{XXX}, test-set correlation \hl{YYYY}).
+namely Expressive Language ( test-set RMSE \hl{12.05}, test-set correlation \hl{0.368}), 
+Gross Motor (test-set RMSE \hl{11.07}, test-set correlation \hl{0.137}) and 
+Visual Reception (test-set RMSE \hl{15.2}, test-set correlation \hl{0.333}).
 Between the subscales, not only the predictability (measured by the test-set figures of merit) is different,

### 2023-10-10 manuscript/main.tex
changed supp table reference
updated brain importance numbers

@@
     of test-set preditions are reported accross models for all brain regions.
-    Complete list of values can be found on Table S2.}
+    Complete list of values can be found on Table S6.}
     \end{center}
@@
 highest average relative importance along all reported 
-(mean = 1.7\%; min = 1.0\%; max = 3.8\%, Figure 4D-E).
+(mean = 1.4\%; min = 0.8\%; max = 3.6\%, Figure 4D-E).
 \emph{B. wexlerae}, which was important in RF models of cognitive function,
@@
 This species was highly important in predicting the relative volume of
-the right anterior cingulate (respectively, \hl{3.0\% and 2.6\%} importances).
+the right anterior cingulate (\hl{2.8\%} relative importance).
 Another example of this pattern is \emph{Coprococcus eutactus}
-which is heavily associated with the left precentral area (\hl{2.7\%} importance). 
+which is heavily associated with the left precentral area (\hl{3.5\%} importance). 
 Finally, \emph{C. comes} has an importance
@@
 clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (\hl{4.0\%} importance);
+is the highest for a microbe in RF models (\hl{3.7\%} importance);
 only age had higher relative importance in any model. At the same time,
@@
 posterior cingulate such as the pars opercularis (importances,
-\hl{Left = 2.3\%, Right = 2.8\%}), along with upper regions like the left
-precentral (2.3\% importance) and paracentral lobes (importances, \hl{Left = 2.6\%, Right = 1.9\%}).
+\hl{Left = 1.9\%, Right = 2.5\%}), along with upper regions like the left
+precentral (2.0\% importance) and paracentral lobes (importances, \hl{Left = 2.0\%, Right = 1.5\%}).
 

### 2023-10-10 manuscript/Table4.jl
include all segments, not just the important ones

@@ -24,7 +24,7 @@ mean_brain_merits = reduce(
     vcat,
     [ DataFrame(:variable => ordered_brain_segments_list[i], :Train_MAPE => mean(brain_models[ordered_brain_segments_list[i]].merits.Train_MAPE), :Test_MAPE => mean(brain_models[ordered_brain_segments_list[i]].merits.Test_MAPE), :Train_Cor => mean(brain_models[ordered_brain_segments_list[i]].merits.Train_Cor), :Test_Cor => mean(brain_models[ordered_brain_segments_list[i]].merits.Test_Cor)) for i in eachindex(ordered_brain_segments_list) ]
 )
-mean_brain_merits = subset(mean_brain_merits, :variable => x -> x .∈ Ref(interesting_segments) )
+# mean_brain_merits = subset(mean_brain_merits, :variable => x -> x .∈ Ref(interesting_segments) ) # Table should show all segments, including the "bad" ones
 
 ## 3. Building the actual table
 

### 2023-10-10 manuscript/TableS6.jl

@@ -22,7 +22,7 @@ tableS2 = mapreduce( vcat, eachindex(ordered_brain_segments_list)) do i
               :Test_MAPE => round(mean(brain_models[ordered_brain_segments_list[i]].merits.Test_MAPE); digits = 4), 
               :Test_Cor => round(mean(brain_models[ordered_brain_segments_list[i]].merits.Test_Cor); digits = 4))
 end
-tableS2 = subset(tableS2, :variable => x -> x .∈ Ref(interesting_segments) )
+# tableS2 = subset(tableS2, :variable => x -> x .∈ Ref(interesting_segments) ) # Table should show all segments, including the "bad" ones
 tableS2.variable = uppercasefirst.(map(x -> replace(x, "-"=>" "), tableS2.variable))
 
 CSV.write(joinpath("manuscript", "assets", "TableS6.csv"), tableS2)

### 2023-10-10 manuscript/main.tex
changed Table 4 to reflect all segments, not only the important ones

@@
       \textbf{Statistic} & \textbf{MAPE} & \textbf{R} \\\hline
-      Mean & 0.069 & 0.206 \\
-      Standard deviation & 0.015 & 0.099 \\
-      Maximum & 0.11 & 0.444 \\
-      75th percentile & 0.076 & 0.262 \\
-      50th percentile & 0.066 & 0.206 \\
-      25th percentile & 0.057 & 0.16 \\
-      Minimum & 0.046 & 0.005 \\\hline
+      Mean & 0.073 & 0.146 \\
+      Standard deviation & 0.029 & 0.131 \\
+      Maximum & 0.205 & 0.598 \\
+      75th percentile & 0.081 & 0.212 \\
+      50th percentile & 0.066 & 0.129 \\
+      25th percentile & 0.057 & 0.073 \\
+      Minimum & 0.036 & -0.133 \\\hline
     \end{tabular}

### 2023-10-11 manuscript/main.tex
remove highlights

@@
 of five probed subscales (Figure 4A), 
-namely Expressive Language ( test-set RMSE \hl{12.05}, test-set correlation \hl{0.368}), 
-Gross Motor (test-set RMSE \hl{11.07}, test-set correlation \hl{0.137}) and 
-Visual Reception (test-set RMSE \hl{15.2}, test-set correlation \hl{0.333}).
+namely Expressive Language ( test-set RMSE 12.05, test-set correlation 0.368), 
+Gross Motor (test-set RMSE 11.07, test-set correlation 0.137) and 
+Visual Reception (test-set RMSE 15.2, test-set correlation 0.333).
 Between the subscales, not only the predictability (measured by the test-set figures of merit) is different,
@@
 This species was highly important in predicting the relative volume of
-the right anterior cingulate (\hl{2.8\%} relative importance).
+the right anterior cingulate (2.8\% relative importance).
 Another example of this pattern is \emph{Coprococcus eutactus}
-which is heavily associated with the left precentral area (\hl{3.5\%} importance). 
+which is heavily associated with the left precentral area (3.5\% importance). 
 Finally, \emph{C. comes} has an importance
@@
 clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (\hl{3.7\%} importance);
+is the highest for a microbe in RF models (3.7\% importance);
 only age had higher relative importance in any model. At the same time,
@@
 posterior cingulate such as the pars opercularis (importances,
-\hl{Left = 1.9\%, Right = 2.5\%}), along with upper regions like the left
-precentral (2.0\% importance) and paracentral lobes (importances, \hl{Left = 2.0\%, Right = 1.5\%}).
+Left = 1.9\%, Right = 2.5\%), along with upper regions like the left
+precentral (2.0\% importance) and paracentral lobes (importances, Left = 2.0\%, Right = 1.5\%).
 

### 2023-10-11 manuscript/main.tex
add subscales to abstract

@@
 we showed that many species,
-including \emph{Alistipes obesi} and \emph{Blautia wexlerae}, were % check RF models for direction of B. wex
+including \emph{Alistipes obesi} and \emph{Blautia wexlerae}, were
 associated with higher cognitive function, while some
-species such as \emph{Ruminococcus gnavus} were more commonly found in % check RF models for direction of R. gnav
+species such as \emph{Ruminococcus gnavus} were more commonly found in
 children with low cognitive scores after controlling for sociodemographic factors.
@@
 as important in predicting cognitive function also dominated the
-feature importance metric for individual brain regions.
-For example, \emph{B. wexlerae} % is this still true?
+feature importance metric for individual brain regions,
+and for specific subscales of cognitive function.
+For example, \emph{B. wexlerae}
 was the most important species in models predicting the size of the parahippocampal region
-in both the left and right hemispheres, while several species from
-the phylum \textit{Bacteroidetes}, including GABA-producing % is this still true?
+in both the left and right hemispheres
+and was among the top predictors of gross motor and expressive language performance.
+Several species from
+the phylum \textit{Bacteroidetes}, including GABA-producing
 \emph{B. ovatus}, were important for predicting

### 2023-10-11 manuscript/main.tex
minor change to intro, Figure 1 numbers vetted

@@
 The early microbiome is characterized by low microbial
-diversity, rapid succession and evolution, and domination by
+diversity, rapid succession and evolution, and is dominated by
 Actinobacteria, particularly the genus \emph{Bifidobacterium},
@@
 (Figure 1G, R\textsuperscript{2} = 0.0099, q \textless{} 0.01),
-as was variation in microbial gene functions (R\textsuperscript{2} = 0.0119, q \textless{} 0.001).
+as was variation in microbial gene functions (R\textsuperscript{2} = 0.0115, q \textless{} 0.001).
 Variation in microbial taxa and genes was not significantly associated with cognitive
@@
 \cite{suzekUniRefComprehensiveNonredundant2007} -
-R\textsuperscript{2} = 0.0263, q \textless{} 0.001), and very strongly
+R\textsuperscript{2} = 0.0258, q \textless{} 0.001), and very strongly
 associated with variation in neuroimaging (MRI) profiles (R\textsuperscript{2} = 0.258, q

### 2023-10-11 manuscript/main.tex
MSEL is only to 36m

@@
         between individual taxa and domain-specific skill assessments
-        from the MSEL for children under 6m (left), 18m to 48m (right),
+        from the MSEL for children under 6m (left), 18m to 36m (right),
         or over across all ages (right)
@@
 we fit multivariate linear models for each species present in at least 10\%
-of samples for children under 6 months, 18 to 48 months, or all ages
+of samples for children under 6 months, 18 to 36 months, or all ages
 with a concurrent MSEL score (Figure 2C).

### 2023-10-11 manuscript/main.tex
complete figure 2
complete figures 3 and 4

@@
 = 0.029), as were genes for synthesizing the 2 carbon SCFA acetate
-(acetate synthesis II, E.S. = -0.409, q = 0.059; acetate synthesis
+(acetate synthesis II, E.S. = -0.491, q = 0.059; acetate synthesis
 III, E.S. = -0.409, q = 0.133).
@@
 glutamate was also significantly enriched in children with
-higher cognitive function scores in this age group (glutamate synthesis II, E.S. = 0.154, 
-q = 0.178). Taken together, these results suggest that
+higher cognitive function scores in this age group (glutamate synthesis II, E.S. = 0.355, 
+q = 0.023). Taken together, these results suggest that
 microbial metabolic activity, particularly the metabolism (synthesis and
@@
 in the first year of life
-could predict \textit{future} cognitive performance (Figure 3B, D).
+could predict \textit{future} cognitive performance measured (Figure 3B, D).
 Models were consistently able to learn microbial species that are associated
 with cognitive performance at a future visit
-(mean time between visits: 21.5 months [sd = 9.9 months] (Figure 3C),
+(mean time between visits: 21.5 months, S.D. = 9.9 months; Figure 3C,
 mean test set correlation: 0.27).
@@
 
-\subsubsection*{Gut microbial taxa are associated with specific cognitive functions} %TODO: add numbers!
+\subsubsection*{Gut microbial taxa are associated with specific cognitive functions}
 
@@
 To accomplish this, we assessed individual subscales of the  
-Mullen Scales of Early Learning (MSEL, birth to 36 months - \textit{see Methods}). 
+MSEL (birth to 36 months - \textit{see Methods}). 
 Taxonomic profiles coupled with demographics exhibited predictive potential towards three
 of five probed subscales (Figure 4A), 
-namely Expressive Language ( test-set RMSE 12.05, test-set correlation 0.368), 
-Gross Motor (test-set RMSE 11.07, test-set correlation 0.137) and 
-Visual Reception (test-set RMSE 15.2, test-set correlation 0.333).
+namely expressive language ( test-set RMSE 12.05, test-set correlation 0.368), 
+gross motor (test-set RMSE 11.07, test-set correlation 0.137) and 
+visual reception (test-set RMSE 15.2, test-set correlation 0.333).
 Between the subscales, not only the predictability (measured by the test-set figures of merit) is different,
@@
 Notably, \textit{Bifidobacterium pseudocatenulatum}, \textit{B. wexlerae} and \textit{E. eligens} are
-substantially more important for the prediction of EL, while
-\textit{R. faecis}, \textit{Streptococcus salivarius} and \textit{F. saccharivorans} are
-distinctly predictive of GM, and
+substantially more important for the prediction of expressive language, while
+\textit{R. faecis}, \textit{Streptococcus salivarius} and \textit{Fusicatenibacter saccharivorans} are
+distinctly predictive of gross motor, and
 \textit{Clostridium innocuum}, \textit{B. vulgatus} 
-stand out on models targeting visual reception.
+stand out on models targeting visual reception (Supplenmentary Tables S2-4).
 
@@
 \emph{B. ovatus},
-\emph{Fusicatenibacter saccharivorans},
+\emph{F. saccharivorans},
 \emph{Ruminococcus torques},
@@
 \emph{Coprococcus comes},
-\emph{Asaccharobacter celatus} and
+\emph{A. celatus} and
 \emph{B. wexlerae}, which combined, account for approximately 10\% of the cumulative relative importances,
@@
 \emph{B. wexlerae}, which was important in RF models of cognitive function, 
-and was the only species to be significantly associated
-with cogntive function in young children using LMs, 
 was also important in models of the entorhinal, fusiform and lingual areas

### 2023-10-11 manuscript/main.tex
complete discussion

@@
 Among these, \emph{A. hadrus} is the variable with 
-highest average relative importance along all reported 
+highest average relative importance along all reported models
 (mean = 1.4\%; min = 0.8\%; max = 3.6\%, Figure 4D-E).
@@
 resonate with previous findings that associate chronic inflammation with
-changes to neural responses in some types of stimulae, notably visual\cite{xieChronicInfNeuro2019}.
+changes to neural responses in some types of stimulae \cite{xieChronicInfNeuro2019}.
 
@@
 \textit{B. wexlerae} and \textit{A. hadrus} to be among the top 15\% important predictors
-on both left and right thalamus and the EL subscale.
+on both left and right thalamus and the expressive language subscale.
 Another example of those findings is a previously reported association between
-the central opercular cortex (COC) and gross motor development \cite{jiangBrainNetworkCognition2023}. In this case, we identified that
-butyrate-producing \textit{Fusicatenibacter saccharivorans} is among the 15\% most important predictors
+the central opercular cortex (COC) and gross motor development \cite{jiangBrainNetworkCognition2023}.
+In this case, we identified that
+butyrate-producing \textit{F. saccharivorans} is among the 15\% most important predictors
 for both the gross motor subscale and three bilateral components of the COC
@@
 Previous results also link to relationships between the caudate and the cerebellum,
-and the Visual Reception score \cite{deoniMyelinationMullenPreprint2022}.
+and the visual reception score \cite{deoniMyelinationMullenPreprint2022}.
 \textit{S. salivarius} was consistently found among the top 20\% predictors for this subscale, along with

### 2023-10-11 manuscript/Table3-future.jl
rearrange tables structure

@@ -1,86 +1,0 @@
-#####
-# Table 3. Benchmark metrics for the cognitive assessment score prediction models. Confidence intervals are calculated from the distribution of metrics from repeated CV at a confidence level of 95%
-#####
-
-using Resonance
-using CategoricalArrays
-using Statistics
-using Chain
-using JLD2
-using MLJ
-using DataFrames.PrettyTables
-
-isdir(tablefiles()) || mkpath(tablefiles())
-
-## 1. Loading the model files
-RandomForestClassifier = MLJ.@load RandomForestClassifier pkg=DecisionTree
-RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
-# concurrent cogScore reression from taxonomic profiles
-regression_futureCogScores_onlydemo = JLD2.load(modelfiles("regression_futureCogScores_onlydemo.jld"))["regression_futureCogScores_onlydemo"]
-regression_futureCogScores_onlytaxa = JLD2.load(modelfiles("regression_futureCogScores_onlytaxa.jld"))["regression_futureCogScores_onlytaxa"]
-regression_futureCogScores_demoplustaxa = JLD2.load(modelfiles("regression_futureCogScores_demoplustaxa.jld"))["regression_futureCogScores_demoplustaxa"]
-regression_futureCogScores_onlyecs = JLD2.load(modelfiles("regression_futureCogScores_onlyecs.jld"))["regression_futureCogScores_onlyecs"]
-regression_futureCogScores_demoplusecs = JLD2.load(modelfiles("regression_futureCogScores_demoplusecs.jld"))["regression_futureCogScores_demoplusecs"]
-## 2. Calculating the Figures of Merit
-
-conf_interval(vv, critical_z=1.96) = critical_z * std(vv) / sqrt(length(vv))
-
-prod_summary_table = vcat(
-    DataFrames.combine(
-        groupby(regression_futureCogScores_onlydemo.merits, :Hyperpar_Idx),
-        :Test_RMSE => (x -> round(mean(x); digits = 2)) => :Test_RMSE_mean,
-        :Test_RMSE => (x -> round(conf_interval(x); digits = 2)) => :Test_RMSE_CI,
-        :Test_Cor => (x -> round(mean(x); digits = 2)) => :Test_Cor_mean,
[…]

### 2023-10-11 manuscript/Table3.jl

@@ -171,6 +171,6 @@ table3 = DataFrame(
 CSV.write(tablefiles("Table3.csv"), table3)
 
 pretty_table(table3;
-    header = [ "Subject Ages (months)", "Microbial feature", "Demo.", "Test set correlation (± C.I.)", "Test set Root-mean-square error (± C.I.)" ],
+    header = [ "Subject Ages (months)", "Microbial feature", "Demo.", "Test set Root-mean-square error (± C.I.)", "Test set correlation (± C.I.)" ],
     backend = Val(:latex)
 )

### 2023-10-11 manuscript/Table4.jl

@@ -1,5 +1,5 @@
 #####
-# Table 4. Summary statistics for the neuroanatomy prediction benchmarks
+# Table 3. Benchmark metrics for the cognitive assessment score prediction models. Confidence intervals are calculated from the distribution of metrics from repeated CV at a confidence level of 95%
 #####
 
 using Resonance
@@ -13,30 +13,74 @@ using DataFrames.PrettyTables
 isdir(tablefiles()) || mkpath(tablefiles())
 
 ## 1. Loading the model files
-
+RandomForestClassifier = MLJ.@load RandomForestClassifier pkg=DecisionTree
 RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
-# concurrent brain volumes regression from taxonomic profiles
-brain_models = JLD2.load(modelfiles("brain_models.jld"))["brain_models"]
-include("manuscript/Figure4-definitions.jl")
-
+# concurrent cogScore reression from taxonomic profiles
+regression_futureCogScores_onlydemo = JLD2.load(modelfiles("regression_futureCogScores_onlydemo.jld"))["regression_futureCogScores_onlydemo"]
+regression_futureCogScores_onlytaxa = JLD2.load(modelfiles("regression_futureCogScores_onlytaxa.jld"))["regression_futureCogScores_onlytaxa"]
+regression_futureCogScores_demoplustaxa = JLD2.load(modelfiles("regression_futureCogScores_demoplustaxa.jld"))["regression_futureCogScores_demoplustaxa"]
+regression_futureCogScores_onlyecs = JLD2.load(modelfiles("regression_futureCogScores_onlyecs.jld"))["regression_futureCogScores_onlyecs"]
+regression_futureCogScores_demoplusecs = JLD2.load(modelfiles("regression_futureCogScores_demoplusecs.jld"))["regression_futureCogScores_demoplusecs"]
 ## 2. Calculating the Figures of Merit
-mean_brain_merits = reduce(
-    vcat,
[…]

### 2023-10-11 manuscript/Table5.jl

@@ -1,0 +1,178 @@
+#####
+# Table 5. Benchmark metrics for the cognitive assessment Mullen subscores
+#####
+
+using Resonance
+using CategoricalArrays
+using Statistics
+using Chain
+using JLD2
+using MLJ
+using DataFrames.PrettyTables
+
+isdir(tablefiles()) || mkpath(tablefiles())
+
+## 1. Loading the model files
+RandomForestClassifier = MLJ.@load RandomForestClassifier pkg=DecisionTree
+RandomForestRegressor = MLJ.@load RandomForestRegressor pkg=DecisionTree
+
+# concurrent cogScore reression from taxonomic profiles
+regression_currentExpressiveLanguages_18to120mo_onlydemo = JLD2.load(modelfiles("regression_currentExpressiveLanguages_18to120mo_onlydemo.jld"))["regression_currentExpressiveLanguages_18to120mo_onlydemo"]
+regression_currentExpressiveLanguages_18to120mo_onlytaxa = JLD2.load(modelfiles("regression_currentExpressiveLanguages_18to120mo_onlytaxa.jld"))["regression_currentExpressiveLanguages_18to120mo_onlytaxa"]
+regression_currentExpressiveLanguages_18to120mo_demoplustaxa = JLD2.load(modelfiles("regression_currentExpressiveLanguages_18to120mo_demoplustaxa.jld"))["regression_currentExpressiveLanguages_18to120mo_demoplustaxa"]
+regression_currentExpressiveLanguages_18to120mo_onlyecs = JLD2.load(modelfiles("regression_currentExpressiveLanguages_18to120mo_onlyecs.jld"))["regression_currentExpressiveLanguages_18to120mo_onlyecs"]
+regression_currentExpressiveLanguages_18to120mo_demoplusecs = JLD2.load(modelfiles("regression_currentExpressiveLanguages_18to120mo_demoplusecs.jld"))["regression_currentExpressiveLanguages_18to120mo_demoplusecs"]
+regression_currentGrossMotors_18to120mo_onlydemo = JLD2.load(modelfiles("regression_currentGrossMotors_18to120mo_onlydemo.jld"))["regression_currentGrossMotors_18to120mo_onlydemo"]
[…]

### 2023-10-11 manuscript/main.tex
incorporated Table 5 for subscales

--- 2023-10-11 manuscript/main.tex
@@
 as important in predicting cognitive function also dominated the
-feature importance metric for individual brain regions,
-and for specific subscales of cognitive function.
-For example, \emph{B. wexlerae}
+feature importance metric for individual brain regions.
+For example, \emph{B. wexlerae} % is this still true?
 was the most important species in models predicting the size of the parahippocampal region
-in both the left and right hemispheres
-and was among the top predictors of gross motor and expressive language performance.
-Several species from
-the phylum \textit{Bacteroidetes}, including GABA-producing
+in both the left and right hemispheres, while several species from
+the phylum \textit{Bacteroidetes}, including GABA-producing % is this still true?
 \emph{B. ovatus}, were important for predicting
@@
 The early microbiome is characterized by low microbial
-diversity, rapid succession and evolution, and is dominated by
+diversity, rapid succession and evolution, and domination by
 Actinobacteria, particularly the genus \emph{Bifidobacterium},
@@
 (Figure 1G, R\textsuperscript{2} = 0.0099, q \textless{} 0.01),
-as was variation in microbial gene functions (R\textsuperscript{2} = 0.0115, q \textless{} 0.001).
+as was variation in microbial gene functions (R\textsuperscript{2} = 0.0119, q \textless{} 0.001).
 Variation in microbial taxa and genes was not significantly associated with cognitive
@@
 \cite{suzekUniRefComprehensiveNonredundant2007} -
-R\textsuperscript{2} = 0.0258, q \textless{} 0.001), and very strongly
+R\textsuperscript{2} = 0.0263, q \textless{} 0.001), and very strongly
 associated with variation in neuroimaging (MRI) profiles (R\textsuperscript{2} = 0.258, q
@@
         between individual taxa and domain-specific skill assessments
-        from the MSEL for children under 6m (left), 18m to 36m (right),
+        from the MSEL for children under 6m (left), 18m to 48m (right),
         or over across all ages (right)
@@
 we fit multivariate linear models for each species present in at least 10\%
-of samples for children under 6 months, 18 to 36 months, or all ages
+of samples for children under 6 months, 18 to 48 months, or all ages
 with a concurrent MSEL score (Figure 2C).
@@
       \hline\hline
-      \textbf{Subject Ages (months)} & \textbf{Microbial feature} & \textbf{Demo.} &  \textbf{Test set correlation (± C.I.)} & \textbf{Test set RMSE (± C.I.)} \\\hline
+      \textbf{Subject Ages (months)} & \textbf{Microbial feature} & \textbf{Demo.} &  \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
       \multirow{5}{*}{0 to 6} & -    & + & 13.37 ± 0.05 & -0.16 ± 0.01 \\ \cline{2-5}
@@
     prediction models. Confidence intervals (C.I.) are calculated from the
-    distribution of metrics from repeated CV at a confidence level of 95.
-    Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
-    Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.}
-    \end{center}
-\end{table}
-
-\begin{table}[!h]
-    \begin{center}
-    \begin{tabular}{|r|r|r|r|r}
-        \hline
-        \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set correlation (± C.I.)} & \textbf{Test set Root-mean-square error (± C.I.)} \\\hline
-        - & + & 20.25 ± 0.07 & 0.43 ± 0.0 \\ \cline{1-4}
-        \multirow{2}{*}{taxa} & - & 22.32 ± 0.08 & 0.07 ± 0.0 \\ \cline{2-4}
-            & + & 21.53 ± 0.08 & 0.27 ± 0.0 \\ \cline{1-4}
-        \multirow{2}{*}{genes} & - & 22.21 ± 0.08 & 0.12 ± 0.0 \\ \cline{2-4}
-            & + & 22.08 ± 0.08 & 0.15 ± 0.0 \\\hline\hline
-    \end{tabular}
-    \caption{\label{tab:rfbench2}Benchmark metrics for the cognitive assessment score
-    future prediction models. Confidence intervals (C.I.) are calculated from the
     distribution of metrics from repeated CV at a confidence level of 95.
@@
 
+\begin{table}[!h]
+    \begin{center}
+    \begin{tabular}{|r|r|r|r|r}
+        \hline
+        \textbf{Microbial feature} & \textbf{Demo.} &  \textbf{Test set Root-mean-square-error (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
+        - & + & 20.25 ± 0.07 & 0.43 ± 0.0 \\ \cline{1-4}
+        \multirow{2}{*}{taxa} & - & 22.32 ± 0.08 & 0.07 ± 0.0 \\ \cline{2-4}
+            & + & 21.53 ± 0.08 & 0.27 ± 0.0 \\ \cline{1-4}
+        \multirow{2}{*}{genes} & - & 22.21 ± 0.08 & 0.12 ± 0.0 \\ \cline{2-4}
+            & + & 22.08 ± 0.08 & 0.15 ± 0.0 \\\hline\hline
+    \end{tabular}
+    \caption{\label{tab:rfbench2}Benchmark metrics for the cognitive assessment score
+    future prediction models. Confidence intervals (C.I.) are calculated from the
+    distribution of metrics from repeated CV at a confidence level of 95.
+    Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
+    Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.}
+    \end{center}
+\end{table}
+
 Nearly all important taxa in RF models that predicted cognitive function
 in children older than 18 months were also important in
-models that predicted future cognitive function, if they were present (Figure 3B).
+models that predicted future cognitive function, if they were present (Figure 3B, Table S5).
 At the same time, many additional species were important for future prediction models
@@
 Taxonomic profiles coupled with demographics exhibited predictive potential towards three
-of five probed subscales (Figure 4A), 
+of five probed subscales (Figure 4A, Table 4), 
 namely Expressive Language ( test-set RMSE 12.05, test-set correlation 0.368), 
@@
 Although certain species such as \textit{B. longum}, \textit{F. prausznitzii} 
-have similar relative importance across all subscales, others are substantially different.
+have similar relative importance across all subscales, others are substantially different (Tables S2-S4).
 Notably, \textit{Bifidobacterium pseudocatenulatum}, \textit{B. wexlerae} and \textit{E. eligens} are
@@
 stand out on models targeting visual reception.
+
+\begin{table}[!h]
+    \begin{center}
+        \begin{tabular}{|r|r|r|r|r|}
+        \hline
+        \textbf{Target Mullen scale} & \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
+        \multirow{5}{*}{Expressive Language} & - & + & 11.17 ± 0.04 & 0.48 ± 0.0 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa}  & - & 12.37 ± 0.04 & 0.25 ± 0.0 \\ \cline{3-5}
+            &                        & + & 12.05 ± 0.04 & 0.37 ± 0.0 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 12.59 ± 0.04 & 0.15 ± 0.0 \\ \cline{3-5}
+            &                        & + & 12.46 ± 0.04 & 0.21 ± 0.0 \\ \cline{1-5}
+        \multirow{5}{*}{Gross Motor} & - & + & 10.53 ± 0.04 & 0.346 ± 0.006 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa}  & - & 11.08 ± 0.04 & 0.134 ± 0.006 \\ \cline{3-5}
+            &                        & + & 11.07 ± 0.04 & 0.137 ± 0.006 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \cline{3-5}
+            &                        & + & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \hline\hline
+        \multirow{5}{*}{Visual reception} & - & + & 14.13 ± 0.04 & 0.462 ± 0.003 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa}  & - & 15.49 ± 0.04 & 0.245 ± 0.004 \\ \cline{3-5}
+            &                        & + & 15.2 ± 0.04 & 0.333 ± 0.004 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 15.97 ± 0.04 & 0.067 ± 0.004 \\ \cline{3-5}
+            &                        & + & 15.84 ± 0.04 & 0.118 ± 0.005 \\ \hline\hline
+        \end{tabular}
+        \caption{\label{tab:subscalebenchmark}Benchmark metrics for the Mullen subscales prediction models.
+        Confidence intervals (C.I.) are calculated from the
+        distribution of metrics from repeated CV at a confidence level of 95.
+        Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
+        Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.
+        Age range, in all cases, was 18-120 months.}
+    \end{center}
+\end{table}
 
@@
 Some brain regions were more readily
-predicted by RF models trained on microbial taxa (Table 4, Table S2), in particular those that were highly correlated with age.
+predicted by RF models trained on microbial taxa (Table 5, Table S2), in particular those that were highly correlated with age.
 These included the L/R lingual gyrus (mean RF correlations, Left =

### 2023-10-11 manuscript/main.tex
complete methods

--- 2023-10-11 manuscript/main.tex
@@
 As a study of neurotypical development, mothers and their children
-were devoid of major risk factors for abnormal or atypical development.
-These included preterm birth (\textless 37 weeks gestation) low birth weight (\textless 1500g),
+exclusion criteria included major risk factors for abnormal or atypical development,
+i.e., preterm birth (\textless 37 weeks gestation) low birth weight (\textless 1500g),
 \textit{in utero} exposure to alcohol, cigarette or illicit substance exposure;
@@
 them from subsequent analysis.
-For this study, 361 typically-developing children between the
-ages of 2.8 months and 10 years old (median age 2 years, 2 months) were
+For this study, 381 typically-developing children between the
+ages of 40 days and 10 years (median age 2 years, 2 months) were
 selected for analysis based on having provided at least one stool sample
@@
 computational tools \cite{beghiniIntegratingTaxonomicFunctional2021}.
-First, KneadData (v0.7.7) was used to perform quality
+First, KneadData (v0.10.0) was used to perform quality
 control of raw sequence reads, such as read trimming and removal of
-reads matching a human genome reference. Next, MetaPhlAn (v3.0.7, using
-database mpa\_v30\_CHOCOPhlAn\_201901) was used to generate taxonomic
+reads matching a human genome reference. Next, MetaPhlAn (v3.1.0, using
+database mpa\_v31\_CHOCOPhlAn\_201901) was used to generate taxonomic
 profiles by aligning reads to a reference database of marker genes.
-Finally, HUMAnN (v3.0.0a4) was used to functionally profile the
+Finally, HUMAnN (v3.6.0) was used to functionally profile the
 metagenomes.
@@
 (for the experiments in Figure 3) or, when available, subscores on the
-Mullen Scales of Early Learning (Figure 4 A-B: Expressive Language, Gross Motor and Visual Reception).
+Mullen Scales of Early Learning (Figure 4 A-B: expressive language, gross motor and visual reception).
 For the concurrent prediction experiments, when multiple viable longitudinal samples were present,

### 2023-10-11 manuscript/main.tex

@@
 Taxonomic profiles coupled with demographics exhibited predictive potential towards three
-of five probed subscales (Figure 4A), 
+of five probed subscales (Figure 4A, Table 5), 
 namely expressive language ( test-set RMSE 12.05, test-set correlation 0.368), 
@@
 stand out on models targeting visual reception (Supplenmentary Tables S2-4).
+
+\begin{table}[!h]
+    \begin{center}
+        \begin{tabular}{|r|r|r|r|r|}
+        \hline
+        \textbf{Target Mullen scale} & \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
+        \multirow{5}{*}{Expressive Language} & - & + & 11.17 ± 0.04 & 0.48 ± 0.0 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa}  & - & 12.37 ± 0.04 & 0.25 ± 0.0 \\ \cline{3-5}
+            &                        & + & 12.05 ± 0.04 & 0.37 ± 0.0 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 12.59 ± 0.04 & 0.15 ± 0.0 \\ \cline{3-5}
+            &                        & + & 12.46 ± 0.04 & 0.21 ± 0.0 \\ \cline{1-5}
+        \multirow{5}{*}{Gross Motor} & - & + & 10.53 ± 0.04 & 0.346 ± 0.006 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa}  & - & 11.08 ± 0.04 & 0.134 ± 0.006 \\ \cline{3-5}
+            &                        & + & 11.07 ± 0.04 & 0.137 ± 0.006 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \cline{3-5}
+            &                        & + & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \hline\hline
+        \multirow{5}{*}{Visual reception} & - & + & 14.13 ± 0.04 & 0.462 ± 0.003 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa}  & - & 15.49 ± 0.04 & 0.245 ± 0.004 \\ \cline{3-5}
+            &                        & + & 15.2 ± 0.04 & 0.333 ± 0.004 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 15.97 ± 0.04 & 0.067 ± 0.004 \\ \cline{3-5}
+            &                        & + & 15.84 ± 0.04 & 0.118 ± 0.005 \\ \hline\hline
+        \end{tabular}
+        \caption{\label{tab:subscalebenchmark}Benchmark metrics for the Mullen subscales prediction models.
+        Confidence intervals (C.I.) are calculated from the
+        distribution of metrics from repeated CV at a confidence level of 95.
+        Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
+        Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.
+        Age range, in all cases, was 18-120 months.}
+    \end{center}
+\end{table}
 

### 2023-10-11 manuscript/main.tex
complete merge
update diff
vanja-found typo

@@
 Interestingly, while no taxa were significantly associated
-with composite scores in chilren under 6 months old,
+with composite scores in children under 6 months old,
 two species from the genus \textit{Streptococcus}
@@
             & \multirow{2}{*}{genes} & - & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \cline{3-5}
-            &                        & + & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \hline\hline
+            &                        & + & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \hline
         \multirow{5}{*}{Visual reception} & - & + & 14.13 ± 0.04 & 0.462 ± 0.003 \\ \cline{2-5}
@@
 occuring in only ~5\% of samples at greater than 2\% abundance
-(in contrast to \textit{A. celatus}, which was in 32\% of samples)
+(in contrast to \textit{A. celatus}, which was in 32\% of samples).
 While we could not find published evidence of neurological roles for \emph{E. lenta},

### 2023-10-11 manuscript/main.tex
rework sentence

@@
 visual reception (test-set RMSE 15.2, test-set correlation 0.333).
-Between the subscales, not only the predictability (measured by the test-set figures of merit) is different,
-but also the relative importances of several taxa (Figure 4B).
+Between the subscales, both the predictability (measured by the test-set figures of merit)
+and the relative importances of several taxa (Figure 4B) is variable.
 Although certain species such as \textit{B. longum}, \textit{F. prausznitzii} 

### 2023-10-11 manuscript/main.tex
T6 reference

@@
 Some brain regions were more readily
-predicted by RF models trained on microbial taxa (Table 5, Table S2), in particular those that were highly correlated with age.
+predicted by RF models trained on microbial taxa (Table 5, Table S6), in particular those that were highly correlated with age.
 These included the L/R lingual gyrus (mean RF correlations, Left =

### 2023-10-11 manuscript/main.tex
pathobionts

@@
 taxa that were important for predicting concurrent cognitive assessment scores after 18 months, notably
-pathobiontic \textit{Enterobacteriaceae} such as
+pathobionts \textit{Enterobacteriaceae} such as
 \textit{Escherichia} and \textit{Klebsiella}. 

### 2023-10-11 manuscript/main.tex
make figures 90% text width

@@
     \centering
+    \small
     \includegraphics[width=0.9\textwidth]{assets/Figure1.png}
@@
     \centering
-    \includegraphics[width=\textwidth]{assets/Figure2.png}
+    \small
+    \includegraphics[width=0.9\textwidth]{assets/Figure2.png}
     \caption{
@@
     \centering
-    \includegraphics[width=\textwidth]{assets/Figure3.png}
+    \small
+    \includegraphics[width=0.9\textwidth]{assets/Figure3.png}
     \caption{
@@
     \centering
-    \includegraphics[width=\textwidth]{assets/Figure4.png}
+    \small
+    \includegraphics[width=0.9\textwidth]{assets/Figure4.png}
     \caption{
@@
 taxa that were important for predicting concurrent cognitive assessment scores after 18 months, notably
-pathobionts \textit{Enterobacteriaceae} such as
+pathobionts in the genus \textit{Enterobacteriaceae} such as
 \textit{Escherichia} and \textit{Klebsiella}. 

### 2023-10-11 manuscript/main.tex

@@
 taxa that were important for predicting concurrent cognitive assessment scores after 18 months, notably
-pathobionts in the genus \textit{Enterobacteriaceae} such as
+pathobionts in the family \textit{Enterobacteriaceae} such as
 \textit{Escherichia} and \textit{Klebsiella}. 

### 2023-10-18 manuscript/main.tex
consolidate citations

@@
 
-\addbibresource{zotero.bib}
-\addbibresource{addons.bib}
+\addbibresource{citations.bib}
 
@@
 and atypical neurocognition and neurological disorders
-\cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021,
-      magnussonRelationshipsDietrelatedChanges2015,mayneris-perxachsMicrobiotaAlterationsProline2022,
-      needhamGutderivedMetaboliteAlters2022},
+\cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021,magnussonRelationshipsDietrelatedChanges2015,mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022},
 we sought to determine if specific taxa or gene functions were associated
@@
 on Random Forests with high-dimensional, low sample size
-datasets \cite{woodruffInflammationAutoreactivityDefine2022}.
+datasets \cite{woodruffChronicInflammationNeutrophil2023a}.
 The resulting fitness-weighted importances were used
@@
 and 3rd party packages
-\cite{bezansonJuliaFreshApproach2017,
-    bonhamMicrobiomeJlBiobakeryUtils2021,
-    danischMakieJlFlexible2021,
-    blaomMLJJuliaPackage2020,
-    ben_sadeghi_2022_7359268,
-    dahua_lin_2023_7695673,
-    douglas_bates_2023_7734970}.
+\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021,blaomMLJJuliaPackage2020,ben_sadeghi_2022_7359268,dahua_lin_2023_7695673,douglas_bates_2023_7734970}.
 Information for replicating the package environment

### 2023-10-18 manuscript/main.tex
fix citations format
fix acknowledgements sections

@@
 \usepackage[utf8]{inputenc}
+\usepackage{enumitem}
 
@@
 backend=biber,
-style=science
+style=numeric,
+maxbibnames=50
 ]{biblatex}
@@
 
-\subsubsection*{Data and code availability}
+\section*{References}
+
+\printbibliography[heading=none]
+
+\section*{Acknowledgements}
+% https://www.science.org/content/page/science-advances-information-authors#acknowledgments
+
+We thank Libusha Kelly (Albert Einstein School of Medicine) and Daphne Koinis Mitchell (Brown University)
+for helpful feedback on the manuscript. The research team also thanks the participants and families
+of the participants who generously shared their time and samples for this project.
+
+\subsubsection*{Data and materials availability}
 
@@
 
-\section*{References}
-
-\printbibliography[heading=none]
-
-\section*{Acknowledgements}
-% https://www.science.org/content/page/science-advances-information-authors#acknowledgments
-
+\subsubsection*{Funding}
 This research was funded by NIH UG3 OD023313 and Wellcome LEAP 1kD.
+
+\subsubsection*{Competing Interests}
 The authors declare that they have no competing interests.
-We thank Libusha Kelly (Albert Einstein School of Medicine) and Daphne Koinis Mitchell (Brown University)
-for helpful feedback on the manuscript. The research team also thanks the participants and families
-of the participants who generously shared their time and samples for this project.
-
+
+\subsubsection*{Author contributions}
 % https://credit.niso.org/
 
-KSB, VKC, CH, and SD were responsible for conceptualization,
-methodology and project administration.
-VKC, SD, and VD were responsible for funding aqusition.
-KSB, SHM, MB, FB, RCL, and JB contributed to data curation.
-Formal analyses were conducted by KSB, GFB, and MB.
-Investigation was performed by KSB, GFB, SHM, EW, FB, RCL, and MB.
-Resources were provided by SHM, EW, FB, RCL, JB, CH, and MB.
-Software and data visualizations were produced by KSB and GFB.
-KSB, MB, SD, VD, and VKC were responsible for supervision.
-Validation was performed by KSB, SHM, JB, and MB.
-KSB and GFB wrote the initial draft, 
-and review and editing were provided by SD, CH, and VKC.
+\begin{itemize}[noitemsep]
+    \item Conceptualization: KSB, VKC, CH, and SD
+    \item Methodology: KSB, VKC, CH, and SD
+    \item Project administraion: KSB, VKC, CH, and SD
+    \item Funding acquisition: VKC, SD, and VD
+    \item Data curation: KSB, GFB, SHM, MB, FB, RCL, and JB
+    \item Formal analyses: KSB, GFB, and MB
+    \item Investigation: KSB, GFB, SHM, EW, FB, RCL, and MB
+    \item Resources: SHM, EW, FB, RCL, JB, CH, and MB
+    \item Software and data visualizations: KSB and GFB
+    \item Supervision: KSB, MB, SD, VD, and VKC 
+    \item Validation: KSB, SHM, JB, and MB
+    \item Initial draft: KSB and GFB
+    \item Revision and editing: KSB, GFB, SD, CH, and VKC
+\end{itemize}
 

### 2023-10-18 manuscript/main.tex
remove urls

@@
 as well as code for data analysis and figure generation, as well as scripts for automated
-download of input files are available on GitHub \cite{kevin_bonham_2023_7647510}.
+download of input files are available on GitHub at https://github.com/KlepacCerajLab/Resonance,
+and archived on Zenodo \cite{kevin_bonham_2023_7647510}.
 

### 2023-10-18 manuscript/main.tex
start on figure and table legend titles

@@
     \caption{
-        \label{fig:1} Study design and overview. (\textbf{A}) Stool samples,
+        \label{fig:1} \textbf{The RESONANCE cohort is a diverse cohort with longitudinal sampling.} (\textbf{A}) Stool samples,
         cognitive assessments and neuroimaging were collected from participants
@@
     \caption{
-        Taxa and gene functional groups are associated with cognitive function.
+        \textbf{Taxa and gene functional groups are associated with cognitive function.}
         (\textbf{A}) Volcano plot of multivariable linear model results showing the
@@
     \end{tabular}
-    \caption{\label{tab:fsea}Feature set enrichment analysis on neuroactive microbial gene sets.
+    \caption{\label{tab:fsea} \textbf{Neuroactive microbial gene sets microbial gene sets are associated with cognitive function}.
     Red and blue text indicates significant (q value \textless 0.2) genesets with a positive or negative
@@
     \caption{
-        Random forest models predict both concurrent and future measured cognitive function.
+        \textbf{Random forest models predict both concurrent and future measured cognitive function.}
         Comparison of RF predictor importance versus linear models for children
@@
     \end{tabular}
-    \caption{\label{tab:rfbench}Benchmark metrics for the cognitive assessment score
+    \caption{\label{tab:rfbench} \textbf{RF models on microbial taxa and genes can learn associations with concurrently measured cognitive performance.}
+    Benchmark metrics for the cognitive assessment score
     prediction models. Confidence intervals (C.I.) are calculated from the
@@
     \end{tabular}
-    \caption{\label{tab:rfbench2}Benchmark metrics for the cognitive assessment score
+    \caption{\label{tab:rfbench2}  \textbf{RF models on microbial taxa and genes can learn associations with future cognitive performance.}
+    Benchmark metrics for the cognitive assessment score
     future prediction models. Confidence intervals (C.I.) are calculated from the
@@
         \end{tabular}
-        \caption{\label{tab:subscalebenchmark}Benchmark metrics for the Mullen subscales prediction models.
+        \caption{\label{tab:subscalebenchmark}  \textbf{RF models on microbial taxa and genes can learn associations with performance in cognitive subdomains.}
+        Benchmark metrics for the Mullen subscales prediction models.
         Confidence intervals (C.I.) are calculated from the
@@
     \end{tabular}
-    \caption{\label{tab:neuro}Summary statistics for the neuroanatomy prediction benchmarks.
+    \caption{\label{tab:neuro}  \textbf{RF models on microbial taxa predicting brain regions have generally good performance.}
     Mean absolute proportional error (MAPE) and correlation coeficient (R)
@@
     \caption{
-        Microbial feature importance in predicting MSEL subscores and brain volumes in children over 18 months of age.
+        \textbf{Microbial features can train RF models to predict MSEL subscores and brain volumes in children over 18 months of age}.
         (\textbf{A}) Test set correlations for RF models predicting MSEL subscale scores,

### 2023-10-18 manuscript/main.tex
start on figure legend revisions

@@
     \caption{\label{tab:rfbench2}  \textbf{RF models on microbial taxa and genes can learn associations with future cognitive performance.}
-    Benchmark metrics for the cognitive assessment score
-    future prediction models. Confidence intervals (C.I.) are calculated from the
+    Confidence intervals (C.I.) are calculated from the
     distribution of metrics from repeated CV at a confidence level of 95.

### 2023-10-18 manuscript/main.tex
fix diff

@@
 
-\subsubsection*{Cohort overview and summary data}
+\subsubsection*{The RESONANCE cohort is part of a study of child development}
 
@@
 and brain volume profiles was similarly driven largely by subject age (Figure 1E-F).
-
-\begin{sidewaysfigure}[!h]
-    \centering
-    \small
-    \includegraphics[width=0.9\textwidth]{assets/Figure1.png}
-    \caption{
-        \label{fig:1} \textbf{The RESONANCE cohort is a diverse cohort with longitudinal sampling.} (\textbf{A}) Stool samples,
-        cognitive assessments and neuroimaging were collected from participants
-        at different ages throughout the first years of life.
-        (\textbf{B}) Cognitive function scores are assessed by
-        different instruments, but may be normalized using full-scale composite scores.
-        (\textbf{C-D}) Principal coordinates analysis using Bray-Curtis dissimilarity on
-        taxonomic profiles demonstrates high beta diversity, with much of the
-        first axis of variation explained by increasing age and alpha diversity.
-        Differences in gene function profiles (\textbf{E}) and neuroimaging (PCA based on
-        the Euclidean distance of brain region volumes) (\textbf{F}) are likewise
-        dominated by changes as children age. (\textbf{G}) Permutation analysis of
-        variance (PERMANOVA) of taxonomic profiles, functional profiles
-        (annotated as UniRef90s) and neuroimaging, against the metadata of
-        interest. Variations in taxonomic and functional profiles explain a
-        modest but significant percent of the variation in cognitive development
-        in children over 18 months of age. (\textbf{H}) Mantel testing of different
-        microbial feature matrices, shows overlapping but distinct patterns of
-        variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
-        used as cut-offs in some following models.
-    }
-
-\end{sidewaysfigure}
 
@@
 
-\begin{sidewaysfigure}[!htb]
-    \centering
-    \small
-    \includegraphics[width=0.9\textwidth]{assets/Figure2.png}
-    \caption{
-        \textbf{Taxa and gene functional groups are associated with cognitive function.}
-        (\textbf{A}) Volcano plot of multivariable linear model results showing the
-        relationship between individual taxa and cognitive function score in
-        children over 18 months of age (left) or across all ages (right),
-        controlling for age, maternal education, and sequencing depth.
-        Taxa that were significant after FDR correction (q
-        \textless{} 0.2) are colored red. (\textbf{B}) For taxa that were significantly
-        associated with cognitive function,
-        heatmaps of correlation with cognitive function in different age groups.
-        (\textbf{C}) Multivariate linear models showing the relationship
-        between individual taxa and domain-specific skill assessments
-        from the MSEL for children under 6m (left), 18m to 36m (right),
-        or over across all ages (right)
-        controlling for age, maternal education, and sequencing depth.
-        Asterisks in (A) and (C) indicate significance (q \textless{} 0.2)
-        after multiple hypothesis correction.
-        (\textbf{D-E}) Enrichment plots for proprionate degradation I gene set
-        used in feature set enrichment analysis (FSEA) in children under 6m (D)
-        or over 18m (E). (\textbf{G}) summary of FSEA
-        results in children under 6m of age (left)
-        over 18 months (middle) or across all ages (right), colored based on the
-        significance of the association. Markers indicate the individual
-        correlation of genes within a gene set,
-        and vertical bars colored bars represent the
-        median correlation of that gene set.
-        Vertical dotted lines within each panel represent the median T statistic
-        for all genes in that age group.
-    }
-    \label{fig:2}
-\end{sidewaysfigure}
-
 Cognitive assessments (MSEL, WPPSI, WISC)
@@
         geneset & E.S. & q value & E.S. & q value \\\hline
-        Acetate syn. I & -0.165 & 0.284 & 0.192 & \color{red}{\textbf{0.145}} \\
-        Acetate syn. II & -0.491 & \color{blue}{\textbf{0.059}} & 0.248 & \color{red}{\textbf{0.15}} \\
-        Acetate syn. III & -0.409 & \color{blue}{\textbf{0.133}} & 0.215 & 0.209 \\
-        Butyrate syn. II & -0.357 & 0.284 & 0.414 & \color{red}{\textbf{0.178}} \\
-        Glutamate syn. II & 0.343 & 0.284 & 0.355 & \color{red}{\textbf{0.023}} \\
-        Propionate deg. I & -0.659 & \color{blue}{\textbf{0.029}} & 0.154 & \color{red}{\textbf{0.178}} \\\hline
+        Acetate syn. I & -0.165 & 0.284 & 0.192 & \textbf{0.145} \\
+        Acetate syn. II & -0.491 & \textbf{0.059} & 0.248 & \textbf{0.15} \\
+        Acetate syn. III & -0.409 & \textbf{0.133} & 0.215 & 0.209 \\
+        Butyrate syn. II & -0.357 & 0.284 & 0.414 & \textbf{0.178} \\
+        Glutamate syn. II & 0.343 & 0.284 & 0.355 & \textbf{0.023} \\
+        Propionate deg. I & -0.659 & \textbf{0.029} & 0.154 & \color{red}{\textbf{0.178}} \\\hline
     \end{tabular}
     \caption{\label{tab:fsea} \textbf{Neuroactive microbial gene sets microbial gene sets are associated with cognitive function}.
-    Red and blue text indicates significant (q value \textless 0.2) genesets with a positive or negative
-    pseudomedian, respectively.}
+    Bold text indicates significant (q value \textless 0.2) genesets with a positive or negative
+    pseudomedian.}
     \end{center}
@@
 or in combination with microbial taxonomic profiles (Table 3).
-
-\begin{sidewaysfigure}[!htb]
-    \centering
-    \small
-    \includegraphics[width=0.9\textwidth]{assets/Figure3.png}
-    \caption{
-        \textbf{Random forest models predict both concurrent and future measured cognitive function.}
-        Comparison of RF predictor importance versus linear models for children
-        over 18 months (\textbf{A}).
-        Colors represent whether the species belong to the group of
-        top-important features that account for 60\% of the cumulative
-        importance on the RF model, if that species was significant (q
-        \textless{} 0.2) in linear models, both, or neither.
-        Comparison of RF predictor importance for microbial features
-        between concurrent and future cognitive function measure
-        (\textbf{B}).
-        Colors represent whether the species belong to the group of
-        top-important features that account for 60\% of the cumulative
-        importance on the RF model only for concurrent prediction,
-        only for the future prediction, or for both models.
-        Features absent of one model are shown on the origin-aligned sideplots.
-        Distribution of ages for each microbiome sampling - cognitive assesment dyad (\textbf{C}).
-        Blue violin plot shows distribution of sampling age,
-        while red violin plot shows distribution of asessment age.
-        Black lines connect ages for each individual subject.      
-        Ranked feature importance for taxa in RF models for
-        both concurrent (\textbf{D}) and future (\textbf{E})
-        cognitive function measurement prediction.
-        Red dots indicate how relative importances change (among taxa)
-        when demographics (sex, SES) are arred as covariables.
-        Abundances and prevalences of taxa that are important
-        for RF models both for concurrent (\textbf{F}) and future (\textbf{G}) assesment prediction.
-        }
-    \label{fig:3}
-\end{sidewaysfigure}
 
@@
 in the first year of life
-could predict \textit{future} cognitive performance measured (Figure 3B, D).
+could predict \textit{future} cognitive performance measured (Figure 3B, D; Table 4).
 Models were consistently able to learn microbial species that are associated
@@
 Some brain regions were more readily
-predicted by RF models trained on microbial taxa (Table 5, Table S6), in particular those that were highly correlated with age.
+predicted by RF models trained on microbial taxa (Tables 5, 6, and S6), in particular those that were highly correlated with age.
 These included the L/R lingual gyrus (mean RF correlations, Left =
@@
 computed after fitness-weighting on the reported brain segments (Figure 4D-E).
-
-\begin{sidewaysfigure}[!htb]
-    \centering
-    \small
-    \includegraphics[width=0.9\textwidth]{assets/Figure4.png}
-    \caption{
-        \textbf{Microbial features can train RF models to predict MSEL subscores and brain volumes in children over 18 months of age}.
-        (\textbf{A}) Test set correlations for RF models predicting MSEL subscale scores,
-        with or without demographics included as predictors.
-        (\textbf{B}) Important taxa in models predicting MSEL subscale scores.
-        (\textbf{C}) Average test-set correlations for prediction of MRI segmentation
-        data from microbiome and demographics on select segments after repeated
-        cross validation. (\textbf{D}) Heatmap of average individual relative taxa
-        importances on each brain segment. Importances are reported as
-        proportions relative to the sum of importances for each model - since
-        every model is trained with 132 features, and even distribution of
-        importance would be 0.75\% for each feature. Segments and taxa ordered
-        by HCA on a select list of species with high ranks on average importances,
-        and their respective highest-load segments (\textbf{E}).
-    }
-    \label{fig:4}
-\end{sidewaysfigure}
 
@@
 
+\subsection*{Figures}
+
+\begin{figure}[!h]
+    \centering
+    \small
+    % \includegraphics[width=0.9\textwidth]{assets/Figure1.png}
+    \caption{
+        \label{fig:1} \textbf{The RESONANCE cohort is a diverse cohort with longitudinal sampling.} (\textbf{A}) Stool samples,
+        cognitive assessments and neuroimaging were collected from participants
+        at different ages throughout the first years of life.
+        (\textbf{B}) Cognitive function scores are assessed by
+        different instruments, but may be normalized using full-scale composite scores.
+        (\textbf{C-D}) Principal coordinates analysis using Bray-Curtis dissimilarity on
+        taxonomic profiles demonstrates high beta diversity, with much of the
+        first axis of variation explained by increasing age and alpha diversity.
+        Differences in gene function profiles (\textbf{E}) and neuroimaging (PCA based on
+        the Euclidean distance of brain region volumes) (\textbf{F}) are likewise
+        dominated by changes as children age. (\textbf{G}) Permutation analysis of
+        variance (PERMANOVA) of taxonomic profiles, functional profiles
+        (annotated as UniRef90s) and neuroimaging, against the metadata of
+        interest. Variations in taxonomic and functional profiles explain a
+        modest but significant percent of the variation in cognitive development
+        in children over 18 months of age. (\textbf{H}) Mantel testing of different
+        microbial feature matrices, shows overlapping but distinct patterns of
+        variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
+        used as cut-offs in some following models.
+    }
+
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \small
+    % \includegraphics[width=0.9\textwidth]{assets/Figure2.png}
+    \caption{
+        \textbf{Taxa and gene functional groups are associated with cognitive function.}
+        (\textbf{A}) Volcano plot of multivariable linear model results showing the
+        relationship between individual taxa and cognitive function score in
+        children over 18 months of age (left) or across all ages (right),
+        controlling for age, maternal education, and sequencing depth.
+        Taxa that were significant after FDR correction (q
+        \textless{} 0.2) are colored red. (\textbf{B}) For taxa that were significantly
+        associated with cognitive function,
+        heatmaps of correlation with cognitive function in different age groups.
+        (\textbf{C}) Multivariate linear models showing the relationship
+        between individual taxa and domain-specific skill assessments
+        from the MSEL for children under 6m (left), 18m to 36m (right),
+        or over across all ages (right)
+        controlling for age, maternal education, and sequencing depth.
+        Asterisks in (A) and (C) indicate significance (q \textless{} 0.2)
+        after multiple hypothesis correction.
+        (\textbf{D-E}) Enrichment plots for proprionate degradation I gene set
+        used in feature set enrichment analysis (FSEA) in children under 6m (D)
+        or over 18m (E). (\textbf{G}) summary of FSEA
+        results in children under 6m of age (left)
+        over 18 months (middle) or across all ages (right), colored based on the
+        significance of the association. Markers indicate the individual
+        correlation of genes within a gene set,
+        and vertical bars colored bars represent the
+        median correlation of that gene set.
+        Vertical dotted lines within each panel represent the median T statistic
+        for all genes in that age group.
+    }
+    \label{fig:2}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \small
+    % \includegraphics[width=0.9\textwidth]{assets/Figure3.png}
+    \caption{
+        \textbf{Random forest models predict both concurrent and future measured cognitive function.}
+        Comparison of RF predictor importance versus linear models for children
+        over 18 months (\textbf{A}).
+        Colors represent whether the species belong to the group of
+        top-important features that account for 60\% of the cumulative
+        importance on the RF model, if that species was significant (q
+        \textless{} 0.2) in linear models, both, or neither.
+        Comparison of RF predictor importance for microbial features
+        between concurrent and future cognitive function measure
+        (\textbf{B}).
+        Colors represent whether the species belong to the group of
+        top-important features that account for 60\% of the cumulative
+        importance on the RF model only for concurrent prediction,
+        only for the future prediction, or for both models.
+        Features absent of one model are shown on the origin-aligned sideplots.
+        Distribution of ages for each microbiome sampling - cognitive assesment dyad (\textbf{C}).
+        Blue violin plot shows distribution of sampling age,
+        while red violin plot shows distribution of asessment age.
+        Black lines connect ages for each individual subject.      
+        Ranked feature importance for taxa in RF models for
+        both concurrent (\textbf{D}) and future (\textbf{E})
+        cognitive function measurement prediction.
+        Red dots indicate how relative importances change (among taxa)
+        when demographics (sex, SES) are arred as covariables.
+        Abundances and prevalences of taxa that are important
+        for RF models both for concurrent (\textbf{F}) and future (\textbf{G}) assesment prediction.
+        }
+    \label{fig:3}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \small
+    % \includegraphics[width=0.9\textwidth]{assets/Figure4.png}
+    \caption{
+        \textbf{Microbial features can train RF models to predict MSEL subscores and brain volumes in children over 18 months of age}.
+        (\textbf{A}) Test set correlations for RF models predicting MSEL subscale scores,
+        with or without demographics included as predictors.
+        (\textbf{B}) Important taxa in models predicting MSEL subscale scores.
+        (\textbf{C}) Average test-set correlations for prediction of MRI segmentation
+        data from microbiome and demographics on select segments after repeated
+        cross validation. (\textbf{D}) Heatmap of average individual relative taxa
+        importances on each brain segment. Importances are reported as
+        proportions relative to the sum of importances for each model - since
+        every model is trained with 132 features, and even distribution of
+        importance would be 0.75\% for each feature. Segments and taxa ordered
+        by HCA on a select list of species with high ranks on average importances,
+        and their respective highest-load segments (\textbf{E}).
+    }
+    \label{fig:4}
+\end{figure}
+
 \end{document}

### 2023-10-19 manuscript/main.tex
update figure and table legend titles

@@
     \end{tabular}
-    \caption{\label{tab:demo}Subject demographics of the ECHO RESONANCE cohort.
+    \caption{\label{tab:demo}\textbf{Subjects in the ECHO RESONANCE cohort come from a diverse cross-section of Rhode Island families.}
             "All" refers to the entire cohort.
@@
         Glutamate syn. II & 0.343 & 0.284 & 0.355 & \textbf{0.023} \\
-        Propionate deg. I & -0.659 & \textbf{0.029} & 0.154 & \color{red}{\textbf{0.178}} \\\hline
+        Propionate deg. I & -0.659 & \textbf{0.029} & 0.154 & \textbf{0.178} \\\hline
     \end{tabular}
@@
     \caption{
-        \label{fig:1} \textbf{The RESONANCE cohort is a diverse cohort with longitudinal sampling.} (\textbf{A}) Stool samples,
+        \label{fig:1} \textbf{The RESONANCE cohort is a diverse cohort with accelerated longitudinal sampling.} (\textbf{A}) Stool samples,
         cognitive assessments and neuroimaging were collected from participants
@@
     \caption{
-        \textbf{Microbial features can train RF models to predict MSEL subscores and brain volumes in children over 18 months of age}.
+        \textbf{Coupled importance of microbes in neuroanatomy and subscores are consistent with known subcortical associations with cognition.}
         (\textbf{A}) Test set correlations for RF models predicting MSEL subscale scores,

### 2023-10-19 manuscript/main.tex
add teaser
update dataset doi

@@
 
-% Teaser
-% This one-sentence summary provides a snapshot of your research for
-% non-specialist readers and will be used to promote your article if accepted for publication.
-% These should complement rather than repeat the title.
-
-% The microorganisms that colonize the gut of humans can influence
-% health and disease in numerous
+\subsection*{Teaser}
+
+The complex community of microorganisms living in the gastrointestinal tract of infants and children
+may affect how their brains develop.
 
@@
 demographics necessary for statistical analyses and machine learning are
-available on the Open Science Framework
-\cite{bonhamECHORESONANCEMicrobiome2022}.
+available on the Data Dryad
+\cite{bonhamDataDryad2023}.
 Data processing, generation of summary statistics, and

### 2023-10-19 manuscript/main.tex
rearrange figure panel labels

@@
 As expected, the greatest differences in microbial taxa were observed across
-child age (Figure 1C-D, PCoA axis 1), with older children primarily
+child age (Figure 1B, E; PCoA axis 1), with older children primarily
 stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
@@
 Overall subject-to-subject variation in gut microbial genes
-and brain volume profiles was similarly driven largely by subject age (Figure 1E-F).
+and brain volume profiles was similarly driven largely by subject age (Figure 1C, F).
 
@@
 that were important for discriminating cognitive function scores
-(Table 3, Figure 3A). Subject age was consistently ranked highly in
+(Table 3, Figure 3A-B, E). Subject age was consistently ranked highly in
 feature importance, which could indicate that decision branches based on
@@
 in the first year of life
-could predict \textit{future} cognitive performance measured (Figure 3B, D; Table 4).
+could predict \textit{future} cognitive performance measured at least 6 months
+after the stool sample was collected (Figure 3C-D, F; Table 4).
 Models were consistently able to learn microbial species that are associated
 with cognitive performance at a future visit
-(mean time between visits: 21.5 months, S.D. = 9.9 months; Figure 3C,
+(mean time between visits: 21.5 months, S.D. = 9.9 months; Figure 3G,
 mean test set correlation: 0.27).
@@
 in children older than 18 months were also important in
-models that predicted future cognitive function, if they were present (Figure 3B, Table S5).
+models that predicted future cognitive function, if they were present (Figure 3B-C, Table S5).
 At the same time, many additional species were important for future prediction models
@@
 were dominated by two species of \textit{Klebsiella},
-\textit{K. pneumoniae} and \textit{K. varicola} (Figure 3D, E).
+\textit{K. pneumoniae} and \textit{K. varicola} (Figure 3C, F).
 Many of the other species that ranked highest in feature importance

### 2023-10-20 manuscript/main.tex
change Figure-> fig.
hack up abstract

--- 2023-10-19 manuscript/main.tex
@@
 \begin{abstract}
-The gastrointestinal tract, its resident microorganisms, and the central
-nervous system are connected by biochemical signaling, also known as the
-"microbiome-gut-brain-axis." Both the human brain and the gut microbiome
-have critical developmental windows in the first years of life,
-raising the possibility that their development is co-occurring and
-likely co-dependent. Emerging evidence implicates gut microorganisms and
-microbiota composition in cognitive outcomes and neurodevelopmental
-disorders (e.g., autism and anxiety), but the influence of gut microbial
-metabolism on typical neurodevelopment has not been explored in detail.
-We investigated the relationship of the microbiome with the neuroanatomy
-and cognitive function of 381 healthy children, demonstrating that
-differences in gut microbial taxa and gene functions are associated with
-overall cognitive function and with differences in the size of
-multiple brain regions.
-Using a combination of multivariate linear and machine learning (ML) models,
-we showed that many species,
-including \emph{Alistipes obesi} and \emph{Blautia wexlerae}, were
-associated with higher cognitive function, while some
-species such as \emph{Ruminococcus gnavus} were more commonly found in
-children with low cognitive scores after controlling for sociodemographic factors.
-Microbial genes for enzymes involved in the metabolism of
-neuroactive compounds, particularly short-chain fatty acids such as
-acetate and propionate, were also associated with cognitive function.
-In addition, ML models were able to use microbial taxa to predict the
-volume of brain regions, and many taxa that were identified
-as important in predicting cognitive function also dominated the
-feature importance metric for individual brain regions,
-and for specific subscales of cognitive function.
-For example, \emph{B. wexlerae}
-was the most important species in models predicting the size of the parahippocampal region
-in both the left and right hemispheres
-and was among the top predictors of gross motor and expressive language performance.
-Several species from
-the phylum \textit{Bacteroidetes}, including GABA-producing
-\emph{B. ovatus}, were important for predicting
-the size of the left accumbens area, but not the right. These
-findings provide potential biomarkers of neurocognition and brain development
-and may lead to the future development of targets for early detection and early
-intervention.
+    Emerging evidence implicates gut microbial metabolism in neurodevelopmental disorders,
+    but its influence on typical neurodevelopment has not been explored in detail.
+    We investigated the relationship between the microbiome and neuroanatomy and cognition of 381 healthy children,
+    demonstrating that differences in microbial taxa and genes are associated
+    with overall cognitive function and the size of brain regions.
+    Using a combination of statistical and machine learning (ML) models,
+    we showed that species including \textit{Alistipes obesi}, \textit{Blautia wexlerae}, and \textit{Ruminococcus gnavus}
+    were enriched or depleted in children with higher cognitive function scores.
+    Microbial metabolism of short-chain fatty acids was also associated with cognitive function.
+    In addition, ML models were able to predict the volume of brain regions from microbial profiles,
+    and taxa that were important in predicting cognitive function
+    were also important for predicting individual brain regions and specific subscales of cognitive function.
+    These findings provide potential biomarkers of neurocognitive development
+    and may enable development of targets for early detection and intervention.
 \end{abstract}
@@
 
-The complex community of microorganisms living in the gastrointestinal tract of infants and children
+The complex community of microorganisms living in the guts of infants and children
 may affect how their brains develop.
@@
 381 healthy and neurotypically-developing children (172 female) between 40 days and 10 years of age
-(Table 1,  Figure 1A, Figure S1)
+(table 1, fig. 1A, fig. S1)
 using a variety of orthogonal microbial and neurocognitive measures.
@@
 (i.e., the MSEL for children under 4 years of age, The WPPSI-3 for 3-6 year-olds,
-and the WISC-V for children older than 6), but were normalized to a common scale (Figure 1B),
+and the WISC-V for children older than 6), but were normalized to a common scale (fig. 1B),
 As expected, the greatest differences in microbial taxa were observed across
-child age (Figure 1C-D, PCoA axis 1), with older children primarily
+child age (fig. 1B, E; PCoA axis 1), with older children primarily
 stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
-high-abundance \emph{Prevotella copri} (Figure S2).
+high-abundance \emph{Prevotella copri} (fig. S2).
 Overall subject-to-subject variation in gut microbial genes
-and brain volume profiles was similarly driven largely by subject age (Figure 1E-F).
+and brain volume profiles was similarly driven largely by subject age (fig. 1C, F).
 
@@
 However, even with this categorization, we find increasing variation
-in gut microbiomes in children over 18 months old (Figure 1G).
+in gut microbiomes in children over 18 months old (fig. 1G).
 We also found that overall variation in microbial species in children over 18 months old
 was significantly associated with variation in cognitive function score
-(Figure 1G, R\textsuperscript{2} = 0.0099, q \textless{} 0.01),
+(fig. 1G, R\textsuperscript{2} = 0.0099, q \textless{} 0.01),
 as was variation in microbial gene functions (R\textsuperscript{2} = 0.0115, q \textless{} 0.001).
@@
 function in children under 6 months, though this may be due to the low
-taxonomic diversity and broad lack of overlap between taxa in infants (Figure 1G).
+taxonomic diversity and broad lack of overlap between taxa in infants (fig. 1G).
 As expected, age was significantly associated with microbial beta
@@
 profiles, both generated from metagenomic sequences, being tightly coupled
-(Figure 1H, p \textless{} 0.001). Other functional
+(fig. 1H, p \textless{} 0.001). Other functional
 groupings (Enzyme commission level4 - ECs
@@
 for the relative abundance of each species that had at least 15\%
-prevalence in a given age group (Figure 2A, N\textsubscript{species} = 116 for 0--120 months, 
+prevalence in a given age group (fig. 2A, N\textsubscript{species} = 116 for 0--120 months, 
 N\textsubscript{species} = 54 for 0--6 months, N\textsubscript{species} = 136 for 18--120 months).
@@
 several mirobial species were significantly enriched
-in children with higher cognitive function scores (Figure 2B),
+in children with higher cognitive function scores (fig. 2B),
 including \emph{Alistipes obesi},
@@
 \cite{ghoshMediterraneanDietIntervention2020}
-(Figure 2B). \textit{Sutterella wadsworthensis} was the only microbial species
+(fig. 2B). \textit{Sutterella wadsworthensis} was the only microbial species
 that was significantly negatively associated with cognitive function score (q value = 0.165).
@@
 of samples for children under 6 months, 18 to 36 months, or all ages
-with a concurrent MSEL score (Figure 2C).
-Several taxa associated with composite scores (Figure 2B)
+with a concurrent MSEL score (fig. 2C).
+Several taxa associated with composite scores (fig. 2B)
 were also associated with one or more subscores, including
@@
 \cite{valles-colomerNeuroactivePotentialHuman2019}
-and concurrent cognitive ability score (Table 2, Figure 2D-F) and found that
+and concurrent cognitive ability score (table 2, fig. 2D-F) and found that
 several metabolic pathways were either significantly enriched or depleted in
@@
 were significantly depleted in children with higher cognitive ability
-scores in children under 6m (Table 2; propionate degradation I,
+scores in children under 6m (table 2; propionate degradation I,
 enrichment score (E.S.) = -0.66, corrected p-value (q)
@@
 sex, and age were included as possible predictors, either alone
-or in combination with microbial taxonomic profiles (Table 3).
+or in combination with microbial taxonomic profiles (table 3).
 
@@
 Species that were important in RF overlapped with those that were
-significant when testing the relationship with LMs (Figure
-3A-B, Table S1A-B); Both analysis identified 
+significant when testing the relationship with LMs (fig.
+3A-B, table S1A-B); Both analysis identified 
 \textit{F. prausnitzii}, % F. prausz was mentioned before, ~ line 395
@@
 that were important for discriminating cognitive function scores
-(Table 3, Figure 3A). Subject age was consistently ranked highly in
+(table 3, fig. 3A-B, E). Subject age was consistently ranked highly in
 feature importance, which could indicate that decision branches based on
@@
 in the first year of life
-could predict \textit{future} cognitive performance measured (Figure 3B, D; Table 4).
+could predict \textit{future} cognitive performance measured at least 6 months
+after the stool sample was collected (fig. 3C-D, F; table 4).
 Models were consistently able to learn microbial species that are associated
 with cognitive performance at a future visit
-(mean time between visits: 21.5 months, S.D. = 9.9 months; Figure 3C,
+(mean time between visits: 21.5 months, S.D. = 9.9 months; fig. 3G,
 mean test set correlation: 0.27).
@@
 in children older than 18 months were also important in
-models that predicted future cognitive function, if they were present (Figure 3B, Table S5).
+models that predicted future cognitive function, if they were present (fig. 3B-C, table S5).
 At the same time, many additional species were important for future prediction models
@@
 were dominated by two species of \textit{Klebsiella},
-\textit{K. pneumoniae} and \textit{K. varicola} (Figure 3D, E).
+\textit{K. pneumoniae} and \textit{K. varicola} (fig. 3C, F).
 Many of the other species that ranked highest in feature importance
@@
 Taxonomic profiles coupled with demographics exhibited predictive potential towards three
-of five probed subscales (Figure 4A, Table 5), 
+of five probed subscales (fig. 4A, table 5), 
 namely expressive language ( test-set RMSE 12.05, test-set correlation 0.368), 
@@
 Between the subscales, both the predictability (measured by the test-set figures of merit)
-and the relative importances of several taxa (Figure 4B) is variable.
+and the relative importances of several taxa (fig. 4B) is variable.
 Although certain species such as \textit{B. longum}, \textit{F. prausznitzii} 
-have similar relative importance across all subscales, others are substantially different (Tables S2-S4).
+have similar relative importance across all subscales, others are substantially different (tables S2-S4).
 Notably, \textit{Bifidobacterium pseudocatenulatum}, \textit{B. wexlerae} and \textit{E. eligens} are
@@
 \textit{Clostridium innocuum}, \textit{B. vulgatus} 
-stand out on models targeting visual reception (Supplenmentary Tables S2-4).
+stand out on models targeting visual reception (tables S2-4).
 
@@
 Some brain regions were more readily
-predicted by RF models trained on microbial taxa (Tables 5, 6, and S6), in particular those that were highly correlated with age.
+predicted by RF models trained on microbial taxa (tables 5, 6, and S6), in particular those that were highly correlated with age.
 These included the L/R lingual gyrus (mean RF correlations, Left =
@@
     of test-set preditions are reported accross models for all brain regions.
-    Complete list of values can be found on Table S6.}
+    Complete list of values can be found on table S6.}
     \end{center}
@@
 was responsible for half of the cumulative importance,
-regardless of model performance (Figure 4C-E, Table S2).
+regardless of model performance (fig. 4C-E, table S2).
 We observed two major patterns of importance distribution from
@@
 \emph{B. wexlerae}, which combined, account for approximately 10\% of the cumulative relative importances,
-computed after fitness-weighting on the reported brain segments (Figure 4D-E).
+computed after fitness-weighting on the reported brain segments (fig. 4D-E).
 
@@
 highest average relative importance along all reported models
-(mean = 1.4\%; min = 0.8\%; max = 3.6\%, Figure 4D-E).
+(mean = 1.4\%; min = 0.8\%; max = 3.6\%, fig. 4D-E).
 \emph{B. wexlerae}, which was important in RF models of cognitive function, 
@@
 In line with this hypothesis, RF models that were able to learn associations
-between early-life microbiomes and future cognitive function scores (Figure 3).
+between early-life microbiomes and future cognitive function scores (fig. 3).
 The top-ranked species for these models included several 
@@
 Clostridiales, as is \emph{R. gnavus}, which we found was negatively
-associated with cognitive function (Figure 2A-B). This kind of
+associated with cognitive function (fig. 2A-B). This kind of
 species-level resolution is typically not possible with amplicon
@@
 genes for the metabolism of SCFAs, both their degradation and synthesis,
-are associated with cognitive function scores (Figure 2). 
+are associated with cognitive function scores (fig. 2). 
 SCFAs are produced by anaerobic fermentation of dietary fiber
@@
 \cite{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
-(Figure S3).
+(fig. S3).
 
@@
 The cognitive assessments were collected as either composite scores
-(for the experiments in Figure 3) or, when available, subscores on the
-Mullen Scales of Early Learning (Figure 4 A-B: expressive language, gross motor and visual reception).
+(for the experiments in fig. 3) or, when available, subscores on the
+Mullen Scales of Early Learning (fig. 4 A-B: expressive language, gross motor and visual reception).
 For the concurrent prediction experiments, when multiple viable longitudinal samples were present,
@@
 the latest future cognitive assessment on the assesment time range.
-Age (in months) was provided as a covariate for all models (Table S3).
+Age (in months) was provided as a covariate for all models (table S3).
 

### 2023-10-20 manuscript/main.tex
remove abreviation in abstract

@@
     with overall cognitive function and the size of brain regions.
-    Using a combination of statistical and machine learning (ML) models,
+    Using a combination of statistical and machine learning models,
     we showed that species including \textit{Alistipes obesi}, \textit{Blautia wexlerae}, and \textit{Ruminococcus gnavus}
@@
     Microbial metabolism of short-chain fatty acids was also associated with cognitive function.
-    In addition, ML models were able to predict the volume of brain regions from microbial profiles,
+    In addition, machine models were able to predict the volume of brain regions from microbial profiles,
     and taxa that were important in predicting cognitive function

### 2023-10-20 manuscript/main.tex
Guilherme comments
short title and rearrange post-text

@@
 
-\subsection*{Teaser}
+\subsubsection*{Short title}
+
+Gut microbes are linked to child neurocognition
+
+\subsubsection*{Teaser}
 
@@
 (i.e., the MSEL for children under 4 years of age, The WPPSI-3 for 3-6 year-olds,
-and the WISC-V for children older than 6), but were normalized to a common scale (fig. 1B),
+and the WISC-V for children older than 6), but were normalized to a common scale (fig. 1D),
 As expected, the greatest differences in microbial taxa were observed across
@@
 significant when testing the relationship with LMs (fig.
-3A-B, table S1A-B); Both analysis identified 
+3A, table S1); Both analysis identified 
 \textit{F. prausnitzii}, % F. prausz was mentioned before, ~ line 395
@@
 that were important for discriminating cognitive function scores
-(table 3, fig. 3A-B, E). Subject age was consistently ranked highly in
+(table 3, fig. 3B, E). Subject age was consistently ranked highly in
 feature importance, which could indicate that decision branches based on
@@
 Clostridiales, as is \emph{R. gnavus}, which we found was negatively
-associated with cognitive function (fig. 2A-B). This kind of
+associated with cognitive function (fig. 3C, F). This kind of
 species-level resolution is typically not possible with amplicon
@@
 
-\subsubsection*{Data and materials availability}
-
-All data needed to evaluate the conclusions in the paper
-are present in the paper and/or the Supplementary Materials.
-Taxonomic and functional microbial profiles, as well as subject
-demographics necessary for statistical analyses and machine learning are
-available on the Data Dryad
-\cite{bonhamDataDryad2023}.
-Data processing, generation of summary statistics, and
-generation of plots was performed using the julia programming language
-and 3rd party packages
-\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021,blaomMLJJuliaPackage2020,ben_sadeghi_2022_7359268,dahua_lin_2023_7695673,douglas_bates_2023_7734970}.
-Information for replicating the package environment
-as well as code for data analysis and figure generation, as well as scripts for automated
-download of input files are available on GitHub at https://github.com/KlepacCerajLab/Resonance,
-and archived on Zenodo \cite{kevin_bonham_2023_7647510}.
-
-\subsubsection*{Funding}
+\section*{Funding}
 This research was funded by NIH UG3 OD023313 and Wellcome LEAP 1kD.
 
-\subsubsection*{Competing Interests}
-The authors declare that they have no competing interests.
-
-\subsubsection*{Author contributions}
+\section*{Author contributions}
 % https://credit.niso.org/
@@
 
-\subsection*{Figures}
+\section*{Competing Interests}
+The authors declare that they have no competing interests.
+\section*{Figures}
+
+\section*{Data and materials availability}
+
+All data needed to evaluate the conclusions in the paper
+are present in the paper and/or the Supplementary Materials.
+Taxonomic and functional microbial profiles, as well as subject
+demographics necessary for statistical analyses and machine learning are
+available on the Data Dryad
+\cite{bonhamDataDryad2023}.
+Data processing, generation of summary statistics, and
+generation of plots was performed using the julia programming language
+and 3rd party packages
+\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021,blaomMLJJuliaPackage2020,ben_sadeghi_2022_7359268,dahua_lin_2023_7695673,douglas_bates_2023_7734970}.
+Information for replicating the package environment
+as well as code for data analysis and figure generation, as well as scripts for automated
+download of input files are available on GitHub at https://github.com/KlepacCerajLab/Resonance,
+and archived on Zenodo \cite{kevin_bonham_2023_7647510}.
 

### 2023-10-23 manuscript/main.tex

@@
         Rosa Cano Lorente,$^{2}$
-        The RESONANCE Consortium,
+        The RESONANCE Consortium$^{\ast}$,
         Curtis Huttenhower,$^{4}$
@@
         Sean Deoni,$^{2}$
-        Vanja Klepac-Ceraj$^{1\ast}$
+        Vanja Klepac-Ceraj$^{1\dagger}$
     }
@@
 \\
-\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  vklepacc@wellesley.edu.}
+\normalsize{$^\ast$The RESONANCE consortium collaborators and affiliations are listed in the supplementary materials.}\\
+\normalsize{$^\dagger$To whom correspondence should be addressed; E-mail:  vklepacc@wellesley.edu.}
 }

### 2023-11-13 manuscript/main.tex
more more more

@@
 style=numeric,
-maxbibnames=50
+maxbibnames=100
 ]{biblatex}
@@
 
+Members of the RESONANCE consortium also contributed to
+project planning, conceptualization, and funding acquisition.
+
 \section*{Competing Interests}
@@
 as well as code for data analysis and figure generation, as well as scripts for automated
-download of input files are available on GitHub at https://github.com/KlepacCerajLab/Resonance,
+download of input files are available on GitHub at https://github.com/Klepac-Ceraj-Lab/Resonance,
 and archived on Zenodo \cite{kevin_bonham_2023_7647510}.

### 2023-11-14 manuscript/main.tex
update curtis refs

@@
         The RESONANCE Consortium$^{\ast}$,
-        Curtis Huttenhower,$^{4}$
+        Curtis Huttenhower,$^{4-7}$
         Muriel Bruchhage,$^{3}$
@@
 \normalsize{$^{4}$Department of Biostatistics, Harvard T.H. Chan School of Public Health, Boston, MA, USA}\\
+\normalsize{$^{5}$Department of Immunology and Infectious Diseases, Harvard T.H. Chan School of Public Health, Boston, MA, USA}\\
+\normalsize{$^{6}$Harvard Chan Microbiome in Public Health Center, Harvard T.H. Chan School of Public Health, Boston, MA, USA}\\
+\normalsize{$^{7}$Associate Member, Broad Institute of MIT and Harvard}\\
 \\

### 2023-11-21 manuscript/main.tex

@@
         Rosa Cano Lorente,$^{2}$
-        The RESONANCE Consortium$^{\ast}$,
+        The RESONANCE Consortium$^{1\dagger}$,
         Curtis Huttenhower,$^{4-7}$
@@
         Sean Deoni,$^{2}$
-        Vanja Klepac-Ceraj$^{1\dagger}$
+        Vanja Klepac-Ceraj$^{\ast}$
     }
@@
 \\
-\normalsize{$^\ast$The RESONANCE consortium collaborators and affiliations are listed in the supplementary materials.}\\
-\normalsize{$^\dagger$To whom correspondence should be addressed; E-mail:  vklepacc@wellesley.edu.}
+\normalsize{$^\dagger$The RESONANCE consortium authors and collaborators and affiliations are listed in the supplementary materials.}\\
+\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  vklepacc@wellesley.edu.}\\
 }
@@
 metagenomes.
+Statistical analyses, including linear regression
+and feature set enrichment analysis were performed
+using the julia programming language and third-party packages
+\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021,blaomMLJJuliaPackage2020,ben_sadeghi_2022_7359268,dahua_lin_2023_7695673,douglas_bates_2023_7734970}.
 
@@
 generation of plots was performed using the julia programming language
-and 3rd party packages
-\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021,blaomMLJJuliaPackage2020,ben_sadeghi_2022_7359268,dahua_lin_2023_7695673,douglas_bates_2023_7734970}.
+and 3rd party packages (see Materials and Methods).
 Information for replicating the package environment

### 2024-02-22 manuscript/submission.tex
permissions

--- 2023-11-13 manuscript/main.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage{enumitem}
 
@@
 backend=biber,
-style=numeric,
-maxbibnames=100
+style=science
 ]{biblatex}
 
-\addbibresource{citations.bib}
+\addbibresource{zotero.bib}
 
@@
         Rosa Cano Lorente,$^{2}$
-        The RESONANCE Consortium$^{\ast}$,
+        The RESONANCE Consortium,
         Curtis Huttenhower,$^{4}$
@@
         Sean Deoni,$^{2}$
-        Vanja Klepac-Ceraj$^{1\dagger}$
+        Vanja Klepac-Ceraj$^{1\ast}$
     }
@@
 \\
-\normalsize{$^\ast$The RESONANCE consortium collaborators and affiliations are listed in the supplementary materials.}\\
-\normalsize{$^\dagger$To whom correspondence should be addressed; E-mail:  vklepacc@wellesley.edu.}
+\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  vklepacc@wellesley.edu.}
 }
@@
 \begin{abstract}
-    Emerging evidence implicates gut microbial metabolism in neurodevelopmental disorders,
-    but its influence on typical neurodevelopment has not been explored in detail.
-    We investigated the relationship between the microbiome and neuroanatomy and cognition of 381 healthy children,
-    demonstrating that differences in microbial taxa and genes are associated
-    with overall cognitive function and the size of brain regions.
-    Using a combination of statistical and machine learning models,
-    we showed that species including \textit{Alistipes obesi}, \textit{Blautia wexlerae}, and \textit{Ruminococcus gnavus}
-    were enriched or depleted in children with higher cognitive function scores.
-    Microbial metabolism of short-chain fatty acids was also associated with cognitive function.
-    In addition, machine models were able to predict the volume of brain regions from microbial profiles,
-    and taxa that were important in predicting cognitive function
-    were also important for predicting individual brain regions and specific subscales of cognitive function.
-    These findings provide potential biomarkers of neurocognitive development
-    and may enable development of targets for early detection and intervention.
+The gastrointestinal tract, its resident microorganisms, and the central
+nervous system are connected by biochemical signaling, also known as the
+"microbiome-gut-brain-axis." Both the human brain and the gut microbiome
+have critical developmental windows in the first years of life,
+raising the possibility that their development is co-occurring and
+likely co-dependent. Emerging evidence implicates gut microorganisms and
+microbiota composition in cognitive outcomes and neurodevelopmental
+disorders (e.g., autism and anxiety), but the influence of gut microbial
+metabolism on typical neurodevelopment has not been explored in detail.
+We investigated the relationship of the microbiome with the neuroanatomy
+and cognitive function of 361 healthy children, demonstrating that
+differences in gut microbial taxa and gene functions are associated with
+overall cognitive function and with differences in the size of
+multiple brain regions.
+Using a combination of multivariate linear and machine learning (ML) models,
+we showed that many species,
+including \emph{Gordonibacter pamelae} and \emph{Blautia wexlerae}, were
+significantly associated with higher cognitive function, while some
+species such as \emph{Ruminococcus gnavus} were more commonly found in
+children with low cognitive scores after controlling for sociodemographic factors.
+Microbial genes for enzymes involved in the metabolism of
+neuroactive compounds, particularly short-chain fatty acids such as
+acetate and propionate, were also associated with cognitive function.
+In addition, ML models were able to use microbial taxa to predict the
+volume of brain regions, and many taxa that were identified
+as important in predicting cognitive function also dominated the
+feature importance metric for individual brain regions. For example, \emph{B.
+wexlerae} was the most important species in models predicting the size of the parahippocampal region
+in both the left and right hemispheres, while several species from
+the phylum \textit{Bacteroidetes}, including GABA-producing
+\emph{B. ovatus}, were important for predicting
+the size of the left accumbens area, but not the right. These
+findings provide potential biomarkers of neurocognition and brain development
+and may lead to the future development of targets for early detection and early
+intervention.
 \end{abstract}
 
-\subsubsection*{Short title}
-
-Gut microbes are linked to child neurocognition
-
-\subsubsection*{Teaser}
-
-The complex community of microorganisms living in the guts of infants and children
-may affect how their brains develop.
+% Teaser
+% This one-sentence summary provides a snapshot of your research for
+% non-specialist readers and will be used to promote your article if accepted for publication.
+% These should complement rather than repeat the title.
+
+% The microorganisms that colonize the gut of humans can influence
+% health and disease in numerous
 
@@
 Many of these bacteria have specialized metabolic capabilities
-for digesting human milk, such as \emph{Bifidobacterium longum}
+for digesting human breast milk, such as \emph{Bifidobacterium longum}
 subsp. \textit{infantis} and \emph{Bacteroides fragilis}
@@
 
-\subsubsection*{The RESONANCE cohort is part of a study of child development}
+\subsubsection*{Cohort overview and summary data}
 
 We investigated the co-development of the brain and microbiome in
-381 healthy and neurotypically-developing children (172 female) between 40 days and 10 years of age
-(table 1, fig. 1A, fig. S1)
+361 healthy and neurotypically-developing children (163 female) between 82 days and 10 years of age
+(Table 1,  Figure 1A, Figure S1)
 using a variety of orthogonal microbial and neurocognitive measures.
@@
 (i.e., the MSEL for children under 4 years of age, The WPPSI-3 for 3-6 year-olds,
-and the WISC-V for children older than 6), but were normalized to a common scale (fig. 1D),
+and the WISC-V for children older than 6), but were normalized to a common scale (Figure 1B),
 As expected, the greatest differences in microbial taxa were observed across
-child age (fig. 1B, E; PCoA axis 1), with older children primarily
+child age (Figure 1C-D, PCoA axis 1), with older children primarily
 stratified into Bacteroidetes-dominant, Firmicutes-dominant, or
-high-abundance \emph{Prevotella copri} (fig. S2).
+high-abundance \emph{Prevotella copri} (Figure S2).
 Overall subject-to-subject variation in gut microbial genes
-and brain volume profiles was similarly driven largely by subject age (fig. 1C, F).
+and brain volume profiles was similarly driven largely by subject age (Figure 1E-F).
+
+\begin{sidewaysfigure}[!h]
+    \centering
+    \includegraphics[width=0.9\textwidth]{assets/Figure1.png}
+    \caption{
+        \label{fig:1} Study design and overview. (\textbf{A}) Stool samples,
+        cognitive assessments and neuroimaging were collected from participants
+        at different ages throughout the first years of life.
+        (\textbf{B}) Cognitive function scores are assessed by
+        different instruments, but may be normalized using full-scale composite scores.
+        (\textbf{C-D}) Principal coordinates analysis using Bray-Curtis dissimilarity on
+        taxonomic profiles demonstrates high beta diversity, with much of the
+        first axis of variation explained by increasing age and alpha diversity.
+        Differences in gene function profiles (\textbf{E}) and neuroimaging (PCA based on
+        the Euclidean distance of brain region volumes) (\textbf{F}) are likewise
+        dominated by changes as children age. (\textbf{G}) Permutation analysis of
+        variance (PERMANOVA) of taxonomic profiles, functional profiles
+        (annotated as UniRef90s) and neuroimaging, against the metadata of
+        interest. Variations in taxonomic and functional profiles explain a
+        modest but significant percent of the variation in cognitive development
+        in children over 18 months of age. (\textbf{H}) Mantel testing of different
+        microbial feature matrices, shows overlapping but distinct patterns of
+        variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
+        used as cut-offs in some following models.
+    }
+
+\end{sidewaysfigure}
 
@@
     \centering
-    \begin{tabular}{|r|r|r|r|r|r|}
+    \begin{tabular}{|r|r|r|r|r|}
       \hline\hline
-      \textbf{Group}              & \textbf{Subgroup}        & \textbf{All} & \textbf{Under 6 months} & \textbf{Over 18 months} & \textbf{Future} \\\hline
-      N subjects                  &                          & 381 & 82 & 277 & 104 \\  \hline
-    \multirow{3}{*}{Samples}      & 1                        & 214 (56.2\%) &  &  &  \\ \cline{2-6}
-                                  & 2                        & 90 (23.6\%) &  &  &  \\ \cline{2-6}
-                                  & \textgreater 2           & 77 (20.2\%) &  &  &  \\ \hline
-    \multirow{3}{*}{Age (months)} & min                      & 1.3 & 2.89 & 18.03 & 1.3 \\ \cline{2-6}
-                                  & max                      & 119.53 & 5.95 & 119.53 & 11.95 \\ \cline{2-6}
-                                  & median                   & 24.62 & 3.64 & 41.59 & 6.08 \\ \hline
-    \multirow{2}{*}{Sex}          & F                        & 172 (45.1\%) & 42 (51.2\%) & 122 (44.0\%) & 47 (45.2\%) \\   \cline{2-6}
-                                  & M                        & 209 (54.9\%) & 40 (48.8\%) & 155 (56.0\%) & 57 (54.8\%) \\  \hline
-    \multirow{6}{*}{Race}         & White                    & 265 (69.6\%) & 48 (58.5\%) & 201 (72.6\%) & 62 (59.6\%) \\   \cline{2-6}
-                                  & Black                    & 40 (10.5\%) & 15 (18.3\%) & 21 (7.6\%) & 17 (16.3\%) \\ \cline{2-6}
-                                  & Indigenous               & 1 (0.3\%) & 0 (0.0\%) & 1 (0.4\%) & 0 (0.0\%) \\ \cline{2-6}
-                                  & Asian                    & 8 (2.1\%) & 2 (2.4\%) & 7 (2.5\%) & 3 (2.9\%) \\ \cline{2-6}
-                                  & Mixed                    & 54 (14.2\%) & 14 (17.1\%) & 38 (13.7\%) & 16 (15.4\%) \\ \cline{2-6}
-                                  & Other                    & 3 (0.8\%) & 1 (1.2\%) & 3 (1.1\%) & 2 (1.9\%) \\ \hline
-    \multirow{6}{*}{Maternal Ed.} & Junior high school       & 2 (0.5\%) & 0 (0.0\%) & 2 (0.7\%) & 0 (0.0\%) \\   \cline{2-6}
-                                  & Some high school         & 6 (1.6\%) & 3 (3.7\%) & 4 (1.4\%) & 2 (1.9\%) \\ \cline{2-6}
-                                  & High school grad         & 39 (10.2\%) & 13 (15.9\%) & 22 (7.9\%) & 17 (16.3\%) \\ \cline{2-6}
-                                  & Some college             & 100 (26.2\%) & 28 (34.1\%) & 60 (21.7\%) & 30 (28.8\%) \\ \cline{2-6}
-                                  & College grad             & 95 (24.9\%) & 19 (23.2\%) & 75 (27.1\%) & 30 (28.8\%) \\ \cline{2-6}
-                                  & Grad/professional school & 126 (33.1\%) & 16 (19.5\%) & 106 (38.3\%) & 23 (22.1\%) \\ \hline\hline
+      \textbf{Group} & \textbf{Subgroup} & \textbf{All} & \textbf{Under 6 months} & \textbf{Over 18 months} \\\hline
+      N subjects &  & 361 & 74 & 257 \\  \hline
+    \multirow{3}{*}{Samples} & 1 & 262 (72.6\%) & 74 (100.0\%) & 203 (79.0\%) \\ \cline{2-5}
+         & 2 & 72 (19.9\%) & 0 (0.0\%) & 46 (17.9\%) \\ \cline{2-5}
+         & \textgreater 2 & 27 (7.5\%) & 0 (0.0\%) & 8 (3.1\%) \\ \hline
+    \multirow{3}{*}{Age (months)} & min & 2.76 & 2.76 & 18.03 \\ \cline{2-5}
+         & max & 119.3 & 6.0 & 119.3 \\ \cline{2-5}
+         & median & 25.79 & 3.66 & 45.92 \\ \hline
+    \multirow{2}{*}{Sex} & F & 163 (45.2\%) & 37 (50.0\%) & 114 (44.4\%) \\   \cline{2-5}
+                         & M & 198 (54.8\%) & 37 (50.0\%) & 143 (55.6\%) \\  \hline
+    \multirow{6}{*}{Race} & White & 252 (69.8\%) & 42 (56.8\%) & 191 (74.3\%) \\   \cline{2-5}
+        & Black & 40 (11.1\%) & 16 (21.6\%) & 21 (8.2\%) \\ \cline{2-5}
+        & Indiginous & 1 (0.3\%) & 0 (0.0\%) & 1 (0.4\%) \\ \cline{2-5}
+        & Asian & 7 (1.9\%) & 1 (1.4\%) & 6 (2.3\%) \\ \cline{2-5}
+        & Mixed & 48 (13.3\%) & 12 (16.2\%) & 29 (11.3\%) \\ \cline{2-5}
+        & Other & 2 (0.6\%) & 1 (1.4\%) & 2 (0.8\%) \\ \hline
+    \multirow{6}{*}{Maternal Ed.} & Junior high school & 1 (0.3\%) & 0 (0.0\%) & 1 (0.4\%) \\   \cline{2-5}
+        & Some high school & 6 (1.7\%) & 3 (4.1\%) & 4 (1.6\%) \\ \cline{2-5}
+        & High school grad & 37 (10.2\%) & 13 (17.6\%) & 16 (6.2\%) \\ \cline{2-5}
+        & Some college & 95 (26.3\%) & 25 (33.8\%) & 57 (22.2\%) \\ \cline{2-5}
+        & College grad & 90 (24.9\%) & 17 (23.0\%) & 69 (26.8\%) \\ \cline{2-5}
+        & Grad/professional school & 123 (34.1\%) & 15 (20.3\%) & 104 (40.5\%) \\\hline\hline
     \end{tabular}
-    \caption{\label{tab:demo}\textbf{Subjects in the ECHO RESONANCE cohort come from a diverse cross-section of Rhode Island families.}
-            "All" refers to the entire cohort.
-            "Under 6 months", "Over 18 months", and "Future" refer specifically
-            to samples that were used included in the respective subgroup analyses.
-    }
+    \caption{\label{tab:demo}Subject demographics of the ECHO RESONANCE cohort.}
 \end{table}
@@
 and atypical neurocognition and neurological disorders
-\cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021,magnussonRelationshipsDietrelatedChanges2015,mayneris-perxachsMicrobiotaAlterationsProline2022,needhamGutderivedMetaboliteAlters2022},
+\cite{liuAlteredGutMicrobiota2019,wanUnderdevelopmentGutMicrobiota2021,
+      magnussonRelationshipsDietrelatedChanges2015,mayneris-perxachsMicrobiotaAlterationsProline2022,
+      needhamGutderivedMetaboliteAlters2022},
 we sought to determine if specific taxa or gene functions were associated
@@
 However, even with this categorization, we find increasing variation
-in gut microbiomes in children over 18 months old (fig. 1G).
+in gut microbiomes in children over 18 months old (Figure 1G).
 We also found that overall variation in microbial species in children over 18 months old
 was significantly associated with variation in cognitive function score
-(fig. 1G, R\textsuperscript{2} = 0.0099, q \textless{} 0.01),
-as was variation in microbial gene functions (R\textsuperscript{2} = 0.0115, q \textless{} 0.001).
+(Figure 1G, R\textsuperscript{2} = 0.0124, q \textless{} 0.001),
+as was variation in microbial gene functions (R\textsuperscript{2} = 0.0119, q \textless{} 0.001).
 Variation in microbial taxa and genes was not significantly associated with cognitive
 function in children under 6 months, though this may be due to the low
-taxonomic diversity and broad lack of overlap between taxa in infants (fig. 1G).
+taxonomic diversity and broad lack of overlap between taxa in infants (Figure 1G).
 As expected, age was significantly associated with microbial beta
-diversity (taxa R\textsuperscript{2} = 0.0207, and  gene functions
+diversity (taxa R\textsuperscript{2} = 0.0215, and  gene functions
 annotated with UniRef90s clusters of 90\% similarity
 \cite{suzekUniRefComprehensiveNonredundant2007} -
-R\textsuperscript{2} = 0.0258, q \textless{} 0.001), and very strongly
-associated with variation in neuroimaging (MRI) profiles (R\textsuperscript{2} = 0.258, q
+R\textsuperscript{2} = 0.0245, q \textless{} 0.001), and very strongly
+associated with variation in neuroimaging profiles (R\textsuperscript{2} = 0.256, q
 \textless{} 0.001).
@@
 profiles, both generated from metagenomic sequences, being tightly coupled
-(fig. 1H, p \textless{} 0.001). Other functional
+(Figure 1H, p \textless{} 0.001). Other functional
 groupings (Enzyme commission level4 - ECs
@@
 cohorts, despite being derived from UniRef90 labels. In children over 18
-months, some variation (taxa - 20.1\%, p \textless{} 0.01;
-gene functions - 15.9\%, p \textless{} 0.05) in neuroimaging
+months, some variation (15.9\%, p \textless{} 0.01) in neuroimaging
 overlapped with microbial measures, though this may be due to the
@@
 for the relative abundance of each species that had at least 15\%
-prevalence in a given age group (fig. 2A, N\textsubscript{species} = 116 for 0--120 months, 
-N\textsubscript{species} = 54 for 0--6 months, N\textsubscript{species} = 136 for 18--120 months).
-No species were significantly associated (q \textless{} 0.20) with
+prevalence in a given age group (Figure 2A, N\textsubscript{species} = 92 for 0--120 months, 
+N\textsubscript{species} = 46 for 0--6 months, N\textsubscript{species} = 97 for 18--120 months).
+Only \emph{Blautia wexlerae} was significantly associated (q value = 0.14, $\beta$ = 0.0015) with
 cognitive function in children under 6 months old after adjusting for
-age and maternal education.
+age and maternal education (Figure 2B).
 By contrast, in children over 18 months of age,
-several mirobial species were significantly enriched
-in children with higher cognitive function scores (fig. 2B),
-including \emph{Alistipes obesi},
-\emph{Asaccharobacter celatus} (also known as \textit{Adlercreutzia equolifaciens} subsp. \textit{celatus}
+several mirobial species were significantly enriched (q-value
+\textless{} 0.20) in children with higher cognitive function scores,
+including \emph{Gordonibacter pamelaeae},
+\emph{Asaccharobacter celatus} and \emph{Adelcreutzia equolifaciens}
+(also known as \textit{Adlercreutzia equolifaciens} subsp. \textit{celatus}
+and \textit{Adlercreutzia equolifaciens} subsp. \textit{equolifaciens} respectively
 \cite{maruoAdlercreutziaEquolifaciensGen2008}),
@@
 \cite{ghoshMediterraneanDietIntervention2020}
-(fig. 2B). \textit{Sutterella wadsworthensis} was the only microbial species
-that was significantly negatively associated with cognitive function score (q value = 0.165).
-
-Cognitive assessments (MSEL, WPPSI, WISC)
-are composites of multiple subscales that assess different cognitive domains.
-While the composite scores are broadly analagous accross ages and test modality,
-the subscales assessed in each test are different.
-To assess whether microbial taxa were associated with
-the development of specific cognitive domains,
-we fit multivariate linear models for each species present in at least 10\%
-of samples for children under 6 months, 18 to 36 months, or all ages
-with a concurrent MSEL score (fig. 2C).
-Several taxa associated with composite scores (fig. 2B)
-were also associated with one or more subscores, including
-\textit{F. prausnitzii} and \textit{A. obesi}.
-Interestingly, while no taxa were significantly associated
-with composite scores in children under 6 months old,
-two species from the genus \textit{Streptococcus}
-(\textit{S. peroris} and \textit{S. mitis}), as well as
-\textit{Bacteroides fragilis}, were positively associated
-with expressive language in this age group.
-\textit{S. mitis} was also associated with expressive language
-when looking at all ages, while \textit{Eisenbergiella massiliensis},
-\textit{Anaerotruncus colihominis}, and an unclassified \textit{Blautia}
-species were negatively associated. % Add more here if they match brain?
+(Figure 2B). \textit{Ruminococcus gnavus} was the only microbial species
+that was significantly negatively associated with cognitive function score (q value = 0.18).
+
+\begin{sidewaysfigure}[!htb]
+    \centering
+    \includegraphics[width=\textwidth]{assets/Figure2.png}
+    \caption{
+        Taxa and gene functional groups are associated with cognitive function.
+        (\textbf{A}) Volcano plot of multivariable linear model results showing the
+        relationship between individual taxa and cognitive function score in
+        children over 18 months of age, controlling for age, maternal education,
+        and sequencing depth. Taxa that were significant after FDR correction (q
+        \textless{} 0.2) are colored red. (\textbf{B}) For taxa that were significantly
+        associated with cognitive function, heatmaps of prevalence, mean
+        relative abundance, and correlation with cognitive function in different
+        age groups. (\textbf{C-F}) Enrichment plots for selected neuroactive gene sets
+        used in feature set enrichment analysis (FSEA). (\textbf{G}) summary of FSEA
+        results across all samples (left) as well as in the under 6 months
+        (middle) and over 18 months (right) subsets, colored based on the
+        significance of the association. Markers indicate the individual
+        correlation of genes within a gene set, and vertical bars represent the
+        median correlation of that gene set.
+    }
+    \label{fig:2}
+\end{sidewaysfigure}
 
@@
 \cite{valles-colomerNeuroactivePotentialHuman2019}
-and concurrent cognitive ability score (table 2, fig. 2D-F) and found that
+and concurrent cognitive ability score (Table 2, Figure 2C-G) and found that
 several metabolic pathways were either significantly enriched or depleted in
-children with higher cognitive ability scores.
-For example, genes for degrading the 3-carbon SCFA, propionate,
+children with higher cognitive ability scores. This was true both when
+considering all ages (0-120 months) together, though the enrichment of most
+pathways was more pronounced either in children under 6 months or those over 18
+months. For example, genes for degrading the 3-carbon SCFA, propionate,
 were significantly depleted in children with higher cognitive ability
-scores in children under 6m (table 2; propionate degradation I,
-enrichment score (E.S.) = -0.66, corrected p-value (q)
-= 0.029), as were genes for synthesizing the 2 carbon SCFA acetate
-(acetate synthesis II, E.S. = -0.491, q = 0.059; acetate synthesis
-III, E.S. = -0.409, q = 0.133).
-Interstingly, in children over 18m,
-these metabolic associations were reversed, with both acetate synthesis
-and propionate degredation being enriched in children
-with higher cognitive scores
-(acetate synthesis I, E.S. = 0.192, q = 0.145;
-acetate synthesis II, E.S. = 0.248, q = 0.15;
-propionate degredation I, E.S. = 0.154, q = 0.178),
-as was synthesis of the 4-carbon SCFA butyrate
-(butyrate synthesis II, E.S. = 0.414, q = 0.178).
-Synthesis of the amino acid
-glutamate was also significantly enriched in children with
-higher cognitive function scores in this age group (glutamate synthesis II, E.S. = 0.355, 
-q = 0.023). Taken together, these results suggest that
+scores across all age groups tested (Table 2; propionate degradation I,
+under 6 months, enrichment score (E.S.) = -0.542, corrected p-value (q)
+= 0.020; over 18 months, E.S. = -0.674, q = 0.041). Interestingly, genes
+for propionate synthesis were also significantly depleted in higher
+scoring children over 18 months (E.S. -0.676, q = 0.023), as were genes
+for synthesizing the 2 carbon SCFA acetate in children under 6 months
+old (acetate synthesis I, E.S. = -0.194, q = 0.153; acetate synthesis
+II, E.S. = -0.342, q = 0.020; acetate synthesis III, E.S. = -0.31, q = 0.052).
+In children over 18 months old, synthesis of menaquinone (vitamin K)
+was also negatively associated with cognitive function score (menaquinone synthesis I,
+E.S. = -0.170, q = 0.0183), while
+genes for the synthesis of the amino acids
+glutamate and tryptophan were significantly enriched in children with
+higher cognitive function scores (glutamate synthesis I, E.S. = 0.242, 
+q = 0.047; tryptophan synthesis, E.S. = 0.119, q = 0.041).
+Taken together, these results suggest that
 microbial metabolic activity, particularly the metabolism (synthesis and
@@
     \begin{tabular}{|r|r|r|r|r|r|r|}
-        \hline
-        \textbf{Age group}& \multicolumn{2}{c|}{\textbf{0 to 6m}} & \multicolumn{2}{c|}{\textbf{18 to 120m}} \\\hline
-        geneset & E.S. & q value & E.S. & q value \\\hline
-        Acetate syn. I & -0.165 & 0.284 & 0.192 & \textbf{0.145} \\
-        Acetate syn. II & -0.491 & \textbf{0.059} & 0.248 & \textbf{0.15} \\
-        Acetate syn. III & -0.409 & \textbf{0.133} & 0.215 & 0.209 \\
-        Butyrate syn. II & -0.357 & 0.284 & 0.414 & \textbf{0.178} \\
-        Glutamate syn. II & 0.343 & 0.284 & 0.355 & \textbf{0.023} \\
-        Propionate deg. I & -0.659 & \textbf{0.029} & 0.154 & \textbf{0.178} \\\hline
+      \hline\hline
+      \textbf{Age group} & \multicolumn{2}{c}{\textbf{0 to 120 months}} & \multicolumn{2}{c}{\textbf{0 to 6 months}} & \multicolumn{2}{c}{\textbf{18 to 120 months}} \\ \hline
+      geneset & E.S. & q value & E.S. & q value & E.S. & q value \\\hline
+      Acetate syn. I & -0.122 & 0.555 & -0.194 & \color{blue}{\textbf{0.153}} & 0.127 & 0.602 \\
+      Acetate syn. II & -0.22 & \color{blue}{\textbf{0.066}} & -0.342 & \color{blue}{\textbf{0.02}} & -0.095 & 0.92 \\
+      Acetate syn. III & -0.21 & \color{blue}{\textbf{0.086}} & -0.31 & \color{blue}{\textbf{0.052}} & -0.086 & 0.92 \\
+      Glutamate deg. I & -0.404 & \color{blue}{\textbf{0.18}} & -0.284 & 0.816 & -0.249 & 0.602 \\
+      Glutamate syn. II & 0.095 & 0.602 & -0.134 & 0.98 & 0.242 & \color{red}{\textbf{0.047}} \\
+      Menaquinone syn. I & -0.085 & 0.563 & -0.083 & 0.816 & -0.17 & \color{blue}{\textbf{0.182}} \\
+      Propionate deg. I & -0.461 & \color{blue}{\textbf{0.13}} & -0.542 & \color{blue}{\textbf{0.02}} & -0.674 & \color{blue}{\textbf{0.041}} \\
+      Propionate syn. I & -0.334 & 0.555 & -0.32 & 0.571 & -0.676 & \color{blue}{\textbf{0.023}} \\
+      Tryptophan syn. & 0.057 & 0.974 & -0.072 & 0.816 & 0.119 & \color{red}{\textbf{0.041}} \\\hline\hline
     \end{tabular}
-    \caption{\label{tab:fsea} \textbf{Neuroactive microbial gene sets microbial gene sets are associated with cognitive function}.
-    Bold text indicates significant (q value \textless 0.2) genesets with a positive or negative
-    pseudomedian.}
+    \caption{\label{tab:fsea}Feature set enrichment analysis on neuroactive microbial gene sets.
+    Red and blue text indicates significant (q value \textless 0.2) genesets with a positive or negative
+    pseudomedian, respectively.}
     \end{center}
@@
 
-\subsubsection*{Gut microbial taxonomic and functional profiles predict concurrently measured cognitive function}
+\subsubsection*{Gut microbial taxonomic and functional profiles predict cognitive function}
 
@@
 sex, and age were included as possible predictors, either alone
-or in combination with microbial taxonomic profiles (table 3).
+or in combination with microbial taxonomic profiles (Table 3).
+
+\begin{sidewaysfigure}[!htb]
+    \centering
+    \includegraphics[width=\textwidth]{assets/Figure3.png}
+    \caption{
+        Random forest models predict concurrently measured cognitive function.
+        Comparison of RF predictor importance versus linear models for children
+        between birth and 6 months old (\textbf{A}), and for those older than 18 months
+        (\textbf{B}). Colors represent whether the species belong to the group of
+        top-important features that account for 60\% of the cumulative
+        importance on the RF model, if that species was significant (q
+        \textless{} 0.2) in linear models, both, or neither. Ranked feature
+        importance for taxa in RF models for children between birth and 6 months
+        old (\textbf{C}), and for those older than 18 months (\textbf{D}). Taxa that are important
+        for RF models both for children under 6 months and for those over 18
+        months (\textbf{E}), only for children under 6 months (\textbf{F}), and only for children
+        over 18 months (\textbf{G})
+    }
+    \label{fig:3}
+\end{sidewaysfigure}
 
@@
       \textbf{Subject Ages (months)} & \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
-      \multirow{5}{*}{0 to 6} & -    & + & 13.37 ± 0.05 & -0.16 ± 0.01 \\ \cline{2-5}
-            & \multirow{2}{*}{taxa}  & - & 12.99 ± 0.04 & -0.11 ± 0.01 \\ \cline{3-5}
-            &                        & + & 13.01 ± 0.04 & -0.13 ± 0.01 \\ \cline{2-5}
-            & \multirow{2}{*}{genes} & - & 12.95 ± 0.04 & -0.05 ± 0.01 \\ \cline{3-5}
-            &                        & + & 12.95 ± 0.04 & -0.05 ± 0.01 \\ \cline{1-5}
-      \multirow{5}{*}{18 to 120} & - & + & 17.06 ± 0.04 & 0.511 ± 0.003 \\ \cline{2-5}
-            & \multirow{2}{*}{taxa}  & - & 18.71 ± 0.04 & 0.347 ± 0.003 \\ \cline{3-5}
-            &                        & + & 18.25 ± 0.04 & 0.429 ± 0.003 \\ \cline{2-5}
-            & \multirow{2}{*}{genes} & - & 18.96 ± 0.04 & 0.299 ± 0.003 \\ \cline{3-5}
-            &                        & + & 18.74 ± 0.04 & 0.341 ± 0.003 \\ \hline\hline
+      \multirow{5}{*}{0 to 6} & - & + & 13.01 ± 0.05 & -0.14 ± 0.01 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa} & - & 12.66 ± 0.05 & -0.1 ± 0.01 \\ \cline{3-5}
+            &       & + & 12.7 ± 0.05 & -0.12 ± 0.01 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 12.56 ± 0.05 & -0.01 ± 0.01 \\ \cline{3-5}
+            &       & + & 12.56 ± 0.05 & -0.02 ± 0.01 \\ \cline{1-5}
+      \multirow{5}{*}{18 to 120} & - & + & 16.27 ± 0.04 & 0.506 ± 0.003 \\ \cline{2-5}
+            & \multirow{2}{*}{taxa} & - & 17.66 ± 0.04 & 0.363 ± 0.003 \\ \cline{3-5}
+            &       & + & 17.29 ± 0.04 & 0.429 ± 0.003 \\ \cline{2-5}
+            & \multirow{2}{*}{genes} & - & 18.01 ± 0.04 & 0.303 ± 0.003 \\ \cline{3-5}
+            &       & + & 17.86 ± 0.04 & 0.333 ± 0.004 \\\hline\hline
     \end{tabular}
-    \caption{\label{tab:rfbench} \textbf{RF models on microbial taxa and genes can learn associations with concurrently measured cognitive performance.}
-    Benchmark metrics for the cognitive assessment score
+    \caption{\label{tab:rfbench}Benchmark metrics for the cognitive assessment score
     prediction models. Confidence intervals (C.I.) are calculated from the
@@
 As with LMs, RF models for children under 6 months old
-performed poorly (mean test-set correlation -0.13, mean RMSE 13.01),
+performed poorly (mean test-set correlation -0.12, mean RMSE 12.70),
 but RFs were consistently able to learn the relationship between taxa
 and cognitive function scores in children over 18 months of age (mean
-test-set correlation 0.429, mean RMSE 18.25).
-Species that were important in RF overlapped with those that were
-significant when testing the relationship with LMs (fig.
-3A, table S1); Both analysis identified 
-\textit{F. prausnitzii}, % F. prausz was mentioned before, ~ line 395
-\textit{E. eligens}, % E. eligens was mentioned before, ~ line 395
-\textit{Parasutterella excrementihominis}, and % Not mentioned before
-\textit{A. obesi} as important % A. obesi was mentioned before, ~ lines 120, 390
-(q \textless 0.2 for LMs, cumulative importance \textless 60\% for RFs).
-Other taxa that were significant in LMs were not identified by RF models
-including \textit{Eggerthella} sp. CAG 298,
-and \textit{B. intestinihominis}. \textit{E. ramulus}
-was just outside our cutoff (64\% cumulative importance). % Flag - 64%?
-RF models also identified an additional 48 species
-that were important for discriminating cognitive function scores
-(table 3, fig. 3B, E). Subject age was consistently ranked highly in
+test-set correlation 0.429, mean RMSE 17.29). For both age groups,
+species that were important in RF overlapped with those that were
+significant when testing the relationship with LMs (Figure
+3A-B, Table S1A-B); except for \emph{Eubacterium ramulus},
+all taxa that were significant using LMs belonged to a group
+responsible for 60\% of the total importance in RF models
+(Table 3, Figure 3A).
+However, the rank-ordering of LM significance and RF importance
+had substantive differences. For example, in children under 6 months,
+only \emph{B. wexlerae} was significantly associated with cognitive function in LMs, but
+was ranked 20th in importance for RF models.
+
+Many taxa highly ranked in importance in RF models in both age
+groups, including several that were identified in LMs for children over
+18 months old, including \emph{R. gnavus} (0--6 months, rank = 13;
+18--120 months, rank = 7) and \emph{G. pamelaeae} (0--6 months, rank =
+16; 18--120 months, rank = 20), while others such as \emph{Allistipes
+finegoldii} were age-group specific. A number of taxa that were important in RF models
+were not statistically significant when using LMs after
+multiple hypothesis correction. However, these taxa often had small nominal
+p-values. For example, \emph{Erysipelatoclostridium ramosum} was the
+most important feature in RF models for children under 6 months old and
+had an LM p-value of 0.04. Subject age was consistently ranked highly in
 feature importance, which could indicate that decision branches based on
@@
 
-\subsubsection*{Gut microbial profiles in the first year of life predict future cogntitive function}
-
-While LMs did not identify any taxa in the 0-6 month range
-that were associated with concurrently measured cognitive function,
-we reasoned that developmental trajectories that begin
-\textit{in utero} may not be immediately altered by
-microbial exposures, such that microbial influences early in life
-may take time to manifest. To test this hypothesis,
-we asked whether RF models trained on taxonomic profiles
-in the first year of life
-could predict \textit{future} cognitive performance measured at least 6 months
-after the stool sample was collected (fig. 3C-D, F; table 4).
-Models were consistently able to learn microbial species that are associated
-with cognitive performance at a future visit
-(mean time between visits: 21.5 months, S.D. = 9.9 months; fig. 3G,
-mean test set correlation: 0.27).
-
-\begin{table}[!h]
-    \begin{center}
-    \begin{tabular}{|r|r|r|r|r}
-        \hline
-        \textbf{Microbial feature} & \textbf{Demo.} &  \textbf{Test set Root-mean-square-error (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
-        - & + & 20.25 ± 0.07 & 0.43 ± 0.0 \\ \cline{1-4}
-        \multirow{2}{*}{taxa} & - & 22.32 ± 0.08 & 0.07 ± 0.0 \\ \cline{2-4}
-            & + & 21.53 ± 0.08 & 0.27 ± 0.0 \\ \cline{1-4}
-        \multirow{2}{*}{genes} & - & 22.21 ± 0.08 & 0.12 ± 0.0 \\ \cline{2-4}
-            & + & 22.08 ± 0.08 & 0.15 ± 0.0 \\\hline\hline
-    \end{tabular}
-    \caption{\label{tab:rfbench2}  \textbf{RF models on microbial taxa and genes can learn associations with future cognitive performance.}
-    Confidence intervals (C.I.) are calculated from the
-    distribution of metrics from repeated CV at a confidence level of 95.
-    Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
-    Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.}
-    \end{center}
-\end{table}
-
-Nearly all important taxa in RF models that predicted cognitive function
-in children older than 18 months were also important in
-models that predicted future cognitive function, if they were present (fig. 3B-C, table S5).
-At the same time, many additional species were important for future prediction models
-that were not important in the models measuring concurrent cognitive function.
-In particular, feature importance for future prediction models
-were dominated by two species of \textit{Klebsiella},
-\textit{K. pneumoniae} and \textit{K. varicola} (fig. 3C, F).
-Many of the other species that ranked highest in feature importance
-were also oportunistic pathogens, or previously reported
-to be linked with disease (eg. \textit{R. gnavus}, \textit{E. coli}).
-
-\subsubsection*{Gut microbial taxa are associated with specific cognitive functions}
-
-Given that the development of different domains of cognitive function
-(e.g., gross motor development vs visual perception) may develop at different times
-and be affected by different molecular signals,
-we sought to determine whether the associations of microbial taxa
-with cognitive function are general across all cognitive subdomains.
-To accomplish this, we assessed individual subscales of the  
-MSEL (birth to 36 months - \textit{see Methods}). 
-Taxonomic profiles coupled with demographics exhibited predictive potential towards three
-of five probed subscales (fig. 4A, table 5), 
-namely expressive language ( test-set RMSE 12.05, test-set correlation 0.368), 
-gross motor (test-set RMSE 11.07, test-set correlation 0.137) and 
-visual reception (test-set RMSE 15.2, test-set correlation 0.333).
-Between the subscales, both the predictability (measured by the test-set figures of merit)
-and the relative importances of several taxa (fig. 4B) is variable.
-Although certain species such as \textit{B. longum}, \textit{F. prausznitzii} 
-have similar relative importance across all subscales, others are substantially different (tables S2-S4).
-Notably, \textit{Bifidobacterium pseudocatenulatum}, \textit{B. wexlerae} and \textit{E. eligens} are
-substantially more important for the prediction of expressive language, while
-\textit{R. faecis}, \textit{Streptococcus salivarius} and \textit{Fusicatenibacter saccharivorans} are
-distinctly predictive of gross motor, and
-\textit{Clostridium innocuum}, \textit{B. vulgatus} 
-stand out on models targeting visual reception (tables S2-4).
-
-\begin{table}[!h]
-    \begin{center}
-        \begin{tabular}{|r|r|r|r|r|}
-        \hline
-        \textbf{Target Mullen scale} & \textbf{Microbial feature} & \textbf{Demo.} & \textbf{Test set RMSE (± C.I.)} & \textbf{Test set correlation (± C.I.)} \\\hline
-        \multirow{5}{*}{Expressive Language} & - & + & 11.17 ± 0.04 & 0.48 ± 0.0 \\ \cline{2-5}
-            & \multirow{2}{*}{taxa}  & - & 12.37 ± 0.04 & 0.25 ± 0.0 \\ \cline{3-5}
-            &                        & + & 12.05 ± 0.04 & 0.37 ± 0.0 \\ \cline{2-5}
-            & \multirow{2}{*}{genes} & - & 12.59 ± 0.04 & 0.15 ± 0.0 \\ \cline{3-5}
-            &                        & + & 12.46 ± 0.04 & 0.21 ± 0.0 \\ \cline{1-5}
-        \multirow{5}{*}{Gross Motor} & - & + & 10.53 ± 0.04 & 0.346 ± 0.006 \\ \cline{2-5}
-            & \multirow{2}{*}{taxa}  & - & 11.08 ± 0.04 & 0.134 ± 0.006 \\ \cline{3-5}
-            &                        & + & 11.07 ± 0.04 & 0.137 ± 0.006 \\ \cline{2-5}
-            & \multirow{2}{*}{genes} & - & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \cline{3-5}
-            &                        & + & 10.98 ± 0.04 & 0.197 ± 0.005 \\ \hline
-        \multirow{5}{*}{Visual reception} & - & + & 14.13 ± 0.04 & 0.462 ± 0.003 \\ \cline{2-5}
-            & \multirow{2}{*}{taxa}  & - & 15.49 ± 0.04 & 0.245 ± 0.004 \\ \cline{3-5}
-            &                        & + & 15.2 ± 0.04 & 0.333 ± 0.004 \\ \cline{2-5}
-            & \multirow{2}{*}{genes} & - & 15.97 ± 0.04 & 0.067 ± 0.004 \\ \cline{3-5}
-            &                        & + & 15.84 ± 0.04 & 0.118 ± 0.005 \\ \hline\hline
-        \end{tabular}
-        \caption{\label{tab:subscalebenchmark}  \textbf{RF models on microbial taxa and genes can learn associations with performance in cognitive subdomains.}
-        Benchmark metrics for the Mullen subscales prediction models.
-        Confidence intervals (C.I.) are calculated from the
-        distribution of metrics from repeated CV at a confidence level of 95.
-        Root mean square error (RMSE) was used to estimate how well the random forest model was able to predict the outcome.
-        Demo. column indicates whether demographic factors such as maternal education, sex and age were included as predictors.
-        Age range, in all cases, was 18-120 months.}
-    \end{center}
-\end{table}
-
 \subsubsection*{Gut microbial taxonomic profiles predict brain structure differences}
 
-Given that individual microbes could be associated with particular domains
-of cognitive function and not others, if there are causal effects of microbial metabolism on cognitive
+If there are causal effects of microbial metabolism on cognitive
 function, they might be reflected in changes in neuroanatomy. We again
@@
 profiles with individual brain regions identified in MRI scans
-(children over 18 months old, N = 124).
+(children over 18 months old, N = 121).
 Some brain regions were more readily
-predicted by RF models trained on microbial taxa (tables 5, 6, and S6), in particular those that were highly correlated with age.
+predicted by RF models trained on microbial taxa (Table 4, Table S2), in particular those that were highly correlated with age.
 These included the L/R lingual gyrus (mean RF correlations, Left =
-0.444, Right = 0.441; importance for age, Left = 6.8\%, Right =
-7.7\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
-0.267, Right = 0.315; importance for age, Left = 3.6\%, Right = 6.2\%).
+0.421, Right = 0.434; importance for age, Left = 7.5\%, Right =
+8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
+0.200, Right = 0.273; importance for age, Left = 3.6\%, Right = 6.3\%).
 In many cases, however, age was not an important variable in
 high-performing models, such as that for the left accumbens area (mean
-RF correlations = 0.275; relative age importances = 1.1\%).
+RF correlations = 0.288; relative age importances = 1.2\%).
 
@@
       \textbf{Statistic} & \textbf{MAPE} & \textbf{R} \\\hline
-      Mean & 0.073 & 0.146 \\
-      Standard deviation & 0.029 & 0.131 \\
-      Maximum & 0.205 & 0.598 \\
-      75th percentile & 0.081 & 0.212 \\
-      50th percentile & 0.066 & 0.129 \\
-      25th percentile & 0.057 & 0.073 \\
-      Minimum & 0.036 & -0.133 \\\hline
+      Mean & 0.073 & 0.137 \\
+      Standard deviation & 0.028 & 0.137 \\
+      Maximum & 0.206 & 0.605 \\
+      75th percentile & 0.082 & 0.195 \\
+      50th percentile & 0.067 & 0.134 \\
+      25th percentile & 0.057 & 0.062 \\
+      Minimum & 0.035 & -0.147 \\\hline\hline
     \end{tabular}
-    \caption{\label{tab:neuro}  \textbf{RF models on microbial taxa predicting brain regions have generally good performance.}
+    \caption{\label{tab:neuro}Summary statistics for the neuroanatomy prediction benchmarks.
     Mean absolute proportional error (MAPE) and correlation coeficient (R)
     of test-set preditions are reported accross models for all brain regions.
-    Complete list of values can be found on table S6.}
+    Complete list of values can be found on Table S2.}
     \end{center}
@@
 brain region models (R = 0.288), as compared to the right which
-had a negligible test-set correlation (R = 0.005).
+had a negative mean test-set correlation (R = -0.041),
+indicating that generalizable models were not identified.
 Feature importance for models of the left accumbens area were dominated
 by three species of \emph{Bacteroides}, \emph{B. vulgatus} (3.3\% importance),
-\emph{B. ovatus} (3.2\% importance), and
-\emph{B. uniformis} (2.8\% importance).
+\emph{B. ovatus} (3.7\% importance), and
+\emph{B. uniformis} (2.9\% importance).
 
@@
 were also identified in LMs and RF models of cognitive function (Figures 2 and 3). 
-In general, for all segments, a consistent number of species between 28 and 42 (median = 38, less than 30\% of the input features)
+In general, for all segments, a consistent number of species between 31 and 45 (median = 37, less than 30\% of the input features)
 was responsible for half of the cumulative importance,
-regardless of model performance (fig. 4C-E, table S2).
+regardless of model performance (Figure 4A-B, Table S2).
 We observed two major patterns of importance distribution from
@@
 just one or two brain segments. Notable cases of the first pattern
-included nine species - 
-\emph{Anaerostipes hadrus},
-\emph{B. vulgatus},
-\emph{B. ovatus},
-\emph{F. saccharivorans},
-\emph{Ruminococcus torques},
-\emph{Eubacterium rectale},
-\emph{Coprococcus comes},
-\emph{A. celatus} and
+included seven species - \emph{Anaerostipes hadrus}, \emph{B.
+vulgatus}, \emph{Fusicatenibacter saccharivorans}, \emph{Ruminococcus
+torques}, \emph{Eubacterium rectale}, \emph{Coprococcus comes} and
 \emph{B. wexlerae}, which combined, account for approximately 10\% of the cumulative relative importances,
-computed after fitness-weighting on the reported brain segments (fig. 4D-E).
+computed after fitness-weighting on the reported brain segments (Figure 4A-B).
+
+\begin{sidewaysfigure}[!htb]
+    \centering
+    \includegraphics[width=\textwidth]{assets/Figure4.png}
+    \caption{
+        Microbial feature importance in predicting brain volumes in children over 18 months of age.
+        (\textbf{A}) Average test-set correlations for prediction of MRI segmentation
+        data from microbiome and demographics on select segments after repeated
+        cross validation. (\textbf{B}) Heatmap of average individual relative taxa
+        importances on each brain segment. Importances are reported as
+        proportions relative to the sum of importances for each model - since
+        every model is trained with 132 features, and even distribution of
+        importance would be 0.75\% for each feature. Segments and taxa ordered
+        by HCA on a select list of species with high ranks on average importances,
+        and their respective highest-load segments (\textbf{C}).
+    }
+    \label{fig:4}
+\end{sidewaysfigure}
 
 Among these, \emph{A. hadrus} is the variable with 
-highest average relative importance along all reported models
-(mean = 1.4\%; min = 0.8\%; max = 3.6\%, fig. 4D-E).
+highest average relative importance along all reported 
+(mean = 1.7\%; min = 1.0\%; max = 3.8\%, Figure 4B-C).
 \emph{B. wexlerae}, which was important in RF models of cognitive function,
+and was the only species to be significantly associated
+with cogntive function in young children using LMs, 
 was also important in models of the entorhinal, fusiform and lingual areas
@@
 cluster contains \emph{B. ovatus} and \emph{B. uniformis},
-but also \emph{Alistipes finegoldii and S. salivarius}.
-
-\emph{A. celatus}, an equol-producing species, is an examples of the second contribution
+but also \emph{Alistipes finegoldii and Streptococcus salivaris}.
+
+Both \emph{A. equolofaciens} and \emph{A. celatus}, two closely-related
+equol-producing species, are examples of the second contribution
 pattern, where a taxon only contributes significantly to a very small number of segments. 
-This species was highly important in predicting the relative volume of
-the right anterior cingulate (2.8\% relative importance).
-Another example of this pattern is \emph{Coprococcus eutactus}
-which is heavily associated with the left precentral area (3.5\% importance). 
+Each of these species were highly important in predicting the relative volume of
+the right anterior cingulate (respectively, 3.0\% and 2.6\% importances).
+Another example of this pattern is \emph{R. gnavus}, which was the only
+species with significant negative association with cognitive function in linear models.
+It is heavily associated with the left pars opercularis (2.7\% importance). 
 Finally, \emph{C. comes} has an importance
@@
 clusters. Its loading on the prediction of the left posterior cingulate
-is the highest for a microbe in RF models (3.7\% importance);
+is the highest for a microbe in RF models (4.0\% importance);
 only age had higher relative importance in any model. At the same time,
@@
 posterior cingulate such as the pars opercularis (importances,
-Left = 1.9\%, Right = 2.5\%), along with upper regions like the left
-precentral (2.0\% importance) and paracentral lobes (importances, Left = 2.0\%, Right = 1.5\%).
+Left = 2.3\%, Right = 2.8\%), along with upper regions like the left
+precentral (2.3\% importance) and paracentral lobes (importances, Left = 2.6\%, Right = 1.9\%).
 
@@
 We identified several species in the family Eggerthelaceae that were
-associated with cognitive function, including \emph{A. celatus}
-(a subspecies of \emph{A. equolofaciens}
+associated with cognitive function, including \emph{G.pamelaeae, A. equolofaciens, A. celatus}
+(formerly regarded a subspecies of \emph{A. equolofaciens}
 \cite{takahashiCompleteGenomeSequence2021}),
@@
 found in soybeans \cite{wangEnantioselectiveSynthesisSEquol2005}.
-This is particularly intriguing, since \emph{A. celatus}
-was also the most important features in RF models of the
+This is particularly intriguing, since both \emph{A. equolofaciens} and \emph{A. celatus}
+were also the most important features in RF models of the
 right anterior cingulate, which has been linked to social cognition
@@
 in this brain region \cite{xiaoEstrogenAnteriorCingulate2013}.
-Another Eggerthelaceae, \emph{Gordonibacter pamelaeae},
-also has a unique metabolism that may impact the brain,
+\emph{G. pamelaeae} also has a unique metabolism that may impact the brain,
 as it can metabolize the polyphenol ellagic acid
@@
 \cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
-Though \textit{G. pamelaeae} was not identified
-as an important feature in RF models,
-it was also relatively low prevalence in our dataset,
-occuring in only ~5\% of samples at greater than 2\% abundance
-(in contrast to \textit{A. celatus}, which was in 32\% of samples).
 While we could not find published evidence of neurological roles for \emph{E. lenta},
@@
 targets for further mechanistic studies.
-
-We did not identify any statistical associations between individual species
-and cognitive performance in our youngest subset (0 to 6 months).
-We hypothesize that this may be because brain developmental trajectories
-begin \textit{in utero}, and microbial influences (that can only begin at birth)
-take additional time to manifest.
-In line with this hypothesis, RF models that were able to learn associations
-between early-life microbiomes and future cognitive function scores (fig. 3).
-The top-ranked species for these models included several 
-taxa that were important for predicting concurrent cognitive assessment scores after 18 months, notably
-pathobionts in the family \textit{Enterobacteriaceae} such as
-\textit{Escherichia} and \textit{Klebsiella}. 
-Representatives of these genera are known to be early and long-term persistent
-colonizers of infant intestine that can, in some settings,
-promote inflammatory response \cite{popeEnterobacEarlyGut2019}.
-By contrast, several inflammation-reducing representatives
-of the \textit{Bifidobacterium} genus,
-from the early HMO-metabolizer \textit{B. longum} to the
-later complex carbohydrate relying \textit{B. pseudocatenulatum} \cite{chungBifidoPsudocatenulatum2021},
-commonly associated with weaning to solid food in early age,
-also showed high relative importance.
-The involvement of most important taxa in different aspects of inflammatory response
-resonate with previous findings that associate chronic inflammation with
-changes to neural responses in some types of stimulae \cite{xieChronicInfNeuro2019}.
 
@@
 
-Our ability to probe individual cognitive subscales provided
-results that also align with previous findings and provide insight to
-potential gut-microbial connections between cognition and neuroanatomy.
-For example, a longitudinal neurocognitive development study pointed at bilateral
-associations between the thalamus and the expressive language
-component of the Mullen scales \cite{chenLongitudinalBrainMullen2023}.
-In line with this finding, we found that taxa such as 
-\textit{B. wexlerae} and \textit{A. hadrus} to be among the top 15\% important predictors
-on both left and right thalamus and the expressive language subscale.
-Another example of those findings is a previously reported association between
-the central opercular cortex (COC) and gross motor development \cite{jiangBrainNetworkCognition2023}.
-In this case, we identified that
-butyrate-producing \textit{F. saccharivorans} is among the 15\% most important predictors
-for both the gross motor subscale and three bilateral components of the COC
-(left and right pars orbitalis, pars opercularis and pars triangularis).
-Previous results also link to relationships between the caudate and the cerebellum,
-and the visual reception score \cite{deoniMyelinationMullenPreprint2022}.
-\textit{S. salivarius} was consistently found among the top 20\% predictors for this subscale, along with
-both sides of the thalamus and cerebellum white matter, along with cerebellar vermal lobules I-X.
-
 As is made apparent by the potentially opposing roles of 
@@
 Clostridiales, as is \emph{R. gnavus}, which we found was negatively
-associated with cognitive function (fig. 3C, F). This kind of
+associated with cognitive function (Figure 2A-B). This kind of
 species-level resolution is typically not possible with amplicon
@@
 genes for the metabolism of SCFAs, both their degradation and synthesis,
-are associated with cognitive function scores (fig. 2). 
+are associated with cognitive function scores (Figure 2). 
 SCFAs are produced by anaerobic fermentation of dietary fiber
@@
 the relationship between microbial metabolism and cognitive development.
-Serum metabolomic analysis may be even more informative,
-since the metabolites measured in stool are not necessarily indicative
-of exposure, but drawing sufficient blood from children,
-especially very young children, is challenging.
 Further, strain-level analysis linking specific gene content in species
@@
 \cite{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
-(fig. S3).
+(Figure S3).
 
 This analysis allowed us to establish links between microbial taxa and
-their functional potential with cognition and brain structure. 
-Although we did not directly test the causal relationships
+their functional potential with cognition and brain structure. Although
+we cannot test causality or the chemistry behind the interactions
 between gut microbial taxa and their genes, the gut, and the brain,
 this study provides clear and statistically significant associations
-that could serve as targets for future efforts in pre-clinical models.
-For example, mouse models of brain development, learning, and executive function
-could be assessed in germ-free animals with the addition of
-metabolically characterized and manipulated microbes.
-Future studies should also focus on characterizing the early-life microbiome
+between the infant and early child gut microbiota and neurocognition.
+Future studies should focus on characterizing the early-life microbiome
 and neurocognitive development across different geographic regions
@@
 All procedures for this study were approved by the local institutional
-review board at Rhode Island Hospital, 
-(Investigating Early Childhood Neurodevelopment 1500991,
-and Prenatal Influences on Fetal and Infant Brain Development 1501119;
-most recent approval for both: 2020-01-15)
-and all experiments adhered to
+review board at Rhode Island Hospital, and all experiments adhered to
 the regulation of the review board. Written informed consent was
@@
 As a study of neurotypical development, mothers and their children
-exclusion criteria included major risk factors for abnormal or atypical development,
-i.e., preterm birth (\textless 37 weeks gestation) low birth weight (\textless 1500g),
-\textit{in utero} exposure to alcohol, cigarette or illicit substance exposure;
+were devoid of major risk factors for abnormal or atypical development.
+These included preterm birth (\textless 37 weeks gestation) low birth weight (\textless 1500g),
+in utero exposure to alcohol, cigarette or illicit substance exposure;
 abnormalities on fetal ultrasound; non-singleton or complicated pregnancy,
@@
 them from subsequent analysis.
-For this study, 381 typically-developing children between the
-ages of 40 days and 10 years (median age 2 years, 2 months) were
+For this study, 361 typically-developing children between the
+ages of 2.8 months and 10 years old (median age 2 years, 2 months) were
 selected for analysis based on having provided at least one stool sample
@@
 computational tools \cite{beghiniIntegratingTaxonomicFunctional2021}.
-First, KneadData (v0.10.0) was used to perform quality
+First, KneadData (v0.7.7) was used to perform quality
 control of raw sequence reads, such as read trimming and removal of
-reads matching a human genome reference. Next, MetaPhlAn (v3.1.0, using
-database mpa\_v31\_CHOCOPhlAn\_201901) was used to generate taxonomic
+reads matching a human genome reference. Next, MetaPhlAn (v3.0.7, using
+database mpa\_v30\_CHOCOPhlAn\_201901) was used to generate taxonomic
 profiles by aligning reads to a reference database of marker genes.
-Finally, HUMAnN (v3.6.0) was used to functionally profile the
+Finally, HUMAnN (v3.0.0a4) was used to functionally profile the
 metagenomes.
@@
 demographics (sex and maternal education as a proxy of socioeconomic
-status) on participants from different age groups.
-
-For the concurrent cognitive assessment score preditions,
-all viable datapoints containing 
-age at data collection, demographics, stool metagenomics and cognitive assessment were employed.
-The cognitive assessments were collected as either composite scores
-(for the experiments in fig. 3) or, when available, subscores on the
-Mullen Scales of Early Learning (fig. 4 A-B: expressive language, gross motor and visual reception).
-For the concurrent prediction experiments, when multiple viable longitudinal samples were present,
-the single earliest viable datapoint inside the age range of the experiment was considered.
-For the future prediction experiments, every demographics-complete subject
-contributed a single datapoint built from the combination of
-the latest metagenome collected within the stool collection time range and
-the latest future cognitive assessment on the assesment time range.
-Age (in months) was provided as a covariate for all models (table S3).
+status) on participants from different age groups. Age (in months) was
+provided as a covariate for all models (Table S3).
 
@@
 on Random Forests with high-dimensional, low sample size
-datasets \cite{woodruffChronicInflammationNeutrophil2023a}.
+datasets \cite{woodruffInflammationAutoreactivityDefine2022}.
 The resulting fitness-weighted importances were used
@@
 
-\section*{References}
-
-\printbibliography[heading=none]
-
-\section*{Acknowledgements}
-% https://www.science.org/content/page/science-advances-information-authors#acknowledgments
-
-We thank Libusha Kelly (Albert Einstein School of Medicine) and Daphne Koinis Mitchell (Brown University)
-for helpful feedback on the manuscript. The research team also thanks the participants and families
-of the participants who generously shared their time and samples for this project.
-
-\section*{Funding}
-This research was funded by NIH UG3 OD023313 and Wellcome LEAP 1kD.
-
-\section*{Author contributions}
-% https://credit.niso.org/
-
-\begin{itemize}[noitemsep]
-    \item Conceptualization: KSB, VKC, CH, and SD
-    \item Methodology: KSB, VKC, CH, and SD
-    \item Project administraion: KSB, VKC, CH, and SD
-    \item Funding acquisition: VKC, SD, and VD
-    \item Data curation: KSB, GFB, SHM, MB, FB, RCL, and JB
-    \item Formal analyses: KSB, GFB, and MB
-    \item Investigation: KSB, GFB, SHM, EW, FB, RCL, and MB
-    \item Resources: SHM, EW, FB, RCL, JB, CH, and MB
-    \item Software and data visualizations: KSB and GFB
-    \item Supervision: KSB, MB, SD, VD, and VKC 
-    \item Validation: KSB, SHM, JB, and MB
-    \item Initial draft: KSB and GFB
-    \item Revision and editing: KSB, GFB, SD, CH, and VKC
-\end{itemize}
-
-Members of the RESONANCE consortium also contributed to
-project planning, conceptualization, and funding acquisition.
-
-\section*{Competing Interests}
-The authors declare that they have no competing interests.
-\section*{Figures}
-
-\section*{Data and materials availability}
+\subsubsection*{Data and code availability}
 
@@
 demographics necessary for statistical analyses and machine learning are
-available on the Data Dryad
-\cite{bonhamDataDryad2023}.
+available on the Open Science Framework
+\cite{bonhamECHORESONANCEMicrobiome2022}.
 Data processing, generation of summary statistics, and
@@
 and 3rd party packages
-\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021,blaomMLJJuliaPackage2020,ben_sadeghi_2022_7359268,dahua_lin_2023_7695673,douglas_bates_2023_7734970}.
+\cite{bezansonJuliaFreshApproach2017,
+    bonhamMicrobiomeJlBiobakeryUtils2021,
+    danischMakieJlFlexible2021,
+    blaomMLJJuliaPackage2020,
+    ben_sadeghi_2022_7359268,
+    dahua_lin_2023_7695673,
+    douglas_bates_2023_7734970}.
 Information for replicating the package environment
 as well as code for data analysis and figure generation, as well as scripts for automated
-download of input files are available on GitHub at https://github.com/Klepac-Ceraj-Lab/Resonance,
-and archived on Zenodo \cite{kevin_bonham_2023_7647510}.
-
-\begin{figure}[!h]
-    \centering
-    \small
-    % \includegraphics[width=0.9\textwidth]{assets/Figure1.png}
-    \caption{
-        \label{fig:1} \textbf{The RESONANCE cohort is a diverse cohort with accelerated longitudinal sampling.} (\textbf{A}) Stool samples,
-        cognitive assessments and neuroimaging were collected from participants
-        at different ages throughout the first years of life.
-        (\textbf{B}) Cognitive function scores are assessed by
-        different instruments, but may be normalized using full-scale composite scores.
-        (\textbf{C-D}) Principal coordinates analysis using Bray-Curtis dissimilarity on
-        taxonomic profiles demonstrates high beta diversity, with much of the
-        first axis of variation explained by increasing age and alpha diversity.
-        Differences in gene function profiles (\textbf{E}) and neuroimaging (PCA based on
-        the Euclidean distance of brain region volumes) (\textbf{F}) are likewise
-        dominated by changes as children age. (\textbf{G}) Permutation analysis of
-        variance (PERMANOVA) of taxonomic profiles, functional profiles
-        (annotated as UniRef90s) and neuroimaging, against the metadata of
-        interest. Variations in taxonomic and functional profiles explain a
-        modest but significant percent of the variation in cognitive development
-        in children over 18 months of age. (\textbf{H}) Mantel testing of different
-        microbial feature matrices, shows overlapping but distinct patterns of
-        variation. Dotted lines in (B) and (C) show 6 and 18 months, which are
-        used as cut-offs in some following models.
-    }
-
-\end{figure}
-
-\begin{figure}[!htb]
-    \centering
-    \small
-    % \includegraphics[width=0.9\textwidth]{assets/Figure2.png}
-    \caption{
-        \textbf{Taxa and gene functional groups are associated with cognitive function.}
-        (\textbf{A}) Volcano plot of multivariable linear model results showing the
-        relationship between individual taxa and cognitive function score in
-        children over 18 months of age (left) or across all ages (right),
-        controlling for age, maternal education, and sequencing depth.
-        Taxa that were significant after FDR correction (q
-        \textless{} 0.2) are colored red. (\textbf{B}) For taxa that were significantly
-        associated with cognitive function,
-        heatmaps of correlation with cognitive function in different age groups.
-        (\textbf{C}) Multivariate linear models showing the relationship
-        between individual taxa and domain-specific skill assessments
-        from the MSEL for children under 6m (left), 18m to 36m (right),
-        or over across all ages (right)
-        controlling for age, maternal education, and sequencing depth.
-        Asterisks in (A) and (C) indicate significance (q \textless{} 0.2)
-        after multiple hypothesis correction.
-        (\textbf{D-E}) Enrichment plots for proprionate degradation I gene set
-        used in feature set enrichment analysis (FSEA) in children under 6m (D)
-        or over 18m (E). (\textbf{G}) summary of FSEA
-        results in children under 6m of age (left)
-        over 18 months (middle) or across all ages (right), colored based on the
-        significance of the association. Markers indicate the individual
-        correlation of genes within a gene set,
-        and vertical bars colored bars represent the
-        median correlation of that gene set.
-        Vertical dotted lines within each panel represent the median T statistic
-        for all genes in that age group.
-    }
-    \label{fig:2}
-\end{figure}
-
-\begin{figure}[!htb]
-    \centering
-    \small
-    % \includegraphics[width=0.9\textwidth]{assets/Figure3.png}
-    \caption{
-        \textbf{Random forest models predict both concurrent and future measured cognitive function.}
-        Comparison of RF predictor importance versus linear models for children
-        over 18 months (\textbf{A}).
-        Colors represent whether the species belong to the group of
-        top-important features that account for 60\% of the cumulative
-        importance on the RF model, if that species was significant (q
-        \textless{} 0.2) in linear models, both, or neither.
-        Comparison of RF predictor importance for microbial features
-        between concurrent and future cognitive function measure
-        (\textbf{B}).
-        Colors represent whether the species belong to the group of
-        top-important features that account for 60\% of the cumulative
-        importance on the RF model only for concurrent prediction,
-        only for the future prediction, or for both models.
-        Features absent of one model are shown on the origin-aligned sideplots.
-        Distribution of ages for each microbiome sampling - cognitive assesment dyad (\textbf{C}).
-        Blue violin plot shows distribution of sampling age,
-        while red violin plot shows distribution of asessment age.
-        Black lines connect ages for each individual subject.      
-        Ranked feature importance for taxa in RF models for
-        both concurrent (\textbf{D}) and future (\textbf{E})
-        cognitive function measurement prediction.
-        Red dots indicate how relative importances change (among taxa)
-        when demographics (sex, SES) are arred as covariables.
-        Abundances and prevalences of taxa that are important
-        for RF models both for concurrent (\textbf{F}) and future (\textbf{G}) assesment prediction.
-        }
-    \label{fig:3}
-\end{figure}
-
-\begin{figure}[!htb]
-    \centering
-    \small
-    % \includegraphics[width=0.9\textwidth]{assets/Figure4.png}
-    \caption{
-        \textbf{Coupled importance of microbes in neuroanatomy and subscores are consistent with known subcortical associations with cognition.}
-        (\textbf{A}) Test set correlations for RF models predicting MSEL subscale scores,
-        with or without demographics included as predictors.
-        (\textbf{B}) Important taxa in models predicting MSEL subscale scores.
-        (\textbf{C}) Average test-set correlations for prediction of MRI segmentation
-        data from microbiome and demographics on select segments after repeated
-        cross validation. (\textbf{D}) Heatmap of average individual relative taxa
-        importances on each brain segment. Importances are reported as
-        proportions relative to the sum of importances for each model - since
-        every model is trained with 132 features, and even distribution of
-        importance would be 0.75\% for each feature. Segments and taxa ordered
-        by HCA on a select list of species with high ranks on average importances,
-        and their respective highest-load segments (\textbf{E}).
-    }
-    \label{fig:4}
-\end{figure}
+download of input files are available on GitHub \cite{kevin_bonham_2023_7647510}.
+
+\section*{References}
+
+\printbibliography[heading=none]
+
+\section*{Acknowledgements}
+% https://www.science.org/content/page/science-advances-information-authors#acknowledgments
+
+This research was funded by NIH UG3 OD023313 and Wellcome LEAP 1kD.
+The authors declare that they have no competing interests.
+We thank Libusha Kelly (Albert Einstein School of Medicine) and Daphne Koinis Mitchell (Brown University)
+for helpful feedback on the manuscript. The research team also thanks the participants and families
+of the participants who generously shared their time and samples for this project.
+
+% https://credit.niso.org/
+
+KSB, VKC, CH, and SD were responsible for conceptualization,
+methodology and project administration.
+VKC, SD, and VD were responsible for funding aqusition.
+KSB, SHM, MB, FB, RCL, and JB contributed to data curation.
+Formal analyses were conducted by KSB, GFB, and MB.
+Investigation was performed by KSB, GFB, SHM, EW, FB, RCL, and MB.
+Resources were provided by SHM, EW, FB, RCL, JB, CH, and MB.
+Software and data visualizations were produced by KSB and GFB.
+KSB, MB, SD, VD, and VKC were responsible for supervision.
+Validation was performed by KSB, SHM, JB, and MB.
+KSB and GFB wrote the initial draft, 
+and review and editing were provided by SD, CH, and VKC.
 

### 2024-02-22 manuscript/Figure2-calculations.jl

@@ -29,7 +29,7 @@ taxa = Resonance.load(TaxonomicProfiles(); timepoint_metadata = seqs) # this can
 species = filter(f-> taxrank(f) == :species, taxa)
 relativeabundance!(species)
 unirefs = Resonance.load(UnirefProfiles(); timepoint_metadata = seqs) # this can take a bit
-ecs = Resonance.load(ECProfiles(); timepoint_metadata = seqs)
+#ecs = Resonance.load(ECProfiles(); timepoint_metadata = seqs)
 kos = Resonance.load(KOProfiles(); timepoint_metadata = seqs)
 
 # metabolites = Resonance.load(MetabolicProfiles(); timepoint_metadata = mdata)

### 2024-02-22 notes/presentation.jl

@@ -1,3 +1,4 @@
+using ColorSchemes: grays
 using Resonance
 using FileIO
 using ThreadsX
@@ -7,6 +8,7 @@ using ColorSchemes
 using Distances
 using MultivariateStats
 using CategoricalArrays
+using Clustering
 using MLJ
 using JLD2
 
@@ -25,6 +27,10 @@ seqs = subset(mdata, "sample"=> ByRow(!ismissing))
 seqs.sample = [s for s in seqs.sample]
 taxa = Resonance.load(TaxonomicProfiles(); timepoint_metadata = seqs)
 taxdf = comm2wide(taxa)
+
+#-
+
+
 
 #-
 
@@ -56,7 +62,7 @@ fig
 
 #-
 
-B = Figure(; resolution=(1200,600))
+B = Figure(; resolution=(700, 350))
 
 bax1 = Axis(B[1,1]; xlabel = "Age (months)", ylabel = "cogScore", xticks=(4:4:24), limits=((2,24), nothing))
 bax2 = Axis(B[1,2]; xlabel = "Age (years)", xticks = (4:2:12), limits=((2,nothing), nothing))
@@ -74,10 +80,10 @@ bleg = let
         return cs[3] # WISC
     end
     ages = mdata.ageMonths[u2y]
-    scatter!(bax1, ages, mdata.cogScore[u2y]; color=colorage.(ages))
+    scatter!(bax1, ages, mdata.cogScore[u2y]; color=[(colorage(age), 0.4) for age in ages])
 
     ages = mdata.ageMonths[o2y]
-    scatter!(bax2, ages ./ 12, mdata.cogScore[o2y]; color=colorage.(ages))
+    scatter!(bax2, ages ./ 12, mdata.cogScore[o2y]; color=[(colorage(age), 0.4) for age in ages])
     
     vlines!(bax1, [6, 18]; linestyle=:dash, color=:gray)
     # TODO: add colors for training/test sets in later models
@@ -515,4 +521,496 @@ Legend(
 save("/home/kevin/Downloads/resonance-importance-18to120.png", CD_subfig)
 
 #-
+
+fig = Figure(; resolution=(1100,700))
+
+axB = Axis(fig[1,2])
+
+let plot_colorset = [t10[1], t10[6], t10[5], (:white, 0.)]
+    Resonance.plot_comparative_rfvsrf_scatterplots!(axB,
+        regression_currentCogScores_18to120mo_onlytaxa,
[…]

### 2025-01-06 manuscript/main.tex

--- 2023-11-13 manuscript/main.tex
@@
 style=numeric,
-maxbibnames=100
+maxbibnames=50
 ]{biblatex}
@@
 
-Members of the RESONANCE consortium also contributed to
-project planning, conceptualization, and funding acquisition.
-
 \section*{Competing Interests}
@@
 as well as code for data analysis and figure generation, as well as scripts for automated
-download of input files are available on GitHub at https://github.com/Klepac-Ceraj-Lab/Resonance,
+download of input files are available on GitHub at https://github.com/KlepacCerajLab/Resonance,
 and archived on Zenodo \cite{kevin_bonham_2023_7647510}.
@@
         \textbf{Random forest models predict both concurrent and future measured cognitive function.}
-        Comparison of RF predictor importance versus linear models for children
-        over 18 months (\textbf{A}).
+        (\textbf{A}) Comparison of RF predictor importance versus linear models for children
+        over 18 months.
         Colors represent whether the species belong to the group of
@@
         \textless{} 0.2) in linear models, both, or neither.
-        Comparison of RF predictor importance for microbial features
-        between concurrent and future cognitive function measure
-        (\textbf{B}).
+        Ranked feature importance for taxa in RF models for
+        both concurrent (\textbf{B}) and future (\textbf{C})
+        cognitive function measurement prediction.
+        Red dots indicate how relative importances change (among taxa)
+        when demographics (sex, SES) are arred as covariables.
+        (\textbf{D}) Comparison of RF predictor importance for microbial features
+        between concurrent and future cognitive function measure.
         Colors represent whether the species belong to the group of
@@
         Features absent of one model are shown on the origin-aligned sideplots.
-        Distribution of ages for each microbiome sampling - cognitive assesment dyad (\textbf{C}).
-        Blue violin plot shows distribution of sampling age,
-        while red violin plot shows distribution of asessment age.
+        Abundances and prevalences of taxa that are important
+        for RF models both for concurrent (\textbf{E}) and future (\textbf{F}) assesment prediction.
+        (\textbf{G}) Distribution of ages for each microbiome sampling - cognitive assesment dyad
+        at sampling age (left),
+        or asessment age (right).
         Black lines connect ages for each individual subject.
-        Ranked feature importance for taxa in RF models for
-        both concurrent (\textbf{D}) and future (\textbf{E})
-        cognitive function measurement prediction.
-        Red dots indicate how relative importances change (among taxa)
-        when demographics (sex, SES) are arred as covariables.
-        Abundances and prevalences of taxa that are important
-        for RF models both for concurrent (\textbf{F}) and future (\textbf{G}) assesment prediction.
         }
